### 2011-09-26 doc/forslag/forslag.tex
new plan, start on paper, some mapping..

\documentclass{article}
%\documentclass[a4paper]{prosper}
%\usepackage{listings}
\usepackage{verbatim}
%\usepackage{graphicx}
%\usepackage{tabularx}
\usepackage[utf8]{inputenc}
%\usepackage{float}

\begin{document}
\title{Proposal for Master Thesis in Computer Science \\
\vspace{2mm}
 \large{A Wide-Coverage Grammar and Parser for Swedish}}
\author{Malin Ahlberg}
\maketitle
\vspace{20mm}

\thispagestyle{empty}
\abstract
%\large{A proposal for a master thesis in Computer Science.}\\
By using existing technologies
and tools, we would like to implement a robust parser for Swedish.
The goal is to be able to parse open domain language and the
parser will be evaluated on an extensive Swedish
treebank, Talbanken. 
A parser like this would be of great use for many natural
language processing applications, such as translation and information retrieval
as well as making semantic representations.

The project is accepted and funded by Center of Language 
Technology, Gothenburg University.

\section*{\center{\normalsize{Description and future work}}}
Making computers able of handling human language is a 
hard problem.
The meaning of a sentence depends not only of which words it consists of, but
also on their syntactic use, how they interact and relate to each other.
For a computer to make sense of natural language, it needs to analyse this 
syntactic structure; it needs a good grammar and parser.
In this project, a robust parser for Swedish will be implemented
using the grammar formalism Grammatical Framework. %\cite{gf} (GF).

The libraries of GF provide a basic grammar for
Swedish, covering the fundamental features of the language, such as morphology and
commonly used syntax. This is suitable 
for building domain specific applications. In those cases the user is not
allowed to freely compose sentences, but has to stay within the bounds of a 
controlled language. Both the vocabulary and grammatical structures are fixed,
meaning that there only is a limit number of ways to write a sentence.
Parsing open domain natural language is a much bigger task, since it involves handling both
standard and non-standard grammatical constructions.

The future work will include extending and enhancing the existing Swedish GF grammar,
importing lexicon and develop techniques for handling unknown words and grammatical
constructions, proper names, idioms, ellipses etc, in order to make the parser robust.
\end{document}

### 2011-09-26 doc/paper/Paper.tex
new plan, start on paper, some mapping..

\documentclass[submission,copyright,creativecommons]{eptcs} %rätt lincens?
\providecommand{\event}{LREC2012}
%\usepackage{breakurl}        
\usepackage{graphicx}

\title{Den stora grammatiken} % ska vi vara anonyma?
\author{Malin Ahlberg
\institute{Gothenburg University, Sweden}
\email{ahlberg.malin@gmail.com}
}
\def\titlerunning{First results}
\def\authorrunning{M. Ahlberg}
\begin{document}
\maketitle

\begin{abstract}
om parsning of free language (take from old report)
but for linguistics, make a wide coverage grammar, robust parser
By using existing technologies
and tools:use three resources, gf, talbanken, saldo
A parser like this would be of great use for many natural
language processing applications, such as translation and information retrieval
as well as making semantic representations.
The goal is to be able to parse open domain language and the
parser will be evaluated on an extensive Swedish
treebank, Talbanken.

This report describes the start of implementing a robust parser for Swedish,
using the grammar formalism Grammatical Framework\cite{gf} (GF).

\end{abstract}

\section{Introduction}

Making computers able of handling human language is a 
hard problem.
The meaning of a sentence depends not only of which words it consists of, but
also on their syntactic use, how they interact and relate to each other.
For a computer to make sense of natural language, it needs to analyse this 
syntactic structure; it needs a good grammar and parser.
Moreover, the parser needs a preprocessor, named entity recognizer and more
tequnices to make it robust.

We are working on an extended implementation of a grammar from Swedish, allowing
complicated constructions such as focusing different parts of the sentences,
using complex relative clauses etc and constructions (byt ord) specific to Swedish.

The libraries of GF provide a basic grammar for
Swedish, covering the fundamental features of the language, such as morphology and
commonly used syntax. This is suitable 
for building domain specific applications. In those cases the user is not
allowed to freely compose sentences, but has to stay within the bounds of a 
controlled language. Both the vocabulary and grammatical structures are fixed,
meaning that there only is a limit number of ways to write a sentence.
Parsing open domain natural language is a much bigger task, since it involves handling both
standard and non-standard grammatical constructions.

The future work will include extending and enhancing the existing Swedish GF grammar,
importing lexicon and develop techniques for handling unknown words and grammatical
constructions, proper names, idioms, ellipses etc, in order to make the parser robust.

By using this, we hope to eventually implement a grammar able to handle free % synonyyym!
Swedish. It would be of great use for many natural
language processing applications, such as translation and information retrieval
as well as making semantic representations.

\subsection{Grammatical Framework}

%compiler inspired approach, 
% om resources: library that implements morphology and syntax rules, länk till papper om dessa
% skriv ord som modular, architectur, reusable (fast stava dem rätt)

% about why we need more, extra module, separate swedish from scandinavian
GF\cite{gf} (Grammatical Framework) is a grammar formalism based on functional programming,
and suitable for multilingual grammar applications. The key idea is to have
one abstract grammar, modelling the structure of the domain. 
In addition, concrete grammars are implementing the abstract in different languages.
Since all languages supported in an application shares a common abstract syntax,
multilingual is easy to support.

In this way the grammar becomes independent, it may be reused in another project
and the implementation may be changed without modifying the application utilizing the grammar.

GF provides resource grammars\cite{resource} implementing information about the morphological and
syntactical rules for 18 different languages. By utilizing these, 
GF is an advantageous tool for writing special-purpose grammars.

The libraries of GF, the resource libraries\cite{resource}, provides
a common abstract and implementations of this in 18 languages.
Since translation always should be possible, the abstract may only contain constructions that
are common to all implemented languages.

  But this also means that all ---- added to the resources
  must be translatable to and expressable in all other language.
 therefor, the resources are very general, non language specific.

Language specific constructs may be
given in the module \verb|Extra.gf|. This may be stylistic changes, idioms, informal
expressions etc.
As an example, the Swedish \verb|Extra| module contains a function for expressing sentences
where the negation is put in focus: \emph{``Inte var jag glad"} (\emph{``Not was I happy"}).

The work to enhance and expand the Swedish GF grammar has already been started
% parsing free language, more work is required and it has been investigated what
%Further, other tools that are needed have been implemented and examined.
%will be needed to make it sufficient and a start on developing tools has beenmade.

\section{Talbanken}
%Wherefrom, how big, who, when, version, used for maltparser
Talbanken\cite{talbanken} is
a Swedish treebank which was put together in the 1970s at Lund University,
consisting of 6316 sentences. Each word is tagged with 
its part of speech, syntactic function and its position in a head dependent
analysis.
In 2005 Talbanken was mordenized \cite{} and enriched with annotation for a
full phrase structure analysis. Talbanken05 was used when training the 
data-driven parser Maltparser \cite{}, which var bra?

\section{Saldo}

SALDO\cite{saldo} is an open source lexicon resource
based on Svenskt Associations Lexikon (SAL). It is
developed at Sprakbanken\cite{} at Gothenburg University
and intended to be used in language technology
research. 
% how it is tagged, other usages, benefits,
% how easy it is to map it to gf
From SALDO, a large GF lexicon, \\
\verb|DictSwe.gf| has earlier been extracted \cite{},
containing 50 000 entries. Using the same updated tecniques,
we would like to reimport the lexicon in order to enlarge,
enhance it. The importing-method should be fast and reliable
enough to be able to always have an fresh version of the dictonary
in GF.
We are using parts of Talbanken for testing and development,
and the parsetree generated by our grammar will eventually be 
compared to those from the treebank.

\section{The lexicon tool}
A tool for automatically acquisition has been created. Making use of the
tags in Talbanken and the paradigms in the Swedish resource grammar, 
it interactivelly generates GF lexicons. The user
verifies tho correctness of a guess from the program, and can either
allow the word to be added to the lexicon, remove it or demand the program
to make another guess.
Altough using very simple tecniques, the test of the tool on verbs 
has a correctness rate of about
70-75\%, and can easily ? be improved.

use together with Extract/FM,

\section{Mapping of trees}
The information from the tags in Talbanken can be used for many xx.
The tools for lexical extraction can be enhanced if we get more information
about which form the word is currently used in. 
We are currently working on a automatical mapping from Talbanken trees to trees in GF format.
This would later enable us to extract possibilities for how often different functions
are used, a feature (!) that would improve the parsing considerably.
The evalution of the parser could also be accomplished by comparing the trees
from the parser and from the mapping-process.
brastartord there are differences in the notation of the sättet träden är uppbyggda på \\
 exempel?? \\

explation of the trees
why we need the mapping : evaluation, adding words, finding missing constructions
difficulties

\section{New things and hard things in the grammar}
Due to the ressemblemences between the Scandinavian languages, they share about 80\% (?)
of their resource code. This is good because blabla.
However, for a bigger Swedish grammar, the implementation of the languages need to be more separated,
in order to avoid ad-hoc solutions and constructions that are standard in
Norweigan and Danish but very rare in Swedish. \\ %skriv om här änna
A number of grammatical constructions have been added the the Swedish grammar.
Those include using the reflexive pronoun \emph{sitt}:\\
\textbf{Han såg sitt hus} \emph{He saw his (own) house} \\
as opposed to \\
\textbf{Han såg hans hus} \emph{He saw his (an other person's) house} \\
% to simple for linguists?
The possibility to put a part of a sentence in focus has also been added such as:\\
\textbf{Glad var han inte} \emph{Happy was he not}
Some ellipses are added, while there is work going on to add trickier ones (eller inte 
ellipser just nu dårå, men vi skriver något fint här med, ja-jaaaa!).

\section{Future Work}
robust parser, ner, preprocessing, adding new words,
probabilties
ellipses, chunk parsing

\section{Conclusion}
hard but nice
why better than statistical

\nocite{*}
\bibliographystyle{eptcs}
\bibliography{ourBib}
\end{document}

### 2011-09-30 doc/gullmarsstrand/pres.tex
presentation

\documentclass[10pt]{beamer} 
%\usefonttheme{structuresmallcapsserif} 
%% \usepackage{beamerthemeshadow}
\usepackage{verbatim} 
%% \usetheme{Pittsburgh}
\usepackage{colortbl}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{cancel}
 \renewcommand{\baselinestretch}{1.5}
     \normalsize

% THIS SHOULD BE HERE!
% No unimportant, irrelevant things. Only information.
% Only code if it is of significance.

% get inspired by:
% Simon Jones, Microsoft research, videos.
% John Hughes article How to give a good research presentation.
\title{A Wide-Coverage Grammar and Parser for Swedish}
\subtitle{\large First results and perspectives}
\author{Malin Ahlberg \\ Gothenburg University}
\date{}

\begin{document}
\maketitle

\begin{frame}
    \framesubtitle{The goal}
    Malin Ahlberg \\
 \pause 
    Gothenburg University
\end{frame}

\begin{frame}
\frametitle{What I'm doing now}
\framesubtitle{} 
mapping
grammar 
saldo
\end{frame}

\begin{frame}
\frametitle{mapping}
\framesubtitle{Example} 
Nice example
\end{frame}

\begin{frame}
\frametitle{mapping}
\framesubtitle{Why, how} 
why it's hard
why it's good
goals
\end{frame}

\begin{frame}
\frametitle{mapping}
\framesubtitle{problems} 
avp
differnt parsings
\end{frame}

\begin{frame}
\frametitle{grammar added}
\framesubtitle{examples} 
PassVP "äpplet åts"
  ("äpplet blev ätet")
ReflNP, ReflSlash

RelNP' "flickan, som inte ätit äpplen"
 (RelNp "flickan , sådan att hon inte åt äpplen")

AdvVPSlash "hon åt redan äpplet" (hon åt äpplet redan")

PPartAP "det är sett","det ätna äpplet"
 (not "det är givet till henne" osv, "minskad efterfrågan")

list of functions
examples
difficulties
\end{frame}

\begin{frame}
\frametitle{Particip}
\framesubtitle{What type does a particip have?} 
attribut- predikativ
en ökad effektiv information (ökat vanligt!!)
en minskad benägenhet
i en gången tids svenska bondesamhälle  4872    
en fritt vuxen natur
centralt belägna, eftertraktade lägenheter
de stegrade hyrorna
fyra representerade kyrkor
en kvalificerad majoritet
misstänkt (av)
kokt potatis (av)
som aldrig är skalad (industriskalad, skalkokt)
vår behandlade potatis
är oroad
beprövad (av)
de levererar den doppad i ...
\end{frame}

\begin{frame}
\frametitle{ReflGenVP}
\framesubtitle{Problem} 
sin egen lilla stubbe -ok 
alla sina barn
all vår undervisning
sin urgamla rätt att... 
hur löses det för vår? (inget problem för det hör ej ihop med subj)

\end{frame}

\begin{frame}
\frametitle{Goals}
\framesubtitle{Goals for christmas} 
finish and evaluate mapping
extend grammar with: relative clauses, ..? what more
new mapping of saldo - making it easy to redo
\end{frame}

\begin{frame}
    \frametitle{The end}
\huge\begin{center}Questions?\end{center}
\end{frame}
\end{comment}
\end{document}

### 2011-09-30 mapping/GraphTree.hs
testing tools

@@ -1,4 +1,5 @@
 module GraphTree where
+import Data.Char
 import Data.Tree
     
 -- Tree helper function: generate a graphviz dot ﬁle depicting it
@@ -6,14 +7,14 @@ dotTree :: Tree String -> String -> String
 dotTree t [] = dotTree t "T"
 dotTree t name = header ++ body ++ footer where
    header = "digraph \"" ++ name ++ "\" {\n"
-   body = dotTreeBody t 
+   body = dotTreeBody t 0
    footer = "\n}"
    -- Tree helper function: generate the body of the digraph
-   dotTreeBody :: Tree String -> String
-   dotTreeBody (Node n [Node x []]) =
-        n ++ " [label=\"" ++ show x ++ "\"]\n"
-   dotTreeBody (Node name xs) =
+   dotTreeBody :: Tree String -> Int ->  String
+   dotTreeBody (Node (n:ns) []) i =
+        (toUpper n : ns) ++show i++ " []" --label=\"" ++ "\"]\n"
+   dotTreeBody (Node name xs) i =
       mynode ++ links ++ bodies where
-        mynode = name ++ " [style=filled,color=lightgrey];\n"
-        links = concat [name ++ " -> " ++ n ++ ";\n" | (Node n _) <- xs]
-        bodies = concat [dotTreeBody (Node n ts) | (Node n ts) <- xs ]
+        mynode = name++show i ++ " [style=filled,color=lightgrey];\n"
+        links = concat [name++show i ++ " -> " ++ n++show (i+1) ++ ";\n" | (Node n _) <- xs]
+        bodies = concat [dotTreeBody (Node n ts) (i+1) | (Node n ts) <- xs ]

### 2011-09-30 mapping/Skript.hs

@@ -1,0 +1,8 @@
+import Data.List
+
+
+maini f = do
+ inp <- readFile f
+ let metas  = map (length . filter (=='?')) (lines inp)
+ return $ map (\x -> (head x,length x)) $ group $ sort metas
+

### 2011-09-30 mapping/Tester.hs

@@ -1,0 +1,86 @@
+module Tester where
+import XMLHelp
+import System.IO.Error
+import System.FilePath.Posix
+import System.Directory
+import System.Environment (getArgs)
+import System.Exit (exitFailure, exitSuccess)
+
+import PGF
+import Data.Maybe
+
+
+type ParsedS = (Sent,[Tree]) 
+type Sent = (Id,String)
+
+sentType = fromJust $ readType "Text"
+pgfFile = "../gf/BigTest.pgf"
+
+main :: String ->  IO ()
+main filePath = do -- do args <- getArgs
+        --  case args of
+        --     filePath:args' -> do
+                sents <- mainF filePath (return . map getSentence)
+                res <- testParse getLang $ concat sents 
+                let noOk = length $ filter ((/=[]). snd) res
+                putStrLn $ "Writing output to file resFile"
+                let outp = parsNo noOk ++ (unlines $ map formatRes res)
+                writeFile "resFile.txt" outp 
+                putStrLn $ "Results written to: resFile"
+                putStrLn $ "Parsed "++show noOk ++" out of totally "++show (length res)
+        --        exitSuccess
+        --     []      -> do
+        --        putStrLn "Give a filename"
+        --        exitFailure
+
+parseS :: PGF -> Language -> Sent -> IO ParsedS
+parseS pgf lang s = do
+  putStrLn $ "try "++snd s
+  let mParsed = parse pgf lang sentType (snd s)
+  putMsg s mParsed
+  return (s,mParsed)
+
+testParse :: Language -> [Sent] -> IO [ParsedS]
+testParse lang ss = do
+  putStrLn "Parsing input..."
+  mpgf <- try (readPGF pgfFile)
+  putStrLn "Input parsed..."
+  case mpgf of
+    Left e    -> error "ingen pgf!!" 
+    Right pgf -> mapM (parseS pgf lang) ss
+
+putMsg :: Sent -> [Tree] -> IO ()
+putMsg s [] = putStrLn $ show s ++ color red " was not parsed"
+putMsg s ps = putStrLn $ show s ++ (color green " was parsed in "
+                                      ++show (length ps)++" ways")
+
+parsNo :: Int -> String
+parsNo n = "Parsed "++show n++"sentences\n"
+
+getLang :: Language
+getLang = fromJust $ readLanguage "BigTestSwe"
+
[…]

### 2011-09-30 mapping/translate.hs

@@ -49,7 +49,7 @@ startState = S {_isReflGenVP = False, _isExist = False}
        -}
 
 main = main' "test.xml" >> return ()
-bigTest = main' "../testSuites/testShortSimpleTwo.xml" >>= writeFile "mappingShort3.txt" . unlines
+bigTest = main' "../testSuites/testShortSimpleTwo.xml" >>= writeFile "mappingShort4.txt" . unlines
 main2 = main' "test2.xml" >> return ()
 mainTest = main' "testSimple.xml" >>= putStrLn . compareRes
 mainT2 = main' "testSimple.xml" >>= putStrLn . unlines
@@ -107,8 +107,10 @@ testa  str = do
   return [(lemma,an,cat) | (lemma,an) <- lookupMorpho morpho str
                    ,let cat = maybe "" (showType []) (functionType pgf lemma)]
 
-paintTree t = do
+paintTree = do
+  s <- fmap concat $ Form.parse "test.xml" 
   pgf <- readPGF $ fst usePGF
+  let t = snd $ head s
   writeFile "tmp_treetest.dot" (dotTree t [])
   rawSystem "dot" ["-Tpdf", "tmp_treetest.dot", "-otesttree.pdf"]
   return ()

### 2011-09-30 doc/gullmarsstrand/pres.tex
presentation

@@
 \framesubtitle{problems} 
+so far 42\% of the easy sentences are totally mapped 
+       12\% not at all
+
 avp
@@
 \framesubtitle{examples} 
+Easy sentences with known words (old testsuite) : 90 out of 118 (76\%) 
 PassVP "äpplet åts"
   ("äpplet blev ätet")
+
 ReflNP, ReflSlash
+  relate to Peter - sig/sig själv, can't say hon ger den till sig
+  (solved?)
 
@@
 vår behandlade potatis
+
 är oroad
@@
 finish and evaluate mapping
-extend grammar with: relative clauses, ..? what more
+extend grammar with: relative clauses, embedded clauses
 new mapping of saldo - making it easy to redo
+evaluation
 \end{frame}

### 2011-10-02 mapping/Tester.hs
update to mapping

@@ -25,7 +25,7 @@ main filePath = do -- do args <- getArgs
                 let noOk = length $ filter ((/=[]). snd) res
                 putStrLn $ "Writing output to file resFile"
                 let outp = parsNo noOk ++ (unlines $ map formatRes res)
-                writeFile "resFile.txt" outp 
+                writeFile "resFile2.txt" outp 
                 putStrLn $ "Results written to: resFile"
                 putStrLn $ "Parsed "++show noOk ++" out of totally "++show (length res)
         --        exitSuccess

### 2011-10-02 mapping/translate.hs

@@ -291,8 +291,23 @@ pCl =
          --(cl1,t) = maybe (e1,cidUttS) (\q -> (mkApp cidQuestIAdv [q,cl],cidQuestVP)) qadv  
          e2 = foldr (\ad e -> mkApp cidAdvS [ad, e]) e1 advs
      return (e2,tmp,pol,t)
-  --`mplus`
-  ---- question; 'vem är det'   
+  `mplus`
+  --- question; 'vem är det'   
+  do (tmp,pol,np,nptyp,vp) <- msum $ map pOVS vForms
+     advs2 <- many pAdv
+     let e0 = foldr (\ad e -> mkApp cidAdvVP [e,ad]) vp advs2
+         cl = constructCl nptyp np e0
+         t  =cidQuestVP -- kolla att det är sant
+         c  = maybe cl (\q -> (mkApp cidQuestIAdv [q,cl])) qadv  
+         e1 = mkApp (clType t) [fromMaybe (mkExpr meta) (isVTense tmp)
+                             ,mkExpr pol
+                             ,c
+                             ]
+               -- if pAdv:s is found after pIAdv, there will be a problem..
+         --(cl1,t) = maybe (e1,cidUttS) (\q -> (mkApp cidQuestIAdv [q,cl],cidQuestVP)) qadv  
+         e2 = foldr (\ad e -> mkApp cidAdvS [ad, e]) e1 advs
+     return (e2,tmp,pol,t)
+
   --sp
   --fv
   --ss
@@ -305,10 +320,14 @@ pSS =
 
 
 pOVS typ = do
-  -- (tmp,s,pol,vp) <- pComplVP typ 
-  return undefined
+  exps <- pCompl typ
+  write $ "found compl in OVS "++show typ
+  (v,t) <- pSlashVP typ "FV"
+  (tmp,s,pol,vp) <- pComplVP typ v t exps
+  (np,nptyp) <- write "looking for SS in OVS" >> parseSubject
+  return (tmp,pol,np,vptyp,vp)
 
-pVSOs = foldr1 (mplus) $ map pVSO vForms
+pVSOs = msum $ map pVSO vForms
 
 pVSO typ = do
   (v,t) <- pSlashVP typ "FV"
@@ -316,7 +335,6 @@ pVSO typ = do
            --inside "FV" (consume >> metaVP' typ)    -- slash should find them if the right tag is there!
   (np,nptyp) <- write "looking for SS" >> parseSubject
[…]

### 2011-10-02 doc/gullmarsstrand/pres.tex
presentation

--- 2011-09-30 doc/gullmarsstrand/pres.tex
@@
 \begin{frame}
-    \framesubtitle{The goal}
-    Malin Ahlberg \\
- \pause 
-    Gothenburg University
+\frametitle{Current work}
+\framesubtitle{} 
+A mapping of Talbanken05 trees to GF trees\\
+Extending the grammar \\
+Re-import saldo\\
 \end{frame}
@@
 \begin{frame}
-\frametitle{What I'm doing now}
-\framesubtitle{} 
-mapping
-grammar 
-saldo
-\end{frame}
-
-\begin{frame}
-\frametitle{mapping}
+\frametitle{Mapping}
 \framesubtitle{Example} 
@@
 \begin{frame}
-\frametitle{mapping}
-\framesubtitle{Why, how} 
-why it's hard
-why it's good
-goals
+\frametitle{Mapping}
+\framesubtitle{Why?} 
+In order to \\
+\begin{itemize}
+\item{evaluate the parser}
+\item{identify grammatical constructions not current in the GF grammar}
+\item{enable good lexical extraction}
+\item{extract probabilites for GF functions}
+\end{itemize}
+% why it's hard (396 hur och hur)
+%goals
 \end{frame}
 
+%\begin{frame}
+%\frametitle{mapping}
+%\framesubtitle{problems} 
+%so far 42\% of the easy sentences are totally mapped 
+%       10\% not at all
+%
+%avp
+%differnt parsings
+%\end{frame}
+
 \begin{frame}
-\frametitle{mapping}
-\framesubtitle{problems} 
-avp
-differnt parsings
+\frametitle{The grammar}
+\framesubtitle{Current work} 
+\begin{tabular}{ll}
+\textbf{PassVP} & V2 $\rightarrow$ VP (should be VPSlash $\rightarrow$ VP) \\
+& \emph{äpplet åts},\emph{boken gavs till honom} \\ % \;
+ %\emph{("äpplet blev ätet")}\\ % boken gavs till honom
+
+\textbf{ReflSlash} & VPSlash $\rightarrow$ VP \\
+& \emph{han ger sina små barn till dem} \\
+% \\ but not \emph{sig själv}, \emph{han äter alla sina äpplen} or \emph{han ger sig till sig}  \\
+  %relate to Peter - sig/sig själv, can't say hon ger sig till sig
+%
+%%RelNP' "flickan, som inte ätit äpplen"
+%% (RelNp "flickan , sådan att hon inte åt äpplen")
+
+% CompAdv here, plus we need ComlVVAdv, Pass? osv
+\textbf{AdvVPSlash} & VPSlash $\rightarrow$ Adv $\rightarrow$ VPSlash \\
+& \emph{hon åt redan äpplet}, \emph{hon är redan här}\\
+% &\emph{("hon åt äpplet redan")} \\
+
+\textbf{PPartAP} & V2 $\rightarrow$ AP \\
+& \emph{det är skrivet},\emph{den skrivna artikeln}
+\end{tabular}\\
+
+Easy sentences with known words (old testsuite) : 92 out of 118 (78\%) \\
+%list of functions
+%examples
+%difficulties
+%sådan
 \end{frame}
 
+%kolla upp dessa i stort lexikon
 \begin{frame}
-\frametitle{grammar added}
-\framesubtitle{examples} 
-PassVP "äpplet åts"
-  ("äpplet blev ätet")
-ReflNP, ReflSlash
-
-RelNP' "flickan, som inte ätit äpplen"
- (RelNp "flickan , sådan att hon inte åt äpplen")
-
-AdvVPSlash "hon åt redan äpplet" (hon åt äpplet redan")
-
-PPartAP "det är sett","det ätna äpplet"
- (not "det är givet till henne" osv, "minskad efterfrågan")
-
-list of functions
-examples
-difficulties
+\frametitle{Perfect Participle}
+\framesubtitle{What type of functions do we need?} 
+PPartAP cannot handle \emph{det är givet till henne} or \emph{gågna åren} \\
+VPSlash, V2, V?\\
+Is the word an adjective or verb? \\
+%attribut- predikativ
+\begin{tabular}{ll}
+Some examples: & \\
+\textbf{V} & \\
+& \emph{försvunnen}, \emph{slocknad} \\
+& \emph{i en gången tids svenska bondesamhälle}  \\% 4872    
+\end{tabular} \\
+\textbf{V2 or V?}
+\emph{en ökad effektiv information} \\
+\emph{en minskad benägenhet} \\
+\textbf{V2}
+\emph{de stegrade hyrorna} \\
+\emph{fyra representerade kyrkor} \\
+\emph{som aldrig är skalad (industriskalad, skalkokt)} \\
+\emph{eftertraktade lägenheter} \\
+\textbf{V3}
+\emph{skickad till mig}
+\textbf{Adj?}
+\emph{en fritt vuxen natur } \\
+\emph{centralt belägna} \\
+\emph{misstänkt } \\% (av)}
+\emph{en kvalificerad majoritet} \\
+\emph{kokt potatis} \\% (av)}
+?
+\emph{vår behandlade potatis} \\
+\emph{är oroad} \\
+\emph{beprövad}  \\ % (av)
+\emph{de levererar den doppad i askorbinsyra} \\
 \end{frame}
 
-\begin{frame}
-\frametitle{Particip}
-\framesubtitle{What type does a particip have?} 
-attribut- predikativ
-en ökad effektiv information (ökat vanligt!!)
-en minskad benägenhet
-i en gången tids svenska bondesamhälle  4872    
-en fritt vuxen natur
-centralt belägna, eftertraktade lägenheter
-de stegrade hyrorna
-fyra representerade kyrkor
-en kvalificerad majoritet
-misstänkt (av)
-kokt potatis (av)
-som aldrig är skalad (industriskalad, skalkokt)
-vår behandlade potatis
-är oroad
-beprövad (av)
-de levererar den doppad i ...
-\end{frame}
-
-\begin{frame}
+\begin{frame}[containsverbatim]
 \frametitle{ReflGenVP}
 \framesubtitle{Problem} 
-sin egen lilla stubbe -ok 
-alla sina barn
-all vår undervisning
-sin urgamla rätt att... 
-hur löses det för vår? (inget problem för det hör ej ihop med subj)
+\begin{tabular}{ll}
+Adjectives are ok: &  \emph{sin gula lilla katt} \\
+% fixa indentering
+& \small{\verb|ReflCN NumSg (AdjCN (PositA yellow\_A)|} \\
+&\small{\verb|(AdjCN (PositA small\_A) (UseN cat\_N)))|}\\
+\end{tabular} \\
+\begin{tabular}{ll}
+But predeterminers are not: & \emph{alla sina barn} \\
+ & \emph{all vår undervisning} \\
+ & \emph{sin urgamla rätt att ...}\\  % N2V?
+\end{tabular} \\
+%hur löses det för vår? (inget problem för det hör ej ihop med subj)
 
@@
 \begin{frame}
+\frametitle{The Grammar}
+\framesubtitle{Future work} 
+Reciprocals pronouns - \emph{varandra} \\
+Add category of words that can be used either as adjectives or determiners - \\
+   \emph{fler katter är där}, \emph{fler är där}, \emph{kattern är fler}\\ %sådana
+Category used as predeterminer or adverbs - \\
+  \emph{bara barn}, \emph{hon bara log} \\
+\emph{en sorts katt}, \emph{katter av en sort}
+den yngsta är gulast
+idioms
+minst, lika
+"Flickan som såg katten är här"
+\textbf{Cleft sentences} är ok!!!
+"Det är till uppträdande i denna miljön som utbildningen närmast syftar"
+"Det är då som det träder i funktion"
+\end{frame}
+\begin{frame}
 \frametitle{Goals}
 \framesubtitle{Goals for christmas} 
-finish and evaluate mapping
-extend grammar with: relative clauses, ..? what more
-new mapping of saldo - making it easy to redo
+\begin{itemize}
+\item{Finish and evaluate mapping}
+\item{Extend grammar}
+    \item{New mapping of saldo - making it easy to keep ut-to-date}
+\item{Evaluation}
+\end{itemize}
 \end{frame}
@@
     \frametitle{The end}
-\huge\begin{center}Questions?\end{center}
+\huge\begin{center}Thanks for listening\end{center}
+\large\begin{center}Questions?\end{center}
 \end{frame}
-\end{comment}
 \end{document}

### 2011-10-09 doc/paper/Paper.tex
paper

@@
-\documentclass[submission,copyright,creativecommons]{eptcs} %rätt lincens?
+\documentclass[submission]{eptcs} %rätt lincens?
 \providecommand{\event}{LREC2012}
+\usepackage[utf8]{inputenc}
 %\usepackage{breakurl}        
@@
 
-\title{Den stora grammatiken} % ska vi vara anonyma?
+\title{First results of developing a ruled based Swedish parser} %A Wide-Coverage Swedish Grammar} % ska vi vara anonyma?
 \author{Malin Ahlberg
-\institute{Gothenburg University, Sweden}
-\email{ahlberg.malin@gmail.com}
+% vad ska stå här eg?
+\institute{Center of Language Technology, Gothenburg University, Sweden}
+%\email{ahlberg.malin@gmail.com}
 }
-\def\titlerunning{First results}
+\def\titlerunning{First results of developing a ruled based Swedish parser}
 \def\authorrunning{M. Ahlberg}
@@
 \begin{abstract}
-om parsning of free language (take from old report)
-but for linguistics, make a wide coverage grammar, robust parser
-By using existing technologies
-and tools:use three resources, gf, talbanken, saldo
-A parser like this would be of great use for many natural
-language processing applications, such as translation and information retrieval
-as well as making semantic representations.
-The goal is to be able to parse open domain language and the
-parser will be evaluated on an extensive Swedish
-treebank, Talbanken.
+This paper describes an on-going work of a rule-based, open source
+parser for Swedish, based on a wide-coverage grammar implemented in
+Grammatical Framework (GF).
+In addition to GF, we use two other main resources; the Swedish treebank Talbanken
+and the electronic lexicon SALDO.
+The grammar will be highly reusable as GF provides libraries for compiling
+grammars to a number of programming languages.
+% allow easy interaction with other programs and grammars
+%
 
-This report describes the start of implementing a robust parser for Swedish,
-using the grammar formalism Grammatical Framework\cite{gf} (GF).
+%>> something very commercial, interesting and outstanding here please!
+
+%R: this phrase "We aim to implement a parser for open domain language
+%and this paper describes our work and results so far as well as the
+%directions of future work." doesn't belong in the abstract because it
+%just describes the normal structure of any paper, without saying why
+%your work are interesting/outstanding, worth-reading-further. Make it
+%sound more commercial, if you want to have it there.
 
@@
 \section{Introduction}
+%R: the first sentence - in "deep robust parser" - why "deep" - grammar
+%formalism, why "robust" - ??
+% afterwards, you don't actually have named-entity recognition, and
+%further more, there's no such thing as "exhaustive lexicon" - you can
+%always switch to another domain and find new words. Further on, the
+%work has nothing to do with statistitics so far, so you can't mention
+%as opening line.  The same goes for parallel chunk parsing.
 
-Making computers able of handling human language is a 
-hard problem.
-The meaning of a sentence depends not only of which words it consists of, but
-also on their syntactic use, how they interact and relate to each other.
-For a computer to make sense of natural language, it needs to analyse this 
-syntactic structure; it needs a good grammar and parser.
-Moreover, the parser needs a preprocessor, named entity recognizer and more
-tequnices to make it robust.
+%R: the first sentence doesn't really make sense because you first
+%refer to a parser and only then to GF and a grammar, but without
+%giving an explanation of how making the grammar would give you a
+%parser for free.
 
-We are working on an extended implementation of a grammar from Swedish, allowing
-complicated constructions such as focusing different parts of the sentences,
-using complex relative clauses etc and constructions (byt ord) specific to Swedish.
+%R: -- the second paragraph - OK but please extend.
+We would like to implement a wide coverage grammar and parser for Swedish.
+We thereby investigate how the grammar formalism Grammatical Framework,
+further described in section \ref{sec:gf}, can be used for open domain parsing.
+By starting from an already existing Swedish grammar written in GF, we get 
+an fundamental description of the language. The framework also provides
+tools such as parsing, generation and
+a well tested interpretation of the parse trees. There are also tools for using
+the grammar in a number of programming languages. 
+%>> , tools for visualizing ...
 
-The libraries of GF provide a basic grammar for
-Swedish, covering the fundamental features of the language, such as morphology and
-commonly used syntax. This is suitable 
-for building domain specific applications. In those cases the user is not
-allowed to freely compose sentences, but has to stay within the bounds of a 
-controlled language. Both the vocabulary and grammatical structures are fixed,
-meaning that there only is a limit number of ways to write a sentence.
-Parsing open domain natural language is a much bigger task, since it involves handling both
-standard and non-standard grammatical constructions.
+From the GF grammar we aim to implement a robust parser. 
+The parser will be evaluated both by experts and by comparing the results of
+parsing sentences in an often used Swedish treebank, Talbanken.
 
-The future work will include extending and enhancing the existing Swedish GF grammar,
-importing lexicon and develop techniques for handling unknown words and grammatical
-constructions, proper names, idioms, ellipses etc, in order to make the parser robust.
+%We aim to implement % The goal is to be able to 
+%a parser for open domain language and 
+%This paper describes our work and results 
+%so far as well as the directions of future work.
 
-By using this, we hope to eventually implement a grammar able to handle free % synonyyym!
-Swedish. It would be of great use for many natural
-language processing applications, such as translation and information retrieval
-as well as making semantic representations.
+\section{Background}
+\subsection{Grammatical Framework}
+\label{sec:gf}
 
-\subsection{Grammatical Framework}
+%R: the GF part - "especially suitable" - it's actually meant for the purpose
+%:)
 
@@
 
+Grammatical Framework\cite{gf} (GF) is a grammar formalism based on functional
+programming, designed for multilingual grammar applications.
+
+The key idea
+is to divide the grammar into abstract and concrete.
+The abstract grammar gives a logical representation of the semantics of the
+domain, modeled as abstract trees. Complications occurring in
+natural languages, such as agreement, case and word order, are abstracted away. 
+%R: what complications do you abstract away ?
+%categories and trees are not interesting as such, but their usage
+%is - why do you need them in the abstract syntax actually ?
+That is, the abstract grammar contains no actual linguistic information, but
+information about what categories are used and how to combine them into trees.
+
+To add a new languages means to give a concrete grammar. This implements
+the abstract grammar and describes how to translate abstract
+trees to strings of the language. The framework enables us to parse strings into 
+abstract trees as well as linearize trees into strings.
+Since all languages supported in an application shares a common abstract syntax,
+we get multilinguality for free.
+
+% obra mening
+The grammar acts as an independent module and may be used in different
+projects.
+%R: more about the GF library - if you have space
+Reusability is further supported by the GF resource grammars\cite{resource},
+a library provided with the GF package, 
+where information about the morphological and
+syntactical rules for more than 20 different languages are implemented.
+As a result, an end user of the framework does not need to describe how to 
+form standard sentences, phrases or decline words, but can focus on the
+particularities for the given domain and choose what kind of 
+utterances that should be allowed.
+
+%so far bara used för vissa saker, 
 % about why we need more, extra module, separate swedish from scandinavian
-GF\cite{gf} (Grammatical Framework) is a grammar formalism based on functional programming,
-and suitable for multilingual grammar applications. The key idea is to have
-one abstract grammar, modelling the structure of the domain. 
-In addition, concrete grammars are implementing the abstract in different languages.
-Since all languages supported in an application shares a common abstract syntax,
-multilingual is easy to support.
+% get inspired by Shafqat.
+GF has so far been used in a number of projects, MOLTO\cite{molto} and
+WebAlt\cite{webalt} to mention a few. 
+All those are special domain applications, dealing with controlled natural
+language.
+This project experiments on other usages of a GF grammar.
 
-In this way the grammar becomes independent, it may be reused in another project
-and the implementation may be changed without modifying the application utilizing the grammar.
-
-GF provides resource grammars\cite{resource} implementing information about the morphological and
-syntactical rules for 18 different languages. By utilizing these, 
-GF is an advantageous tool for writing special-purpose grammars.
-
-The libraries of GF, the resource libraries\cite{resource}, provides
-a common abstract and implementations of this in 18 languages.
-Since translation always should be possible, the abstract may only contain constructions that
-are common to all implemented languages.
-
-  But this also means that all ---- added to the resources
-  must be translatable to and expressable in all other language.
- therefor, the resources are very general, non language specific.
-
-Language specific constructs may be
-given in the module \verb|Extra.gf|. This may be stylistic changes, idioms, informal
-expressions etc.
-As an example, the Swedish \verb|Extra| module contains a function for expressing sentences
-where the negation is put in focus: \emph{``Inte var jag glad"} (\emph{``Not was I happy"}).
-
-The work to enhance and expand the Swedish GF grammar has already been started
 % parsing free language, more work is required and it has been investigated what
@@
 
-\section{Talbanken}
+\subsection{Talbanken}
 %Wherefrom, how big, who, when, version, used for maltparser
 Talbanken\cite{talbanken} is
-a Swedish treebank which was put together in the 1970s at Lund University,
-consisting of 6316 sentences. Each word is tagged with 
-its part of speech, syntactic function and its position in a head dependent
-analysis.
-In 2005 Talbanken was mordenized \cite{} and enriched with annotation for a
-full phrase structure analysis. Talbanken05 was used when training the 
-data-driven parser Maltparser \cite{}, which var bra?
+a Swedish treebank put together in the 1970s at Lund University.
+In 2005 it was modernized by Nivre, Nilsson and Hall\cite{talbanken05} and
+enriched with annotation for a
+full phrase structure analysis. 
+Talbanken contains both written and spoken Swedish. For this project only
+the prose material, consisting of 6316 sentences will be used.
+This part was also used when training the data-driven parser Maltparser \cite{malt}. \\
+We are using parts of Talbanken for testing, development and evaluation.
+%and the parsetree generated by our grammar will eventually be 
+%compared to those from the treebank.
 
-\section{Saldo}
-
+\subsection{Saldo}
+%R: Saldo part - why would you like to reimport ? "enlarge and enhance"
+%- how do you enhance it ?
 SALDO\cite{saldo} is an open source lexicon resource
-based on Svenskt Associations Lexikon (SAL). It is
-developed at Sprakbanken\cite{} at Gothenburg University
-and intended to be used in language technology
+based on Svenskt Associationslexikon (SAL). It is
+developed at Språkbanken at Gothenburg University
+and intended for usage in language technology
 research. 
@@
 % how easy it is to map it to gf
-From SALDO, a large GF lexicon, \\
-\verb|DictSwe.gf| has earlier been extracted \cite{},
-containing 50 000 entries. Using the same updated tecniques,
-we would like to reimport the lexicon in order to enlarge,
-enhance it. The importing-method should be fast and reliable
-enough to be able to always have an fresh version of the dictonary
+From SALDO, a large GF lexicon
+has earlier been extracted,
+containing 50 000 entries.
+By updating this technique, we would like
+to reimport SALDO, which has since then been developed.
+%and we would like
+%to reimport the it using the same
+%updated techniques. %in order to enlarge our lexicon,
+%we  in order to enlarge and
+%GF version. 
+The importing method should be fast and reliable
+enough to be allow us to always have a fresh version of the dictionary
 in GF.
-We are using parts of Talbanken for testing and development,
-and the parsetree generated by our grammar will eventually be 
-compared to those from the treebank.
 
-\section{The lexicon tool}
-A tool for automatically acquisition has been created. Making use of the
-tags in Talbanken and the paradigms in the Swedish resource grammar, 
-it interactivelly generates GF lexicons. The user
-verifies tho correctness of a guess from the program, and can either
+\section{Work in progress}
+\subsection{A tool for lexical acquisition}
+%R: lexicon extraction part - more about everything, comparison with
+%Extract and Functional Morphology, evaluation of the lexicon builder -
+%"supervised lexicon extraction"(common term for this - google it if
+%you want to find related work)
+
+To enlarge the lexicon, 
+a tool for automatically acquisition has been created. It has 
+been tested on verbs with good results. It makes use of
+the \emph{smart paradigm} given in the Swedish resource grammar.
+The smart paradigm accepts verbs in present tense indicative form and decides
+ which paradigm it most likely belong to. If needed, it also accepts more
+verb forms showing the correct inflection.
+%which Making use of the
+
+By combining this method with the information from the tags in Talbanken,
+%and the paradigms in the Swedish resource grammar, 
+the tool interactively generates GF lexicons. 
+Given a list of words, it iteratively
+tries to figure out how to conjugate each of them. If several forms of a word is 
+given, the program will try to identify the one that carries the most linguistic
+information, put this in a form recognized by the smart paradigm and ask GF to output
+a table with the resulting inflection table. 
+If this table contains all other forms from the input list,
+the program will ask the  user to
+validate the claimed conjunction. The user may now either
 allow the word to be added to the lexicon, remove it or demand the program
 to make another guess.
-Altough using very simple tecniques, the test of the tool on verbs 
-has a correctness rate of about
-70-75\%, and can easily ? be improved.
+Although using simple techniques, the tool 
+manage to correctly guess 70-75\% of the given lemmas when tested on
+verbs. \\
+For further extractions, there are also earlier developed tools such as
+Extract\cite{extract} for supervised lexicon extraction and
+FM\cite{fm} for programming lexical resources. Extract uses a similar 
+method but does not have access to the GF smart paradigm.
 
-use together with Extract/FM,
+%% Write more here!!
+\subsection{Mapping of trees}
+%R: mapping - very nice - more about it
+The information from the tags in Talbanken can be used for many purposes.
+We are currently working on an automatic transformation of Talbanken trees 
+to trees in GF format. The translation makes use of the POS tags as well as
+the syntactic information and the mapping has so far turned out to be unambiguous.
 
-\section{Mapping of trees}
-The information from the tags in Talbanken can be used for many xx.
-The tools for lexical extraction can be enhanced if we get more information
-about which form the word is currently used in. 
-We are currently working on a automatical mapping from Talbanken trees to trees in GF format.
-This would later enable us to extract possibilities for how often different functions
-are used, a feature (!) that would improve the parsing considerably.
-The evalution of the parser could also be accomplished by comparing the trees
-from the parser and from the mapping-process.
-brastartord there are differences in the notation of the sättet träden är uppbyggda på \\
- exempel?? \\
+The information given by this mapping may be used for the lexical extraction
+tools. Those can be enhanced if they are given more data about which form a word is
+currently used in. 
+The translated trees will also enable us to extract possibilities for how often
+different functions are used, a feature that would enable disambiguation.
+Furthermore, the translation makes it easy to identify grammatical constructions
+missing from the GF grammar and shows how the GF analysis differs from the one made
+in Talbanken.
 
-explation of the trees
-why we need the mapping : evaluation, adding words, finding missing constructions
-difficulties
+%>> by looking on how often a gf function is used
+%<< that would improve the parsing considerably.
+Another important use of the mapping is evaluation. By comparing the trees
+from the parser and from the transformer, we get an interesting way of
+accomplishing an evaluation.
+%An evaluation of the parser
+%could also be accomplished by  \\
+%<< brastartord there are differences in the notation of the sättet träden är uppbyggda på \\
+%>>maybe something about how it's going and why it's hard?
+%>> exempel?? \\
 
-\section{New things and hard things in the grammar}
-Due to the ressemblemences between the Scandinavian languages, they share about 80\% (?)
-of their resource code. This is good because blabla.
-However, for a bigger Swedish grammar, the implementation of the languages need to be more separated,
-in order to avoid ad-hoc solutions and constructions that are standard in
-Norweigan and Danish but very rare in Swedish. \\ %skriv om här änna
+%explation of the trees
+%why we need the mapping : evaluation, adding words, finding missing constructions
+%difficulties
+
+\subsection{Development of the grammar}
+%R: New things and hard things in the grammar" - this goes hand in
+%hand with more explanations about resource library and functors for
+%families, also emphasize the words that you refer to in the example -
+%bold font or smth equally noticeable. Explain more why these
+%constructions weren't there before.
+
+For multilingual GF applications, translation should always be possible, and
+therefore the abstract may only contain constructions that
+are common to all implemented languages.
+%>> resources for controlled language. does not aim for covering whole language
+%>> googla efter bra citat
+This means that everything added to the resources
+must be expressable in all other languages.
+Therefore, the resources have to be general and non language specific.
+Language specific constructs such as stylistic changes, idioms, informal
+expressions etc may be given in a special module.
+This module, together with the Swedish resource grammar has been the starting
+point of this project.
+
+Since many of the languages in the GF library reassemble each other grammatically,
+they can share much of their grammar implementations. This is usually done by using a
+\verb|Functor|, which lets a number of languages share parts of their implementation. 
+In addition to simply avoiding code duplication, this technique aids the code maintenance.
+For Swedish, about 85\% of the resource code is shared with the other Scandinavian
+languages.
+%due to the resemblances between the languages, they share
+%>> using a functor for sharing code, for when having the same structure but
+%>> but only different words. 
+For a bigger and still well-defined Swedish grammar however, the implementation of
+the languages needs to be more separated. % in order to avoid ad-hock solutions.
+%<<in order to avoid ad-hoc solutions and 
+Some constructions that are standard in
+Norwegian and Danish may be very rare in Swedish and while extending the
+grammar there may occur situations where a deeper restructuring of the
+implementation is needed. \\
 A number of grammatical constructions have been added the the Swedish grammar.
-Those include using the reflexive pronoun \emph{sitt}:\\
-\textbf{Han såg sitt hus} \emph{He saw his (own) house} \\
+%>> this is language specific, therefore not in common abstract
+Those include constructs that are generally not present in other languages,
+such as the reflexive pronoun \emph{sitt}: \\
+\emph{Han såg \textbf{sitt} hus} $\;$ (\emph{He saw his (own) house}) $\;$
 as opposed to \\
-\textbf{Han såg hans hus} \emph{He saw his (an other person's) house} \\
-% to simple for linguists?
-The possibility to put a part of a sentence in focus has also been added such as:\\
-\textbf{Glad var han inte} \emph{Happy was he not}
-Some ellipses are added, while there is work going on to add trickier ones (eller inte 
-ellipser just nu dårå, men vi skriver något fint här med, ja-jaaaa!).
+\emph{Han såg \textbf{hans} hus} $\;$ (\emph{He saw his (an other person's) house}) \\
+%>> an stylistic change that has been added, not needed for expressable new things (typ) 
+%>> not critical, does not have to be in resource grammar
+%>> also not expressable in all languages
+The possibility to put a part of a sentence in focus has also been added:\\
+\emph{Glad var han inte.} $\;$ (\emph{Happy was he not}). \\
+This sort of rephrasing which has little effect on the logical representation
+is commonly not given by the resource grammar.
 
+In order to keep the grammar as clean and to keep it from being too allowing,
+much care is taken to develop this as neatly as possible. 
+%<<Some ellipses are added, while there is work going on to add trickier ones (eller inte 
+%<<ellipser just nu dårå, men vi skriver något fint här med, ja-jaaaa!).
+
+%% Rephrase
+\section{Evaluation}
+%R: you need especially an evaluation part where you can also argue
+%that the method is better than statistics.
+Compared to a statistical parser which operates like a black box,
+a rule-based one is not only theoretically interesting,
+but could also give a more explanatory output since the rules
+are given their names by a human and hence acts as informative labels.
+Chosen parts of the parser or grammar may also be used in other applications,
+which may be dealing with controlled natural languages. Automatic language
+generation from the grammar is also provided by GF.
+
+We intend to evaluate the parser both automatically - by comparing the output
+the translated trees from Talbanken - and manually by an expert in Swedish grammar.
+
+%% more explaination of probabilities
 \section{Future Work}
-robust parser, ner, preprocessing, adding new words,
-probabilties
-ellipses, chunk parsing
 
+After having developed the grammar and lexicon, we plan to 
+make the parser robust by using techniques such as chunk parsing, named
+entity recognition, methods for handling unknown words and
+grammatical constructions, idioms, ellipses etc.
+
+It is already possible to add probabilities for GF functions,
+which serves as a ranking when several parse trees are possible.
+By adding dependency probabilities, we would like to improve the
+disambiguation.
+
+% ha med det som inte nämnts innan
+% ha en jämförelse med stat.parse?
+% är den deep?
 \section{Conclusion}
-hard but nice
-why better than statistical
+We intend to implement a robust deep parser for Swedish, by first developing
+a large scale grammar and then equip this with named entity recognition, a
+voluminous lexicon, statistical information for disambiguation,
+a method of parallel chunk parsing etc.
+So far we have been working on the grammar, the lexical resources and
+a translation of the treebank Talbanken to GF.
+The usage of Grammatical Framework gives us the advent age to start from
+a well-defined system of describing grammar, as well as tools for
+using the grammar in combination with programming languages like Haskell
+and Java.
+
+\section{Acknowledgments}
+Special thanks to Aarne Ranta, Ramona Enache, Elisabet Engdahl, Krasimir
+Angelov and Lars Borin for their help and support.
 
@@
 \bibliographystyle{eptcs}
-\bibliography{ourBib}
+\bibliography{ourbib}
 \end{document}

### 2011-10-10 doc/paper/Paper.tex
things

@@
 
-\title{First results of developing a ruled based Swedish parser} %A Wide-Coverage Swedish Grammar} % ska vi vara anonyma?
+\title{First results of developing a Swedish parser based on a wide coverage grammar} %A Wide-Coverage Swedish Grammar} % ska vi vara anonyma?
 \author{Malin Ahlberg
@@
 }
-\def\titlerunning{First results of developing a ruled based Swedish parser}
+\def\titlerunning{First results of developing a grammar based Swedish parser}
 \def\authorrunning{M. Ahlberg}
@@
 This paper describes an on-going work of a rule-based, open source
-parser for Swedish, based on a wide-coverage grammar implemented in
-Grammatical Framework (GF).
-In addition to GF, we use two other main resources; the Swedish treebank Talbanken
-and the electronic lexicon SALDO.
-The grammar will be highly reusable as GF provides libraries for compiling
+parser for Swedish. The central component will be a wide-coverage grammar
+implemented in the grammar formalism Grammatical Framework (GF). The
+resulting parser will be the first deep Swedish parser based on a grammar and
+will give a full syntactic analysis of the input sentences.
+In addition to GF, we use two other main resources; the Swedish
+treebank Talbanken and the electronic lexicon SALDO.
+% ha med detta?
+The grammar will be highly reusable, chosen 
+parts of the it can be used for applications dealing with controlled 
+or free natural language. GF also provides libraries for compiling
 grammars to a number of programming languages.
@@
 % 
-
+%>>
 %>> something very commercial, interesting and outstanding here please!
@@
 
+Swedish is a North Germanic language spoken by approximately 10 million people.
+It is a SVO language, and although the word order is relatively strict,
+inverted order is commonly used
+to indicate questions or to emphazise different parts of the sentence. \\
+Language technology for Swedish is an area of much interesting research.
+%on  are currently being conducted. \\
 %R: -- the second paragraph - OK but please extend.
-We would like to implement a wide coverage grammar and parser for Swedish.
-We thereby investigate how the grammar formalism Grammatical Framework,
-further described in section \ref{sec:gf}, can be used for open domain parsing.
+Our goal is to implement a wide coverage grammar and parser for Swedish.
+We thereby investigate how the grammar formalism Grammatical Framework
+%further described in section \ref{sec:gf}, 
+can be used for open domain parsing.
 By starting from an already existing Swedish grammar written in GF, we get 
@@
 tools such as parsing, generation and
-a well tested interpretation of the parse trees. There are also tools for using
-the grammar in a number of programming languages. 
+a well tested interpretation of the parse trees. Furthermore, there are tools
+for using the grammar in a number of programming languages. 
 %>> , tools for visualizing ...
@@
 
+This paper will briefly introduce GF, Talbanken and SALDO in section \ref{sec:background}
+and describe our current work in section \ref{sec:progress}.
 %We aim to implement % The goal is to be able to 
@@
 \section{Background}
+\label{sec:background}
 \subsection{Grammatical Framework}
@@
 
-Grammatical Framework\cite{gf} (GF) is a grammar formalism based on functional
+Grammatical Framework\cite{ranta-2011} (GF) is a grammar formalism based on functional
 programming, designed for multilingual grammar applications.
@@
 The key idea
-is to divide the grammar into abstract and concrete.
-The abstract grammar gives a logical representation of the semantics of the
-domain, modeled as abstract trees. Complications occurring in
+is to divide the grammar into abstract and concrete parts. %% parts?
+The abstract grammar gives a logical representation of the semantics,
+%of the domain, 
+modeled as abstract trees. Complications occurring in
 natural languages, such as agreement, case and word order, are abstracted away. 
@@
 
-% obra mening
 The grammar acts as an independent module and may be used in different
@@
 %R: more about the GF library - if you have space
-Reusability is further supported by the GF resource grammars\cite{resource},
+Reusability is further supported by the separation between resource grammars
+and appliciation grammars. The resource grammars is 
 a library provided with the GF package, 
@@
 syntactical rules for more than 20 different languages are implemented.
-As a result, an end user of the framework does not need to describe how to 
-form standard sentences, phrases or decline words, but can focus on the
-particularities for the given domain and choose what kind of 
-utterances that should be allowed.
+Hence a writer of a application grammar can start her work at a higher
+level and does not need to describe how to 
+form standard sentences, phrases or decline words.
+%, but can focus on the
+%particularities for the given domain and choose what kind of 
+%utterances that should be allowed.
 
@@
 % get inspired by Shafqat.
-GF has so far been used in a number of projects, MOLTO\cite{molto} and
-WebAlt\cite{webalt} to mention a few. 
+GF has so far been used in a number of projects, MOLTO\cite{molto}, TALK\cite{talk}
+and WebAlt\cite{webalt} to mention a few. 
 All those are special domain applications, dealing with controlled natural
 language.
-This project experiments on other usages of a GF grammar.
+This project experiments on other usage of a GF grammar - to use it for open
+domain language.
 
@@
 In 2005 it was modernized by Nivre, Nilsson and Hall\cite{talbanken05} and
-enriched with annotation for a
-full phrase structure analysis. 
-Talbanken contains both written and spoken Swedish. For this project only
-the prose material, consisting of 6316 sentences will be used.
+enriched with annotation for a full phrase structure analysis. 
+We are using the treebank for development and evaluation.
+
+Although Talbanken contains both written and spoken Swedish,
+%For this project 
+only the prose material, consisting of 6316 sentences will be used in this project.
 This part was also used when training the data-driven parser Maltparser \cite{malt}. \\
-We are using parts of Talbanken for testing, development and evaluation.
 %and the parsetree generated by our grammar will eventually be 
@@
 \section{Work in progress}
+\label{sec:progress}
 \subsection{A tool for lexical acquisition}
@@
 the \emph{smart paradigm} given in the Swedish resource grammar.
-The smart paradigm accepts verbs in present tense indicative form and decides
- which paradigm it most likely belong to. If needed, it also accepts more
-verb forms showing the correct inflection.
+The smart paradigm acts as functions that given one form of a word can
+infer which paradigm it most likely belong to.
+The smart paradigm for verbs accepts words in present tense indicative form.
+If needed, it also accepts more verb forms showing the correct inflection.
 %which Making use of the
@@
 Extract\cite{extract} for supervised lexicon extraction and
-FM\cite{fm} for programming lexical resources. Extract uses a similar 
-method but does not have access to the GF smart paradigm.
+FM\cite{fm} for programming lexical resources.
+%Extract uses a similar 
+%method but defines the paradigms itself.
+%rules are given in another format.
 
@@
 For multilingual GF applications, translation should always be possible, and
-therefore the abstract may only contain constructions that
+therefore the resource abstract may only contain constructions that
 are common to all implemented languages.
@@
 %>> googla efter bra citat
-This means that everything added to the resources
-must be expressable in all other languages.
+%This means that everything added to the resources must be expressable in all other languages.
 Therefore, the resources have to be general and non language specific.
@@
 For Swedish, about 85\% of the resource code is shared with the other Scandinavian
-languages.
+languages. % >> When trying to achieve a lightweight, usable code? Not
 %due to the resemblances between the languages, they share
@@
 %>> but only different words. 
-For a bigger and still well-defined Swedish grammar however, the implementation of
-the languages needs to be more separated. % in order to avoid ad-hock solutions.
+However, if we aim for a deeper and more covering analysis of Swedish,
+%For a bigger and still well-defined Swedish grammar however,
+the implementation of the languages needs to be more separated. 
+% in order to avoid ad-hock solutions.
 %<<in order to avoid ad-hoc solutions and 
-Some constructions that are standard in
-Norwegian and Danish may be very rare in Swedish and while extending the
-grammar there may occur situations where a deeper restructuring of the
-implementation is needed. \\
+%Some constructions that are standard in
+%Norwegian and Danish may be very rare in Swedish and while extending the
+%grammar there may occur situations where a deeper restructuring of the
+%implementation is needed. \\
 A number of grammatical constructions have been added the the Swedish grammar.
@@
 such as the reflexive pronoun \emph{sitt}: \\
-\emph{Han såg \textbf{sitt} hus} $\;$ (\emph{He saw his (own) house}) $\;$
+\emph{Han såg \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw his (own) house}) $\;$
 as opposed to \\
@@
 
-In order to keep the grammar as clean and to keep it from being too allowing,
+In order to keep the grammar as clean and to keep it from allowing syntactical errors,
 much care is taken to develop this as neatly as possible. 
@@
 %<<ellipser just nu dårå, men vi skriver något fint här med, ja-jaaaa!).
+
+\section{Related work}
+There are other parser for Swedish parser such as the the statistical MaltParser\cite{malt}
+which was trained on Talbanken. 
+CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is a  based on finite state cascades,
+whereas shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
+a context free grammar. Both GTA and CassSwe operates on POS tagged text.
+%deep Swedish Core Language Parser,
+%shallow Swedish Constraint Grammar.
 
@@
 make the parser robust by using techniques such as chunk parsing, named
-entity recognition, methods for handling unknown words and
-grammatical constructions, idioms, ellipses etc.
+entity recognition, methods for handling unknown 
+grammatical constructions such as idioms and ellipses. %etc.
 
@@
 \section{Acknowledgments}
-Special thanks to Aarne Ranta, Ramona Enache, Elisabet Engdahl, Krasimir
-Angelov and Lars Borin for their help and support.
+The work has been funded by Center of Language Technology.
+I would also like to give special thanks to Aarne Ranta, Ramona Enache,
+Elisabet Engdahl, Krasimir Angelov and Lars Borin for their help and support.
 

### 2011-10-11 doc/gullmarsstrand/pres.tex
work on paper

@@
 \usepackage{verbatim} 
-%% \usetheme{Pittsburgh}
+\usetheme{Singapore}
+%\usetheme{Pittsburgh}
+%\usetheme{Montpellier}
 \usepackage{colortbl}
@@
 \usepackage{cancel}
+\setbeamercolor{alerted text}{fg=lightgray,bg=white}
  \renewcommand{\baselinestretch}{1.5}
@@
 % John Hughes article How to give a good research presentation.
-\title{A Wide-Coverage Grammar and Parser for Swedish}
-\subtitle{\large First results and perspectives}
+\title{A Wide-Coverage Grammar for Swedish}
+%\subtitle{\large }
 \author{Malin Ahlberg \\ Gothenburg University}
@@
  \begin{frame}
+  \frametitle{Grammatical Framework}
+  \framesubtitle{Introduction}
+  A grammar formalism based on functional programming \\
+  \vspace{5mm}
+  \pause
+  One abstract grammar   \\
+  \pause
+  One concrete grammar for each language \\
+ \end{frame}
+
+ \begin{frame}
+  \frametitle{Grammatical Framework}
+  \framesubtitle{Resource grammars}
+  The resource grammars covers about 20 languages \\
+  \vspace{5mm}
+  \pause
+  Extra module for language specific constructions
+ \end{frame}
+
+\begin{frame}
+\frametitle{The project}
+\framesubtitle{} 
+A parser for Swedish using:
+\begin{itemize}
+\item{Grammatical Framework}
+\item{SALDO}
+\item{Talbanken}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{Current work}
@@
 A mapping of Talbanken05 trees to GF trees\\
+  \pause
 Extending the grammar \\
-Re-import saldo\\
+  \pause
+Re-import SALDO\\
 \end{frame}
@@
 \framesubtitle{Example} 
-Nice example
-\end{frame}
-
-\begin{frame}
-\frametitle{Mapping}
-\framesubtitle{Why?} 
-In order to \\
-\begin{itemize}
-\item{evaluate the parser}
-\item{identify grammatical constructions not current in the GF grammar}
-\item{enable good lexical extraction}
-\item{extract probabilites for GF functions}
-\end{itemize}
-% why it's hard (396 hur och hur)
-%goals
+\small{\textbf{Talbanken05}}\hspace{160pt}
+\small{\textbf{GF abstract tree}}
+\includegraphics[width=130pt,height=200pt]{taltree.pdf}
+\includegraphics[width=170pt,height=200pt]{gftree.pdf}
 \end{frame}
@@
 %\end{frame}
+\begin{frame}
+
+\frametitle{Mapping}
+\framesubtitle{Benefits} 
+\begin{itemize}
+\item{Evaluate the parser}
+\pause
+\item{Identify grammatical constructions not current in the GF grammar}
+\pause
+\item{Enable good lexical extraction}
+\pause
+\item{Extract probabilities for GF functions}
+\end{itemize}
+% why it's hard (396 hur och hur)
+%goals
+\end{frame}
 
@@
 \begin{tabular}{ll}
-\textbf{PassVP} & V2 $\rightarrow$ VP (should be VPSlash $\rightarrow$ VP) \\
-& \emph{äpplet åts},\emph{boken gavs till honom} \\ % \;
- %\emph{("äpplet blev ätet")}\\ % boken gavs till honom
+\textbf{ReflGenVP} & VPSlash  $\rightarrow$ ReflCN $\rightarrow$ VP \\
+& \emph{han ger sina små barn till dem} \\
+% \\ but not \emph{sig själv}, \emph{han äter alla sina äpplen} or \emph{han ger sig till sig}  \\
+  %relate to Peter - sig/sig själv, can't say hon ger sig till sig
 
-\textbf{ReflSlash} & VPSlash $\rightarrow$ VP \\
-& \emph{han ger sina små barn till dem} \\
+\alert{\textbf{PassVP}}
+  & \alert{V2 $\rightarrow$ VP} \\
+& \alert{\emph{äpplet åts},}\alert{\emph{boken gavs till honom}} \\
+
+\alert{\textbf{AdvVPSlash}} & \alert{VPSlash $\rightarrow$ Adv $\rightarrow$ VPSlash} \\
+& \alert{\emph{hon åt redan äpplet}, \emph{hon är redan här}}\\
+
+\alert{\textbf{PPartAP}} & \alert{V2 $\rightarrow$ AP} \\
+& \alert{\emph{det är skrivet},\emph{den skrivna artikeln}}
+\end{tabular}\\
+
+\end{frame}
+
+\begin{frame}[containsverbatim]
+%% vad ska det heta???
+\frametitle{Reflexive genitive verb phrase}
+\framesubtitle{Problems} 
+\begin{tabular}{ll}
+Can say:&  \emph{han ger den till sin lilla katt} \\
+ &        \emph{han ger sin egen katt till honom} \\
+\end{tabular} \\
+% fixa indentering
+%\begin{tabular}{ll}
+%\indent \small{\verb|ReflCN NumSg (AdjCN (PositA small_A) (UseN cat_N)))|} & \\
+%\end{tabular} \\
+
+\begin{tabular}{ll}
+But can not say: & \emph{alla sina barn} \\
+ & \emph{all sin undervisning} \\
+ & \emph{han ger sina pengar till sina barn}
+% & \emph{sin urgamla rätt att ...}\\  % N2V?
+\end{tabular} \\
+%hur löses det för vår? (inget problem för det hör ej ihop med subj)
+
+\end{frame}
+
+%kolla upp dessa i stort lexikon
+\begin{frame}
+ \renewcommand{\baselinestretch}{1.5}
+\frametitle{The grammar}
+\framesubtitle{Current work} 
+\begin{tabular}{ll}
+\alert{\textbf{ReflGenVP}} & \alert{VPSlash $\rightarrow$ ReflCN $\rightarrow$ VP} \\
+& \alert{\emph{han ger sina små barn till dem}} \\
 % \\ but not \emph{sig själv}, \emph{han äter alla sina äpplen} or \emph{han ger sig till sig}  \\
@@
 %% (RelNp "flickan , sådan att hon inte åt äpplen")
+{\textbf{PassVP}} & V2 $\rightarrow$ VP \\%  (should be VPSlash $\rightarrow$ VP) \\
+& \emph{äpplet åts} \\ % \;
+ %\emph{("äpplet blev ätet")}\\ % boken gavs till honom
+
+% CompAdv here, plus we need ComlVVAdv, Pass? osv
+\alert{\textbf{AdvVPSlash}} & \alert{VPSlash $\rightarrow$ Adv $\rightarrow$ VPSlash} \\
+& \alert{\emph{hon åt redan äpplet}, \emph{hon är redan här}}\\
+% &\emph{("hon åt äpplet redan")} \\
+
+\alert{\textbf{PPartAP}} & \alert{V2 $\rightarrow$ AP} \\
+& \alert{\emph{det är skrivet},\emph{den skrivna artikeln}}
+\end{tabular}\\
+
+\end{frame}
+
+\begin{frame}[containsverbatim]
+\frametitle{Passive Form}
+\framesubtitle{Problems} 
+Only accepts verbs of type \verb|V2| \\
+Cannot say \; \emph{Boken gavs till honom}\\
+The type should be \verb|VPSlash -> VP|, but verb phrases does not carry
+information about the passive form\\
+
+\end{frame}
+
+\begin{frame}
+\frametitle{The grammar}
+\framesubtitle{Current work} 
+\begin{tabular}{ll}
+\alert{\textbf{ReflGenVP}} & \alert{VPSlash $\rightarrow$ ReflCN $\rightarrow$ VP} \\
+& \alert{\emph{han ger sina små barn till dem}} \\
+
+ %\emph{("äpplet blev ätet")}\\ % boken gavs till honom
+\alert{\textbf{PassVP}}
+  & \alert{V2 $\rightarrow$ VP} \\ % (should be VPSlash $\rightarrow$ VP)} \\
+& \alert{\emph{äpplet åts},}\alert{\emph{boken gavs till honom}} \\
 
@@
 
-\textbf{PPartAP} & V2 $\rightarrow$ AP \\
-& \emph{det är skrivet},\emph{den skrivna artikeln}
+\alert{\textbf{PPartAP}} & \alert{V2 $\rightarrow$ AP} \\
+& \alert{\emph{det är skrivet},\emph{den skrivna artikeln}}
 \end{tabular}\\
-
-Easy sentences with known words (old testsuite) : 92 out of 118 (78\%) \\
 %list of functions
@@
 
-%kolla upp dessa i stort lexikon
 \begin{frame}
-\frametitle{Perfect Participle}
-\framesubtitle{What type of functions do we need?} 
-PPartAP cannot handle \emph{det är givet till henne} or \emph{gågna åren} \\
-VPSlash, V2, V?\\
-Is the word an adjective or verb? \\
-%attribut- predikativ
-\begin{tabular}{ll}
-Some examples: & \\
-\textbf{V} & \\
-& \emph{försvunnen}, \emph{slocknad} \\
-& \emph{i en gången tids svenska bondesamhälle}  \\% 4872    
-\end{tabular} \\
-\textbf{V2 or V?}
-\emph{en ökad effektiv information} \\
-\emph{en minskad benägenhet} \\
-\textbf{V2}
-\emph{de stegrade hyrorna} \\
-\emph{fyra representerade kyrkor} \\
-\emph{som aldrig är skalad (industriskalad, skalkokt)} \\
-\emph{eftertraktade lägenheter} \\
-\textbf{V3}
-\emph{skickad till mig}
-\textbf{Adj?}
-\emph{en fritt vuxen natur } \\
-\emph{centralt belägna} \\
-\emph{misstänkt } \\% (av)}
-\emph{en kvalificerad majoritet} \\
-\emph{kokt potatis} \\% (av)}
-?
-\emph{vår behandlade potatis} \\
-\emph{är oroad} \\
-\emph{beprövad}  \\ % (av)
-\emph{de levererar den doppad i askorbinsyra} \\
-\end{frame}
-
-\begin{frame}[containsverbatim]
-\frametitle{ReflGenVP}
-\framesubtitle{Problem} 
-\begin{tabular}{ll}
-Adjectives are ok: &  \emph{sin gula lilla katt} \\
-% fixa indentering
-& \small{\verb|ReflCN NumSg (AdjCN (PositA yellow\_A)|} \\
-&\small{\verb|(AdjCN (PositA small\_A) (UseN cat\_N)))|}\\
-\end{tabular} \\
-\begin{tabular}{ll}
-But predeterminers are not: & \emph{alla sina barn} \\
- & \emph{all vår undervisning} \\
- & \emph{sin urgamla rätt att ...}\\  % N2V?
-\end{tabular} \\
-%hur löses det för vår? (inget problem för det hör ej ihop med subj)
-
+\frametitle{Adverbs describing VPSlash}
+\framesubtitle{Example} 
+\small{\textbf{AdvVPSlash}}\hspace{160pt}
+\small{\textbf{AdvVP}}
+\includegraphics[width=170pt,height=200pt]{advVP.pdf}
+\includegraphics[width=170pt,height=200pt]{advVP2.pdf}
 \end{frame}
@@
 \begin{frame}
+\frametitle{The grammar}
+\framesubtitle{Current work} 
+\begin{tabular}{ll}
+\alert{\textbf{ReflGenVP}} & \alert{VPSlash $\rightarrow$ ReflCN $\rightarrow$ VP} \\
+& \alert{\emph{han ger sina små barn till dem}} \\
+
+ %\emph{("äpplet blev ätet")}\\ % boken gavs till honom
+\alert{\textbf{PassVP}}
+  & \alert{V2 $\rightarrow$ VP} \\ % (should be VPSlash $\rightarrow$ VP)} \\
+& \alert{\emph{äpplet åts},}\alert{\emph{boken gavs till honom}} \\
+
+% CompAdv here, plus we need ComlVVAdv, Pass? osv
+\alert{\textbf{AdvVPSlash}} & \alert{VPSlash $\rightarrow$ Adv $\rightarrow$ VPSlash} \\
+& \alert{\emph{hon åt redan äpplet}, \emph{hon är redan här}}\\
+% &\emph{("hon åt äpplet redan")} \\
+
+\textbf{PPartAP} & V2 $\rightarrow$ AP \\
+& \emph{det är skrivet},\emph{den skrivna artikeln}
+\end{tabular}\\
+\pause
+%Easy sentences with known words (old testsuite) : 92 out of 118 (78\%) \\
+%list of functions
+%examples
+%difficulties
+%sådan
+\end{frame}
+
+\begin{frame}
+ \renewcommand{\baselinestretch}{1.0}
+\frametitle{Perfect Participle}
+\framesubtitle{What type of functions do we need?} 
+PPartAP cannot yet handle \emph{det är givet till henne} or \emph{de gågna åren} \\
+% Adjective or verb? \\
+%attribut- predikativ
+\pause
+Some examples:
+\begin{itemize}
+\item{ \textbf{V} \small{\emph{försvunnen}}, \small{\emph{slocknad}}}
+\pause
+\item{\textbf{V2 or V?} \small{\emph{en ökad effektiv information}},
+\small{\emph{en minskad benägenhet}}} \\
+\pause
+\item{\textbf{V2}
+\small{\emph{de stegrade hyrorna}},
+\small{\emph{som aldrig är skalad}}, % (industriskalad, skalkokt)}} \\
+\small{\emph{eftertraktade lägenheter}}} \\
+\pause
+\item{\textbf{V3} \small{\emph{de är skickade till mig}}}
+\pause
+\item{\textbf{VS} \small{\emph{han blev ombedd att gå}}}
+\pause
+\item{\textbf{Adj?}
+\small{\emph{en fritt vuxen natur }},
+\small{\emph{centralt belägna}},
+\small{\emph{misstänkt }}}% \\% (av)}
+\end{itemize}
+\pause
+
+VPSlash, V2, V?\\
+%\emph{en kvalificerad majoritet} \\
+%\emph{kokt potatis}} \\% (av)}
+%& \emph{i en gången tids svenska bondesamhälle}  \\% 4872    
+%?
+%\emph{vår behandlade potatis} \\
+%\emph{är oroad} \\
+%\emph{beprövad}  \\ % (av)
+%\emph{de levererar den doppad i askorbinsyra} 
+\end{frame}
+
+\begin{frame}
+ \renewcommand{\baselinestretch}{1.0}
 \frametitle{The Grammar}
 \framesubtitle{Future work} 
-Reciprocals pronouns - \emph{varandra} \\
-Add category of words that can be used either as adjectives or determiners - \\
-   \emph{fler katter är där}, \emph{fler är där}, \emph{kattern är fler}\\ %sådana
-Category used as predeterminer or adverbs - \\
-  \emph{bara barn}, \emph{hon bara log} \\
-\emph{en sorts katt}, \emph{katter av en sort}
-den yngsta är gulast
-idioms
-minst, lika
-"Flickan som såg katten är här"
-\textbf{Cleft sentences} är ok!!!
-"Det är till uppträdande i denna miljön som utbildningen närmast syftar"
-"Det är då som det träder i funktion"
+Many words are treated differently in GF and Talbanken. \\
+\begin{tabular}{lll}
+& \textbf{Talbanken} & \textbf{GF} \\
+\emph{mycket} & Pronoun, adverb &  Determiner and adverb\\
+& &\emph{mycket blir bättre}, \emph{hon äter mycket}\\
+\pause
+\emph{bara} & Adverb & Predeterminer and AdV \\ %% nice name here!!
+& &\emph{bara barn är små}, \emph{hon bara log}, \emph{hon log bara} \\
+\end{tabular} \\
+\vspace{0.5cm}
+\pause
+Pronouns in Talbanken:
+\emph{fler},\emph{båda}\\
+   \emph{det finns fler upplysningar}, \emph{fler är där}, \emph{analfabeterna blir fler} %sådana
 \end{frame}
+
+\begin{frame}
+ \renewcommand{\baselinestretch}{1.0}
+\frametitle{The Grammar}
+\framesubtitle{Future work} 
+\begin{itemize}
+\item{Reciprocal pronouns - \emph{varandra }}
+\pause
+\item{\emph{Slags, sorts}: \emph{en sort\textbf{s} katt}, \emph{katter av en sort}}
+\pause
+\item{\emph{Den yngsta    är gladast}}
+\pause
+\item{\emph{De  arbetande är tröttast}}
+\pause
+\item{Idioms, AVP (adverb phrases)\\
+\emph{till exempel, i stället,i fortsättningen}}
+%\item{\emph{minst}, \emph{lika}}
+\end{itemize}
+\end{frame}
+
 \begin{frame}
 \frametitle{Goals}
-\framesubtitle{Goals for christmas} 
+\framesubtitle{Goals for Christmas} 
 \begin{itemize}
 \item{Finish and evaluate mapping}
+\pause
 \item{Extend grammar}
-    \item{New mapping of saldo - making it easy to keep ut-to-date}
+\pause
 \item{Evaluation}
+\pause
+\item{New mapping of SALDO - making it easy to keep up-to-date}
+\pause
+\item{?}
 \end{itemize}
@@
     \frametitle{The end}
-\huge\begin{center}Thanks for listening\end{center}
-\large\begin{center}Questions?\end{center}
+\large\begin{center}Thanks for listening\end{center}
 \end{frame}

### 2011-10-11 doc/paper/Paper.tex
work on paper

@@
 \usepackage[utf8]{inputenc}
-%\usepackage{breakurl}        
 \usepackage{graphicx}
 
-\title{First results of developing a Swedish parser based on a wide coverage grammar} %A Wide-Coverage Swedish Grammar} % ska vi vara anonyma?
+\title{First results of developing a Swedish parser based on a wide coverage grammar}
 \author{Malin Ahlberg
-% vad ska stå här eg?
-\institute{Center of Language Technology, Gothenburg University, Sweden}
-%\email{ahlberg.malin@gmail.com}
+\institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
 }
@@
 \begin{abstract}
-This paper describes an on-going work of a rule-based, open source
+This paper describes on-going work of a rule-based, open source
 parser for Swedish. The central component will be a wide-coverage grammar
-implemented in the grammar formalism Grammatical Framework (GF). The
+implemented in the formalism Grammatical Framework (GF). The
 resulting parser will be the first deep Swedish parser based on a grammar and
 will give a full syntactic analysis of the input sentences.
-In addition to GF, we use two other main resources; the Swedish
+In addition to GF, we use two other main resources: the Swedish
 treebank Talbanken and the electronic lexicon SALDO.
@@
 % allow easy interaction with other programs and grammars
-% 
-%>>
 %>> something very commercial, interesting and outstanding here please!
-
-%R: this phrase "We aim to implement a parser for open domain language
-%and this paper describes our work and results so far as well as the
-%directions of future work." doesn't belong in the abstract because it
-%just describes the normal structure of any paper, without saying why
-%your work are interesting/outstanding, worth-reading-further. Make it
-%sound more commercial, if you want to have it there.
 
@@
 \section{Introduction}
-%R: the first sentence - in "deep robust parser" - why "deep" - grammar
-%formalism, why "robust" - ??
-% afterwards, you don't actually have named-entity recognition, and
-%further more, there's no such thing as "exhaustive lexicon" - you can
-%always switch to another domain and find new words. Further on, the
-%work has nothing to do with statistitics so far, so you can't mention
-%as opening line.  The same goes for parallel chunk parsing.
-
-%R: the first sentence doesn't really make sense because you first
-%refer to a parser and only then to GF and a grammar, but without
-%giving an explanation of how making the grammar would give you a
-%parser for free.
-
 Swedish is a North Germanic language spoken by approximately 10 million people.
@@
 inverted order is commonly used
-to indicate questions or to emphazise different parts of the sentence. \\
+to indicate questions or to emphasize different parts of the sentence. \\
+%Olga Caprotti October 10, 2011 10:52 PM
+%you should have a sentence maybe on how old of a research area it is. "much interesting" is rather weak.
 Language technology for Swedish is an area of much interesting research.
 %on  are currently being conducted. \\
-%R: -- the second paragraph - OK but please extend.
 Our goal is to implement a wide coverage grammar and parser for Swedish.
 We thereby investigate how the grammar formalism Grammatical Framework
-%further described in section \ref{sec:gf}, 
 can be used for open domain parsing.
 By starting from an already existing Swedish grammar written in GF, we get 
-an fundamental description of the language. The framework also provides
+a fundamental description of the language. The framework also provides
 tools such as parsing, generation and
@@
 The parser will be evaluated both by experts and by comparing the results of
-parsing sentences in an often used Swedish treebank, Talbanken.
+parsing sentences in an commonly used Swedish treebank, Talbanken.
 
-This paper will briefly introduce GF, Talbanken and SALDO in section \ref{sec:background}
-and describe our current work in section \ref{sec:progress}.
+This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
+and describe our current work in Section \ref{sec:progress}.
 %We aim to implement % The goal is to be able to 
@@
 
-%R: the GF part - "especially suitable" - it's actually meant for the purpose
-%:)
-
 %compiler inspired approach, 
-% om resources: library that implements morphology and syntax rules, länk till papper om dessa
 % skriv ord som modular, architectur, reusable (fast stava dem rätt)
@@
 Grammatical Framework\cite{ranta-2011} (GF) is a grammar formalism based on functional
-programming, designed for multilingual grammar applications.
+programming, designed for multilingual applications.
 
-The key idea
-is to divide the grammar into abstract and concrete parts. %% parts?
+The key idea is to divide the grammar into abstract and concrete parts. %% parts?
 The abstract grammar gives a logical representation of the semantics,
-%of the domain, 
 modeled as abstract trees. Complications occurring in
 natural languages, such as agreement, case and word order, are abstracted away. 
-%R: what complications do you abstract away ?
-%categories and trees are not interesting as such, but their usage
-%is - why do you need them in the abstract syntax actually ?
 That is, the abstract grammar contains no actual linguistic information, but
@@
 
-To add a new languages means to give a concrete grammar. This implements
+To add a new language amounts to give a concrete grammar, that implements
 the abstract grammar and describes how to translate abstract
-trees to strings of the language. The framework enables us to parse strings into 
+trees into strings of the language. The framework enables us to parse strings into 
 abstract trees as well as linearize trees into strings.
-Since all languages supported in an application shares a common abstract syntax,
+Since all languages supported in an application share a common abstract syntax,
 we get multilinguality for free.
@@
 projects.
-%R: more about the GF library - if you have space
 Reusability is further supported by the separation between resource grammars
-and appliciation grammars. The resource grammars is 
-a library provided with the GF package, 
-where information about the morphological and
-syntactical rules for more than 20 different languages are implemented.
-Hence a writer of a application grammar can start her work at a higher
+and application grammars. The resource library provided with GF, 
+implements morphological and
+syntactical rules for more than 20 different languages.
+Hence a writer of an application grammar can start her work at a higher
 level and does not need to describe how to 
@@
 
-%so far bara used för vissa saker, 
-% about why we need more, extra module, separate swedish from scandinavian
-% get inspired by Shafqat.
 GF has so far been used in a number of projects, MOLTO\cite{molto}, TALK\cite{talk}
@@
 language.
-This project experiments on other usage of a GF grammar - to use it for open
-domain language.
+This project takes a different approach by using GF for open domain language.
+%experiments on other usage of a GF grammar - to use it for open
+%domain language.
 
@@
 \subsection{Talbanken}
-%Wherefrom, how big, who, when, version, used for maltparser
 Talbanken\cite{talbanken} is
-a Swedish treebank put together in the 1970s at Lund University.
+a Swedish treebank assembled in the 1970s at Lund University.
 In 2005 it was modernized by Nivre, Nilsson and Hall\cite{talbanken05} and
@@
 Although Talbanken contains both written and spoken Swedish,
-%For this project 
-only the prose material, consisting of 6316 sentences will be used in this project.
+only the prose material, consisting of 6316 sentences, will be used in this project.
 This part was also used when training the data-driven parser Maltparser \cite{malt}. \\
@@
 \subsection{Saldo}
-%R: Saldo part - why would you like to reimport ? "enlarge and enhance"
-%- how do you enhance it ?
 SALDO\cite{saldo} is an open source lexicon resource
-based on Svenskt Associationslexikon (SAL). It is
+based on Svenskt Associationslexikon. It is
 developed at Språkbanken at Gothenburg University
@@
 research. 
-% how it is tagged, other usages, benefits,
-% how easy it is to map it to gf
-From SALDO, a large GF lexicon
-has earlier been extracted,
-containing 50 000 entries.
-By updating this technique, we would like
-to reimport SALDO, which has since then been developed.
-%and we would like
-%to reimport the it using the same
-%updated techniques. %in order to enlarge our lexicon,
-%we  in order to enlarge and
-%GF version. 
+From SALDO, a large GF lexicon has earlier been extracted,
+containing 50 000 entries. By updating this technique, we want
+to reimport SALDO, which has been developed since.
 The importing method should be fast and reliable
-enough to be allow us to always have a fresh version of the dictionary
+enough to allow us to always have a fresh version of the dictionary
 in GF.
@@
 \subsection{A tool for lexical acquisition}
-%R: lexicon extraction part - more about everything, comparison with
-%Extract and Functional Morphology, evaluation of the lexicon builder -
-%"supervised lexicon extraction"(common term for this - google it if
-%you want to find related work)
-
 To enlarge the lexicon, 
@@
 The smart paradigm acts as functions that given one form of a word can
-infer which paradigm it most likely belong to.
-The smart paradigm for verbs accepts words in present tense indicative form.
+infer which paradigm it most likely belongs to.
+For verbs, the paradigm accepts words in present tense indicative form.
 If needed, it also accepts more verb forms showing the correct inflection.
-%which Making use of the
 
 By combining this method with the information from the tags in Talbanken,
-%and the paradigms in the Swedish resource grammar, 
 the tool interactively generates GF lexicons. 
 Given a list of words, it iteratively
-tries to figure out how to conjugate each of them. If several forms of a word is 
+tries to figure out how to conjugate each of them. If several forms of a word are 
 given, the program will try to identify the one that carries the most linguistic
 information, put this in a form recognized by the smart paradigm and ask GF to output
-a table with the resulting inflection table. 
-If this table contains all other forms from the input list,
+a table with the resulting inflection. 
+If the table contains all other conjugations from the input list,
 the program will ask the  user to
 validate the claimed conjunction. The user may now either
-allow the word to be added to the lexicon, remove it or demand the program
-to make another guess.
+allow the word to be added to the lexicon, remove it or request another guess.
 Although using simple techniques, the tool 
-manage to correctly guess 70-75\% of the given lemmas when tested on
-verbs. \\
+manages to correctly guess 70-75\% of the given lemmas when tested on
+verbs.\\
 For further extractions, there are also earlier developed tools such as
-Extract\cite{extract} for supervised lexicon extraction and
-FM\cite{fm} for programming lexical resources.
+Extract for supervised lexicon extraction and
+FM\cite{MarkusForsberg2007} for programming lexical resources.
 %Extract uses a similar 
@@
 
-The information given by this mapping may be used for the lexical extraction
-tools. Those can be enhanced if they are given more data about which form a word is
-currently used in. 
+The information given by this mapping, namely data about which form a word is
+currently used in, may be used for the lexical extraction
+tools.% Those can be enhanced if they are given more data. 
 The translated trees will also enable us to extract possibilities for how often
@@
 
-%>> by looking on how often a gf function is used
-%<< that would improve the parsing considerably.
-Another important use of the mapping is evaluation. By comparing the trees
-from the parser and from the transformer, we get an interesting way of
+Another important use of the mapping is evaluation. Comparing the trees
+from the parser and from the transformer, is an interesting way of
+% an straightforward
 accomplishing an evaluation.
-%An evaluation of the parser
-%could also be accomplished by  \\
-%<< brastartord there are differences in the notation of the sättet träden är uppbyggda på \\
-%>>maybe something about how it's going and why it's hard?
-%>> exempel?? \\
 
-%explation of the trees
-%why we need the mapping : evaluation, adding words, finding missing constructions
-%difficulties
+% add evaluation
 
 \subsection{Development of the grammar}
-%R: New things and hard things in the grammar" - this goes hand in
-%hand with more explanations about resource library and functors for
-%families, also emphasize the words that you refer to in the example -
-%bold font or smth equally noticeable. Explain more why these
-%constructions weren't there before.
-
-For multilingual GF applications, translation should always be possible, and
-therefore the resource abstract may only contain constructions that
-are common to all implemented languages.
-%>> resources for controlled language. does not aim for covering whole language
-%>> googla efter bra citat
-%This means that everything added to the resources must be expressable in all other languages.
-Therefore, the resources have to be general and non language specific.
+For multilingual GF applications, translation should always be possible.
+%, and therefore the resource abstract may only contain constructions that
+%are common to all implemented languages.
+Therefore, the resources have to be general and language independent.
 Language specific constructs such as stylistic changes, idioms, informal
@@
 
-Since many of the languages in the GF library reassemble each other grammatically,
+Since many of the languages in the GF library resemble each other grammatically,
 they can share much of their grammar implementations. This is usually done by using a
@@
 However, if we aim for a deeper and more covering analysis of Swedish,
-%For a bigger and still well-defined Swedish grammar however,
-the implementation of the languages needs to be more separated. 
-% in order to avoid ad-hock solutions.
+the implementation of the languages needs to be more independent.
 %<<in order to avoid ad-hoc solutions and 
@@
 %implementation is needed. \\
-A number of grammatical constructions have been added the the Swedish grammar.
-%>> this is language specific, therefore not in common abstract
-Those include constructs that are generally not present in other languages,
+A number of grammatical constructions have been added the the Swedish grammar
+include constructs that are generally not present in other languages,
 such as the reflexive pronoun \emph{sitt}: \\
@@
 \emph{Han såg \textbf{hans} hus} $\;$ (\emph{He saw his (an other person's) house}) \\
-%>> an stylistic change that has been added, not needed for expressable new things (typ) 
 %>> not critical, does not have to be in resource grammar
@@
 The possibility to put a part of a sentence in focus has also been added:\\
-\emph{Glad var han inte.} $\;$ (\emph{Happy was he not}). \\
-This sort of rephrasing which has little effect on the logical representation
+\emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
+This sort of rephrasing, which has little effect on the logical representation,
 is commonly not given by the resource grammar.
 
-In order to keep the grammar as clean and to keep it from allowing syntactical errors,
-much care is taken to develop this as neatly as possible. 
-%<<Some ellipses are added, while there is work going on to add trickier ones (eller inte 
-%<<ellipser just nu dårå, men vi skriver något fint här med, ja-jaaaa!).
+In order to keep the grammar clean and to keep it syntactically correct,
+much care is taken to develop it as neatly as possible.
 
 \section{Related work}
-There are other parser for Swedish parser such as the the statistical MaltParser\cite{malt}
+There are other parsers for Swedish such as the statistical MaltParser\cite{malt}
 which was trained on Talbanken. 
-CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is a  based on finite state cascades,
-whereas shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
-a context free grammar. Both GTA and CassSwe operates on POS tagged text.
-%deep Swedish Core Language Parser,
-%shallow Swedish Constraint Grammar.
+CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
+whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
+a context free grammar. Both GTA and CassSwe operate on POS tagged text.
 
-%% Rephrase
 \section{Evaluation}
-%R: you need especially an evaluation part where you can also argue
-%that the method is better than statistics.
 Compared to a statistical parser which operates like a black box,
@@
 but could also give a more explanatory output since the rules
-are given their names by a human and hence acts as informative labels.
+are given their names by a human and hence act as informative labels.
 Chosen parts of the parser or grammar may also be used in other applications,
-which may be dealing with controlled natural languages. Automatic language
+which deal with controlled natural languages. Automatic language
 generation from the grammar is also provided by GF.
 
-We intend to evaluate the parser both automatically - by comparing the output
+We intend to evaluate the parser both automatically - by comparing the output of
 the translated trees from Talbanken - and manually by an expert in Swedish grammar.
@@
 make the parser robust by using techniques such as chunk parsing, named
-entity recognition, methods for handling unknown 
+entity recognition and methods for handling unknown 
 grammatical constructions such as idioms and ellipses. %etc.
@@
 It is already possible to add probabilities for GF functions,
-which serves as a ranking when several parse trees are possible.
-By adding dependency probabilities, we would like to improve the
+which rank the parse trees.
+By adding dependency probabilities, we aim to improve the
 disambiguation.
 
-% ha med det som inte nämnts innan
-% ha en jämförelse med stat.parse?
-% är den deep?
+% ha med inte det som inte nämnts innan
 \section{Conclusion}
@@
 a translation of the treebank Talbanken to GF.
-The usage of Grammatical Framework gives us the advent age to start from
+% adventage to what??
+The usage of Grammatical Framework gives us the advantage to start from
 a well-defined system of describing grammar, as well as tools for
@@
 I would also like to give special thanks to Aarne Ranta, Ramona Enache,
-Elisabet Engdahl, Krasimir Angelov and Lars Borin for their help and support.
+Elisabet Engdahl, Krasimir Angelov, Olga Caprotti and Lars Borin for their help
+and support.
 

### 2011-10-12 doc/paper/Paper.tex
paper!

@@
 
-\title{First results of developing a Swedish parser based on a wide coverage grammar}
-\author{Malin Ahlberg
+\title{First results of developing a Swedish parser based on a wide-coverage grammar}
+\author{Malin Ahlberg \& Ramona Enache
+%\and
+%Co Author \qquad\qquad 
 \institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
@@
 \begin{abstract}
-This paper describes on-going work of a rule-based, open source
+This paper describes on going work of a rule-based, open-source
 parser for Swedish. The central component will be a wide-coverage grammar
 implemented in the formalism Grammatical Framework (GF). The
-resulting parser will be the first deep Swedish parser based on a grammar and
+resulting parser will be the first deep Swedish grammar-based parser  and
 will give a full syntactic analysis of the input sentences.
@@
 The grammar will be highly reusable, chosen 
-parts of the it can be used for applications dealing with controlled 
-or free natural language. GF also provides libraries for compiling
+parts of the it can be used in applications dealing with
+%controlled or free
+natural language. GF also provides libraries for compiling
 grammars to a number of programming languages.
@@
 \section{Introduction}
-Swedish is a North Germanic language spoken by approximately 10 million people.
+Swedish is a North-Germanic language spoken by approximately 10 million people.
 It is a SVO language, and although the word order is relatively strict,
 inverted order is commonly used
-to indicate questions or to emphasize different parts of the sentence. \\
-%Olga Caprotti October 10, 2011 10:52 PM
-%you should have a sentence maybe on how old of a research area it is. "much interesting" is rather weak.
-Language technology for Swedish is an area of much interesting research.
-%on  are currently being conducted. \\
-Our goal is to implement a wide coverage grammar and parser for Swedish.
-We thereby investigate how the grammar formalism Grammatical Framework
-can be used for open domain parsing.
-By starting from an already existing Swedish grammar written in GF, we get 
-a fundamental description of the language. The framework also provides
-tools such as parsing, generation and
-a well tested interpretation of the parse trees. Furthermore, there are tools
-for using the grammar in a number of programming languages. 
-%>> , tools for visualizing ...
-
-From the GF grammar we aim to implement a robust parser. 
+in subordinate clauses, to indicate questions or to emphasize different parts of the sentence. \\
+Our goal is to implement a wide coverage grammar and parser for Swedish
+using the grammar formalism GF. 
+% better 
+We implement a Swedish grammar and use GF to compile it to a parser.
+From this, we aim to implement a robust parser. 
+We thereby investigate how GF
+can be used for open-domain parsing.
 The parser will be evaluated both by experts and by comparing the results of
@@
 
-%compiler inspired approach, 
-% skriv ord som modular, architectur, reusable (fast stava dem rätt)
+%städa upp detta stycke!!
 
@@
 The abstract grammar gives a logical representation of the semantics,
+% shorten this part
 modeled as abstract trees. Complications occurring in
@@
 information about what categories are used and how to combine them into trees.
-
-To add a new language amounts to give a concrete grammar, that implements
+To add a new language it amounts to give a concrete grammar, that implements
 the abstract grammar and describes how to translate abstract
@@
 abstract trees as well as linearize trees into strings.
-Since all languages supported in an application share a common abstract syntax,
-we get multilinguality for free.
+%Since all languages supported in an application share a common abstract syntax,
+%we get multilinguality for free.
 
@@
 form standard sentences, phrases or decline words.
-%, but can focus on the
-%particularities for the given domain and choose what kind of 
-%utterances that should be allowed.
+
+Since many of the languages in the GF library resemble each other grammatically,
+they can share much of their grammar implementations. This is usually done by using a
+\verb|Functor|, which lets a number of languages share parts of their implementation. 
+In addition to simply avoiding code duplication, this technique aids the code maintenance.
 
@@
 This project takes a different approach by using GF for open domain language.
-%experiments on other usage of a GF grammar - to use it for open
-%domain language.
-
-% parsing free language, more work is required and it has been investigated what
-%Further, other tools that are needed have been implemented and examined.
-%will be needed to make it sufficient and a start on developing tools has beenmade.
+While the resources have to be general and language independent,
+language specific constructs such as stylistic changes, idioms, informal
+expressions are given in a special module.
+This module, together with the Swedish resource grammar has been our starting
+point. % of this project.
+%By starting from an already existing Swedish grammar written in GF,
+We get a basic description of the language and the framework provides
+tools such as parsing, generation and
+a well-tested interpretation of the parse trees. Furthermore, there are tools
+for using the grammar in a number of programming languages like Haskell
+and Java.
 
@@
 To enlarge the lexicon, 
-a tool for automatically acquisition has been created. It has 
-been tested on verbs with good results. It makes use of
+we have created a tool for automatic acquisition. We have
+tested it on verbs with good results. It makes use of
 the \emph{smart paradigm} given in the Swedish resource grammar.
-The smart paradigm acts as functions that given one form of a word can
+The smart paradigm is a function that given one form of a word can
 infer which paradigm it most likely belongs to.
@@
 verbs.\\
-For further extractions, there are also earlier developed tools such as
-Extract for supervised lexicon extraction and
-FM\cite{MarkusForsberg2007} for programming lexical resources.
 %Extract uses a similar 
@@
 tools.% Those can be enhanced if they are given more data. 
-The translated trees will also enable us to extract possibilities for how often
+The translated trees will also enable us to extract probabilities for how often
 different functions are used, a feature that would enable disambiguation.
@@
 
-Another important use of the mapping is evaluation. Comparing the trees
-from the parser and from the transformer, is an interesting way of
-% an straightforward
-accomplishing an evaluation.
+Another important use of the mapping is evaluation of the parser, which can be
+accomplished by comparing the parse trees and the trees from the transformer.
+%is an interesting way of
+%% an straightforward
+%accomplishing an evaluation.
 
@@
 \subsection{Development of the grammar}
-For multilingual GF applications, translation should always be possible.
-%, and therefore the resource abstract may only contain constructions that
-%are common to all implemented languages.
-Therefore, the resources have to be general and language independent.
-Language specific constructs such as stylistic changes, idioms, informal
-expressions etc may be given in a special module.
-This module, together with the Swedish resource grammar has been the starting
-point of this project.
-
-Since many of the languages in the GF library resemble each other grammatically,
-they can share much of their grammar implementations. This is usually done by using a
-\verb|Functor|, which lets a number of languages share parts of their implementation. 
-In addition to simply avoiding code duplication, this technique aids the code maintenance.
-For Swedish, about 85\% of the resource code is shared with the other Scandinavian
+For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
 languages. % >> When trying to achieve a lightweight, usable code? Not
@@
 \emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
-This sort of rephrasing, which has little effect on the logical representation,
-is commonly not given by the resource grammar.
+Emphasizing words or phrases by putting them in the beginning of a sentence
+is commonly used in Swedish.
+This sort of rephrasing is however not given by the resource grammar, since it
+has little effect on the logical representation.
 
@@
 
+%compare!!
 \section{Related work}
-There are other parsers for Swedish such as the statistical MaltParser\cite{malt}
-which was trained on Talbanken. 
+Language technology for Swedish is an area of much interesting research.
+There are tools for lexical extractions, such as
+Extract for supervised lexicon extraction and
+FM\cite{MarkusForsberg2007} for programming lexical resources.
+Among other parsers for Swedish, the statistical MaltParser\cite{malt}
+which was trained on Talbanken, is worth mentioning. 
 CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
@@
 
-\section{Evaluation}
-Compared to a statistical parser which operates like a black box,
-a rule-based one is not only theoretically interesting,
-but could also give a more explanatory output since the rules
-are given their names by a human and hence act as informative labels.
+\section{Evaluation and future work}
+%Compared to a statistical parser which operates like a black box,
+%a rule-based one is not only theoretically interesting,
+%but could also give a more explanatory output since the rules
+%are given their names by a human and hence act as informative labels.
+We are building a parser which is based on grammatical rules rather than
+statistics. The parser will therefore validate that
+a given sentence is grammatically correct according to the rules defined.
 Chosen parts of the parser or grammar may also be used in other applications,
@@
 
-We intend to evaluate the parser both automatically - by comparing the output of
-the translated trees from Talbanken - and manually by an expert in Swedish grammar.
-
 %% more explaination of probabilities
-\section{Future Work}
-
 After having developed the grammar and lexicon, we plan to 
@@
 grammatical constructions such as idioms and ellipses. %etc. 
-
 It is already possible to add probabilities for GF functions,
@@
 
-% ha med inte det som inte nämnts innan
+We intend to evaluate the parser both automatically - by comparing the output of
+the translated trees from Talbanken - and manually by an expert in Swedish grammar.
+Also the intermediate results are evaluated by our expert.
+
 \section{Conclusion}
 We intend to implement a robust deep parser for Swedish, by first developing
-a large scale grammar and then equip this with named entity recognition, a
+a large scale-grammar. The grammar can be compiled into a
+parser by GF, and we will equip this with named entity recognition, a
 voluminous lexicon, statistical information for disambiguation,
@@
 a well-defined system of describing grammar, as well as tools for
-using the grammar in combination with programming languages like Haskell
-and Java.
+parsing.
 
@@
 The work has been funded by Center of Language Technology.
-I would also like to give special thanks to Aarne Ranta, Ramona Enache,
+We would also like to give special thanks to Aarne Ranta, % Ramona Enache,
 Elisabet Engdahl, Krasimir Angelov, Olga Caprotti and Lars Borin for their help

### 2011-10-13 doc/paper/Paper.tex
paper, evaluation, mapping, plannig.

@@
 
-\title{First results of developing a Swedish parser based on a wide-coverage grammar}
+\title{First results of developing a Swedish parser based on a wide-coverage grammar\\
+        eller Combining resources into a Swedish parser
+        eller Bootstrapping a Swedish parser \\
+        eller A grammar-based Swedish parser from already existing resources}
 \author{Malin Ahlberg \& Ramona Enache
@@
 parser for Swedish. The central component will be a wide-coverage grammar
-implemented in the formalism Grammatical Framework (GF). The
-resulting parser will be the first deep Swedish grammar-based parser  and
-will give a full syntactic analysis of the input sentences.
+implemented in the formalism Grammatical Framework (GF), from which we 
+derive?? a deep parser.
 In addition to GF, we use two other main resources: the Swedish
 treebank Talbanken and the electronic lexicon SALDO.
+The resulting parser will 
+will give a full syntactic analysis of the input sentences.
+Since all the parser will be freely available, 
 % ha med detta?
-The grammar will be highly reusable, chosen 
-parts of the it can be used in applications dealing with
+it will be highly reusable and chosen parts of the the grammar can be used in
+applications dealing with
 %controlled or free
@@
 
+%compare!!
+\section{Related work}
+Language technology for Swedish is an area of much interesting research.
+There are tools for lexical extractions, such as
+Extract for supervised lexicon extraction and
+FM\cite{MarkusForsberg2007} for programming lexical resources.
+There are two other deep grammar-based parser for Swedis,
+SICS (Manny Raynen, ) and ? (Mats Wirén,?). Unlike our parser, 
+they are not freely commercialized.
+Among other parsers for Swedish, the statistical MaltParser\cite{malt}
+which was trained on Talbanken, is worth mentioning. 
+CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
+whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
+a context free grammar. Both GTA and CassSwe operate on POS tagged text.
+
 \section{Work in progress}
@@
 
-%compare!!
-\section{Related work}
-Language technology for Swedish is an area of much interesting research.
-There are tools for lexical extractions, such as
-Extract for supervised lexicon extraction and
-FM\cite{MarkusForsberg2007} for programming lexical resources.
-Among other parsers for Swedish, the statistical MaltParser\cite{malt}
-which was trained on Talbanken, is worth mentioning. 
-CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
-whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
-a context free grammar. Both GTA and CassSwe operate on POS tagged text.
-
 \section{Evaluation and future work}

### 2011-10-13 mapping/Convert.hs

@@ -47,11 +47,11 @@ mainEt src
                                 [withInputEncoding utf8
                                 , withRemoveWS yes] src
 	        >>>
-	        arrIO mappEvaluations --simplify --process   
+	        arrIO mappEvaluations --simplify --process  -- (return . take 100)  
 	        >>>
             xpickleDocument xpSentences
                             [withIndent yes,
-                             withInputEncoding utf8] "evaluation")
+                             withInputEncoding utf8] "TalbankenBeg.xml")
 
       return ()
 
@@ -66,13 +66,13 @@ simplify = arrIO (\x -> do putStrLn (show x)
                 DT.trace (word w++pos w) $ W (XMLHelp.id w) simpleW (pos w) 
 
 mappEvaluations x =
-  do pgf <- readPGF "../gf/BigTest.pgf"
+  do pgf <- readPGF "../gf/Big.pgf"
      let morpho  = buildMorpho pgf language
-         Just language = readLanguage "BigTestSwe"
+         Just language = readLanguage "BigSwe"
          nice  = filter (not . hasBadTag) x
          short = sortBy (comparing ws) nice
          good  = map (replaceWords morpho) short 
-     return $ drop 400 $ take 500 good
+     return $ take 100 nice --drop 400 $ take 500 good
  where replaceWords morpho s = 
           Sent (idS s) (rootS s) (map (g morpho) (XMLHelp.words s)) (info s) (ws s)
        g morpho w = let smallW = map toLower (word w) 
@@ -95,9 +95,11 @@ extractSentences = do
   ss <- mainF talbanken (return . map getSentence')
   return $ concat ss
 
-badTags = ["NAC","XP","AVP","CAP","CAVP","CNP","CONJP","CPP","CS","CVP","CXP",
-            "CJ","XX","XT","XF","XA","DB","IC","ID","IG","IQ","IR","IS","IT",
-            "ET","UK","++"]
+badTags = ["NAC","XP","AVP","PU"-- ,"CAP","CAVP","CNP","CONJP","CPP","CS","CVP","CXP",
+           -- "CJ"
+           ,"XX"--,"XT","XF","XA","DB",
+           ,"IC","IG","IQ","IR","IS","IT"]
+            --,"ID","ET","UK","++"]
 isBad = any (`elem` ["*"]) . snd
 prettyPrint (id,s) = id++"\t"++ Data.List.unwords s
 

### 2011-10-13 mapping/translate.hs

@@ -44,12 +44,12 @@ type PMonad = (RWS () [String] S)
 
 $(mkLabels [''S])
 
-test = True
+test = False
 usePGF = testGr
 testGr = ("../gf/BigTest.pgf","BigTestSwe")
-bigGr  = ("../gf/BigNew.pgf","BigNewSwe")
+bigGr  = ("../gf/Big.pgf","BigSwe")
 lang   = fromJust $ readLanguage "BigTestSwe"
-paint  = False
+paint  = True
 
 startState :: S 
 startState = S {_isReflGenVP = False, _isExist = False
@@ -62,8 +62,10 @@ mapp f    = main' f >> return ()
 main      = main' "test.xml" >> return ()
 bigTest   = main' "../testSuites/testShortSimpleTwo.xml" 
             >>= writeFile "mappingShort6.txt" . unlines
-evaluation = main' "EvalMappSuite2.xml"
-             >>= writeFile "Evalresult.txt" . unlines
+evaluation = evaluations "EvalMappSuite2.xml" "Evalresult.txt" 
+evaluations test to = 
+    main' test 
+    >>= writeFile to . unlines 
 main2      = main' "test2.xml" >> return ()
 mainTest   = main' "testSimple.xml" >>= putStrLn . compareRes
 mainT2     = main' "testSimple.xml" >>= putStrLn . unlines
@@ -87,10 +89,11 @@ main' fil  = do
       when test $ putStrLn $ unlines trace
       hPutStrLn stdout (showExpr [] e)
       when paint $ do
-        writeFile "tmp_tree.dot" (graphvizAbstractTree pgf (True,False) e) --(graphvizParseTree pgf lang e) 
+        writeFile "tmp_tree.dot" --(graphvizAbstractTree pgf (True,False) e)
+                                   (graphvizParseTree pgf lang e) 
         rawSystem "dot" ["-Tpdf", "tmp_tree.dot"
-                        --, "-otrees/tree"++showAlign l'++".pdf"]
-                        , "-o"++dropExtension fil++"GF.pdf"]
+                        , "-otrees/tree"++showAlign l'++"GFparsX.pdf"]
+                        --, "-o"++dropExtension fil++"GF.pdf"]
         return ()
       hPutStrLn stderr (show ((fromIntegral cn' / fromIntegral co') * 100))
       --return (showExpr [] e)
@@ -104,16 +107,17 @@ main' fil  = do
[…]

### 2011-10-13 doc/paper/Paper.tex
new version of paper

--- 2011-10-12 doc/paper/Paper.tex
@@
 
-\title{First results of developing a Swedish parser based on a wide-coverage grammar}
+\title{First results of developing a Swedish parser based on a wide-coverage grammar\\
+        or Combining resources into a grammar-driven Swedish parser \\
+        or Bootstrapping a Swedish parser \\
+        or A grammar-based Swedish parser from already existing resources}
 \author{Malin Ahlberg \& Ramona Enache
@@
 This paper describes on going work of a rule-based, open-source
-parser for Swedish. The central component will be a wide-coverage grammar
-implemented in the formalism Grammatical Framework (GF). The
-resulting parser will be the first deep Swedish grammar-based parser  and
-will give a full syntactic analysis of the input sentences.
+parser for Swedish. The central component is a wide-coverage grammar
+implemented in the formalism Grammatical Framework (GF), which
+can be compiled to a % from which we derive?? 
+parser.
 In addition to GF, we use two other main resources: the Swedish
 treebank Talbanken and the electronic lexicon SALDO.
-% ha med detta?
-The grammar will be highly reusable, chosen 
-parts of the it can be used in applications dealing with
+The resulting parser will 
+give a full syntactic analysis of the input sentences.
+It will be highly reusable, freely available
+and as GF provides libraries for compiling
+grammars to a number of programming languages,
+chosen parts of the the 
+grammar may be used in various NLP applications. % dealing with
 %controlled or free
-natural language. GF also provides libraries for compiling
-grammars to a number of programming languages.
+%natural language. 
 % allow easy interaction with other programs and grammars
@@
 Swedish is a North-Germanic language spoken by approximately 10 million people.
-It is a SVO language, and although the word order is relatively strict,
-inverted order is commonly used
-in subordinate clauses, to indicate questions or to emphasize different parts of the sentence. \\
-Our goal is to implement a wide coverage grammar and parser for Swedish
-using the grammar formalism GF. 
+It is a verb-second language: the second constituent of a declarative main
+clause must consist of a verb.
+The first constituent of the clause is usually made up by the subject,
+although it likewise could consist of adverbial phrases or objects.
+In questions and subordinate clauses, inverted word order occurs.
+
+Our goal is to implement a wide-coverage grammar and parser for Swedish
+using the grammar formalism GF, and 
+thereby investigate how GF can be used for open-domain parsing.
 % better 
-We implement a Swedish grammar and use GF to compile it to a parser.
-From this, we aim to implement a robust parser. 
-We thereby investigate how GF
-can be used for open-domain parsing.
-The parser will be evaluated both by experts and by comparing the results of
-parsing sentences in an commonly used Swedish treebank, Talbanken.
+We compile a large-scale grammar to a parser
+and combine it with the extensive lexicon SALDO. The Swedish treebank 
+Talbanken provides manually tagged trees which we use for improving and evaluating
+the grammar. The parser will additionally
+be evaluated by experts. \\
+%s and by comparing the results of parsing sentences in a commonly used Swedish treebank, Talbanken.
+We choose our resources so that both our grammar and parser can 
+be freely available and open-source.
 
 This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
-and describe our current work in Section \ref{sec:progress}.
+and describe our results so far in Section \ref{sec:progress}.
 %We aim to implement % The goal is to be able to 
@@
 Grammatical Framework\cite{ranta-2011} (GF) is a grammar formalism based on functional
-programming, designed for multilingual applications.
+programming. %, designed for multilingual applications.
 
@@
 % shorten this part
-modeled as abstract trees. Complications occurring in
-natural languages, such as agreement, case and word order, are abstracted away. 
-That is, the abstract grammar contains no actual linguistic information, but
-information about what categories are used and how to combine them into trees.
-To add a new language it amounts to give a concrete grammar, that implements
-the abstract grammar and describes how to translate abstract
-trees into strings of the language. The framework enables us to parse strings into 
+modeled as abstract trees.
+%Complications occurring in natural languages, such as agreement, case and word
+%order, are abstracted away. 
+%That is, 
+%It contains no actual linguistic information, but
+%information about what categories are used and how to combine them into trees.
+A concrete grammar tells how to translate the abstract trees to a given language,
+and deals with issues such as word order, case and agreement. 
+%implements the abstract 
+%To add a new language it amounts to give a concrete grammar, that implements
+%the abstract grammar and describes how to translate abstract
+%trees into strings of the language.
+The framework enables us to parse strings into 
 abstract trees as well as linearize trees into strings.
@@
 Reusability is further supported by the separation between resource grammars
-and application grammars. The resource library provided with GF, 
+and application grammars. The resource library provided with GF 
 implements morphological and
@@
 level and does not need to describe how to 
-form standard sentences, phrases or decline words.
+form standard sentences, phrases or inflect words.
 
 Since many of the languages in the GF library resemble each other grammatically,
-they can share much of their grammar implementations. This is usually done by using a
-\verb|Functor|, which lets a number of languages share parts of their implementation. 
+they can share much of their implementations. This is usually done by using a
+\verb|Functor|. % , which lets a number of languages share parts of their implementation. 
 In addition to simply avoiding code duplication, this technique aids the code maintenance.
@@
 This project takes a different approach by using GF for open domain language.
-While the resources have to be general and language independent,
-language specific constructs such as stylistic changes, idioms, informal
-expressions are given in a special module.
-This module, together with the Swedish resource grammar has been our starting
-point. % of this project.
+%While the resources have to be general and language independent,
+%language specific constructs such as stylistic changes, idioms, informal
+%expressions are given in a special module.
+%This module, together with 
+The Swedish resource grammar has been our starting
+point, % of this project.
 %By starting from an already existing Swedish grammar written in GF,
-We get a basic description of the language and the framework provides
+from which We get a basic description of the language. The framework provides
 tools such as parsing, generation and
 a well-tested interpretation of the parse trees. Furthermore, there are tools
-for using the grammar in a number of programming languages like Haskell
+for using a grammar in a number of programming languages like Haskell
 and Java.
@@
 research. 
-From SALDO, a large GF lexicon has earlier been extracted,
-containing 50 000 entries. By updating this technique, we want
-to reimport SALDO, which has been developed since.
+A large GF lexicon has been extracted from SALDO,
+containing 50 000 entries. By updating this technique, we will
+reimport SALDO, which has been developed since.
 The importing method should be fast and reliable
@@
 
-\section{Work in progress}
+%compare!!
+\section{Related work}
+Language technology for Swedish is an area of much interesting research.
+There are two other deep grammar-based parser for Swedish:
+SICS (Raynen and Gambäck, 1992) based on Core Language Engine and ? (Mats Wirén,?).
+None of them are freely available.
+Among other parsers for Swedish, the statistical MaltParser\cite{malt}
+which was trained on Talbanken, is worth mentioning. 
+CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
+whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
+a context free grammar. Both GTA and CassSwe operate on POS tagged text.\\
+Extract and FM\cite{MarkusForsberg2007} are tools for supervised lexicon
+extraction compatible with GF.
+%For lexical extractions, there are tools like
+%Extract for supervised lexicon extraction and
+%FM for programming lexical resources.
+
+\section{Results and progress}
 \label{sec:progress}
@@
 allow the word to be added to the lexicon, remove it or request another guess.
-Although using simple techniques, the tool 
-manages to correctly guess 70-75\% of the given lemmas when tested on
-verbs.\\
+
+The tool has been tested on verbs. Although using simple techniques it 
+manages to assign the correct paradigm to 70-75\% of the given lemmas.
+Smaller test has shown that out of the accepted lemmas, the correct guess is
+made directly in 75\% of the cases. 
+%is % estimated
+% whereas the user has to reject the others to force a new guess. %rest are found after an incorrect guess rejected by the user.\\
 %Extract uses a similar 
@@
 The information given by this mapping, namely data about which form a word is
-currently used in, may be used for the lexical extraction
+currently used in, may be used by the lexical extraction
 tools.% Those can be enhanced if they are given more data. 
-The translated trees will also enable us to extract probabilities for how often
-different functions are used, a feature that would enable disambiguation.
+The translated trees enable us to extract probabilities for how often
+different functions are used, a feature that will enable disambiguation.
 Furthermore, the translation makes it easy to identify grammatical constructions
@@
 in Talbanken.
-
 Another important use of the mapping is evaluation of the parser, which can be
 accomplished by comparing the parse trees and the trees from the transformer.
+
+We have so far concentrated on shorter sentences, without idioms and
+conjunction. 
+Since GF demands information 
+that are not given by the Talbanken tags - such as the valency of verbs - a
+full translation requires that all words used are already known to GF. 
+The evaluation of trees of this sort shows that the mapping
+can restore 85\% of the nodes. 
+%, given that all words are
+%known to the grammar.
+If the we lift the restrictions, we can currently restore 64\% of the tags.
+%mapping, a number that would probably increase substansially when working with a bigger lexicon.
 %is an interesting way of
@@
 %implementation is needed. \\
-A number of grammatical constructions have been added the the Swedish grammar
-include constructs that are generally not present in other languages,
-such as the reflexive pronoun \emph{sitt}: \\
+The resource grammar gives a good start and our present grammar 
+covers constructions such as declarative sentences with normal or inverted
+word order, questions, passives, imperatives, relative clauses, cleft
+constructions etc.
+A number of constructs that are generally not present in other languages and therefore not given
+by the resources, have been added. Those include the use of the reflexive pronoun \emph{sitt}: \\
 \emph{Han såg \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw his (own) house}) $\;$
@@
 %>> also not expressable in all languages
-The possibility to put a part of a sentence in focus has also been added:\\
+Fronting a words or sentences are very common in Swedish and are now allowed by the grammar:\\
+%the possibility for this has been added to the grammar:\\
 \emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
-Emphasizing words or phrases by putting them in the beginning of a sentence
-is commonly used in Swedish.
 This sort of rephrasing is however not given by the resource grammar, since it
@@
 much care is taken to develop it as neatly as possible.
-
-%compare!!
-\section{Related work}
-Language technology for Swedish is an area of much interesting research.
-There are tools for lexical extractions, such as
-Extract for supervised lexicon extraction and
-FM\cite{MarkusForsberg2007} for programming lexical resources.
-Among other parsers for Swedish, the statistical MaltParser\cite{malt}
-which was trained on Talbanken, is worth mentioning. 
-CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
-whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
-a context free grammar. Both GTA and CassSwe operate on POS tagged text.
 
@@
 %% more explaination of probabilities
-After having developed the grammar and lexicon, we plan to 
-make the parser robust by using techniques such as chunk parsing, named
-entity recognition and methods for handling unknown 
+We aim to make the parser robust by equipping it with techniques such as chunk
+parsing, named entity recognition and methods for handling unknown 
 grammatical constructions such as idioms and ellipses. %etc. 
@@
 \section{Conclusion}
-We intend to implement a robust deep parser for Swedish, by first developing
-a large scale-grammar. The grammar can be compiled into a
-parser by GF, and we will equip this with named entity recognition, a
-voluminous lexicon, statistical information for disambiguation,
-a method of parallel chunk parsing etc.
-So far we have been working on the grammar, the lexical resources and
-a translation of the treebank Talbanken to GF.
+%We intend to implement a robust deep parser for Swedish, by first developing
+%a large scale-grammar. 
+By combining GF with a large lexicon, we acquire a deep parser for 
+Swedish. All parts, including the underlying grammar, are open-source and can be reused.
+We have developed tools for extending the lexicon with words from Talbanken,
+and we show how to make use of the information in the manually tagged
+treebank.
+The usage of GF gives us the advantage to start from
+a well-defined system for describing grammar, as well as tools for
+parsing.
+
+%We have shown how to combine existing resources, we build a large-scale grammar for Swedish.
+%The grammar can be compiled into a
+%parser by GF, and we equip this with a extensive lexicon. 
+%and we will equip this with named entity recognition, a
+%voluminous lexicon, statistical information for disambiguation,
+%a method of parallel chunk parsing etc.
+%So far we have been working on the grammar, the lexical resources and
+%a translation of the treebank Talbanken to GF.
 % adventage to what??
-The usage of Grammatical Framework gives us the advantage to start from
-a well-defined system of describing grammar, as well as tools for
-parsing.
 

### 2011-10-14 mapping/translate.hs
test files

@@ -45,11 +45,11 @@ type PMonad = (RWS () [String] S)
 $(mkLabels [''S])
 
 test = False
-usePGF = testGr
+usePGF = bigGr --testGr
 testGr = ("../gf/BigTest.pgf","BigTestSwe")
 bigGr  = ("../gf/Big.pgf","BigSwe")
 lang   = fromJust $ readLanguage "BigTestSwe"
-paint  = True
+paint  = False
 
 startState :: S 
 startState = S {_isReflGenVP = False, _isExist = False
@@ -264,6 +264,8 @@ penn =
 --     --,"PT"  cant parse 'sjálv'
       ,"RA" :-> inAdv
       ,"SP" :-> do a <- pAdj         --undersok om fler sp kan ha denna
+                        `mplus`
+                        pAdA 
                    write ("adj return"++show a)
                    returnApp cidCompAP [a]
                 `mplus`

### 2011-10-14 doc/paper/Paper.tex
final paper, mapping text, updated plan

@@
 
-\title{First results of developing a Swedish parser based on a wide-coverage grammar\\
-        or Combining resources into a grammar-driven Swedish parser \\
-        or Bootstrapping a Swedish parser \\
-        or A grammar-based Swedish parser from already existing resources}
-\author{Malin Ahlberg \& Ramona Enache
+\title{%First results of developing a Swedish parser based on a wide-coverage grammar\\
+        Combining language resources into a grammar-driven Swedish parser} 
+        %or Bootstrapping a Swedish parser \\
+        %or A grammar-based Swedish parser from already existing resources}
+\author{Malin Ahlberg \qquad\qquad Ramona Enache
 %\and
@@
 }
-\def\titlerunning{First results of developing a grammar based Swedish parser}
+\def\titlerunning{Combining language resources into a grammar-driven Swedish parser}
 \def\authorrunning{M. Ahlberg}
@@
 \begin{abstract}
-This paper describes on going work of a rule-based, open-source
+This paper describes on-going work of a rule-based, open-source
 parser for Swedish. The central component is a wide-coverage grammar
-implemented in the formalism Grammatical Framework (GF), which
-can be compiled to a % from which we derive?? 
+implemented in the Grammatical Framework formalism (GF), from which
+we obtain a %can be compiled to a % from which we derive?? 
 parser.
 In addition to GF, we use two other main resources: the Swedish
-treebank Talbanken and the electronic lexicon SALDO.
+treebank Talbanken and the electronic SALDO lexicon.
 The resulting parser will 
 give a full syntactic analysis of the input sentences.
-It will be highly reusable, freely available
+It will be highly reusable, freely available,
 and as GF provides libraries for compiling
@@
 clause must consist of a verb.
-The first constituent of the clause is usually made up by the subject,
+The first constituent of the clause is usually made up of the subject,
 although it likewise could consist of adverbial phrases or objects.
@@
 Our goal is to implement a wide-coverage grammar and parser for Swedish
-using the grammar formalism GF, and 
+using the GF formalism, and 
 thereby investigate how GF can be used for open-domain parsing.
@@
 %s and by comparing the results of parsing sentences in a commonly used Swedish treebank, Talbanken.
-We choose our resources so that both our grammar and parser can 
+We chose our resources so that both our grammar and parser can 
 be freely available and open-source.
@@
 
-The key idea is to divide the grammar into abstract and concrete parts. %% parts?
+The key idea is to divide a grammar into abstract and concrete parts. %% parts?
 The abstract grammar gives a logical representation of the semantics,
@@
 %information about what categories are used and how to combine them into trees.
-A concrete grammar tells how to translate the abstract trees to a given language,
-and deals with issues such as word order, case and agreement. 
+The concrete grammars tell how to translate the abstract trees to a given language,
+and deal with issues such as word order, case and agreement. 
 %implements the abstract 
@@
 
-The grammar acts as an independent module and may be used in different
-projects.
+The grammar acts as an independent module.
+%and may be used in different projects.
 Reusability is further supported by the separation between resource grammars
@@
 syntactical rules for more than 20 different languages.
-Hence a writer of an application grammar can start her work at a higher
+Hence the writer of an application grammar can start her work at a higher
 level and does not need to describe how to 
@@
 \verb|Functor|. % , which lets a number of languages share parts of their implementation. 
-In addition to simply avoiding code duplication, this technique aids the code maintenance.
+This avoids code duplication and aids the code maintenance.
 
@@
 %This module, together with 
-The Swedish resource grammar has been our starting
+Using the Swedish resource grammar as our starting
 point, % of this project.
 %By starting from an already existing Swedish grammar written in GF,
-from which We get a basic description of the language. The framework provides
+we get a basic description of the language. The framework provides
 tools such as parsing, generation and
 a well-tested interpretation of the parse trees. Furthermore, there are tools
-for using a grammar in a number of programming languages like Haskell
+for using GF grammars in a number of programming languages like Haskell
 and Java.
@@
 containing 50 000 entries. By updating this technique, we will
-reimport SALDO, which has been developed since.
+re-import SALDO, which has been developed since.
 The importing method should be fast and reliable
@@
 There are two other deep grammar-based parser for Swedish:
-SICS (Raynen and Gambäck, 1992) based on Core Language Engine and ? (Mats Wirén,?).
-None of them are freely available.
+SICS (Raynen and Gambäck, 1992) based on Core Language Engine and a
+unification-based chart parser (Mats Wirén, 1988),
+though none of them are freely available.
 Among other parsers for Swedish, the statistical MaltParser\cite{malt}
@@
 we have created a tool for automatic acquisition. We have
-tested it on verbs with good results. It makes use of
+tested it on verbs, with good results. It makes use of
 the \emph{smart paradigm} given in the Swedish resource grammar.
-The smart paradigm is a function that given one form of a word can
+The smart paradigm is a function which given one form of a word, can
 infer which paradigm it most likely belongs to.
@@
 By combining this method with the information from the tags in Talbanken,
-the tool interactively generates GF lexicons. 
+the tool interactively generates GF lexical entries. 
 Given a list of words, it iteratively
@@
 the program will ask the  user to
-validate the claimed conjunction. The user may now either
+validate the claimed paradigm. The user may now either
 allow the word to be added to the lexicon, remove it or request another guess.
@@
 manages to assign the correct paradigm to 70-75\% of the given lemmas.
-Smaller test has shown that out of the accepted lemmas, the correct guess is
-made directly in 75\% of the cases. 
+A smaller test has shown that out of the accepted lemmas, the correct guess is
+made directly in 75\% of the cases, whereas the user has to reject one or more
+guesses for 25\%. 
 %is % estimated
@@
 to trees in GF format. The translation makes use of the POS tags as well as
-the syntactic information and the mapping has so far turned out to be unambiguous.
+the syntactic information, and the mapping has so far turned out to be unambiguous.
 
-The information given by this mapping, namely data about which form a word is
-currently used in, may be used by the lexical extraction
-tools.% Those can be enhanced if they are given more data. 
+The mapping gives information about which form a word is
+currently used in and this may be used by the lexical extraction
+tools. % Those can be enhanced if they are given more data. 
 The translated trees enable us to extract probabilities for how often
@@
 Since GF demands information 
-that are not given by the Talbanken tags - such as the valency of verbs - a
+that is not given by the Talbanken tags -- such as the valency of verbs -- a
 full translation requires that all words used are already known to GF. 
@@
 %known to the grammar.
-If the we lift the restrictions, we can currently restore 64\% of the tags.
+If the we lift the restrictions, we can currently restore 63\% of the tags.
 %mapping, a number that would probably increase substansially when working with a bigger lexicon.
@@
 %>> but only different words. 
-However, if we aim for a deeper and more covering analysis of Swedish,
+However, if we aim for a deeper and more comprehensive analysis of Swedish,
 the implementation of the languages needs to be more independent.
@@
 constructions etc.
-A number of constructs that are generally not present in other languages and therefore not given
-by the resources, have been added. Those include the use of the reflexive pronoun \emph{sitt}: \\
+A number of constructs that are generally not present in other languages and
+therefore not given by the resources, have also been added. These include the use of
+the reflexive pronoun \emph{sitt}: \\
 \emph{Han såg \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw his (own) house}) $\;$
 as opposed to \\
-\emph{Han såg \textbf{hans} hus} $\;$ (\emph{He saw his (an other person's) house}) \\
+\emph{Han såg \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
 %>> not critical, does not have to be in resource grammar
 %>> also not expressable in all languages
-Fronting a words or sentences are very common in Swedish and are now allowed by the grammar:\\
+Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
 %the possibility for this has been added to the grammar:\\
 \emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
-This sort of rephrasing is however not given by the resource grammar, since it
+This sort of rephrasing is not given by the resource grammar, since it
 has little effect on the logical representation.
 
-In order to keep the grammar clean and to keep it syntactically correct,
-much care is taken to develop it as neatly as possible.
+%In order to keep the grammar clean and to keep it syntactically correct,
+%much care is taken to develop it as neatly as possible.
 
@@
 a given sentence is grammatically correct according to the rules defined.
-Chosen parts of the parser or grammar may also be used in other applications,
+Chosen parts of the parser or grammar may be used in other applications
 which deal with controlled natural languages. Automatic language
@@
 By adding dependency probabilities, we aim to improve the
-disambiguation.
+potential for disambiguation.
 
-We intend to evaluate the parser both automatically - by comparing the output of
-the translated trees from Talbanken - and manually by an expert in Swedish grammar.
-Also the intermediate results are evaluated by our expert.
+We intend to evaluate the parser both automatically -- by comparing the output of
+the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
+%an expert in Swedish grammar.
+She also evaluates the intermediate results.
 
@@
 By combining GF with a large lexicon, we acquire a deep parser for 
-Swedish. All parts, including the underlying grammar, are open-source and can be reused.
+Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
 We have developed tools for extending the lexicon with words from Talbanken,
@@
 treebank.
-The usage of GF gives us the advantage to start from
+The usage of GF allows us to start from
 a well-defined system for describing grammar, as well as tools for
@@
 We would also like to give special thanks to Aarne Ranta, % Ramona Enache,
-Elisabet Engdahl, Krasimir Angelov, Olga Caprotti and Lars Borin for their help
+Elisabet Engdahl, Krasimir Angelov, Olga Caprotti, Lars Borin and John Camilleri for their help
 and support.

### 2011-10-14 saldo/FM/dist/build/dalin/dalin-tmp/Dict/Lex.hs
krasimirs code

@@ -1,0 +1,339 @@
+{-# OPTIONS -fglasgow-exts -cpp #-}
+{-# LINE 3 "src/lib/Dict/Lex.x" #-}
+
+{-# OPTIONS -fno-warn-incomplete-patterns #-}
+module Dict.Lex where
+import UTF8
+
+
+#if __GLASGOW_HASKELL__ >= 603
+#include "ghcconfig.h"
+#elif defined(__GLASGOW_HASKELL__)
+#include "config.h"
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import Data.Array
+import Data.Char (ord)
+import Data.Array.Base (unsafeAt)
+#else
+import Array
+import Char (ord)
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import GHC.Exts
+#else
+import GlaExts
+#endif
+alex_base :: AlexAddr
+alex_base = AlexA# "\xf8\xff\xff\xff\xfc\xff\xff\xff\xfd\xff\xff\xff\xdb\xff\xff\xff\xdc\xff\xff\xff\x00\x00\x00\x00\xf9\xff\xff\xff\xfa\xff\xff\xff\xdd\xff\xff\xff\xde\xff\xff\xff\xdf\xff\xff\xff\x12\x00\x00\x00\x00\x00\x00\x00\xf8\x00\x00\x00\x78\x01\x00\x00\x00\x00\x00\x00\x73\x00\x00\x00\x93\x00\x00\x00\x31\x00\x00\x00\x4e\x00\x00\x00"#
+
+alex_table :: AlexAddr
[…]

### 2011-10-14 saldo/FM/dist/build/dalin/dalin-tmp/Dict/Par.hs

@@ -1,0 +1,609 @@
+{-# OPTIONS_GHC -fno-warn-overlapping-patterns #-}
+{-# OPTIONS -fglasgow-exts -cpp #-}
+{-# OPTIONS -fno-warn-incomplete-patterns -fno-warn-overlapping-patterns #-}
+module Dict.Par where
+import Dict.Abs
+import Dict.Lex
+import Dict.ErrM
+#if __GLASGOW_HASKELL__ >= 503
+import qualified Data.Array as Happy_Data_Array
+#else
+import qualified Array as Happy_Data_Array
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import qualified GHC.Exts as Happy_GHC_Exts
+#else
+import qualified GlaExts as Happy_GHC_Exts
+#endif
+
+-- parser produced by Happy Version 1.18.4
+
+newtype HappyAbsSyn  = HappyAbsSyn HappyAny
+#if __GLASGOW_HASKELL__ >= 607
+type HappyAny = Happy_GHC_Exts.Any
+#else
+type HappyAny = forall a . a
+#endif
+happyIn10 :: (Ident) -> (HappyAbsSyn )
+happyIn10 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn10 #-}
+happyOut10 :: (HappyAbsSyn ) -> (Ident)
+happyOut10 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut10 #-}
+happyIn11 :: (Integer) -> (HappyAbsSyn )
+happyIn11 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn11 #-}
+happyOut11 :: (HappyAbsSyn ) -> (Integer)
+happyOut11 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut11 #-}
+happyIn12 :: (String) -> (HappyAbsSyn )
+happyIn12 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyIn12 #-}
+happyOut12 :: (HappyAbsSyn ) -> (String)
+happyOut12 x = Happy_GHC_Exts.unsafeCoerce# x
+{-# INLINE happyOut12 #-}
+happyIn13 :: (Dictionary) -> (HappyAbsSyn )
[…]

### 2011-10-14 saldo/FM/src/dalin/AdjRulesSw.hs

@@ -1,0 +1,157 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : AdjRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+-- Swedish adjective functions
+-------------------------------------------------------------------------
+module AdjRulesSw where
+
+import TypesSw
+import General
+import NounRulesSw(mk_case)
+import GenRulesSw
+import List(nub)
+
+no_neutrum :: Adjective -> Adjective
+no_neutrum f = missing f [(AF (Pos (Strong ASgNeutr))  c) | c <- values]
+
+no_genitive :: Adjective -> Adjective
+no_genitive f = missing f [(AF v Gen) | v <- values]
+
+no_masc :: Adjective -> Adjective
+no_masc f = missing f $ [AF (Pos (Weak (AxSg Sex_Masc))) v | v <- values] ++ [AF (Super (SupWeak Sex_Masc)) v | v <- values]
+
+mk_adj_pos :: [String] -> [String] -> [String] -> [String] -> AdjFormPos -> [String]
+mk_adj_pos liten litet lilla sma a = case a of
+  Strong gn  -> case gn of
+    ASgUtr   -> liten
+    ASgNeutr -> litet
+    APl      -> sma
+  Weak sn -> case sn of
+    AxSg g -> giveValues [lilla, lille] g
+    AxPl -> sma
+ where
+   lille = let x = map (\s -> if (last s == 'a') then tk 1 s else s) lilla in 
+           -- avoid bra -> bre, få -> fe
+             if (and (map (any is_vowel) x)) then nub (map (+?"e") x) else lilla
+
+mk_adjective :: String -> String -> String -> String -> String -> String -> String -> Adjective
+mk_adjective liten litet lilla sma mindre minst minsta = 
+    mk_adjective_v [liten] [litet] [lilla] [sma] [mindre] [minst] [minsta]
+
+mk_adjective_v :: [String] -> [String] -> [String] -> [String] -> [String] -> [String] -> [String] -> Adjective
+mk_adjective_v liten litet lilla sma mindre minst minsta AdjComp = strings $ concat [[x,x +? "-"] | x@(_:_) <- liten]
[…]

### 2011-10-14 saldo/FM/src/dalin/BuildSw.hs

@@ -1,0 +1,12 @@
+module BuildSw (
+                module NounBuildSw,
+                module AdjBuildSw,
+                module VerbBuildSw,
+                module OtherBuildSw
+               ) where
+
+import NounBuildSw
+import AdjBuildSw
+import VerbBuildSw
+import OtherBuildSw
+

### 2011-10-14 saldo/FM/src/dalin/EncodeSw.hs

@@ -1,0 +1,243 @@
+module EncodeSw where
+
+import Char
+import Frontend
+
+import qualified Data.Map as Map
+
+sw_encodings = Map.insert "parole" (Just parole_pos,Just parole_inhs,Just parole_param) (Map.insert "SUC" (Just suc_pos,Just suc_inhs,Just suc_param) Map.empty)
+parole_pos  _ = ""
+
+parole_inhs _ = []
+
+parole_param (pos,p,inhs) = 
+ case pos of
+   "pm" -> case words p of
+             xs | elem "nom" xs -> "NP00N@0S"
+             xs | elem "gen" xs -> "NP00G@0S"
+             _ -> parole_param ("nn",p,inhs)
+   "ab" -> case p of
+             "invar" -> "RG0S"
+             "pos"   -> "RGPS"
+             "komp"  -> "RGCS"
+             "super" -> "RGSS"
+             "c"     -> "RG0C"
+             "sms"   -> "RG0C"
+             _ -> error $ "ab " ++ p
+   "av" -> case p of
+             "pos indef sg f nom"     -> "AQPUSNIS"
+             "pos indef sg f gen"     -> "AQPUSGIS"
+             "pos indef sg m nom"     -> "AQPUSNIS"
+             "pos indef sg m gen"     -> "AQPUSGIS"
+             "pos indef sg n nom"     -> "AQPNSNIS"
+             "pos indef sg n gen"     -> "AQPNSNIS"
+             "pos indef pl nom"       -> "AQPNSNIS"
+             "pos indef pl gen"       -> "AQP0PG0S"
+             "pos def sg no_masc nom" -> "AQP0SNDS"
+             "pos def sg no_masc gen" -> "AQP0SGDS"
+             "pos def sg masc nom"    -> "AQPMSNDS"
+             "pos def sg masc gen"    -> "AQPMSGDS"
+             "pos def pl nom"         -> "AQP0PN0S"
+             "pos def pl gen"         -> "AQP0PG0S"
+             "komp nom"               -> "AQC00N0S"
+             "komp gen"               -> "AQC00G0S"
+             "super indef nom"        -> "AQS0PNIS"
+             "super def no_masc nom"  -> "AQS00NDS"
+             "super def masc nom"     -> "AQSMSNDS"
+             "super def masc gen"     -> "AQSMSGDS"
+             "c"                      -> "AQC0000C"
[…]

### 2011-10-14 saldo/FM/src/dalin/GenRulesSw.hs

@@ -1,0 +1,183 @@
+module GenRulesSw where
+
+import General
+import List
+import qualified Data.Set as Set
+
+type Suffixes = [((String -> String), String)]
+
+e :: String -> ((String -> String),String)
+e s = (id,s) 
+
+apply_suffixes :: String -> Suffixes -> [String]
+apply_suffixes stem xs = [f stem ++ suff | (f,suff) <- xs]
+
+lift :: [a] -> [[a]]
+lift  [] = []
+lift   s = [s]
+
+keep :: Int -> (String -> String) -> String -> String
+keep n f s = f pre ++ suf
+ where
+  (pre,suf) = case splitAt n (reverse s) of
+               (rs,rp) -> (reverse rp, reverse rs)
+
+drop_last_vowel = dv
+
+dsuff :: String -> String -> String
+dsuff suff s 
+ | isPrefixOf (reverse suff) (reverse s) = tk (length suff) s
+ | otherwise = s
+
+dv :: String -> String
+dv w = case find_stem_vowel w of
+        (seg,e,l) -> seg ++ l
+
+
+dvu :: String -> String
+dvu w = case find_stem_vowel w of
+          (seg,e,l) -> (ungeminate_m_n seg) ++ l
+
+fr :: String -> String
+fr s = case reverse s of
+         ('f':v:_) -> s ++ "v"
+         _ -> s
+
+fv s = if (dp 2 s == "fv") then tk 1 s else s
+
+dpe :: String -> String
+dpe w = case find_stem_vowel w of
+         (seg,e,l) | slast e == 'e' && length l < 2 -> seg ++ init e ++ l
+         _ -> w
+
+slast x@(_:_) = last x
+slast _ = '_'
+
+dpeu :: String -> String
+dpeu w = case find_stem_vowel w of
+           (seg,e,l) | slast e == 'e' && length l < 2 -> ungeminate_m_n seg ++ init e ++ l
+           _ -> ungeminate_m_n w
+
+dpl :: Char -> String -> String
+dpl c s = case span (/= c) (reverse s) of
+           (b,_:r) -> reverse (b ++ r)
+           (b,[])  -> reverse b
+
+drop_second_last :: String -> String
+drop_second_last s = (tk 2 s) ++ (dp 1 s)
+
+is_vowel :: Char -> Bool
+is_vowel c = Set.member c vowel 
+
+vowel = Set.fromList "aeiouyåäöAEIOUYÅÄÖ"
+
[…]

### 2011-10-14 saldo/FM/src/dalin/OtherBuildSw.hs

@@ -1,0 +1,210 @@
+module OtherBuildSw where
+
+import General
+import Dictionary
+import TypesSw
+import RulesSw
+import GenRulesSw
+
+-- Proper names
+{-
+pm_m :: String -> String -> Entry
+pm_m tax s = entryI (mk_pm s) [prValue MascGen,tax] 
+
+pm_f :: String -> String -> Entry
+pm_f tax s = entryI (mk_pm s) [prValue FemGen,tax] 
+
+pm_f_alice :: String -> String -> Entry
+pm_f_alice tax s = entryI (mk_pm_alice s) [prValue FemGen,tax] 
+
+pm_h :: String -> String -> Entry
+pm_h tax s = entryI (mk_pm s) [prValue Human,tax] 
+
+pm_w :: String -> String -> Entry
+pm_w tax s = entryI (mk_pm s) [prValue PNeutr,tax] 
+
+pm_p :: String -> String -> Entry
+pm_p tax s = entryI (mk_pm s) [prValue GPl,tax] 
+
+pm_v :: String -> String -> Entry
+pm_v tax s = entryI (mk_pm s) [prValue Pend,tax] 
+
+pm_u :: String -> String -> Entry
+pm_u tax s = entryI (mk_pm s) [prValue Utr,tax]
+
+pm_n :: String -> String -> Entry
+pm_n tax s = entryI (mk_pm s) [prValue Neutr,tax]
+
+pma_n :: String -> String -> Entry
+pma_n tax s = entryI (mk_pma s) [prValue Neutr,tax]
+
+pma_u :: String -> String -> Entry
+pma_u tax s = entryI (mk_pma s) [prValue Utr,tax]
+
+pma_h :: String -> String -> Entry
+pma_h tax s = entryI (mk_pma s) [prValue Human,tax]
+
+pma_w :: String -> String -> Entry
+pma_w tax s = entryI (mk_pma s) [prValue PNeutr,tax] 
+
+pma_m :: String -> String -> Entry
+pma_m tax s = entryI (mk_pma s) [prValue PNeutr,tax] 
+
+-- multi-word proper names
+
+pmm_n :: String -> String -> Entry
+pmm_n tax s = entryI (mk_pmm0 (words s)) [prValue Neutr,tax]
+
+pmm_u :: String -> String -> Entry
+pmm_u tax s = entryI (mk_pmm0 (words s)) [prValue Utr,tax]
+
[…]

### 2011-10-14 saldo/FM/src/dalin/OtherRulesSw.hs

@@ -1,0 +1,192 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : OtherRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+--
+-------------------------------------------------------------------------
+module OtherRulesSw where
+
+import GenRulesSw
+import General  
+import TypesSw  
+import AdjRulesSw(av1blek)
+import NounRulesSw(mk_case,mk_case_v)
+
+-- adverbs
+
+no_comp :: Adverb -> Adverb
+no_comp f = missing f [AdComp, AdSMS]
+
+mk_adverb :: [String] -> [String] -> [String] -> Adverb
+mk_adverb bra battre bast af = strings $
+   case af of
+    (AdverbForm Posit)  -> bra
+    (AdverbForm Compar) -> battre
+    (AdverbForm Superl) -> bast
+    AdComp              -> concat [[x,x +? "-"] | x <- bra]
+    AdSMS               -> [x +? "-" | x <- bra]
+
+abfrämst :: String -> Adverb
+abfrämst främst = no_comp $ mk_adverb [] [] [främst]
+
+abfint :: String -> Adverb
+abfint = adjective_to_adverb . av1blek . tk 1
+
+abfort fort = mk_adverb [fort] [fort +? "are"] [fort +? "ast"]
+
+
+adjective_to_adverb :: Adjective -> Adverb
+adjective_to_adverb adj =  
+    giveValues  [adj (AF p Nom) | p <- [Pos (Strong (ASgNeutr)), 
+                                        Comp, Super SupStrong]]
+
+ab_i_no_comp :: String -> AdverbInv
+ab_i_no_comp s AdverbInvForm = mkStr s
+ab_i_no_comp s _             = nonExist
+
+abm :: String -> AdverbMInv
+abm till_exempel _ = mkStr $ till_exempel
+
+-- pronouns
+
+mk_pron_jag :: (Suffixes,Suffixes,Suffixes,Suffixes,Suffixes) -> String -> PronPN
+mk_pron_jag (jag,mig,min,mitt,mina) s c = strings $ case c of
+  PNom            -> suff jag
+  PAcc            -> suff mig
[…]

### 2011-10-14 saldo/FM/src/fsv/PrintSw.hs

@@ -1,0 +1,112 @@
+module PrintSw where
+
+import List(intersperse,nub,inits)
+import Util
+import Dictionary
+import General
+import Char
+import Data.Bits ((.&.))
+
+print_table = [("multi",multi_printer),("rdf_sb",prRDF_sb),("flat_xml",prFlatXML)]
+
+-- | Print JSON
+multi_printer :: Dictionary -> String
+multi_printer = concat . nub . concat . map prOne . filter is_multi. unDict
+ where prOne  (id,stem,para,cat,inhs,tbl,_) = 
+           [conc [p "word" x, p "head" stem, p "pos" cat, p "param" "frag", pl "inhs" inhs, p "id" id, p "p" para, p "attr" (show attr)] |  (a,(attr,s)) <- (merge tbl), str <- unStr s, length (words str) > 1, x <- map unwords $ gen (words str)]
+       merge            [] = []
+       merge           [x] = [x]
+       merge (x@(param1,(a,s)):y@(param2,(a',s1)):xs) = 
+           case (number param1, number param2) of
+             ((v,n1,n2),(v',n1',n2')) | v == v' && n1' > n1 -> merge ((param2,(a',mkStr (unwords (unStr s ++ unStr s1)))):xs)
+             _ -> x:merge (y:xs)
+       number p = case map (map (read :: String -> Int) . words . map f) (filter (isDigit.head) (words p)) of
+                    [[v1,n1,n2]] -> (v1,n1,n2)
+       f x
+         | elem x "-:" = ' ' 
+         | otherwise = x
+       gen :: [String] -> [[String]]
+       gen []     =  []
+       gen (x:xs) =  filter (\xs -> not (null xs || one xs)) (inits (x:xs)) ++ gen xs
+       one [_] = True
+       one _ = False
+       pl s []   = quote s ++ ":" ++ "[]"
+       pl s xs   = quote s ++ ":[" ++ (concat (intersperse "," (map quote xs))) ++"]"
+       p s1 s2 = quote s1 ++ ":" ++ quote s2
+       conc xs = '{':(concat (intersperse "," xs)) ++ "}\n"
+       is_multi (_,stem,para,cat,_,_,_) = length (words stem) > 1
+
+-- | Print RDF
+prRDF_sb :: Dictionary -> String
[…]

### 2011-10-14 saldo/FM/src/fsv/TestGen.hs

@@ -1,0 +1,134 @@
+module TestGen where
+
+import List
+import Frontend
+import Char
+import Maybe
+
+is_a_word  = isJust . tword
+
+not_a_word = not . isJust . tword
+
+p_pos t = takeWhile (/='_') (tpara t)
+
+p_annotation t =  
+    case dropWhile (/='_') (tpara t) of
+      (_:xs) -> takeWhile (/='_') xs
+
+p_word t = case dropWhile (/='_') (tpara t) of
+             (_:xs) -> case dropWhile (/='_') xs of
+                         (_:ys) -> map f ys
+ where f '_' = ' '
+       f x   = x
+
+id_pos   t = f (tid t)
+ where f [] = error "invalid id: " ++ (tid t)
+       f ('.':'.':'.':xs) = f ('.':'.':xs)
+       f ('.':'.':xs)     = takeWhile (/='.') xs
+       f (x:xs) = f xs 
+
+id_lemma t = f (tid t)
+ where f [] = error "invalid id: " ++ (tid t)
+       f ('.':'.':'.':xs) = '.':f ('.':'.':xs)
+       f ('.':'.':xs)     = []
+       f (x:xs)           = x : f xs 
+
+norm_id_lemma t = remove_number $ reverse $ i
+  where i = id_lemma t
+        remove_number (x:xs) 
+         | isDigit x = case dropWhile isDigit xs of
+                         ('_':zs@(_:_)) -> reverse zs
+                         _              -> i
+         | otherwise = i
+
+head_to_lemma_id = add_underscore . symbols . unspace . thead
+ where unspace s = [if (x == ' ') then '_' else x | x <- s]
+       symbols = concat . map tr
+       add_underscore (x:xs)
+        | isDigit x || elem x "-·" = ('_':x:xs)
+        | otherwise = (x:xs)
+       tr c = maybe [c] id $ lookup c 
+               [(':',"·3A"), ('\'',"·27"), ('/',"·2F"), ('+',"·2B"), 
+                (',',"·2C"),('°',"·B0"), ('%',"·25"),('=',"·3D"),('?',"·3F")] 
+
+is_multi_word t = length pos > 2 && last pos == 'm' 
+ where pos = tcat t
+
+is_abbr t = last (tcat t) == 'a'
+
+is_vowel c = elem c "aeiouyåäöAEIOUYÅÄÖüÜàÀæÆøØ"
+
[…]

### 2011-10-14 saldo/FM/src/lib/CTrie.hs

@@ -1,0 +1,141 @@
+{-# LANGUAGE ForeignFunctionInterface #-}
+----------------------------------------------------------------------
+-- |
+-- Module      : Dictionary
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+-- > CVS $Date: 2006/10/15 15:54:16 $
+-- > CVS $Author: markus $
+-- > CVS $Revision: 1.19 $
+--
+-- A Trie ADT for Functional Morphology 
+-----------------------------------------------------------------------------
+module CTrie (
+              buildTrie, 
+              buildTrieDict, 
+              buildTrieDictSynt,
+              buildTrieWordlist,
+              trie_lookup, 
+              isInTrie, 
+              decompose
+             ) where
+
+import System.IO
+import System.IO.Unsafe
+import List
+import Foreign.C
+import Compound
+import qualified Data.Set as Set 
+import Data.Char (isDigit)
+import Monad(when)
+import Dictionary(dict2fullform,dict2idlex,Dictionary,FullFormLex)
+
+foreign import ccall "trie_lib.h build"        build     :: CString -> IO()
+foreign import ccall "trie_lib.h lookup_t"     lookup_t  :: CString -> IO()
+foreign import ccall "trie_lib.h start"        start     :: IO()
+foreign import ccall "trie_lib.h stop"         stop      :: IO()
+foreign import ccall "trie_lib.h next"         next      :: IO CString 
+foreign import ccall "trie_lib.h more"         more      :: IO CInt
+foreign import ccall "trie_lib.h getNumber"    getNumber :: CString -> CInt
+foreign import ccall "trie_lib.h in_t"         in_t      :: CString -> CInt
+foreign import ccall "trie_lib.h empty"        empty     :: IO()
+foreign import ccall "trie_lib.h insert_t"     insert_t  :: CString -> CString -> IO()
+foreign import ccall "trie_lib.h reversed"     reversed  :: IO()
+foreign import ccall "trie_lib.h no_count"     no_count  :: IO()
+
+type Attr = Int
+
+-----------------------------------------------------------------------
+{- | Constructs a C-trie from a file containing a fullform lexicon. -}
+buildTrie :: FilePath -> Bool -> Bool -> IO ()
+buildTrie f cb b = do fc <- newCString f
+                      when b reversed
[…]

### 2011-10-14 saldo/FM/src/lib/CommonMain.hs

@@ -1,0 +1,390 @@
+module CommonMain where
+
+import DictToDictionary
+import Print
+import List(intersperse,isPrefixOf,sortBy,nub)
+import System(getArgs, getEnv)
+import GeneralIO
+import General
+import IO
+import qualified Data.Set as Set
+import System(getProgName)
+import Dictionary
+import Frontend
+import Char
+import Dict.ErrM
+import Print
+import Monad
+import Command
+import Util
+import Combine
+--import Net
+import qualified Data.Map as Map
+
+-- import Tokenize
+import qualified CTrie
+
+data AnaType = 
+    Normal                |
+    NoAnalysis            |
+    AnaEval               |
+    FilterLexiconNo       |
+    FilterLexiconNoComp   |
+    FilterLexiconComp
+    
+gfTypes :: Language a => a -> String
+gfTypes l = "types." ++ name l ++ ".gf"
+
+readDicts :: Language a => a -> [FilePath] -> (Bool,Bool,Bool) -> IO (Dictionary,Int)
+readDicts l fs (undefcheck,argccheck,unusedcheck) = do output
+                                                       readDicts' l fs
+ where
+  output 
+   | length fs > 1 = prErr $ "\nprocessing dictionaries in files: " ++ (unwords fs) 
+   | null fs       = prErr $ "\nno dictionary loaded"
+   | otherwise     = prErr $ "\nprocessing dictionary in file " ++ (unwords fs)
+  readDicts' l [] = return $ (internDict l, 0)
+  readDicts' l (f:fs) = do (d,n) <- readDicts' l fs
+                           res   <- parseDict l f (undefcheck,argccheck,unusedcheck)
+                           case res of
+                             Ok (d1,n1) -> return $ (unionDictionary d d1,n+n1)
+                             Bad s      -> do prErr s
+                                              return (d,n)
+
+prStatistics :: Language a => a -> Int -> IO()
+prStatistics l n =
+ do let is = size (internDict l)
+    prErr $ print_lang l ++ print_paradigms l ++ print_size (n+is,n,is) ++ "\n"
+
+print_lang l = "language id: " ++ (name l) ++ "\n"
+
+print_paradigms l = case paradigmCount l of
+                     0 -> "no paradigms\n"
+                     1 -> "1 paradigm\n"
+                     n -> show n ++ " paradigms\n"
+
+print_size (0,n,isz)  = "no/empty dictionary"
[…]

### 2011-10-14 saldo/FM/src/lib/Dict/Lex.hs

@@ -1,0 +1,291 @@
+{-# LANGUAGE CPP #-}
+{-# LINE 3 "Lex.x" #-}
+
+{-# OPTIONS -fno-warn-incomplete-patterns #-}
+module Dict.Lex where
+import UTF8
+
+
+#if __GLASGOW_HASKELL__ >= 603
+#include "ghcconfig.h"
+#elif defined(__GLASGOW_HASKELL__)
+#include "config.h"
+#endif
+#if __GLASGOW_HASKELL__ >= 503
+import Data.Array
+import Data.Char (ord)
+import Data.Array.Base (unsafeAt)
+#else
+import Array
+import Char (ord)
+#endif
+alex_base :: Array Int Int
+alex_base = listArray (0,19) [-8,-4,-3,-37,-36,0,-7,-6,-35,-34,-33,18,0,248,376,0,115,147,49,78]
+
+alex_table :: Array Int Int
+alex_table = listArray (0,631) [0,11,11,11,11,11,-1,-1,1,7,7,7,8,0,0,0,0,0,0,0,0,0,0,0,11,0,16,11,11,11,11,11,12,12,0,0,12,3,7,6,19,19,19,19,19,19,19,19,19,19,11,12,0,0,0,0,0,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,17,0,0,0,0,17,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,10,0,12,4,5,0,0,0,0,0,-1,19,19,19,19,19,19,19,19,19,19,0,0,0,0,0,17,0,0,0,0,0,0,0,15,0,0,0,0,0,0,0,-1,0,17,0,0,0,0,0,17,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,15,0,0,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,18,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,13,18,13,13,13,13,13,13,13,13,-1,-1,-1,-1,-1,-1,-1,-1,-1,
[…]

### 2011-10-14 saldo/FM/src/lib/Dict/Par.hs

@@ -1,0 +1,685 @@
+{-# OPTIONS_GHC -w #-}
+{-# OPTIONS -fno-warn-incomplete-patterns -fno-warn-overlapping-patterns #-}
+module Dict.Par where
+import Dict.Abs
+import Dict.Lex
+import Dict.ErrM
+
+-- parser produced by Happy Version 1.18.6
+
+data HappyAbsSyn 
+	= HappyTerminal (Token)
+	| HappyErrorToken Int
+	| HappyAbsSyn10 (Ident)
+	| HappyAbsSyn11 (Integer)
+	| HappyAbsSyn12 (String)
+	| HappyAbsSyn13 (Dictionary)
+	| HappyAbsSyn14 ([Entry])
+	| HappyAbsSyn15 ([Arg])
+	| HappyAbsSyn16 ([Term])
+	| HappyAbsSyn17 (Entry)
+	| HappyAbsSyn18 (Term)
+	| HappyAbsSyn19 (Arg)
+
+{- to allow type-synonyms as our monads (likely
+ - with explicitly-specified bind and return)
+ - in Haskell98, it seems that with
+ - /type M a = .../, then /(HappyReduction M)/
+ - is not allowed.  But Happy is a
+ - code-generator that can just substitute it.
+type HappyReduction m = 
+	   Int 
+	-> (Token)
+	-> HappyState (Token) (HappyStk HappyAbsSyn -> [(Token)] -> m HappyAbsSyn)
+	-> [HappyState (Token) (HappyStk HappyAbsSyn -> [(Token)] -> m HappyAbsSyn)] 
+	-> HappyStk HappyAbsSyn 
+	-> [(Token)] -> m HappyAbsSyn
+-}
+
+action_0,
+ action_1,
+ action_2,
+ action_3,
+ action_4,
+ action_5,
+ action_6,
+ action_7,
+ action_8,
+ action_9,
+ action_10,
+ action_11,
+ action_12,
+ action_13,
+ action_14,
+ action_15,
+ action_16,
+ action_17,
+ action_18,
+ action_19,
+ action_20,
+ action_21,
+ action_22,
+ action_23,
+ action_24,
+ action_25,
+ action_26,
+ action_27,
+ action_28,
+ action_29,
+ action_30,
+ action_31,
+ action_32,
+ action_33,
+ action_34,
+ action_35,
+ action_36,
+ action_37 :: () => Int -> ({-HappyReduction (Err) = -}
+	   Int 
+	-> (Token)
[…]

### 2011-10-14 saldo/FM/src/lib/General.hs

@@ -1,0 +1,220 @@
+module General where
+
+import List (isPrefixOf,union,nub)
+import SharedString
+
+-- infixr 5 +/
+
+-- | list of word forms. 
+newtype Str = Str [String]   
+  deriving (Show, Eq, Ord)
+
+-- | Inflection table
+type Table  a = [(a,Str)]  
+
+-- | Finite inflection function
+type Finite a = a -> Str   
+
+-- | The finite parameter class  
+class (Eq a, Show a) => Param a where
+  values  :: [a]
+  value   :: Int -> a
+  value0  :: a
+  prValue :: a -> String
+  value n = values !! n
+  value0  = value 0
+  prValue = show
+
+-- | Token type
+data Tok = W String                   | 
+           A (String,String)          | -- dealing with ambiguous casing.
+           AA (String,String, String) | -- KALLE, Kalle, kalle
+	   P String                   | 
+           PD String                  |
+           BL                         | -- blank line.
+           D String
+  deriving (Eq, Show)
+
+applytok f t = case t of
+                W s -> W (f s)
+                A (s1,s2) -> A (f s1, f s2)
+                AA (s1,s2,s3) -> AA (f s1,f s2,f s3)
+                P  s          -> P (f s)
+                D  s          -> D (f s)
+
+-- | Compound forms
+type Attr = Int
+
+-- | default compound value
+noComp :: Attr 
+noComp = 0
+
+-- | Promote String to Str
+mkStr :: String -> Str
+mkStr     [] = nonExist
+mkStr (x:xs) = Str [(x:xs)]
+
+-- | Sharing of Str
+shareStr :: Str -> Str
+shareStr (Str xs) = Str $ map shareString xs
+
+-- | Translate Str to [String]
+unStr :: Str -> [String]
+unStr (Str xs) = xs
+
+-- | Translate a [String] to Str
+strings :: [String] -> Str
+strings = Str . nub . filter (not . null)
+
+-- | Apply function to 'a' and promote the resulting String to Str.
+mkStr1 :: (a -> String) -> a -> Str
+mkStr1 =  (mkStr .)
+
+-- | Apply function to all variants.
+mapStr :: (String -> String) -> Str -> Str
+mapStr f (Str xs) = Str (map f xs)
+
+-- | Union of two Str
+unionStr :: Str -> Str -> Str
[…]

### 2011-10-14 saldo/FM/src/lib/GeneralIO.hs

@@ -1,0 +1,106 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : GeneralIO
+-- Maintainer  : Markus Forsberg
+-- Stability   : 
+-- Portability : 
+--
+-- Top-level functions: 
+-- * reading/writing morphology databases
+-- * writing Lexicon, Tables, GF, XFST, Latex 
+-- * analysis/synthesis (Trie)
+--
+-----------------------------------------------------------------------------
+module GeneralIO where
+
+import Print
+import Dict.GetDict(getEntry)
+import DictToDictionary
+import General
+import Dictionary
+import IO
+import Command (Comp(..))
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+import Frontend
+import List (nub,sort,intersperse)
+import Maybe(fromJust)
+import Char
+import Dict.ErrM
+import Util
+import Compound
+import Monad(when)
+import qualified CTrie
+
+type Stem      = String
+type Id        = String
+
+analysis :: Language a => a -> Comp -> Maybe CompDesc -> ((String,String) -> [(String,String)]) -> String -> [[String]]
+analysis l mc f sandhi s =  post_filter mc $ map (map snd) $ CTrie.decompose f sandhi s
+    where f' = if mc == None then Just [[One (word_attr l)]] else f
+
+post_filter Min xs = case sort_length Set.empty xs of
+                              ys@(r:_) -> let n = length r in takeWhile (\s -> length s == n) ys
+                              _         -> []
+post_filter Max xs = case sort_length_rev Set.empty xs of
+                       ys@(r:_) -> let n = length r in takeWhile (\s -> length s == n) ys
+                       _         -> []
+post_filter _           xs = xs -- all, or default
+
+
+
+--       post_filter (Length n) xs  = filter (\s -> and (map (fun n) s)) xs
+-- [x | x <- xs, and $ map (fun n) x]
+--       fun n x = case words x of
+--                   (x:_) -> length x >= n
+--                   _     -> False
+
+lookupId :: String -> [String]
+lookupId s = nub $ [identifier xs | xs <-  map snd (CTrie.trie_lookup False s)]
+   where identifier ('\"':'i':'d':'\"':':':'\"':xs) = case span (/='\"') xs of
+                                                        (i,_) -> i
[…]

### 2011-10-14 saldo/FM/src/lib/Print.hs

@@ -1,0 +1,651 @@
+module Print where
+
+import General
+import Dictionary
+import List (intersperse,sort)
+import IO
+import Char
+import Util
+import Data.Bits((.&.))
+import qualified Data.Set as Set
+import qualified Data.Map as Map
+
+type Ident = String
+
+-- printing morphological objects as strings
+
+-- | Print word forms separated with '/'
+prStr :: Str -> String
+prStr = concat . intersperse "/" . unStr
+
+-- | similar prStr, but output '*' for nonExist 
+prAlts :: Str -> String
+prAlts ss = 
+    case unStr ss of
+     [] -> "*"
+     ys -> unwords $ intersperse "/" ys
+
+-- | Create a constant table
+consTable :: Str -> Table String
+consTable s = [("INVAR", s)]
+
+-- | Create an attributed constant table
+consTableW :: Str -> [(String,(Attr,Str))]
+consTableW s = [("INVAR", (noComp,s))]
+
+-- | Output a 'showed' inflection function
+putFun0 :: Param a => (a -> Str) -> IO ()
+putFun0 = putStr . unlines . map show . prTable . table
+
+-- | Output an inflection function
+putFun :: Param a => (a -> Str) -> IO ()
+putFun = putStr . unlines . map pr . prTable . table where
+  pr (a,ss) = a ++ " : " ++ prAlts ss
+
+-- | Print a parameter value without hierarchy (= parentheses)
+prFlat :: String -> String
+prFlat = filter (flip notElem "()")
+
+-- | Show all values for the first parameter
+prFirstForm :: Param a => Table a -> String
+prFirstForm = prStr . firstForm
+
+-- | Show one value for the first parameter (used in dictionary)
+prDictForm :: Param a => Table a -> String
+prDictForm = prDictStr . firstForm
+
+-- | Another Str printing function
+prDictStr :: Str -> String
+prDictStr t = case unStr t of
+  s:_ -> s
+  []  -> "NONE"
+
+-- | Print dictionary without attributes.
+prDictionary :: Dictionary -> String
+prDictionary = concat . intersperse ("\n") . map (unlines . prOne) . unDict
+ where
+  prOne (id,stem,para,typ,inhs,infl,_) = 
[…]

### 2011-10-14 saldo/FM/src/lib/Util.hs

@@ -1,0 +1,46 @@
+module Util where
+
+import List
+import IO
+import qualified Data.Set as Set
+
+quote :: String -> String
+quote s = ('\"':quot s)
+ where quot    [] = "\""
+       quot ('\"':xs) = '\\':'\"':quot xs
+       quot (x:xs)  = x:quot xs
+
+splitWith :: Eq a => a -> [a] -> Maybe ([a],[a])
+splitWith a xs = 
+    case (span (/=a) xs) of
+      (as,(_:bs)) -> return (as,bs)
+      _           -> Nothing
+
+splitList :: Eq a => a -> [a] -> [[a]]
+splitList a [] = []
+-- splitList 
+
+-- | Print to stderr.
+prErr :: String -> IO()
+prErr s =  hPutStr stderr (s ++ "\n")
+
+prStd :: String -> IO()
+prStd s = hPutStr stdout (s ++ "\n")
+
+sort_length l = sortBy f
+ where  check xs = or [Set.member s l | s <- xs]
+        f x y = case (length x, length y) of
+                 (n1,n2) | check x && not (check y) -> LT
+                         | not (check x) && check y -> GT
+                         | n1 < n2 -> LT
+                         | n1 > n2 -> GT
+                 _                 -> EQ
+
+sort_length_rev l = sortBy f
+ where check xs = or [Set.member s l | s <- xs]
+       f x y = case (length x, length y) of
+                 (n1,n2) | check x && not (check y) -> GT
+                       | not (check x) && check y -> LT
+                       | n1 < n2 -> GT
+                       | n1 > n2 -> LT
+                 _               -> EQ

### 2011-10-14 saldo/FM/src/saldo/Attr.hs

@@ -1,0 +1,27 @@
+module Attr where
+
+import General(Attr)
+
+h_attr :: Attr
+h_attr  = 0 
+
+init_attr :: Attr
+init_attr  = 1
+
+medial_attr :: Attr
+medial_attr = 2
+
+w_attr :: Attr
+w_attr  = 3 
+
+wp_attr :: Attr
+wp_attr = 4
+
+d_attr :: Attr
+d_attr  = 5
+
+c_attr :: Attr
+c_attr = 6
+
+s_attr :: Attr
+s_attr = 7

### 2011-10-14 saldo/FM/src/saldo/CommandsSw.hs

@@ -1,0 +1,2630 @@
+
+module CommandsSw where
+
+import BuildSw
+import Frontend
+import TypesSw
+import GenRulesSw
+import General
+import Dictionary
+import Attr
+import Maybe(catMaybes)
+import qualified Data.Map as Map
+
+commands :: [(String, [String], [String] -> Entry)]
+commands = sal ++ sal_tail
+
+suffix :: Int -> Entry -> String ->Entry
+suffix n e s = map_wordforms ((tk n s)++) e
+
+unDEFINED :: String -> Entry
+unDEFINED _ = emptyEntry
+
+noc :: (String -> Entry) -> (String -> Entry)
+noc f = \s -> remove_param "sms" $ remove_param "ci" $ remove_param "cm" $ f s
+
+sal_tail :: [(String, [String], [String] -> Entry)]
+sal_tail =
+ [
+  paradigm_h "pm_fph_kleopatra"              ["Kleopatra"]            $ pm_f "ph",
+  paradigm_h "pm_uls_storgatan"              ["Storgatan"]            $ pm_u "ls",
+  paradigm_h "pm_ulf_centralen"              ["Centralen"]            $ pm_u "lf",
+  paradigm_h "pm_noe_harvard"                ["Harvard"]              $ pm_n "oe",
+  paradigm_h "pm_utz_bambara"                ["Bambara"]              $ pm_u "tz",
+  paradigm_h "pm_vlg_nordsjön"               ["Nordsjön"]             $ pm_v "lg",
+  paradigm_h "pm_uoc_operan"                 ["Operan"]               $ pm_u "oc",
+  paradigm_h "pm_ueh_upplysningen"           ["Upplysningen"]         $ pm_u "eh",
+  paradigm_h "pm_uaa_viggen"                 ["Viggen"]               $ pm_u "aa",
+  paradigm_h "pm_naw_titanic"                ["Titanic"]              $ pm_n "aw",
+  paradigm_h "pm_upm_audhumbla"              ["Audhumbla"]            $ pm_u "pm",
+  paradigm_h "pm_uog_polisen"                ["Polisen"]              $ pm_u "og",
+  paradigm_h "pm_uaa_camel"                  ["Camel"]                $ pm_u "aa",
[…]

### 2011-10-14 saldo/FM/src/saldo/GenRulesSw.hs

@@ -1,0 +1,219 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : GenRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+--
+-- General functions for Swedish
+-------------------------------------------------------------------------
+module GenRulesSw (
+		   e, apply_suffixes, Suffixes,
+                   dv, dvu,keep,dsuff,
+                   drop_last_vowel,
+                   ds,ds_drop,fl,
+                     compound_s,
+                   -- drop_last_vowel,
+                    lift,
+                    is_vowel,
+                    is_voiced,
+                    if_vowel,
+                    is_consonant,
+                    drop_final_e,
+                    ungeminate,
+                    ungeminate_m_n,
+                    verb_ungeminate,
+                    verb_ungeminate_m_n,
+                    drop_second_last,
+                    insert_second_last,
+                    geminate,
+                    find_stem_vowel,
+                    vc,vct,
+                    dpl,
+                    mmn,mmr,
+                    triplecons,
+                    umlaut
+                  ) where
+
+import General
+import List
+import qualified Data.Set as Set
+
+type Suffixes = [((String -> String), String)]
+
+e :: String -> ((String -> String),String)
+e s = (id,s) 
+
+apply_suffixes :: String -> Suffixes -> [String]
+apply_suffixes stem xs = [f stem ++ suff | (f,suff) <- xs]
+
+lift :: [a] -> [[a]]
+lift  [] = []
+lift   s = [s]
+
+keep :: Int -> (String -> String) -> String -> String
+keep n f s = f pre ++ suf
+ where
+  (pre,suf) = case splitAt n (reverse s) of
+               (rs,rp) -> (reverse rp, reverse rs)
+
+drop_last_vowel = dv
+
+dsuff :: String -> String -> String
+dsuff suff s 
+ | isPrefixOf (reverse suff) (reverse s) = tk (length suff) s
+ | otherwise = s
+
+dv :: String -> String
+dv w = case find_stem_vowel w of
[…]

### 2011-10-14 saldo/FM/src/saldo/Multiwords.hs

@@ -1,0 +1,24 @@
+module Multiwords where
+
+import Frontend
+import Dictionary
+import List
+import General
+import Attr
+
+multiwords :: Dictionary -> Dictionary
+multiwords = dictionary . map process_entry . unDict
+ where process_entry (id,stem,para,cat,inhs,tbl,e) = 
+           (id,stem,para,cat,inhs,
+              if (length (words stem) > 1) then split_multiwords tbl else tbl,e) 
+       split_multiwords tbl = concat [frag_seq param a pos mw | (param,(a,str)) <- tbl, 
+                                                               (mw,pos) <- zip (unStr str) [1..]]
+       hole = Nothing
+       frag_seq param a var s =
+           let ws  = words s
+               wsl = length ws
+               f n = case hole of
+                       Nothing -> ""
+                       Just position -> if n < position then "pre" else "post"
+            in [(unwords $ filter (not.null) [param,f n, show var ++ ":" ++ show n ++ "-" ++ show wsl],
+                 (w_attr,mkStr fragment)) | (n,fragment) <- zip [1..] ws]

### 2011-10-14 saldo/FM/src/saldo/NounBuildSw.hs

@@ -1,0 +1,785 @@
+module NounBuildSw where
+
+import General
+import Dictionary
+import GenRulesSw
+import TypesSw
+import RulesSw
+import Char
+import Attr
+
+substantive :: Substantive -> Genus -> Entry
+substantive n g = entryI ((hyphenate_compounds.geminator) n) [prValue g]
+
+substUtrum :: Substantive -> Entry
+substUtrum s = substantive s Utr
+
+substNeutrum :: Substantive -> Entry
+substNeutrum s = substantive s Neutr
+
+substPl :: Substantive -> Entry
+substPl s = substantive s GPl
+
+substDPl :: Substantive -> Entry
+substDPl s = substantive s GDPl
+
+substVack :: Substantive -> Entry
+substVack s = substantive s Pend
+
+substMasc :: Substantive -> Entry
+substMasc s = substantive s Utr
+
+noun_f :: Int -> Genus -> (Suffixes,Suffixes,Suffixes,Suffixes) -> (String -> Entry)
+noun_f n g (apa,apan,apor,aporna) s = noun_compound n g (apa,apan,apor,aporna,apa,apa) s 
+
+noun_compound_ng :: Int -> Genus -> (Suffixes,Suffixes,Suffixes,Suffixes,Suffixes,Suffixes) -> (String -> Entry)
+noun_compound_ng n g (apa,apan,apor,aporna,ap,ap2) s = substantive (missing n_f [(SF n d Gen) | n <- values, d <- values]) g 
+  where n_f p = (mk_substantive_v (suff apa) (suff apan) (suff apor) (suff aporna) (suff ap) (suff ap2) (suff ap)) p
+        suff    = apply_suffixes (tk n s)
+
+noun_compound :: Int -> Genus -> (Suffixes,Suffixes,Suffixes,Suffixes,Suffixes,Suffixes) -> (String -> Entry)
+noun_compound n g (apa,apan,apor,aporna,ap,ap2) s = substantive n_f g 
+  where n_f p = (mk_substantive_v (suff apa) (suff apan) (suff apor) (suff aporna) (suff ap) (suff ap2) (suff ap)) p
+        suff    = apply_suffixes (tk n s)
+
[…]

### 2011-10-14 saldo/FM/src/saldo/NounRulesSw.hs

@@ -1,0 +1,828 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : NounRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+--
+-- Swedish noun functions
+-------------------------------------------------------------------------
+module NounRulesSw where
+
+import GenRulesSw
+import General
+import TypesSw
+import Char(isDigit)
+import List
+
+type NSg    = String
+type NSgD   = String 
+type NPl    = String
+type NPlD   = String
+type NComp  = String
+type NDeriv = String 
+
+-- General Noun functions
+
+mk_substantive :: NSg -> NSgD -> NPl -> NPlD -> NComp -> NDeriv -> Substantive
+mk_substantive apa apan apor aporna ap_comp ap_deriv = 
+ variants (mk_substantive_v (lift apa) (lift apan) (lift apor) 
+                     (lift aporna) (lift ap_comp) (lift ap_comp) (lift ap_deriv)) [(MedComposite,strings (compound_s ap_comp))]
+
+mk_substantive_comp :: NSg -> NSgD -> NPl -> NPlD -> NComp -> NComp -> NDeriv -> Substantive
+mk_substantive_comp apa apan apor aporna ap_comp ap_comp2 ap_deriv a = 
+    mk_substantive_v (lift apa) (lift apan) (lift apor) 
+                     (lift aporna) (lift ap_comp) (lift ap_comp2) (lift ap_deriv) a
+
+mk_substantive_v :: [NSg] -> [NSgD] -> [NPl] -> [NPlD] -> [NComp] -> [NComp] -> [NDeriv]-> Substantive
+mk_substantive_v apa apan apor aporna ap_comp ap_comp2 ap_deriv a = 
+        case a of
+          SF n s c -> strings $ mk_case_v c $ 
+                       case (n,s) of
+                         (Sg,Indef) -> apa
+                         (Sg,Def)   -> apan
+                         (Pl,Indef) -> apor
+                         (Pl,Def)   -> aporna
+          InitComposite    -> strings ap_comp
+          MedComposite     -> strings ap_comp2
+          SMS              -> nonExist
+          Deriv            -> strings ap_deriv
+
[…]

### 2011-10-14 saldo/FM/src/saldo/OtherBuildSw.hs

@@ -1,0 +1,213 @@
+module OtherBuildSw where
+
+import General
+import Dictionary
+import TypesSw
+import RulesSw
+import GenRulesSw
+
+-- Proper names
+
+pm_m :: String -> String -> Entry
+pm_m tax s = entryI (mk_pm s) [prValue MascGen,tax] 
+
+pm_f :: String -> String -> Entry
+pm_f tax s = entryI (mk_pm s) [prValue FemGen,tax] 
+
+pm_f_alice :: String -> String -> Entry
+pm_f_alice tax s = entryI (mk_pm_alice s) [prValue FemGen,tax] 
+
+pm_h :: String -> String -> Entry
+pm_h tax s = entryI (mk_pm s) [prValue Human,tax] 
+
+pm_w :: String -> String -> Entry
+pm_w tax s = entryI (mk_pm s) [prValue PNeutr,tax] 
+
+pm_p :: String -> String -> Entry
+pm_p tax s = entryI (mk_pm s) [prValue GPl,tax] 
+
+pm_v :: String -> String -> Entry
+pm_v tax s = entryI (mk_pm s) [prValue Pend,tax] 
+
+pm_u :: String -> String -> Entry
+pm_u tax s = entryI (mk_pm s) [prValue Utr,tax]
+
+pm_n :: String -> String -> Entry
+pm_n tax s = entryI (mk_pm s) [prValue Neutr,tax]
+
+pma_n :: String -> String -> Entry
+pma_n tax s = entryI (mk_pma s) [prValue Neutr,tax]
+
+pma_u :: String -> String -> Entry
+pma_u tax s = entryI (mk_pma s) [prValue Utr,tax]
+
+pma_h :: String -> String -> Entry
+pma_h tax s = entryI (mk_pma s) [prValue Human,tax]
+
+pma_w :: String -> String -> Entry
+pma_w tax s = entryI (mk_pma s) [prValue PNeutr,tax] 
+
+pma_m :: String -> String -> Entry
+pma_m tax s = entryI (mk_pma s) [prValue PNeutr,tax] 
+
+-- multi-word proper names
+
+pmm_n :: String -> String -> Entry
+pmm_n tax s = entryI (mk_pmm0 (words s)) [prValue Neutr,tax]
+
+pmm_u :: String -> String -> Entry
+pmm_u tax s = entryI (mk_pmm0 (words s)) [prValue Utr,tax]
+
[…]

### 2011-10-14 saldo/FM/src/saldo/OtherRulesSw.hs

@@ -1,0 +1,212 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : OtherRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+--
+-------------------------------------------------------------------------
+module OtherRulesSw where
+
+import GenRulesSw
+import General  
+import TypesSw  
+import AdjRulesSw(av1blek)
+import NounRulesSw(mk_case,mk_case_v)
+
+-- adverbs
+
+no_comp :: Adverb -> Adverb
+no_comp f = missing f [AdComp, AdSMS]
+
+mk_adverb :: [String] -> [String] -> [String] -> Adverb
+mk_adverb bra battre bast af = strings $
+   case af of
+    (AdverbForm Posit)  -> bra
+    (AdverbForm Compar) -> battre
+    (AdverbForm Superl) -> bast
+    AdComp              -> concat [[x,x +? "-"] | x <- bra]
+    AdSMS               -> [x +? "-" | x <- bra]
+
+mk_adverb_c :: [String] -> [String] -> [String] -> [String] -> Adverb
+mk_adverb_c bra battre bast comp af = strings $
+   case af of
+    (AdverbForm Posit)  -> bra
+    (AdverbForm Compar) -> battre
+    (AdverbForm Superl) -> bast
+    AdComp              -> concat [[x,x +? "-"] | x <- comp]
+    AdSMS               -> [x +? "-" | x <- comp]
+
+abfrämst :: String -> Adverb
+abfrämst främst = no_comp $ mk_adverb [] [] [främst]
+
+abfint :: String -> Adverb
+abfint = adjective_to_adverb . av1blek . tk 1
+
+abfort fort = mk_adverb [fort] [fort +? "are"] [fort +? "ast"]
+
+
+adjective_to_adverb :: Adjective -> Adverb
+adjective_to_adverb adj =  
+    giveValues  [adj (AF p Nom) | p <- [Pos (Strong (ASgNeutr)), 
+                                        Comp, Super SupStrong]]
+
+ab_i_no_comp :: String -> AdverbInv
+ab_i_no_comp s AdverbInvForm = mkStr s
[…]

### 2011-10-14 saldo/FM/src/saldo/PrintSw.hs

@@ -1,0 +1,115 @@
+module PrintSw where
+
+import List(intersperse,nub,inits)
+import Util
+import Dictionary
+import General
+import Char
+import Data.Bits ((.&.))
+
+print_table = [("multi",multi_printer),("rdf_sb",prRDF_sb),("flat_xml",prFlatXML)]
+
+multi_printer :: Dictionary -> String
+multi_printer = concat . nub . map prOne . multi
+ where prOne (x,stem,cat,inhs,id,para,attr) = conc [p "word" x, p "head" stem, p "pos" cat, p "param" "frag", pl "inhs" inhs, p "id" id, p "p" para, p "attr" attr]
+       pl s []   = quote s ++ ":" ++ "[]"
+       pl s xs   = quote s ++ ":[" ++ (concat (intersperse "," (map quote xs))) ++"]"
+       p s1 s2 = quote s1 ++ ":" ++ quote s2
+       conc xs = '{':(concat (intersperse "," xs)) ++ "}\n"
+
+multi = concat . map prOne . filter is_multi. unDict
+ where prOne  (id,stem,para,cat,inhs,tbl,_) = 
+           [(x,stem, cat, inhs, id, para, (show attr)) |  (a,(attr,s)) <- (merge tbl), str <- unStr s, length (words str) > 1, x <- map unwords $ gen (words str)]
+       merge            [] = []
+       merge           [x] = [x]
+       merge (x@(param1,(a,s)):y@(param2,(a',s1)):xs) = 
+           case (number param1, number param2) of
+             ((v,n1,n2),(v',n1',n2')) | v == v' && n1' > n1 -> merge ((param2,(a',mkStr (unwords (unStr s ++ unStr s1)))):xs)
+             _ -> x:merge (y:xs)
+       number p = case map (map (read :: String -> Int) . words . map f) (filter (isDigit.head) (words p)) of
+                    [[v1,n1,n2]] -> (v1,n1,n2)
+       f x
+         | elem x "-:" = ' ' 
+         | otherwise = x
+       gen :: [String] -> [[String]]
+       gen []     =  []
+       gen (x:xs) =  filter (\xs -> not (null xs || one xs)) (inits (x:xs)) ++ gen xs
+       one [_] = True
+       one _ = False
[…]

### 2011-10-14 saldo/FM/src/saldo/TestGenSw.hs

@@ -1,0 +1,134 @@
+module TestGenSw where
+
+import List
+import Frontend
+import Char
+import Maybe
+
+is_a_word  = isJust . tword
+
+not_a_word = not . isJust . tword
+
+p_pos t = takeWhile (/='_') (tpara t)
+
+p_annotation t =  
+    case dropWhile (/='_') (tpara t) of
+      (_:xs) -> takeWhile (/='_') xs
+
+p_word t = case dropWhile (/='_') (tpara t) of
+             (_:xs) -> case dropWhile (/='_') xs of
+                         (_:ys) -> map f ys
+ where f '_' = ' '
+       f x   = x
+
+id_pos   t = f (tid t)
+ where f [] = [] -- "invalid id: " ++ (tid t)
+       f ('.':'.':'.':xs) = f ('.':'.':xs)
+       f ('.':'.':xs)     = takeWhile (/='.') xs
+       f (x:xs) = f xs 
+
+id_lemma t = f (tid t)
+ where f [] = [] -- error "invalid id: " ++ (tid t)
+       f ('.':'.':'.':xs) = '.':f ('.':'.':xs)
+       f ('.':'.':xs)     = []
+       f (x:xs)           = x : f xs 
+
+norm_id_lemma t = remove_number $ reverse $ i
+  where i = id_lemma t
+        remove_number (x:xs) 
+         | isDigit x = case dropWhile isDigit xs of
+                         ('_':zs@(_:_)) -> reverse zs
+                         _              -> i
+         | otherwise = i
+
+head_to_lemma_id = add_underscore . symbols . unspace . thead
+ where unspace s = [if (x == ' ') then '_' else x | x <- s]
+       symbols = concat . map tr
+       add_underscore (x:xs)
+        | isDigit x || elem x "-·" = ('_':x:xs)
+        | otherwise = (x:xs)
+       tr c = maybe [c] id $ lookup c 
+               [('&',"·26"),(':',"·3A"), ('\'',"·27"), ('/',"·2F"), ('+',"·2B"), 
+                (',',"·2C"),('°',"·B0"), ('%',"·25"),('=',"·3D"),('?',"·3F")] 
+
+is_multi_word t = length pos > 2 && last pos == 'm' 
+ where pos = tcat t
+
+is_abbr t = last (tcat t) == 'a'
+
[…]

### 2011-10-14 saldo/FM/src/saldo/TestSw.hs

@@ -1,0 +1,330 @@
+module TestSw (sal_id_exceptions, tests) where
+
+import General
+import List
+import Frontend
+import Char
+import Maybe
+import TestGenSw
+import qualified Data.Set as Set
+
+
+sal_id_exceptions :: [(String,String)]
+sal_id_exceptions = [("lova_2..vb.1","lova_1..vb.1"),
+                     ("kola_2..vb.1","kola_1..vb.1"),
+                     ("förslag_2..nn.1","förslag_1..nn.1"),
+                     ("förlag_2..nn.1","förlag_1..nn.1"),
+                     ("fris_2..nn.1","fris_1..nn.1"),
+                     ("bildrulle_2..nn.1","bildrulle_1..nn.1"),
+                     ("jaja_2..in.1","jaja_1..in.1"),
+                     ("oskyld_2..av.1","oskyld_1..av.1"),
+                    ("väl_2..ab.1","väl_1..ab.1"),
+                    ("förut_2..ab.1","förut_1..ab.1"),
+                     ("förklistra_2..vb.1","förklistra_1..vb.1"),
+                     ("förklistring_2..nn.1","förklistring_1..nn.1")
+                    ]
+
+tests :: (PositiveTests, NegativeTests)
+tests = (pos,neg)
+ where 
+ pos = pos_gen_rules ++ pos_noun_rules ++ pos_verb_rules ++ pos_adj_rules
+ neg = neg_gen_rules ++ neg_noun_rules ++ neg_verb_rules ++ neg_adj_rules
+
+-- General rules
+pos_gen_rules = [pos_id_para, multi_word, head_lemma,paradigm_pos]
+
+neg_gen_rules = [three_of_same, symbols, mmn,mmr,nnt,sms_missing_hyphen,end_with_h]
+
+-- Is POS in lemma id = paradigm id?
+pos_id_para :: TestInput -> Result
+pos_id_para t 
+ | not ( is_a_word t) = pass
+ | p_pos t == id_pos t = pass
+ | otherwise           = message t "lemma_id pos - paradigm pos mismatch"
+ 
+-- Have the multi words more than one word?
+multi_word :: TestInput -> Result
+multi_word t
+ | not (is_multi_word t)                             = pass
+ | length (words (thead t)) > 1                      = pass
+ | otherwise                                         = message t "singleton multi word"
+
+--  Is head == lemma id head?
[…]

### 2011-10-14 saldo/FM/src/saldo/VerbRulesSw.hs

@@ -1,0 +1,386 @@
+----------------------------------------------------------------------
+-- |
+-- Module      : VerbRulesSw
+-- Maintainer  : Markus Forsberg
+-- Stability   : (stability)
+-- Portability : (portability)
+--
+--
+-- Swedish verb functions
+-------------------------------------------------------------------------
+
+module VerbRulesSw where
+
+import GenRulesSw
+import NounRulesSw(mk_case)
+import AdjRulesSw(mk_adj_pos)
+import General
+import TypesSw
+
+infer_pret_part :: String -> AdjFormPos -> [String]
+infer_pret_part bunden' = mk_adj_pos [bunden] [bundet] [bundna] [bundna] where
+  (bundet,bundna) = case reverse bunden of
+    'd':'a':_ -> (bunde ++ "t",  bunden ++ "e")
+    'n':'e':_ -> (bund  ++ "et", verb_ungeminate_m_n bund ++ "na") -- fun-na, kom-na
+    'd':'d':_ -> (bund  ++ "tt", bunden ++ "a") 
+    't':'t':_ -> (bunden,        bunden ++ "a")
+    'd':_     -> (bunde ++ "t",  bunden ++ "a")
+    _         -> (bunden,        bunden ++ "a")
+  bunden = verb_ungeminate bunden'
+  bunde = verb_ungeminate $ tk 1 bunden
+  bund  = verb_ungeminate $ tk 2 bunden
+  
+ppvar :: Int -> Verb -> String -> Verb
+ppvar n v s = variants_or_excepts v [(VF (Pres Ind SForm), strings xs)]
+ where ws = unStr (v (VF Imper))
+       variants_or_excepts = if all (=="s") (map (dp 1) ws) then excepts else variants
+       xs  = [w ++ "es" | w <- ws, not (vow w || single_m w)]
+       single_m x = case reverse x of
+                      ('m':m:_) -> m /= 'm'
+                      _         -> False
+       vow [] = False
+       vow w@(_:_) = is_vowel (last w) 
+
+
+-- if (is_vowel (last (tk n s))) then [] else [e "es"]
+
+mk_verb :: String -> String -> String -> String -> String -> String -> String -> String -> Verb
+mk_verb finna finner finne finn fann funne funnit funnen = 
[…]

### 2011-10-14 saldo/make-saldo.hs

@@ -1,0 +1,326 @@
+import Prelude hiding (lex)
+import System.IO
+import System.Cmd
+import System.Exit
+import System.Directory
+import Data.Char
+import Data.Maybe
+import Data.List
+import Data.IORef
+import qualified Data.Map as Map
+import Text.Read.Lex(Lexeme(..),lex)
+import Text.ParserCombinators.ReadP
+import Debug.Trace
+
+import qualified PGF as PGF
+
+import UTF8
+import CommandsSw
+import Frontend
+import TestSw
+import Dictionary(Entry,set_lemma_id,get_lemma_id,unDict)
+import qualified Dict.Abs as Abs
+import EncodeSw
+import Dict.ErrM(Err(..))
+import DictToDictionary
+import General(Str(..))
+
+main = do
+  hSetBuffering stdout NoBuffering
+  putStr "Reading saldo_2.0/saldo_20v00.txt ... "
+  res   <- parseDict SAL "saldo_2.0/saldo_20v00.txt" (False,False,False)
+  saldo <- case res of
+             Ok (dict,_) -> return $ Map.fromList [(mkGFName (get_lemma_id entry), entry) | entry <- unDict dict]
+             Bad msg     -> fail msg
+  putStrLn ""
+
+  cnc_file <- createGF saldo
+  tmp_saldo <- compileGF
+  loop tmp_saldo saldo cnc_file
+  where
+    loop tmp_saldo saldo cnc_file = do
+
+      (cnc_file,changes) <- updateGF tmp_saldo saldo cnc_file
+      tmp_saldo <- compileGF
+      if changes == 0
+        then return ()
+        else loop tmp_saldo saldo cnc_file
+
+-------------------------------------------------------------------
+-- Bootstrap initial version of GF
+-------------------------------------------------------------------
+
+createGF saldo = do
+  putStr "Creating GF source for SALDO ... "
+  print [(inft) | (id,lemma,_,cat,_, inft,_) <- Map.elems saldo, lemma == "betyg"]
+  let cnc_file = [(id,gf_cat,[[lemma]],"",paradigms) | (id,lemma,_,cat,_, inft,_) <- Map.elems saldo
+                                                     , (fm_cat,gf_cat,_,paradigms) <- catMap, fm_cat == cat
[…]

### 2011-10-24 doc/paper/Paper.tex
newer paper

@@
 treebank Talbanken and the electronic SALDO lexicon.
-The resulting parser will 
-give a full syntactic analysis of the input sentences.
+The resulting parser 
+gives a full syntactic analysis of the input sentences.
 It will be highly reusable, freely available,
@@
 This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
-and describe our results so far in Section \ref{sec:progress}.
+and describe our results in Section \ref{sec:progress}.
 %We aim to implement % The goal is to be able to 
@@
 
-The grammar acts as an independent module.
+The grammar acts as an independent module and
 %and may be used in different projects.
-Reusability is further supported by the separation between resource grammars
+reusability is further supported by the separation between resource grammars
 and application grammars. The resource library provided with GF 
@@
 they can share much of their implementations. This is usually done by using a
-\verb|Functor|. % , which lets a number of languages share parts of their implementation. 
-This avoids code duplication and aids the code maintenance.
+\verb|Functor|, % , which lets a number of languages share parts of their implementation. 
+avoids code duplication and aids the code maintenance.
 
@@
 \subsection{Talbanken}
-Talbanken\cite{talbanken} is
-a Swedish treebank assembled in the 1970s at Lund University.
-In 2005 it was modernized by Nivre, Nilsson and Hall\cite{talbanken05} and
-enriched with annotation for a full phrase structure analysis. 
-We are using the treebank for development and evaluation.
-
+For development and evaluation, we use the Swedish treebank
+Talbanken\cite{talbanken}.
+It was assembled in the 1970s at Lund University and modernized
+in 2005 by Nivre, Nilsson and Hall\cite{talbanken05} and
+enriched with annotation for a full phrase structure analysis.  \\
 Although Talbanken contains both written and spoken Swedish,
@@
 research. 
-A large GF lexicon has been extracted from SALDO,
-containing 50 000 entries. By updating this technique, we will
-re-import SALDO, which has been developed since.
+We are developing tools for extracting GF lexicons
+from SALDO, which have resulted in a dictionary containing
+over 98000 entries including information about verbs that
+are reflexive or needs a particle.
 The importing method should be fast and reliable
@@
 There are two other deep grammar-based parser for Swedish:
-SICS (Raynen and Gambäck, 1992) based on Core Language Engine and a
-unification-based chart parser (Mats Wirén, 1988),
+Swedish Core Language Engine (Raynen and Gambäck, 1992) and a
+unification-based chart parser\cite{wiren},
+%correct reference?
 though none of them are freely available.
-Among other parsers for Swedish, the statistical MaltParser\cite{malt}
-which was trained on Talbanken, is worth mentioning. 
+Among other parsers for Swedish, the statistical MaltParser\cite{malt},
+trained on Talbanken, is worth mentioning. 
 CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
@@
 %known to the grammar.
-If the we lift the restrictions, we can currently restore 63\% of the tags.
-%mapping, a number that would probably increase substansially when working with a bigger lexicon.
+If we lift the restrictions, we can restore 63\% of the tags.
+%mapping, a number that would probably increase substantially when working with a bigger lexicon.
 %is an interesting way of
@@
 %>> not critical, does not have to be in resource grammar
-%>> also not expressable in all languages
+%>> also not expressible in all languages
 Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
@@
 
-%% more explaination of probabilities
+%% more explanation of probabilities
 We aim to make the parser robust by equipping it with techniques such as chunk

### 2011-10-24 doc/paper/Paper.tex
latest paper

@@
 \begin{abstract}
-This paper describes on-going work of a rule-based, open-source
+This paper describes on-going work on a rule-based, open-source
 parser for Swedish. The central component is a wide-coverage grammar
-implemented in the Grammatical Framework formalism (GF), from which
+implemented in the GF formalism (Grammatical Framework), from which
 we obtain a %can be compiled to a % from which we derive?? 

### 2011-11-04 doc/report/dispostion.tex
planning and disposition

--- 2011-10-24 doc/paper/Paper.tex
@@
-\documentclass[submission]{eptcs} %rätt lincens?
-\providecommand{\event}{LREC2012}
+\documentclass{article} 
 \usepackage[utf8]{inputenc}
@@
 
-\title{%First results of developing a Swedish parser based on a wide-coverage grammar\\
-        Combining language resources into a grammar-driven Swedish parser} 
+% ska gå att förstå och locka till läsning
+%ska tydligt beskriva vad rapporten handlar om
+%ska innehålla vettiga sökord
+\title{A large scale grammar and parser for Swedish}
+      %First results of developing a Swedish parser based on a wide-coverage grammar\\
+        %Combining language resources into a grammar-driven Swedish parser} 
         %or Bootstrapping a Swedish parser \\
         %or A grammar-based Swedish parser from already existing resources}
-\author{Malin Ahlberg \qquad\qquad Ramona Enache
-%\and
-%Co Author \qquad\qquad 
+\author{Malin Ahlberg
 \institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
@@
 \begin{abstract}
-This paper describes on-going work on a rule-based, open-source
-parser for Swedish. The central component is a wide-coverage grammar
-implemented in the GF formalism (Grammatical Framework), from which
-we obtain a %can be compiled to a % from which we derive?? 
-parser.
-In addition to GF, we use two other main resources: the Swedish
-treebank Talbanken and the electronic SALDO lexicon.
-The resulting parser 
-gives a full syntactic analysis of the input sentences.
-It will be highly reusable, freely available,
-and as GF provides libraries for compiling
-grammars to a number of programming languages,
-chosen parts of the the 
-grammar may be used in various NLP applications. % dealing with
-%controlled or free
-%natural language. 
-% allow easy interaction with other programs and grammars
-%>> something very commercial, interesting and outstanding here please!
-
-\end{abstract}
+Vad är en parser? Varför vill vi ha den? Vad vill vi lägga till existerande GF?
+Vad har jag gjort? Vilka delar tog tid? Vilka metoder har jag använt?
+Vad vill vi komma till? Vad jag jag kommit fram till?
+% ca 10 minuter eller 200 ord
+%om problem, metod, resultat och ev. konsekvenser.
+%Sammanfattningen ska kunna läsas fristående!
 
 \section{Introduction}
+% Varför valde du ämnet? Varför är ämnet intressant?
+% Vilka frågor behandlar du?
+Vad är svenska, hur funkar GF, vad kan GF inte klara, hur bra var GF
+på talbanken? 
+Vad bestämde vi oss för att göra och inte göra?
+
+Making computers able of handling human language is a 
+hard problem.
+The meaning of a sentence depends not only of which words it consists of, but
+also on their syntactic use, how they interact and relate to each other.
+For a computer to make sense of natural language, it needs to analyse this 
+syntactic structure; it needs a good grammar and parser.
+In this project I have been aiming at implementing a robust parser for Swedish
+using the grammar formalism Grammatical Framework. %\cite{gf} (GF).
+
+The libraries of GF provide a basic grammar for Swedish, covering the
+fundamental features of the language, such as morphology and commonly used
+syntax. This is suitable for building domain specific applications. In those
+cases the user is not allowed to freely compose sentences, but has to stay
+within the bounds of a controlled language. Both the vocabulary and grammatical
+structures are fixed, meaning that there only is a limit number of ways to
+write a sentence.  Parsing open domain natural language is a much bigger task,
+since it involves handling both standard and non-standard grammatical
+constructions.
+
+I have mainly focused on three parts: extending the lexicon, the grammar and
+finding ways to compare the analyse given by GF to the one given by Talbanken. 
+%Vad vill du göra med dessa frågor?
+%Har du avgränsat ämnet?
+
+%Vilka metoder finns? Vilken har du valt och varför?
+%Vilka antaganden har du gjort?
+
+Our goal is to implement a wide-coverage grammar and parser for Swedish
+using the GF formalism, and 
+thereby investigate how GF can be used for open-domain parsing.
+We compile a large-scale grammar to a parser
+and combine it with the extensive lexicon SALDO. The Swedish treebank 
+Talbanken provides manually tagged trees which we use for improving and evaluating
+the grammar. The parser will additionally
+be evaluated by experts. \\
+We chose our resources so that both our grammar and parser can 
+be freely available and open-source.
+
+This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
+and describe our results in Section \ref{sec:progress}.
+
+\subsection{Swedish}
+
 Swedish is a North-Germanic language spoken by approximately 10 million people.
@@
 
-Our goal is to implement a wide-coverage grammar and parser for Swedish
-using the GF formalism, and 
-thereby investigate how GF can be used for open-domain parsing.
-% better 
-We compile a large-scale grammar to a parser
-and combine it with the extensive lexicon SALDO. The Swedish treebank 
-Talbanken provides manually tagged trees which we use for improving and evaluating
-the grammar. The parser will additionally
-be evaluated by experts. \\
-%s and by comparing the results of parsing sentences in a commonly used Swedish treebank, Talbanken.
-We chose our resources so that both our grammar and parser can 
-be freely available and open-source.
-
-This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
-and describe our results in Section \ref{sec:progress}.
-%We aim to implement % The goal is to be able to 
-%a parser for open domain language and 
-%This paper describes our work and results 
-%so far as well as the directions of future work.
-
-\section{Background}
-\label{sec:background}
-\subsection{Grammatical Framework}
-\label{sec:gf}
-
-%städa upp detta stycke!!
+\subsection{GF}
 
@@
 
-The key idea is to divide a grammar into abstract and concrete parts. %% parts?
+The key idea is to divide a grammar into abstract and concrete parts. 
 The abstract grammar gives a logical representation of the semantics,
-% shorten this part
 modeled as abstract trees.
@@
 
+\subsection{Saldo}
+SALDO\cite{saldo} is an open source lexicon resource
+based on Svenskt Associationslexikon. It provides information about
+semantical relationships
+as well as full morphological analyses.
+It is
+developed at Språkbanken at Gothenburg University
+and intended for usage in language technology
+research.  
+% semantics of saldo, history, structure
+
 \subsection{Talbanken}
@@
 
-\subsection{Saldo}
-SALDO\cite{saldo} is an open source lexicon resource
-based on Svenskt Associationslexikon. It is
-developed at Språkbanken at Gothenburg University
-and intended for usage in language technology
-research. 
-We are developing tools for extracting GF lexicons
-from SALDO, which have resulted in a dictionary containing
-over 98000 entries including information about verbs that
-are reflexive or needs a particle.
-The importing method should be fast and reliable
-enough to allow us to always have a fresh version of the dictionary
-in GF.
-
-%compare!!
-\section{Related work}
+\subsection{Related work}
 Language technology for Swedish is an area of much interesting research.
@@
 \label{sec:progress}
+% Beskriv metoden.
+% Beskriv hur du omsatte metoden i verkligheten, dvs. ditt
+%i jobb/empirin/undersökningen o. dyl.
+
+\subsection{Lexical parts}
+\subsection{Importing SALDO}
+We are developing tools for extracting GF lexicons
+from SALDO, which have resulted in a dictionary containing
+over 100 000 entries. Since the word class analyses differs between
+Saldo and GF, we have focused on verbs, nouns, adjectives, and adverbs.
+including information about verbs that
+are reflexive or needs a particle.
+The importing method should be fast and reliable
+enough to allow us to always have a fresh version of the dictionary
+in GF.
+% nameing convention
+% unknown forms of saldo - how to handle
+% other word classes
+% description: se krasses paper
+% how to make use of other lists (idioms, use talbanken to see how pronouns are used)
+% evaluation
+
 \subsection{A tool for lexical acquisition}
+% about how this will be enhanced with more tag info, about
+% a noun guesser, use it while parsing
 To enlarge the lexicon, 
@@
 
-%% Write more here!!
-\subsection{Mapping of trees}
-%R: mapping - very nice - more about it
+\subsection{Mapping}
 The information from the tags in Talbanken can be used for many purposes.
@@
 %accomplishing an evaluation.
-
-% add evaluation
 
@@
 
-%In order to keep the grammar clean and to keep it syntactically correct,
-%much care is taken to develop it as neatly as possible.
+\section{Discussion}
+% Kommentera tabeller både i text och bildtext.
 
-\section{Evaluation and future work}
-%Compared to a statistical parser which operates like a black box,
-%a rule-based one is not only theoretically interesting,
-%but could also give a more explanatory output since the rules
-%are given their names by a human and hence act as informative labels.
-We are building a parser which is based on grammatical rules rather than
-statistics. The parser will therefore validate that
-a given sentence is grammatically correct according to the rules defined.
-Chosen parts of the parser or grammar may be used in other applications
-which deal with controlled natural languages. Automatic language
-generation from the grammar is also provided by GF.
+%Tolkning av fakta, dvs. slutsatser. Du ska göra
+% tolkningarna – inte läsaren!
+% 
+% Utvärdera slutsatserna och sätt in dem i ett större
+% sammanhang. Markera att detta är dina åsikter.
+% 
+% Ev. förslag grundade på slutsatserna, dvs. konsekvenser av
+% ditt arbete.
 
-%% more explanation of probabilities
+We intend to evaluate the parser both automatically -- by comparing the output of
+the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
+%an expert in Swedish grammar.
+She also evaluates the intermediate results.
+
+\section{Future work}
 We aim to make the parser robust by equipping it with techniques such as chunk
@@
 potential for disambiguation.
-
-We intend to evaluate the parser both automatically -- by comparing the output of
-the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
-%an expert in Swedish grammar.
-She also evaluates the intermediate results.
-
-\section{Conclusion}
-%We intend to implement a robust deep parser for Swedish, by first developing
-%a large scale-grammar. 
-By combining GF with a large lexicon, we acquire a deep parser for 
-Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
-We have developed tools for extending the lexicon with words from Talbanken,
-and we show how to make use of the information in the manually tagged
-treebank.
-The usage of GF allows us to start from
-a well-defined system for describing grammar, as well as tools for
-parsing.
-
-%We have shown how to combine existing resources, we build a large-scale grammar for Swedish.
-%The grammar can be compiled into a
-%parser by GF, and we equip this with a extensive lexicon. 
-%and we will equip this with named entity recognition, a
-%voluminous lexicon, statistical information for disambiguation,
-%a method of parallel chunk parsing etc.
-%So far we have been working on the grammar, the lexical resources and
-%a translation of the treebank Talbanken to GF.
-% adventage to what??
-
-\section{Acknowledgments}
-The work has been funded by Center of Language Technology.
-We would also like to give special thanks to Aarne Ranta, % Ramona Enache,
-Elisabet Engdahl, Krasimir Angelov, Olga Caprotti, Lars Borin and John Camilleri for their help
-and support.
-
-\nocite{*}
-\bibliographystyle{eptcs}
-\bibliography{ourbib}
-\end{document}

### 2011-11-04 doc/report/dispostion2.tex
planning and disposition

\documentclass{article} 
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

% ska gå att förstå och locka till läsning
%ska tydligt beskriva vad rapporten handlar om
%ska innehålla vettiga sökord
\title{Dispositon for master thesis \\
      \small{A large scale grammar and parser for Swedish}}
      %First results of developing a Swedish parser based on a wide-coverage grammar\\
        %Combining language resources into a grammar-driven Swedish parser} 
        %or Bootstrapping a Swedish parser \\
        %or A grammar-based Swedish parser from already existing resources}
\author{Malin Ahlberg
\institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
}
\begin{document}
\maketitle

\section{Introduction} - 
Short introduction to parsing, GF, swedish, controlled natural language vs.
free language.
Why we want a parser, what we can get from GF and what we can't get.
Which parts we are focusing on and why.

\section{Background} - 
Longer introduction to GF, Talbanken and Saldo. Other parsers and related work.

\section{Progress} - 
One part about how we have imported saldo, problems and results, how we deal
with words that were not imported. Use of lexical aquisition tools.

One part about the mapping from Talbanken to GF, description. This part explains
the tags in Talbanken and gives examples of how they are translated.

One part about the grammar. What parts that I have been working on and why, 
what constructions I wanted to add, examples of sentences that should work
and of sentences that shouldn't be covered. How to separate Swedish from
Scandinavian, bigger reconstructions in the grammar.

\section{Results and discussion}
Evaluation methods and results from all parts.

Discussion of the results and methods, why are the results good, why are they not better.
What other methods could have been used? What did we/I expect and what happend?

\section{Future Work}
what is still left to do? Description and direction of future work:
how to make the parser robust, what is still needed in the grammar, 
how to get valency information, make the parser faster.
Getting and using probabilities.

\section{Conclusion}
My conclusions about the work.

### 2011-11-11 doc/report/dispostion2.tex

@@
 
+\begin{document}
 % ska gå att förstå och locka till läsning
@@
              \small{A large scale grammar and parser for Swedish}}
-      %First results of developing a Swedish parser based on a wide-coverage grammar\\
-        %Combining language resources into a grammar-driven Swedish parser} 
-        %or Bootstrapping a Swedish parser \\
-        %or A grammar-based Swedish parser from already existing resources}
-\author{Malin Ahlberg
-\institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
-}
-\begin{document}
+\author{Malin Ahlberg}
 \maketitle
 
-\section{Introduction} - 
-Short introduction to parsing, GF, swedish, controlled natural language vs.
+\section{Introduction} 
+Short introduction to parsing, GF, Swedish, controlled natural language vs.
 free language.
-Why we want a parser, what we can get from GF and what we can't get.
-Which parts we are focusing on and why.
+Why we want a large scale grammar and a parser, what we can get from GF and
+what we can't get. Which parts we are focusing on and why.
 
-\section{Background} - 
+\section{Background}  
 Longer introduction to GF, Talbanken and Saldo. Other parsers and related work.
 
-\section{Progress} - 
+\section{Progress}  
+\begin{itemize}
+\item{
 One part about how we have imported saldo, problems and results, how we deal
 with words that were not imported. Use of lexical aquisition tools. 
-
-One part about the mapping from Talbanken to GF, description. This part explains
-the tags in Talbanken and gives examples of how they are translated.
-
-One part about the grammar. What parts that I have been working on and why, 
+}
+\item{
+One part about the mapping from Talbanken to GF, description. About the purpose.
+This part explains
+the tags in Talbanken and gives examples of how they are translated,
+examples of what can be parsed and what cannot.
+}
+\item{
+One part about the grammar. The state of the original grammar, what was covered and not. 
+What parts that I have been working on and why, 
 what constructions I wanted to add, examples of sentences that should work
@@
 Scandinavian, bigger reconstructions in the grammar. 
+}
+\end{itemize}
 
@@
 \section{Future Work}
-what is still left to do? Description and direction of future work:
+What is still left to do? Description and direction of future work:
 how to make the parser robust, what is still needed in the grammar, 
@@
 My conclusions about the work.
+\end{document}

### 2011-11-15 doc/report/report.tex
reports

\documentclass{article} 
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\begin{document}
% ska gå att förstå och locka till läsning
%ska tydligt beskriva vad rapporten handlar om
%ska innehålla vettiga sökord
\title{A large scale grammar and parser for Swedish}
\author{Malin Ahlberg}
\maketitle
\newpage

\tableofcontents
\newpage

\abstract{
% sf dream to be able to work with npl. Parser important, statistics working
% on n-grams fail, closer to the human brain with rules? 
% Describe the basic.. of Swedish, in a format compatible with 20 other languages.
% Combine other resources, create free software.
% Equivalent for English and other languages coming.
}

\newpage

\section{Acknowledgements}
% Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc

\newpage

\section{Introduction} 
What is parsing, flat strings to trees, what is deep parsing. What is controlled language,
what is robust parsing and why we need it for free langugae, benefits of a grammar.
GF is grammar formalism, many languages, many projects for controlled lang.
GF grammar for Swedish, how big, not intended for parsing, but for building
application grammars. Experiments of using the grammar for free parsing, methods
being developped for this. Benefits of starting from GF. Free parsing is
interesting. To do this, we need a better (bigger grammar), a large scale
lexicon, and an evaluation method. This project has explored those things. The
grammar also have other usages, important in itself, so we wanted to make it as
covering as possible without being overallowing or too ambiguous.

Other work exist, but none is freely available. We want to create a open source
product, therefor can only use resources that allows this. Work can continue,
will not disappear, be made use of in other projects.
What we could not cover by this project, what have been covered before (in
the spring project), and why we choose those parts.

\section{Background}  
Too long way to build a parser/grammar from scratch. Idea to combine existing parts. 
The project is heavily dependening on three resources, will describe them here.
\subsection{GF}
Most important part. Been developed for ? years, inspiration from Haskell, ML,
and also Prolog, Agda. Can be used with other programming langauges. 
Main ideas, abstract and concrete, translation, parse- and abstract trees.
What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
ideas of anaphores, montague? dependent types, ontologies.
For this project, only (mostly) important to model Swedish without abandoning the
resources too much. Use of the big grammar for translation, why it might be hard with
semantics and lexicon.
Research about how to parse with gf by Krasimir, experiments with English.
Example of a piece with Swedish, the reader gets familiar with how it looks.

\subsection{Talbanken}
For evaluating and testing, comparing our parser we needed a reliable source.
Talbanken manually anotade by ... in .. . Consists of four parts : ... . 
Also used by Maltparser and was developed in 2005 by Nivre, deeper trees.
Different parts (written, spoken) and deep/flat. Differences in mapping
on flat/deep?
Differences in analyse, do we want a similar? 
Uses of Talbanken in project, mapping to evaluate and test.  
Look at output from Maltparser, could also comparing our parse trees to this.

\subsection{Saldo}
The last(..) part is a big lexicon. The electronical Saldo is developped at GU.
Built from sal,
Contains much more information,
graph, word relationships, semantics, but also morphological info.
Example of a word, how the info looks.
About size and being maintained. Other usages?
Large, freely available and compatible with GF.

\section{Related work}
Computational ling is active area for Swedish, there are many other Swedish parser,
but grammars?
Many of the parsers not ruled based, the most known one Malt ... .
However has been two, about CLE and what happend to it. Logic and translation
to Swedish, The spoken lang. translator.
Other Sw. parsers and grammars, evaluation of them? 
Other grammar frameworks.
% related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,
% the multilingual treebank.

\section{Progress}  
 The purpose has been to prepare the grammar for large scale parsing.
 Develop method used for English to fit Swedish, pynta the sw grammar with
 language specific constructions and create methods for lexicon.
 Earlier have done tool, but the lexicon was not big enough to be able
 to fix it in this way. Tool may instead be used for complementing.

\subsection{Saldo}
Saldo has been imported before, ref, but the code needed updates to not miss words.
Have also been made more robust, print logs and list of all words that wasn't
imported together with a reason. Reson could be that it cannot construct all forms
or that the word class shouldn't be in the gf lexicon.

To understand this ... Example of gf and saldo table. Explanation of the algorithm,
how the table will be calculated. How saldo is divided to save memory, one part fail, the others 
wont, also possible to restart. Naming convention.
There are a number of differences between and saldo Till skillnad från GF,
saldo does not contain forms that aren't used gf generates all forms. A form
needed to generate the gf table may not be given in saldo such as singular
forms for 'byxor'. Also contains idioms, 'hålla huvudet kallt' and has
different grammatical analyse, a lot of pronouns. Example. 
Other word classes we do not import and why.
Those should not be in the standard gf lexicon but maybe somewhere else.
The list can be used for adding other kinds of lexicons - idioms, have pos-tags.
Also to analyse what's wrong, or use the words with a word guesser. The code
has been adapted for this.
For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
category. How this works, look at tags. Small amount, can be done manually.

The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
have any information. The earlier explained techniques may be used for this in Talbanken.
To get even bigger material, use Korp, but then no guarantee.

%see notes to add more about this, pronouns and the importing itself
%choices for adjectives vs. verbs for particips
The importing can be redone anytime.

\subsection{The grammar}
Although we can't expect to get a full coverage, we would like a tillfredsställande
coverage of standard Swedish. May also be interesting for other purposes.
Explain what's in the resources, constructions present in most languages. 
Swedish shares with Nor and Dan. Extra module, covering things like sw is a
v2-lang, focusing parts of sentences. Sw also makes use of passive s-form a lot.
kommer att - for scand, extra tempus, example
passiv s: but now need bigger change in the grammar. 
embedded - relative
pronouns/Quant
Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
           fixing this include restructuring all
Det-utr

to change position of adv:
normal: after verb 'han äpplet redan'
now want: 'äter redan äpplet'
          'har redan ätit äpplet' 
          'har ätit redan äpplet' :(

but have other words like 'bara' which can be put in other positions.
  but needs to restruct whole of VP, since no field is befor fin.
Also needed categories for 'pronouns' wich can act in different ways.
and 'antal','sorter' - need indetermined, antal plural, sorter either.
Should be able to modify by adjective, and they must be determined.

The VP require many changes, to add passive, PPart, bara.
About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
for the last one, we need VPSlash to allow adverb.

varandra and sin/sitt/sig. Not subject, but NP.

here or in robust: SupCl, det as VP or VV or S for S (remember Foc)

\subsubsection{Testing}
Hard do test, can't verify. Regression testing. Talbanken is still too hard.
Test how many parse trees that are generated.
Some results.

% One part about the grammar. The state of the original grammar, what was covered and not. 
% What parts that I have been working on and why, 
% what constructions I wanted to add, examples of sentences that should work
% and of sentences that shouldn't be covered. How to separate Swedish from
% Scandinavian, bigger reconstructions in the grammar. 
% sammanhängande explanation of the swedish grammar, choices I have made, why
% regression testing

\subsection{Mapping}
To connect GF with other annotations, to evaluate. Identify missing constructs,
missing words. More reasons from presentation.
Similar for English and Penntree, but PennTree and Talbanken different, plus swedish
makes a different. More possible word orders (or at least more commonly used).
How we try to build up a 'parse tree', a S with a NP VP for example. 
How verbs look for complements (or inverted order), how metas are added.
Description of most important tags from Talbanken, their translation.
Description of the ones most hard to translate (particles, flat sentences)
Example of nice translation and one with many metas, and one that doesn't work.
Developed by adding one mapping for each tag, and by examples.
How we cannot cover all cases. The testing: regression testing again,
very easy to destroy what was working before. 
How we need valency information.
Evaluation.

%One part about the mapping from Talbanken to GF, description. About the purpose.
%This part explains
%the tags in Talbanken and gives examples of how they are translated,
%examples of what can be parsed and what cannot.
%about how to make the code work for swedish, differences
%regression testing

\section{Results and discussion}
Results from mapping, can probably be improved, bit by bit, depends much on words.
Valencys would therefor help.  
Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
results. Mostly compound nouns, which we can't expect in the dictionary.

Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
in the process to verify the solutions. Testing trees against talbanken.

Discussion of the results and methods, why are the results good, why are they not better.
What other methods could have been used? What did we/I expect and what happend?

\section{Future Work}
What is still left to do? Description and direction of future work:
how to make the parser robust, elipses and long sentences, names and numbers.
What is still needed in the grammar, how important that is. 
how to get valency information, description of extracting thise.
make the parser faster by using smaller lexicon.
Lexicon tool on the fly.
Getting and using probabilities, to disambiguate.
process list from saldo.
Use multiligual treebank?

\section{Conclusion}
There is obviously a lot do still, will never get finish with NPL.
Grammar may need even bigger changes, to allow enough but not too much.
Different from writing a grammar for generation.
Results from mapping shows that the translation is doable, but becomes harder
since talbanken don't have formal rules.
The saldo shows how easy to make use of utomstående resources, promising if
we want to extract other information. 
Hard, but not finished!

\end{document}

### 2011-11-15 doc/report/report.tex
documentation and report

@@
 makes a different. More possible word orders (or at least more commonly used).
+One deep and one flatter version of Talbanken, deep contains ca 5 more tags.
+The flat was used since the deep one
+isn't deep in the same way as GF. Example with CNP (s5). Some
+more info about VPs are given, but do not think this will give any relevant information.
+Talbanken also contains frequent .. tags XP, XX osv. We cannot expect to translate
+this to anything but metas.
 How we try to build up a 'parse tree', a S with a NP VP for example. 
@@
 How we need valency information.
+If we don't take the verb valency into account, allow all verbs to have Gf category V,
+we get a slightly better result,
+from 62.5\% to 65.4\%.  (see EvalText). The complemnt of the verbs should tell which category
+they are anyway.
 Evaluation.
@@
 in the process to verify the solutions. Testing trees against talbanken.
+The big lexicon makes it very slow, eats all memory
 

### 2011-11-16 testSuites/ParseAndCompare.hs
passive and FocAdv working

@@ -6,10 +6,16 @@ import System.Environment (getArgs)
 import System.Process
 import Prelude hiding (compare)
 
+{-
+-- To create a new standard
+main = do
+  run "treebankTest.xml" False "newTreebank.txt"
+  -}
+
 main = do
   (file:arg) <- getArgs
   run file False "tmp.tmp"
-  correct <- readFile "treebank.txt"
+  correct <- readFile "newTreebank.txt"
   have    <- readFile "tmp.tmp"
   compare correct have
   rawSystem "rm" ["tmp.tmp"]
@@ -19,8 +25,8 @@ compare = zipWithM_ rate `on` parseResult
   where rate (i,n) (j,m) | i /= j = putStrLn $ color red 
                                        $ "Not the same sentence! "++show i++" "++show j
                          | n == m = putStrLn $ color green $ "Sentence "++show i++" ok."
-                         | n == 0 = putStrLn $ color turkos $ "Sentence "++show i++"can now be parsed."
-                         | n <  m = putStrLn $ color turkos 
+                         | n == 0 = putStrLn $ color pink $ "Sentence "++show i++"can now be parsed."
+                         | n <  m = putStrLn $ color pink 
                                        $ "Sentence "++show i++" can be parsed in "++show m++" ways!"
                                           ++ " (before only "++show n++")"
 

### 2011-11-16 testSuites/ParseTrees.hs

@@ -1,0 +1,9 @@
+import Parsing
+import System.Environment (getArgs)
+
+main = do
+  (file:arg) <- getArgs
+  run file True "Res.txt"
+  putStrLn "See result in Res.txt!"
+
+

### 2011-11-16 testSuites/Parsing.hs

@@ -86,7 +86,7 @@ addTime   = modify $ \s -> s {time = succ (time s)}
 count     = modify $ \s -> s {total = 1 + total s}
 
 putTimeMsg s   = output s blue   " was timed out" 
-putEmptyMsg s  = output s turkos " was rejected"  
+putEmptyMsg s  = output s pink " was rejected"  
 putLexMsg s ws strict | strict = output s red 
                                   (" was not parsed (unknown words: "
                                   ++unwords ws++")" )

### 2011-11-17 doc/report/report.tex
passive and FocAdv working
reflexive, updates to bara, formal object

@@
 %ska innehålla vettiga sökord
-\title{A large scale grammar and parser for Swedish}
+\title{Scaling up a GF grammar and prepare it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 Example of a piece with Swedish, the reader gets familiar with how it looks.
+Also show how information 'disappears', become strings.
 
@@
 Although we can't expect to get a full coverage, we would like a tillfredsställande
-coverage of standard Swedish. May also be interesting for other purposes.
+coverage of standard Swedish. Been working on a grammar fragment based on Talbanken,
+like cool people like Montague. May also be interesting for other purposes.
 Explain what's in the resources, constructions present in most languages. 
@@
 
-The VP require many changes, to add passive, PPart, bara.
+The VP require many changes, to add passive (no), PPart, bara.
 About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
@@
 varandra and sin/sitt/sig. Not subject, but NP.
+work pretty well by letting all subject depend on the subject in the phrase.
+Extra info, but true 'han såg sin mammas bil', they same could be applied to
+APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin kompis'. (ComparAdvAdj)
+The current version is nice and works for recursive genitivs, but atm you can't focus
+this kind of object, then a ClSlash need info about the subject. FunRP also tricky, in this
+case RP needs to give more information.
 
@@
 Hard do test, can't verify. Regression testing. Talbanken is still too hard.
+While developing: regression, 150 sentences to start with. Made longer, both good and bad,
+to see how many ways things can be parsed.
 Test how many parse trees that are generated.
@@
 Use multiligual treebank?
+sTop ambiguities by fixing so that Gen of a quant isn't a quant.
 

### 2011-11-17 doc/report/report.tex
report

@@
-\documentclass{article} 
+\documentclass{report} 
 \usepackage[utf8]{inputenc}
 \usepackage{graphicx}
+\usepackage[margin=1.5in]{geometry}
 
@@
 % Describe the basic.. of Swedish, in a format compatible with 20 other languages.
+% example of sentence and how it can be parsed, stepping out of the controlled lang.
 % Combine other resources, create free software.
 % Equivalent for English and other languages coming.
+% We aim for free parsing, this is preparation and up-scaling, finding difficulties.
 }
@@
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 % Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc
@@
 \newpage
-
+\chapter{Introduction}
 \section{Introduction} 
@@
 The project is heavily dependening on three resources, will describe them here.
+
 \subsection{GF}
@@
 Main ideas, abstract and concrete, translation, parse- and abstract trees.
+Categories, valency built in. Parameters, fields.
 What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
@@
 Also show how information 'disappears', become strings.
+
+\subsection{Swedish}
+Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
+Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
+for what will be written in the grammar part.
 
@@
 % related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,
-% the multilingual treebank.
+% the multilingual treebank, Homse&Hinchcliff, Aarne i LILT
 
-\section{Progress}  
+\chapter{Proceedings}
+%\section{Proceedings med bättre ord}  
  The purpose has been to prepare the grammar for large scale parsing.
@@
 
-\subsection{Saldo}
-Saldo has been imported before, ref, but the code needed updates to not miss words.
-Have also been made more robust, print logs and list of all words that wasn't
-imported together with a reason. Reson could be that it cannot construct all forms
-or that the word class shouldn't be in the gf lexicon.
+\section{Saldo}
+In saldo there is also semantic info, so
+We need the numbers from saldo if there are to lemmas with the same
+identifier. sluta\_V and sluta2\_V (slöt). However gf does not want more than one table
+if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
+We are just interested in morphology, so far. Dont want too big lexicon.
 
-To understand this ... Example of gf and saldo table. Explanation of the algorithm,
-how the table will be calculated. How saldo is divided to save memory, one part fail, the others 
-wont, also possible to restart. Naming convention.
+Saldo has been imported before, ref, but the code needed updates to not miss
+many important words.
+
+The basic algorith is the same. To understand this need to look at differences.
+Example of gf and saldo table.
+Explanation of the algorithm: map categories, try each paradigm, compare the table of forms
+created. If several forms in saldo, gf should pick on of them. Saldo has forms for
+compounds, gf does not. A grammar is written, the correct ones saved, the others 
+retried. Try all paradigms which can be formed from saldo. 
 There are a number of differences between and saldo Till skillnad från GF,
@@
 different grammatical analyse, a lot of pronouns. Example.
+
+The new version have also been made more robust, print logs and list of all
+words that wasn't imported together with a reason. Reson could be that it
+cannot construct all forms or that the word class shouldn't be in the gf lexicon.
+Costs very much memory, the saldo file is big, the generated lexicon is big.
+Saldo is divided to save memory. Errors should just be reported, but if something oförutsett
+happens and one part fail, the separation ensures that the others wont, also
+possible to restart. Problems with naming convention, special
+characters (å,ä,ö,apostrop,entrecôte.) needs to be translated , but avoid
+classhes 'kaeltisk' 'kältisk', gf will crash by this, so we add a 1
+in the end whenever we have changed a letter.
+
 Other word classes we do not import and why.
@@
 
-\subsection{The grammar}
+\section{The grammar}
 Although we can't expect to get a full coverage, we would like a tillfredsställande
@@
 kommer att - for scand, extra tempus, example
-passiv s: but now need bigger change in the grammar. 
-embedded - relative
-pronouns/Quant
+passiv s: but wich verbs can be made passive? 
+embedded - relative, 'sådan att' not really nice for normal swedish, and no agreement
+between subject 'en katt sådan att det regnar'.
+pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
+has to decide, one category for each?
 Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
            fixing this include restructuring all
-Det-utr
+Det-utr. not effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
+ Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
+ so added one more. Leads to ambiguities in some cases, but considering that we would want to use
+ this for translation, neutr and utr may differ in the other language.
 
+-- or is this really just AdVs? 
 to change position of adv:
@@
           'har redan ätit äpplet' 
-          'har ätit redan äpplet' :(
+          'har ätit redan äpplet' :( which doesn't work either
 
@@
   but needs to restruct whole of VP, since no field is befor fin.
+  Can also act before the noun or as AdV. Examples.
+\verb|s SPres Simul Neg Main : han bara sover inte alltid|
+  have added a field in VP, a0, show table.
+  makes some weird constructions when using modal verbs, 
+  \verb|s SPres Anter Pos Main : han bara har alltid sovit|
+  but rather than changing to 'han har bara sovit', we keep it. Is a possible
+  tolkning and the other case is covered by using it as an AdV, we don't want
+  ambiguities.
+
 Also needed categories for 'pronouns' wich can act in different ways.
 and 'antal','sorter' - need indetermined, antal plural, sorter either.
-Should be able to modify by adjective, and they must be determined.
+Should be able to modify by a godtyckligt antal adjective, and they must be determined.
+'(ett stort antal) människor' and not 'ett stort (antal människor)'
+Don't want them as a NPs or CNs since we don't want other Ns to be used this way
+'en flicka kaffe', this is then Appos? (Or do we?). Should also not be utsatta for what other
+NPs or CN can be. Dessutom, we want to make use of the comp-fiel. 'mamma till', but is this really
+the same??
 
@@
 varandra and sin/sitt/sig. Not subject, but NP.
-work pretty well by letting all subject depend on the subject in the phrase.
-Extra info, but true 'han såg sin mammas bil', they same could be applied to
-APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin kompis'. (ComparAdvAdj)
-The current version is nice and works for recursive genitivs, but atm you can't focus
-this kind of object, then a ClSlash need info about the subject. FunRP also tricky, in this
-case RP needs to give more information.
+do not want to copy all rules for another category. Would like depent types!
+work pretty well by letting all NP depend on the subject in the phrase.
+The current version is nice and works for recursive genitivs!
+Extra info, but maybe this is just the case about NPs.
+'han såg sin mammas bil', they same could be applied to
+APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin
+kompis'. (ComparAdvAdj)
+Atm you can't focus this kind of object, then a ClSlash need info about the
+subject. FunRP also tricky, in this case RP needs to give more information.
+Kort sagt, the information about the subject must be spread
+in different parts of the sentence. 
+Could also decide to always produc 'sin' -> jag såg sin bil
+to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
+in third person, and add nonexist for others. But this way we can also get rid
+of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
+'jag såg din bil'.
+
+In general, many categories need more parameters and fields, tex Quant can be used as NP
+and therefor needs to depend on Case, de -> dem.
 
@@
 
-\subsubsection{Testing}
+\subsection{Testing}
 Hard do test, can't verify. Regression testing. Talbanken is still too hard.
@@
 
-\subsection{Mapping}
+\section{Mapping}
 To connect GF with other annotations, to evaluate. Identify missing constructs,
@@
 
+\chapter{Summary}
 \section{Results and discussion}
@@
 What is still needed in the grammar, how important that is. 
-how to get valency information, description of extracting thise.
+how to get valency information 'sitter och läser', description of extracting this
+or handling it by the robust parser or the grammar itself.
 make the parser faster by using smaller lexicon.
@@
 Use multiligual treebank?
-sTop ambiguities by fixing so that Gen of a quant isn't a quant.
+stop ambiguities by fixing so that Gen of a quant isn't a quant, or better, that all fields
+have a genitive field.
 

### 2011-11-18 doc/report/report.tex
report

@@
 \usepackage[margin=1.5in]{geometry}
+\usepackage{lingmacros}
 
@@
 %ska innehålla vettiga sökord
-\title{Scaling up a GF grammar and prepare it for cool stuff for Swedish}
+\title{Scaling up a GF grammar and prepare it for cool stuff and for Swedish}
 \author{Malin Ahlberg}
@@
 Main ideas, abstract and concrete, translation, parse- and abstract trees.
-Categories, valency built in. Parameters, fields.
+Example of abstract.
+Categories, valency built in (problems with ställa - på, under, vid). Parameters, fields.
+Example of a piece with Swedish, the reader gets familiar with how it looks.
+Also show how information 'disappears', become strings.
 What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
@@
 Research about how to parse with gf by Krasimir, experiments with English.
-Example of a piece with Swedish, the reader gets familiar with how it looks.
-Also show how information 'disappears', become strings.
 
@@
 for what will be written in the grammar part.
+Since the s-passive is the most common one in written Swedish \ref{SAG-34-1}, it has
 
@@
 like cool people like Montague. May also be interesting for other purposes.
+\subsection{What was in the resources}
 Explain what's in the resources, constructions present in most languages. 
 Swedish shares with Nor and Dan. Extra module, covering things like sw is a
-v2-lang, focusing parts of sentences. Sw also makes use of passive s-form a lot.
+v2-lang, focusing parts of sentences. 
+kommer att.
+
+\subsection{Added}
+Summary of what have been added and motivation.
+
 kommer att - for scand, extra tempus, example
-passiv s: but wich verbs can be made passive? 
-embedded - relative, 'sådan att' not really nice for normal swedish, and no agreement
-between subject 'en katt sådan att det regnar'.
+
+%%PASSIV
+Is passive also more used in Swedish?
+There are two ways of forming passive verb phrases in Swedish: the 
+periphrastic passive, formed by using a modale auxiliary verb 
+\enumsentence{Bilden blev tagen av  av rymdsonden Galileo}
+\enumsentence{The picture was taken by spacecraft Galileo.}
+or by adding an \emph{s} to
+the verb:
+\enumsentence{Bilden togs av rymdsonden Galileo}
+\enumsentence{The picture [took+s] by spacecraft Galileo.}
+This second variant, the s-passive, is commonly used in Swedish, some research
+suggest that it is used in more than 80\% of the times \ref{laanemets}.
+The other scandinavian languages prefers the periphrastic passive, especially in spoken langauge.
+\enumsentence{Bildet ble tatt av romsonden (NO)}
+The resource grammar for Scandinavian therefore allowed/expressed constructions using auxiliary verb. \\
+\verb|PassV2 : V2  -> VP | \\
+\verb|         äta -> blev äten| \\
+This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
+turned into comlpete verb phrases; they no longer need an object.
+The s-passive has now been added as the standard case for the Swedish GF
+grammar. Periphrastics passive is still allowed, but as an alternative rather
+than the default.
+The grammar also allows using three-place verbs \verb|V3| , as \emph{give} to
+form passives:
+\enumsentence{Betyget 3 ges till många elever}
+\label{ex:passV3}
+by letting the first object be `removed' and given by the subject. 
+Sentences where the second object is put first in the sentence, like 
+\enumsentence{Till många elever ges betyget 3}
+is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
+parse this by applying \verb|FocObj|.
+%%
+
+%% Clauses
+\subsection{The clause-level}
+The resource grammar already provided a number of ways to express relative and embedded clauses, such as
+\enumsentence{Frågan är vilken färg den har\\
+                   Pojken, som är blyg, tystnar\\
+                   Han såg kunden som tyckte om sallad\\
+                   Jag tänkte på huset vilket hon bodde i} % better example!!
+
+Since those constructions are inspired by other languages, some of them may sound quite
+non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
+and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
+\emph{`She sleeps, which is good'} as
+\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
+embedded in the relative clause. To fix this, an extra parameter was needed to tell whether
+\emph{`som'} or \emph{`vilket'} should be used. \\
+Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
+The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
+\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
+An more natural sounding alternative would be
+\enumsentence{Jag vill ha en sådan katt som inte fäller.}
+and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
+'en katt sådan att det regnar'. \\
+
+Formal subjects \ref{SAG-19} is often used in Swedish
+\enumsentence{Det sitter en fågel på taket}
+\emph{`There is a bird sitting on the roof'}
+or, literally
+\emph{`It sits a bird on the roof'}.\\
+\emph{`Det'} has the position of the subject, and the egentligt subject, 
+\emph{`en fågel'} the one of an object.
+Transitive verbs may not be used like this
+\enumsentence{*Det äter en fågel på taket.}
+unless their in passive form
+\enumsentence{Det äts en fågel på taket.}
+There are also restrictions on the nounphrase of the subject, it can only
+be determined in certain ways
+\enumsentence{*Det sitter den fågel på taket. \\
+              **Det sitter fåglarna på taket.}
+This distinction is not easily captured in the formal rules of GF. 
+Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
+not be used, but the determiners \emph{`många'}, \emph{`några'} and the quantifier
+\emph{`ingen'} work well.
+Semantic difference, atm we allow any NP. \\
+
+'det är kul att..'
+%%
+%embedded - relative, 'sådan att' not really nice for normal swedish, and no agreement
+%between subject 'en katt sådan att det regnar'.
 pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
@@
 'han såg sin mammas bil', they same could be applied to
-APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin
-kompis'. (ComparAdvAdj)
+APs for
+\enumsentence{Han är längre än sin kompis}
+\label{ex:ReflAP}
+ or Adv for
+\enumsentence{Han leker oftare än sin kompis}
+ (ComparAdvAdj)
 Atm you can't focus this kind of object, then a ClSlash need info about the
@@
 'jag såg din bil'.
+Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
+not 'alla sina', usw. Want it to handle conjuction 
+\enumsentence{Han såg sin hund och sitt hus}
+etc.
+Test to example \ref{ex:ReflAP}
 
@@
 have a genitive field.
+More fields in the grammar? Or leave it all to the robust parser
 

### 2011-11-21 doc/report/Refl.tex
report, and planning

--- 2011-11-17 doc/report/report.tex
@@
-\documentclass{article} 
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}
+varandra and sin/sitt/sig. Not subject, but NP.
+Want the following:
+\enumsentence{Han såg sina barns skor.}
+\enumsentence{Sina vantar hade han glömt på tåget.}
+but not
+\enumsentence{*Sina vantar var kvar på tåget.}
+\enumsentence{Han ger sina pengar till sina syskon.}
+\enumsentence{*Jag ger sina pengar till sina syskon.}
+\enumsentence{Han är längre än sin kompis.}
+\enumsentence{Hon tyckte om alla sina elever.}
+\enumsentence{Han såg sin hund och sitt hus}
+\label{ex:refGen}
+\enumsentence{Han är längre än sin kompis}
+\label{ex:ReflAP}
 
-\begin{document}
-% ska gå att förstå och locka till läsning
-%ska tydligt beskriva vad rapporten handlar om
-%ska innehålla vettiga sökord
-\title{Scaling up a GF grammar and prepare it for cool stuff for Swedish}
-\author{Malin Ahlberg}
-\maketitle
-\newpage
+Suggest a connection between the subject and the object.
+Could be combined with 'sig', a ReflNP without another noun.
+Make a distinction between objects and subjects?
+do not want to copy all rules for another category. Would like depent types!
+work pretty well by letting all NP depend on the subject in the phrase.
+The current version is nice and works for recursive genitivs!
+Extra info, but maybe this is just the case about NPs.
+Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
+*Slash* : VPSlash -> NPR -> VP ;
+Keep number of fields down, but introduces another category, and one more
+label to the parse tree, is it interesting? How to combine it with genitive 
+without making it ambiguous (Han såg barnens hund).
+What if we add alot of other rules that uses NP, will they all need to be
+duplicated then?
 
-\tableofcontents
-\newpage
-
-\abstract{
-% sf dream to be able to work with npl. Parser important, statistics working
-% on n-grams fail, closer to the human brain with rules? 
-% Describe the basic.. of Swedish, in a format compatible with 20 other languages.
-% Combine other resources, create free software.
-% Equivalent for English and other languages coming.
-}
-
-\newpage
-
-\section{Acknowledgements}
-% Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc
-
-\newpage
-
-\section{Introduction} 
-What is parsing, flat strings to trees, what is deep parsing. What is controlled language,
-what is robust parsing and why we need it for free langugae, benefits of a grammar.
-GF is grammar formalism, many languages, many projects for controlled lang.
-GF grammar for Swedish, how big, not intended for parsing, but for building
-application grammars. Experiments of using the grammar for free parsing, methods
-being developped for this. Benefits of starting from GF. Free parsing is
-interesting. To do this, we need a better (bigger grammar), a large scale
-lexicon, and an evaluation method. This project has explored those things. The
-grammar also have other usages, important in itself, so we wanted to make it as
-covering as possible without being overallowing or too ambiguous.
-
-Other work exist, but none is freely available. We want to create a open source
-product, therefor can only use resources that allows this. Work can continue,
-will not disappear, be made use of in other projects.
-What we could not cover by this project, what have been covered before (in
-the spring project), and why we choose those parts.
-
-\section{Background}  
-Too long way to build a parser/grammar from scratch. Idea to combine existing parts. 
-The project is heavily dependening on three resources, will describe them here.
-\subsection{GF}
-Most important part. Been developed for ? years, inspiration from Haskell, ML,
-and also Prolog, Agda. Can be used with other programming langauges. 
-Main ideas, abstract and concrete, translation, parse- and abstract trees.
-What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
-ideas of anaphores, montague? dependent types, ontologies.
-For this project, only (mostly) important to model Swedish without abandoning the
-resources too much. Use of the big grammar for translation, why it might be hard with
-semantics and lexicon.
-Research about how to parse with gf by Krasimir, experiments with English.
-Example of a piece with Swedish, the reader gets familiar with how it looks.
-Also show how information 'disappears', become strings.
-
-\subsection{Talbanken}
-For evaluating and testing, comparing our parser we needed a reliable source.
-Talbanken manually anotade by ... in .. . Consists of four parts : ... . 
-Also used by Maltparser and was developed in 2005 by Nivre, deeper trees.
-Different parts (written, spoken) and deep/flat. Differences in mapping
-on flat/deep?
-Differences in analyse, do we want a similar? 
-Uses of Talbanken in project, mapping to evaluate and test.  
-Look at output from Maltparser, could also comparing our parse trees to this.
-
-\subsection{Saldo}
-The last(..) part is a big lexicon. The electronical Saldo is developped at GU.
-Built from sal,
-Contains much more information,
-graph, word relationships, semantics, but also morphological info.
-Example of a word, how the info looks.
-About size and being maintained. Other usages?
-Large, freely available and compatible with GF.
-
-\section{Related work}
-Computational ling is active area for Swedish, there are many other Swedish parser,
-but grammars?
-Many of the parsers not ruled based, the most known one Malt ... .
-However has been two, about CLE and what happend to it. Logic and translation
-to Swedish, The spoken lang. translator.
-Other Sw. parsers and grammars, evaluation of them? 
-Other grammar frameworks.
-% related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,
-% the multilingual treebank.
-
-\section{Progress}  
- The purpose has been to prepare the grammar for large scale parsing.
- Develop method used for English to fit Swedish, pynta the sw grammar with
- language specific constructions and create methods for lexicon.
- Earlier have done tool, but the lexicon was not big enough to be able
- to fix it in this way. Tool may instead be used for complementing.
-
-\subsection{Saldo}
-Saldo has been imported before, ref, but the code needed updates to not miss words.
-Have also been made more robust, print logs and list of all words that wasn't
-imported together with a reason. Reson could be that it cannot construct all forms
-or that the word class shouldn't be in the gf lexicon.
-
-To understand this ... Example of gf and saldo table. Explanation of the algorithm,
-how the table will be calculated. How saldo is divided to save memory, one part fail, the others 
-wont, also possible to restart. Naming convention.
-There are a number of differences between and saldo Till skillnad från GF,
-saldo does not contain forms that aren't used gf generates all forms. A form
-needed to generate the gf table may not be given in saldo such as singular
-forms for 'byxor'. Also contains idioms, 'hålla huvudet kallt' and has
-different grammatical analyse, a lot of pronouns. Example. 
-Other word classes we do not import and why.
-Those should not be in the standard gf lexicon but maybe somewhere else.
-The list can be used for adding other kinds of lexicons - idioms, have pos-tags.
-Also to analyse what's wrong, or use the words with a word guesser. The code
-has been adapted for this.
-For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
-category. How this works, look at tags. Small amount, can be done manually.
-
-The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
-have any information. The earlier explained techniques may be used for this in Talbanken.
-To get even bigger material, use Korp, but then no guarantee.
-
-%see notes to add more about this, pronouns and the importing itself
-%choices for adjectives vs. verbs for particips
-The importing can be redone anytime.
-
-\subsection{The grammar}
-Although we can't expect to get a full coverage, we would like a tillfredsställande
-coverage of standard Swedish. Been working on a grammar fragment based on Talbanken,
-like cool people like Montague. May also be interesting for other purposes.
-Explain what's in the resources, constructions present in most languages. 
-Swedish shares with Nor and Dan. Extra module, covering things like sw is a
-v2-lang, focusing parts of sentences. Sw also makes use of passive s-form a lot.
-kommer att - for scand, extra tempus, example
-passiv s: but now need bigger change in the grammar. 
-embedded - relative
-pronouns/Quant
-Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
-           fixing this include restructuring all
-Det-utr
-
-to change position of adv:
-normal: after verb 'han äpplet redan'
-now want: 'äter redan äpplet'
-          'har redan ätit äpplet' 
-          'har ätit redan äpplet' :(
-
-but have other words like 'bara' which can be put in other positions.
-  but needs to restruct whole of VP, since no field is befor fin.
-Also needed categories for 'pronouns' wich can act in different ways.
-and 'antal','sorter' - need indetermined, antal plural, sorter either.
-Should be able to modify by adjective, and they must be determined.
-
-The VP require many changes, to add passive (no), PPart, bara.
-About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
-for the last one, we need VPSlash to allow adverb.
-
-varandra and sin/sitt/sig. Not subject, but NP.
-work pretty well by letting all subject depend on the subject in the phrase.
-Extra info, but true 'han såg sin mammas bil', they same could be applied to
-APs for 'han är längre än sin kompis' or Adv for 'han leker oftare än sin kompis'. (ComparAdvAdj)
-The current version is nice and works for recursive genitivs, but atm you can't focus
-this kind of object, then a ClSlash need info about the subject. FunRP also tricky, in this
-case RP needs to give more information.
-
-here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
-
-\subsubsection{Testing}
-Hard do test, can't verify. Regression testing. Talbanken is still too hard.
-While developing: regression, 150 sentences to start with. Made longer, both good and bad,
-to see how many ways things can be parsed.
-Test how many parse trees that are generated.
-Some results.
-
-% One part about the grammar. The state of the original grammar, what was covered and not. 
-% What parts that I have been working on and why, 
-% what constructions I wanted to add, examples of sentences that should work
-% and of sentences that shouldn't be covered. How to separate Swedish from
-% Scandinavian, bigger reconstructions in the grammar. 
-% sammanhängande explanation of the swedish grammar, choices I have made, why
-% regression testing
-
-\subsection{Mapping}
-To connect GF with other annotations, to evaluate. Identify missing constructs,
-missing words. More reasons from presentation.
-Similar for English and Penntree, but PennTree and Talbanken different, plus swedish
-makes a different. More possible word orders (or at least more commonly used).
-One deep and one flatter version of Talbanken, deep contains ca 5 more tags.
-The flat was used since the deep one
-isn't deep in the same way as GF. Example with CNP (s5). Some
-more info about VPs are given, but do not think this will give any relevant information.
-Talbanken also contains frequent .. tags XP, XX osv. We cannot expect to translate
-this to anything but metas.
-How we try to build up a 'parse tree', a S with a NP VP for example. 
-How verbs look for complements (or inverted order), how metas are added.
-Description of most important tags from Talbanken, their translation.
-Description of the ones most hard to translate (particles, flat sentences)
-Example of nice translation and one with many metas, and one that doesn't work.
-Developed by adding one mapping for each tag, and by examples.
-How we cannot cover all cases. The testing: regression testing again,
-very easy to destroy what was working before. 
-How we need valency information.
-If we don't take the verb valency into account, allow all verbs to have Gf category V,
-we get a slightly better result,
-from 62.5\% to 65.4\%.  (see EvalText). The complemnt of the verbs should tell which category
-they are anyway.
-Evaluation.
-
-%One part about the mapping from Talbanken to GF, description. About the purpose.
-%This part explains
-%the tags in Talbanken and gives examples of how they are translated,
-%examples of what can be parsed and what cannot.
-%about how to make the code work for swedish, differences
-%regression testing
-
-\section{Results and discussion}
-Results from mapping, can probably be improved, bit by bit, depends much on words.
-Valencys would therefor help.  
-Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
-results. Mostly compound nouns, which we can't expect in the dictionary.
-
-Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
-in the process to verify the solutions. Testing trees against talbanken.
-The big lexicon makes it very slow, eats all memory
-
-Discussion of the results and methods, why are the results good, why are they not better.
-What other methods could have been used? What did we/I expect and what happend?
-
-\section{Future Work}
-What is still left to do? Description and direction of future work:
-how to make the parser robust, elipses and long sentences, names and numbers.
-What is still needed in the grammar, how important that is. 
-how to get valency information, description of extracting thise.
-make the parser faster by using smaller lexicon.
-Lexicon tool on the fly.
-Getting and using probabilities, to disambiguate.
-process list from saldo.
-Use multiligual treebank?
-sTop ambiguities by fixing so that Gen of a quant isn't a quant.
-
-\section{Conclusion}
-There is obviously a lot do still, will never get finish with NPL.
-Grammar may need even bigger changes, to allow enough but not too much.
-Different from writing a grammar for generation.
-Results from mapping shows that the translation is doable, but becomes harder
-since talbanken don't have formal rules.
-The saldo shows how easy to make use of utomstående resources, promising if
-we want to extract other information. 
-Hard, but not finished!
-
-\end{document}
+'han såg sin mammas bil', they same could be applied to
+APs for \ref{ex:ReflAP} or Adv for
+\enumsentence{Han leker oftare än sin kompis}
+ (ComparAdvAdj)
+Atm you can't focus this kind of object, then a ClSlash need info about the
+subject. FunRP also tricky (and what does it mean, we can say
+Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
+Kort sagt, the information about the subject must be spread
+in different parts of the sentence. 
+Could also decide to always produc 'sin' -> jag såg sin bil
+to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
+in third person, and add nonexist for others. But this way we can also get rid
+of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
+'jag såg din bil'.
+Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
+not 'alla sina', usw. Want it to handle conjuction as in \ref{ex:refGen}
+etc.

### 2011-11-21 doc/report/report.tex
report, and planning

@@
 %ska innehålla vettiga sökord
-\title{Scaling up a GF grammar and prepare it for cool stuff and for Swedish}
+\title{Scaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 \abstract{
+% This is the report of the work of expanding a GF grammar and preparing
+% it for parsing of free Swedish. A lexicon consisting of more than 100 000 words
+% has been extracted and we have constructed an automatic mapping between the
+% annotation of the treebank Talbanken and GF. 
+% The goal is to build a robust parser dealing with uncontrolled natural
+% language, a parser which, for example, can parse all of Talbanken.
+% Outside of this project, an equivalent GF project for English is being worked upon,
+% as are theqniques for making the parser robust.
+% With this in mind, we have extended the current GF grammar to adapt it for
+% Swedish specifically so that it can parse the an extended set of basic
+% constructions. We have completed a method for importing the eletronical 
+% lexicon Saldo to a GF format and upprätat en connection mellan oss och Talbanken.
+
 % sf dream to be able to work with npl. Parser important, statistics working
@@
 \section{Background}  
-Too long way to build a parser/grammar from scratch. Idea to combine existing parts. 
-The project is heavily dependening on three resources, will describe them here.
+To build a parser or a grammar from scratch is a heavy work. The idea for this
+project is to combine already existing parts into a bigger and more expressive grammar
+as well as a large scale GF lexicon.
+As all GF grammars, this one can be used for parsing, and we develop it by getting
+examples, ideas and test material from a treebank.
+The project is hence heavily dependening on three resources, which will be described
+in the following sections.
 
 \subsection{GF}
-Most important part. Been developed for ? years, inspiration from Haskell, ML,
-and also Prolog, Agda. Can be used with other programming langauges. 
-Main ideas, abstract and concrete, translation, parse- and abstract trees.
-Example of abstract.
-Categories, valency built in (problems with ställa - på, under, vid). Parameters, fields.
-Example of a piece with Swedish, the reader gets familiar with how it looks.
-Also show how information 'disappears', become strings.
-What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
-ideas of anaphores, montague? dependent types, ontologies.
-For this project, only (mostly) important to model Swedish without abandoning the
-resources too much. Use of the big grammar for translation, why it might be hard with
-semantics and lexicon.
-Research about how to parse with gf by Krasimir, experiments with English.
+The main component of the the project is the Grammatical Framework\ref{gfbok} (GF). It is
+a functional grammar formalism based on Martin-Löf type theory.
+It has been developed for about 20? years and are inspired by programming langauges
+like Haskell\footnote{}, ML\footnote{}, Prolog\footnote{} and Agda\footnote{}. Grammars written in GF can be used together with languages 
+like Java, Java script and Haskell and works for both parsing and generation.
+
+GF is developed with multilinguality in mind and makes an important distiction between
+\textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
+shared by many languages, it models the semantics of the domain without blanda in
+language specific features such as agreement or word order.
+
+\begin{figure}[h]
+\begin{verbatim}
+ abstract TestGrammar {
+  cat N ; V ; S ;
+
+  fun 
+    Pred : N -> V -> S ;
+ }
+\end{verbatim}
+\caption{A small example of an abstract syntax}
+\end{figure}
+
+The example shows an abstract grammar defining three \textit{Categories},
+one for nouns, one for verbs and one for sentences. An abstract grammar also gives
+the types of functions. In this case we have \verb|Pred| which
+tells us that by taking a noun and a verb we can form a sentence. It does not
+give any information of how this is done, or how any of the categories should look.
+maybe continue example with swedish and english, have DetNP : N -> N wich
+gives 'the cat', 'katten'. Thereby introduce tables, parameters, fields.
+Show abstract tree.
+
+We now have two implementations of the abstract. The resulting GF grammar is able to
+both parse a string to the abstract tree and to go in the other direction; to produce
+a string of natural language given a abstract tree. This step is called linearization.
+Translation is a consequence of this. Picture of sw. string -> tree -> eng. tree.
+
+Valency information is normally built into the GF categories. Add V2 to the example.
+Point out how N + V can be glued together to a string, the components cannot be seen
+anymore.
+
+A GF grammar usually covers a very restricted set of language, where few ambiguities
+occur and much care can be taken to preserving the semantics between the different languages.
+By using a controlled language like this,  good and reliable translation can be
+ensured. Some examples of projects using GF are the math .. \cite{olga}, the
+Online Syllogism Solver\cite{malin!}.  The framework is also used in the
+european project MOLTO for online translation between upto 15 language.\\
+
+One of the best... features of the framework is the resource libraries. Provided in the 
+GF package, those gives a good start for all who want to write an application grammar.
+Instead of letting the users reimplementing the morphological and the fundamental
+syntactic structures of a language for each application, all this can be found
+as libraries. So far, 20 languages are implemented all sharing the same abstract syntax.
+They describe how to conjugate words, and how to construct phrases and sentences.
+There are also a small test lexicon included, containing some hundered of common words.\\
+
+As an open source-project, GF is constantely being developed and improved. New
+languges are added, the compiler is being improved, ways of använad det på ett
+lätt sätt are added and the possiblities to use GF in different environments
+increased. There are  research about how to make better use of the dependent
+types, for solving syllogism or generating natural language via Montague
+semantics.
+Further, experiments on using GF for free parsing has been conducted (Angelov,2010?).
+This is also a future goal for our project and one of the delmål has been to
+model an important part of Swedish in GF. The starting point was the resource
+grammar and the result should still have a strong connection to this.
+
+%Main ideas, abstract and concrete, translation, parse- and abstract trees.
+%Example of abstract.
+%We say that this is a function that takes as argument and returns..
+%Categories, valency built in (problems with ställa - på, under, vid). Parameters, fields.
+%Example of a piece with Swedish, the reader gets familiar with how it looks.
+%Also show how information 'disappears', become strings.
+%What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
+%ideas of anaphores, montague? dependent types, ontologies, cool pictures by Krasse.
+%
+%The resources, why, how many, how to use, projects. functors.
+%
+%Still under big development, evolving, growing.
+%For this project, only (mostly) important to model Swedish without abandoning the
+%resources too much. Use of the big grammar for translation, why it might be hard with
+%semantics and lexicon.
+%Research about how to parse with gf by Krasimir, experiments with English.
 
@@
 for what will be written in the grammar part.
-Since the s-passive is the most common one in written Swedish \ref{SAG-34-1}, it has
+Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag är stor, I am big
+Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
 
 \subsection{Talbanken}
-For evaluating and testing, comparing our parser we needed a reliable source.
-Talbanken manually anotade by ... in .. . Consists of four parts : ... . 
-Also used by Maltparser and was developed in 2005 by Nivre, deeper trees.
-Different parts (written, spoken) and deep/flat. Differences in mapping
-on flat/deep?
-Differences in analyse, do we want a similar? 
-Uses of Talbanken in project, mapping to evaluate and test.  
-Look at output from Maltparser, could also comparing our parse trees to this.
+For testing and evaluation of the grammar and lexicon, we needed to be able to
+compare them against a reliable source.
+Being an freely avaible, manually annotade, large scale treebank,
+Talbanken\cite{talbanken} was perfect for this ändamål.
+It was created in the 1970s at Lund University. In total it consists of four
+parts. Two of them are transcribitions of spoken language, one a collection of
+text written by high school (?) students, and the last one, section P,
+consists of professionally written Swedish gathered from newspapers and ... .\\
+Talbanken was also used to train the Swedish version of the Maltparser\cite{mal}
+and was därmed reutgiven in a more modern version\cite{nivre}. This work
+was done by Nivre and ... in 2005. They released it in some different XML-formats,
+where the trees had been made deeper and smt more. \\
+In this project, as well as for the Malt parser, the more than 6000 sentences
+in section P has been used.
+The treebank has served as an inspiration and an evaluation source throughout this
+project. An automatic mapping between its trees and the abstract trees from GF has been
+done, which will be explained in section \ref{sec:Mapping}.
+%Differences in analyse, do we want a similar? 
+%Uses of Talbanken in project, mapping to evaluate and test.  
+%Look at output from Maltparser, could also comparing our parse trees to this.
 
 \subsection{Saldo}
-The last(..) part is a big lexicon. The electronical Saldo is developped at GU.
-Built from sal,
-Contains much more information,
-graph, word relationships, semantics, but also morphological info.
-Example of a word, how the info looks.
-About size and being maintained. Other usages?
-Large, freely available and compatible with GF.
+A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
+large electronic lexicon developed and maintained at Gothenburg University. It is
+built upon Svenskt Associations Lexicon and contains information about modern
+Swedish and the words are equipped with both sematic and a morphological
+information. The user can find compounding analyses, examples of usage in corpora,
+and graphs of sematically connected words.\\
+Example of katt? \\
+For our purpose however, only the morphological data was needed.\\
+Example of morpho for katt.\\
+Saldo provides this in XML format under LPGL. The data can be processed and translated to GF format as described in 
+section \ref{sec:prog.saldo}.
+Other usages?
 
@@
 \section{Saldo}
+\label{sec:prog.saldo}
 In saldo there is also semantic info, so
@@
 
-Saldo has been imported before, ref, but the code needed updates to not miss
-many important words.
+Saldo has been imported before (Krasimir, 2009?), but the code needed updates
+to not miss many important words.
 
@@
 Although we can't expect to get a full coverage, we would like a tillfredsställande
+Explain what's in the resources, constructions present in most languages.
+
+\subsection{GF for Swedish}
+What was in the resources
+description of the existing grammar, the types, the possibilities
+VP, NP.
+valency built in (problems with ställa - på, under, vid).
+Determiners, quants, predets, The difference from how pronouns are regarded. 
+Variants, nonexist.
 coverage of standard Swedish. Been working on a grammar fragment based on Talbanken,
 like cool people like Montague. May also be interesting for other purposes.
-\subsection{What was in the resources}
-Explain what's in the resources, constructions present in most languages. 
 Swedish shares with Nor and Dan. Extra module, covering things like sw is a
 v2-lang, focusing parts of sentences.
-kommer att.
 
@@
 kommer att - for scand, extra tempus, example
+example table
 
@@
 This second variant, the s-passive, is commonly used in Swedish, some research
-suggest that it is used in more than 80\% of the times \ref{laanemets}.
+suggest that it is used in more than 80\% of the times \cite{laanemets}.
 The other scandinavian languages prefers the periphrastic passive, especially in spoken langauge.
@@
 and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
-'en katt sådan att det regnar'. \\
+'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
+needed.\\
 
-Formal subjects \ref{SAG-19} is often used in Swedish
+Formal subjects \cite{SAG-19} is often used in Swedish
 \enumsentence{Det sitter en fågel på taket}
@@
 \emph{`It sits a bird on the roof'}.\\
-\emph{`Det'} has the position of the subject, and the egentligt subject, 
+\emph{`Det'} has the position of the subject, and the real subject, 
 \emph{`en fågel'} the one of an object.
@@
               **Det sitter fåglarna på taket.}
-This distinction is not easily captured in the formal rules of GF. 
-Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
-not be used, but the determiners \emph{`många'}, \emph{`några'} and the quantifier
-\emph{`ingen'} work well.
-Semantic difference, atm we allow any NP. \\
+The distinction lays in the definitness of the noun itself, (not the whole noun phrase..)
+it must be indefinite, even though determiners are allowed. Jfr
+\enumsentence{Det sitter några fåglar på taket\\
+              **Det sitter samtliga fåglar på taket}
+%%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
+%%not be used, but the determiners \emph{`många'}, \emph{`några'} and the quantifier
+%%\emph{`ingen'} work well.
+%%Semantic difference, atm we allow any NP. \\
+The grammar handles this by looking at the determiner used
+\verb|FormalSub : SimpleVP -> Det -> NP -> Cl ;|
+The function combines a verb phrase a determiner and a noun phrase in to a clause.
+\verb|SimpleVP| is a normal VP, but with the restriction that the main verb must be
+intransitiv, a \verb|V|, (and passives?). If the  determiner requires a noun in determinened
+form, the clause is put to NONEXIST.
 
-'det är kul att..'
+'det är kul att..' Näst vanligaste två-ordskombinationen i talad svenska.
 %%
@@
           'har redan ätit äpplet' 
-          'har ätit redan äpplet' :( which doesn't work either
+          'har ätit redan äpplet' :( which doesn't work either :)
 
@@
 
+%% Refl 
+Could varandra be thought of as reflexive for 'oss'?
 varandra and sin/sitt/sig. Not subject, but NP.
+Want the following:
+\enumsentence{Han såg sina barns skor.}
+\enumsentence{Sina vantar hade han glömt på tåget.}
+but not
+\enumsentence{*Sina vantar var kvar på tåget.}
+\enumsentence{Han ger sina pengar till sina syskon.}
+\enumsentence{*Jag ger sina pengar till sina syskon.}
+\enumsentence{Han är längre än sin kompis.}
+\enumsentence{Hon tyckte om alla sina elever.}
+\enumsentence{Han såg sin hund och sitt hus}
+\label{ex:refGen}
+\enumsentence{Han är längre än sin kompis}
+\label{ex:ReflAP}
+
+Suggest a connection between the subject and the object.
+Could be combined with 'sig', a ReflNP without another noun.
+Make a distinction between objects and subjects?
 do not want to copy all rules for another category. Would like depent types!
@@
 Extra info, but maybe this is just the case about NPs.
+Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
+*Slash* : VPSlash -> NPR -> VP ;
+Keep number of fields down, but introduces another category, and one more
+label to the parse tree, is it interesting? How to combine it with genitive 
+without making it ambiguous (Han såg barnens hund).
+What if we add alot of other rules that uses NP, will they all need to be
+duplicated then?
+
 'han såg sin mammas bil', they same could be applied to
-APs for
-\enumsentence{Han är längre än sin kompis}
-\label{ex:ReflAP}
- or Adv for
+APs for example \ref{ex:ReflAP} or Adv for
 \enumsentence{Han leker oftare än sin kompis}
@@
 Atm you can't focus this kind of object, then a ClSlash need info about the
-subject. FunRP also tricky, in this case RP needs to give more information.
+subject. FunRP also tricky (and what does it mean, we can say
+Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
 Kort sagt, the information about the subject must be spread
@@
 Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
-not 'alla sina', usw. Want it to handle conjuction 
-\enumsentence{Han såg sin hund och sitt hus}
+not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
 etc.
-Test to example \ref{ex:ReflAP}
+%%
 
@@
 Test how many parse trees that are generated.
+Look at tables for a VP tex.
 Some results.
@@
 \section{Mapping}
+\label{sec:Mapping}
 To connect GF with other annotations, to evaluate. Identify missing constructs,
@@
 they are anyway.
-Evaluation.
+Evaluation. -> Works for both flat and deep, but with sligthly better results for the flat one.
 

### 2011-11-22 doc/report/report.tex
report

@@
 
+%Functor
 %Main ideas, abstract and concrete, translation, parse- and abstract trees.
@@
 \subsection{Swedish}
-Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
-Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
-for what will be written in the grammar part.
-Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag är stor, I am big
-Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
+Swedish is a North-Germanic language spoken,
+closely related to Norweigan and Danish. The languages share most of their
+grammatical structures and are mutual intelligibility. Swedish is also one
+a official language in Finland and approximately 10 million people speak Swedish.
+Charasterictic for Swedish is the fact that it is a verb-second language: the
+second constituent of a declarative main clause must consist of a verb.
+The first constituent of the clause is usually made up of the subject,
+although it likewise could consist of adverbial phrases or objects that are
+being empathazied. \\
+
+\enumsentence{
+\begin{tabular}{lll}
+  Johan & gick & sakta på gatan.\\
+\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
+  Sakta & gick & Johan på gatan.\\
+\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
+  På gatan & gick & Johan sakta.\\
+\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
+\end{tabular}}
+
+Inverted word order marks questions
+\enumsentence{Gick Johan på gatan?\\
+        \emph{Did Johan walk on the street?}}
+The word order in subordinate clauses in also slightly different,
+\enumsentence{Jag såg Anna. Hon hälsade \textbf{inte}.\\
+        \emph{I saw Anna. She did not say hello.}\\
+              Jag såg att Anna \textbf{inte} hälsade.\\
+        \emph{I saw that Anna did not say hello.}}
+
+Some other features typical for Swedish is it's use of passive, especially the
+\textit{s-passive}, which is formed by adding an \emph{s} at the end of a verb
+\enumsentence{Han jagade\textbf{s} av ett lejon.\\
+        \emph{He was hunted by a lion.}}
+Jfr.
+\enumsentence{Ett lejon jagade honom.\\
+        \emph{A lion was hunting him.}}
+
+Constructions with 'det är/var' ('it is/was') is very common\cite{H&H309d}:
+\enumsentence{Det var roligt att höra.\\
+        \emph{I'm glad to hear that.}}
+'Det' \emph{it} is also used as formal subject for intransitive verbs,
+with the real subject put in the position of an object.
+\enumsentence{Det står en älg på fältet.\\
+        \emph{There is a mouse on the field.}}
+
+The Scandinavian language have reflexive pronouns\cite{H&H310}
+and special reflexive forms of the possessive pronouns for the 3rd
+person\cite{H&H319}. Those are distinct from the normal 3rd person forms.
+\enumsentence{
+\begin{tabular}{ll}
+Han slog sig. & Han såg sitt barn.\\
+\vspace{3mm}
+\emph{He hit him self.} & \emph{He saw his (own) child.}\\
+Han slog honom. & Han såg hans barn.\\
+\emph{He hit him (another person).}&\emph{He saw his (another person's) child.}
+\end{tabular}}
+The 1st and 2nd persons use the normal personal pronoun in object form.
+
+%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
+%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
+%for what will be written in the grammar part.
+%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
+%är stor, I am big
+%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
 
@@
 Explain what's in the resources, constructions present in most languages. 
-
-\subsection{GF for Swedish}
-What was in the resources
-description of the existing grammar, the types, the possibilities
-VP, NP.
-valency built in (problems with ställa - på, under, vid).
-Determiners, quants, predets, The difference from how pronouns are regarded. 
-Variants, nonexist.
-coverage of standard Swedish. Been working on a grammar fragment based on Talbanken,
+Been working on a grammar fragment based on Talbanken,
 like cool people like Montague. May also be interesting for other purposes.
@@
 v2-lang, focusing parts of sentences.
+
+\subsection{GF for Swedish}
+The GF resource grammars contains a fundamental description of Swedish,
+covering the morphology, many phrases and clauses.
+Covers sentences like
+han har inte ätit två gula äpplen idag
+word order, agreement, tempus, adverbs, adjectivs...
+and relatives
+han såg katten han tyckte om
+Also info in the Extra module, where more language specific constructions
+can be found.
+honom ville hon inte tänka på.
+
+Before describing what has been added, will give a short introductions to the
+the cats of resource grammar and the sw ipmlementation in particular.
+
+Most important now is NP and VP.
+Starting from N, get to CN what is it, what can we do now.
+To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
+but allow variantion in.. 
+What is the difference? Some words on differences for pronouns.
+Example, use vartiants {}.
+There are also Quants, which .... 
+Once we have a NP, can use Predets. What they are.
+
+For verbs, but also others there is an important distinction for valency.
+Either directly to VP if V, or needs to be given arguments. Can then add adverbs.
+Maybe talk about ställa.
+
+%Besides the categories NP,VP,AP there are also the import Det and Quant, Predet.
+
+%What was in the resources
+%description of the existing grammar, the types, the possibilities
+%VP, NP.
+%valency built in (problems with ställa - på, under, vid).
+%Determiners, quants, predets, The difference from how pronouns are regarded. 
+%Variants, nonexist.
+%coverage of standard Swedish.
 
@@
 
-Formal subjects \cite{SAG-19} is often used in Swedish
+Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish
 \enumsentence{Det sitter en fågel på taket}
@@
 \enumsentence{Det äts en fågel på taket.}
+%det finns/saknas/fattas
 There are also restrictions on the nounphrase of the subject, it can only
@@
 \enumsentence{Det sitter några fåglar på taket\\
-              **Det sitter samtliga fåglar på taket}
+              **Det sitter vardera fåglarna på taket}
 %%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
@@
 \label{ex:ReflAP}
+% hon ber sin katt gå - sin subject in clause.
 

### 2011-11-23 doc/report/report.tex
report

@@
 \section{Introduction} 
-What is parsing, flat strings to trees, what is deep parsing. What is controlled language,
-what is robust parsing and why we need it for free langugae, benefits of a grammar.
-GF is grammar formalism, many languages, many projects for controlled lang.
-GF grammar for Swedish, how big, not intended for parsing, but for building
-application grammars. Experiments of using the grammar for free parsing, methods
-being developped for this. Benefits of starting from GF. Free parsing is
-interesting. To do this, we need a better (bigger grammar), a large scale
-lexicon, and an evaluation method. This project has explored those things. The
-grammar also have other usages, important in itself, so we wanted to make it as
-covering as possible without being overallowing or too ambiguous.
+Computers want to 'understand' language. Not good with flat strings.
+We need to translate strings of natural language to a richer sturture,
+a parse tree. So far there is no
+freely availbe ruled-based parsing giving a deep analyse for Swedish.
+%There are parsers which put only put part-of-speach tags to
+%the text, and others that gives a deeper analyse. 
+There are statistical parser, and parser which only are deal with controlled language.
+The GF parser are one of those. 
+%What is parsing, flat strings to trees, example picture, what is deep parsing.
+So far stable and good when working with controlled languges, but recent experiments
+have showed that it is also possible to parse free language.
+To do this for Swedish, we need to  .. the resource grammar needed to covering
+the language specific constructions. Further a lagre lexicon is also needed, as well
+as a method for evaluation.
+During this project, we have worked on and explored those things and some problems identified.
+We have extended the Swedish resource grammar in GF, to give a better coverage and imported
+a large lexicon. Work to compare the trees from Talbanken to GF and to translate them.
+
+%What is controlled language,
+%what is robust parsing and why we need it for free langugae, benefits of a grammar.
+%GF is grammar formalism, many languages, many projects for controlled lang.
+%GF grammar for Swedish, how big, not intended for parsing, but for building
+%application grammars. Experiments of using the grammar for free parsing, methods
+%being developped for this. Benefits of starting from GF. Free parsing is
+%interesting. To do this, we need a better (bigger grammar), a large scale
+%lexicon, and an evaluation method. This project has explored those things. The
+%grammar also have other usages, important in itself, so we wanted to make it as
+%covering as possible without being overallowing or too ambiguous.
 
 Other work exist, but none is freely available. We want to create a open source
-product, therefor can only use resources that allows this. Work can continue,
+product, therefore can only use resources that allows this. Work can continue,
 will not disappear, be made use of in other projects.
@@
 Could be combined with 'sig', a ReflNP without another noun.
+-- man ser sig!!
 Make a distinction between objects and subjects?
@@
 
+% hard examples like
+% ger sig till henne -> ReflSlash ge henne in gf
+% questions with objects that should be moved
+
 %One part about the mapping from Talbanken to GF, description. About the purpose.

### 2011-11-23 mapping/Idents.hs
update to mapping

@@ -41,7 +41,11 @@ cidPNeg        = mkCId "PNeg"
 cidComplSlash  = mkCId "ComplSlash"
 cidSlashV2a    = mkCId "SlashV2a"
 cidSlashV2A    = mkCId "SlashV2A"
+cidSlash2V3    = mkCId "Slash2V3"
+cidSlash3V3    = mkCId "Slash3V3"
+cidSlashV2V    = mkCId "SlashV2V"
 cidComplVS     = mkCId "ComplVS"
+cidComplVQ     = mkCId "ComplVQ"
 cidUseV        = mkCId "UseV"
 cidAdVVP       = mkCId "AdVVP"
 cidAdvVP       = mkCId "AdvVP"

### 2011-11-23 mapping/Translate.hs

@@ -30,9 +30,16 @@ data SentenceType = Q | D | F
 data VPForm  = Cop | Sup | VV | VA 
              | V | V2 | V2A | V2Pass 
              | Fut | FutKommer
-             | VS         
-                                            
+             | VS  
+             | V3  | VQ | V2V | V2S  --new
+
+
+
   deriving (Eq,Show)
+vForms    = [Cop,Sup,Fut,FutKommer,VV,VA,V2A,V2,V2Pass,VS,V,  V3,VQ,V2V,V2S]
+gfvForms  = ["VV","VA","V2A","V2","VS","V","V3","VQ","V2V","V2S"] -- need more?
+
+
 data S = S { _isReflGenVP  :: Bool
            , _isExist      :: Bool
            , _iquant       :: Bool
@@ -45,7 +52,7 @@ data S = S { _isReflGenVP  :: Bool
 type PMonad = (RWS () [String] S)
 
 
-test = True
+test = False
 usePGF = testGr
 testGr = ("../gf/BigTest.pgf","BigTestSwe")
 bigGr  = ("../gf/Big.pgf","BigSwe")
@@ -559,8 +566,6 @@ isVTenseForm a (VTense t) = t == a
 isVTenseForm _ _          = False
 
 
-vForms     = [Cop,Sup,Fut,FutKommer,VV,VA,V2A,V2,V2Pass,VS,V]
-gfvForms  = ["VV","VA","V2A","V2","VS","V","V3"] -- need more?
 
 pSlashVP V typ =
  do (t,v) <-inside typ $ pVerb "VV" "V"
@@ -612,6 +617,9 @@ pSlashVP FutKommer typ =
     return (mkExpr meta,t)
 pSlashVP VS typ = 
  do (t,v) <- inside typ $ pVerb "VV" "VS"
+    return (mkExpr v,t)
+pSlashVP v typ = 
+ do (t,v) <- inside typ $ pVerb "VV" (show typ)
     return (mkExpr v,t)
 
 mkTmp = mkTmp' cidASimul 
@@ -775,16 +783,55 @@ pComplVP FutKommer q vp t (pol,exps,_) = do
   let vp1  = maybe vp0 (\a -> mkApp cidAdvVPSlash [vp0,a]) adv 
   return (mkTmp t,cidASimul,pol,vp1)
  
-pComplVP VS q vp t (pol,exps,_) = do
-  comp <- getComplement VS q exps
[…]

### 2011-11-24 doc/report/report.tex
report

@@
 \usepackage{lingmacros}
+\usepackage{subfig}
 
@@
 \section{Introduction} 
-Computers want to 'understand' language. Not good with flat strings.
+Ultimate goal: want computers to 'understand' language. Not good with flat strings.
 We need to translate strings of natural language to a richer sturture,
@@
 and was därmed reutgiven in a more modern version\cite{nivre}. This work
-was done by Nivre and ... in 2005. They released it in some different XML-formats,
+was done by Nivre and ... in 2005. They released it in Malt and Tiger XML-formats,
 where the trees had been made deeper and smt more. \\
@@
 \label{sec:Mapping}
-To connect GF with other annotations, to evaluate. Identify missing constructs,
-missing words. More reasons from presentation.
-Similar for English and Penntree, but PennTree and Talbanken different, plus swedish
-makes a different. More possible word orders (or at least more commonly used).
-One deep and one flatter version of Talbanken, deep contains ca 5 more tags.
-The flat was used since the deep one
-isn't deep in the same way as GF. Example with CNP (s5). Some
-more info about VPs are given, but do not think this will give any relevant information.
-Talbanken also contains frequent .. tags XP, XX osv. We cannot expect to translate
-this to anything but metas.
-How we try to build up a 'parse tree', a S with a NP VP for example. 
-How verbs look for complements (or inverted order), how metas are added.
-Description of most important tags from Talbanken, their translation.
-Description of the ones most hard to translate (particles, flat sentences)
-Example of nice translation and one with many metas, and one that doesn't work.
-Developed by adding one mapping for each tag, and by examples.
-How we cannot cover all cases. The testing: regression testing again,
-very easy to destroy what was working before. 
-How we need valency information.
-If we don't take the verb valency into account, allow all verbs to have Gf category V,
-we get a slightly better result,
-from 62.5\% to 65.4\%.  (see EvalText). The complemnt of the verbs should tell which category
-they are anyway.
-Evaluation. -> Works for both flat and deep, but with sligthly better results for the flat one.
+Talbanken contains much valuable information about phrase structures and word 
+usage. One part of this project has focused on extracting this information
+and to translate it to GF notaiton. We have constructed a mapping, which 
+automatically transforms trees in Talbanken format to GF abstract syntax trees.
+Figure \ref{fig:translationtrees} shows an exapmle of a visualized tree
+from Talbanken05 and the result of translating it to GF.
+Both the POS tags as well as the syntactic information are needed for the translation.
+
+\begin{figure}[h]
+\centering
+\subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
+\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
+\caption{Trees for the sentence ``Katten på bilen blir större".}
+\label{fig:translationtrees}
+\end{figure}
+
+The mapping connects GF with another annotations and gives us means to evaluate
+our parser. By both parsing a talbanken sentence and transforming its annotaded tree, we can
+easily inspect if the results are equal.
+The mapping also makes it clear which grammatical constructions that are still missing
+from the GF grammar and shows how the GF analysis differs from the one made
+in Talbanken.
+If there are words missing from our dictionary, the rich
+pos-tags may help us to automatically find the correct declination and add it to the
+lexicon. Further, our parser will need probabilities, see section
+\ref{sec:future}, of how often a function is used.\\
+%missing words. Lexical aquisation, probabilities, see section .
+%The mapping gives information about which form a word is
+%currently used in and this may be used by the lexical extraction
+%tools. % Those can be enhanced if they are given more data. 
+%The translated trees enable us to extract those probabilities. %The mapping gives us a source for this.
+%Furthermore, the translation makes it easy to identify grammatical constructions
+%missing from the GF grammar Another important use of the mapping is evaluation of the parser, which can be
+%accomplished by comparing the parse trees and the trees from the transformer.
+
+A program for translating trees to GF syntax has been done before (Angelov,).
+It translated trees from the English Penn Treebank. By starting this progam, we now have
+a translation that works for the quite differently annotated Talbanken. The differences between
+Swedish and English syntax also meant that many rules had to be changed.\\
+
+The mapping works for Talbanken05 in Tiger XML-format. There are three kinds of
+tags: categories, edge labels and pos tags. 
+The categories give syntactic information such as \verb|S|,\verb|NP| and \verb|VP|.
+The edge labels give similair information with tags as \verb|SS| for subject, 
+\verb|OO| for object. There are more than 300 different pos tags. The high
+number is due to the detailedness, for example there are almost 50 tags for
+nouns, excluding proper names. The tags show definitiness, case, whether the
+word is a compound. Some words also have their own tags.\\
+
+\begin{tabular}{lll}
+\textbf{Tag} & \textbf{Example} & \textbf{Explanation} \\
+\hline
+\textbf{SV} & ska & The verb \emph{shall}\\
+\textbf{WVIV} & vilja & The verb \emph{want} in infinitive\\
+\textbf{KVSP} & kommit att & The verb \emph{will} in supine\\
+\textbf{GVPT  PA} & gjordes & The verb \emph{do, make}, present, passive \\
+\textbf{NNDDHSGG} & familjemedlemmarnas & Noun, definite, compound (person), genitive \\
+\textbf{PU} & *, 1), a) & List item  ...\\
+\end{tabular}\\
+
+%Pos continues
+%to either new tree or a word. Example S -> PP -> (Hd -> Id , Hd -> Id). Show
+%corresponding GF tree.
+
+The translation works sectionwise, ie, if it finds a tree of category \verb|S|,
+it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
+can find those, they can be combine with the GF function \verb|PredVP|.
+At the lowest level in the tree we look up the word in the GF lexicon.
+It figure \ref{pic:tbtree}, the verb is followed by an adjective as a present
+participle, \verb|SP|. GF admits \emph{bli} to have type \verb|VA|, a verb
+with an adjective as complement, and the translator can apply the gf function
+\verb|ComplVA| to combine the verb with its object. \\
+It is not often the case that we can first find the subject, then the verb and 
+finally the objects. Often the order is Verb-Subject-Object or Object-Subject-Verb. \\
+During the mapping process, the translator also needs to find out whether the sentence is
+negated and what tempus is used. If it finds two trees of type, say, \verb|VP| but does not
+have any rules for combining them, the \textit{meta function} \emph{'?'} will be used. 
+Figure xx shows an exapmle of this.\\
+
+%How we try to build up a 'abstract syntax tree', a S with a NP VP for example. 
+%Continue with dividing VP into Verb Objs. More possible word orders (or at
+%least more commonly used). 'katten åt sin mat fort' 'åt katten sin mat fort' 'sin mat åt katten fort'
+% If can't find, add a '?', a meta.
+%This is also what happens when the translator finds a combination it doesn't know about.
+%Example of tree with meta.
+
+%Description of most important tags from Talbanken, their translation? appendix?
+The structural differences between GF and Talbanken cause some translation problems.
+On example is that Talbanken has no valencey information but for GF this is the key when
+constructing for example verb phrases. The grammar bestämmer that a transitive \verb|V2|
+needs by two objects. The distinction is not always clear in Swedish where 
+most transitive verbs can also be used as intransitive. Both sentences in
+example \ref{ex:mappTransitive}.
+\enumsentence{Han sitter och läser.\\ Han sitter och läser en bok.} \label{ex:mappTransitive}
+See section \ref{sec:futureValency} for a longer discussion about this.
+For this reason, the mapping will allow words to be used in other ways than their
+GF valencies prescribes.
+
+Another problem can be instantiated by sentence \ref{ex:mappMan}, containing the generic pronoun \emph{man}.
+\enumsentence{Man uppskattar dem.\\
+        \emph{They are appreciated}}\label{ex:mappMan} %  (s216).
+In talbanken, \emph{man} is simply a personal pronoun, 
+in GF it causes the whole clause to get the form \verb|GenericCl|, which has no
+subject visible in the abstract tree. See figure \ref{fig:mappMan}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=80mm]{man.pdf}
+\caption{Abstract tree for ``Man uppskattar dem".}
+\label{fig:mappMan}
+\end{figure}
+In this case, it is not possible to simply glue the subject to the verb with \verb|PredVP|.
+%Therefore, the translator must change the type of the 
+%sentence when inside the SS.
+
+%förklara bättre eller ta bort??
+A similair issue arises form the use of verb particles. If a verb with a particle is considered
+as one klump in GF, which means that the particle has no effect on the abstract tree. Its presence
+is announced by the verb itself. The lexicon may contain a verb both with and without a particle,
+for example \emph{tänka} (think) and \emph{tänka till} (consider?).
+\begin{verbatim}
+taenka_V  : V ;
+taenka_till_V : V ;
+\end{verbatim}
+When the translator looks up a the form \emph{tänka}, both of the two verbs
+will be among the results. The first option will be selected, and if the
+particle isn't directly attatched to the verb
+%There is, at this point, no easy way of knowing which one of them wants a particle and the
+the transformation might confuse the two words. This is also the case when translating
+all sorts of idioms that GF considers to be just one bit.\\
+
+%Difficulties in looking up 'idioms' like this
+%in the gf lexicon. Therefore the translator might get confused.
+%Sentences are also differntly deep. Example of a conjuncted sentence.
+
+On of the reasons why \textit{metas} are put in the output, is the 
+use of tags like \\
+\begin{figure}[h]
+\begin{tabular}{ll}
+\textbf{NAC} & Not a constituent\\
+\textbf{XP} & Other (non-coordinated) phrase\\
+\textbf{DB} & Doubled function\\
+\textbf{PU} & List item\\
+\textbf{XX} & Unclassifiable part-of-speech\\
+\end{tabular}\label{fig:mapBadtag}
+\end{figure}
+
+They occure quite frequently in the treebank and are always translated
+to \emph{?}.
+%How we cannot cover all cases.
+
+\subsection{Results and evalutaion}
+The flat version of Talbanken05 was used for developping the mapping, but
+the program works for both versions.
+The deeper contains six more tags, but they are not always useful for our       
+purpose. Figuer \ref{fig:mappDeepFlat} shows the difference for a conjunction.
+% show piece of s5 here!
+%\begin{figure}[h]
+%\centering
+%\subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
+%\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
+%\caption{Ttrees for the sentence ``Katten på bilen blir större".}
+%\label{fig:translationtrees}
+%\end{figure}
+
+Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
+which can group an infinite verb and an object abverbial together into a \verb|VP|.
+This information can however be extracted from the flat version, and the results
+get slightly better when using the flat version. \\
+
+%At least one rule  and also most of the
+The development was mostly example-driven, and at least one rule for each
+Talbanken-tag has been implemented.
+Shorter sentences have been prioritized in order to get a coverage of the most
+fudamental constructions. Further,  
+regression testing was used during the development. % in order to  very easy to destroy what was working before.
+
+When evaluating the mapping, the results depend a lot on the restrictions on the
+input. 
+The focus has been on shorter sentences, with no idioms or conjunction. We are
+not interested in sentences made up by list, since they are meant for making
+a nice graphic look in folders etc.\\
+
+If we
+assure that the lexicon know the correct word class for all lemmas involved, we
+can restore more than 85\% of the nodes in the original tree.
+If we lift all the restrictions excluding the \verb|PU|, we get
+%When tested the 100 first sentences from Talbanken, we get
+65\% coverage.
+If we test randomly collected sentences that does not contain and any of the tags listed
+in figure \ref{fig:mapBadtag} 72\% can be restored. \\
+
+\begin{tabular}{ll}
+No list items & 65\%\\
+No special tags or punctiation & 72\%\\
+Short sentences with known words & 85\%\\
+\end{tabular}
+
+%If we don't take the verb valency into account, allow all verbs to have Gf category V,
+%we get a slightly better result,
+%from 62.5\% to 65.4\%.  (see EvalText). The complemnt of the verbs should tell which category
+%they are anyway.
+%Evaluation. -> Works for both flat and deep, but with sligthly better results for the flat one.
 
@@
 % questions with objects that should be moved
-
-%One part about the mapping from Talbanken to GF, description. About the purpose.
+% vilken katt såg du
 %This part explains
@@
 \section{Future Work}
+\label{sec:future}
 What is still left to do? Description and direction of future work:
+Handling unknown words and , handling unknown grammar, 
 how to make the parser robust, elipses and long sentences, names and numbers.
 What is still needed in the grammar, how important that is. 
-how to get valency information 'sitter och läser', description of extracting this
-or handling it by the robust parser or the grammar itself.
+\subsection{Valency}
+\label{sec:futureValency}
+how to get valency information 'sitter och läser',
+is this doable for swedish?
+description of extracting this from talbanken/korp
+or handling it by the robust parser or the grammar itself (leaveout obj?).
 make the parser faster by using smaller lexicon.
 Lexicon tool on the fly.
+\subsection{Probabilites}
+\label{sec:futureProbabilities}
 Getting and using probabilities, to disambiguate.

### 2011-11-28 doc/report/report.tex
backup of docs, not newer than showup-branch

@@
 \usepackage{lingmacros}
+\usepackage{verbatim}
 \usepackage{subfig}
+% exempel: 2) Inte vet jag
+%              not know I
+%             I don't know
 
@@
 
+\subsection{Swedish}
+Swedish is a North-Germanic language spoken,
+closely related to Norweigan and Danish. The languages share most of their
+grammatical structures and are mutual intelligibility. Swedish is also 
+an official language in Finland and spoken by approximately 10 million people.
+A list of things:
+\subsubsection*{Verb second}
+Swedish is a verb-second language: the
+second constituent of a declarative main clause must consist of a verb.
+The normal word order is Subject-Verb-Object, but any syntactic category can be fronted.
+The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
+meaning, although the fronted part of sentence \ref{ex:swedish-ovs} are being emphasized.
+\enumsentence{
+\shortex{4}
+{Jag & ser & dig & inte.}
+{I &  see & you & not.}
+{`I  don't see you'.}} \label{ex:swedish-svo}
+\enumsentence{
+\shortexnt{4}
+{Dig  & ser &  jag & inte.}
+{You& I& see & not.}}\label{ex:swedish-ovs} 
+\enumsentence{
+\shortexnt{4}
+{Inte & ser& jag &dig.}
+{Not&  see& I&   you.}} \label{ex:swedish-avso}
+
+%\shortexm{3}
+%{hej} {hi}{hhe} {bing}{bong}{salt}
+%; adverbials, objects
+%Alla syntaktiska kategorier kan be fronted utom obetonade satsadverbial (ju,väl).
+%The first constituent of the clause is usually made up of the subject,
+%although it likewise could consist of adverbial phrases or objects that are
+%being empathazied. \\
+%\enumsentence{
+%\begin{tabular}{lll}
+%  Johan & gick & sakta på gatan.\\
+%\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
+%  Sakta & gick & Johan på gatan.\\
+%\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
+%  På gatan & gick & Johan sakta.\\
+%\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
+%\end{tabular}}
+
+Inverted word order marks questions
+\enumsentence{
+\shortex{4}
+{Såg & jag & dig & inte?}
+{Saw  & I&  you& not?}
+{`Didn't I see you?'}}
+%\enumsentence{Gick Johan på gatan?\\
+%        \emph{Did Johan walk on the street?}}
+The word order in subordinate clauses in also slightly different,
+%\enumsentence{
+%\begin{tabular}{ll}
+%{\shortex{7}
+%{Jag & såg&  Anna.&  Hon & såg & inte&  mig.} 
+%{ & & &  & & &}}
+%\shortex{7}
+%{Jag såg att Anna \textbf{inte} såg mig.}\\
+%\shortex{7}
+%{I  saw Anna. She saw not   me.}  & {I   saw that Anna not saw me}. \\
+%`I saw Anna. $\;$ She did not see me.'& `I saw that Anna did not see me'.
+%}
+\enumsentence{Main sentences:\\
+{\shortex{7}{Jag & såg & Anna. & Hon & såg & {\bf inte} & mig }
+{ I & saw & Anna. & She & saw & {\bf not} & me}
+{`I saw Anna. She didn't see me'}}}
+\enumsentence{Subordinate sentence:\\
+\shortex{7}{Jag & såg & att& Anna& \textbf{inte}& såg& mig.}
+            {I &  saw & that& Anna& \textbf{not}& saw& me.}
+            {`I saw that Anna didn't see me'}}
+
+\subsubsection*{Passive voice}
+Passive voice is often used in Swedish, especially the
+\textit{s-passive}, which is formed by adding an \emph{s} at the end of a verb.
+\enumsentence{
+\shortex{5}{Han & jagade\textbf{s} & av & ett & lejon.}
+        {He & hunted+\textbf{s} & by & a & lion.}
+        {`He was hunted by a lion.'}}
+cf.
+\enumsentence{
+\shortex{5}{Ett & lejon & jagade & honom.}
+        {A &lion &hunted &him.}
+        {`A lion was hunting him.'}}
+
+\subsubsection*{Impersonal constructions}
+Constructions with 'det är/var' ('it is/was') is very common\cite{H&H309d}:
+\enumsentence{
+\shortex{6}{Det & var & roligt & att & höra.}
+{It &was & nice &to &hear.}
+{`I'm glad to hear that.'}}
+'Det' (\emph{it}) is also used as formal subject %for intransitive verbs,
+with the real subject put in the position of an object.
+\enumsentence{
+\shortex{6}{Det &står &en &älg &på &fältet.}
+{ It & stands &a & moose &on &the field.}
+{`There is a moose on the field.'}}
+
+\subsubsection*{Reflexive pronouns}
+The Scandinavian language have reflexive pronouns\cite{H&H310}
+and special reflexive forms of the possessive pronouns for the 3rd
+person\cite{H&H319}. Those are distinct from the normal 3rd person forms.
+\enumsentence{
+\begin{tabular}{ll}
+Han slog sig. & Han såg sitt barn.\\
+\vspace{3mm}
+\emph{He hit him self.} & \emph{He saw his (own) child.}\\
+Han slog honom. & Han såg hans barn.\\
+\emph{He hit him (another person).}&\emph{He saw his (another person's) child.}
+\end{tabular}}
+The 1st and 2nd persons use the normal personal pronoun in object form.
+
+%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
+%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
+%for what will be written in the grammar part.
+%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
+%är stor, I am big
+%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
+
+\subsection{Saldo}
+A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
+large electronic lexicon developed and maintained at Gothenburg University. It is
+built upon Svenskt Associations Lexicon and contains information about modern
+Swedish and the words are equipped with both sematic and a morphological
+information. The user can find compounding analyses, examples of usage in corpora,
+and graphs of sematically connected words.\\
+Example of katt? \\
+For our purpose however, only the morphological data was needed.\\
+Example of morpho for katt.\\
+Saldo provides this in XML format under LPGL. The data can be processed and
+translated to GF format as described in 
+section \ref{sec:prog.saldo}.
+%Other usages?
+
 \subsection{GF}
@@
 
-GF is developed with multilinguality in mind and makes an important distiction between
+Being developed with multilinguality in mind, GF makes an important distiction between
 \textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
@@
 \begin{verbatim}
- abstract TestGrammar {
+ abstract TestGrammar = {
   cat N ; V ; S ;
@@
     Pred : N -> V -> S ;
+    cat_N : N ;
+    sleep_V : V ;
  }
@@
 
-The example shows an abstract grammar defining three \textit{Categories},
+The example shows an abstract grammar defining three categories, % \verb|Categories|,
 one for nouns, one for verbs and one for sentences. An abstract grammar also gives
 the types of functions. In this case we have \verb|Pred| which
-tells us that by taking a noun and a verb we can form a sentence. It does not
-give any information of how this is done, or how any of the categories should look.
-maybe continue example with swedish and english, have DetNP : N -> N wich
-gives 'the cat', 'katten'. Thereby introduce tables, parameters, fields.
-Show abstract tree.
+tells us that by taking a noun and a verb we can form a sentence. At this stage, 
+no information of how this is done is given. The grammars also defines two words,
+\verb|cat| and \verb|sleep|.
+
+% or how any of the categories should look.
+\begin{figure}[h]
+\begin{verbatim}
+concrete TestGrammarSwe of TestGrammar = {
+  lincat N, V, S = Str ;
+
+  lin Pred n v   = n ++ v ;
+      cat_N   = "katten"  ;
+      sleep_V = "sover"  ;
+}
+\end{verbatim}
+\caption{A Swedish concrete grammar}
+\label{fig:gfSweCnc1}
+\end{figure}
+
+Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
+for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|
+and \verb|Pred| simply glues the two strings ``katten" and ``sover" together:
+\verb|Pred cat sleep = "katten sover"|.\\
+Let's look at a more complicated example, where nouns can be used in both
+plural and singular. To do this, we add a category \verb|NP| to the
+abstract. \verb|N| now means a noun
+in any of the number, and we introduce two funciton, \verb|NSg : N -> NP| and
+\verb|NPl : N -> Pl| which sets the number.
+\begin{figure}[h!]
+\begin{verbatim}              
+ abstract TestGrammar = {          concrete TestGrammarSwe of TestGrammar = {
+  cat N ; NP ; V ; S ;               lincat V, S, NP = Str ;
+                                            N = {s : Num => Str} ;
+  fun                                lin   
+    Pred : NP -> V -> S ;              Pred n v = n ++ v ;
+    NSg : N -> NP ;                    NPl n = n.s ! Pl ;
+    NPl : N -> NP ;                    NSg n = n.s ! Sg ;
+    cat_N : N ;                        cat_N = {s = table {Sg => "katten" ;
+    sleep_V : V ;                                          Pl => "katterna"}};
+ }                                     sleep_V = "sover"  ;
+                                     param Num = Sg | Pl ;
+                                     }
+\end{verbatim}           
+\caption{A modified grammar}
+\label{fig:gfTest2}
+\end{figure}
+Figure \ref{fig:gfTest2} introduces some new concepts: records,tables and parameters.
+In the concrete syntax, \verb|N| is defined to be a record consisting of the field
+\verb|s|. The type of \verb|s| shows that it is a table, which given a parameter of type \verb|Num| returns
+a string. \verb|Num| is defined as a parameter, which can either have value \verb|Sg|
+or \verb|Pl|. In \verb|NPl| and \verb|NSg|, \verb|n.s| means that we use the
+field \verb|s| of \verb|n|, and then the selection operator \verb|!| is used to
+select a branch in the table.
+
+If we want to implement a English version of the grammar, we encounter another
+problem: the verb form depends on the noun phrase. The \verb|NP| needs to carry
+information about which number it is in, and \verb|Pred| can pass this on to the
+verb. Hence, the \verb|V| also needs to be a table, depending on the number.
+\begin{figure}[h!]
+\begin{verbatim}              
+concrete TestGrammarEng of TestGrammar = {
+  lincat S = Str ;
+         V = {s : Num => Str} ;
+         N = {s : Num => Str} ;
+         NP = {s : Str ; num = Num} ;
+  lin   
+    Pred n v = n.s ++ v.s ! n.num ;
+    NPl n = {n.s ! Pl ; num = Pl} ;
+    NSg n = {n.s ! Sg ; num = Sp} ;
+    cat_N = {s = table {Sg => "the cat" ;
+                        Pl => "the cats"}};
+    sleep_V = {s = table {Sg => "sleeps" ; 
+                          Pl => "sleep"}};
+  param Num = Sg | Pl ;
+  }
+\end{verbatim}           
+\caption{English implementation}
+\label{fig:gfTestEng}
+\end{figure}
+
+%maybe continue example with swedish and english, have DetNP : N -> N wich
+%gives 'the cat', 'katten'. Thereby introduce tables, parameters, fields.
+%Show abstract tree.
 
@@
 a string of natural language given a abstract tree. This step is called linearization.
-Translation is a consequence of this. Picture of sw. string -> tree -> eng. tree.
+Translation is a consequence of this, we can parse a Swedish string and then 
+linearize the resulting abstract tree to English.
 
@@
 
-A GF grammar usually covers a very restricted set of language, where few ambiguities
-occur and much care can be taken to preserving the semantics between the different languages.
+A GF grammar usually covers a very restricted set of language where few ambiguities
+occur and much care can be taken to preserving the semantics of the
+different languages.
 By using a controlled language like this,  good and reliable translation can be
@@
 
-One of the best... features of the framework is the resource libraries. Provided in the 
-GF package, those gives a good start for all who want to write an application grammar.
-Instead of letting the users reimplementing the morphological and the fundamental
-syntactic structures of a language for each application, all this can be found
-as libraries. So far, 20 languages are implemented all sharing the same abstract syntax.
-They describe how to conjugate words, and how to construct phrases and sentences.
+\subsubsection{The resource library}
+The resource library\cite{gf-resource} is a very useful part of the framework. It is provided in the 
+GF package. %, they give a good start for all who want to write an application grammar.
+Instead of letting the users of the framework reimplement the morphological and
+the fundamental syntactic structures of a language for each application, all
+this can be found as libraries. So far there are 20 languages in the library, and
+they all share the same abstract syntax.
+The resource grammars describe how to decline words and how to construct
+phrases and sentences.
 There are also a small test lexicon included, containing some hundered of common words.\\
@@
 
-\subsection{Swedish}
-Swedish is a North-Germanic language spoken,
-closely related to Norweigan and Danish. The languages share most of their
-grammatical structures and are mutual intelligibility. Swedish is also one
-a official language in Finland and approximately 10 million people speak Swedish.
-Charasterictic for Swedish is the fact that it is a verb-second language: the
-second constituent of a declarative main clause must consist of a verb.
-The first constituent of the clause is usually made up of the subject,
-although it likewise could consist of adverbial phrases or objects that are
-being empathazied. \\
-
-\enumsentence{
-\begin{tabular}{lll}
-  Johan & gick & sakta på gatan.\\
-\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
-  Sakta & gick & Johan på gatan.\\
-\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
-  På gatan & gick & Johan sakta.\\
-\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
-\end{tabular}}
-
-Inverted word order marks questions
-\enumsentence{Gick Johan på gatan?\\
-        \emph{Did Johan walk on the street?}}
-The word order in subordinate clauses in also slightly different,
-\enumsentence{Jag såg Anna. Hon hälsade \textbf{inte}.\\
-        \emph{I saw Anna. She did not say hello.}\\
-              Jag såg att Anna \textbf{inte} hälsade.\\
-        \emph{I saw that Anna did not say hello.}}
-
-Some other features typical for Swedish is it's use of passive, especially the
-\textit{s-passive}, which is formed by adding an \emph{s} at the end of a verb
-\enumsentence{Han jagade\textbf{s} av ett lejon.\\
-        \emph{He was hunted by a lion.}}
-Jfr.
-\enumsentence{Ett lejon jagade honom.\\
-        \emph{A lion was hunting him.}}
-
-Constructions with 'det är/var' ('it is/was') is very common\cite{H&H309d}:
-\enumsentence{Det var roligt att höra.\\
-        \emph{I'm glad to hear that.}}
-'Det' \emph{it} is also used as formal subject for intransitive verbs,
-with the real subject put in the position of an object.
-\enumsentence{Det står en älg på fältet.\\
-        \emph{There is a mouse on the field.}}
-
-The Scandinavian language have reflexive pronouns\cite{H&H310}
-and special reflexive forms of the possessive pronouns for the 3rd
-person\cite{H&H319}. Those are distinct from the normal 3rd person forms.
-\enumsentence{
-\begin{tabular}{ll}
-Han slog sig. & Han såg sitt barn.\\
-\vspace{3mm}
-\emph{He hit him self.} & \emph{He saw his (own) child.}\\
-Han slog honom. & Han såg hans barn.\\
-\emph{He hit him (another person).}&\emph{He saw his (another person's) child.}
-\end{tabular}}
-The 1st and 2nd persons use the normal personal pronoun in object form.
-
-%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
-%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
-%for what will be written in the grammar part.
-%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
-%är stor, I am big
-%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
+%inte vet jag. foc + v + resten
 
@@
 %Look at output from Maltparser, could also comparing our parse trees to this.
-
-\subsection{Saldo}
-A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
-large electronic lexicon developed and maintained at Gothenburg University. It is
-built upon Svenskt Associations Lexicon and contains information about modern
-Swedish and the words are equipped with both sematic and a morphological
-information. The user can find compounding analyses, examples of usage in corpora,
-and graphs of sematically connected words.\\
-Example of katt? \\
-For our purpose however, only the morphological data was needed.\\
-Example of morpho for katt.\\
-Saldo provides this in XML format under LPGL. The data can be processed and translated to GF format as described in 
-section \ref{sec:prog.saldo}.
-Other usages?
 
@@
 and relatives
-han såg katten han tyckte om
 Also info in the Extra module, where more language specific constructions
 can be found.
+jag ser dig
+dig ser jag
+dig har jag inte sett
+jag har inte sett dig.
+how the parse tree is generated, the finite verb, the subject, the rest. usw.
+han såg katten han tyckte om
+för vem måste jag akta mig
+vem måste jag akta mig för
+
 honom ville hon inte tänka på.
@@
 Maybe talk about ställa.
+Particles and prepositions, can't be seen on the trees.
+tycka om /tycka om -> om vad tycker du inte. fast bra
 
@@
 \enumsentence{The picture [took+s] by spacecraft Galileo.}
-This second variant, the s-passive, is commonly used in Swedish, some research
+This second variant, the s-passive, is commonly used in Swedish, some studies
 suggest that it is used in more than 80\% of the times \cite{laanemets}.
-The other scandinavian languages prefers the periphrastic passive, especially in spoken langauge.
+For the other scandinavian languages, the s-passive can not be used in all tenses for all words.
+The periphrastic passive is preferred, especially in spoken langauge.
 \enumsentence{Bildet ble tatt av romsonden (NO)}
@@
 'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
-needed.\\
+needed.
+du är här, vilket jag vet.
+i de äpplen/det äpple du äter bor en mask/ som du ser.\\
 
+\subsection{Presentering, Impersonal constantiation(?), Topicalisering}
 Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish
@@
 
+finnas: här får man ha definit: det finns de som inte vill. Samma RelCl som
+jag just la till, de äppplen. (de + relsats)
+
 'det är kul att..' Näst vanligaste två-ordskombinationen i talad svenska.
@@
 %embedded - relative, 'sådan att' not really nice for normal swedish, and no agreement
-%between subject 'en katt sådan att det regnar'.
+%between subject 'en katt sådan att det regnar'. Kan säga 'en sån katt som sover, med 'sådan' + Rel
+%har lagt till Det äpple + Rel (RelCNNP). *Det äpple är rött. men Det äpple jag äter är rött.
+%
+
 pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
@@
 etc.
+varandra, works the same but has no case for Sg.
+Johan vill att jag rakar sig* -> honom  
 %%
@@
 The deeper contains six more tags, but they are not always useful for our       
-purpose. Figuer \ref{fig:mappDeepFlat} shows the difference for a conjunction.
+purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
 % show piece of s5 here!
-%\begin{figure}[h]
-%\centering
-%\subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
-%\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
-%\caption{Ttrees for the sentence ``Katten på bilen blir större".}
-%\label{fig:translationtrees}
-%\end{figure}
+\begin{figure}[h]
+\centering
+\subfloat[Deep]{\label{pic:mappDeep}\includegraphics[width=20mm]{nofile.png}}
+\subfloat[Flat]{\label{pic:mappFlat}\includegraphics[width=20mm]{nofile.png}}
+\caption{...}
+\label{fig:mappDeepFlat}
+\end{figure}
 
@@
 results. Mostly compound nouns, which we can't expect in the dictionary.
+Is too slow, can't be used for parsing right away
 
@@
 What is still needed in the grammar, how important that is. 
+\subsection{Lexicon/Parallell parsing}
+Speed up.
 \subsection{Valency}

### 2011-11-28 doc/report/report.tex
report

@@
 \subsection{Swedish}
-Swedish is a North-Germanic language spoken,
+Swedish\cite{SAG:intro} is a North-Germanic language,
 closely related to Norweigan and Danish. The languages share most of their
@@
 an official language in Finland and spoken by approximately 10 million people.
-A list of things:
+Swedish syntax are often similair to English, but have some characteristics:
 \subsubsection*{Verb second}
@@
 second constituent of a declarative main clause must consist of a verb.
-The normal word order is Subject-Verb-Object, but any syntactic category can be fronted.
+The normal word order is subject-verb-object, but any syntactic category can be fronted\cite{H&H:1027}.
+Topicalisation is very common, especially for temporal and spatial adverbial phrases.
 The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
-meaning, although the fronted part of sentence \ref{ex:swedish-ovs} are being emphasized.
+meaning, although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being emphasized.
 \enumsentence{
 \shortex{4}
-{Jag & ser & dig & inte.}
-{I &  see & you & not.}
-{`I  don't see you'.}} \label{ex:swedish-svo}
+{Du & ser & mig & inte.}
+{You & see & me & not.}
+{`You  don't see me'.}} \label{ex:swedish-svo}
 \enumsentence{
 \shortexnt{4}
-{Dig  & ser &  jag & inte.}
-{You& I& see & not.}}\label{ex:swedish-ovs} 
+{Mig  & ser &  du & inte.}
+{Me & see & you & not.}}\label{ex:swedish-ovs} 
 \enumsentence{
 \shortexnt{4}
-{Inte & ser& jag &dig.}
-{Not&  see& I&   you.}} \label{ex:swedish-avso}
+{Inte & ser& du &mig.}
+{Not&  see& you& me.}} \label{ex:swedish-avso}
 
@@
 \shortex{4}
-{Såg & jag & dig & inte?}
-{Saw  & I&  you& not?}
-{`Didn't I see you?'}}
+{Såg & du & mig & inte?}
+{Saw  & you&  me& not?}
+{`Didn't you see me?'}}
 %\enumsentence{Gick Johan på gatan?\\
@@
 \textit{s-passive}, which is formed by adding an \emph{s} at the end of a verb.
-\enumsentence{
-\shortex{5}{Han & jagade\textbf{s} & av & ett & lejon.}
+\enumsentence{Passive \hspace{39mm} Active\\
+\shortexm{11}{Han & jagade\textbf{s} & av & ett & lejon.}
         {He & hunted+\textbf{s} & by & a & lion.}
-        {`He was hunted by a lion.'}}
-cf.
-\enumsentence{
-\shortex{5}{Ett & lejon & jagade & honom.}
+        {`He was hunted by a lion.'}
+{Ett & lejon & jagade & honom.}
         {A &lion &hunted &him.}
         {`A lion was hunting him.'}}
+
+%cf.
+%\enumsentence{
+%\shortex{5}{Ett & lejon & jagade & honom.}
+%        {A &lion &hunted &him.}
+%        {`A lion was hunting him.'}}
 
@@
 \begin{tabular}{ll}
-Han slog sig. & Han såg sitt barn.\\
+Han slog \textbf{sig}. & Han såg \textbf{sitt} barn.\\
 \vspace{3mm}
 \emph{He hit him self.} & \emph{He saw his (own) child.}\\
-Han slog honom. & Han såg hans barn.\\
+Han slog \textbf{honom}. & Han såg \textbf{hans} barn.\\
 \emph{He hit him (another person).}&\emph{He saw his (another person's) child.}
@@
 The GF resource grammars contains a fundamental description of Swedish,
-covering the morphology, many phrases and clauses.
-Covers sentences like
-han har inte ätit två gula äpplen idag
-word order, agreement, tempus, adverbs, adjectivs...
-and relatives
-Also info in the Extra module, where more language specific constructions
-can be found.
+covering the morphology, many phrases and clauses. Most of the implementation
+is shared with the other scandinavian languages.
+Apart form descriping morphology, the resource grammar deals with
+word order, agreement, tempus, basic conjunction etc.
+Sentence \ref{gfSwe:parseable} produces the parse in figure \ref{gfSwe:parsetree}.
+%Covers sentences like
+% better sentence where the verb phrase is split up?
+\enumsentence{\shortexnt{8}
+{Har& han& inte& ätit& de & gula& äpplena& idag}
+{He& has& not& eaten& the& yellow& apples& today}} %correct!!!
+\label{gfSwe:parseable}
+\begin{figure}[h]
+\includegraphics[width=60mm]{gfTree.pdf}
+\caption{Parse tree for ``Han har inte ätit de gula äpplena idag".}%correct
+\label{gfSwe:parsetree}
+\end{figure}
+Even though the verb phrase \emph{``har inte ätit..} is discontinuous, the whole
+phrase is still treated as one constituent in GF. The parts
+are connected in the tree, and the subject \emph{``han"} is put between the
+finite verb and the rest of the phrase. In the code, this is implemented by using a record.
+The linearization rule for questions -- clauses with inverted order -- picks the 
+finite form: \\
+\verb|table { Inv => verb.fin ++ subj ++ neg ++ verb.inf ++ compl ; ... | \\
+Relative clauses are also covered: 
+\enumsentence{Hon ser pojken som sover\\
+              She sees the boy who sleeps}
+
+In addition to the basic resource grammar, additional constructions had been added
+to the module \verb|Extra|. The functions given here do not have to be translatable
+to all other language, but are meant to cover language specific constructions.
+The \verb|Extra| module included functions for topologication, although all
+of them were not implemented for Swedish.
+%word order, agreement, tempus, adverbs, adjectivs...
+%and relatives
+%Also info in the Extra module, where more language specific constructions
+%can be found.
+Check wich were here first:
 jag ser dig

### 2011-11-29 doc/report/report.tex
report

@@
 The GF resource grammars contains a fundamental description of Swedish,
-covering the morphology, many phrases and clauses. Most of the implementation
+covering the morphology and many phrase and clause constructions. Most of the implementation
 is shared with the other scandinavian languages.
@@
 word order, agreement, tempus, basic conjunction etc.
-Sentence \ref{gfSwe:parseable} produces the parse in figure \ref{gfSwe:parsetree}.
+\enumsentence{\shortexnt{8}
+{Har& han& inte& ätit& de & gula& äpplena& idag?}
+{Has& he& not& eaten& the& yellow& apples& today?}} %correct!!!
+\label{gfSwe:parseable}
+The grammar can construct and parse the sentence in example \ref{gfSwe:parseable}
+and the parse tree is represented as in figure \ref{gfSwe:parsetree}.
 %Covers sentences like
 % better sentence where the verb phrase is split up?
-\enumsentence{\shortexnt{8}
-{Har& han& inte& ätit& de & gula& äpplena& idag}
-{He& has& not& eaten& the& yellow& apples& today}} %correct!!!
-\label{gfSwe:parseable}
 \begin{figure}[h]
@@
 \end{figure}
-Even though the verb phrase \emph{``har inte ätit..} is discontinuous, the whole
+Even though the verb phrase \emph{``har inte ätit.."} is discontinuous, the whole
 phrase is still treated as one constituent in GF. The parts
 are connected in the tree, and the subject \emph{``han"} is put between the
-finite verb and the rest of the phrase. In the code, this is implemented by using a record.
-The linearization rule for questions -- clauses with inverted order -- picks the 
-finite form: \\
-\verb|table { Inv => verb.fin ++ subj ++ neg ++ verb.inf ++ compl ; ... | \\
+finite verb and the rest of the phrase. At the code level, this is implemented
+using records.
+The linearization rule for questions -- clauses with inverted order -- picks 
+the field for the finite verb and appends it to the subject followed by
+the negation, infinite part and the complement of the verb phrase: \\
+\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
+
 Relative clauses are also covered: 
-\enumsentence{Hon ser pojken som sover\\
-              She sees the boy who sleeps}
-
+\enumsentence{\shortexnt{5}{Hon & ser & pojken & som  &sover}
+              {She & sees & the boy & who & sleeps}}
+              %{`She sees the sleeping boy'}}
+
+\enumsentence{\shortexnt{6}{Han & ser &katten & han  &tycker & om}
+              {He & sees & the cat & he & likes &}}
+              %{`She sees the sleeping boy'}}
 In addition to the basic resource grammar, additional constructions had been added
@@
 to all other language, but are meant to cover language specific constructions.
-The \verb|Extra| module included functions for topologication, although all
-of them were not implemented for Swedish.
+%For example, the \verb|Extra| module included
+Among those were functions for topologication, although not all
+of them were implemented for Swedish.
 %word order, agreement, tempus, adverbs, adjectivs...
@@
 %can be found.
-Check wich were here first:
-jag ser dig
-dig ser jag
-dig har jag inte sett
-jag har inte sett dig.
-how the parse tree is generated, the finite verb, the subject, the rest. usw.
-han såg katten han tyckte om
-för vem måste jag akta mig
-vem måste jag akta mig för
-
-honom ville hon inte tänka på.
+For example, \verb|FocObj| fronted the object as in sentence \ref{gfSwe:apple}.
+\enumsentence{\shortex{6}
+{Det & äpplet & vill & jag & inte & ha}
+{That & apple & want & I & not &}
+{I don't want that apple}}\label{gfSwe:apple}
+The module provided functions for preposition stranding:
+\enumsentence{\begin{tabular}{ll}
+Stranded preposition & cf.\\
+Vem måste jag akta mig för? & För vem måste jag akta mig?\\
+Who do I need to watch out for? & For whom do I need to wath out?\\
+\end{tabular}}
+
+%jag ser dig
+%dig ser jag
+%dig har jag inte sett
+%jag har inte sett dig.
+%how the parse tree is generated, the finite verb, the subject, the rest. usw.
+%han såg katten han tyckte om
+%för vem måste jag akta mig
+%vem måste jag akta mig för
+
+%honom ville hon inte tänka på.
 
@@
 What is the difference? Some words on differences for pronouns.
+To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
 Example, use vartiants {}.
@@
 %%PASSIV
-Is passive also more used in Swedish?
+\subsubsection{Passive}
 There are two ways of forming passive verb phrases in Swedish: the 
-periphrastic passive, formed by using a modale auxiliary verb 
-\enumsentence{Bilden blev tagen av  av rymdsonden Galileo}
-\enumsentence{The picture was taken by spacecraft Galileo.}
+\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
+\enumsentence{ \shortexnt{6}
+{Bilden &blev& tagen& av & rymdsonden & Galileo}
+{The picture & was & taken & by & the spacecraft & Galileo}}
+\label{gfPass:peri-pass}
 or by adding an \emph{s} to
-the verb:
-\enumsentence{Bilden togs av rymdsonden Galileo}
-\enumsentence{The picture [took+s] by spacecraft Galileo.}
-This second variant, the s-passive, is commonly used in Swedish, some studies
+the verb: \\
+\enumsentence{\shortex{5}
+{Bilden & togs&  av & rymdsonden&  Galileo}
+{The picture&  took+\textbf{s}&  by&the  spacecraft& Galileo}
+{``The picture was taken by the spacecraft Galileo"}}
+\label{gfPass:s-pass}
+This second variant, the \textbf{s-passive}, is commonly used in Swedish, some studies
 suggest that it is used in more than 80\% of the times \cite{laanemets}.
-For the other scandinavian languages, the s-passive can not be used in all tenses for all words.
-The periphrastic passive is preferred, especially in spoken langauge.
-\enumsentence{Bildet ble tatt av romsonden (NO)}
-The resource grammar for Scandinavian therefore allowed/expressed constructions using auxiliary verb. \\
+It is not as common in the other scandinavian languages however, 
+and it can not be used in all tenses for all words. The norweigan 
+translation of both sentence \ref{gfPass:peri-pass} and sentence  \ref{gfPass:s-pass} is
+%The periphrastic passive is preferred, especially in spoken langauge.
+\enumsentence{Bildet ble tatt av romsonden Galileo}
+The resource grammar for Scandinavian therefore expressed the passive using auxiliary verb. \\
 \verb|PassV2 : V2  -> VP | \\
-\verb|         äta -> blev äten| \\
+\verb|         ta -> blev tagen| \\
 This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
@@
 than the default.
-The grammar also allows using three-place verbs \verb|V3| , as \emph{give} to
+The grammar also allows using three-place verbs \verb|V3| , as \emph{give}, to
 form passives:
-\enumsentence{Betyget 3 ges till många elever}
+\enumsentence{\shortex{6}
+{Betyget & 3 &ges &till &många &elever}
+{Grade & 3 & give+\textbf{s}& to & many & students}
+{``Grade 3 is given to many students"}}
 \label{ex:passV3}
 by letting the first object be `removed' and given by the subject. 
-Sentences where the second object is put first in the sentence, like 
+A remark can be made about sentences where the second object is put first in the sentence, like 
 \enumsentence{Till många elever ges betyget 3}
-is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
+This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
 parse this by applying \verb|FocObj|.
@@
 %% Clauses
-\subsection{The clause-level}
-The resource grammar already provided a number of ways to express relative and embedded clauses, such as
-\enumsentence{Frågan är vilken färg den har\\
-                   Pojken, som är blyg, tystnar\\
-                   Han såg kunden som tyckte om sallad\\
-                   Jag tänkte på huset vilket hon bodde i} % better example!!
-
-Since those constructions are inspired by other languages, some of them may sound quite
-non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
-and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
-\emph{`She sleeps, which is good'} as
-\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
-embedded in the relative clause. To fix this, an extra parameter was needed to tell whether
-\emph{`som'} or \emph{`vilket'} should be used. \\
-Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
-The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
-\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
-An more natural sounding alternative would be
-\enumsentence{Jag vill ha en sådan katt som inte fäller.}
-and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
-'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
-needed.
-du är här, vilket jag vet.
-i de äpplen/det äpple du äter bor en mask/ som du ser.\\
-
-\subsection{Presentering, Impersonal constantiation(?), Topicalisering}
-Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish
-\enumsentence{Det sitter en fågel på taket}
-\emph{`There is a bird sitting on the roof'}
-or, literally
-\emph{`It sits a bird on the roof'}.\\
+\subsubsection*{Presentering, Impersonal constructions, Topicalisering}
+Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish.
+\enumsentence{\shortex{6}
+{Det & sitter & en & fågel & på & taket}
+{It & sits & a & bird & on & the roof}
+{``There is a bird sitting on the roof''}}
 \emph{`Det'} has the position of the subject, and the real subject, 
@@
 Transitive verbs may not be used like this
-\enumsentence{*Det äter en fågel på taket.}
+\enumsentence{
+*Det äter en fågel på taket\\
+ ''There is a bird eating on the roof"}
+
 unless their in passive form
-\enumsentence{Det äts en fågel på taket.}
-%det finns/saknas/fattas
+\enumsentence{
+Det äts en fågel på taket \\
+A bird is being eaten on the roof}
+A very common example of this is sentences with the verbs \emph{finnas},
+\emph{saknas} and \emph{fattas}.
+\enumsentence{\shortex{3}
+{Det & finns & kaffe}
+{It & exist & coffee}
+{''There is coffee"}}
+\enumsentence{\shortex{3}
+{Det & saknas&  kaffe}
+{It & misses & coffee}
+{''There is no coffee"}}
+\enumsentence{\shortex{3}
+{Det & fattas & kaffe}
+{It & misses & coffee}
+{''There is no coffee"}}
+A special GF category was needed for this, the \verb|SimpleVP|.
+
 There are also restrictions on the nounphrase of the subject, it can only
 be determined in certain ways
-\enumsentence{*Det sitter den fågel på taket. \\
-              **Det sitter fåglarna på taket.}
-The distinction lays in the definitness of the noun itself, (not the whole noun phrase..)
-it must be indefinite, even though determiners are allowed. Jfr
+\enumsentence{\begin{tabular}{ll}
+*Det sitter den fågel på taket. & **Det sitter fåglarna på taket.\\
+There are that bird sitting on the roof & There are the birds sitting on the roof \\
+\end{tabular}}
+The distinction lays in the definitness of the noun: 
+it must be indefinite, even though some determiners are allowed.
 \enumsentence{Det sitter några fåglar på taket\\
@@
 %%Semantic difference, atm we allow any NP. \\
-The grammar handles this by looking at the determiner used
-\verb|FormalSub : SimpleVP -> Det -> NP -> Cl ;|
+The GF grammar does not make any distinction between indefinite and definite noun
+phrases. Therefore this is handled by looking at the determiner. \\
+\verb|FormalSub : SimpleVP -> Det -> NP -> Cl ;|\\
 The function combines a verb phrase a determiner and a noun phrase in to a clause.
-\verb|SimpleVP| is a normal VP, but with the restriction that the main verb must be
-intransitiv, a \verb|V|, (and passives?). If the  determiner requires a noun in determinened
-form, the clause is put to NONEXIST.
+If the determiner requires a noun in determinened form, the clause is put to NONEXIST.
+However, this is not an ideal solution. There are other examples of when we
+need to know the definitness of the noun phrase, for example when using
+the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
+or when using some predeterminers (see section `Quantifiers')
 
@@
 
+\subsubsection{Quantifiers}
 pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
@@
  this for translation, neutr and utr may differ in the other language.
+
+ 'de flesta de gula hästarna'
 
@@
 %% Refl 
+\subsubsection*{Reflexive pronouns}
 Could varandra be thought of as reflexive for 'oss'?
@@
 here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
+\subsubsection*{Relative clauses}
+The resource grammar already provided a number of ways to express relative and embedded clauses, such as
+\enumsentence{Frågan är vilken färg den har\\
+                   Pojken, som är blyg, tystnar\\
+                   Han såg kunden som tyckte om sallad\\
+                   Jag tänkte på huset i vilket hon bodde} % better example!!
+
+Kanske mindre om detta, skriv om vad jag nu la till.
+om 'de äpplen'
+Since those constructions are inspired by other languages, some of them may sound quite
+non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
+and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
+\emph{`She sleeps, which is good'} as
+\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
+embedded in the relative clause. To fix this, an extra parameter was needed to tell whether
+\emph{`som'} or \emph{`vilket'} should be used. \\
+Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
+The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
+\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
+An more natural sounding alternative would be
+\enumsentence{Jag vill ha en sådan katt som inte fäller.}
+and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
+'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
+needed.
+du är här, vilket jag vet.
+i de äpplen/det äpple du äter bor en mask/ som du ser.\\
 

### 2011-11-29 doc/report/report.tex
report

@@
 {`You  don't see me'.}} \label{ex:swedish-svo}
+\vspace{-3mm}
 \enumsentence{
@@
 {Me & see & you & not.}}\label{ex:swedish-ovs} 
+\vspace{-3mm}
 \enumsentence{
@@
 
+Could varandra be thought of as reflexive for 'oss'?
 %Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
@@
 \begin{figure}[h]
-\includegraphics[width=60mm]{gfTree.pdf}
+\includegraphics[width=70mm,height=70mm]{apples.pdf}
 \caption{Parse tree for ``Han har inte ätit de gula äpplena idag".}%correct
@@
 {''There is no coffee"}}
-A special GF category was needed for this, the \verb|SimpleVP|.
-
-There are also restrictions on the nounphrase of the subject, it can only
-be determined in certain ways
+
+A special GF category, \verb|SimpleVP|, was needed for verb phrases of the type
+described above.
+
+There are also restrictions on the subject.
 \enumsentence{\begin{tabular}{ll}
-*Det sitter den fågel på taket. & **Det sitter fåglarna på taket.\\
-There are that bird sitting on the roof & There are the birds sitting on the roof \\
+*Det sitter den fågeln på taket. & *Det sitter fågeln på taket.\\
+``There are that bird sitting on the roof'' & ``There are the bird sitting on the roof'' \\
 \end{tabular}}
-The distinction lays in the definitness of the noun: 
-it must be indefinite, even though some determiners are allowed.
+The noun must be in indefinite form, even though some determiners are allowed.
 \enumsentence{Det sitter några fåglar på taket\\
-              **Det sitter vardera fåglarna på taket}
+              ``There are some birds on the roof''}
+\enumsentence{*Det sitter fåglarna på taket\\
+``There are the birds on the roof''}
 %%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
@@
 phrases. Therefore this is handled by looking at the determiner. \\
-\verb|FormalSub : SimpleVP -> Det -> NP -> Cl ;|\\
+\verb|FormalSub : SimpleVP -> Det -> CN -> Cl ;|\\
 The function combines a verb phrase a determiner and a noun phrase in to a clause.
@@
 the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
-or when using some predeterminers (see section `Quantifiers')
-
-finnas: här får man ha definit: det finns de som inte vill. Samma RelCl som
-jag just la till, de äppplen. (de + relsats)
-
-'det är kul att..' Näst vanligaste två-ordskombinationen i talad svenska.
+or when using some predeterminers (see section `Quantifiers').
+
+%finnas: här får man ha definit: det finns de som inte vill. Samma RelCl som
+%jag just la till, de äppplen. (de + relsats)
+
+%'det är kul att..' Näst vanligaste två-ordskombinationen i talad svenska.
 %%
@@
 
-\subsubsection{Quantifiers}
-pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
-has to decide, one category for each?
-Genitive -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form
-           fixing this include restructuring all
-Det-utr. not effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
+\subsubsection{Quantifiers, determiners and predeterminers}
+The GF analyse differs sometimes. Experiments of how to extract
+this information from Talbanken by using Saldo as a starting poin.
+One of the biggest differences between the GF analysis and the one made in Talbanken
+is related to pronouns. While GF only recognizes the personal pronouns, 
+`sådan', `fler`, `somliga' etc. are all regarded as pronouns in Talbanken and Saldo.
+In the GF analyse they may be quantifiers, determiners or predeterminers. 
+All pronouns from Saldo was listed and by searching Talbanken for which tags
+they got we could sort out the ones acting either as
+determiners, as `sådana' in sentence \ref{ex:QuantDet}, or as adjectives as
+`sådan' in sentence \ref{ex:QuantAdj}. 
+\enumsentence{Sådana katter vill jag ha}\label{ex:QuantDet}
+\vspace{-3mm}
+\enumsentence{Han är sådan}\label{ex:QuantAdj}
+\vspace{-3mm}
+\enumsentence{Sådant är inte bra}\label{ex:QuantDet2}
+Since all determiners can be used alone as noun phrases, sentence
+\ref{ex:QuantDet2} is also accepted.
+%what is the difference really???
+%The techniques of searching Talbanken can be used
+%for this purpose as well. 
+%combining the information in Saldo and Talbanken
+%pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
+%has to decide, one category for each?
+
+\subsubsection{Other small things}
+
+\textbf{Genitive} -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
+           fixing this include restructuring all???
+\textbf{Det-utr}. not effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
  Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
@@
 
+\textbf{finitness}
  'de flesta de gula hästarna'
 
+\textbf{advs}
 -- or is this really just AdVs? 
@@
 
+\textbf{bara}
 but have other words like 'bara' which can be put in other positions.
@@
 
+\textbf{antalet}
 Also needed categories for 'pronouns' wich can act in different ways.
@@
 
-The VP require many changes, to add passive (no), PPart, bara.
+\textbf{VP}
+The VP require many changes PPart, bara.
 About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
@@
 
-%% Refl 
+\textbf{SuperlA}
+
 \subsubsection*{Reflexive pronouns}
-Could varandra be thought of as reflexive for 'oss'?
-varandra and sin/sitt/sig. Not subject, but NP.
+%varandra and sin/sitt/sig.
+Both sig and sin
+Not subject, but NP.
 Want the following:
@@
 Kanske mindre om detta, skriv om vad jag nu la till.
-om 'de äpplen'
+om 'de äpplen', which are also used in 'finnas'
 Since those constructions are inspired by other languages, some of them may sound quite

### 2011-11-30 doc/report/report.tex
report

@@
 \subsubsection*{Reflexive pronouns}
+\label{swe:refl}
 The Scandinavian language have reflexive pronouns\cite{H&H310}
@@
 large electronic lexicon developed and maintained at Gothenburg University. It is
-built upon Svenskt Associations Lexicon and contains information about modern
-Swedish and the words are equipped with both sematic and a morphological
+built upon Svenskt Associations Lexicon and contains information about more than 
+120 000 modern Swedish words.
+The words are equipped with both sematic and a morphological
 information. The user can find compounding analyses, examples of usage in corpora,
 and graphs of sematically connected words.\\
-Example of katt? \\
-For our purpose however, only the morphological data was needed.\\
-Example of morpho for katt.\\
+\newpage % only until saldo can be on same page as the pic
+\begin{figure}[h!]
+\centering
+\subfloat[Morphological info for ``katt"]{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
+\hspace{20mm}
+\subfloat[Graph showing the hyponyms of ``katt"]{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
+\caption{Information in Saldo}
+\label{fig:saldo}
+\end{figure}
+For our purpose however, only the morphological data was needed.
 Saldo provides this in XML format under LPGL. The data can be processed and
@@
 
-\subsection{GF}
+\subsection{Grammatical Framework}
 The main component of the the project is the Grammatical Framework\ref{gfbok} (GF). It is
@@
 
-Being developed with multilinguality in mind, GF makes an important distiction between
+An important feature of GF is
+%Being developed with multilinguality in mind, GF makes an important 
+the distiction between
 \textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
-shared by many languages, it models the semantics of the domain without blanda in
+shared by many languages, it models the semantics of the domain without concern for
 language specific features such as agreement or word order.
@@
 \caption{A small example of an abstract syntax}
+\label{fig:gfAbstract1}
 \end{figure}
 
-The example shows an abstract grammar defining three categories, % \verb|Categories|,
+The example in figure \ref{fig:gfAbstract1} shows an abstract grammar defining three categories, % \verb|Categories|,
 one for nouns, one for verbs and one for sentences. An abstract grammar also gives
-the types of functions. In this case we have \verb|Pred| which
+the types of functions. In our case we have \verb|Pred| which
 tells us that by taking a noun and a verb we can form a sentence. At this stage, 
 no information of how this is done is given. The grammars also defines two words,
-\verb|cat| and \verb|sleep|.
+the noun \verb|cat| and the verb \verb|sleep|.
 
@@
 Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
-for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|
+for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|,
 and \verb|Pred| simply glues the two strings ``katten" and ``sover" together:
 \verb|Pred cat sleep = "katten sover"|.\\
-Let's look at a more complicated example, where nouns can be used in both
+We get a more complicated example if we allow the nouns to be used in both
 plural and singular. To do this, we add a category \verb|NP| to the
@@
          N = {s : Num => Str} ;
-         NP = {s : Str ; num = Num} ;
+         NP = {s : Str ; num : Num} ;
   lin   
     Pred n v = n.s ++ v.s ! n.num ;
-    NPl n = {n.s ! Pl ; num = Pl} ;
-    NSg n = {n.s ! Sg ; num = Sp} ;
+    NPl n = {s = n.s ! Pl ; num = Pl} ;
+    NSg n = {s = n.s ! Sg ; num = Sg} ;
     cat_N = {s = table {Sg => "the cat" ;
@@
 We now have two implementations of the abstract. The resulting GF grammar is able to
-both parse a string to the abstract tree and to go in the other direction; to produce
-a string of natural language given a abstract tree. This step is called linearization.
+both parse a string to a abstract tree and to go in the other direction; to produce
+a string of natural language given an abstract tree. This step is called linearization.
 Translation is a consequence of this, we can parse a Swedish string and then 
@@
 
-Valency information is normally built into the GF categories. Add V2 to the example.
-Point out how N + V can be glued together to a string, the components cannot be seen
-anymore.
+Valency information is normally built into the GF categories. To indicate that
+a verb is transitive and needs an object, we use the category \verb_V2_.
+\begin{verbatim}
+abstract:
+cat V2 ; 
+eat_V2 : V2 ;
+rat_N ;
+UseV2 : V2 -> NP -> V ;
+
+concrete:
+eat_V2 = "äta" ;
+rat_N = {s = table {Sg => "råttan" ;
+                    Pl => "råttorna"}};
+UseV2 v np = v ++ np ;
+\end{verbatim}
+
+The example above now allows ``Katten äter råttan", but not ``Katten äter".
+\verb|UseV2| concatenate the strings ``äter" and ``råttan". Note that if we
+wanted to add negation, ``katten äter inte råttan", or question "äter katten råttan",
+we cannot glue the verb to the object at once. Instead we would have to use a table,
+with one field for the verb and one for the object.
 
@@
 
-\subsection{GF for Swedish}
+\subsection{The Swedish resource grammar}
 The GF resource grammars contains a fundamental description of Swedish,
@@
 \includegraphics[width=70mm,height=70mm]{apples.pdf}
-\caption{Parse tree for ``Han har inte ätit de gula äpplena idag".}%correct
+\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}%correct
 \label{gfSwe:parsetree}
@@
 %% Clauses
-\subsubsection*{Presentering, Impersonal constructions, Topicalisering}
+\subsubsection{Impersonal constructions}%Presentering, , Topicalisering
+\label{sec:Formal}
 Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish.
@@
 the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
-or when using some predeterminers (see section `Quantifiers').
+or when using some predeterminers (see section `Quantifiers'). Not true, not the same :(
 
@@
 
-\subsubsection*{Reflexive pronouns}
+\subsubsection{Reflexive pronouns}
 %varandra and sin/sitt/sig.
-Both sig and sin
-Not subject, but NP.
-Want the following:
-\enumsentence{Han såg sina barns skor.}
-\enumsentence{Sina vantar hade han glömt på tåget.}
+An important piece in a Swedish grammar is the reflexive pronouns and
+%both \emph{sin} and , eg. \emph{sin}.
+the possessive pronouns as described in section \ref{swe:refl}.
+Those make up noun phrases wich are not allowed in subjects, 
+and we want to be able to parse to following sentences:
+\enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
+He saw his children's shoes.}
+\enumsentence{Sina vantar hade han glömt på tåget.\\
+His gloves, he had forgotten on the train.}
+\enumsentence{Han ger sina pengar till sina syskon.\\
+He gives his money to his children.}
+\enumsentence{Han är längre än sin kompis.\\
+He is taller than his friend.} \label{ex:ReflAP}
+\vspace{-2mm}
+\enumsentence{Han är här oftare än sin kompis.\\
+He is here more often than his friend.} \label{ex:ReflAdv}
+\vspace{-2mm}
+\enumsentence{Hon tyckte om skolan och alla sina elever.\\
+She liked the school and all her students.}\label{ex:ReflConj}
+\enumsentence{Han såg sina få böcker och sin penna.\\
+He saw his few books and his pencil.}\label{ex:ReflConj2}
+\vspace{-2mm}
+\enumsentence{Hon ber alla sina kompisar att gå\\
+She asks all her friends to leave}\label{ex:ReflPredet}
+\enumsentence{Jag vill att han rakar sig.\\
+I want him to shave.}
 but not
-\enumsentence{*Sina vantar var kvar på tåget.}
-\enumsentence{Han ger sina pengar till sina syskon.}
-\enumsentence{*Jag ger sina pengar till sina syskon.}
-\enumsentence{Han är längre än sin kompis.}
-\enumsentence{Hon tyckte om alla sina elever.}
-\enumsentence{Han såg sin hund och sitt hus}
-\label{ex:refGen}
-\enumsentence{Han är längre än sin kompis}
-\label{ex:ReflAP}
-% hon ber sin katt gå - sin subject in clause.
-
-Suggest a connection between the subject and the object.
-Could be combined with 'sig', a ReflNP without another noun.
--- man ser sig!!
-Make a distinction between objects and subjects?
-do not want to copy all rules for another category. Would like depent types!
-work pretty well by letting all NP depend on the subject in the phrase.
-The current version is nice and works for recursive genitivs!
-Extra info, but maybe this is just the case about NPs.
-Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
-*Slash* : VPSlash -> NPR -> VP ;
-Keep number of fields down, but introduces another category, and one more
-label to the parse tree, is it interesting? How to combine it with genitive 
-without making it ambiguous (Han såg barnens hund).
-What if we add alot of other rules that uses NP, will they all need to be
-duplicated then?
-
-'han såg sin mammas bil', they same could be applied to
-APs for example \ref{ex:ReflAP} or Adv for
-\enumsentence{Han leker oftare än sin kompis}
- (ComparAdvAdj)
-Atm you can't focus this kind of object, then a ClSlash need info about the
-subject. FunRP also tricky (and what does it mean, we can say
-Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
-Kort sagt, the information about the subject must be spread
-in different parts of the sentence. 
-Could also decide to always produc 'sin' -> jag såg sin bil
-to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
-in third person, and add nonexist for others. But this way we can also get rid
-of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
-'jag såg din bil'.
-Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
-not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
-etc.
-varandra, works the same but has no case for Sg.
-Johan vill att jag rakar sig* -> honom  
+\enumsentence{*Sina vantar var kvar på tåget.\\
+Her gloves were left on the train.}\label{ex:ReflSubj}
+\enumsentence{*Jag ger sina pengar till sina syskon.\\
+I give his money to his siblings. }
+\enumsentence{*Han vill att jag rakar sig.\\
+He wants me to shave himself.}
+
+The reflexive pronouns require another subject whith which they agree in
+number and person. Apart from this restriction, they may be used as normal
+noun phrases and conjuncted (sentence \ref{ex:ReflConj} and \ref{ex:ReflConj2}),
+determined (sentence \ref{ex:ReflPredet}) and used as genitives (sentence
+\ref{ex:ReflGen}) as such.
+%Could be combined with 'sig', a ReflNP without another noun.
+In short, what is wanted can be summerized as follows:
+\begin{figure}
+\begin{enumerate}
+\item{
+Construct a noun phrase from a common noun phrase \\
+katt $\rightarrow$ sin katt} % \\
+\item{Modify it like a \verb|NP|\\
+sina katter $\rightarrow$ alla sina katter}
+\item{Use it only as object\\
+sin katt $\rightarrow$ han såg sin katt }
+\end{enumerate}
+\caption{Rules for noun phrases with reflexive possessive pronouns}
+\label{fig:reflrestrict}
+\end{figure}
+If we ignore the second requirement, special functions for using common nouns
+together with reflexive possessive pronouns could be introduced:
+\begin{verbatim}
+ReflVP: CN -> VPSlash -> VP ;
+        mat + äta     -> äta sin mat 
+\end{verbatim}
+This way, using the phrases as subject is excluded, but neither of sentence
+\ref{ex:ReflConj}, \ref{ex:ReflConj2} or \ref{ex:ReflPredet} is allowed.\\
+The simplest way of fixing this is to treat the phrases as normal \verb_NP_s.
+We can add a rule for using the possessives with common noun phrases:
+\begin{verbatim}
+ReflNP: CN -> NP ;
+        mat -> sin mat 
+PredetNP : Predet -> NP       -> NP ;
+           alla   + sina barn -> alla sina barn
+\end{verbatim}
+But by keeping them as \verb_NP_s we lose information, and there is no way of
+keeping them from being used as subjects. We have ignored the third restriction
+and are allowing sentences like \ref{ex:ReflSubj}.
+
+The solution introduces a new object type, \verb_Obj_, which is
+identical to normal \verb_NP_ except for that it depends on the agreement of
+the subject. All functions using noun phrase as objects are changed to dealing
+with objects instead.
+\begin{verbatim}
+ReflVP: Obj     -> VPSlash -> VP ;
+     -- sin mat +  äta     -> äta sin mat 
+NPtoObj : NP  -> Obj ;
+PredVP  : NP -> VP -> Cl ;  -- NP used as subject
+\end{verbatim}
+Any normal noun phrase may also be used as an object, whereas subjects only
+can be made up by \verb-NP-s.
+Since the resource grammar did not make any distinction between
+subjects and objects, the new Swedish grammar needed to be separated from the old
+structure. 
+Further, the dependence on the subject needs to be applied to adverbial phrases and
+adjective phrases as well. The adjetive phrase in sentence \ref{ex:ReflAP} and
+the adverbial phrase in sentence \ref{ex:ReflAdv} are examples of this; if the
+subject was changed to 2nd person, they  would need to change too.
+\enumsentence{\textbf{Du} är längre än \textbf{din} kompis.\\
+You are taller than your friend.} \label{ex:ReflAP2}
+\vspace{-2mm}
+\enumsentence{*Du är här oftare än sin kompis.\\
+You are here more often than his friend.} \label{ex:ReflAdv2}
+Adverbial phrases may also be parts of new noun phrases
+\enumsentence{\shortex{3}
+{Taket & på sitt hus &}
+{\scriptsize{\textbf{NP +}} & \scriptsize{\textbf{(Adv with Obj)}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{The roof of his house}}
+\enumsentence{\shortex{3}
+{Sitt hus & på berget &}
+{\scriptsize{\textbf{Obj +}} & \scriptsize{\textbf{Adv}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{His house on the mountain}}
+
+The first restriction in figure \ref{fig:reflrestrict} should be slightly
+modified, to allow certain kind of \verb|NP|s to be used.
+\begin{figure}[h]
+\begin{enumerate}
+\item{
+Construct a noun phrase from a common noun phrase or certain noun phrases \\
+få katter $\rightarrow$ sina få katter} 
+\end{enumerate}
+\caption{Modification of \ref{fig:reflrestrict}}
+\end{figure}
+
+The kind of noun phrases that can be used seems to coincide with the ones
+that can be used with formal subjects (see section `Impersonal constructions').
+
+% agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
+% F1,2,3 : NP -> VP -> Cl ; NP needs gender (Masc,Fem,Utr,Neutr)
+
+%Make a distinction between objects and subjects?
+%do not want to copy all rules for another category. Would like depent types!
+%work pretty well by letting all NP depend on the subject in the phrase.
+%The current version is nice and works for recursive genitivs!
+%Extra info, but maybe this is just the case about NPs.
+%Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
+%*Slash* : VPSlash -> NPR -> VP ;
+%Keep number of fields down, but introduces another category, and one more
+%label to the parse tree, is it interesting? How to combine it with genitive 
+%without making it ambiguous (Han såg barnens hund).
+%What if we add alot of other rules that uses NP, will they all need to be
+%duplicated then?
+
+%'han såg sin mammas bil', they same could be applied to
+%APs for example \ref{ex:ReflAP} or Adv for
+%\enumsentence{Han leker oftare än sin kompis}
+% (ComparAdvAdj)
+%Atm you can't focus this kind of object, then a ClSlash need info about the
+%subject. FunRP also tricky (and what does it mean, we can say
+%Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
+%Kort sagt, the information about the subject must be spread
+%in different parts of the sentence. 
+%Could also decide to always produc 'sin' -> jag såg sin bil
+%to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
+%in third person, and add nonexist for others. But this way we can also get rid
+%of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
+%'jag såg din bil'.
+%Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
+%not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
+%etc.
+%varandra, works the same but has no case for Sg.
 %%
@@
 here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
-\subsubsection*{Relative clauses}
+\subsubsection{Relative clauses}
 The resource grammar already provided a number of ways to express relative and embedded clauses, such as

### 2011-12-01 doc/report/report.tex
report and pictures

@@
 \usepackage{subfig}
+\usepackage{wrapfig}
 % exempel: 2) Inte vet jag
@@
 \subsection{Swedish}
-Swedish\cite{SAG:intro} is a North-Germanic language,
+Swedish\cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
 closely related to Norweigan and Danish. The languages share most of their
@@
 \subsubsection*{Verb second}
-Swedish is a verb-second language: the
+Swedish is a verb-second language\cite[p. 116]{gunlog}: the
 second constituent of a declarative main clause must consist of a verb.
-The normal word order is subject-verb-object, but any syntactic category can be fronted\cite{H&H:1027}.
+The normal word order is subject-verb-object, but any syntactic category can be fronted\cite[\textsection 1027]{H&H}.
 Topicalisation is very common, especially for temporal and spatial adverbial phrases.
@@
 \subsubsection*{Impersonal constructions}
-Constructions with 'det är/var' ('it is/was') is very common\cite{H&H309d}:
+Constructions with 'det är/var' ('it is/was') is very common\cite[\textsection 309d]{H&H}:
 \enumsentence{
@@
 \label{swe:refl}
-The Scandinavian language have reflexive pronouns\cite{H&H310}
-and special reflexive forms of the possessive pronouns for the 3rd
-person\cite{H&H319}. Those are distinct from the normal 3rd person forms.
+The Scandinavian language have special reflexive pronouns
+and possessive pronouns for the 3rd
+person\cite[\textsection 310 \& 319]{H&H}. Those are distinct from the normal 3rd person forms.
 \enumsentence{
@@
 \subsection{Grammatical Framework}
-The main component of the the project is the Grammatical Framework\ref{gfbok} (GF). It is
+The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
 a functional grammar formalism based on Martin-Löf type theory.
 It has been developed for about 20? years and are inspired by programming langauges
-like Haskell\footnote{}, ML\footnote{}, Prolog\footnote{} and Agda\footnote{}. Grammars written in GF can be used together with languages 
+like Haskell\cite{haskell},
+ML\cite{ml}, Prolog\cite{bratko} and
+Agda\cite{agda}.
+Grammars written in GF can be used together with languages 
 like Java, Java script and Haskell and works for both parsing and generation.
@@
 By using a controlled language like this,  good and reliable translation can be
-ensured. Some examples of projects using GF are the math .. \cite{olga}, the
-Online Syllogism Solver\cite{malin!}.  The framework is also used in the
-european project MOLTO for online translation between upto 15 language.\\
+ensured. Some examples of projects using GF are WebAlt\cite{webalt} and something else.
+%, the Online Syllogism Solver\cite{malin!}. 
+The framework is also used in the
+european project MOLTO\cite{molto} for online translation between upto 15 language.\\
 
@@
 semantics.
-Further, experiments on using GF for free parsing has been conducted (Angelov,2010?).
+Further, experiments on using GF for free parsing has been conducted \cite[\textsection 4.1]{gfMech} 
 This is also a future goal for our project and one of the delmål has been to
@@
 parts. Two of them are transcribitions of spoken language, one a collection of
-text written by high school (?) students, and the last one, section P,
-consists of professionally written Swedish gathered from newspapers and ... .\\
-Talbanken was also used to train the Swedish version of the Maltparser\cite{mal}
-and was därmed reutgiven in a more modern version\cite{nivre}. This work
-was done by Nivre and ... in 2005. They released it in Malt and Tiger XML-formats,
-where the trees had been made deeper and smt more. \\
+text written by high school students, and the last one, section P,
+consists of professionally written Swedish gathered from newspapers, brochures and textbooks.\\
+Talbanken was also used to train the Swedish version of the Maltparser\cite{malt}
+and was thus redistributed in a more modern version\cite{talbanken05} This work
+was done by Nivre, Hall and Nilsson in 2005. They released it in Malt and Tiger XML-formats,
+where the trees had been made deeper and more detailed. \\
 In this project, as well as for the Malt parser, the more than 6000 sentences
@@
 but grammars?
-Many of the parsers not ruled based, the most known one Malt ... .
-However has been two, about CLE and what happend to it. Logic and translation
-to Swedish, The spoken lang. translator.
+Many of the parsers not ruled based, the most known one Malt\cite{malt} ... .
+However has been two, about CLE\cite{cle} and what happend to it. Logic and translation
+to Swedish, The spoken lang. translator\cite{cle2}. Wiren \cite{wiren}, cassSwe \cite{casswe}
 Other Sw. parsers and grammars, evaluation of them? 
@@
 %\section{Proceedings med bättre ord}  
-The purpose has been to prepare the grammar for large scale parsing.
-Develop method used for English to fit Swedish, pynta the sw grammar with
-language specific constructions and create methods for lexicon.
-Earlier have done tool, but the lexicon was not big enough to be able
-to fix it in this way. Tool may instead be used for complementing.
+The purpose of this project has been to prepare the grammar for free parsing.
+This has meant to develop earlier techniques to fit for Swedish, create methods
+for achieving and keeping a extensive lexicon and to adapt the existing
+resource grammar to fit for more language specific constrictions.
+The methodology and results will be presented in the following section.
 
@@
 \label{sec:prog.saldo}
-In saldo there is also semantic info, so
-We need the numbers from saldo if there are to lemmas with the same
-identifier. sluta\_V and sluta2\_V (slöt). However gf does not want more than one table
-if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
-We are just interested in morphology, so far. Dont want too big lexicon.
-
-Saldo has been imported before (Krasimir, 2009?), but the code needed updates
-to not miss many important words.
-
-The basic algorith is the same. To understand this need to look at differences.
-Example of gf and saldo table.
+Saldo is compatible with GF, which makes it easy to transform the morphological
+lexicon into our format. The lexicon has been imported earlier (Krasimir,2009?)
+but the code needed updates to not miss many important words.
+Experiments have been made to use a interactive tool for lexical aquisation,
+but the dictionary was still much to small. The aquisation tool may still
+be used for complenting the lexicon.
+
+The basic algorithm for porting Saldo is the same as Angelov's. It makes use of
+the smart paradigms in GF : explanation.
+%Find the paper!
+%To understand it, we first need to look at differences between Saldo and a GF lexicon.
+The program will try to feed as few forms as possible to the GF smart paradigm.
+It starts from the easiest case, compare the results and either saves the code or 
+makes another try using more word forms.
+Figure \ref{pic:TabVax} shows a the information given by Saldo and by GF respectively.\\
+\begin{figure}[h]
+  \begin{center}
+\subfloat[Saldo]{\label{pic:saldoVax}\includegraphics[height=90mm]{saldoVax.png}}
+\hspace{5mm}
+\subfloat[GF]{\label{pic:gfVax}\includegraphics[height=60mm]{gfVax.png}}
+\caption{Inflection table for the verb ``växa" (grow)}
+\label{pic:TabVax}
+  \end{center}
+\end{figure}
+We can see that there are more saldo forms and some are only in GF.
+
 Explanation of the algorithm: map categories, try each paradigm, compare the table of forms
@@
 different grammatical analyse, a lot of pronouns. Example.
+
+In Saldo there is also semantic info, so
+We need the numbers from saldo if there are to lemmas with the same
+identifier. växa (bli större), växa (vara en planta) vs. sluta\_V and sluta2\_V
+(slöt). However gf does not want more than one table
+if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
+We are just interested in morphology, so far. Dont want too big lexicon.
 
@@
 
+\subsection{Results}
+How many words, how many Talbanken words are not here now? Why?
+
 \section{The grammar}
+\label{sec:prog.grammar}
 Although we can't expect to get a full coverage, we would like a tillfredsställande
@@
 tycka om /tycka om -> om vad tycker du inte. fast bra
+about slashes\cite{gazdar}
 
@@
 \label{sec:Formal}
-Formal subjects \cite{SAG-19,H&H-309d} is often used in Swedish.
+Formal subjects \cite[19]{SAG} is often used in Swedish.
 \enumsentence{\shortex{6}
@@
 
+In general, many categories need more parameters and fields, tex Quant can be used as NP
+and therefor needs to depend on Case, de -> dem.
+
+here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
 \subsubsection{Reflexive pronouns}
@@
 
-In general, many categories need more parameters and fields, tex Quant can be used as NP
-and therefor needs to depend on Case, de -> dem.
-
-here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
 \subsubsection{Relative clauses}
@@
 
-A program for translating trees to GF syntax has been done before (Angelov,).
-It translated trees from the English Penn Treebank. By starting this progam, we now have
+A translation of the English Penn Treebank to the GF syntax has previosly been done \cite{gfpenn}.
+By starting this progam, we now have
 a translation that works for the quite differently annotated Talbanken. The differences between
@@
 
+\bibliographystyle{ieeetr}
+\bibliography{masterbib}
 \end{document}

### 2011-12-01 doc/report/report.tex
reort

@@
 they all share the same abstract syntax.
-The resource grammars describe how to decline words and how to construct
-phrases and sentences.
+The resource grammars describe how to construct phrases and sentences and how to
+decline words. The latter is done by smart paradigms: functions that analyses
+some given forms of a word to find out how the inflection table should look.
+For example, the declination of many Swedish nouns can be determined by looking
+at the singular form only. This is the case for ``flicka"??, the correct declination
+can be guessed by noting that the word ends with an ``a". Like most Swedish nouns
+ending with an ``a", it follows the first?? declination. For others, like
+"hjärta", also the plural form "hjärtan" is needed, even though it ends
+with an ``a". The worst case is nouns that need four forms, both singular and
+plural in definite and indefinite form.
 There are also a small test lexicon included, containing some hundered of common words.\\
@@
 
-\section{Saldo}
+\section{Importing Saldo}
 \label{sec:prog.saldo}

### 2011-12-02 doc/report/report.tex
report!

@@
 %ska innehålla vettiga sökord
-\title{Scaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Title of the report} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 \end{figure}
-For our purpose however, only the morphological data was needed.
-Saldo provides this in XML format under LPGL. The data can be processed and
-translated to GF format as described in 
+
+If one word
+have different meanings, there will be separate entries for each of them.
+The word ``växa" (``grow") for example, has two meanings: \emph{getting bigger}
+\enumsentence{Gräset växte varje dag\\
+The grass grew each day}
+or \emph{being a plant}. 
+\enumsentence{Det växte en gammal ek där.\\
+There stood an old oak.}
+The word ``växa" is however declined the same way for both interpretations of the word.
+Hence Saldo's morpological lexicon, provided under LPGL in XML format,
+only keeps one entry of ``växa".
+For the homonyms that does not share the whole inflection tables, different entries are 
+kept in the morphological lexicon.
+
+For our purpose, only the morphological lexcion was needed.
+The data can be processed and translated to GF format as described in 
 section \ref{sec:prog.saldo}.
-%Other usages?
 
@@
 they all share the same abstract syntax.
+There are also a small test lexicon included, containing some hundered of common words.
+
 The resource grammars describe how to construct phrases and sentences and how to
@@
 For example, the declination of many Swedish nouns can be determined by looking
-at the singular form only. This is the case for ``flicka"??, the correct declination
-can be guessed by noting that the word ends with an ``a". Like most Swedish nouns
-ending with an ``a", it follows the first?? declination. For others, like
-"hjärta", also the plural form "hjärtan" is needed, even though it ends
-with an ``a". The worst case is nouns that need four forms, both singular and
+at the singular form only. 
+\begin{wrapfigure}{l}{0.4\textwidth}
+\begin{tabular}{| l | l |}
+\hline
+1st declination & 5th declination \\
+\hline
+flick\textbf{a}    &     hjärt\textbf{a}   \\
+flick\textbf{an}    &    hjärt\textbf{at}  \\
+flic\textbf{kor}    &    hjärt\textbf{an}  \\
+flic\textbf{korna}  &    hjärt\textbf{ana} \\
+\hline
+\end{tabular}
+\end{wrapfigure}
+This is the case for ``flicka", the correct declination
+can be guessed by using the first declination. % and noting that the word ends with an ``a". 
+%Like most Swedish nouns ending with an ``a", it follows the first declination.
+For others, like "hjärta", also the plural form "hjärtan" is needed.
+%, even though it ends with an ``a". 
+The worst case is nouns that need four forms, both singular and
 plural in definite and indefinite form.
-There are also a small test lexicon included, containing some hundered of common words.\\
-
+
+\vspace{5mm}
 As an open source-project, GF is constantely being developed and improved. New
@@
 Being an freely avaible, manually annotade, large scale treebank,
-Talbanken\cite{talbanken} was perfect for this ändamål.
+Talbanken\cite{talbanken} was perfect for our purpose.
 It was created in the 1970s at Lund University. In total it consists of four
@@
 
-The basic algorithm for porting Saldo is the same as Angelov's. It makes use of
-the smart paradigms in GF : explanation.
+The basic algorithm for porting Saldo is the same as Angelov's, it works by
+making use of the smart paradigms in GF.
 %Find the paper!
 %To understand it, we first need to look at differences between Saldo and a GF lexicon.
-The program will try to feed as few forms as possible to the GF smart paradigm.
-It starts from the easiest case, compare the results and either saves the code or 
-makes another try using more word forms.
+The program will try to create a lexicon by finding a number of forms for every
+word and feeding them to the smart paradigms. As few forms as possible should be
+used, and the resulting table generated by GF should be equivalent to the one found
+in Saldo. This process is done iteratively,
+starting from the easiest case. For each attempt the results are compared and when
+the correct way of using a word has been found, the code is saved.
 Figure \ref{pic:TabVax} shows a the information given by Saldo and by GF respectively.\\
@@
 \end{figure}
-We can see that there are more saldo forms and some are only in GF.
-
-Explanation of the algorithm: map categories, try each paradigm, compare the table of forms
-created. If several forms in saldo, gf should pick on of them. Saldo has forms for
-compounds, gf does not. A grammar is written, the correct ones saved, the others 
-retried. Try all paradigms which can be formed from saldo. 
-There are a number of differences between and saldo Till skillnad från GF,
-saldo does not contain forms that aren't used gf generates all forms. A form
-needed to generate the gf table may not be given in saldo such as singular
-forms for 'byxor'. Also contains idioms, 'hålla huvudet kallt' and has
-different grammatical analyse, a lot of pronouns. Example.
-
-In Saldo there is also semantic info, so
-We need the numbers from saldo if there are to lemmas with the same
-identifier. växa (bli större), växa (vara en planta) vs. sluta\_V and sluta2\_V
-(slöt). However gf does not want more than one table
-if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
-We are just interested in morphology, so far. Dont want too big lexicon.
-
-The new version have also been made more robust, print logs and list of all
-words that wasn't imported together with a reason. Reson could be that it
-cannot construct all forms or that the word class shouldn't be in the gf lexicon.
-Costs very much memory, the saldo file is big, the generated lexicon is big.
-Saldo is divided to save memory. Errors should just be reported, but if something oförutsett
-happens and one part fail, the separation ensures that the others wont, also
-possible to restart. Problems with naming convention, special
-characters (å,ä,ö,apostrop,entrecôte.) needs to be translated , but avoid
-classhes 'kaeltisk' 'kältisk', gf will crash by this, so we add a 1
-in the end whenever we have changed a letter.
-
-Other word classes we do not import and why.
-Those should not be in the standard gf lexicon but maybe somewhere else.
-The list can be used for adding other kinds of lexicons - idioms, have pos-tags.
-Also to analyse what's wrong, or use the words with a word guesser. The code
-has been adapted for this.
-For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
-category. How this works, look at tags. Small amount, can be done manually.
-
-The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
-have any information. The earlier explained techniques may be used for this in Talbanken.
-To get even bigger material, use Korp, but then no guarantee.
+The tables does not overlap completely, there are some more forms in Saldo's table (eg. the 
+compounding form ``väx-")
+while the one generated in GF contains some that are not in Saldo (eg. ``vuxits").
+Saldo may also contain variats: ``växt" and ``vuxit" are both supine form.
+As far as possible, the program makes up for this by only comparing the overlapping forms
+and only requiring that GF generatens one variant whenever alternatives are given. 
+The program continues to generate new lexicons as long as there are more
+paradigms to choose from and compare the results after each run.
+
+The program has been made more robust than the previous version, it prints
+logs and finally lists of the words that were not imported. The logs provides
+information about which paradigms that were used for each word and the result
+from comparing the inflection tables.
+During the porting of Saldo, some renaming of the lemmas was needed.
+Special characters like \emph{å,ä,ö} should not occure in GF function names.
+Those are translated into \emph{aa,ae,oe} respectively. 
+The Saldo identifier \verb|äta..vb| is thus turned into the GF identifier \verb|aeta_V|.
+Since there might be two lemmas that share the same
+identifier after the translation, a number is added to the name whenever a translation
+have been done: 
+\begin{tabular}{lll}
+kältisk & $\rightarrow$ & kaeltisk\_1 \\
+kaeltisk & $\rightarrow$ & kaeltisk \\
+entrecôte &$\rightarrow$ & entrecoote\_1 \\ 
+\end{tabular}
+
+%Problems with naming convention, special
+%characters (å,ä,ö,apostrop,entrecôte.) needs to be translated , but avoid
+%classhes 'kaeltisk' 'kältisk', gf will crash by this, so we add a 1
+%in the end whenever we have changed a letter.
+%,é
+%Costs very much memory, the saldo file is big, the generated lexicon is big.
+%Saldo is divided to save memory. Errors should just be reported, but if something oförutsett
+%happens and one part fail, the separation ensures that the others wont, also
+%possible to restart.
+
+%Explanation of the algorithm: map categories, try each paradigm, compare the table of forms
+%created. If several forms in saldo, gf should pick on of them. Saldo has forms for
+%compounds, gf does not. A grammar is written, the correct ones saved, the others 
+%retried. Try all paradigms which can be formed from saldo. 
+%In Saldo there is also semantic info, so
+%We need the numbers from saldo if there are to lemmas with the same
+%identifier. växa (bli större), växa (vara en planta) vs. sluta\_V and sluta2\_V
+%(slöt). However gf does not want more than one table
+%if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
+%We are just interested in morphology, so far. Dont want too big lexicon.
 
@@
 \subsection{Results}
-How many words, how many Talbanken words are not here now? Why?
+All words from Saldo were not be imported. There are a number of reasons
+for this, the most obvious one is that we do not want all categories from
+Saldo in the GF lexicon. Prepositions(\emph{pp}) and numerals(\emph{nl}) etc. are
+assumed to be known already,
+and are usually not kept in the standard lexicon, but in modules for the basic
+structures of the language. Categories involving several words(\emph{vbm,nnm})
+need to be handeld differently, usually as idioms. Pronouns(\emph{pn}) are not  
+analysed the same way, see section \ref{sec:swegf}. Some experiments on finding
+the right GF-category for those have been done, see \ref{sec:gf.quant}.
+Six categories were imported: \\
+\begin{tabular}{l|lll}
+& Saldo tag & GF category & Example \\
+\hline
+ Adverb & \textbf{ab} &\textbf{Adv} & ofta \\
+ Adjective&\textbf{av} &    \textbf{A} & gul\\
+ Noun & \textbf{nn} &\textbf{N} & hus\\
+ Verb & \textbf{vb} &\textbf{V} & springa\\
+ Reflexive verbs  &\textbf{vbm}& \textbf{V2} & raka sig\\
+ Verbs with particles &\textbf{vbm}& \textbf{V}  & hitta på\\
+\end{tabular}\\
+
+%
+%There are a number of differences between and saldo Till skillnad från GF,
+%saldo does not contain forms that aren't used gf generates all forms. A form
+%needed to generate the gf table may not be given in saldo such as singular
+%forms for 'byxor'. Also contains idioms, 'hålla huvudet kallt' and has
+%different grammatical analyse, a lot of pronouns. Example. 
+One reason for why words of the categories listed above could not be imported
+is that Saldo does not contain information about word forms that are not used.
+In some cases those forms are needed by the smart paradigm. 
+This was a common reason why nouns could not be imported. 
+Consider for example
+the plural tantum %\cite[44]{SAG}
+noun ``glasögon" (``glasses"). Since the smart paradigm
+requires a singular form, this lemma could not be added to the lexicon. ``Glasögon" could
+be added manually, by using the somewhat incorrect singular form ``glasöga" (``glass-eye").
+The same problem occured for verbs that always are used in passive form:
+``andas" (``breathe"), ``frodas" (``flourish"), ``lyckas" (``succeeed"). This
+was the case for close to 80\% of the failing verbs of type \verb_vb_.
+
+In some cases the smart paradigms could not generate the correct declination,
+although this happend seldom. % frysa,fisa passive preterium
+
+The resulting dictionary contains more than 100 000 entries, compared to the total size
+of Saldo this is approximately 80\%. When testing the coverage of Talbanken,
+we found that 3058 words were still missing. In addition, there were 1817 words
+which were given different labels in GF than in Talbanken. Many of those are
+acceptable, such as the ones in table \ref{tab:saldodiff} and reflects the difference 
+made in the analyses, but others are examples of words that are still missing from
+the lexicon.
+\begin{figure}[h]
+\begin{tabular}{lll}
+Word & Talbanken tag & GF category \\
+\hline
+måste & MVPS          & VV \\
+allting & POTP & N \\
+få & POZP & Det \\
+\end{tabular}
+\caption{}
+\label{tab:saldodiff}
+\end{figure}
+Among the words that were completely missing, more than 2000 was found to be nouns,
+many of which were compounds.
+
+The lexicon does not contain any valency information since Saldo does not give any such
+information. Valencies are however crucial for GF, and by using the results from the mapping
+(see section \ref{sec:Mapping}) it should be possible to extract this information about this.
+%The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
+%have any information. The earlier explained techniques may be used for this in Talbanken.
+%To get even bigger material, use Korp, but then no guarantee.
 
@@
 \label{sec:prog.grammar}
-Although we can't expect to get a full coverage, we would like a tillfredsställande
-Explain what's in the resources, constructions present in most languages. 
-Been working on a grammar fragment based on Talbanken,
-like cool people like Montague. May also be interesting for other purposes.
-Swedish shares with Nor and Dan. Extra module, covering things like sw is a
-v2-lang, focusing parts of sentences.
+Although we can't expect to get a full coverage, we aim for a high-quality 
+grammar with satisfactory coverage.
+We have been working on a grammar fragment which have been adapted to cover
+constructions used in Talbanken. The starting point has been the Swedish
+resource grammar and the new grammar should still be compatible with this.
+A description of the Swedish implementation of the resource grammar is given
+in the next section, while section \ref{sec:Added} shows and explains the
+development during this project.
+%Explain what's in the resources, constructions present in most languages. 
+%Been working on a grammar fragment based on Talbanken,
+%like cool people like Montague. May also be interesting for other purposes.
+%Swedish shares with Nor and Dan. Extra module, covering things like sw is a
+%v2-lang, focusing parts of sentences.
 
 \subsection{The Swedish resource grammar}
+\label{sec:swegf}
 The GF resource grammars contains a fundamental description of Swedish,
@@
 {Har& han& inte& ätit& de & gula& äpplena& idag?}
-{Has& he& not& eaten& the& yellow& apples& today?}} %correct!!!
+{Has& he& not& eaten& the& yellow& apples& today?}} 
 \label{gfSwe:parseable}
@@
 \includegraphics[width=70mm,height=70mm]{apples.pdf}
-\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}%correct
+\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}
 \label{gfSwe:parsetree}
@@
 \subsection{Added}
+\label{sec:Added}
 Summary of what have been added and motivation.
@@
 \subsubsection{Quantifiers, determiners and predeterminers}
+\label{sec:gf.quant}
 The GF analyse differs sometimes. Experiments of how to extract
@@
 \section{Results and discussion}
-Results from mapping, can probably be improved, bit by bit, depends much on words.
-Valencys would therefor help.  
-Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
-results. Mostly compound nouns, which we can't expect in the dictionary.
-Is too slow, can't be used for parsing right away
-
-Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
-in the process to verify the solutions. Testing trees against talbanken.
-The big lexicon makes it very slow, eats all memory
-
-Discussion of the results and methods, why are the results good, why are they not better.
-What other methods could have been used? What did we/I expect and what happend?
+%Results from mapping, can probably be improved, bit by bit, depends much on words.
+%Valencys would therefor help.  
+%Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
+%results. Mostly compound nouns, which we can't expect in the dictionary.
+%Is too slow, can't be used for parsing right away
+%
+%Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
+%in the process to verify the solutions. Testing trees against talbanken.
+%The big lexicon makes it very slow, eats all memory
+%
+%Discussion of the results and methods, why are the results good, why are they not better.
+%What other methods could have been used? What did we/I expect and what happend?
 
@@
 What is still needed in the grammar, how important that is. 
-\subsection{Lexicon/Parallell parsing}
-Speed up.
+\subsection{Parallell parsing/Chunk parsing}
+\subsection{Lexicon}
+%Speed up!
+%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
+%Also to analyse what's wrong, or use the words with a word guesser. The code
+%has been adapted for this.
+%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
+%category. How this works, look at tags. Small amount, can be done manually.
+
 \subsection{Valency}
 \label{sec:futureValency}
-how to get valency information 'sitter och läser',
-is this doable for swedish?
-description of extracting this from talbanken/korp
-or handling it by the robust parser or the grammar itself (leaveout obj?).
-make the parser faster by using smaller lexicon.
-Lexicon tool on the fly.
+%how to get valency information 'sitter och läser',
+%is this doable for swedish?
+%description of extracting this from talbanken/korp
+%or handling it by the robust parser or the grammar itself (leaveout obj?).
+%make the parser faster by using smaller lexicon.
+%Lexicon tool on the fly.
+
 \subsection{Probabilites}
 \label{sec:futureProbabilities}
-Getting and using probabilities, to disambiguate.
-process list from saldo.
-Use multiligual treebank?
-stop ambiguities by fixing so that Gen of a quant isn't a quant, or better, that all fields
-have a genitive field.
-More fields in the grammar? Or leave it all to the robust parser
+%Getting and using probabilities, to disambiguate.
+%process list from saldo.
+%Use multiligual treebank?
+%stop ambiguities by fixing so that Gen of a quant isn't a quant, or better, that all fields
+%have a genitive field.
+%More fields in the grammar? Or leave it all to the robust parser
 
 \section{Conclusion}
-There is obviously a lot do still, will never get finish with NPL.
-Grammar may need even bigger changes, to allow enough but not too much.
-Different from writing a grammar for generation.
-Results from mapping shows that the translation is doable, but becomes harder
-since talbanken don't have formal rules.
-The saldo shows how easy to make use of utomstående resources, promising if
-we want to extract other information. 
-Hard, but not finished!
+%There is obviously a lot do still, will never get finish with NPL.
+%Grammar may need even bigger changes, to allow enough but not too much.
+%Different from writing a grammar for generation.
+%Results from mapping shows that the translation is doable, but becomes harder
+%since talbanken don't have formal rules.
+%The saldo shows how easy to make use of utomstående resources, promising if
+%we want to extract other information. 
+%Hard, but not finished!
 

### 2011-12-05 doc/report/report.tex
dependent types, comments on report

@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 % Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc
-
+% Funding!
+
+% check references! lncs is good maybe. lnai
 \newpage
@@
 \section{Introduction} 
+>>This work is part of bigger project, spring and next spring.
+>>Connection between grammar & parser: GF, every grammar defines a parser.
 Ultimate goal: want computers to 'understand' language. Not good with flat strings.
-We need to translate strings of natural language to a richer sturture,
+We need to translate strings of natural language to a richer structure,
 a parse tree. So far there is no
-freely availbe ruled-based parsing giving a deep analyse for Swedish.
+freely available ruled-based parsing giving a deep analyse for Swedish.
 %There are parsers which put only put part-of-speach tags to
@@
 %What is parsing, flat strings to trees, example picture, what is deep parsing.
-So far stable and good when working with controlled languges, but recent experiments
+So far stable and good when working with controlled languages, but recent experiments
 have showed that it is also possible to parse free language.
 To do this for Swedish, we need to  .. the resource grammar needed to covering
-the language specific constructions. Further a lagre lexicon is also needed, as well
+the language specific constructions. Further a large lexicon is also needed, as well
 as a method for evaluation.
@@
 \section{Background}  
+>>This work is part of bigger project, spring and next spring.
+>>Connection between grammar & parser: GF, every grammar defines a parser.
 To build a parser or a grammar from scratch is a heavy work. The idea for this
@@
 examples, ideas and test material from a treebank.
-The project is hence heavily dependening on three resources, which will be described
+The project is hence heavily depending on three resources, which will be described
 in the following sections.
@@
 Swedish\cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
-closely related to Norweigan and Danish. The languages share most of their
+closely related to Norwegian and Danish. The languages share most of their
 grammatical structures and are mutual intelligibility. Swedish is also 
 an official language in Finland and spoken by approximately 10 million people.
-Swedish syntax are often similair to English, but have some characteristics:
+Swedish syntax are often similar to English, but have some characteristics:
 \subsubsection*{Verb second}
@@
 \subsection{Saldo}
+>> syntactic information too. Mention that GF does not have this, only lexical or
+>> semantic. Not 1-1.
 A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
@@
 120 000 modern Swedish words.
-The words are equipped with both sematic and a morphological
+The words are equipped with both semantic and a morphological
 information. The user can find compounding analyses, examples of usage in corpora,
-and graphs of sematically connected words.\\
+and graphs of semantically connected words.\\
 \newpage % only until saldo can be on same page as the pic
@@
 The word ``växa" is however declined the same way for both interpretations of the word.
-Hence Saldo's morpological lexicon, provided under LPGL in XML format,
+Hence Saldo's morphological lexicon, provided under LPGL in XML format,
 only keeps one entry of ``växa".
@@
 
-For our purpose, only the morphological lexcion was needed.
+For our purpose, only the morphological lexicon was needed.
 The data can be processed and translated to GF format as described in 
@@
 \subsection{Grammatical Framework}
+>> more here. First: grammar formalism with Martin-Löf.
+>> snd: programming language, type checking and logic: prolog (check this)
+>> dependent and implicit types : agda. (read gf book and krasses book)
+>> pgf\ref{}, runtime system with parsing, lin, random generation osv.
 The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
 a functional grammar formalism based on Martin-Löf type theory.
-It has been developed for about 20? years and are inspired by programming langauges
+It has been developed for about 20? years and are inspired by programming languages
 like Haskell\cite{haskell},
@@
 
+>> The most important!!
 An important feature of GF is
 %Being developed with multilinguality in mind, GF makes an important 
-the distiction between
+the distinction between
 \textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
@@
 abstract. \verb|N| now means a noun
-in any of the number, and we introduce two funciton, \verb|NSg : N -> NP| and
-\verb|NPl : N -> Pl| which sets the number.
+in any of the number, and we introduce two function, \verb|NSg : N -> NP| and
+\verb|NPl : N -> NP| which sets the number.
 \begin{figure}[h!]
@@
 
+>>not very restricted, not few ambiguities.
+>>the scope is controlled language, but possible ambiguities (discuss this in future
+>>work, saldo, conclusion). Ambiguities: from multilinguality, or inherited:
+>>syntactical (man with the telescope), 
+>>lexical (youSg/Pl, växa).
 A GF grammar usually covers a very restricted set of language where few ambiguities
@@
 ensured. Some examples of projects using GF are WebAlt\cite{webalt} and something else.
+>> key, talk, hats, bom (see references in Ramonas thesis).
 %, the Online Syllogism Solver\cite{malin!}. 
 The framework is also used in the
-european project MOLTO\cite{molto} for online translation between upto 15 language.\\
+European project MOLTO\cite{molto} for online translation between up to 15 language.\\
 
 \subsubsection{The resource library}
+>>Best part? Biggest library?
+>>Most general resource, first large grammar, covers fundamental syntactics.
 The resource library\cite{gf-resource} is a very useful part of the framework. It is provided in the 
@@
 they all share the same abstract syntax.
-There are also a small test lexicon included, containing some hundered of common words.
+There are also a small test lexicon included, containing some hundred of common words.
 
@@
 flick\textbf{an}    &    hjärt\textbf{at}  \\
-flic\textbf{kor}    &    hjärt\textbf{an}  \\
-flic\textbf{korna}  &    hjärt\textbf{ana} \\
+flick\textbf{or}    &    hjärt\textbf{an}  \\
+flick\textbf{orna}  &    hjärt\textbf{ana} \\
 \hline
@@
 \vspace{5mm}
-As an open source-project, GF is constantely being developed and improved. New
-languges are added, the compiler is being improved, ways of använad det på ett
-lätt sätt are added and the possiblities to use GF in different environments
+>> references here! to everything. Remove uncertain, add ontology, aarne paper about
+>> donkey, find aarnes paper in ramona's thesis (french), hats formal methods,
+>> bom.
+As an open source-project, GF is constantly being developed and improved. New
+languages are added, the compiler is being improved, ways of använad det på ett
+lätt sätt are added and the possibilities to use GF in different environments
 increased. There are  research about how to make better use of the dependent
@@
 compare them against a reliable source.
-Being an freely avaible, manually annotade, large scale treebank,
+Being an freely available, manually annotated, large scale treebank,
 Talbanken\cite{talbanken} was perfect for our purpose.
@@
 \section{Related work}
+>> Much more here, important!! comparisons.
 Computational ling is active area for Swedish, there are many other Swedish parser,
@@
 Many of the parsers not ruled based, the most known one Malt\cite{malt} ... .
-However has been two, about CLE\cite{cle} and what happend to it. Logic and translation
+However has been two, about CLE\cite{cle} and what happened to it. Logic and translation
 to Swedish, The spoken lang. translator\cite{cle2}. Wiren \cite{wiren}, cassSwe \cite{casswe}
@@
 but the code needed updates to not miss many important words.
-Experiments have been made to use a interactive tool for lexical aquisation,
-but the dictionary was still much to small. The aquisation tool may still
-be used for complenting the lexicon.
+Experiments have been made to use a interactive tool for lexical acquisition,
+but the dictionary was still much to small. The acquisition tool may still
+be used for complementing the lexicon.
 
@@
 As far as possible, the program makes up for this by only comparing the overlapping forms
-and only requiring that GF generatens one variant whenever alternatives are given. 
+and only requiring that GF generates one variant whenever alternatives are given. 
 The program continues to generate new lexicons as long as there are more
@@
 During the porting of Saldo, some renaming of the lemmas was needed.
-Special characters like \emph{å,ä,ö} should not occure in GF function names.
+Special characters like \emph{å,ä,ö} should not occur in GF function names.
 Those are translated into \emph{aa,ae,oe} respectively. 
@@
 structures of the language. Categories involving several words(\emph{vbm,nnm})
-need to be handeld differently, usually as idioms. Pronouns(\emph{pn}) are not  
+need to be handled differently, usually as idioms. Pronouns(\emph{pn}) are not  
 analysed the same way, see section \ref{sec:swegf}. Some experiments on finding
@@
  Verb & \textbf{vb} &\textbf{V} & springa\\
- Reflexive verbs  &\textbf{vbm}& \textbf{V2} & raka sig\\
- Verbs with particles &\textbf{vbm}& \textbf{V}  & hitta på\\
+ Reflexive verbs  &\textbf{vbm}& \textbf{V} & raka sig\\
+ Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till\\
 \end{tabular}\\
@@
 be added manually, by using the somewhat incorrect singular form ``glasöga" (``glass-eye").
-The same problem occured for verbs that always are used in passive form:
+The same problem occurred for verbs that always are used in passive form:
 ``andas" (``breathe"), ``frodas" (``flourish"), ``lyckas" (``succeeed"). This
@@
 In some cases the smart paradigms could not generate the correct declination,
-although this happend seldom. % frysa,fisa passive preterium
+although this happened seldom. % frysa,fisa passive preterium
 
@@
 \subsection{The Swedish resource grammar}
+>> no subjects or object, analyse is more from the lexicon,
+>> compare to others, ask elisabet.
 \label{sec:swegf}
@@
 covering the morphology and many phrase and clause constructions. Most of the implementation
-is shared with the other scandinavian languages.
-Apart form descriping morphology, the resource grammar deals with
-word order, agreement, tempus, basic conjunction etc.
+is shared with the other Scandinavian languages.
+Apart form describing morphology, the resource grammar deals with
+word order, agreement, tense, basic conjunction etc.
 \enumsentence{\shortexnt{8}
@@
 Vem måste jag akta mig för? & För vem måste jag akta mig?\\
-Who do I need to watch out for? & For whom do I need to wath out?\\
+Who do I need to watch out for? & For whom do I need to watch out?\\
 \end{tabular}}
@@
 Before describing what has been added, will give a short introductions to the
-the cats of resource grammar and the sw ipmlementation in particular.
+the cats of resource grammar and the sw implementation in particular.
 
@@
 To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
-but allow variantion in.. 
+but allow variation in.. 
 What is the difference? Some words on differences for pronouns.
 To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
-Example, use vartiants {}.
+Example, use variants {}.
 There are also Quants, which .... 
@@
 
-kommer att - for scand, extra tempus, example
+kommer att - for scand, extra tense, example
 example table
@@
 \subsubsection{Passive}
+>> why not VPSlash here??
 There are two ways of forming passive verb phrases in Swedish: the 
@@
 suggest that it is used in more than 80\% of the times \cite{laanemets}.
-It is not as common in the other scandinavian languages however, 
-and it can not be used in all tenses for all words. The norweigan 
+It is not as common in the other Scandinavian languages however, 
+and it can not be used in all tenses for all words. The Norwegian 
 translation of both sentence \ref{gfPass:peri-pass} and sentence  \ref{gfPass:s-pass} is
@@
 This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
-turned into comlpete verb phrases; they no longer need an object.
+turned into complete verb phrases; they no longer need an object.
 The s-passive has now been added as the standard case for the Swedish GF
-grammar. Periphrastics passive is still allowed, but as an alternative rather
+grammar. Periphrastic passive is still allowed, but as an alternative rather
 than the default.
@@
 The function combines a verb phrase a determiner and a noun phrase in to a clause.
-If the determiner requires a noun in determinened form, the clause is put to NONEXIST.
+If the determiner requires a noun in determined form, the clause is put to NONEXIST.
 However, this is not an ideal solution. There are other examples of when we
-need to know the definitness of the noun phrase, for example when using
+need to know the definiteness of the noun phrase, for example when using
 the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
@@
 The GF analyse differs sometimes. Experiments of how to extract
-this information from Talbanken by using Saldo as a starting poin.
+this information from Talbanken by using Saldo as a starting point.
 One of the biggest differences between the GF analysis and the one made in Talbanken
@@
 
-\textbf{finitness}
+\textbf{finiteness}
  'de flesta de gula hästarna'
@@
 but have other words like 'bara' which can be put in other positions.
-  but needs to restruct whole of VP, since no field is befor fin.
+  but needs to reconstruct whole of VP, since no field is before fin.
   Can also act before the noun or as AdV. Examples.
@@
 \textbf{antalet}
-Also needed categories for 'pronouns' wich can act in different ways.
+Also needed categories for 'pronouns' which can act in different ways.
 and 'antal','sorter' - need indetermined, antal plural, sorter either.
-Should be able to modify by a godtyckligt antal adjective, and they must be determined.
+Should be able to modify by a arbitrary antal adjective, and they must be determined.
 '(ett stort antal) människor' and not 'ett stort (antal människor)'
@@
 the possessive pronouns as described in section \ref{swe:refl}.
-Those make up noun phrases wich are not allowed in subjects, 
+Those make up noun phrases which are not allowed in subjects, 
 and we want to be able to parse to following sentences:
+
+>> remove stupid examples, add 'hon och sin vän åkte tåg'
 \enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
@@
 
-The reflexive pronouns require another subject whith which they agree in
+The reflexive pronouns require another subject with which they agree in
 number and person. Apart from this restriction, they may be used as normal
@@
 %Could be combined with 'sig', a ReflNP without another noun.
-In short, what is wanted can be summerized as follows:
+In short, what is wanted can be summarized as follows:
 \begin{figure}
@@
 Further, the dependence on the subject needs to be applied to adverbial phrases and
-adjective phrases as well. The adjetive phrase in sentence \ref{ex:ReflAP} and
+adjective phrases as well. The adjective phrase in sentence \ref{ex:ReflAP} and
 the adverbial phrase in sentence \ref{ex:ReflAdv} are examples of this; if the
@@
 \section{Mapping}
+>> no subjects or object, analyse is more from the lexicon,
+>> compare to others, ask elisabet.
 \label{sec:Mapping}
@@
 usage. One part of this project has focused on extracting this information
-and to translate it to GF notaiton. We have constructed a mapping, which 
+and to translate it to GF notation. We have constructed a mapping, which 
 automatically transforms trees in Talbanken format to GF abstract syntax trees.
-Figure \ref{fig:translationtrees} shows an exapmle of a visualized tree
+Figure \ref{fig:translationtrees} shows an example of a visualized tree
 from Talbanken05 and the result of translating it to GF.
@@
 
+>> add parse tree, and describe connection between parse and abstract
+>> show how we keep analyse
 \begin{figure}[h]
@@
 The mapping connects GF with another annotations and gives us means to evaluate
-our parser. By both parsing a talbanken sentence and transforming its annotaded tree, we can
+our parser. By both parsing a Talbanken sentence and transforming its annotated tree, we can
 easily inspect if the results are equal.
@@
 If there are words missing from our dictionary, the rich
-pos-tags may help us to automatically find the correct declination and add it to the
+POS-tags may help us to automatically find the correct declination and add it to the
 lexicon. Further, our parser will need probabilities, see section
@@
 
-A translation of the English Penn Treebank to the GF syntax has previosly been done \cite{gfpenn}.
-By starting this progam, we now have
+A translation of the English Penn Treebank to the GF syntax has previously been done \cite{gfpenn}.
+By starting this from program, we now have
 a translation that works for the quite differently annotated Talbanken. The differences between
@@
 The categories give syntactic information such as \verb|S|,\verb|NP| and \verb|VP|.
-The edge labels give similair information with tags as \verb|SS| for subject, 
+The edge labels give similar information with tags as \verb|SS| for subject, 
 \verb|OO| for object. There are more than 300 different pos tags. The high
 number is due to the detailedness, for example there are almost 50 tags for
-nouns, excluding proper names. The tags show definitiness, case, whether the
+nouns, excluding proper names. The tags show definiteness, case, whether the
 word is a compound. Some words also have their own tags.\\
@@
 
-The translation works sectionwise, ie, if it finds a tree of category \verb|S|,
+The translation works section-wise, ie, if it finds a tree of category \verb|S|,
 it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
@@
 During the mapping process, the translator also needs to find out whether the sentence is
-negated and what tempus is used. If it finds two trees of type, say, \verb|VP| but does not
+negated and what tense is used. If it finds two trees of type, say, \verb|VP| but does not
 have any rules for combining them, the \textit{meta function} \emph{'?'} will be used. 
-Figure xx shows an exapmle of this.\\
+Figure xx shows an example of this.\\
 
@@
 The structural differences between GF and Talbanken cause some translation problems.
-On example is that Talbanken has no valencey information but for GF this is the key when
+On example is that Talbanken has no valence information but for GF this is the key when
 constructing for example verb phrases. The grammar bestämmer that a transitive \verb|V2|
@@
         \emph{They are appreciated}}\label{ex:mappMan} %  (s216).
-In talbanken, \emph{man} is simply a personal pronoun, 
+In Talbanken, \emph{man} is simply a personal pronoun, 
 in GF it causes the whole clause to get the form \verb|GenericCl|, which has no
@@
 %förklara bättre eller ta bort??
-A similair issue arises form the use of verb particles. If a verb with a particle is considered
+A similar issue arises form the use of verb particles. If a verb with a particle is considered
 as one klump in GF, which means that the particle has no effect on the abstract tree. Its presence
@@
 will be among the results. The first option will be selected, and if the
-particle isn't directly attatched to the verb
+particle isn't directly attached to the verb
 %There is, at this point, no easy way of knowing which one of them wants a particle and the
@@
 
-They occure quite frequently in the treebank and are always translated
+They occur quite frequently in the treebank and are always translated
 to \emph{?}.
@@
 
-\subsection{Results and evalutaion}
-The flat version of Talbanken05 was used for developping the mapping, but
+\subsection{Results and evaluation}
+>> more! replace words, how much better results (both grammar and mapping) 
+>> more results
+The flat version of Talbanken05 was used for developing the mapping, but
 the program works for both versions.
@@
 Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
-which can group an infinite verb and an object abverbial together into a \verb|VP|.
+which can group an infinite verb and an object adverbial together into a \verb|VP|.
 This information can however be extracted from the flat version, and the results
@@
 Shorter sentences have been prioritized in order to get a coverage of the most
-fudamental constructions. Further,  
+fundamental constructions. Further,  
 regression testing was used during the development. % in order to  very easy to destroy what was working before.
@@
 No list items & 65\%\\
-No special tags or punctiation & 72\%\\
+No special tags or punctuation & 72\%\\
 Short sentences with known words & 85\%\\
@@
 \label{sec:future}
+>> divide into consequences and future plans.
 What is still left to do? Description and direction of future work:
 Handling unknown words and , handling unknown grammar, 
-how to make the parser robust, elipses and long sentences, names and numbers.
+how to make the parser robust, ellipses and long sentences, names and numbers.
 What is still needed in the grammar, how important that is. 
-\subsection{Parallell parsing/Chunk parsing}
+\subsection{Parallel parsing/Chunk parsing}
 \subsection{Lexicon}
@@
 
-\subsection{Probabilites}
+\subsection{Probabilities}
 \label{sec:futureProbabilities}
@@
 \section{Conclusion}
+>> why is this significant, other grammars can use this Swedish grammar.
+>> well known (comp. ling) trade-off between depth & coverage. for now depth, robustness in in future
+>> work leads to coverage.
 %There is obviously a lot do still, will never get finish with NPL.

### 2011-12-05 saldo/Saldoer.hs

@@ -30,7 +30,6 @@ Translates saldom.xml into GF dictionary.
 To change input format, just replace parseDict
 The main fuction is extract, which could operate alone or together with
 SaldoMain
-Based on code by Krasimir.
 -----------------------------------------------------------------------------}
 
 inputFile = "lillsaldo.xml"
@@ -68,7 +67,7 @@ extract select name inputFile n  = do
   mst <- runErrorT $ execStateT (createGF saldo
                     >> compileGF
                     >> loop
-                    -- >> printGFFinal
+                    >> printGFFinal
                     >> cleanUp)  (initState n select name)
   er <- case mst of 
              Left er  -> return er
@@ -126,7 +125,7 @@ findGrammar (id,E pos table) =  do
 okCat "V2" = hasPrep  
 okCat "VR" = isRefl
 okCat _    = const True
-findA w | part `elem` preps = ")\""++part++"\"" --first paranthesis to close mkV
+findA w | part `elem` preps = ")\""++part++"\"" --paranthesis to close mkV
         | otherwise         = ""
  where part = findsndWord w
 
@@ -209,7 +208,7 @@ checkForms paramMap fm_t gf_t entry@(G id cat lemmas _ _  _)
         -- to do: if the comparative forms for an adjective doesn't exist, add compounda
            then do report ("all forms for "++id++" not found" )
                    getNextLemma (G id cat lemmas b f ps)
-           else replace (G id cat [catMaybes forms] a (specialF pre f) ps)
+           else replace (G id cat [catMaybes forms] a {-(specialF pre-} f ps)
       where
         specialF "mk3A" _ = ("mk3A","") --- to be done more nicely -- can probably be removed now, but test first
         specialF x           f = f
@@ -252,8 +251,10 @@ compileGF = do
 -------------------------------------------------------------------
 
 mkGFName :: String -> String -> String
-mkGFName id' cat = name++"_"++cat
+mkGFName id' cat = name++"_"++toGFcat cat
   where
+       toGFcat "VR" = "V"
+       toGFcat  v   = v
        dash2us '-' = '_'
        dash2us x = x
        num x = if isDigit (head' "isDigit" x) then 'x':x else x
@@ -263,7 +264,7 @@ mkGFName id' cat = name++"_"++cat
               $ map dash2us 
[…]

### 2011-12-06 doc/report/report.tex
documentation

@@
 Other grammar frameworks.
+HPSG\cite{hpsg} and lingo matrix\cite{lingomatrix}, English, Japanese, German, Greece, Spanish, Norweigan.
+also Swedish: BiTSE\cite{stymne}, Stymne 2006, Swedish English translation, focusing on verb phrames, built on HPSG.
+S\o rgaard and Haugereid, Scandinavian Grammar Matrix\cite{scandmatrix}
+SweCore ??
+Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English: ongoing to get widecoverage grammar
+under GPL. Translation to Korean.
 % related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,

### 2011-12-06 doc/report/report.tex
references in report

--- 2011-12-05 doc/report/report.tex
@@
 >>Connection between grammar & parser: GF, every grammar defines a parser.
+>> reusablilty, freely available.
+>> the big amount of data, need language technology, machine translation
 Ultimate goal: want computers to 'understand' language. Not good with flat strings.
@@
 second constituent of a declarative main clause must consist of a verb.
-The normal word order is subject-verb-object, but any syntactic category can be fronted\cite[\textsection 1027]{H&H}.
+The normal word order is subject-verb-object, but any syntactic category can be
+fronted\cite[\textsection 1027]{H&H}.
 Topicalisation is very common, especially for temporal and spatial adverbial phrases.
@@
 The word order in subordinate clauses in also slightly different,
+>> and stricter
 %\enumsentence{
@@
 >> more here. First: grammar formalism with Martin-Löf.
->> snd: programming language, type checking and logic: prolog (check this)
+>> snd: functional and programming language, type checking and logic: prolog (check this)
 >> dependent and implicit types : agda. (read gf book and krasses book)
@@
 By using a controlled language like this,  good and reliable translation can be
-ensured. Some examples of projects using GF are WebAlt\cite{webalt} and something else.
+ensured. Some examples of projects using GF are WebAlt\cite{webalt},
+a project which aims to develop language independent material for mathematics problems,
+and the formal software verification tool KeY\cite{key}, in wich GF is used for translating specification
+to natural language.
 >> key, talk, hats, bom (see references in Ramonas thesis).
@@
 types, for solving syllogism or generating natural language via Montague
-semantics.
-Further, experiments on using GF for free parsing has been conducted \cite[\textsection 4.1]{gfMech} 
+semantics\cite{montague}.
+Further, experiments on using GF for free parsing has been conducted
+\cite[\textsection 4.1]{gfMech} 
 This is also a future goal for our project and one of the delmål has been to
@@
 Being an freely available, manually annotated, large scale treebank,
+>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
+>> for large scale projects.
 Talbanken\cite{talbanken} was perfect for our purpose.
@@
 However has been two, about CLE\cite{cle} and what happened to it. Logic and translation
-to Swedish, The spoken lang. translator\cite{cle2}. Wiren \cite{wiren}, cassSwe \cite{casswe}
+to Swedish, S-CLE (Gambäck,1997) full analyse but coverage is too limited (nivres papper), The
+spoken lang. translator\cite{cle2}. Wiren \cite{wiren}. bit more information than
+just pos: cassSwe \cite{casswe}
 Other Sw. parsers and grammars, evaluation of them? 
 Other grammar frameworks.
+SweCG (birn) annotation, parser Swedish FDG (voultanien,2001) which is not freely available.
+finding subject, objects, suubject complements (SC) and
+to link them to their proper regents (main verbs) : over 90\% both precision and recall.
+uses a dependency grammar, but "The functional description of adverb phrases
+and prepositional phrases (e.g. agent, source, goal, benefactive, time) remains
+to be described in a future version." 
+HPSG\cite{hpsg} and lingo matrix\cite{lingomatrix}, English, Japanese, German,
+Greece, Spanish, Norweigan (NorSource, Hellan 2003),( not famous: SweCore
+(Ahrenberg) based on NorSource)
+also Swedish: BiTSE\cite{stymne}, Stymne 2006, Swedish English translation,
+focusing on verb phrames, built on HPSG.
+S\o rgaard and Haugereid, Scandinavian Grammar Matrix\cite{scandmatrix}
+Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English: ongoing to get widecoverage grammar
+under GPL. Translation to Korean.
 % related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,
@@
 tycka om /tycka om -> om vad tycker du inte. fast bra
-about slashes\cite{gazdar}
+'jag tar med mig den till Bo' : ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;
+but is really 'till' part of verb?
+VP is divided corresponding to satsschema (diderichsen) (renewed by SAG),
+otherwise use slash.
+"Speed up", instead of movement used slashes\cite{gazdar}.
+Slashes good for question (vilken katt ser du idag)
+relatives (katten som du tittar på) and topicalisation
+The grammar has VP,Cl, SSlash, all missing NP.
 

### 2011-12-06 saldo/Saldoer.hs
notes from counting talbanken/saldo

@@ -122,13 +122,16 @@ findGrammar (id,E pos table) =  do
                                               | (gf_cat,(f,f'),paradigms) <- xs]
                       modify $ \s -> s {retries = cnc++retries s}
 
-okCat "VP" = hasPrep  
+--- particles can be prepositions (hitta på), adverbs (åka hem), nouns (åka hem)... 
+--  we first check for reflexives, and accept the rest as particles.
+okCat "VP" = hasPart
 okCat "VR" = isRefl
 okCat _    = const True
 findA w | part `elem` preps = ")\""++part++"\"" --paranthesis to close mkV
         | otherwise         = ""
  where part = findsndWord w
 
+hasPart = all isAlpha . findsndWord
 hasPrep = (`elem` preps) . findsndWord
 isRefl  = (=="sig") . findsndWord
 
@@ -304,8 +307,8 @@ catMap  =
   , (pack "nn",   "N", map (first pack) nounParamMap, ("mkN",""), nounParadigmList)
   -- particles were V2. Why? -"dirV2 (partV (mkV",")"
   -- VR should not be V2 either.
-  , (pack "vbm", "VP", map (first pack) verbPParamMap, ("partV (mkV",""), verbPParadigmList)
   , (pack "vbm", "VR", map (first pack) verbRParamMap, ("reflV (mkV",""), verbRParadigmList)
+  , (pack "vbm", "VP", map (first pack) verbPParamMap, ("partV (mkV",""), verbPParadigmList)
   ]
   
 -- For prepositions, not run automatically

### 2011-12-07 doc/report/report.tex
notes from counting talbanken/saldo
report

@@
 %ska innehålla vettiga sökord
-\title{Title of the report} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Scaling up a GF grammar for free parsing} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 >>This work is part of bigger project, spring and next spring.
->>Connection between grammar & parser: GF, every grammar defines a parser.
+>>Connection between grammar and parser: GF, every grammar defines a parser.
 >> reusablilty, freely available.
 >> the big amount of data, need language technology, machine translation
+>> parsing: assingn of morph. and syntactic structures.
+>>Connection between grammar and parser: GF, every grammar defines a parser.
 Ultimate goal: want computers to 'understand' language. Not good with flat strings.
@@
 will not disappear, be made use of in other projects.
-What we could not cover by this project, what have been covered before (in
-the spring project), and why we choose those parts.
+%What we could not cover by this project, what have been covered before (in
+%the spring project), and why we choose those parts.
+To build a parser or a grammar from scratch is a heavy work and we
+therefore combine already existing parts.
 
 \section{Background}  
->>This work is part of bigger project, spring and next spring.
->>Connection between grammar & parser: GF, every grammar defines a parser.
-To build a parser or a grammar from scratch is a heavy work. The idea for this
-project is to combine already existing parts into a bigger and more expressive grammar
-as well as a large scale GF lexicon.
+The work described in this is thesis is part of a bigger project which aims at
+using GF for parsing free Swedish. 
+In previous work \ref{??} the start on extending the Swedish GF grammar was
+made and a tool for lexical aquistion were developed.
+We now construct a bigger
+and more expressive grammar as well as a large scale GF lexicon.
 As all GF grammars, this one can be used for parsing, and we develop it by getting
@@
 The project is hence heavily depending on three resources, which will be described
-in the following sections.
+in this section.
 
@@
 \subsection{Saldo}
->> syntactic information too. Mention that GF does not have this, only lexical or
->> semantic. Not 1-1.
 A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
@@
 120 000 modern Swedish words.
-The words are equipped with both semantic and a morphological
+The words are equipped with semantic, syntactic and a morphological
 information. The user can find compounding analyses, examples of usage in corpora,
@@
 \subsection{Grammatical Framework}
->> more here. First: grammar formalism with Martin-Löf.
+The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
+a grammar formalism based on Martin-Löf type theory\cite{martinlof}. GF is
+designed for multilingual applications and is equivalent to Parallel Multiple
+Context-Free Grammars (Seki \& al. 1991), it is hence stronger than mildly
+contex-free grammars.
+
+GF is a functional programming languages, inspired by programming languages
+like ML\cite{ml}, Haskell\cite{haskell}, Prolog\cite{bratko} and
+Agda\cite{agda}. 
+It has been developed for about 20 years 
 >> snd: functional and programming language, type checking and logic: prolog (check this)
 >> dependent and implicit types : agda. (read gf book and krasses book)
->> pgf\ref{}, runtime system with parsing, lin, random generation osv.
-The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
-a functional grammar formalism based on Martin-Löf type theory.
-It has been developed for about 20? years and are inspired by programming languages
-like Haskell\cite{haskell},
-ML\cite{ml}, Prolog\cite{bratko} and
-Agda\cite{agda}.
-Grammars written in GF can be used together with languages 
-like Java, Java script and Haskell and works for both parsing and generation.
-
->> The most important!!
-An important feature of GF is
+
+A grammar written in GF can be used for both parsing and generation.
+The GF package also provides various tools for working with and using the grammars:
+a compiler, an interpreter and a runtime system. The interpreter, the GF shell, lets
+you try and test your grammar by using commands for parsing, vizualization of parse trees,
+random generation, word alignment, morphological quizzes etc.
+The compiler lets you compile your grammar into portable grammar format (PGF)\ref{pgf}.
+There are libraries for using PGF files with Java, Java script and Haskell. The grammars
+may also be used in Android applications.
+and are inspired by.\\
+
+The key feature in GF is the distinction between
+\textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
+shared by many languages and can be implemented by a numeber of concrete grammars,
+each representing one language. The abstract syntax models the semantics of the
+domain without concern for language specific features such as agreement or word
+order.
+
 %Being developed with multilinguality in mind, GF makes an important 
-the distinction between
-\textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
-shared by many languages, it models the semantics of the domain without concern for
-language specific features such as agreement or word order.
-
 \begin{figure}[h]
@@
 a project which aims to develop language independent material for mathematics problems,
-and the formal software verification tool KeY\cite{key}, in wich GF is used for translating specification
-to natural language.
->> key, talk, hats, bom (see references in Ramonas thesis).
+and the formal software verification tool KeY\cite{key}, in wich GF is used for
+translating specification to natural language.
+%>> key, talk, hats, bom (see references in Ramonas thesis).
 %, the Online Syllogism Solver\cite{malin!}. 
@@
 \subsubsection{The resource library}
->>Best part? Biggest library?
->>Most general resource, first large grammar, covers fundamental syntactics.
-The resource library\cite{gf-resource} is a very useful part of the framework. It is provided in the 
-GF package. %, they give a good start for all who want to write an application grammar.
-Instead of letting the users of the framework reimplement the morphological and
-the fundamental syntactic structures of a language for each application, all
-this can be found as libraries. So far there are 20 languages in the library, and
-they all share the same abstract syntax.
-There are also a small test lexicon included, containing some hundred of common words.
+%>>Best part? Biggest library?
+%>>Most general resource, first large grammar, covers fundamental syntactics.
+The GF package provides a powerful resource library\cite{gf-resource}, which covers the
+fundamental morphology and syntactics for more than 20 languages. %is a very useful part of the framework. It is provided in the 
+%, they give a good start for all who want to write an application grammar.
+%of a language for each application, all
+%this can be found as libraries. So far there are 20 languages in the library, and
+The languages all share the same abstract and they constitute  
+a very useful resource when implementing application grammars.
+There is also a small test lexicon included, containing some hundred of common
+words.
 
@@
 
+A GF parse tree has two visualized representations, the \textit{parse tree} and
+the \textit{abstract tree}, as shown in figure \ref{fig:gfPandAtrees}.
+>> have other treees! One in english.
+\begin{figure}[h]
+\centering
+\subfloat[GF parse tree]{\label{pic:gfPtree}\includegraphics[width=50mm]{gfPTree.pdf}}
+\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=60mm]{gfTree.pdf}}
+\caption{Parse tree and abstract tree for the sentence ``Katten på bilen blir större".}
+\label{fig:gfPandAtrees}
+\end{figure}
+The abstract trees shows all functions GF needs to use
+to parse the sentence and is consequently the same for all concrete syntaxes.
+The parse tree (\ref{pic:gftree}) shows the words and which types the functions assign
+to them.
+The abstract tree shows us that the function
+\verb_DetCN_ is applied the \verb-UseN cat_N- and the information
+definitive (\verb-DefArt-) and singular (\verb-NumSg-). In the Swedish parse tree
+we see that \emph{``katten"}, which is in definite form singular, is noun \verb-N-
+which is turned into a \verb-CN- and finally a \verb-NP-. In the English parse
+tree, the form of \emph{``cat"} expresses singular while the definite article
+\emph{``the"} is explicit and therefore shown in the parse tree.
+
 \vspace{5mm}
->> references here! to everything. Remove uncertain, add ontology, aarne paper about
->> donkey, find aarnes paper in ramona's thesis (french), hats formal methods,
->> bom.
 As an open source-project, GF is constantly being developed and improved. New
-languages are added, the compiler is being improved, ways of använad det på ett
-lätt sätt are added and the possibilities to use GF in different environments
+languages are added, the compiler is being improved, ways of using it in more 
+effecient and easier ways are added
+and the possibilities to use GF in different environments
 increased. There are  research about how to make better use of the dependent
-types, for solving syllogism or generating natural language via Montague
-semantics\cite{montague}.
+types, for using ontologies\cite{ramona}(>>check!!) or generating natural
+language via Montague semantics\cite{montague}.
 Further, experiments on using GF for free parsing has been conducted
-\cite[\textsection 4.1]{gfMech} 
-This is also a future goal for our project and one of the delmål has been to
+\cite[\textsection 4.1]{gfMech}.
+This is also a future goal for our project and one of the subgoal has been to
 model an important part of Swedish in GF. The starting point was the resource
@@
 %Main ideas, abstract and concrete, translation, parse- and abstract trees.
-%Example of abstract.
 %We say that this is a function that takes as argument and returns..
@@
 Being an freely available, manually annotated, large scale treebank,
->> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
->> for large scale projects.
 Talbanken\cite{talbanken} was perfect for our purpose.
@@
 and was thus redistributed in a more modern version\cite{talbanken05} This work
-was done by Nivre, Hall and Nilsson in 2005. They released it in Malt and Tiger XML-formats,
+was done by Nivre, Hall and Nilsson in 2005. They released it in Malt and Tiger
+XML-formats\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html},
 where the trees had been made deeper and more detailed. \\
@@
 However has been two, about CLE\cite{cle} and what happened to it. Logic and translation
-to Swedish, S-CLE (Gambäck,1997) full analyse but coverage is too limited (nivres papper), The
-spoken lang. translator\cite{cle2}. Wiren \cite{wiren}. bit more information than
+to Swedish, S-CLE (Gambäck,1997) full analyse and semantics, translation,
+but coverage is too limited (nivres papper), The
+spoken lang. translator\cite{cle2}. Wiren \cite{wiren}?? chartparser, but for swedish. bit more information than
 just pos: cassSwe \cite{casswe}
@@
 also Swedish: BiTSE\cite{stymne}, Stymne 2006, Swedish English translation,
+and one for the common parts of scandinavian (Haugereid etc) in Nordisk sprog.
 focusing on verb phrames, built on HPSG.
@@
 \label{sec:prog.saldo}
+>> syntactic information too. Mention that GF does not have this, only lexical or
+>> semantic. Not 1-1.
 Saldo is compatible with GF, which makes it easy to transform the morphological
@@
 of Saldo this is approximately 80\%. When testing the coverage of Talbanken,
-we found that 3058 words were still missing. In addition, there were 1817 words
+we found that around 3000 words were still missing (excluding the one tagged as
+numerals and proper names, but some names are still there). In addition, there were 1817 words
 which were given different labels in GF than in Talbanken. Many of those are
@@
 
-\subsection{Added}
+\subsection{Development of the grammar}
 \label{sec:Added}
@@
 >> compare to others, ask elisabet.
+>> diffs: valencies, pp (gf ingen skillnad direkt/indirekt i träd, pga multiling)
+>> mamba very used and tested for swedish, standard.
 \label{sec:Mapping}
 Talbanken contains much valuable information about phrase structures and word 
-usage. One part of this project has focused on extracting this information
-and to translate it to GF notation. We have constructed a mapping, which 
+usage. It is analysed with the MAMBA annotation scheme (Telemann, 1974) which is 
+%>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
+%>> for large scale projects.
+%annotation scheme
+commonly used, and hence very well-tested, for describing Swedish.
+The deepened analyse in Talbanken05 includes the lexical MAMBA layer. 
+One part of this project has focused translating trees from Talbanken to GF.
+We hence get a comparision between the two annotaitons and at the same time
+we get ways to extract and translate information to GF notation. We have
+constructed a mapping, which 
 automatically transforms trees in Talbanken format to GF abstract syntax trees.
-Figure \ref{fig:translationtrees} shows an example of a visualized tree
-from Talbanken05 and the result of translating it to GF.
-Both the POS tags as well as the syntactic information are needed for the translation.
-
->> add parse tree, and describe connection between parse and abstract
->> show how we keep analyse
+Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
+of the sentence \emph{``The cat on the car gets bigger"} and the result of translating it to GF.
+During the translation all the information given in the abstract tree
+(\ref{pic:gftree}) needs to be extracted. We need to know that \emph{``katten"} is a noun
+used in definite form (\verb-DetArt-), singular (\verb-NumSg-). Even though the
+notation is changed by the transformation, the original analyse from Talbanken,
+that ``på bilen" forms a subtree which is attatched to ``katten" for example,
+is preserved.
+%produces a \verb-NP- from ``katten" this means that
+%DetCN : Quant                  -> CN    -> NP
+%DetCN  (DetQuant DefArt NumSg)    cat_N = katten
+Both the POS tags as well as the syntactic information are needed.
+
 \begin{figure}[h]
-\centering
+\hspace{-10mm}
 \subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
 \subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
+\subfloat[GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
 \caption{Trees for the sentence ``Katten på bilen blir större".}
@@
 By starting this from program, we now have
-a translation that works for the quite differently annotated Talbanken. The differences between
+a translation that works for the differently annotated Talbanken. The differences between
 Swedish and English syntax also meant that many rules had to be changed.\\
 
-The mapping works for Talbanken05 in Tiger XML-format. There are three kinds of
-tags: categories, edge labels and pos tags. 
-The categories give syntactic information such as \verb|S|,\verb|NP| and \verb|VP|.
+The mapping works for Talbanken05 in Tiger XML-format.
+There are three kinds of tags: categories, edge labels and POS tags. 
+The categories give syntactic information such as \verb|S|, \verb|NP| and \verb|VP|.
 The edge labels give similar information with tags as \verb|SS| for subject, 
@@
 \textbf{NNDDHSGG} & familjemedlemmarnas & Noun, definite, compound (person), genitive \\
-\textbf{PU} & *, 1), a) & List item  ...\\
+\textbf{PU} & * $\;$ 1) $\;$ a) & List item\\
 \end{tabular}\\
@@
 
-The translation works section-wise, ie, if it finds a tree of category \verb|S|,
+The translation works section-wise, ie. if it finds a tree of category \verb|S|,
 it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
@@
 It figure \ref{pic:tbtree}, the verb is followed by an adjective as a present
-participle, \verb|SP|. GF admits \emph{bli} to have type \verb|VA|, a verb
-with an adjective as complement, and the translator can apply the gf function
+participle, \verb|SP|. GF admits \emph{``bli"} to have type \verb|VA|, a verb
+with an adjective as complement, and the translator can apply the GF function
 \verb|ComplVA| to combine the verb with its object. \\
-It is not often the case that we can first find the subject, then the verb and 
+It is not always the case that we first find the subject, then the verb and 
 finally the objects. Often the order is Verb-Subject-Object or Object-Subject-Verb. \\
@@
 negated and what tense is used. If it finds two trees of type, say, \verb|VP| but does not
-have any rules for combining them, the \textit{meta function} \emph{'?'} will be used. 
+have any rules for combining them, the \textit{meta function} \verb-?- will be used. 
 Figure xx shows an example of this.\\
@@
 %Description of most important tags from Talbanken, their translation? appendix?
-The structural differences between GF and Talbanken cause some translation problems.
-On example is that Talbanken has no valence information but for GF this is the key when
-constructing for example verb phrases. The grammar bestämmer that a transitive \verb|V2|
-needs by two objects. The distinction is not always clear in Swedish where 
-most transitive verbs can also be used as intransitive. Both sentences in
-example \ref{ex:mappTransitive}.
-\enumsentence{Han sitter och läser.\\ Han sitter och läser en bok.} \label{ex:mappTransitive}
-See section \ref{sec:futureValency} for a longer discussion about this.
-For this reason, the mapping will allow words to be used in other ways than their
-GF valencies prescribes.
-
-Another problem can be instantiated by sentence \ref{ex:mappMan}, containing the generic pronoun \emph{man}.
-\enumsentence{Man uppskattar dem.\\
-        \emph{They are appreciated}}\label{ex:mappMan} %  (s216).
-In Talbanken, \emph{man} is simply a personal pronoun, 
-in GF it causes the whole clause to get the form \verb|GenericCl|, which has no
+\subsubsection{Differences in the notation}
+Some structural differences between GF and Talbanken were made obvious during the
+translation. \\
+%Since the translation is done word by word, ambiguities may confuse the translator.\\
+%shown during the translation.
+%cause some translation problems.
+In Talbanken the valency information is given implicitly by the complements
+of a word. If a verb is followed by an object, \verb-OO-, which contains a
+\verb-S-, we can conclude that this is a sentence-complemnt verb.
+In GF, the valency information is given for each entry in the lexicon.
+A sentence-complement verb has the type \verb-VS- and from this we know 
+that the function \verb-ComplVS- should be used to combine the verb and with a 
+sentence.
+A similar issue arises form the use of verb particles. A verb with a
+particle is considered as one constituent in GF, which means that the particle
+has no effect on the abstract tree. Its presence is announced by the verb
+itself, while Talbanken treats them as separated units.
+%In Talbanken, verbs and particles are treated as separate units,
+%and the particle is marked with the tag \verb-PL- and attached to. 
+%Talbanken has no valence information while this is crucial for GF.
+%The grammar requires a adjective-complement verb,
+%\verb|VA|, to be used with one adjectival phrase. For \emph{``bli"} in example
+%\ref{pic:tbtree}, we know that we need to find an adjective phrase.
+%%know what kind of verb we are looking for
+%%when consulting the GF lexicon.
+%If the GF lexicon assings different types to verb, the correct type 
+%should be showed by the complements.
+%This way %\enumsentence{Han gillar det\\He likes it} 
+The mapping cannot depend on the valencies given in GF only, however, 
+since the distinction between different types of verbs is not always clear in Swedish.
+The grammar rules becomes too strict for accepting most Talbanken sentences.
+For example, most transitive verbs can also be used as intransitive. Both sentences in
+example \ref{ex:mappTransitive} are grammatically correct.
+\enumsentence{
+\shortexm{11}
+{Han& sitter& och &läser.}
+{He& sits& and &reads.}{}
+{Han& sitter& och& läser& en& bok.}
+{He& sits& and& reads& a& book.}{}}
+\label{ex:mappTransitive}
+See section \ref{sec:futureValency} for a longer discussion about this.\\
+%the Talbankenwill allow words to be used in other ways than their
+%GF valencies prescribes.
+Another problem can be instantiated by sentence \ref{ex:mappMan}, containing
+the generic pronoun \emph{``man"}.
+\enumsentence{
+\shortex{3}{Man & uppskattar & dem}
+{One & appreciates & them}
+{``They are appreciated"}}\label{ex:mappMan} %  (s216).
+In Talbanken, \emph{``man"} is simply a personal pronoun, 
+in GF it is represented by using the function \verb|GenericCl|, which 
+turns a \verb-VP- into a clause and has no
 subject visible in the abstract tree. See figure \ref{fig:mappMan}.
@@
 \end{figure}
-In this case, it is not possible to simply glue the subject to the verb with \verb|PredVP|.
+When translating a senctence like this, it is thus not possible to simply glue the
+subject to the verb with \verb|PredVP|.\\
 %Therefore, the translator must change the type of the 
@@
 
-%förklara bättre eller ta bort??
-A similar issue arises form the use of verb particles. If a verb with a particle is considered
-as one klump in GF, which means that the particle has no effect on the abstract tree. Its presence
-is announced by the verb itself. The lexicon may contain a verb both with and without a particle,
-for example \emph{tänka} (think) and \emph{tänka till} (consider?).
-\begin{verbatim}
-taenka_V  : V ;
-taenka_till_V : V ;
-\end{verbatim}
-When the translator looks up a the form \emph{tänka}, both of the two verbs
-will be among the results. The first option will be selected, and if the
-particle isn't directly attached to the verb
-%There is, at this point, no easy way of knowing which one of them wants a particle and the
-the transformation might confuse the two words. This is also the case when translating
-all sorts of idioms that GF considers to be just one bit.\\
-
+%The lexicon may contain a verb both with and without a particle,
+%for example \emph{tänka} (``think") and \emph{tänka till} (``consider").
+%\begin{verbatim}
+%taenka_V  : V ;
+%taenka_till_V : V ;
+%\end{verbatim}
+%When the translator looks up a the form \emph{tänka}, both of the two verbs
+%will be among the results. %It is not known until later which one 
+%%The first option will be selected, and if the
+%%particle isn't directly attached to the verb
+%%There is, at this point, no easy way of knowing which one of them wants a particle and the
+%the transformation might confuse the two words. This is also the case when translating
+%all sorts of idioms that GF considers to be just one bit.\\
+%
 %Difficulties in looking up 'idioms' like this
@@
 
-On of the reasons why \textit{metas} are put in the output, is the 
-use of tags like \\
+One of the reasons why \textit{metas} are put in the output, is the 
+use of the tags show in table \ref{fig:mapBadtag}.
 \begin{figure}[h]
@@
 \textbf{XX} & Unclassifiable part-of-speech\\
-\end{tabular}\label{fig:mapBadtag}
+\end{tabular}
+\caption{}\label{fig:mapBadtag}
 \end{figure}
-
-They occur quite frequently in the treebank and are always translated
-to \emph{?}.
+The analyse in Talbanken makes a difference between subjects
+and objects, which is not needed in GF (see section \ref{sec:swegf}).
+Senctences like
+\enumsentence{För stora krav.\\
+``Too high demands."}
+hence makes use of the tags \textbf{XX} and \textbf{NP}, since it is not obvious whether
+the noun phrase is used as subject or an object.
+The tags shown above occur quite frequently in the treebank and are always translated
+to metas. 
 %How we cannot cover all cases.
@@
 >> why is this significant, other grammars can use this Swedish grammar.
->> well known (comp. ling) trade-off between depth & coverage. for now depth, robustness in in future
+>> well known (comp. ling) trade-off between depth and coverage. for now depth, robustness in in future
 >> work leads to coverage.
@@
 
-\bibliographystyle{ieeetr}
+\bibliographystyle{lncs}
 \bibliography{masterbib}

### 2011-12-07 gf/testgf.sh

@@ -1,0 +1,4 @@
+echo "p \"${1}\" | vp | wf -file=test.tmp" | gf -run BigTestSwe.gf
+dot -Tpdf test.tmp > test.pdf 
+xpdf test.pdf &
+ 

### 2011-12-08 doc/report/report.tex
report

@@
 \chapter{Introduction}
-\section{Introduction} 
->>This work is part of bigger project, spring and next spring.
->>Connection between grammar and parser: GF, every grammar defines a parser.
->> reusablilty, freely available.
->> the big amount of data, need language technology, machine translation
->> parsing: assingn of morph. and syntactic structures.
->>Connection between grammar and parser: GF, every grammar defines a parser.
-Ultimate goal: want computers to 'understand' language. Not good with flat strings.
-We need to translate strings of natural language to a richer structure,
-a parse tree. So far there is no
-freely available ruled-based parsing giving a deep analyse for Swedish.
+%\section{Introduction} 
+The grammar formalism GF has so far been succesfull used for 
+controlled languges, but recent experiments have showed
+that it is also possible to use the framework for parsing free language.
+Parsing, that is the task of automatically identifiyng morpohlogical and
+syntactical structures to text, is subject to increasing interest, especially
+considering the steadily growing amount of data available online. 
+So far there is no freely available ruled-based parser that gives a deep analyse for
+Swedish.  There are other parsers, but the fact that they are not freely available does not only
+restrict the usage but also their possiblities of being developed. Our goal is
+to create a open source Swedish grammar and parser that can be continuosly enhanced and
+be made use of in other projects.
+To build a parser or a grammar from scratch is a heavy work and we
+therefore combine already existing parts.
+We utgår från GF resource grammar and therefore have 
+The usage of GF allows us to start from
+a well-defined system for describing grammar, as well as a strong connection to
+more than 20 other langugages implentented in the framework.
+The grammar defines a parser, which accepts all constructions defined by the grammar.
+
+We hence needed to extend the Swedish grammar to cover
+the language specific constructions. Further a large lexicon needed, as well
+as a method for evaluation.
+During this project, we have worked on and explored those things and some problems identified.
+%We have extended the Swedish resource grammar in GF, to give a better coverage and imported
+%a large lexicon. Work to compare the trees from Talbanken to GF and to translate them.
+
+%Ultimate goal: want computers to 'understand' language. Not good with flat strings.
+%We need to translate strings of natural language to a richer structure,
+%a parse tree. 
 %There are parsers which put only put part-of-speach tags to
 %the text, and others that gives a deeper analyse. 
-There are statistical parser, and parser which only are deal with controlled language.
-The GF parser are one of those. 
+%There are statistical parser, and parser which only are deal with controlled language.
+%The GF parser are one of those. 
 %What is parsing, flat strings to trees, example picture, what is deep parsing.
-So far stable and good when working with controlled languages, but recent experiments
-have showed that it is also possible to parse free language.
-To do this for Swedish, we need to  .. the resource grammar needed to covering
-the language specific constructions. Further a large lexicon is also needed, as well
-as a method for evaluation.
-During this project, we have worked on and explored those things and some problems identified.
-We have extended the Swedish resource grammar in GF, to give a better coverage and imported
-a large lexicon. Work to compare the trees from Talbanken to GF and to translate them.
 
@@
 
-Other work exist, but none is freely available. We want to create a open source
-product, therefore can only use resources that allows this. Work can continue,
-will not disappear, be made use of in other projects.
 %What we could not cover by this project, what have been covered before (in
 %the spring project), and why we choose those parts.
-To build a parser or a grammar from scratch is a heavy work and we
-therefore combine already existing parts.
 
@@
 \subsection{Grammatical Framework}
+>> extensnion to CFG, same expressiveness as TAG and HG. parse in polynomial time.
+>> incremental parsing.
 The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
 a grammar formalism based on Martin-Löf type theory\cite{martinlof}. GF is
-designed for multilingual applications and is equivalent to Parallel Multiple
-Context-Free Grammars (Seki \& al. 1991), it is hence stronger than mildly
-contex-free grammars.
+designed for multilingual applications and the formalism is equivalent to
+Parallel Multiple Context-Free Grammars (Seki \& al. 1991), it is hence
+stronger than mildly contex-free grammars.
 
@@
 Agda\cite{agda}. 
+>> GF ~ 15 years. PGF= runtime systm
 It has been developed for about 20 years 
@@
 There are libraries for using PGF files with Java, Java script and Haskell. The grammars
+>> Haskell only complete, also or C\# and C. 
 may also be used in Android applications.
@@
 increased. There are  research about how to make better use of the dependent
-types, for using ontologies\cite{ramona}(>>check!!) or generating natural
+types, for using ontologies\cite{ramona} or generating natural
 language via Montague semantics\cite{montague}.
@@
 >> Much more here, important!! comparisons.
+>> HPSG and LFG are context sensitive
 Computational ling is active area for Swedish, there are many other Swedish parser,
@@
 that can be used with formal subjects (see section `Impersonal constructions').
-
+>> subject controlling verbs
+>> why comlicated with doubled types, show simplicity (elegant) with dep.types, and
+>> some problems (many fields).
+>> samesame as Obj, but conjunction works too, sin is Quant but what i sig?
+>> (show circle (sig : NP -> GenNP sig -> sin??)
 % agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
@@
 >> work leads to coverage.
+>> reusablilty, freely available.
 %There is obviously a lot do still, will never get finish with NPL.

### 2011-12-09 doc/report/report.tex
report with notes from meeting

@@
 
-Could varandra be thought of as reflexive for 'oss'?
+Mention varandra. Same idea.
 %Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
@@
 >> incremental parsing.
+>> how parse trees work, not correct complete representation.
+>> not a constituent structure. compare to lfg, fstructure, cstructure.
+>> explain how they are created, graphviz and how a user uses them.
 The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
@@
 \textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
+>> internal structure
+>> compare fstructure in LFG
 shared by many languages and can be implemented by a numeber of concrete grammars,
@@
 The same problem occurred for verbs that always are used in passive form:
+>> not passive!! deponent, s-verbs!!
 ``andas" (``breathe"), ``frodas" (``flourish"), ``lyckas" (``succeeed"). This
@@
 \verb|         ta -> blev tagen| \\
+>> frånta, erbjuda works. add. check vp in urdu
+>> how are they used with reflexives??
 This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
@@
 >> (show circle (sig : NP -> GenNP sig -> sin??)
+>> how we can add dependent type for Num to use 'varandra'
+>> CN should have dettyp too, 'avståndet till sin bil'
 % agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
@@
 
-Kanske mindre om detta, skriv om vad jag nu la till.
-om 'de äpplen', which are also used in 'finnas'
-Since those constructions are inspired by other languages, some of them may sound quite
+A few problematic things can be illuminated:
+Since the constructions are inspired by other languages, some of them may sound quite
 non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
@@
 \emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
-embedded in the relative clause. To fix this, an extra parameter was needed to tell whether
-\emph{`som'} or \emph{`vilket'} should be used. \\
+embedded in the relative clause. This is easily fixed, by adding an extra
+parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
 Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
@@
 \enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
-An more natural sounding alternative would be
-\enumsentence{Jag vill ha en sådan katt som inte fäller.}
-and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
-'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
-needed.
-du är här, vilket jag vet.
-i de äpplen/det äpple du äter bor en mask/ som du ser.\\
+This relative clause in the sentence is, in GF, put together by two constituents: \emph{`` katt"}
+of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause). They are
+glued together by \emph{``sådan att"} and \emph{``katt"} should agree with \emph{``den"}. 
+\enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
+However, this is not the case and sentence \ref{sent:detfaller} is currently accepted by the grammar.
+
+The grammar has been extended to accept
+\enumsentence{\shortex{7}
+{\textbf{De uppfattningar}& som& förs& fram ...} 
+{The opinions& that& put & forward ...}
+{``The opinions presented ..."}}
+where \emph{``de"} is used together with a indefinite noun. This is only correct for 
+relative clauses or together with the verb \emph{``finnas"}.
+In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
+\enumsentence{\textbf{De uppfattningarna} förs fram.\\
+``The opinions are presented."}\label{sent:debocker}
+% also ok in finnas!!
+
+%An more natural sounding alternative would be
+%\enumsentence{Jag vill ha en sådan katt som inte fäller.}
+%and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
+%'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
+%needed.
+%du är här, vilket jag vet.
+%i de äpplen/det äpple du äter bor en mask/ som du ser.\\
 

### 2011-12-11 doc/report/report.tex

@@
 \usepackage{wrapfig}
+\usepackage{tipa}
 % exempel: 2) Inte vet jag
@@
 
-\tableofcontents
-\newpage
-
 \abstract{
-% This is the report of the work of expanding a GF grammar and preparing
-% it for parsing of free Swedish. A lexicon consisting of more than 100 000 words
-% has been extracted and we have constructed an automatic mapping between the
-% annotation of the treebank Talbanken and GF. 
+This is the report of the work of expanding and enhancing a GF grammar and preparing
+it for parsing of free Swedish. 
+Outside of this project, an equivalent project for English is being worked upon,
+as are tehcniques for making the parser robust.
+With this in mind, we have extended the current GF grammar to adapt it for
+Swedish specifically so that it can parse an extended set of basic
+constructions. We have completed a method for importing the electronical 
+lexicon Saldo to a GF format and established a connection to the treebank Talbanken.
+The imported lexicon consists of more than 100 000 words
+and the grammar now covers ??\% of the xx part of Talbanken. 
+% The mapping between the annotations of Talbanken and GF is automatic. 
 % The goal is to build a robust parser dealing with uncontrolled natural
-% language, a parser which, for example, can parse all of Talbanken.
-% Outside of this project, an equivalent GF project for English is being worked upon,
-% as are theqniques for making the parser robust.
-% With this in mind, we have extended the current GF grammar to adapt it for
-% Swedish specifically so that it can parse the an extended set of basic
-% constructions. We have completed a method for importing the eletronical 
-% lexicon Saldo to a GF format and upprätat en connection mellan oss och Talbanken.
+%language, a parser which can parse all of Talbanken.
 
@@
 \newpage
+
+\tableofcontents
+\newpage
 \chapter{Introduction}
-%\section{Introduction} 
+>> short intro to GF
 The grammar formalism GF has so far been succesfull used for 
@@
 considering the steadily growing amount of data available online. 
-So far there is no freely available ruled-based parser that gives a deep analyse for
-Swedish.  There are other parsers, but the fact that they are not freely available does not only
-restrict the usage but also their possiblities of being developed. Our goal is
+So far there is no freely available grammar-driven parser that give a deep analyse for
+Swedish. The fact that the existing rule-based Swedish
+parsers are not freely available does not only restrict their
+usage but also their possiblities of being developed. Our
+goal is
 to create a open source Swedish grammar and parser that can be continuosly enhanced and
-be made use of in other projects.
-To build a parser or a grammar from scratch is a heavy work and we
-therefore combine already existing parts.
-We utgår från GF resource grammar and therefore have 
-The usage of GF allows us to start from
+be made use of in other projects.\\
+To build a parser or a grammar from scratch is a heavy work ~~ and we
+therefore combine already existing parts of ??.
+We start from GF resource grammar and thereby get
+%The usage of GF allows us to start from
 a well-defined system for describing grammar, as well as a strong connection to
 more than 20 other langugages implentented in the framework.
-The grammar defines a parser, which accepts all constructions defined by the grammar.
-
+>>The grammar defines a parser, which accepts all constructions defined by the grammar.
+
+\section{Aims}
+>> enhance gf...
+The purpose of this project has been to prepare the GF grammar for free parsing.
+This has meant to develop earlier techniques to fit for Swedish, create methods
+for achieving and keeping a extensive lexicon and to adapt the existing
+resource grammar to fit for more language specific constrictions.
 We hence needed to extend the Swedish grammar to cover
-the language specific constructions. Further a large lexicon needed, as well
-as a method for evaluation.
-During this project, we have worked on and explored those things and some problems identified.
+the language specific constructions. Further a large lexicon was
+needed, as well as a method for evaluation.
+During this project, we have worked on and explored those things and
+while identifying problems and future work.
+
+\section{Outline}
+This thesis is divided into 6 chapters. We start by giving some background
+information to areas relevant to the project. %Chapter \ref{sec:background} 
+Grammatical Framework is presented in section \ref{sec:gf} and a deeper
+presentation of the Swedish resource grammar is given in section \ref{sec:swegf}.
+Section \ref{sec:swedish} gives an introduction Swedish and
+brief presentations of Saldo and Talbanken are found in section \ref{sec:saldo} and
+\ref{sec:talbanken} respectively.
+A summary of related work can be found in section \ref{sec:related}.
+
+Chapters \ref{sec:prog.saldo} - \ref{sec:Mapping} present the methodology,
+implementation and some results. Finally, conclusions, an evaluation and a list
+of future work are presented in chapter \ref{sec:results}.
+
 %We have extended the Swedish resource grammar in GF, to give a better coverage and imported
@@
 
-\section{Background}  
+\chapter{Background}  
+\label{sec:background}
 The work described in this is thesis is part of a bigger project which aims at
@@
 
-\subsection{Swedish}
+\section{Swedish}
+\label{sec:swedish}
+>> move, have GF first?\\
 Swedish\cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
 closely related to Norwegian and Danish. The languages share most of their
-grammatical structures and are mutual intelligibility. Swedish is also 
-an official language in Finland and spoken by approximately 10 million people.
-Swedish syntax are often similar to English, but have some characteristics:
-\subsubsection*{Verb second}
-Swedish is a verb-second language\cite[p. 116]{gunlog}: the
+grammatical structures and are mutually intelligible. Swedish is also 
+an official language in Finland and altogether spoken by approximately 10 ~~
+9?Check SAG million people.
+Swedish syntax is often similar to English, but have some characteristics ~~:
+
+\subsection*{Post-nominal articles} 
+Unlike most of the world's languages Swedish express not only the number but also
+the definitness of a noun by suffixes. The endings are decided by the gender of the noun,
+which may be \textit{neuter} or \textit{common}.\\
+%(or \textit{utrum})
+
+\begin{tabular}{l|lll}
+&\textbf{Definite}& \textbf{Indefinite}& \textbf{Gender} \\
+ \hline
+\textbf{Singular} & {en katt} & {katten} & \textit{common}\\
+& \emph{a cat} & \emph{the cat} &\\
+\textbf{Plural} & katter & katterna &\\
+ &\emph{cats} & \emph{the cats} &\\
+ \hline
+\textbf{Singular} & {ett hus} & {huset} & \textit{neuter}\\
+ & \emph{a house} & \emph{the house} & \\
+\textbf{Plural} & hus & husen & \\
+& \emph{houses} & \emph{the houses} & \\
+\end{tabular}\\
+
+The definiteness can be accentuated by using the definite articles.\\
+
+\begin{tabular}{llll}
+den katten & det huset & de katterna & de husen \\
+\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
+\end{tabular}\\
+
+The definite articles are required when the noun is modified by adjectival phrase, sentence
+\ref{sent:trottkatt}b is hence not grammatically correct.
+%\enumsentence{\shortexnt{4}
+%{Den& trötta& katten & sov}
+%{the& tired& cat+\textbf{the} & slept}}
+\enumsentence{\begin{tabular}{llllll}
+a.& Katten sov. & b.& *Trötta katten sov &c. &Den trötta katten sov\\
+&The cat slept. & &~~The tired cat slept& &The tired cat slept\\
+\end{tabular}}\label{sent:trottkatt}
+
+%can be superdoubled
+%example: den där boken
+
+\subsection*{Verb second}
+Swedish is a verb-second language\cite[p.116]{gunlog}: the
 second constituent of a declarative main clause must consist of a verb.
@@
 The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
-meaning, although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being emphasized.
+meaning, but vary in how the content is presented.
+%~~ although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being
+%emphasized.
 \enumsentence{
 \shortex{4}
-{Du & ser & mig & inte.}
-{You & see & me & not.}
+{Du & ser &  inte &mig.}
+%empha glossning!! ha samma överallt, ' eller ''
+{you & see & not & me}
 {`You  don't see me'.}} \label{ex:swedish-svo}
@@
 {Mig  & ser &  du & inte.}
-{Me & see & you & not.}}\label{ex:swedish-ovs} 
+{me & see & you & not}}\label{ex:swedish-ovs} 
 \vspace{-3mm}
@@
 {Inte & ser& du &mig.}
-{Not&  see& you& me.}} \label{ex:swedish-avso}
-
-%\shortexm{3}
-%{hej} {hi}{hhe} {bing}{bong}{salt}
-%; adverbials, objects
+{not&  see& you& me}} \label{ex:swedish-avso}
+
 %Alla syntaktiska kategorier kan be fronted utom obetonade satsadverbial (ju,väl).
-%The first constituent of the clause is usually made up of the subject,
-%although it likewise could consist of adverbial phrases or objects that are
-%being empathazied. \\
 %\enumsentence{
@@
 {Såg & du & mig & inte?}
-{Saw  & you&  me& not?}
+{saw  & you&  me& not}
 {`Didn't you see me?'}}
@@
 The word order in subordinate clauses in also slightly different,
->> and stricter
-%\enumsentence{
-%\begin{tabular}{ll}
-%{\shortex{7}
-%{Jag & såg&  Anna.&  Hon & såg & inte&  mig.} 
-%{ & & &  & & &}}
-%\shortex{7}
-%{Jag såg att Anna \textbf{inte} såg mig.}\\
-%\shortex{7}
-%{I  saw Anna. She saw not   me.}  & {I   saw that Anna not saw me}. \\
-%`I saw Anna. $\;$ She did not see me.'& `I saw that Anna did not see me'.
-%}
+%>> and stricter
 \enumsentence{Main sentences:\\
-{\shortex{7}{Jag & såg & Anna. & Hon & såg & {\bf inte} & mig }
-{ I & saw & Anna. & She & saw & {\bf not} & me}
-{`I saw Anna. She didn't see me'}}}
+{\shortex{8}{Jag & såg & Anna & men & hon & såg & {\bf inte} & mig }
+{ I & saw & Anna & but & she & saw & {\bf not} & me}
+{`I saw Anna but she didn't see me'}}}
 \enumsentence{Subordinate sentence:\\
-\shortex{7}{Jag & såg & att& Anna& \textbf{inte}& såg& mig.}
-            {I &  saw & that& Anna& \textbf{not}& saw& me.}
-            {`I saw that Anna didn't see me'}}
-
-\subsubsection*{Passive voice}
-Passive voice is often used in Swedish, especially the
-\textit{s-passive}, which is formed by adding an \emph{s} at the end of a verb.
-\enumsentence{Passive \hspace{39mm} Active\\
-\shortexm{11}{Han & jagade\textbf{s} & av & ett & lejon.}
-        {He & hunted+\textbf{s} & by & a & lion.}
-        {`He was hunted by a lion.'}
-{Ett & lejon & jagade & honom.}
-        {A &lion &hunted &him.}
-        {`A lion was hunting him.'}}
-
+\shortex{7}{Jag & förstod & att& Anna& \textbf{inte}& såg& mig.}
+            {I &  understood & that& Anna& \textbf{not}& saw& me}
+            {`I understood that Anna didn't see me'}}
+
+\subsection*{Passive voice}
+There are two ways of forming passive verb phrases in Swedish: the 
+\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
+%>> better with animate subjects
+\enumsentence{
+\shortex{5}
+{Tjuven &blev& tagen& av & polisen}
+{the thief & was & taken & by & the police }
+{`The thief was arrested by the police'}}
+\label{gfPass:peri-pass}
+and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
+\enumsentence{Passive \hspace{45mm} Active\\
+\shortexm{12}
+{Tjuven & togs&  av & polisen}
+{the thief&  took+\textbf{s}&  by&the police}
+{`The thief was arrested by the police'}
+{&&Polisen & tog & tjuven}
+{&&the police & took & the thief}
+{\hspace{-9mm}`The police arrested the thief'}}
+\label{gfPass:s-pass}
+The s-passive is the most commonly used one for both written
+and Swedish \cite[??]{SAG} and dominates especially when the subject is inanimate.
 %cf.
@@
 
-\subsubsection*{Impersonal constructions}
-Constructions with 'det är/var' ('it is/was') is very common\cite[\textsection 309d]{H&H}:
+\subsection*{Impersonal constructions}
+Constructions with `det är/var' (\emph{`it is/was'}) is very common in Swedish \cite[\textsection 309d]{H&H}:
 \enumsentence{
 \shortex{6}{Det & var & roligt & att & höra.}
-{It &was & nice &to &hear.}
+{it &was & nice &to &hear.}
 {`I'm glad to hear that.'}}
-'Det' (\emph{it}) is also used as formal subject %for intransitive verbs,
-with the real subject put in the position of an object.
+'Det' (\emph{it}) is also used as formal subject in presentational constructions where
+the real subject is put in the position of an object.
 \enumsentence{
 \shortex{6}{Det &står &en &älg &på &fältet.}
-{ It & stands &a & moose &on &the field.}
-{`There is a moose on the field.'}}
-
-\subsubsection*{Reflexive pronouns}
+{ it & stands &a & moose &on &the field}
+{`There is a moose in the field.'}}
+
+\subsection*{Reflexive pronouns}
 \label{swe:refl}
@@
 and possessive pronouns for the 3rd
-person\cite[\textsection 310 \& 319]{H&H}. Those are distinct from the normal 3rd person forms.
+person\cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
 \enumsentence{
-\begin{tabular}{ll}
-Han slog \textbf{sig}. & Han såg \textbf{sitt} barn.\\
-\vspace{3mm}
-\emph{He hit him self.} & \emph{He saw his (own) child.}\\
-Han slog \textbf{honom}. & Han såg \textbf{hans} barn.\\
-\emph{He hit him (another person).}&\emph{He saw his (another person's) child.}
+\begin{tabular}{llll}
+a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
+&\emph{He hit him self.} && \emph{He saw his (own) child.}
 \end{tabular}}
-The 1st and 2nd persons use the normal personal pronoun in object form.
-
-Mention varandra. Same idea.
+\enumsentence{
+\begin{tabular}{llll}
+a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
+&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
+\end{tabular}}
+The 1st and 2nd persons use the normal personal pronoun in object form as reflexive
+pronouns.
+
+%Mention varandra. Same idea.
 %Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
@@
 
-\subsection{Saldo}
+\section{Saldo}
+\label{sec:saldo}
 A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
 large electronic lexicon developed and maintained at Gothenburg University. It is
-built upon Svenskt Associations Lexicon and contains information about more than 
+built on Svenskt Associations Lexicon and contains information about more than 
 120 000 modern Swedish words.
-The words are equipped with semantic, syntactic and a morphological
+The ~~words are equipped ~~ with semantic, syntactic and a morphological
 information. The user can find compounding analyses, examples of usage in corpora,
 and graphs of semantically connected words.\\
-\newpage % only until saldo can be on same page as the pic
 \begin{figure}[h!]
@@
 
-If one word
-have different meanings, there will be separate entries for each of them.
-The word ``växa" (``grow") for example, has two meanings: \emph{getting bigger}
+If one word has different meanings, there will be separate entries for each of
+them.  The word ``växa" (``grow") for example, has two meanings: \emph{getting
+bigger}
 \enumsentence{Gräset växte varje dag\\
@@
 only keeps one entry of ``växa".
-For the homonyms that does not share the whole inflection tables, different entries are 
+For the homonyms that do not share the whole inflection tables, different entries are 
 kept in the morphological lexicon.
@@
 
-\subsection{Grammatical Framework}
->> extensnion to CFG, same expressiveness as TAG and HG. parse in polynomial time.
->> incremental parsing.
->> how parse trees work, not correct complete representation.
->> not a constituent structure. compare to lfg, fstructure, cstructure.
->> explain how they are created, graphviz and how a user uses them.
-The main component of the the project is the Grammatical Framework\cite{gfbok} (GF). It is
+\section{Grammatical Framework}
+\label{sec:gf}
+The main component of the the project is the Grammatical Framework (GF) \cite{gfbok}, %It is
 a grammar formalism based on Martin-Löf type theory\cite{martinlof}. GF is
@@
 Parallel Multiple Context-Free Grammars (Seki \& al. 1991), it is hence
-stronger than mildly contex-free grammars.
-
-GF is a functional programming languages, inspired by programming languages
-like ML\cite{ml}, Haskell\cite{haskell}, Prolog\cite{bratko} and
-Agda\cite{agda}. 
->> GF ~ 15 years. PGF= runtime systm
-It has been developed for about 20 years 
->> snd: functional and programming language, type checking and logic: prolog (check this)
->> dependent and implicit types : agda. (read gf book and krasses book)
-
+stronger than mildly contex-free grammars and can express whatever tree
+adjoining grammars (Joshni,xx) and HG?? can.  --very confusing here!!
+
+GF is a strongly typed functional programming languages, inspired by
+%programming languages like 
+ML \cite{ml} and Haskell \cite{haskell}. Its type
+checking system is inspired by \textlambda Prolog\cite{bratko} and the
+dependent types by Agda\cite{agda}. 
+>> check this!!
+The first version of the framework was implemented in 19xx in xx and has
+since then fatt en trogen skara användare, med headquarter 
+at Gothenburg University.
+
+%> dependent and implicit types : agda. (read gf book and krasses book)
 A grammar written in GF can be used for both parsing and generation.
+The parsing algorithm is incremental and has polynomial time and space
+complexity \cite{gfMech}. >>check!
 The GF package also provides various tools for working with and using the grammars:
-a compiler, an interpreter and a runtime system. The interpreter, the GF shell, lets
-you try and test your grammar by using commands for parsing, vizualization of parse trees,
+a compiler, an interpreter and a runtime system.
+The grammars can be compiled into portable grammar format (PGF)\ref{pgf},
+which can be used with %There are libraries for using PGF files with
+Java, Java script, C and Haskell. 
+%>> Haskell only complete, also or C\# and C.  %may also be used in Android applications.
+The interpreter, the GF shell, allows
+the user to try and test her grammar by using commands for parsing, vizualization of parse trees,
 random generation, word alignment, morphological quizzes etc.
-The compiler lets you compile your grammar into portable grammar format (PGF)\ref{pgf}.
-There are libraries for using PGF files with Java, Java script and Haskell. The grammars
->> Haskell only complete, also or C\# and C. 
-may also be used in Android applications.
-and are inspired by.\\
-
+%The following example shows how to parse a sentence and visualizing the parse tree.
+%, by using the \textit{pipe} \verb-|-, sending the result to 
+\begin{figure}[h]
+\begin{verbatim} 
+> parse "jag ser dig" | visualize_parse
+> parse "jag ser dig" | visualize_tree
+\end{verbatim}
+\caption{Example of how the GF shell can be used. The first command will parse the given
+sentence and visualize the resulting parse tree and the second will visualize the 
+abstract syntax tree.}
+\label{fig:shellvp}
+\end{figure}
+>> test vp -format=pdf -view=evince
+>> test vt -format=pdf -view=evince -nofun
+
+The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
+Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
+%a visualization of a abstract syntax tree and a parse tree.
+\begin{figure}[h]
+>> example, typ : i see you. 
+\caption{Visualized GF trees}
+\label{fig:gftree1}
+\end{figure}
+The reader should keep in mind that the visualized parse trees is not a
+complete representation.
+%>> how parse trees work, not correct complete representation.
+%>> not a constituent structure. compare to lfg, fstructure, cstructure.
+%>> explain how they are created, graphviz and how a user uses them.
+%>> fotnoote that they are not complete
+
+\subsection{Writing a GF grammar}
 The key feature in GF is the distinction between
-\textit{abstract} and \textit{concrete} syntaxes. The abstract is meant to be
->> internal structure
->> compare fstructure in LFG
-shared by many languages and can be implemented by a numeber of concrete grammars,
-each representing one language. The abstract syntax models the semantics of the
-domain without concern for language specific features such as agreement or word
-order.
+\textit{abstract} and \textit{concrete} syntax. The abstract syntax represents
+the internal structure and models the semantics without concern for language
+specific features such as agreement or word order.
+One abstract grammar can be implemented by a number of concrete grammars, each
+representing one language. As a comparison, the abstract and concrete syntax
+may be thought of as fstructures and cstructures in LFG (x,x).
 
@@
 
+>> move much of this to Swedish GF part?
 Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
 for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|,
-and \verb|Pred| simply glues the two strings ``katten" and ``sover" together:
+and \verb|Pred| simply glues the two strings ``katten" and ``sover" together:\\
 \verb|Pred cat sleep = "katten sover"|.\\
@@
 plural and singular. To do this, we add a category \verb|NP| to the
+>> reformulate so that I don't sound stupid
 abstract. \verb|N| now means a noun
@@
 %>> key, talk, hats, bom (see references in Ramonas thesis).
+>> use talk here and move key to other place??
 %, the Online Syllogism Solver\cite{malin!}. 
@@
 
-\subsubsection{The resource library}
+\subsection{The resource library}
 %>>Best part? Biggest library?
@@
 The resource grammars describe how to construct phrases and sentences and how to
-decline words. The latter is done by smart paradigms: functions that analyses
+decline words. The latter is done by smart paradigms: functions that analyse
 some given forms of a word to find out how the inflection table should look.
@@
 and the possibilities to use GF in different environments
-increased. There are  research about how to make better use of the dependent
+increased. There are research on how to make better use of the dependent
 types, for using ontologies\cite{ramona} or generating natural
@@
 model an important part of Swedish in GF. The starting point was the resource
-grammar and the result should still have a strong connection to this.
+grammar and the to which the result has a strong connection.
 
@@
 
-\subsection{Talbanken}
+\section{Talbanken}
+\label{sec:talbanken}
 For testing and evaluation of the grammar and lexicon, we needed to be able to
 compare them against a reliable source.
-Being an freely available, manually annotated, large scale treebank,
+Being a freely available, manually annotated, large scale treebank,
 Talbanken\cite{talbanken} was perfect for our purpose.
@@
 parts. Two of them are transcribitions of spoken language, one a collection of
-text written by high school students, and the last one, section P,
+text written by high school students, and one, section P,
 consists of professionally written Swedish gathered from newspapers, brochures and textbooks.\\
 Talbanken was also used to train the Swedish version of the Maltparser\cite{malt}
-and was thus redistributed in a more modern version\cite{talbanken05} This work
-was done by Nivre, Hall and Nilsson in 2005. They released it in Malt and Tiger
+and was thus redistributed in a more modern version\cite{talbanken05}.
+%This work was done by Nivre, Hall and Nilsson in 2005.
+It is released in Malt and Tiger
 XML-formats\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html},
@@
 \section{Related work}
+\label{sec:related}
 >> Much more here, important!! comparisons.
@@
 
-\chapter{Proceedings}
-%\section{Proceedings med bättre ord}  
-The purpose of this project has been to prepare the grammar for free parsing.
-This has meant to develop earlier techniques to fit for Swedish, create methods
-for achieving and keeping a extensive lexicon and to adapt the existing
-resource grammar to fit for more language specific constrictions.
-The methodology and results will be presented in the following section.
-
-\section{Importing Saldo}
+\chapter{Importing Saldo}
 \label{sec:prog.saldo}
+\section{Implementation}
 >> syntactic information too. Mention that GF does not have this, only lexical or
@@
 Saldo is compatible with GF, which makes it easy to transform the morphological
-lexicon into our format. The lexicon has been imported earlier (Krasimir,2009?)
+lexicon into our format. The lexicon has been imported earlier (Angelov,2009?)
 but the code needed updates to not miss many important words.
@@
 
-\subsection{Results}
+\section{Results}
 All words from Saldo were not be imported. There are a number of reasons
 for this, the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions(\emph{pp}) and numerals(\emph{nl}) etc. are
-assumed to be known already,
+Saldo in the GF lexicon. Prepositions(\emph{pp}) and numerals(\emph{nl}) etc.
+~~are assumed to be known already~~,
 and are usually not kept in the standard lexicon, but in modules for the basic
@@
 analysed the same way, see section \ref{sec:swegf}. Some experiments on finding
-the right GF-category for those have been done, see \ref{sec:gf.quant}.
+the right GF-category for those have been done, see ~~~~~\ref{sec:gf.quant}.
 Six categories were imported: \\
+>> english examples
 \begin{tabular}{l|lll}
@@
 One reason for why words of the categories listed above could not be imported
-is that Saldo does not contain information about word forms that are not used.
+is that Saldo does not contain information about ~~word forms that are not used~~.
 In some cases those forms are needed by the smart paradigm. 
@@
 requires a singular form, this lemma could not be added to the lexicon. ``Glasögon" could
-be added manually, by using the somewhat incorrect singular form ``glasöga" (``glass-eye").
+be added manually, by using the ~~somewhat incorrect,annan betydelse~~ singular form ``glasöga" (``glass-eye").
 The same problem occurred for verbs that always are used in passive form:
@@
 In some cases the smart paradigms could not generate the correct declination,
-although this happened seldom. % frysa,fisa passive preterium
+although this happened seldom. 
+>> exapmles: % frysa,fisa passive preterium
 
@@
 numerals and proper names, but some names are still there). In addition, there were 1817 words
+>> give reasons! should look good, not bad!
 which were given different labels in GF than in Talbanken. Many of those are
@@
 Among the words that were completely missing, more than 2000 was found to be nouns,
-many of which were compounds.
+many of which were compounds, and this is hard because ...~~~
 
@@
 (see section \ref{sec:Mapping}) it should be possible to extract this information about this.
+>> about what? how? strategy?
 %The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
@@
 
-\section{The grammar}
+\chapter{The grammar}
 \label{sec:prog.grammar}
-Although we can't expect to get a full coverage, we aim for a high-quality 
+Although we can't expect to get a full coverage, we aim for a ~~high-quality~~
 grammar with satisfactory coverage.
-We have been working on a grammar fragment which have been adapted to cover
+We have been developing a grammar fragment which have been adapted to cover
 constructions used in Talbanken. The starting point has been the Swedish
 resource grammar and the new grammar should still be compatible with this.
-A description of the Swedish implementation of the resource grammar is given
-in the next section, while section \ref{sec:Added} shows and explains the
-development during this project.
+%A description of the Swedish implementation of the resource grammar is given
+%in the next section, while
+Before describing what has been added, will give a short introductions to the
+the cats of resource grammar and the sw implementation in particular.
+This is done in section \ref{sec:swegf} while section \ref{sec:Added} shows and
+explains the development during this project.
 %Explain what's in the resources, constructions present in most languages. 
@@
 
-\subsection{The Swedish resource grammar}
+\section{The Swedish resource grammar}
+>> start by easy examples, jag ser inte dig, du ser inte mig osv. 
+>> plocka från svenska delen 1 2 4 
+>> show simple trees
+>> explain all in Sw. section.
+>> show abstract and concrete tree, compare. (or in Gf sect)
 >> no subjects or object, analyse is more from the lexicon,
@@
 word order, agreement, tense, basic conjunction etc.
+
+In figure \ref{fig:gftree1} we saw how the sentence ``Jag ser dig" is parsed.
+Discuss questions, how are word order handled here? Maybe negation and anterior.
+Let's look on how post-determination (sec ref) is implemented, we know that the
+smart paradigms can handle the noun declination, now show 'katten' vs. 'den trötta katten'.
+A N carries the inflection information, turned in to CN by setting number. Example, type
+of N and CN. A CN may then be modified by APs, set field for this. When turned into NP, 
+definitness is set, depends on the isMod -field. NP has case as a parameter.
+\verb|DetNP : Det -> CN -> NP ; |
+To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
+but allow variation in.. 
+To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
+Is a quantifier, hence does not require number, is given later.
+%What is the difference?
+In gf, only personal pronouns are considered to belong to PN. The rests (example from
+saldo, which is pronoun in SAG as well) are either Quantifiers or Dets.
+Once we have a NP, can use Predets. What they are.\\
+
+For verbs, but also others, there is an important distinction for valency.
+Encoded in the lexicon. Categories as V,V2,VS...
+Either directly to VP if V, or needs to be given arguments. Can then add adverbs.
+%Maybe talk about ställa.
+Particles and prepositions, can't be seen on the trees. Can now see why the parse
+trees are not complete. difference shows when fronting.
+tycka om /tycka om -> om vad tycker du inte. fast bra
+Swedish verb may have up to five arguments according to Stymne
+'jag tar med mig den till Bo' : ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;
+but is really 'till' part of verb?
+VP is divided corresponding to satsschema (diderichsen) (renewed by SAG).
+If the object is not given immediatly, the grammar uses slash.
+"Speed up", instead of movement used slashes\cite{gazdar}.
+Slashes good for question (vilken katt ser du idag)
+relatives (katten som du tittar på) and topicalisation
+The grammar has VP,Cl, SSlash, all missing NP.
+
+Example, use variants {}.
+
+>> change dot file to have VP NP in correct order
 \enumsentence{\shortexnt{8}
@@
 \enumsentence{\shortexnt{5}{Hon & ser & pojken & som  &sover}
-              {She & sees & the boy & who & sleeps}}
+              {she & sees & the boy & who & sleeps}}
               %{`She sees the sleeping boy'}}
@@
 \enumsentence{\shortexnt{6}{Han & ser &katten & han  &tycker & om}
-              {He & sees & the cat & he & likes &}}
+              {he & sees & the cat & he & likes &}}
               %{`She sees the sleeping boy'}}
@@
 %For example, the \verb|Extra| module included
-Among those were functions for topologication, although not all
+Among those were functions for topicalisation, although not all
 of them were implemented for Swedish.
@@
 {Det & äpplet & vill & jag & inte & ha}
-{That & apple & want & I & not &}
-{I don't want that apple}}\label{gfSwe:apple}
+{that & apple & want & I & not &}
+{`I don't want that apple'}}\label{gfSwe:apple}
 The module provided functions for preposition stranding:
@@
 %how the parse tree is generated, the finite verb, the subject, the rest. usw.
-%han såg katten han tyckte om
-%för vem måste jag akta mig
-%vem måste jag akta mig för
 
 %honom ville hon inte tänka på.
-
-Before describing what has been added, will give a short introductions to the
-the cats of resource grammar and the sw implementation in particular.
-
-Most important now is NP and VP.
-Starting from N, get to CN what is it, what can we do now.
-To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
-but allow variation in.. 
-What is the difference? Some words on differences for pronouns.
-To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
-Example, use variants {}.
-There are also Quants, which .... 
-Once we have a NP, can use Predets. What they are.
-
-For verbs, but also others there is an important distinction for valency.
-Either directly to VP if V, or needs to be given arguments. Can then add adverbs.
-Maybe talk about ställa.
-Particles and prepositions, can't be seen on the trees.
-tycka om /tycka om -> om vad tycker du inte. fast bra
-'jag tar med mig den till Bo' : ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;
-but is really 'till' part of verb?
-VP is divided corresponding to satsschema (diderichsen) (renewed by SAG),
-otherwise use slash.
-"Speed up", instead of movement used slashes\cite{gazdar}.
-Slashes good for question (vilken katt ser du idag)
-relatives (katten som du tittar på) and topicalisation
-The grammar has VP,Cl, SSlash, all missing NP.
-
-%Besides the categories NP,VP,AP there are also the import Det and Quant, Predet.
 
@@
 %valency built in (problems with ställa - på, under, vid).
-%Determiners, quants, predets, The difference from how pronouns are regarded. 
 %Variants, nonexist.
@@
 
-\subsection{Development of the grammar}
+\section{Development of the grammar}
 \label{sec:Added}
@@
 
+\subsection{A second future tense: ``Kommer att"}
 kommer att - for scand, extra tense, example
@@
 >> why not VPSlash here??
-There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
-\enumsentence{ \shortexnt{6}
-{Bilden &blev& tagen& av & rymdsonden & Galileo}
-{The picture & was & taken & by & the spacecraft & Galileo}}
-\label{gfPass:peri-pass}
-or by adding an \emph{s} to
-the verb: \\
-\enumsentence{\shortex{5}
-{Bilden & togs&  av & rymdsonden&  Galileo}
-{The picture&  took+\textbf{s}&  by&the  spacecraft& Galileo}
-{``The picture was taken by the spacecraft Galileo"}}
-\label{gfPass:s-pass}
-This second variant, the \textbf{s-passive}, is commonly used in Swedish, some studies
-suggest that it is used in more than 80\% of the times \cite{laanemets}.
+Passive voice is often used in Swedish, especially the
+\textit{s-passive}.
+\enumsentence{
+\shortex{5}{Uppsatsen & skrev\textbf{s} & av & en & student.}
+        {the essay & wrote+\textbf{s} & by & a & student}
+        {`The essay was written by a student.'}}\label{sent:skrevs}
+%{b. &En & student & skrev & uppsatsen.}
+%        {&a &student &wrote &the essay}
+%        {~~`A student wrote the essay.'}}
+Some studies suggest that the s-passive is used in more than 80\% of the times \cite{laanemets}.
 It is not as common in the other Scandinavian languages however, 
 and it can not be used in all tenses for all words. The Norwegian 
-translation of both sentence \ref{gfPass:peri-pass} and sentence  \ref{gfPass:s-pass} is
+translation of sentence \ref{sent:skrevs} is:
 %The periphrastic passive is preferred, especially in spoken langauge.
-\enumsentence{Bildet ble tatt av romsonden Galileo}
-The resource grammar for Scandinavian therefore expressed the passive using auxiliary verb. \\
+\enumsentence{\begin{tabular}{ll}
+Oppgaven ble skrevet av en student & [NO]\\
+Uppsatsen blev skriven av en student& [SE]\\
+\end{tabular}}
+The corresponding Swedish sentence is acceptable, but not as natural sounding as sentence
+\ref{sent:skrevs}.
+The resource grammar for Scandinavian therefore ~~expressed~~ the passive using auxiliary verb. \\
+>> change to:
 \verb|PassV2 : V2  -> VP | \\
@@
 than the default.
-The grammar also allows using three-place verbs \verb|V3| , as \emph{give}, to
+>> error!!! 
+\verb|PassVPSlash : VPSlash  -> VP | \\
+\verb|              ta -> togs| \\
+\verb|              erbjöd -> erbjöds| \\
+The grammar also allows not only V2 but all verb phrases missing an object to 
 form passives:
-\enumsentence{\shortex{6}
-{Betyget & 3 &ges &till &många &elever}
-{Grade & 3 & give+\textbf{s}& to & many & students}
-{``Grade 3 is given to many students"}}
-\label{ex:passV3}
-by letting the first object be `removed' and given by the subject. 
-A remark can be made about sentences where the second object is put first in the sentence, like 
-\enumsentence{Till många elever ges betyget 3}
-This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
-parse this by applying \verb|FocObj|.
+\enumsentence{First place in two-place verbs\shortex{3}
+{Jobbet & erbjöds & henne }
+{the job & offered+\textbf{s}& henne}
+{``The job was offered to her"}}
+\label{ex:passV32}
+\enumsentence{Second place in two-place verbs\shortex{3}
+{Hon & erbjöds & jobbet}
+{she & offered+\textbf{s}& the job}
+{``She was offered the job"}}
+\label{ex:passV33}
+\enumsentence{\shortex{3}
+{Huset & målades & rött}
+{the house & painted+\textbf{s}& red}
+{``The house was painted red"}}
+\label{ex:passV2A}
+
+%r
+%\enumsentence{\shortex{6}
+%{Betyget & 3 &ges &till &många &elever}
+%{Grade & 3 & give+\textbf{s}& to & many & students}
+%{``Grade 3 is given to many students"}}
+%\label{ex:passV3}
+%by letting the first object be `removed' and given by the subject. 
+%
+%A remark can be made about sentences where the second object is put first in the sentence, like 
+%\enumsentence{Till många elever ges betyget 3}
+%This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
+%parse this by applying \verb|FocObj|.
 %%
@@
 Transitive verbs may not be used like this
-\enumsentence{
-*Det äter en fågel på taket\\
- ''There is a bird eating on the roof"}
+>> elisabet disagrees: use det äter en fågel ost på taket
+\enumsentence{\shortexnt{6}
+{*Det & gillar & en & fågel & på & taket}
+{It & likes & a & bird & on & the roof"}}
 
 unless their in passive form
-\enumsentence{
-Det äts en fågel på taket \\
-A bird is being eaten on the roof}
+\enumsentence{\shortex{5}
+{Det&dricks&mycket & öl& nuförtiden}
+{It&drink+s&much & beer& nowadays}
+{A lot of beer is being drunk these days}}
 A very common example of this is sentences with the verbs \emph{finnas},
-\emph{saknas} and \emph{fattas}.
+\emph{saknas} and \emph{fattas}. > engelska.
 \enumsentence{\shortex{3}
@@
 described above.
-
+>> read cooper and say smt smart here
 There are also restrictions on the subject.
@@
 
-\textbf{advs}
+\subsubsection{advs}
 -- or is this really just AdVs? 
@@
           'har ätit redan äpplet' :( which doesn't work either :)
-
-\textbf{bara}
-but have other words like 'bara' which can be put in other positions.
-  but needs to reconstruct whole of VP, since no field is before fin.
-  Can also act before the noun or as AdV. Examples.
-\verb|s SPres Simul Neg Main : han bara sover inte alltid|
-  have added a field in VP, a0, show table.
-  makes some weird constructions when using modal verbs, 
-  \verb|s SPres Anter Pos Main : han bara har alltid sovit|
+\subsubsection{Focusing adverbs}
+GF separates between two categories of adverbs, the normal ones \verb-Adv-, 
+eg. `nu' (\emph{`now'}) and
+special ones (name!) \verb-AdV- like `alltid' (\emph{`always'}).
+The \verb-AdV- attaches directly to the verb and have a special
+field in the \verb-VP--table.
+\enumsentence{\shortex{4}
+{Jag & äter & aldrig & fisk}
+{I & eat & never & fish}
+{`I never eat fish'}}
+\enumsentence{\shortex{4}
+{Jag & äter & fisk& nu }
+{I & eat & fish& now }
+{`I eat fish now'}}
+
+\begin{wrapfigure}{l}{0.3\textwidth}
+\begin{tabular}{|l|}
+\hline
+bara \\
+inte ens \\
+till och med \\
+\hline
+\end{tabular}
+\caption{Focusing adverbs}
+\label{fig:focadv}
+\end{wrapfigure}\\
+The focusing adverbs, like `bara' (\emph{`only' or `just'}) may put in both? of this
+positions and additionally, they may be used before the finite verb.
+
+\enumsentence{\shortex{3}
+{Han & bara & log }
+{he & only & smiles }
+{`He just smiles'}}\\
+
+In order to implement this, a new field was needed in the \verb-VP--table,
+which may be put before the finite verb.\\
+\verb|s SPres Simul Pos Main : han bara sover|\\
+compare to negation, \verb-AdV- and \verb-Adv-\\
+\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
+example in Sub or Inv, 
+When used together with modal verbs, the new field results in
+some not very natural sounding constructions:\\
+\verb|s SPres Anter Pos Main : han bara har sovit|\\
 but rather than changing to 'han har bara sovit', we keep it. Is a possible
@@
 ambiguities.
+
+The foucusing adverbs may also be used as predeterminers:
+\enumsentence{\shortex{7}
+{Det & är & bara & barn & får & göra & så }
+{it & is & only & childern & may & do & so }
+{`Only children are allowed to do that'}}
+This is handled by the rule:
+\verb|PredetAdvF : AdvFoc -> Predet ;|
 
@@
 >> remove stupid examples, add 'hon och sin vän åkte tåg'
+>> use self instead of his.
 \enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
@@
 
-The reflexive pronouns require another subject with which they agree in
+The reflexive pronouns require an antecedent with which they agree in
 number and person. Apart from this restriction, they may be used as normal
-noun phrases and conjuncted (sentence \ref{ex:ReflConj} and \ref{ex:ReflConj2}),
+noun phrases and ~~conjuncted~~ (sentence \ref{ex:ReflConj} and \ref{ex:ReflConj2}),
 determined (sentence \ref{ex:ReflPredet}) and used as genitives (sentence
@@
 
-The solution introduces a new object type, \verb_Obj_, which is
-identical to normal \verb_NP_ except for that it depends on the agreement of
-the subject. All functions using noun phrase as objects are changed to dealing
+Another solution introduces a new object type, \verb_Obj_, which is
+identical to normal \verb_NP_ except for that it depends on 
+the antecedent. All functions using noun phrase as objects are changed to dealing
 with objects instead.
@@
 >> CN should have dettyp too, 'avståndet till sin bil'
-% agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
+
+>>Mention varandra. Same idea.
+>>agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
 % F1,2,3 : NP -> VP -> Cl ; NP needs gender (Masc,Fem,Utr,Neutr)
@@
 \enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-However, this is not the case and sentence \ref{sent:detfaller} is currently accepted by the grammar.
+However, this is not the case and sentence \ref{sent:detfaller} is currently
+accepted by the grammar.
 
@@
 
-\subsection{Testing}
+\section{Testing}
 Hard do test, can't verify. Regression testing. Talbanken is still too hard.
@@
 
-\section{Mapping}
+\chapter{Mapping}
 >> no subjects or object, analyse is more from the lexicon,
@@
 >> mamba very used and tested for swedish, standard.
+>> focus on what we DO cover.
 \label{sec:Mapping}
@@
 \textbf{NNDDHSGG} & familjemedlemmarnas & Noun, definite, compound (person), genitive \\
-\textbf{PU} & * $\;$ 1) $\;$ a) & List item\\
+\textbf{PU} & * \; 1) \; a) & List item\\
 \end{tabular}\\
@@
 %GF valencies prescribes.
-Another problem can be instantiated by sentence \ref{ex:mappMan}, containing
+Another problem can be illustrated by sentence \ref{ex:mappMan}, containing
 the generic pronoun \emph{``man"}.
@@
 
-\subsection{Results and evaluation}
+\section{Results and evaluation}
 >> more! replace words, how much better results (both grammar and mapping) 
 >> more results
+>> motivation, why so bad??
 The flat version of Talbanken05 was used for developing the mapping, but
@@
 
-\chapter{Summary}
-\section{Results and discussion}
+\chapter{Results and discussion}
+\label{sec:results}
+\section{Evaluation}
 %Results from mapping, can probably be improved, bit by bit, depends much on words.
@@
 
-\bibliographystyle{lncs}
+\bibliographystyle{apalike}
 \bibliography{masterbib}

### 2011-12-11 doc/report/report.tex
report on gf

@@
 >> move, have GF first?\\
-Swedish\cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
+Swedish \cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
 closely related to Norwegian and Danish. The languages share most of their
@@
 \subsection*{Verb second}
-Swedish is a verb-second language\cite[p.116]{gunlog}: the
+Swedish is a verb-second language \cite[p.116]{gunlog}: the
 second constituent of a declarative main clause must consist of a verb.
 The normal word order is subject-verb-object, but any syntactic category can be
-fronted\cite[\textsection 1027]{H&H}.
+fronted \cite[\textsection 1027]{H&H}.
 Topicalisation is very common, especially for temporal and spatial adverbial phrases.
@@
 and possessive pronouns for the 3rd
-person\cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
+person \cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
 \enumsentence{
@@
 \label{sec:saldo}
-A good parser needs a good lexicon. We have used of Saldo\cite{saldo}, a
+A good parser needs a good lexicon. We have used of Saldo \cite{saldo}, a
 large electronic lexicon developed and maintained at Gothenburg University. It is
@@
 The main component of the the project is the Grammatical Framework (GF) \cite{gfbok}, %It is
-a grammar formalism based on Martin-Löf type theory\cite{martinlof}. GF is
+a grammar formalism based on Martin-Löf type theory \cite{martinlof}. GF is
 designed for multilingual applications and the formalism is equivalent to
@@
 ML \cite{ml} and Haskell \cite{haskell}. Its type
-checking system is inspired by \textlambda Prolog\cite{bratko} and the
-dependent types by Agda\cite{agda}. 
+checking system is inspired by \textlambda Prolog \cite{bratko} and the
+dependent types by Agda \cite{agda}. 
 >> check this!!
@@
 \begin{verbatim} 
-> parse "jag ser dig" | visualize_parse
-> parse "jag ser dig" | visualize_tree
+> parse "jag ser katten" | visualize_parse
+> parse "jag ser katten" | visualize_tree
 \end{verbatim}
@@
 Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
-%a visualization of a abstract syntax tree and a parse tree.
+A GF parse tree has two visualized representations, the \textit{parse tree} and
+the \textit{abstract tree}.
 \begin{figure}[h]
->> example, typ : i see you. 
-\caption{Visualized GF trees}
+\centering
+\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=60mm]{gfTree.pdf}}
+\subfloat[GF parse tree]{\label{pic:gfStree}\includegraphics[width=50mm]{gfPTree.pdf}}
+\subfloat[GF parse tree]{\label{pic:gfEtree}\includegraphics[width=50mm]{gfETree.pdf}}
+\caption{Parse tree and abstract tree for the sentence ``Jag ser katten".}
 \label{fig:gftree1}
 \end{figure}
+The abstract trees shows all functions GF needs to use
+to parse the sentence.
+% do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
+The parse tree (\ref{pic:gfStree}) shows the words and which types the functions assign
+to them.
 The reader should keep in mind that the visualized parse trees is not a
 complete representation.
+If the end node of a  branch has no corresponding word, it is excluded from
+the visualized parse tree. This is the case for the definite article in Swedish,
+and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
+In the corresponding English parse, figure \ref{pic:gfEtree},
+the article \emph{``the"} is explicit and therefore shown in the parse tree.
+
 %>> how parse trees work, not correct complete representation.
 %>> not a constituent structure. compare to lfg, fstructure, cstructure.
-%>> explain how they are created, graphviz and how a user uses them.
-%>> fotnoote that they are not complete
+
+%>>the scope is controlled language, but possible ambiguities (discuss this in future
+%>>work, saldo, conclusion). Ambiguities: from multilinguality, or inherited:
+%>>syntactical (man with the telescope), lexical (youSg/Pl, växa).
+%>> have ambiguity section. 
+The scope of GF grammars is usually controlled language. By
+restricting the coverage, the number of ambiguities
+can be limited and we can ensure that the semantics is preserved during
+translation. The ambiguities arising from multilinguality and the inherited will of
+course remain, but limited. %Example of domainspecific word?
+The use of controlled language gives a possibility of good and reliable translation.
+Some examples of projects using GF are WebAlt \cite{webalt},
+a project which aims to develop language independent material for mathematics problems,
+and the formal software verification tool KeY \cite{key}, in wich GF is used for
+translating specification to natural language.
+%>> key, talk, hats, bom (see references in Ramonas thesis).
+%>> use talk here and move key to other place??
+The framework is also used in the
+European project MOLTO \cite{molto} for online translation between up to 15 language.\\
 
@@
 
->> move much of this to Swedish GF part?
+%>> move much of this to Swedish GF part?
 Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
@@
 We get a more complicated example if we allow the nouns to be used in both
-plural and singular. To do this, we add a category \verb|NP| to the
->> reformulate so that I don't sound stupid
-abstract. \verb|N| now means a noun
+plural and singular. We add a category \verb|NP| to the
+abstract. Note that \verb-NP- is normally used for nouns with a fixed definitness, but for 
+now we concentrate on the number only. Hence \verb|N| now means a noun
 in any of the number, and we introduce two function, \verb|NSg : N -> NP| and
@@
 In the concrete syntax, \verb|N| is defined to be a record consisting of the field
-\verb|s|. The type of \verb|s| shows that it is a table, which given a parameter of type \verb|Num| returns
-a string. \verb|Num| is defined as a parameter, which can either have value \verb|Sg|
+\verb|s|. The type of \verb|s| shows that it is a table, which given a
+parameter of type \verb|Num| returns a string. \verb|Num| is defined as a
+parameter, which can either have value \verb|Sg|
 or \verb|Pl|. In \verb|NPl| and \verb|NSg|, \verb|n.s| means that we use the
@@
 
->>not very restricted, not few ambiguities.
->>the scope is controlled language, but possible ambiguities (discuss this in future
->>work, saldo, conclusion). Ambiguities: from multilinguality, or inherited:
->>syntactical (man with the telescope), 
->>lexical (youSg/Pl, växa).
-A GF grammar usually covers a very restricted set of language where few ambiguities
-occur and much care can be taken to preserving the semantics of the
-different languages.
-By using a controlled language like this,  good and reliable translation can be
-ensured. Some examples of projects using GF are WebAlt\cite{webalt},
-a project which aims to develop language independent material for mathematics problems,
-and the formal software verification tool KeY\cite{key}, in wich GF is used for
-translating specification to natural language.
-%>> key, talk, hats, bom (see references in Ramonas thesis).
->> use talk here and move key to other place??
-%, the Online Syllogism Solver\cite{malin!}. 
-The framework is also used in the
-European project MOLTO\cite{molto} for online translation between up to 15 language.\\
-
 \subsection{The resource library}
@@
 %>>Most general resource, first large grammar, covers fundamental syntactics.
-The GF package provides a powerful resource library\cite{gf-resource}, which covers the
+The GF package provides a powerful resource library \cite{gf-resource}, which covers the
 fundamental morphology and syntactics for more than 20 languages. %is a very useful part of the framework. It is provided in the 
@@
 
-A GF parse tree has two visualized representations, the \textit{parse tree} and
-the \textit{abstract tree}, as shown in figure \ref{fig:gfPandAtrees}.
->> have other treees! One in english.
-\begin{figure}[h]
-\centering
-\subfloat[GF parse tree]{\label{pic:gfPtree}\includegraphics[width=50mm]{gfPTree.pdf}}
-\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=60mm]{gfTree.pdf}}
-\caption{Parse tree and abstract tree for the sentence ``Katten på bilen blir större".}
-\label{fig:gfPandAtrees}
-\end{figure}
-The abstract trees shows all functions GF needs to use
-to parse the sentence and is consequently the same for all concrete syntaxes.
-The parse tree (\ref{pic:gftree}) shows the words and which types the functions assign
-to them.
-The abstract tree shows us that the function
-\verb_DetCN_ is applied the \verb-UseN cat_N- and the information
-definitive (\verb-DefArt-) and singular (\verb-NumSg-). In the Swedish parse tree
-we see that \emph{``katten"}, which is in definite form singular, is noun \verb-N-
-which is turned into a \verb-CN- and finally a \verb-NP-. In the English parse
-tree, the form of \emph{``cat"} expresses singular while the definite article
-\emph{``the"} is explicit and therefore shown in the parse tree.
-
-\vspace{5mm}
+Section \ref{sec:swegf} will give a more thorough description of the Swedish resource grammar.
+
+\subsection{Frontiers of Grammatical Framework}
 As an open source-project, GF is constantly being developed and improved. New
@@
 increased. There are research on how to make better use of the dependent
-types, for using ontologies\cite{ramona} or generating natural
-language via Montague semantics\cite{montague}.
+types, for using ontologies for reasoning \cite{ramona} or generating natural
+language via Montague semantics \cite{montague}.
 Further, experiments on using GF for free parsing has been conducted
@@
 model an important part of Swedish in GF. The starting point was the resource
-grammar and the to which the result has a strong connection.
+grammar to which the result has a strong connection.
 
@@
 Being a freely available, manually annotated, large scale treebank,
-Talbanken\cite{talbanken} was perfect for our purpose.
+Talbanken \cite{talbanken} was perfect for our purpose.
 It was created in the 1970s at Lund University. In total it consists of four
@@
 consists of professionally written Swedish gathered from newspapers, brochures and textbooks.\\
-Talbanken was also used to train the Swedish version of the Maltparser\cite{malt}
-and was thus redistributed in a more modern version\cite{talbanken05}.
+Talbanken was also used to train the Swedish version of the Maltparser \cite{malt}
+and was thus redistributed in a more modern version \cite{talbanken05}.
 %This work was done by Nivre, Hall and Nilsson in 2005.
@@
 but grammars?
-Many of the parsers not ruled based, the most known one Malt\cite{malt} ... .
-However has been two, about CLE\cite{cle} and what happened to it. Logic and translation
+Many of the parsers not ruled based, the most known one Malt \cite{malt} ... .
+However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
 to Swedish, S-CLE (Gambäck,1997) full analyse and semantics, translation,
 but coverage is too limited (nivres papper), The
-spoken lang. translator\cite{cle2}. Wiren \cite{wiren}?? chartparser, but for swedish. bit more information than
+spoken lang. translator \cite{cle2}. Wiren \cite{wiren}?? chartparser, but for swedish. bit more information than
 just pos: cassSwe \cite{casswe}
@@
 to be described in a future version." 
-HPSG\cite{hpsg} and lingo matrix\cite{lingomatrix}, English, Japanese, German,
+HPSG \cite{hpsg} and lingo matrix \cite{lingomatrix}, English, Japanese, German,
 Greece, Spanish, Norweigan (NorSource, Hellan 2003),( not famous: SweCore
 (Ahrenberg) based on NorSource)
-also Swedish: BiTSE\cite{stymne}, Stymne 2006, Swedish English translation,
+also Swedish: BiTSE \cite{stymne}, Stymne 2006, Swedish English translation,
 and one for the common parts of scandinavian (Haugereid etc) in Nordisk sprog.
 focusing on verb phrames, built on HPSG.
-S\o rgaard and Haugereid, Scandinavian Grammar Matrix\cite{scandmatrix}
+S\o rgaard and Haugereid, Scandinavian Grammar Matrix \cite{scandmatrix}
 Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English: ongoing to get widecoverage grammar
@@
 Consider for example
-the plural tantum %\cite[44]{SAG}
+the plural tantum % \cite[44]{SAG}
 noun ``glasögon" (``glasses"). Since the smart paradigm
@@
 If the object is not given immediatly, the grammar uses slash.
-"Speed up", instead of movement used slashes\cite{gazdar}.
+"Speed up", instead of movement used slashes \cite{gazdar}.
 Slashes good for question (vilken katt ser du idag)
@@
 {he & only & smiles }
-{`He just smiles'}}\\
+{`He just smiles'}}
 
@@
 \label{sec:futureProbabilities}
+% discuss ambiguites, will always have some and the big lexicon will increase them.
+% exapmle 'vad händer under äktenskapet?'
 %Getting and using probabilities, to disambiguate.

### 2011-12-11 doc/report/report.tex
related work section

@@
 \chapter{Introduction}
->> short intro to GF
+%>> short intro to GF
 The grammar formalism GF has so far been succesfull used for 
@@
 To build a parser or a grammar from scratch is a heavy work ~~ and we
-therefore combine already existing parts of ??.
+therefore combine already existing resources.
 We start from GF resource grammar and thereby get
@@
 more than 20 other langugages implentented in the framework.
->>The grammar defines a parser, which accepts all constructions defined by the grammar.
+%>>The grammar defines a parser, which accepts all constructions defined by the grammar.
 
 \section{Aims}
->> enhance gf...
+%>> enhance gf...
 The purpose of this project has been to prepare the GF grammar for free parsing.
@@
 a grammar formalism based on Martin-Löf type theory \cite{martinlof}. GF is
-designed for multilingual applications and the formalism is equivalent to
-Parallel Multiple Context-Free Grammars (Seki \& al. 1991), it is hence
-stronger than mildly contex-free grammars and can express whatever tree
-adjoining grammars (Joshni,xx) and HG?? can.  --very confusing here!!
+designed for multilingual applications and the formalism is 
+%equivalent to Parallel Multiple Context-Free Grammars (Seki \& al. 1991), an instance of
+%generalized context free grammar (GCFG) it is hence
+stronger than mildly contex-free grammars and hence has the same expressiveness
+as Tree Adjoining Grammars (Joshi,1975) and Head Grammars (Pollard,1984).
 
@@
 %programming languages like 
-ML \cite{ml} and Haskell \cite{haskell}. Its type
-checking system is inspired by \textlambda Prolog \cite{bratko} and the
-dependent types by Agda \cite{agda}. 
->> check this!!
-The first version of the framework was implemented in 19xx in xx and has
-since then fatt en trogen skara användare, med headquarter 
-at Gothenburg University.
+ML \cite{ml} and Haskell \cite{haskell}. The logic and reasoning 
+are inspired by \textlambda Prolog \cite{prolog} and 
+by Agda \cite{agda}, with which GF also shares its type checking algorithm.
+The first version of the framework was implemented by the Xerox Research Center
+in Grenoble and is now mainly developed in Gothenburg. On of the biggest
+achivements is a library covering the 
+basic morphological and syntactical structures of more than
+20 languages (see section \ref{sec:resources}).
 
@@
 The parsing algorithm is incremental and has polynomial time and space
-complexity \cite{gfMech}. >>check!
+complexity \cite{gfMech}. 
 The GF package also provides various tools for working with and using the grammars:
 a compiler, an interpreter and a runtime system.
-The grammars can be compiled into portable grammar format (PGF)\ref{pgf},
+The grammars can be compiled into portable grammar format (PGF) \cite{pgf},
 which can be used with %There are libraries for using PGF files with
@@
 \begin{verbatim} 
-> parse "jag ser katten" | visualize_parse
-> parse "jag ser katten" | visualize_tree
+> parse "jag ser katten" | visualize_parse -format=pdf -view=evince
+> parse "jag ser katten" | visualize_tree  -format=pdf -view=evince
 \end{verbatim}
@@
 \end{figure}
->> test vp -format=pdf -view=evince
->> test vt -format=pdf -view=evince -nofun
+\begin{figure}[h]
+\centering
+\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
+\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
+\caption{Parse tree and abstract tree for the sentence ``Jag ser katten".}
+\label{fig:gftree1}
+\end{figure}
+
+% nice: test vt -format=pdf -view=evince -nofun
+\begin{wrapfigure}{l}{0.3\textwidth}
+\includegraphics[width=20mm]{gfETree.png}
+\caption{English parse tree}
+\label{pic:gfEtree}
+\end{wrapfigure}
 
@@
 the \textit{abstract tree}.
-\begin{figure}[h]
-\centering
-\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=60mm]{gfTree.pdf}}
-\subfloat[GF parse tree]{\label{pic:gfStree}\includegraphics[width=50mm]{gfPTree.pdf}}
-\subfloat[GF parse tree]{\label{pic:gfEtree}\includegraphics[width=50mm]{gfETree.pdf}}
-\caption{Parse tree and abstract tree for the sentence ``Jag ser katten".}
-\label{fig:gftree1}
-\end{figure}
 The abstract trees shows all functions GF needs to use
@@
 % do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
-The parse tree (\ref{pic:gfStree}) shows the words and which types the functions assign
-to them.
+The parse tree (\ref{pic:gfStree}) shows the words and which types
+the functions assign to them.
 The reader should keep in mind that the visualized parse trees is not a
@@
 and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
-In the corresponding English parse, figure \ref{pic:gfEtree},
-the article \emph{``the"} is explicit and therefore shown in the parse tree.
+
+In the corresponding English parse tree, figure \ref{pic:gfEtree},
+the article \emph{``the"} is explicit and therefore shown as a quantifier in the parse tree.\\
 
@@
 can be limited and we can ensure that the semantics is preserved during
-translation. The ambiguities arising from multilinguality and the inherited will of
-course remain, but limited. %Example of domainspecific word?
+translation. The inhereted ambiguities and the ones arising from multilinguality will
+of course remain, but can be controlled more easily. %Example of domainspecific word?
 The use of controlled language gives a possibility of good and reliable translation.
@@
 representing one language. As a comparison, the abstract and concrete syntax
-may be thought of as fstructures and cstructures in LFG (x,x).
+may be thought of as fstructures and cstructures in Lexical Functional Grammar (Bresnan and Kaplan, 1982).
 
@@
 \subsection{The resource library}
+\label{sec:resources}
 %>>Best part? Biggest library?
@@
 plural in definite and indefinite form.
-
 Section \ref{sec:swegf} will give a more thorough description of the Swedish resource grammar.
@@
 \label{sec:related}
->> Much more here, important!! comparisons.
->> HPSG and LFG are context sensitive
-Computational ling is active area for Swedish, there are many other Swedish parser,
-but grammars?
-Many of the parsers not ruled based, the most known one Malt \cite{malt} ... .
-However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
-to Swedish, S-CLE (Gambäck,1997) full analyse and semantics, translation,
-but coverage is too limited (nivres papper), The
-spoken lang. translator \cite{cle2}. Wiren \cite{wiren}?? chartparser, but for swedish. bit more information than
-just pos: cassSwe \cite{casswe}
-Other Sw. parsers and grammars, evaluation of them? 
-Other grammar frameworks.
-SweCG (birn) annotation, parser Swedish FDG (voultanien,2001) which is not freely available.
-finding subject, objects, suubject complements (SC) and
-to link them to their proper regents (main verbs) : over 90\% both precision and recall.
-uses a dependency grammar, but "The functional description of adverb phrases
-and prepositional phrases (e.g. agent, source, goal, benefactive, time) remains
-to be described in a future version." 
-HPSG \cite{hpsg} and lingo matrix \cite{lingomatrix}, English, Japanese, German,
-Greece, Spanish, Norweigan (NorSource, Hellan 2003),( not famous: SweCore
-(Ahrenberg) based on NorSource)
-also Swedish: BiTSE \cite{stymne}, Stymne 2006, Swedish English translation,
-and one for the common parts of scandinavian (Haugereid etc) in Nordisk sprog.
-focusing on verb phrames, built on HPSG.
-S\o rgaard and Haugereid, Scandinavian Grammar Matrix \cite{scandmatrix}
-Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English: ongoing to get widecoverage grammar
-under GPL. Translation to Korean.
+There has been much research about parsing Swedish. One of the most well known
+parser is the data-driven Malt parser \cite{malt}, which has been trained on 
+Talbanken. 
+There are also a number of grammar-based parsers, although none is still freely avaible.
+CassSwe \cite{casswe} is a cascaded finite state parser giving a syntactic
+analysis.
+The Swedish Constraint Grammar 
+(Birn,1998) gives a shallow syntactic analysis.
+Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
+(Tapanainen and Järvinen,1997), an extension of the Constraint Gramman
+formalism, and produces a dependecy structure focusing on finding the nominal
+arguments. \\
+%for the  finding subject, objects, suubject complements (SC) and
+%to link them to their proper regents (main verbs) : over 90\% both precision and recall.
+
+%Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English:
+%ongoing to get widecoverage grammar
+%under GPL. Translation to Korean.
+
+The Lingo Matrix \cite{lingomatrix}, is a starter-kit for building Head-Driven Phrase
+Structure Grammars \cite{hpsg} (HPSG) providing comptability with tools for
+parsing, evaluation, semantic representations etc.
+There is a collection of implementations done in this framework?, describing ie.
+English, Japanese and German. %, Greece, Spanish etc. 
+The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
+Scandiavian, while Norsource (Hellan, 2003) describes Norweigan. A Swedish version
+was based upon this (SweCore, Ahrenberg) which covers the morphology and some
+differences between Swedish and Norweigan. Further, there is the BiTSE \cite{stymne}
+grammar, also implemented using the Lingo Matrix,
+which focuses on verb frames and translation of those by using Minimal Recursion
+Semantics \cite{mrs} as an interlingua. \\
+It should be pointed out that grammars implemented using Lingo Matrix
+do not share any common abstract, like the resources grammars of GF?\\
+
+The Swedish verison \cite{gamback} of the Core Language Enigne (CLE) \cite{cle}
+gives a full syntacitc analysis as well as semantics represented in xx. A
+translation for conversations?? to English was implemented. The work was further developed in the spoken
+language blabla (xx,xx), but is no longer available. The coverage of the Swedish
+CLE is also reported to be very limited (ref).\\
+
+In the TAG formalism (Joshi,1975), there are projects on getting open source, widecoverage grammars
+for ie. English and Korean, but, to our knowledge, not for Swedish.  
+%However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
+%to Swedish, S-CLE (Gambäck,1997) full analyse and semantics, translation,
+%but coverage is too limited (nivres papper), The
+%spoken lang. translator \cite{cle2}.
+
+%>> Much more here, important!! comparisons.
+%>> HPSG and LFG are context sensitive
+%Computational ling is active area for Swedish, there are many other Swedish parser,
+%but grammars?
+%Many of the parsers not ruled based, the most known one Malt \cite{malt} ... .
+%Wiren \cite{wiren}?? chartparser, but for
+%swedish. bit more information than just pos: cassSwe \cite{casswe}
+%Other Sw. parsers and grammars, evaluation of them? 
+%Other grammar frameworks.
+%SweCG (birn) annotation, parser Swedish FDG  (voultanien,2001)which is not
+%freely available.
+%finding subject, objects, subject complements (SC) and
+%to link them to their proper regents (main verbs) : over 90\% both precision and recall.
+%uses a dependency grammar, but "The functional description of adverb phrases
+%and prepositional phrases (e.g. agent, source, goal, benefactive, time) remains
+%to be described in a future version." 
+%HPSG \cite{hpsg} and lingo matrix , English, Japanese, German,
+%Greece, Spanish, Norweigan (NorSource, Hellan 2003),( not famous: SweCore
+%(Ahrenberg) based on NorSource)
+%also Swedish: BiTSE \cite{stymne}, Stymne 2006, Swedish English translation,
+%and one for the common parts of scandinavian (Haugereid etc) in Nordisk sprog.
+%focusing on verb phrames, built on HPSG.
+%S\o rgaard and Haugereid, Scandinavian Grammar Matrix \cite{scandmatrix}
 % related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,

### 2011-12-12 doc/report/report.tex
better version of ch.1&2

@@
 \chapter{Introduction}
-%>> short intro to GF
+% more intro to gf? Good starting sentence.
 The grammar formalism GF has so far been succesfull used for 
@@
 considering the steadily growing amount of data available online. 
-So far there is no freely available grammar-driven parser that give a deep analyse for
-Swedish. The fact that the existing rule-based Swedish
-parsers are not freely available does not only restrict their
-usage but also their possiblities of being developed. Our
+So far there is no freely available grammar-driven parser that gives a deep analyse for
+Swedish. The fact that %the existing rule-based Swedish
+a parser or grammar is not freely available does not only restrict its
+usage but also its possiblities of being developed. Our
 goal is
 to create a open source Swedish grammar and parser that can be continuosly enhanced and
-be made use of in other projects.\\
-To build a parser or a grammar from scratch is a heavy work ~~ and we
-therefore combine already existing resources.
+made use of in other projects.\\
+To build a parser or a grammar from scratch would be not only time conusming but
+also a great theoretical and techniqual challenge. In order to not reinvent the wheel
+we develop our parser by combining well-tested resources.
 We start from GF resource grammar and thereby get
@@
 a well-defined system for describing grammar, as well as a strong connection to
-more than 20 other langugages implentented in the framework.
+more than 20 other langugages implentented in the framework. Further, we use
+the extensive lexicon Saldo and the treebank Talbanken.
 %>>The grammar defines a parser, which accepts all constructions defined by the grammar.
@@
 \section{Outline}
-This thesis is divided into 6 chapters. We start by giving some background
+The thesis is divided into 6 chapters. We start by giving some background
 information to areas relevant to the project. %Chapter \ref{sec:background} 
-Grammatical Framework is presented in section \ref{sec:gf} and a deeper
-presentation of the Swedish resource grammar is given in section \ref{sec:swegf}.
-Section \ref{sec:swedish} gives an introduction Swedish and
+Grammatical Framework is presented in section \ref{sec:gf} while a more profound
+description of the Swedish resource grammar is given in section \ref{sec:swegf}.
+Section \ref{sec:swedish} gives an introduction Swedish, and
 brief presentations of Saldo and Talbanken are found in section \ref{sec:saldo} and
@@
 using GF for parsing free Swedish. 
-In previous work \ref{??} the start on extending the Swedish GF grammar was
-made and a tool for lexical aquistion were developed.
+In previous work\footnote{??}, the start on extending the Swedish GF grammar was
+made and a tool for lexical aquistion was developed.
 We now construct a bigger
 and more expressive grammar as well as a large scale GF lexicon.
-As all GF grammars, this one can be used for parsing, and we develop it by getting
-examples, ideas and test material from a treebank.
+As all GF grammars, this one defines a parser, and we develop it by getting
+examples, ideas and test material from the treebank Talbanken.
 The project is hence heavily depending on three resources, which will be described
@@
 
-\section{Swedish}
-\label{sec:swedish}
->> move, have GF first?\\
-Swedish \cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
-closely related to Norwegian and Danish. The languages share most of their
-grammatical structures and are mutually intelligible. Swedish is also 
-an official language in Finland and altogether spoken by approximately 10 ~~
-9?Check SAG million people.
-Swedish syntax is often similar to English, but have some characteristics ~~:
-
-\subsection*{Post-nominal articles} 
-Unlike most of the world's languages Swedish express not only the number but also
-the definitness of a noun by suffixes. The endings are decided by the gender of the noun,
-which may be \textit{neuter} or \textit{common}.\\
-%(or \textit{utrum})
-
-\begin{tabular}{l|lll}
-&\textbf{Definite}& \textbf{Indefinite}& \textbf{Gender} \\
- \hline
-\textbf{Singular} & {en katt} & {katten} & \textit{common}\\
-& \emph{a cat} & \emph{the cat} &\\
-\textbf{Plural} & katter & katterna &\\
- &\emph{cats} & \emph{the cats} &\\
- \hline
-\textbf{Singular} & {ett hus} & {huset} & \textit{neuter}\\
- & \emph{a house} & \emph{the house} & \\
-\textbf{Plural} & hus & husen & \\
-& \emph{houses} & \emph{the houses} & \\
-\end{tabular}\\
-
-The definiteness can be accentuated by using the definite articles.\\
-
-\begin{tabular}{llll}
-den katten & det huset & de katterna & de husen \\
-\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
-\end{tabular}\\
-
-The definite articles are required when the noun is modified by adjectival phrase, sentence
-\ref{sent:trottkatt}b is hence not grammatically correct.
-%\enumsentence{\shortexnt{4}
-%{Den& trötta& katten & sov}
-%{the& tired& cat+\textbf{the} & slept}}
-\enumsentence{\begin{tabular}{llllll}
-a.& Katten sov. & b.& *Trötta katten sov &c. &Den trötta katten sov\\
-&The cat slept. & &~~The tired cat slept& &The tired cat slept\\
-\end{tabular}}\label{sent:trottkatt}
-
-%can be superdoubled
-%example: den där boken
-
-\subsection*{Verb second}
-Swedish is a verb-second language \cite[p.116]{gunlog}: the
-second constituent of a declarative main clause must consist of a verb.
-The normal word order is subject-verb-object, but any syntactic category can be
-fronted \cite[\textsection 1027]{H&H}.
-Topicalisation is very common, especially for temporal and spatial adverbial phrases.
-The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
-meaning, but vary in how the content is presented.
-%~~ although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being
-%emphasized.
-\enumsentence{
-\shortex{4}
-{Du & ser &  inte &mig.}
-%empha glossning!! ha samma överallt, ' eller ''
-{you & see & not & me}
-{`You  don't see me'.}} \label{ex:swedish-svo}
-\vspace{-3mm}
-\enumsentence{
-\shortexnt{4}
-{Mig  & ser &  du & inte.}
-{me & see & you & not}}\label{ex:swedish-ovs} 
-\vspace{-3mm}
-\enumsentence{
-\shortexnt{4}
-{Inte & ser& du &mig.}
-{not&  see& you& me}} \label{ex:swedish-avso}
-
-%Alla syntaktiska kategorier kan be fronted utom obetonade satsadverbial (ju,väl).
-%\enumsentence{
-%\begin{tabular}{lll}
-%  Johan & gick & sakta på gatan.\\
-%\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
-%  Sakta & gick & Johan på gatan.\\
-%\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
-%  På gatan & gick & Johan sakta.\\
-%\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
-%\end{tabular}}
-
-Inverted word order marks questions
-\enumsentence{
-\shortex{4}
-{Såg & du & mig & inte?}
-{saw  & you&  me& not}
-{`Didn't you see me?'}}
-%\enumsentence{Gick Johan på gatan?\\
-%        \emph{Did Johan walk on the street?}}
-The word order in subordinate clauses in also slightly different,
-%>> and stricter
-\enumsentence{Main sentences:\\
-{\shortex{8}{Jag & såg & Anna & men & hon & såg & {\bf inte} & mig }
-{ I & saw & Anna & but & she & saw & {\bf not} & me}
-{`I saw Anna but she didn't see me'}}}
-\enumsentence{Subordinate sentence:\\
-\shortex{7}{Jag & förstod & att& Anna& \textbf{inte}& såg& mig.}
-            {I &  understood & that& Anna& \textbf{not}& saw& me}
-            {`I understood that Anna didn't see me'}}
-
-\subsection*{Passive voice}
-There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
-%>> better with animate subjects
-\enumsentence{
-\shortex{5}
-{Tjuven &blev& tagen& av & polisen}
-{the thief & was & taken & by & the police }
-{`The thief was arrested by the police'}}
-\label{gfPass:peri-pass}
-and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
-\enumsentence{Passive \hspace{45mm} Active\\
-\shortexm{12}
-{Tjuven & togs&  av & polisen}
-{the thief&  took+\textbf{s}&  by&the police}
-{`The thief was arrested by the police'}
-{&&Polisen & tog & tjuven}
-{&&the police & took & the thief}
-{\hspace{-9mm}`The police arrested the thief'}}
-\label{gfPass:s-pass}
-The s-passive is the most commonly used one for both written
-and Swedish \cite[??]{SAG} and dominates especially when the subject is inanimate.
-%cf.
-%\enumsentence{
-%\shortex{5}{Ett & lejon & jagade & honom.}
-%        {A &lion &hunted &him.}
-%        {`A lion was hunting him.'}}
-
-\subsection*{Impersonal constructions}
-Constructions with `det är/var' (\emph{`it is/was'}) is very common in Swedish \cite[\textsection 309d]{H&H}:
-\enumsentence{
-\shortex{6}{Det & var & roligt & att & höra.}
-{it &was & nice &to &hear.}
-{`I'm glad to hear that.'}}
-'Det' (\emph{it}) is also used as formal subject in presentational constructions where
-the real subject is put in the position of an object.
-\enumsentence{
-\shortex{6}{Det &står &en &älg &på &fältet.}
-{ it & stands &a & moose &on &the field}
-{`There is a moose in the field.'}}
-
-\subsection*{Reflexive pronouns}
-\label{swe:refl}
-The Scandinavian language have special reflexive pronouns
-and possessive pronouns for the 3rd
-person \cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
-\enumsentence{
-\begin{tabular}{llll}
-a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
-&\emph{He hit him self.} && \emph{He saw his (own) child.}
-\end{tabular}}
-\enumsentence{
-\begin{tabular}{llll}
-a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
-&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
-\end{tabular}}
-The 1st and 2nd persons use the normal personal pronoun in object form as reflexive
-pronouns.
-
-%Mention varandra. Same idea.
-%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
-%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
-%for what will be written in the grammar part.
-%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
-%är stor, I am big
-%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
-
-\section{Saldo}
-\label{sec:saldo}
-A good parser needs a good lexicon. We have used of Saldo \cite{saldo}, a
-large electronic lexicon developed and maintained at Gothenburg University. It is
-built on Svenskt Associations Lexicon and contains information about more than 
-120 000 modern Swedish words.
-The ~~words are equipped ~~ with semantic, syntactic and a morphological
-information. The user can find compounding analyses, examples of usage in corpora,
-and graphs of semantically connected words.\\
-\begin{figure}[h!]
-\centering
-\subfloat[Morphological info for ``katt"]{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
-\hspace{20mm}
-\subfloat[Graph showing the hyponyms of ``katt"]{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
-\caption{Information in Saldo}
-\label{fig:saldo}
-\end{figure}
-
-If one word has different meanings, there will be separate entries for each of
-them.  The word ``växa" (``grow") for example, has two meanings: \emph{getting
-bigger}
-\enumsentence{Gräset växte varje dag\\
-The grass grew each day}
-or \emph{being a plant}. 
-\enumsentence{Det växte en gammal ek där.\\
-There stood an old oak.}
-The word ``växa" is however declined the same way for both interpretations of the word.
-Hence Saldo's morphological lexicon, provided under LPGL in XML format,
-only keeps one entry of ``växa".
-For the homonyms that do not share the whole inflection tables, different entries are 
-kept in the morphological lexicon.
-
-For our purpose, only the morphological lexicon was needed.
-The data can be processed and translated to GF format as described in 
-section \ref{sec:prog.saldo}.
-
 \section{Grammatical Framework}
 \label{sec:gf}
-The main component of the the project is the Grammatical Framework (GF) \cite{gfbok}, %It is
+The main component of the project is the Grammatical Framework (GF) \cite{gfbok}, %It is
 a grammar formalism based on Martin-Löf type theory \cite{martinlof}. GF is
-designed for multilingual applications and the formalism is 
+designed for multilingual applications and represents a formalism
 %equivalent to Parallel Multiple Context-Free Grammars (Seki \& al. 1991), an instance of
 %generalized context free grammar (GCFG) it is hence
-stronger than mildly contex-free grammars and hence has the same expressiveness
+stronger than mildly contex-free grammars. The framework has  hence the same expressiveness
 as Tree Adjoining Grammars (Joshi,1975) and Head Grammars (Pollard,1984).
 
-GF is a strongly typed functional programming languages, inspired by
+GF is a strongly typed functional programming language, inspired by
 %programming languages like 
-ML \cite{ml} and Haskell \cite{haskell}. The logic and reasoning 
-are inspired by \textlambda Prolog \cite{prolog} and 
+ML \cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
+and the built functionality for logic and reasoning 
+is inspired by \textlambda Prolog \cite{prolog} and 
 by Agda \cite{agda}, with which GF also shares its type checking algorithm.
@@
 The grammars can be compiled into portable grammar format (PGF) \cite{pgf},
-which can be used with %There are libraries for using PGF files with
-Java, Java script, C and Haskell. 
+supported by %There are libraries for using PGF files with
+Java, Java script, C and Haskell libraries. 
 %>> Haskell only complete, also or C\# and C.  %may also be used in Android applications.
 The interpreter, the GF shell, allows
-the user to try and test her grammar by using commands for parsing, vizualization of parse trees,
+the user to test grammars by commands for parsing, vizualization of parse trees,
 random generation, word alignment, morphological quizzes etc.
@@
 % nice: test vt -format=pdf -view=evince -nofun
-\begin{wrapfigure}{l}{0.3\textwidth}
+\begin{wrapfigure}{l}{0.22\textwidth}
 \includegraphics[width=20mm]{gfETree.png}
@@
 \end{wrapfigure}
-
 The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
 Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
-A GF parse tree has two visualized representations, the \textit{parse tree} and
-the \textit{abstract tree}.
-The abstract trees shows all functions GF needs to use
-to parse the sentence.
+%A GF parse tree has two visualized representations, the \textit{parse tree} and
+%the \textit{abstract tree}.
+The \textit{parse tree} (\ref{pic:gfStree}) shows which types
+that are assigned to the words and phrases. 
+The \textit{abstract tree} shows which functions GF uses
+to parse a sentence.
 % do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
-The parse tree (\ref{pic:gfStree}) shows the words and which types
-the functions assign to them.
 The reader should keep in mind that the visualized parse trees is not a
-complete representation.
-If the end node of a  branch has no corresponding word, it is excluded from
-the visualized parse tree. This is the case for the definite article in Swedish,
-and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
-
+complete representation. It does not reproduce all information from the abstract tree,
+%If the end node of a  branch has no corresponding word, it is excluded from
+%the visualized parse tree. This is the case for the 
+eg. the definitnes and number of the noun \emph{`katten'}, which is shown as
+\verb-DetQuant DefArt NumSg- in the abstract tree.
+%and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
 In the corresponding English parse tree, figure \ref{pic:gfEtree},
-the article \emph{``the"} is explicit and therefore shown as a quantifier in the parse tree.\\
+the article \emph{``the"} is explicit and therefore shows as a quantifier in the parse tree.\\
 
@@
 %>> have ambiguity section. 
+\newpage %as long as english picture bråks.
 The scope of GF grammars is usually controlled language. By
@@
 can be limited and we can ensure that the semantics is preserved during
-translation. The inhereted ambiguities and the ones arising from multilinguality will
+translation. Inhereted ambiguities and the ones arising from multilinguality will
 of course remain, but can be controlled more easily. %Example of domainspecific word?
-The use of controlled language gives a possibility of good and reliable translation.
-  Some examples of projects using GF are WebAlt \cite{webalt},
-a project which aims to develop language independent material for mathematics problems,
-and the formal software verification tool KeY \cite{key}, in wich GF is used for
-translating specification to natural language.
+The use of controlled language thus gives a possibility of good and reliable translation
+and GF has succesfully been used in project like %Some examples of projects using GF are 
+WebAlt \cite{webalt}, a project which aims to develop language independent
+material for mathematics problems, and the formal software verification tool
+KeY \cite{key}, in wich GF is used for translating specification to natural language.
 %>> key, talk, hats, bom (see references in Ramonas thesis).
@@
 specific features such as agreement or word order.
-One abstract grammar can be implemented by a number of concrete grammars, each
+An abstract grammar can be implemented by a set of concrete grammars, each
 representing one language. As a comparison, the abstract and concrete syntax
@@
 The example in figure \ref{fig:gfAbstract1} shows an abstract grammar defining three categories, % \verb|Categories|,
-one for nouns, one for verbs and one for sentences. An abstract grammar also gives
-the types of functions. In our case we have \verb|Pred| which
-tells us that by taking a noun and a verb we can form a sentence. At this stage, 
-no information of how this is done is given. The grammars also defines two words,
-the noun \verb|cat| and the verb \verb|sleep|.
+one for nouns, one for verbs and one for sentences. The abstract grammar also gives
+the function types. In this case we have \verb|Pred|, which tells us that by
+taking a noun and a verb we can form a sentence.  No information of how this is
+done is given, at this stage. The grammars also defines two words, the noun
+\verb|cat_N| and the verb \verb|sleep_V|.
 
@@
 We get a more complicated example if we allow the nouns to be used in both
-plural and singular. We add a category \verb|NP| to the
-abstract. Note that \verb-NP- is normally used for nouns with a fixed definitness, but for 
-now we concentrate on the number only. Hence \verb|N| now means a noun
-in any of the number, and we introduce two function, \verb|NSg : N -> NP| and
-\verb|NPl : N -> NP| which sets the number.
+plural and singular. We add a category \verb|N'| to the
+abstract, which represents a noun with a fixed number,
+%Note that \verb-N'- is normally used for nouns with a fixed definitness, but for 
+%now we concentrate on the number only. Hence \verb|N| now means a noun in any of the number,
+and we introduce two function, \verb|NSg : N -> N'| and
+\verb|N'l : N -> N'| which sets the number.
 \begin{figure}[h!]
@@
  abstract TestGrammar = {          concrete TestGrammarSwe of TestGrammar = {
-  cat N ; NP ; V ; S ;               lincat V, S, NP = Str ;
+  cat N ; N' ; V ; S ;               lincat V, S, N' = Str ;
                                             N = {s : Num => Str} ;
   fun                                lin   
-    Pred : NP -> V -> S ;              Pred n v = n ++ v ;
-    NSg : N -> NP ;                    NPl n = n.s ! Pl ;
-    NPl : N -> NP ;                    NSg n = n.s ! Sg ;
+    Pred : N' -> V -> S ;              Pred n v = n ++ v ;
+    NSg : N -> N' ;                    NPl n = n.s ! Pl ;
+    NPl : N -> N' ;                    NSg n = n.s ! Sg ;
     cat_N : N ;                        cat_N = {s = table {Sg => "katten" ;
@@
 \end{figure}
-Figure \ref{fig:gfTest2} introduces some new concepts: records,tables and parameters.
+Figure \ref{fig:gfTest2} introduces some new concepts: records, tables and parameters.
 In the concrete syntax, \verb|N| is defined to be a record consisting of the field
-\verb|s|. The type of \verb|s| shows that it is a table, which given a
-parameter of type \verb|Num| returns a string. \verb|Num| is defined as a
-parameter, which can either have value \verb|Sg|
-or \verb|Pl|. In \verb|NPl| and \verb|NSg|, \verb|n.s| means that we use the
-field \verb|s| of \verb|n|, and then the selection operator \verb|!| is used to
-select a branch in the table.
-
-If we want to implement a English version of the grammar, we encounter another
-problem: the verb form depends on the noun phrase. The \verb|NP| needs to carry
-information about which number it is in, and \verb|Pred| can pass this on to the
-verb. Hence, the \verb|V| also needs to be a table, depending on the number.
+\verb|s|. The type of \verb|s|, \verb-Num => Str- shows that it is a table, which given a
+parameter of type \verb|Num| returns a string. \verb|Num| is defined 
+%as a parameter, which can
+to either have value \verb|Sg| or \verb|Pl|. 
+The dot (\verb-.-) is used for projection and the bang (\verb-!-) as a selection operator.
+%In \verb|NPl| and \verb|NSg|,
+\verb|n.s ! Sg| thus means that we use the branch for \verb-Sg- in
+field \verb|s| of \verb|n|.
+%, and then the selection operator \verb|!| is used to select a branch in the table.
+
+When implementing a English version of the grammar, we encounter another
+problem: the verb form depends on the number of the noun. The \verb|N'| needs to carry
+information about which its number and \verb|Pred| must pass this on to the
+verb. Finally, the type of \verb|V| needs a table, depending on the number.
 \begin{figure}[h!]
@@
          N = {s : Num => Str} ;
-         NP = {s : Str ; num : Num} ;
+         N' = {s : Str ; num : Num} ;
   lin   
@@
 
-We now have two implementations of the abstract. The resulting GF grammar is able to
-both parse a string to a abstract tree and to go in the other direction; to produce
+We now have two implementations of the abstract. The resulting GF grammar is able both to
+parse a string to a abstract tree and to go in the other direction; to produce
 a string of natural language given an abstract tree. This step is called linearization.
@@
 linearize the resulting abstract tree to English.
-
-Valency information is normally built into the GF categories. To indicate that
-a verb is transitive and needs an object, we use the category \verb_V2_.
-\begin{verbatim}
-abstract:
-cat V2 ; 
-eat_V2 : V2 ;
-rat_N ;
-UseV2 : V2 -> NP -> V ;
-
-concrete:
-eat_V2 = "äta" ;
-rat_N = {s = table {Sg => "råttan" ;
-                    Pl => "råttorna"}};
-UseV2 v np = v ++ np ;
-\end{verbatim}
-
-The example above now allows ``Katten äter råttan", but not ``Katten äter".
-\verb|UseV2| concatenate the strings ``äter" and ``råttan". Note that if we
-wanted to add negation, ``katten äter inte råttan", or question "äter katten råttan",
-we cannot glue the verb to the object at once. Instead we would have to use a table,
-with one field for the verb and one for the object.
 
@@
 %>>Most general resource, first large grammar, covers fundamental syntactics.
-The GF package provides a powerful resource library \cite{gf-resource}, which covers the
-fundamental morphology and syntactics for more than 20 languages. %is a very useful part of the framework. It is provided in the 
+The GF package provides a advantageous resource library \cite{gf-resource}, covering the
+fundamental morphology and syntactics of more than 20 languages.
+%is a very useful part of the framework. It is provided in the 
 %, they give a good start for all who want to write an application grammar.
@@
 %this can be found as libraries. So far there are 20 languages in the library, and
-The languages all share the same abstract and they constitute  
-a very useful resource when implementing application grammars.
-There is also a small test lexicon included, containing some hundred of common
-words.
+The languages all share the same abstract and are valuable
+%they constitute  a very useful resource 
+when implementing application grammars. There is also a small test lexicon
+included, containing a few hundred common words.
 
@@
 languages are added, the compiler is being improved, ways of using it in more 
-effecient and easier ways are added
+effecient and easygoing manners are added
 and the possibilities to use GF in different environments
-increased. There are research on how to make better use of the dependent
-types, for using ontologies for reasoning \cite{ramona} or generating natural
+increased. There are research on how to make more use of the dependent
+types, for reasoning by using ontologies \cite{ramona} or generating natural
 language via Montague semantics \cite{montague}.
@@
 compare them against a reliable source.
-Being a freely available, manually annotated, large scale treebank,
+Being a freely available, manually annotated, large-scale treebank,
 Talbanken \cite{talbanken} was perfect for our purpose.
-It was created in the 1970s at Lund University. In total it consists of four
+%It was created in the 1970s at Lund University.
+In total it consists of four
 parts. Two of them are transcribitions of spoken language, one a collection of
@@
 Talbanken was also used to train the Swedish version of the Maltparser \cite{malt}
-and was thus redistributed in a more modern version \cite{talbanken05}.
+and was then redistributed in a more modern version, Talbanken05 \cite{talbanken05}.
 %This work was done by Nivre, Hall and Nilsson in 2005.
-It is released in Malt and Tiger
-XML-formats\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html},
+It is released in Malt\footnote{http://w3.msi.vxu.se/~nivre/research/MaltXML.html} 
+and Tiger\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html},
+XML-formats
 where the trees had been made deeper and more detailed. \\
@@
 %Look at output from Maltparser, could also comparing our parse trees to this.
+
+\section{Saldo}
+\label{sec:saldo}
+A good parser needs a good lexicon. We have used of Saldo \cite{saldo}, a
+large electronic lexicon developed and maintained at Gothenburg University. It is
+built on Svenskt Associations Lexicon and contains information about more than 
+120 000 modern Swedish words.
+For each word there is semantical, syntactical and a morphological
+information. The user can find compounding analyses, examples of usage in corpora,
+and graphs of semantically connected words.\\
+\begin{figure}[h!]
+\centering
+\subfloat[Morphological info for ``katt"]{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
+\hspace{20mm}
+\subfloat[Graph showing the hyponyms of ``katt"]{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
+\caption{Information in Saldo}
+\label{fig:saldo}
+\end{figure}
+
+Words with multiple meanings are separated inte different entries.
+%If one word has different meanings, there will be separate entries for each of them. 
+The word ``växa" (``grow") for example, has two meanings: \emph{getting
+bigger}
+\enumsentence{Gräset växte varje dag\\
+`The grass grew each day'}
+or \emph{being a plant}. 
+\enumsentence{Det växte en gammal ek där.\\
+`There stood an old oak.'}
+The word ``växa" is however declined the same way for both interpretations of the word.
+Hence Saldo's morphological lexicon, provided under LPGL in XML format,
+only keeps one entry of ``växa".
+For the homonyms that do not share the whole inflection tables, different entries are 
+kept in the morphological lexicon.
+
+For our purpose, only the morphological lexicon was needed.
+The data can be processed and translated to GF format as described in 
+section \ref{sec:prog.saldo}.
+
+\section{Swedish}
+\label{sec:swedish}
+Swedish \cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
+closely related to Norwegian and Danish. The languages share most of their
+grammatical structures and are mutually intelligible. Swedish is also 
+one of the official languages in Finland and altogether spoken by approximately 9
+million people.
+Swedish syntax is often similar to English, but the  morphology is richer and the
+word order slightly more intricate.
+
+\subsection*{Post-nominal articles} 
+Unlike most of the world's languages Swedish express not only the number but also
+the definitness of a noun by suffixes. The endings are decided by the noun's gender,
+%which may be 
+\textit{neuter} or \textit{common}.\\
+%(or \textit{utrum}) 
+\begin{tabular}{l|lll}
+&\textbf{Definite}& \textbf{Indefinite}& \textbf{Gender} \\
+ \hline
+\textbf{Singular} & {en katt} & {katten} & \textit{common}\\
+& \emph{a cat} & \emph{the cat} &\\
+\textbf{Plural} & katter & katterna &\\
+ &\emph{cats} & \emph{the cats} &\\
+ \hline
+\textbf{Singular} & {ett hus} & {huset} & \textit{neuter}\\
+ & \emph{a house} & \emph{the house} & \\
+\textbf{Plural} & hus & husen & \\
+& \emph{houses} & \emph{the houses} & \\
+\end{tabular}\\
+
+The definiteness can be accentuated by using the definite articles.\\
+\begin{tabular}{llll}
+den katten & det huset & de katterna & de husen \\
+\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
+\end{tabular}\\
+
+A definite article or a determiner is necessary when the noun is modified by adjectival
+phrase; sentence \ref{sent:trottkatt}a is not grammatically correct.
+Since adjectives also marks definitness, it is marked on three places
+in sentence \ref{sent:trottkatt}b. %, the definiteness is 
+%The definiteness in then marked by the article, on the noun and on the adjective.
+
+\enumsentence{\begin{tabular}[t]{@{}*{10}{l@{\ }}}   %\begin{tabular}[t]{@{}*{lllllll}}
+a. & *Gamla          & katten        &sov. &&b. &Den             &gamla            & katten         &sov.\\
+   & $[${\sc+def}$]$ & $[${\sc+def}$]$&    &&  & $[${\sc+def}$]$ & $[${\sc+def}$]$ & $[${\sc+def}$]$&  \\
+   &~~`The old       & cat           &slept.'&& &`The            & old             & cat            &slept.'\\
+\end{tabular}}\label{sent:trottkatt}
+However, for some determiners, ie. `min' (\emph{`my'}), the noun should be in indefinite form.
+%the definitness of the noun is set by the determiner. For some,
+%like `min' , the noun should be used in indefinite form.
+
+\enumsentence{\longexnt{3}{3}
+{Min &trötta& katt.} %&& Den där & trötta &katten}
+{`My & tired & cat'}
+{ $[${\sc+def}$]$ &$[${\sc+def}$]$& $[${\sc-def}$]$} % . && $[${\sc+def}$]$  & $[${\sc+def}$]$ &$[${\sc+def}$]$}
+{}}
+
+%a.& Katten& sov. & b
+% &$[${\sc+def}$]$ & %can be superdoubled
+%  &`The cat& slept.'%example: den där boken
+
+\subsection*{Verb second}
+Swedish is a verb-second language \cite[p.116]{gunlog}: the
+second constituent of a declarative main clause must consist of a verb.
+The normal word order is subject-verb-object, but any syntactic category can be
+fronted \cite[\textsection 1027]{H&H}.
+Topicalisation is very common, especially for temporal and spatial adverbial phrases.
+The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
+meaning, but vary in how the content is presented.
+%~~ although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being
+%emphasized.
+\enumsentence{
+\shortex{4}
+{Du & ser &  inte &mig.}
+%empha glossning!! ha samma överallt, ' eller ''
+{\emph{you} & \emph{see} & \emph{not} &\emph{ me}}
+{`You  don't see me'.}} \label{ex:swedish-svo}
+\vspace{-3mm}
+\enumsentence{
+\shortexnt{4}
+{Mig  & ser &  du & inte.}
+{\emph{me} & \emph{see} & \emph{you} &\emph{ not}}}\label{ex:swedish-ovs} 
+\vspace{-3mm}
+\enumsentence{
+\shortexnt{4}
+{Inte & ser& du &mig.}
+{\emph{not}&  \emph{see}& \emph{you}& \emph{me}}} \label{ex:swedish-avso}
+
+%Alla syntaktiska kategorier kan be fronted utom obetonade satsadverbial (ju,väl).
+%\enumsentence{
+%\begin{tabular}{lll}
+%  Johan & gick & sakta på gatan.\\
+%\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
+%  Sakta & gick & Johan på gatan.\\
+%\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
+%  På gatan & gick & Johan sakta.\\
+%\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
+%\end{tabular}}
+
+Inverted word order marks questions
+\enumsentence{
+\shortex{4}
+{Såg & du & mig & inte?}
+{\emph{saw}  & \emph{you}&  \emph{me}& \emph{not}}
+{`Didn't you see me?'}}
+%\enumsentence{Gick Johan på gatan?\\
+%        \emph{Did Johan walk on the street?}}
+The word order in subordinate clauses in also slightly modified,
+%>> and stricter
+\enumsentence{Main sentences:\\
+{\shortex{8}{Jag & såg & Anna & men & hon & såg & {\bf inte} & mig }
+{ \emph{I} & \emph{saw} & \emph{Anna} & \emph{but} & \emph{she} & \emph{saw} & {\bf \emph{not}} &\emph{ me}}
+{`I saw Anna but she didn't see me'}}}
+\enumsentence{Subordinate sentence:\\
+\shortex{7}{Jag & förstod & att& Anna& \textbf{inte}& såg& mig.}
+            {\emph{I} &  \emph{understood} & \emph{that}& \emph{Anna}& \textbf{\emph{not}}& \emph{saw}& \emph{me}}
+            {`I understood that Anna didn't see me'}}
+
+\subsection*{Passive voice}
+There are two ways of forming passive verb phrases in Swedish: the 
+\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
+%>> better with animate subjects
+\enumsentence{
+\shortex{5}
+{Tjuven &blev& tagen& av & polisen}
+{\emph{the} \emph{thief} & \emph{was} & \emph{taken} & \emph{by} & \emph{the police} }
+{`The thief was arrested by the police'}}
+\label{gfPass:peri-pass}
+and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
+\enumsentence{Passive \hspace{45mm} Active\\
+\shortexm{12}
+{Tjuven & togs&  av & polisen}
+{\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
+{`The thief was arrested by the police'}
+{&&Polisen & tog & tjuven}
+{&&\emph{the police} & \emph{took} & \emph{the thief}}
+{\hspace{-9mm}`The police arrested the thief'}}
+\label{gfPass:s-pass}
+The s-passive is the most commonly used one for both written
+and Swedish \cite[Pass. \textsection 1]{SAG} and dominates especially when the subject is inanimate.
+%cf.
+%\enumsentence{
+%\shortex{5}{Ett & lejon & jagade & honom.}
+%        {A &lion &hunted &him.}
+%        {`A lion was hunting him.'}}
+
+\subsection*{Impersonal constructions}
+Constructions with `det är/var' (\emph{`it is/was'}) is very common in Swedish \cite[\textsection 309d]{H&H}:
+\enumsentence{
+\shortex{6}{Det & var & roligt & att & höra.}
+{\emph{it} &\emph{was} & \emph{nice} &\emph{to} &\emph{hear}.}
+{`I'm glad to hear that.'}}
+'Det' (\emph{it}) is also used as formal subject in presentational constructions where
+the real subject is put in the position of an object.
+\enumsentence{
+\shortex{6}{Det &står &en &älg &på &fältet.}
+{ \emph{it} & \emph{stands} &\emph{a} & \emph{moose} &\emph{on} &\emph{the field}}
+{`There is a moose in the field.'}}
+
+\subsection*{Reflexive pronouns}
+\label{swe:refl}
+The Scandinavian language have special reflexive pronouns
+and possessive pronouns for the 3rd
+person \cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
+\enumsentence{
+\begin{tabular}{llll}
+a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
+&\emph{He hit him self.} && \emph{He saw his (own) child.}
+\end{tabular}}
+\enumsentence{
+\begin{tabular}{llll}
+a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
+&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
+\end{tabular}}
+The 1st and 2nd persons use the normal personal pronoun in object form as reflexive
+pronouns.
+
+%Mention varandra. Same idea.
+%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
+%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
+%for what will be written in the grammar part.
+%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
+%är stor, I am big
+%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
 
@@
 
-The Lingo Matrix \cite{lingomatrix}, is a starter-kit for building Head-Driven Phrase
+The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
 Structure Grammars \cite{hpsg} (HPSG) providing comptability with tools for
-parsing, evaluation, semantic representations etc.
-There is a collection of implementations done in this framework?, describing ie.
+parsing, evaluation, semantic representations etc and  
+translation is supported by using Minimal Recursion
+Semantics \cite{mrs} as an interlingua. \\
+There is a collection of grammars implemented in this framework, giving broad-coverage
+description to ie.  %broad-coverage
 English, Japanese and German. %, Greece, Spanish etc. 
@@
 grammar, also implemented using the Lingo Matrix,
-which focuses on verb frames and translation of those by using Minimal Recursion
-Semantics \cite{mrs} as an interlingua. \\
-It should be pointed out that grammars implemented using Lingo Matrix
-do not share any common abstract, like the resources grammars of GF?\\
+which focuses on verb frames and translation of those to English. \\
+
+%It should be pointed out that grammars implemented using Lingo Matrix
+%do not share any common abstract, like the resources grammars of GF?\\
 
@@
 \label{sec:prog.saldo}
-\section{Implementation}
->> syntactic information too. Mention that GF does not have this, only lexical or
->> semantic. Not 1-1.
 Saldo is compatible with GF, which makes it easy to transform the morphological
-lexicon into our format. The lexicon has been imported earlier (Angelov,2009?)
+lexicon into the GF format. The lexicon has been imported earlier (Angelov,2009)
 but the code needed updates to not miss many important words.
@@
 
-The basic algorithm for porting Saldo is the same as Angelov's, it works by
-making use of the smart paradigms in GF.
+\section{Implementation}
+%>> syntactic information too. Mention that GF does not have this, only lexical or
+%>> semantic. Not 1-1.
+The basic algorithm for importing Saldo was implemented by Angelov (2009?)
+and  produces code for a GF lexcion.
+For each word in Saldo, it decides which forms should be used as input
+to the the GF smart paradigms. For a verb, this will mean giving
+the present form. \\
+
+\begin{figure}[h]
+\verb-mkV "knyter" ;-
+\caption{First output for the irregular? verb `knyta'}
+\label{fig:saldoknyt}
+\end{figure}
+
+When compiling the resulting lexicon, it will produce an inflection table
+for every entry. The tables are compared to the information given
+in Saldo, if the tables are equal the code for the word is saved. If the table
+is erronous, another try is made
+by giving more forms to the smart pardigm.
+For example \ref{fig:saldoknyt}, the smart paradigm will fail to calculate the
+correct inflection table. In the next try both the present and the past tense
+is given:\\
+
+\begin{figure}[h]
+\verb-mkV "knyter" "knöt" ;-
+\caption{Second output for the verb `knyta'}
+\label{fig:saldoknyt2}
+\end{figure}
+The program is run iteratively until the GF table matches the one given in Saldo,
+or until there are no more ways of using the smart paradigm. The verb 'knyta'
+will need tree forms:\\
+
+\verb-mkV "knyter" "knöt" "knutit"-\\
+
 %Find the paper!
-%To understand it, we first need to look at differences between Saldo and a GF lexicon.
-The program will try to create a lexicon by finding a number of forms for every
-word and feeding them to the smart paradigms. As few forms as possible should be
-used, and the resulting table generated by GF should be equivalent to the one found
-in Saldo. This process is done iteratively,
-starting from the easiest case. For each attempt the results are compared and when
-the correct way of using a word has been found, the code is saved.
+%The program will try to create a lexicon by finding a number of forms for every
+%word and feeding them to the smart paradigms. As few forms as possible should be
+%used, and the resulting table generated by GF should be equivalent to the one found
+%in Saldo. This process is done iteratively,
+%starting from the easiest case. For each attempt the results are compared and when
+%the correct way of using a word has been found, the code is saved.
+
 Figure \ref{pic:TabVax} shows a the information given by Saldo and by GF respectively.\\
@@
 saldo, which is pronoun in SAG as well) are either Quantifiers or Dets.
+Difference, how the determinern can set the definiteness of the noun and
+the adjective.
 Once we have a NP, can use Predets. What they are.\\
+
+Valency information is normally built into the GF categories. To indicate that
+a verb is transitive and needs an object, we use the category \verb_V2_.
+\begin{verbatim}
+Abstract:
+cat V2 ; 
+eat_V2 : V2 ;
+rat_N ;
+UseV2 : V2 -> NP -> V ;
+
+Concrete:
+eat_V2 = "äta" ;
+rat_N = {s = table {Sg => "råttan" ;
+                    Pl => "råttorna"}};
+UseV2 v np = v ++ np ;
+\end{verbatim}
+
+The example above allows ``Katten äter råttan", but not ``Katten äter".
+%\verb|UseV2| concatenate the strings ``äter" and ``råttan".
+Note that if we wanted to add negation, ``katten äter inte råttan", or question
+"äter katten råttan", we cannot glue the object directly to the verb as done in
+\verb-UseV2-. Instead we would have to use a table, with one field for the verb
+and one for the object.
 
@@
 described above.
->> read cooper and say smt smart here
 There are also restrictions on the subject.
@@
 \end{tabular}}
+>> depends on the determiner, must require a DIndef? noun! (Cooper,1986)
 The noun must be in indefinite form, even though some determiners are allowed.
@@
 
+>> allowed by the presence of a restrictive relative 
+>> only definite article (determinative pronoun), not determiner
 The grammar has been extended to accept
@@
 {``The opinions presented ..."}}
-where \emph{``de"} is used together with a indefinite noun. This is only correct for 
+where the definite artcile \emph{``de"} is used together with a indefinite
+noun. This is only correct for the definite articles and not determiners
+and requires that the noun phrase is followud by a restrictive
 relative clauses or together with the verb \emph{``finnas"}.
@@
 What is still needed in the grammar, how important that is. 
+\subsection{Grammar}
+köpa bil 
+object shift
 \subsection{Parallel parsing/Chunk parsing}

### 2011-12-13 doc/report/report.tex
breport: better results-sections, mapping, saldo

@@
 The framework is also used in the
-European project MOLTO \cite{molto} for online translation between up to 15 language.\\
+European project MOLTO \cite{Molto} for online translation between up to 15 language.\\
 
@@
 XML-formats
-where the trees had been made deeper and more detailed. \\
+where the trees had been made deeper and more detailed while still containing
+the lexical MAMBA layer. \\
 In this project, as well as for the Malt parser, the more than 6000 sentences
@@
 \label{sec:related}
-There has been much research about parsing Swedish. One of the most well known
-parser is the data-driven Malt parser \cite{malt}, which has been trained on 
+%There has been much research about parsing Swedish. INTRO!!
+A well known parser is the data-driven Malt parser \cite{malt}, which has been trained on 
 Talbanken. 
 There are also a number of grammar-based parsers, although none is still freely avaible.
-CassSwe \cite{casswe} is a cascaded finite state parser giving a syntactic
-analysis.
-The Swedish Constraint Grammar 
-(Birn,1998) gives a shallow syntactic analysis.
+The cascaded finite state parser CassSwe \cite{casswe} and
+The Swedish Constraint Grammar (Birn,1998) 
+give syntactic analyses. %birn: shallow
 Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
@@
 Structure Grammars \cite{hpsg} (HPSG) providing comptability with tools for
-parsing, evaluation, semantic representations etc and  
-translation is supported by using Minimal Recursion
+parsing, evaluation, semantic representations etc.
+Translation is supported by using Minimal Recursion
 Semantics \cite{mrs} as an interlingua. \\
 There is a collection of grammars implemented in this framework, giving broad-coverage
-description to ie.  %broad-coverage
+description to %broad-coverage
 English, Japanese and German. %, Greece, Spanish etc. 
@@
 Scandiavian, while Norsource (Hellan, 2003) describes Norweigan. A Swedish version
-was based upon this (SweCore, Ahrenberg) which covers the morphology and some
+was based upon this (SweCore, Ahrenberg) covering the morphology and some
 differences between Swedish and Norweigan. Further, there is the BiTSE \cite{stymne}
 grammar, also implemented using the Lingo Matrix,
-which focuses on verb frames and translation of those to English. \\
+which focuses on describing and translating verb frames.\\ % and translating them to English. \\
 
@@
 
-The Swedish verison \cite{gamback} of the Core Language Enigne (CLE) \cite{cle}
+The Swedish verison of the Core Language Enigne (CLE) \cite{gamback} % \cite{cle}
 gives a full syntacitc analysis as well as semantics represented in xx. A
 translation for conversations?? to English was implemented. The work was further developed in the spoken
-language blabla (xx,xx), but is no longer available. The coverage of the Swedish
+language blabla (xx,xx). Unfortunally, it is no longer available. The coverage of the Swedish
 CLE is also reported to be very limited (ref).\\
@@
 for this, the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions(\emph{pp}) and numerals(\emph{nl}) etc.
-~~are assumed to be known already~~,
-and are usually not kept in the standard lexicon, but in modules for the basic
-structures of the language. Categories involving several words(\emph{vbm,nnm})
-need to be handled differently, usually as idioms. Pronouns(\emph{pn}) are not  
-analysed the same way, see section \ref{sec:swegf}. Some experiments on finding
-the right GF-category for those have been done, see ~~~~~\ref{sec:gf.quant}.
-Six categories were imported: \\
->> english examples
-\begin{tabular}{l|lll}
+Saldo in the GF lexicon. Prepositions(\emph{pp}), numerals(\emph{nl}),
+personal pronouns etc.
+are assumed to be present in the resource grammars and should not be added again.
+Saldo contains may other pronouns,
+but they are not analysed the same way in GF (see section \ref{sec:swegf}) and
+can thus not simply be extracted -- we first need to find their correct GF category.
+Some experiments on finding
+the right GF-category for those have been done using Talbanken, see p. \pageref{sec:gf.quant}.
+
+%and are usually not kept in the standard lexicon, but in modules for the basic
+%structures of the language.
+Categories involving several words (\emph{vbm, nnm})
+are usuallly handled as idioms and should be given in a separate lexicon. In
+total six categories were imported: \\
+
+\begin{tabular}{|l|lll|}
+\hline
 & Saldo tag & GF category & Example \\
 \hline
- Adverb & \textbf{ab} &\textbf{Adv} & ofta \\
- Adjective&\textbf{av} &    \textbf{A} & gul\\
- Noun & \textbf{nn} &\textbf{N} & hus\\
- Verb & \textbf{vb} &\textbf{V} & springa\\
- Reflexive verbs  &\textbf{vbm}& \textbf{V} & raka sig\\
- Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till\\
+ Adverb & \textbf{ab} &\textbf{Adv} & ofta (\emph{often})\\
+ Adjective&\textbf{av} &    \textbf{A} & gul (\emph{yellow})\\
+ Noun & \textbf{nn} &\textbf{N} & hus (\emph{house})\\
+ Verb & \textbf{vb} &\textbf{V} & springa (\emph{run})\\
+ Reflexive verbs  &\textbf{vbm}& \textbf{V} & raka sig (\emph{shave})\\
+ Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till (\emph{perk})\\
+\hline
 \end{tabular}\\
@@
 %different grammatical analyse, a lot of pronouns. Example. 
-One reason for why words of the categories listed above could not be imported
-is that Saldo does not contain information about ~~word forms that are not used~~.
-In some cases those forms are needed by the smart paradigm. 
-This was a common reason why nouns could not be imported. 
+Most, but not all words of the categories listed above have been imported.
+One reason why the importing phase would fail 
+is that Saldo, unlike GF, only contains the acctually used word forms.
+% does not contain information about ~~word forms that are not used~~.
+For technical reasons, the smart paradigm might need forms never used.
+%for this is Deponent verbs
+%are for example given with their (often non-existing) active form, al
+%In some cases those forms are needed by the smart paradigm. 
 Consider for example
 the plural tantum % \cite[44]{SAG} 
-noun ``glasögon" (``glasses"). Since the smart paradigm
-requires a singular form, this lemma could not be added to the lexicon. ``Glasögon" could
-be added manually, by using the ~~somewhat incorrect,annan betydelse~~ singular form ``glasöga" (``glass-eye").
-The same problem occurred for verbs that always are used in passive form:
->> not passive!! deponent, s-verbs!!
-``andas" (``breathe"), ``frodas" (``flourish"), ``lyckas" (``succeeed"). This
-was the case for close to 80\% of the failing verbs of type \verb_vb_.
-
-In some cases the smart paradigms could not generate the correct declination,
-although this happened seldom. 
->> exapmles: % frysa,fisa passive preterium
-
-The resulting dictionary contains more than 100 000 entries, compared to the total size
-of Saldo this is approximately 80\%. When testing the coverage of Talbanken,
-we found that around 3000 words were still missing (excluding the one tagged as
-numerals and proper names, but some names are still there). In addition, there were 1817 words
->> give reasons! should look good, not bad!
-which were given different labels in GF than in Talbanken. Many of those are
+noun \emph{``glasögon"} (\emph{``glasses"}).
+The smart paradigm requires a singular form, and since the program couldnot
+find this in Saldo, there was no way of adding the lemma to the lexicon. 
+When the program failed to import a noun, this was often the explanaiton.
+%bbMany nouns fai, \emph{``kläder"} (\emph{``clothes"} etc.
+Words of this type could be added manually, for \emph{``glasögon"} by using the
+ostensibly correct singular form ``glasöga", although this
+has another meaning (``glass-eye").
+The same problem occurred for the irregular s-verbs,
+(\emph{``synas"} \emph{(``show")} or \emph{umgås} \emph{(``socialize")})
+which made up 61.5\% of the failing verbs of type \verb_vb_.\\
+%``frodas" (``flourish"), ``lyckas" (``succeeed").
+In a few cases the smart paradigms could not generate the correct declination.\\
+
+%although this happened seldom. The passive preteritum is an exapmle of a form 
+%oftenly because the one form could not be generated. anbd for passive preteritum could not be guessed.
+%that seems hard to guess for some verb.
+% >> should we add smth for this?
+%\begin{tabular}{3}
+%&\textbf{Saldo}& \textbf{GF}
+%\textbf{Passive preteritum} & fryses & frys \\
+%\end{tabular}
+%>> exapmles: % frysa,fisa passive preterium
+
+The resulting dictionary contains more than 100 000 entries, approximately 80\% (retest!)
+of the total size of Saldo.
+%compared to the total size of Saldo this is .
+When testing the coverage of Talbanken,
+we found that there are around 2500 wordforms were still missing, excluding the one
+tagged as names and numbers. This number may seem very high, but %considering that
+4/5 of the wordforms are compounds. We should also take into consideration that 
+we cannot automatically find out how many actually stem from the same word, or
+how many abbreviation that are present (Talbanken also contains a small number of 
+misspellings, probably are enumerated among our missing words). The majority
+of the words are only used once.\\
+
+\begin{tabular}{|l|l|}
+\hline
+Missing words & $\sim$ 2500 wordforms\\
+Missing words ignoring compounds & $\sim$ 500 wordforms\\
+Missing words used more than once & $\sim$ 500 wordforms\\
+Missing words used more than once ignoring compounds & $\sim$ 150 wordforms\\
+\hline
+\end{tabular}\\
+
+%       The table suggests that many 
+%       >> 2537 without IDs and only the ones built of Alpha,only 552 without compounds, still contains abbrev.
+%       >> 494 without abbr. spelling errors
+%       >> 142 more then one time.
+%       numerals and proper names, but some names are still there). In addition, there were 1817 words
+%       >> 223 were IDS.++146 HS.
+%       >> 426 more than one time, with compounds.
+%       >> give reasons! should look good, not bad!
+A list of words that were given different labels in GF than in Talbanken has been
+composed, consisting of about 1600 entries. Many of those are
 acceptable, such as the ones in table \ref{tab:saldodiff} and reflects the difference 
@@
 \begin{figure}[h]
-\begin{tabular}{lll}
+\begin{tabular}{|lll|}
+\hline
 Word & Talbanken tag & GF category \\
@@
 få & POZP & Det \\
+\hline
 \end{tabular}
@@
 \end{figure}
-Among the words that were completely missing, more than 2000 was found to be nouns,
-many of which were compounds, and this is hard because ...~~~
 
 The lexicon does not contain any valency information since Saldo does not give any such
-information. Valencies are however crucial for GF, and by using the results from the mapping
-(see section \ref{sec:Mapping}) it should be possible to extract this information about this.
->> about what? how? strategy?
+information. Valencies are however crucial for GF, and it remains as future work to
+use the results from the mapping (see section \ref{sec:Mapping})
+to extract information about how words are normally used and how to automatically build
+this into the lexicon. 
 %The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
@@
 \chapter{Mapping}
+\label{sec:Mapping}
 >> no subjects or object, analyse is more from the lexicon,
@@
 >> mamba very used and tested for swedish, standard.
->> focus on what we DO cover.
-\label{sec:Mapping}
+>> focus on what we DO cover.\\
 Talbanken contains much valuable information about phrase structures and word 
@@
 commonly used, and hence very well-tested, for describing Swedish.
-The deepened analyse in Talbanken05 includes the lexical MAMBA layer. 
-One part of this project has focused translating trees from Talbanken to GF.
+One part of this project has focused translating trees from Talbanken to GF
+by constructing a mapping, which 
+automatically transforms trees in Talbanken format to GF abstract syntax trees.
 We hence get a comparision between the two annotaitons and at the same time
-we get ways to extract and translate information to GF notation. We have
-constructed a mapping, which 
-automatically transforms trees in Talbanken format to GF abstract syntax trees.
+we get methods to extract and translate information to GF notation. 
 Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
 of the sentence \emph{``The cat on the car gets bigger"} and the result of translating it to GF.
-During the translation all the information given in the abstract tree
-(\ref{pic:gftree}) needs to be extracted. We need to know that \emph{``katten"} is a noun
+Both the POS tags as well as the syntactic information are needed
+during the translation, so that all the information given in the abstract tree
+(\ref{pic:gftree}) can be extracted. We need to know that \emph{``katten"} is a noun
 used in definite form (\verb-DetArt-), singular (\verb-NumSg-). Even though the
-notation is changed by the transformation, the original analyse from Talbanken,
-that ``på bilen" forms a subtree which is attatched to ``katten" for example,
+notation is translated, the original analyse from Talbanken
 is preserved.
+\emph{``på bilen"} should still form a subtree attatched to \emph{``katten"}.
 %produces a \verb-NP- from ``katten" this means that
@@
 %DetCN  (DetQuant DefArt NumSg)    cat_N = katten
-Both the POS tags as well as the syntactic information are needed.
 
@@
 \subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
-\subfloat[GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
+\subfloat[The corresponding GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
 \caption{Trees for the sentence ``Katten på bilen blir större".}
@@
 
-The mapping connects GF with another annotations and gives us means to evaluate
-our parser. By both parsing a Talbanken sentence and transforming its annotated tree, we can
-easily inspect if the results are equal.
-The mapping also makes it clear which grammatical constructions that are still missing
+The mapping %connects GF with another annotations and 
+gives us means to evaluate our parser. By both parsing a Talbanken sentence and
+transforming its annotated tree, we can easily inspect if the results are
+equal.
+Additionally, the mapping clearly shows which grammatical constructions that are still missing
 from the GF grammar and shows how the GF analysis differs from the one made
@@
 
-A translation of the English Penn Treebank to the GF syntax has previously been done \cite{gfpenn}.
-By starting this from program, we now have
-a translation that works for the differently annotated Talbanken. The differences between
-Swedish and English syntax also meant that many rules had to be changed.\\
-
-The mapping works for Talbanken05 in Tiger XML-format.
-There are three kinds of tags: categories, edge labels and POS tags. 
-The categories give syntactic information such as \verb|S|, \verb|NP| and \verb|VP|.
-The edge labels give similar information with tags as \verb|SS| for subject, 
-\verb|OO| for object. There are more than 300 different pos tags. The high
-number is due to the detailedness, for example there are almost 50 tags for
-nouns, excluding proper names. The tags show definiteness, case, whether the
-word is a compound. Some words also have their own tags.\\
+The mapping is based on 
+a translation of the English Penn Treebank to the GF syntax \cite{gfpenn}.
+By modifying this program, we now have a translation that works for
+the Tiger XML-format of
+Talbanken. Adaption was required for the differences in annotation as well as 
+syntactic differences between Swedish and English.
+
+%The mapping works for Talbanken05 in Tiger XML-format.
+\section{The Talbanken annotation}
+Talbanken05 uses three kinds of tags: categories, edge labels and pos-tags. 
+While the pos-tags are reserved for words, the categories give syntactic information
+for phrases, such as \verb|S|, \verb|NP| and \verb|VP|.
+The edge labels give the grammatical functions with tags as \verb|SS| for subject, 
+\verb|OO| for object. Nearly 400 different tags are used, of which more than
+300 different are pos-tags. The high number is due to the detailedness, for
+example there are almost 50 tags for nouns, excluding proper names. The tags
+show definiteness, case and whether the word is a compound. Some words also have
+their own tags.\\
 
@@
 
+\section{The translation process}
 The translation works section-wise, ie. if it finds a tree of category \verb|S|,
@@
 %Description of most important tags from Talbanken, their translation? appendix?
-\subsubsection{Differences in the notation}
+\subsection{Differences in the notation}
 Some structural differences between GF and Talbanken were made obvious during the
@@
 A similar issue arises form the use of verb particles. A verb with a
-particle is considered as one constituent in GF, which means that the particle
-has no effect on the abstract tree. Its presence is announced by the verb
-itself, while Talbanken treats them as separated units.
+particle is treated as two independent units in Talbanken but in GF they are
+considered to be one constituent. % which means that 
+The particle has therefore no effect on the abstract tree, its presence is
+announced by the verb itself.
 %In Talbanken, verbs and particles are treated as separate units,
@@
 %This way %\enumsentence{Han gillar det\\He likes it} 
-The mapping cannot depend on the valencies given in GF only, however, 
-since the distinction between different types of verbs is not always clear in Swedish.
-The grammar rules becomes too strict for accepting most Talbanken sentences.
-For example, most transitive verbs can also be used as intransitive. Both sentences in
-example \ref{ex:mappTransitive} are grammatically correct.
-\enumsentence{
-\shortexm{11}
-{Han& sitter& och &läser.}
-{He& sits& and &reads.}{}
-{Han& sitter& och& läser& en& bok.}
-{He& sits& and& reads& a& book.}{}}
+When constructing the abstract trees, we must of course follow and use the
+GF analysis. However,
+the mapping cannot depend on the valencies given in GF only, 
+since the distinction between different types of verbs is not always clear in Swedish;
+most transitive verbs can also be used as intransitive. Both sentences in
+example \ref{ex:mappTransitive} are grammatically correct, and
+the grammar rules are simply too strict for accepting most Talbanken sentences.
+\enumsentence{\shortexm{13}
+{a. &Han& sitter& och &läser.}
+{&\emph{he}& \emph{sits}& \emph{and} &\emph{reads}}{}
+{b. &Han& sitter& och& läser& en& bok.}
+{&\emph{he}& \emph{sits}& \emph{and}& \emph{reads}& \emph{a}& \emph{book}}{}}
 \label{ex:mappTransitive}
@@
 %GF valencies prescribes.
-Another problem can be illustrated by sentence \ref{ex:mappMan}, containing
+
+Another annotational difference that turned out to be problematic for the
+translation can be illustrated by sentence \ref{ex:mappMan}, containing
 the generic pronoun \emph{``man"}.
@@
 \shortex{3}{Man & uppskattar & dem}
-{One & appreciates & them}
+{\emph{one} & \emph{appreciates} & \emph{them}}
 {``They are appreciated"}}\label{ex:mappMan} %  (s216).
+
 In Talbanken, \emph{``man"} is simply a personal pronoun, 
-in GF it is represented by using the function \verb|GenericCl|, which 
-turns a \verb-VP- into a clause and has no
-subject visible in the abstract tree. See figure \ref{fig:mappMan}.
+in GF it is represented by using the function \verb|GenericCl|, see figure
+\ref{fig:mappMan}. \verb|GenericCl| turns a \verb-VP- into a clause and has no
+subject visible in the abstract tree. 
 \begin{figure}[h]
@@
 When translating a senctence like this, it is thus not possible to simply glue the
-subject to the verb with \verb|PredVP|.\\
+subject to the verb with \verb|PredVP| and the normal top-down? approach fails.\\
 %Therefore, the translator must change the type of the 
@@
 
-One of the reasons why \textit{metas} are put in the output, is the 
+\section{Results and evaluation}
+>> more! replace words, how much better results (both grammar and mapping) 
+>> more results
+>> motivation, why so bad??\\
+The development was mostly example-driven, and at least one rule for each
+Talbanken-tag has been implemented.
+Shorter sentences have been prioritized in order to get a coverage of the most
+fundamental constructions. 
+Regression testing was used to ensure that no blabla was lost during the development. % in order to  very easy to destroy what was working before.
+
+The flat version of Talbanken05 was used for developing the mapping, but
+the program works for both the flat and the deep Tiger-XML versions.
+The deeper contains six more tags, although they are not always useful for our       
+purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
+% show piece of s5 here!
+\begin{figure}[h]
+\centering
+\subfloat[Deep]{\label{pic:mappDeep}\includegraphics[width=20mm]{nofile.png}}
+\subfloat[Flat]{\label{pic:mappFlat}\includegraphics[width=20mm]{nofile.png}}
+\caption{...}
+\label{fig:mappDeepFlat}
+\end{figure}
+
+Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
+which can group an infinite verb and an object adverbial together into a \verb|VP|.
+This information can however be extracted from the flat version, and the results
+get slightly better when using the flat version. \\
+
+When evaluating the mapping, the results depend a lot on the restrictions we
+put on the input. 
+One of the reasons why a Talbanken node cannot be translated, is the 
 use of the tags show in table \ref{fig:mapBadtag}.
+The \textbf{PU} tag is used for graphic listings, and not for fluent text.
+In GF there is natuarally no corresponding function, since
+%We are not interested in sentences made up by list, since 
+the listings are meant for making the text look nice in
+folders etc. The tags \textbf{XX} and \textbf{NP} are often used since
+Talbanken (unlike GF) makes a difference between subjects
+and objects. %(which is not needed in GF (see section \ref{sec:swegf}))
+The analyse of senctences like
+\enumsentence{För stora krav.\\
+``Too high demands."}
+contains the tags \textbf{XX} and \textbf{NP}, since it is not obvious
+whether the noun phrase is used as subject or an object.
+The tags shown above occur quite frequently in the treebank and are always translated
+to metas, which lowers our result. \\
+%How we cannot cover all cases. 
 \begin{figure}[h]
@@
 \end{figure}
-The analyse in Talbanken makes a difference between subjects
-and objects, which is not needed in GF (see section \ref{sec:swegf}).
-Senctences like
-\enumsentence{För stora krav.\\
-``Too high demands."}
-hence makes use of the tags \textbf{XX} and \textbf{NP}, since it is not obvious whether
-the noun phrase is used as subject or an object.
-The tags shown above occur quite frequently in the treebank and are always translated
-to metas. 
-%How we cannot cover all cases.
-
-\section{Results and evaluation}
->> more! replace words, how much better results (both grammar and mapping) 
->> more results
->> motivation, why so bad??
-The flat version of Talbanken05 was used for developing the mapping, but
-the program works for both versions.
-The deeper contains six more tags, but they are not always useful for our       
-purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
-% show piece of s5 here!
-\begin{figure}[h]
-\centering
-\subfloat[Deep]{\label{pic:mappDeep}\includegraphics[width=20mm]{nofile.png}}
-\subfloat[Flat]{\label{pic:mappFlat}\includegraphics[width=20mm]{nofile.png}}
-\caption{...}
-\label{fig:mappDeepFlat}
-\end{figure}
-
-Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
-which can group an infinite verb and an object adverbial together into a \verb|VP|.
-This information can however be extracted from the flat version, and the results
-get slightly better when using the flat version. \\
-
-%At least one rule  and also most of the
-The development was mostly example-driven, and at least one rule for each
-Talbanken-tag has been implemented.
-Shorter sentences have been prioritized in order to get a coverage of the most
-fundamental constructions. Further,  
-regression testing was used during the development. % in order to  very easy to destroy what was working before.
-
-When evaluating the mapping, the results depend a lot on the restrictions on the
-input. 
-The focus has been on shorter sentences, with no idioms or conjunction. We are
-not interested in sentences made up by list, since they are meant for making
-a nice graphic look in folders etc.\\
-
-If we
-assure that the lexicon know the correct word class for all lemmas involved, we
+
+Our focus has been on shorter sentences, with no idioms or conjunction.
+If we assure that the lexicon know the correct word class for all lemmas
+involved, we
 can restore more than 85\% of the nodes in the original tree.
@@
 No list items & 65\%\\
-No special tags or punctuation & 72\%\\
+No special punctuation or bad tags& 72\%\\
 Short sentences with known words & 85\%\\
@@
 \subsection{Lexicon}
+compounds
 %Speed up!

### 2011-12-13 doc/report/report.tex
report on mapping

@@
 This has meant to develop earlier techniques to fit for Swedish, create methods
-for achieving and keeping a extensive lexicon and to adapt the existing
+for achieving and keeping a large-scale lexicon and to adapt the existing
 resource grammar to fit for more language specific constrictions.
-We hence needed to extend the Swedish grammar to cover
-the language specific constructions. Further a large lexicon was
-needed, as well as a method for evaluation.
-During this project, we have worked on and explored those things and
-while identifying problems and future work.
-
+Three subgoals has been identified:
+\begin{itemize}
+\item Enhancing the Swedish GF grammar
+\item Importing the lexicon Saldo
+\item Create a translation between Talbanken and GF
+\end{itemize}
+%We hence needed to extend the Swedish grammar to cover
+%the language specific constructions. Further a large lexicon was
+%needed, as well as a method for evaluation.
+%During this project, we have worked on and explored those things and
+%while identifying problems and future work.
+
+\newpage % until it looks nicer
 \section{Outline}
@@
 The first version of the framework was implemented by the Xerox Research Center
-in Grenoble and is now mainly developed in Gothenburg. On of the biggest
+in Grenoble and it is now mainly developed in Gothenburg. One of the biggest
 achivements is a library covering the 
@@
 
-Words with multiple meanings are separated inte different entries.
+The semantical aspect of Saldo require that words with multiple meanings are
+separated into different entries.
 %If one word has different meanings, there will be separate entries for each of them. 
@@
 `There stood an old oak.'}
-The word ``växa" is however declined the same way for both interpretations of the word.
-Hence Saldo's morphological lexicon, provided under LPGL in XML format,
+``växa" is however declined the same way for both interpretations of the word.
+Hence \textit{Saldo's morphological lexicon}, provided under LPGL in XML format,
 only keeps one entry of ``växa".
 For the homonyms that do not share the whole inflection tables, different entries are 
-kept in the morphological lexicon.
-
-For our purpose, only the morphological lexicon was needed.
+kept also in the morphological lexicon.
+
+For our purpose, only this morphological section was needed.
 The data can be processed and translated to GF format as described in 
@@
 
-The lexicon does not contain any valency information since Saldo does not give any such
-information. Valencies are however crucial for GF, and it remains as future work to
-use the results from the mapping (see section \ref{sec:Mapping})
-to extract information about how words are normally used and how to automatically build
-this into the lexicon. 
+%Since Saldo does not give any valency information, neither does the imported lexicon.
+%The lexicon does not contain any information.
+Valency information, which is crucial for GF, is not given in Saldo and
+hence not in the imported lexicon. It remains as future work to find methods 
+%using the results from the mapping (see section \ref{sec:Mapping})
+to extract this information from Talbanken and to automatically build
+it into the lexicon. 
 %The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles

### 2011-12-14 doc/report/report.tex
report: pics and mapping text

@@
 %There has been much research about parsing Swedish. INTRO!!
-A well known parser is the data-driven Malt parser \cite{malt}, which has been trained on 
-Talbanken. 
-There are also a number of grammar-based parsers, although none is still freely avaible.
+Many years of interesting research has lead to many interesting language
+technology tools for Swedish.
+An example is the well-known parser is the data-driven Malt parser \cite{malt},
+which has been trained on Talbanken. 
+There are also a number of grammar-based parsers, although none is freely avaible.
 The cascaded finite state parser CassSwe \cite{casswe} and
@@
 There is a collection of grammars implemented in this framework, giving broad-coverage
-description to %broad-coverage
+descriptions of %broad-coverage
 English, Japanese and German. %, Greece, Spanish etc. 
@@
 was based upon this (SweCore, Ahrenberg) covering the morphology and some
-differences between Swedish and Norweigan. Further, there is the BiTSE \cite{stymne}
-grammar, also implemented using the Lingo Matrix,
+differences between Swedish and Norweigan. Further, there is the BiTSE 
+grammar \cite{stymne}, also implemented using the Lingo Matrix,
 which focuses on describing and translating verb frames.\\ % and translating them to English. \\
@@
 how many abbreviation that are present (Talbanken also contains a small number of 
-misspellings, probably are enumerated among our missing words). The majority
+spelling errors, probably are enumerated among our missing words). The majority
 of the words are only used once.\\
@@
 Missing words & $\sim$ 2500 wordforms\\
-Missing words ignoring compounds & $\sim$ 500 wordforms\\
+Missing words, ignoring compounds & $\sim$ 500 wordforms\\
 Missing words used more than once & $\sim$ 500 wordforms\\
-Missing words used more than once ignoring compounds & $\sim$ 150 wordforms\\
+Missing words used more than once, ignoring compounds & $\sim$ 150 wordforms\\
 \hline
@@
 \label{sec:prog.grammar}
-Although we can't expect to get a full coverage, we aim for a ~~high-quality~~
-grammar with satisfactory coverage.
+We can't expect to get a grammar giving full coverage, instead we aim for a
+grammar that gives
+a deep analysis while having a satisfactory coverage.
 We have been developing a grammar fragment which have been adapted to cover
@@
 \label{sec:Mapping}
->> no subjects or object, analyse is more from the lexicon,
->> compare to others, ask elisabet.
->> diffs: valencies, pp (gf ingen skillnad direkt/indirekt i träd, pga multiling)
->> mamba very used and tested for swedish, standard.
->> focus on what we DO cover.\\
+%>> no subjects or object, analyse is more from the lexicon,
+%>> compare to others, ask elisabet.
+%>> diffs: valencies, pp (gf ingen skillnad direkt/indirekt i träd, pga multiling)
+%>> mamba very used and tested for swedish, standard.
+%>> focus on what we DO cover.\\
 Talbanken contains much valuable information about phrase structures and word 
@@
 Both the POS tags as well as the syntactic information are needed
-during the translation, so that all the information given in the abstract tree
+during the translation, so that all the information shown in the abstract tree
 (\ref{pic:gftree}) can be extracted. We need to know that \emph{``katten"} is a noun
@@
 is preserved.
-\emph{``på bilen"} should still form a subtree attatched to \emph{``katten"}.
+\emph{``På bilen"} should still form a subtree attatched to \emph{``katten"}.
 %produces a \verb-NP- from ``katten" this means that
@@
 equal.
-Additionally, the mapping clearly shows which grammatical constructions that are still missing
+Additionally, the mapping shows which grammatical constructions that are still missing
 from the GF grammar and shows how the GF analysis differs from the one made
@@
 POS-tags may help us to automatically find the correct declination and add it to the
-lexicon. Further, our parser will need probabilities, see section
-\ref{sec:future}, of how often a function is used.\\
+lexicon. Further, our parser will need probabilities (see section
+\ref{sec:future}) of how often a function is used. The GF treebank we
+achieve from the translation is a good source for this information.\\
 %missing words. Lexical aquisation, probabilities, see section .
@@
 The mapping is based on 
-a translation of the English Penn Treebank to the GF syntax \cite{gfpenn}.
+a translation of the English Penn Treebank \cite{gfpenn}.
 By modifying this program, we now have a translation that works for
@@
 for phrases, such as \verb|S|, \verb|NP| and \verb|VP|.
-The edge labels give the grammatical functions with tags as \verb|SS| for subject, 
-\verb|OO| for object. Nearly 400 different tags are used, of which more than
+The edge labels give the grammatical functions: \verb|SS| for subject, 
+\verb|OO| for object etc. Nearly 400 different tags are used, of which more than
 300 different are pos-tags. The high number is due to the detailedness, for
@@
 \section{The translation process}
-The translation works section-wise, ie. if it finds a tree of category \verb|S|,
-it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
-can find those, they can be combine with the GF function \verb|PredVP|.
-At the lowest level in the tree we look up the word in the GF lexicon.
-It figure \ref{pic:tbtree}, the verb is followed by an adjective as a present
-participle, \verb|SP|. GF admits \emph{``bli"} to have type \verb|VA|, a verb
-with an adjective as complement, and the translator can apply the GF function
-\verb|ComplVA| to combine the verb with its object. \\
-It is not always the case that we first find the subject, then the verb and 
-finally the objects. Often the order is Verb-Subject-Object or Object-Subject-Verb. \\
-During the mapping process, the translator also needs to find out whether the sentence is
-negated and what tense is used. If it finds two trees of type, say, \verb|VP| but does not
-have any rules for combining them, the \textit{meta function} \verb-?- will be used. 
-Figure xx shows an example of this.\\
+The implementation consists of a set of rules that translates
+one node of a Talbanken tree to a GF function, possibly by describing how the
+nodes subtree should look. 
+\begin{verbatim}
+map S  = np <- map NP
+         vp <- map VP
+         return (PredVP np vp)
+\end{verbatim}
+This rule tells that we may map the category \verb-S- to the function 
+\verb-PredVP-, given that we find a noun phrase and a verb phrase in
+the subtree of \verb-S-.
+
+At every leaf a word is given, %When a leaf is reached,and ato see if
+and by using the knowledge from the pos-tag, the word is looked up in the GF
+lexicon. If the word is not found
+a \textit{meta} variable -- \textit{?} -- will be put in its place.
+The meta variable is also used when we have found two translatable trees
+but no rule that allows them to be combined.
+
+Lets look at the example in figure
+\ref{pic:tbtree}. The verb \emph{``bli"} is
+followed by an
+adjective as a present participle, \verb|SP|. GF admits \emph{``bli"} to have
+type \verb|VA|, a verb with an adjective as complement, and the translator can
+apply the GF function \verb|ComplVA| to combine the verb with its object. \\
+
+The translation covers complex noun phrases which may consists of pronouns,
+reflexive objects or common nouns and thereby contain predeterminers, 
+determiners, quantifiers and adjectival modifiers. Special rules were needed
+for each category of verb in order to find the right number of arguments.
+The mapping also covers changed word order, such as questions and topicalisation.
+
+%The translation works section-wise, ie. if it finds a tree of category 
+%it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
+%can find those, they can be combine with the GF function \verb|PredVP|.
+%It is not always the case that we first find the subject, then the verb and 
+%finally the objects. Often the order is Verb-Subject-Object or Object-Subject-Verb. \\
+%During the mapping process, the translator also needs to find out whether the sentence is
+%negated and what tense is used. If it finds two trees of type, say, \verb|VP| but does not
+%have any rules for combining them, the \textit{meta function} \verb-?- will be used. 
+%Figure xx shows an example of this.\\
 
@@
 Some structural differences between GF and Talbanken were made obvious during the
-translation. \\
+translation.
 %Since the translation is done word by word, ambiguities may confuse the translator.\\
@@
 announced by the verb itself.
+
+\begin{wrapfigure}{l}{0.22\textwidth}
+\includegraphics[width=20mm]{metatree.pdf}
+\caption{Meta tree}
+\label{pic:gfEtree}
+\end{wrapfigure}
 %In Talbanken, verbs and particles are treated as separate units,
@@
 example \ref{ex:mappTransitive} are grammatically correct, and
-the grammar rules are simply too strict for accepting most Talbanken sentences.
+the grammar rules are simply too strict for accepting many Talbanken sentences.
 \enumsentence{\shortexm{13}
@@
 When translating a senctence like this, it is thus not possible to simply glue the
-subject to the verb with \verb|PredVP| and the normal top-down? approach fails.\\
+subject to the verb with \verb|PredVP|. For each similair GF construction, an extra ruled
+would be needed to get full coverage. \\ %and the normal top-down? approach fails.\\
 %Therefore, the translator must change the type of the 
@@
 \section{Results and evaluation}
->> more! replace words, how much better results (both grammar and mapping) 
->> more results
->> motivation, why so bad??\\
+%>> more! replace words, how much better results (both grammar and mapping) 
+%>> more results
+%>> motivation, why so bad??\\
 The development was mostly example-driven, and at least one rule for each
@@
 fundamental constructions. 
-Regression testing was used to ensure that no blabla was lost during the development. % in order to  very easy to destroy what was working before.
+Regression testing was used to ensure that a new rule wouldn't violate the old ones
+or overgenerate to allow trees to be translated in an erronous way.
 
@@
 
-When evaluating the mapping, the results depend a lot on the restrictions we
+When evaluating the mapping, the results strongly depend on which restrictions we
 put on the input. 
-One of the reasons why a Talbanken node cannot be translated, is the 
+One of the reasons why a node cannot be translated, is the 
 use of the tags show in table \ref{fig:mapBadtag}.
 The \textbf{PU} tag is used for graphic listings, and not for fluent text.
-In GF there is natuarally no corresponding function, since
+In our grammar there is natuarally no corresponding function, 
 %We are not interested in sentences made up by list, since 
 the listings are meant for making the text look nice in
-folders etc. The tags \textbf{XX} and \textbf{NP} are often used since
+folders etc and out of scope for the grammar itself. The tags \textbf{XX} and \textbf{NP} are often used since
 Talbanken (unlike GF) makes a difference between subjects
@@
 whether the noun phrase is used as subject or an object.
-The tags shown above occur quite frequently in the treebank and are always translated
+The tags shown in table \ref{fig:mapBadtag} occur quite frequently in the treebank and are always translated
 to metas, which lowers our result. \\
@@
 65\% coverage.
-If we test randomly collected sentences that does not contain and any of the tags listed
+If we test randomly collected sentences that does not contain any of the tags listed
 in figure \ref{fig:mapBadtag} 72\% can be restored. \\
 
-\begin{tabular}{ll}
+\begin{tabular}{|ll|}
+\hline
 No list items & 65\%\\
@@
 Short sentences with known words & 85\%\\
-\end{tabular}
+\hline
+\end{tabular}\\
+
+We believe that the results could be enhanced by simply adding more rules 
+and in this way get a wider coverage. There are many special cases that require
+special rules. Since we are not aware of any formal specification of how the tags may be
+combined in Talbanken, the only way of finding all possibilities are to manually look
+for different patterns. Another option would be to make the mapping more
+robust, but the robustness must not interfere with the correctness.
 

### 2011-12-14 mapping/GraphTree.hs
dot file test

@@ -22,3 +22,10 @@ dotTree t name = header ++ body ++ footer where
    correct '?' = 'Q'
    correct ' ' = '_'
    correct x   = x
+
+
+test :: Tree String
+test = Node "UseCl" [Node "TTAnt" [Node "TPres" [], Node "ASimul" []] , 
+                     Node "PPos" [],Node "PredVP" [Node "everybody_NP" [],
+                      Node "ComplSlash" [Node "SlashV2a" [Node "eat_V2" []]],
+                       Node "?" [Node "PrepNP" [Node "on_Prep" [], Node "?" []]]]]

### 2011-12-14 doc/report/report.tex
newer report

@@
 and the grammar now covers ??\% of the xx part of Talbanken. 
+66\% av easyeasy.
 % The mapping between the annotations of Talbanken and GF is automatic. 
@@
 \subsection{A second future tense: ``Kommer att"}
+Swedish and Norwegian have two ways of forming future tense. The standard
+form 'ska' is included in the resources as the translation to
+'will'. The second tense 'kommer att' is also very commonly used, and
+desereved to be covered by our grammar.
+Komma is used when we believe that we are sure that it will happen:
+'det kommer bli mörkt snart'
+and ska more of planning
+'jag ska äta mat nu'.
+since we believe it will happend, it can not be as in :
+'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
+
+Since two out of the three Scandinavian languages share this tempus, it was 
+added to the Scandiniavin extra module.
+"jeg kommer til å savne deg"
+
+A table of result can be overview in figure \ref{fig:kommeratt}
+\begin{figure}[h]
+\begin{verbatim}
+s SFutKommer Simul Pos Main : jag kommer att se henne
+s SFutKommer Simul Pos Inv  : kommer jag att se henne
+s SFutKommer Simul Pos Sub  : jag kommer att se henne
+s SFutKommer Simul Neg Main : jag kommer inte att se henne
+s SFutKommer Simul Neg Inv  : kommer jag inte att se henne
+s SFutKommer Simul Neg Sub  : jag inte kommer att se henne
+s SFutKommer Anter Pos Main : jag kommer att ha sett henne
+s SFutKommer Anter Pos Inv  : kommer jag att ha sett henne
+s SFutKommer Anter Pos Sub  : jag kommer att ha sett henne
+s SFutKommer Anter Neg Main : jag kommer inte att ha sett henne
+s SFutKommer Anter Neg Inv  : kommer jag inte att ha sett henne
+s SFutKommer Anter Neg Sub  : jag inte kommer att ha sett henne
+\end{verbatim}
+\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}
+\label{gfSwe:parsetree}
+\end{figure}
+
 kommer att - for scand, extra tense, example
@@
 -- or is this really just AdVs? 
+-- redan is Adv in abstract. Wrong!!
+-- what about 'i regel'?
 to change position of adv:
@@
 >> CN should have dettyp too, 'avståndet till sin bil'
+\verb-
+PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
+  ajaj unknown here!!
 
@@
 \section{Testing}
-Hard do test, can't verify. Regression testing. Talbanken is still too hard.
-While developing: regression, 150 sentences to start with. Made longer, both good and bad,
-to see how many ways things can be parsed.
-Test how many parse trees that are generated.
-Look at tables for a VP tex.
-Some results.
+For testing has been carried out by using a treebank, consisting of sentences used in 
+sub-testsuites. This way we make sure that we doesn't decrease the coverage and will
+notice if the grammar starts to overgenerate. Also have some sentences that should not
+be parsed. 
+When adding new constructions, the table-function is very helpful, shows how a word will
+turn out in differnet cases. all tenses, for komma.
 
@@
 
-\chapter{Mapping}
+\chapter{Extracting a GF treebank} %bootstraping/extracting a GF treebank
 \label{sec:Mapping}
@@
 \section{Evaluation}
+The grammar in it self is not at all robust, and one unknown word will cause a whole sentence to 
+fail. When parsing text, we get low results, for this reason. Not yet interesting to talk about
+coverage, but focus on quality and ability to scale up. Our goal standarn is so far Talbanken,
+but considering that the low coverage of the grammar without statisticts, a comparison would not be
+describing anything, to little indata.
+Instead we have the luck to have an expert and her presence has increased the standard.
+
 %Results from mapping, can probably be improved, bit by bit, depends much on words.

### 2011-12-14 doc/report/report.tex
report: structuring grammar part

@@
 \usepackage[utf8]{inputenc}
+\usepackage{ amssymb }
 \usepackage{graphicx}
@@
 \subsection*{Post-nominal articles} 
+\label{sec:swedishnoun}
 Unlike most of the world's languages Swedish express not only the number but also
@@
 \label{sec:prog.saldo}
-Saldo is compatible with GF, which makes it easy to transform the morphological
-lexicon into the GF format. The lexicon has been imported earlier (Angelov,2009)
-but the code needed updates to not miss many important words.
-Experiments have been made to use a interactive tool for lexical acquisition,
-but the dictionary was still much to small. The acquisition tool may still
-be used for complementing the lexicon.
+The lexicon provided with the GF resources is far too small for open-domain parsing.
+Experiments have been made to use an interactive tool for lexical acquisition,
+%but the dictionary was still much to small. The acquisition tool may still
+but this should be used for complementing rather than creating the lexicon.
+This section describes the process of importing Saldo, which is 
+compatible with GF, and easily translated to GF format.
+As Saldo is continously updated, the importing process has been designed to be fast
+and stable enough to be redone any time.
 
@@
 For each word in Saldo, it decides which forms should be used as input
-to the the GF smart paradigms. For a verb, this will mean giving
+to the the GF smart paradigms. For a verb, this will in most cases mean giving
 the present form. \\
@@
 \verb-mkV "knyter" ;-
-\caption{First output for the irregular? verb `knyta'}
+\caption{First code produced for the verb `knyta'}
 \label{fig:saldoknyt}
@@
 As far as possible, the program makes up for this by only comparing the overlapping forms
-and only requiring that GF generates one variant whenever alternatives are given. 
-The program continues to generate new lexicons as long as there are more
-paradigms to choose from and compare the results after each run.
-
-The program has been made more robust than the previous version, it prints
-logs and finally lists of the words that were not imported. The logs provides
-information about which paradigms that were used for each word and the result
-from comparing the inflection tables.
-During the porting of Saldo, some renaming of the lemmas was needed.
-Special characters like \emph{å,ä,ö} should not occur in GF function names.
-Those are translated into \emph{aa,ae,oe} respectively. 
-The Saldo identifier \verb|äta..vb| is thus turned into the GF identifier \verb|aeta_V|.
-Since there might be two lemmas that share the same
-identifier after the translation, a number is added to the name whenever a translation
-have been done: 
+and only requiring that GF generates one variant whenever alternatives are given. \\
+%The program continues to generate new lexicons as long as there are more
+%paradigms to choose from and compare the results after each run.
+
+The program has been made more robust than the previous version. It also prints
+log files providing
+information about the process of importing each word: which paradigms that have been
+tried, the results from the comparisons of the inflection tables and finally listings
+of the words that could not be imported.
+
+Each entry in Saldo has an identifier, eg. \verb-äta..vb-, which are used as 
+constant names in the GF lexicon. The identifiers given i Saldo may contain
+special characters that should be avoided in constants in GF. The importation therefore
+needed some renaming. The Swedish letters \emph{å,ä,ö} are translated into
+\emph{aa,ae,oe} respectively. 
+\begin{verbatim}äta..vb  -> aeta_V \end{verbatim}
+The translation may cause two or more lemmas to share the same
+GF identifier, and to avoid name collition a number is added to the name
+whenever a translation have been done: 
 \begin{tabular}{lll}
@@
 entrecôte &$\rightarrow$ & entrecoote\_1 \\ 
-\end{tabular}
+\end{tabular}\\
 
@@
 %choices for adjectives vs. verbs for particips
-The importing can be redone anytime.
 
@@
 \label{sec:prog.grammar}
-We can't expect to get a grammar giving full coverage, instead we aim for a
-grammar that gives
-a deep analysis while having a satisfactory coverage.
-We have been developing a grammar fragment which have been adapted to cover
-constructions used in Talbanken. The starting point has been the Swedish
-resource grammar and the new grammar should still be compatible with this.
-%A description of the Swedish implementation of the resource grammar is given
-%in the next section, while
-Before describing what has been added, will give a short introductions to the
-the cats of resource grammar and the sw implementation in particular.
-This is done in section \ref{sec:swegf} while section \ref{sec:Added} shows and
+An important part of this project has been to developed a grammar and
+adapting it to cover
+constructions used in Talbanken. A grammar can never be expected to give
+full coverage of a language, but we aim for a grammar fragment which gives a deep
+analysis of the most important Swedish constructions.
+The starting point has been the GF resource grammar and our new grammar is
+still compatible with this.
+Before describing the work progress we will give an introduction to the
+resource grammars in general and to the Swedish implementation in particular.
+This is found in section \ref{sec:swegf} while section \ref{sec:Added} shows and
 explains the development during this project.
-%Explain what's in the resources, constructions present in most languages. 
+
+%We can't expect to get a grammar giving full coverage, instead we aim for a
+%grammar that gives
+%a deep analysis while having a satisfactory coverage.
+%The starting point has been the Swedish
+%resource grammar and the new grammar should still be compatible with this.
+%%A description of the Swedish implementation of the resource grammar is given
+%%in the next section, while
+%Before describing what has been added, will give a short introductions to the
+%the cats of resource grammar and the sw implementation in particular.
+%This is done in section%Explain what's in the resources, constructions present in most languages. 
 %Been working on a grammar fragment based on Talbanken,
@@
 >> no subjects or object, analyse is more from the lexicon,
->> compare to others, ask elisabet.
+>> compare to others, ask elisabet.\\
 \label{sec:swegf}
@@
 
-In figure \ref{fig:gftree1} we saw how the sentence ``Jag ser dig" is parsed.
-Discuss questions, how are word order handled here? Maybe negation and anterior.
-Let's look on how post-determination (sec ref) is implemented, we know that the
-smart paradigms can handle the noun declination, now show 'katten' vs. 'den trötta katten'.
-A N carries the inflection information, turned in to CN by setting number. Example, type
-of N and CN. A CN may then be modified by APs, set field for this. When turned into NP, 
-definitness is set, depends on the isMod -field. NP has case as a parameter.
-\verb|DetNP : Det -> CN -> NP ; |
-To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
-but allow variation in.. 
-To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
-Is a quantifier, hence does not require number, is given later.
-%What is the difference?
+Section \ref{gf:swe} contained an easy example with nouns. We used the categories \verb-N-
+and \verb-N'- to separated between nouns with repsectively without a set number.
+The full analysis includes a separation between the categories \verb-N- -- simple nouns --
+and 
+\verb-CN-, common noun phrases which may be modified by adjectival phrases or conjucted.
+A \verb-CN- may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
+the number and definitness. Since
+all morphological information is generated by the smart paradigms and contained in the \verb-N-
+the construction of a \verb-NP- more or less consists of putting the pieces together.
+The two basic articles are \verb-IndefArt- and \verb-DefArt- which
+uses the post-modified form of the noun: katten.
+However, we need to consider the rules described in section \ref{sec:swedishnoun}. If we have
+a common noun phrase consisting of the parts \emph{`liten'} (\emph{`small'}) and \emph{`katt'}
+(\emph{`cat'}), there are three ways they can be combined:
+\begin{verbatim}
+DetNP : Det -> CN -> NP ;
+        en    liten katt     -- DIndef
+        den   lilla katten   -- DDef Def
+        min   lilla katt     -- DDef Indef
+\end{verbatim}        
+Which combination that shoud be used is decided by the determiner. Hence all determiners
+in our grammar must keep information of which definitness they allow. They record field
+has a field storing this and the parameters are called \verb-DIndef-, 
+\verb-DDef Indef- and \verb-DDef Def-. The determiners have arguments telling
+which number the noun phrase should be in.
+We also need to make sure that we do not allow constructions as given in
+\ref{sent:trottkatt}a.
+The definite article  \emph{``en"}is considered to be a quantifier, in plural it
+is the either \emph{``de"} or nothing:
+\begin{verbatim}
+DetNP : Det              -> CN        -> NP ;
+        definite, plural +  katt      =  katter
+        definite, plural +  stor katt =  de stora katterna
+\end{verbatim}        
+This difference depends on the common noun phrase, which must carry information
+about whether it has been modified or not.
+The resource grammar have differenciate between quantifiers -- \verb-Quant-,
+determiners \verb-Det- and predeterminers. Predeterminers modify \verb-NP-s
+while the other to modifies \verb-CN-s. The differences are shown in table
+\ref{tab:detquant}
+\begin{figure}[h]
+\begin{tabular}{l|lll}
+                        &\textbf{Has number}& \textbf{Has definitness}& Example  \\ 
+\hline
+\textbf{Predeterminers} & --  & --  & alla katter, all maten \\ 
+\textbf{Quantifiers}&-- &\checkmark  & min katt, mina katter, *min katten \\ 
+\textbf{Determiners}   &\checkmark &\checkmark & varje katt, *varje katten  \\ 
+\end{tabular}
+\caption{}\label{tab:detquant}
+\end{figure}
+
 In gf, only personal pronouns are considered to belong to PN. The rests (example from
 saldo, which is pronoun in SAG as well) are either Quantifiers or Dets.
-Difference, how the determinern can set the definiteness of the noun and
-the adjective.
-Once we have a NP, can use Predets. What they are.\\
+
+%In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
+%ser dig".
+%Discuss questions, how are word order handled here? Maybe negation and anterior.
+%Let's look on how post-determination (sec ref) is implemented, we know that the
+%smart paradigms can handle the noun declination, now show 'katten' vs. 'den trötta katten'.
+%A N carries the inflection information, turned in to CN by setting number. Example, type
+%of N and CN. A CN may then be modified by APs, set field for this. When turned into NP, 
+%definitness is set, depends on the isMod -field. NP has case as a parameter.
+%\verb|DetNP : Det -> CN -> NP ; |
+%To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
+%but allow variation in.. 
+%To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
+%Is a quantifier, hence does not require number, is given later.
+%What is the difference?
+%Difference, how the determinern can set the definiteness of the noun and
+%the adjective. Ref (Cooper,1986)
+%Once we have a NP, can use Predets. What they are.\\
 
@@
 
-Since two out of the three Scandinavian languages share this tempus, it was 
+Since two out of the three Scandinavian languages share this tense, it was 
 added to the Scandiniavin extra module.
@@
 \end{verbatim}
-\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}
+\caption{The covered and accepted usage of future modal verb 'komma'}
 \label{gfSwe:parsetree}
 \end{figure}
-
-kommer att - for scand, extra tense, example
-example table
 
@@
 \emph{`en fågel'} the one of an object.
+
 Transitive verbs may not be used like this
->> elisabet disagrees: use det äter en fågel ost på taket
-\enumsentence{\shortexnt{6}
-{*Det & gillar & en & fågel & på & taket}
-{It & likes & a & bird & on & the roof"}}
+\enumsentence{\shortexnt{7}
+{*Det & äter & en & fågel & frön & på & taket}
+{it & eats & a & bird & seeds & on & the roof}}
 
@@
 A very common example of this is sentences with the verbs \emph{finnas},
-\emph{saknas} and \emph{fattas}. > engelska.
+(\emph{exist})
+\emph{saknas} (\emph{miss}) and \emph{fattas}(\emph{lack}).
 \enumsentence{\shortex{3}
@@
 {Det & fattas & kaffe}
-{It & misses & coffee}
+{It & lacks & coffee}
 {''There is no coffee"}}
@@
 described above.
-There are also restrictions on the subject.
+There are also restrictions on the real subject.
 \enumsentence{\begin{tabular}{ll}
@@
 \end{tabular}}
->> depends on the determiner, must require a DIndef? noun! (Cooper,1986)
-The noun must be in indefinite form, even though some determiners are allowed.
-\enumsentence{Det sitter några fåglar på taket\\
-              ``There are some birds on the roof''}
-\enumsentence{*Det sitter fåglarna på taket\\
-``There are the birds on the roof''}
+Expressed in GF terms, the determiner must have definitness DIndef. That is, all determiners
+that requireds the noun and the adjective to be used in indefinite form may be used.
+%The quantifier must have determiner must havenoun must be in indefinite form,
+%even though some determiners are allowed.
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. & Det sitter några fåglar på taket    & & b. & *Det sitter fåglarna på taket\\
+   & ``There are some birds on the roof''& &    & ~``There are the birds on the roof''\\
+\end{tabular}}\label{sent:faglar}
+Sentence \ref{sent:gamlafagel}b exemplifies a determiner that uses the noun in indefinite form,
+but the adjective in definite form.
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. & Det sitter en gammal fågel på taket.& & b. & * Det sitter min gamla fågel på taket.\\
+   & ``There are some birds on the roof''& &    &~ ``There is my old bird on the roof.''\\
+\end{tabular}}\label{sent:gamlafagel}
+
 %%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
@@
 %%Semantic difference, atm we allow any NP. \\
-The GF grammar does not make any distinction between indefinite and definite noun
-phrases. Therefore this is handled by looking at the determiner. \\
+To form a clause with formal subject, we hence need to inspect the determiner.
 \verb|FormalSub : SimpleVP -> Det -> CN -> Cl ;|\\
@@
 If the determiner requires a noun in determined form, the clause is put to NONEXIST.
-However, this is not an ideal solution. There are other examples of when we
-need to know the definiteness of the noun phrase, for example when using
-the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
-or when using some predeterminers (see section `Quantifiers'). Not true, not the same :(
+However, this is not an ideal solution, but it cannot be seen on the type level
+which definiteness a noun phrase has and we therefore need to do it at the runtime? level.
+
+% move to future?
+As a future direction it would be intresting to examine if the grammars noun phrases
+should have more information on the type level. In the reflexive objects are added in (see p.
+\pageref{swe:refl}) dependent types are used for showing if a phrase can be used as
+subject. We would also like to produce the sentences, where `av' should be used
+only when the noun has an explicit article or determiner.
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. &De flesta hästarna  && b.& De flesta \textbf{av} de där hästarna.\\
+ &``Most of the horses"  && b.&  ``Most of those horses."\\
+\end{tabular}}\label{sent:mosthorses}
+
+%There are other examples of when we need to know the definiteness of the noun phrase 
+%%, for example when using %the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
+%or when using some predeterminers (see section `Quantifiers'). Not true, not the same :(
 
@@
 
-\textbf{Genitive} -problem 1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
+\textbf{Genitive}
+the extramodule contained GenNP for costructing genitvives. In Swedish
+done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
+severely overgenerating when started, and had to be restricted. Cycle:
+GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP
+      katt -> kattens      ->          kattens           det är kattens
+Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
+but still much better then the overgeneration.
+Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
+min with GenNP applied n number of times. solutions: bind s.
+or: A field should also been added to Quants, so that they can have a gen-form
+ : somligas.
+1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
            fixing this include restructuring all???
-\textbf{Det-utr}. not effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
+
+\textbf{Det-utr}. no effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
  Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
@@
 %both \emph{sin} and , eg. \emph{sin}.
-the possessive pronouns as described in section \ref{swe:refl}.
+the possessive pronouns as described in p. \pageref{swe:refl}.
 Those make up noun phrases which are not allowed in subjects, 
@@
 
->> remove stupid examples, add 'hon och sin vän åkte tåg'
->> use self instead of his.
 \enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
-He saw his children's shoes.}
+He saw \sc{self's} children's shoes.}
 \enumsentence{Sina vantar hade han glömt på tåget.\\
-His gloves, he had forgotten on the train.}
+\sc{self's} gloves, he had forgotten on the train.}
 \enumsentence{Han ger sina pengar till sina syskon.\\
-He gives his money to his children.}
+He gives \sc{self's} money to \sc{self's} children.}
 \enumsentence{Han är längre än sin kompis.\\
-He is taller than his friend.} \label{ex:ReflAP}
+He is taller than \sc{self's} friend.} \label{ex:ReflAP}
 \vspace{-2mm}
 \enumsentence{Han är här oftare än sin kompis.\\
-He is here more often than his friend.} \label{ex:ReflAdv}
+He is here more often than \sc{self's} friend.} \label{ex:ReflAdv}
 \vspace{-2mm}
 \enumsentence{Hon tyckte om skolan och alla sina elever.\\
-She liked the school and all her students.}\label{ex:ReflConj}
+She liked the school and all \sc{self's} students.}\label{ex:ReflConj}
 \enumsentence{Han såg sina få böcker och sin penna.\\
-He saw his few books and his pencil.}\label{ex:ReflConj2}
+He saw \sc{self's}his few books and \sc{self's} pencil.}\label{ex:ReflConj2}
 \vspace{-2mm}
 \enumsentence{Hon ber alla sina kompisar att gå\\
-She asks all her friends to leave}\label{ex:ReflPredet}
+She asks all \sc{self's} friends to leave}\label{ex:ReflPredet}
 \enumsentence{Jag vill att han rakar sig.\\
-I want him to shave.}
+I want him to shave \sc{self's}.}
 but not
 \enumsentence{*Sina vantar var kvar på tåget.\\
-Her gloves were left on the train.}\label{ex:ReflSubj}
+\sc{self's} gloves were left on the train.}\label{ex:ReflSubj}
+\enumsentence{*Hon och sin kompis läser en bok.\\
+She and \sc{self's} friend are reading a book. }
 \enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give his money to his siblings. }
+I give \sc{self's} money to \sc{self's}his siblings. }
 \enumsentence{*Han vill att jag rakar sig.\\
-He wants me to shave himself.}
+He wants me to shave \sc{self's}.}
 
@@
 >> CN should have dettyp too, 'avståndet till sin bil'
-\verb-
-PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
+\verb- PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
   ajaj unknown here!!
@@
 
->> allowed by the presence of a restrictive relative 
->> only definite article (determinative pronoun), not determiner
 The grammar has been extended to accept
@@
 
+% compared to penn treebank: 82 tags. talbanken: at least 130
+% 99.1 % of Penn sentences have 6 or fewer clauses
+
 %If we don't take the verb valency into account, allow all verbs to have Gf category V,
@@
 compounds
-%Speed up!
+Speed up! : perhaps have one lexicon for morph. analyse and one for parsing.
+exchange words.
 %The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.

### 2011-12-15 doc/report/report.tex
report on grammar

@@
 Outside of this project, an equivalent project for English is being worked upon,
-as are tehcniques for making the parser robust.
+as are techniques for making the parser robust.
 With this in mind, we have extended the current GF grammar to adapt it for
 Swedish specifically so that it can parse an extended set of basic
-constructions. We have completed a method for importing the electronical 
+constructions. We have completed a method for importing the electronic 
 lexicon Saldo to a GF format and established a connection to the treebank Talbanken.
@@
 % more intro to gf? Good starting sentence.
-The grammar formalism GF has so far been succesfull used for 
-controlled languges, but recent experiments have showed
+The grammar formalism GF has so far been successful used for 
+controlled languages, but recent experiments have showed
 that it is also possible to use the framework for parsing free language.
-Parsing, that is the task of automatically identifiyng morpohlogical and
+Parsing, that is the task of automatically identifying morphological and
 syntactical structures to text, is subject to increasing interest, especially
@@
 a parser or grammar is not freely available does not only restrict its
-usage but also its possiblities of being developed. Our
+usage but also its possibilities of being developed. Our
 goal is
-to create a open source Swedish grammar and parser that can be continuosly enhanced and
+to create a open source Swedish grammar and parser that can be continuously enhanced and
 made use of in other projects.\\
-To build a parser or a grammar from scratch would be not only time conusming but
-also a great theoretical and techniqual challenge. In order to not reinvent the wheel
+To build a parser or a grammar from scratch would be not only time consuming but
+also a great theoretical and technical challenge. In order to not reinvent the wheel
 we develop our parser by combining well-tested resources.
@@
 a well-defined system for describing grammar, as well as a strong connection to
-more than 20 other langugages implentented in the framework. Further, we use
+more than 20 other languages implemented in the framework. Further, we use
 the extensive lexicon Saldo and the treebank Talbanken.
@@
 resource grammar to fit for more language specific constrictions.
-Three subgoals has been identified:
+Three sub-goals has been identified:
 \begin{itemize}
@@
 In previous work\footnote{??}, the start on extending the Swedish GF grammar was
-made and a tool for lexical aquistion was developed.
+made and a tool for lexical acquisition was developed.
 We now construct a bigger
@@
 %generalized context free grammar (GCFG) it is hence
-stronger than mildly contex-free grammars. The framework has  hence the same expressiveness
+stronger than mildly context-free grammars. The framework has  hence the same expressiveness
 as Tree Adjoining Grammars (Joshi,1975) and Head Grammars (Pollard,1984).
@@
 in Grenoble and it is now mainly developed in Gothenburg. One of the biggest
-achivements is a library covering the 
+achievements is a library covering the 
 basic morphological and syntactical structures of more than
@@
 The interpreter, the GF shell, allows
-the user to test grammars by commands for parsing, vizualization of parse trees,
+the user to test grammars by commands for parsing, visualization of parse trees,
 random generation, word alignment, morphological quizzes etc.
@@
 %the visualized parse tree. This is the case for the 
-eg. the definitnes and number of the noun \emph{`katten'}, which is shown as
+eg. the definiteness and number of the noun \emph{`katten'}, which is shown as
 \verb-DetQuant DefArt NumSg- in the abstract tree.
@@
 can be limited and we can ensure that the semantics is preserved during
-translation. Inhereted ambiguities and the ones arising from multilinguality will
+translation. Inherited ambiguities and the ones arising from multilinguality will
 of course remain, but can be controlled more easily. %Example of domainspecific word?
 The use of controlled language thus gives a possibility of good and reliable translation
-and GF has succesfully been used in project like %Some examples of projects using GF are 
+and GF has successfully been used in project like %Some examples of projects using GF are 
 WebAlt \cite{webalt}, a project which aims to develop language independent
 material for mathematics problems, and the formal software verification tool
-KeY \cite{key}, in wich GF is used for translating specification to natural language.
+KeY \cite{key}, in which GF is used for translating specification to natural language.
 %>> key, talk, hats, bom (see references in Ramonas thesis).
@@
 and we introduce two function, \verb|NSg : N -> N'| and
-\verb|N'l : N -> N'| which sets the number.
+\verb|N'Pl : N -> N'| which sets the number.
 \begin{figure}[h!]
@@
 languages are added, the compiler is being improved, ways of using it in more 
-effecient and easygoing manners are added
+efficient and easygoing manners are added
 and the possibilities to use GF in different environments
@@
 120 000 modern Swedish words.
-For each word there is semantical, syntactical and a morphological
+For each word there is semantic, syntactical and a morphological
 information. The user can find compounding analyses, examples of usage in corpora,
@@
 
-The semantical aspect of Saldo require that words with multiple meanings are
+The semantic aspect of Saldo require that words with multiple meanings are
 separated into different entries.
@@
 Unlike most of the world's languages Swedish express not only the number but also
-the definitness of a noun by suffixes. The endings are decided by the noun's gender,
+the definiteness of a noun by suffixes. The endings are decided by the noun's gender,
 %which may be 
@@
 phrase; sentence \ref{sent:trottkatt}a is not grammatically correct.
-Since adjectives also marks definitness, it is marked on three places
+Since adjectives also marks definiteness, it is marked on three places
 in sentence \ref{sent:trottkatt}b. %, the definiteness is 
@@
 There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using a modale auxiliary verb \\
+\textbf{periphrastic passive}, formed by using a modal auxiliary verb \\
 %>> better with animate subjects
@@
 which has been trained on Talbanken. 
-There are also a number of grammar-based parsers, although none is freely avaible.
+There are also a number of grammar-based parsers, although none is freely available.
 The cascaded finite state parser CassSwe \cite{casswe} and
@@
 Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
-(Tapanainen and Järvinen,1997), an extension of the Constraint Gramman
-formalism, and produces a dependecy structure focusing on finding the nominal
+(Tapanainen and Järvinen,1997), an extension of the Constraint Grammar
+formalism, and produces a dependency structure focusing on finding the nominal
 arguments. \\
@@
 The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
-Structure Grammars \cite{hpsg} (HPSG) providing comptability with tools for
+Structure Grammars \cite{hpsg} (HPSG) providing comparability with tools for
 parsing, evaluation, semantic representations etc.
@@
 The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
-Scandiavian, while Norsource (Hellan, 2003) describes Norweigan. A Swedish version
+Scandinavian, while Norsource (Hellan, 2003) describes Norwegian. A Swedish version
 was based upon this (SweCore, Ahrenberg) covering the morphology and some
-differences between Swedish and Norweigan. Further, there is the BiTSE 
+differences between Swedish and Norwegian. Further, there is the BiTSE 
 grammar \cite{stymne}, also implemented using the Lingo Matrix,
@@
 
-The Swedish verison of the Core Language Enigne (CLE) \cite{gamback} % \cite{cle}
-gives a full syntacitc analysis as well as semantics represented in xx. A
+The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
+gives a full syntactic analysis as well as semantics represented in xx. A
 translation for conversations?? to English was implemented. The work was further developed in the spoken
-language blabla (xx,xx). Unfortunally, it is no longer available. The coverage of the Swedish
+language blabla (xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
 CLE is also reported to be very limited (ref).\\
 
-In the TAG formalism (Joshi,1975), there are projects on getting open source, widecoverage grammars
+In the TAG formalism (Joshi,1975), there are projects on getting open source, wide-coverage grammars
 for ie. English and Korean, but, to our knowledge, not for Swedish.  
@@
 compatible with GF, and easily translated to GF format.
-As Saldo is continously updated, the importing process has been designed to be fast
+As Saldo is continuously updated, the importing process has been designed to be fast
 and stable enough to be redone any time.
@@
 The basic algorithm for importing Saldo was implemented by Angelov (2009?)
-and  produces code for a GF lexcion.
+and  produces code for a GF lexicon.
 For each word in Saldo, it decides which forms should be used as input
@@
 in Saldo, if the tables are equal the code for the word is saved. If the table
-is erronous, another try is made
-by giving more forms to the smart pardigm.
+is erroneous, another try is made
+by giving more forms to the smart paradigm.
 For example \ref{fig:saldoknyt}, the smart paradigm will fail to calculate the
@@
 while the one generated in GF contains some that are not in Saldo (eg. ``vuxits").
-Saldo may also contain variats: ``växt" and ``vuxit" are both supine form.
+Saldo may also contain variants: ``växt" and ``vuxit" are both supine form.
 As far as possible, the program makes up for this by only comparing the overlapping forms
@@
 \section{The Swedish resource grammar}
->> start by easy examples, jag ser inte dig, du ser inte mig osv. 
->> plocka från svenska delen 1 2 4 
->> show simple trees
->> explain all in Sw. section.
->> show abstract and concrete tree, compare. (or in Gf sect)
->> no subjects or object, analyse is more from the lexicon,
->> compare to others, ask elisabet.\\
+%>> start by easy examples, jag ser inte dig, du ser inte mig osv. 
+%>> plocka från svenska delen 1 2 4 
+%>> show simple trees
+%>> explain all in Sw. section.
+%>> show abstract and concrete tree, compare. (or in Gf sect)
+%>> no subjects or object, analyse is more from the lexicon,
+%>> compare to others, ask elisabet.\\
 \label{sec:swegf}
@@
 
+All categories need tables covering the forms the phrase can be used in. They also
+need to carry information about inherent features that may matter to other phrases.
 Section \ref{gf:swe} contained an easy example with nouns. We used the categories \verb-N-
@@
 
-In gf, only personal pronouns are considered to belong to PN. The rests (example from
-saldo, which is pronoun in SAG as well) are either Quantifiers or Dets.
-
-%In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
-%ser dig".
-%Discuss questions, how are word order handled here? Maybe negation and anterior.
-%Let's look on how post-determination (sec ref) is implemented, we know that the
-%smart paradigms can handle the noun declination, now show 'katten' vs. 'den trötta katten'.
-%A N carries the inflection information, turned in to CN by setting number. Example, type
-%of N and CN. A CN may then be modified by APs, set field for this. When turned into NP, 
-%definitness is set, depends on the isMod -field. NP has case as a parameter.
-%\verb|DetNP : Det -> CN -> NP ; |
-%To turn in into a NP need to fix determination, by Det. Dets have .. fixed,
-%but allow variation in.. 
-%To form \emph{`fågeln'}, the determiner (DetQuant DefArt NumSg) is used.
-%Is a quantifier, hence does not require number, is given later.
-%What is the difference?
-%Difference, how the determinern can set the definiteness of the noun and
-%the adjective. Ref (Cooper,1986)
-%Once we have a NP, can use Predets. What they are.\\
-
-Valency information is normally built into the GF categories. To indicate that
-a verb is transitive and needs an object, we use the category \verb_V2_.
+In gf, only personal pronouns are considered to belong to PN. The rests, eg.
+sådan (\emph{``such"}) are either Quantifiers or Dets.
+
+In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
+ser katten". The verb ``såg" (see) takes one object: ``katten". Valency information
+is encoded into the GF categories. The lexicon is therefore a very important part of the
+grammar. 
 \begin{verbatim}
-Abstract:
-cat V2 ; 
-eat_V2 : V2 ;
-rat_N ;
-UseV2 : V2 -> NP -> V ;
-
-Concrete:
-eat_V2 = "äta" ;
-rat_N = {s = table {Sg => "råttan" ;
-                    Pl => "råttorna"}};
-UseV2 v np = v ++ np ;
+see_V2 ; 
+SlashV2a : V2 -> VPSlash ;
+ComplSlash : VPSlash -> NP -> VP ;
 \end{verbatim}
-
-The example above allows ``Katten äter råttan", but not ``Katten äter".
-%\verb|UseV2| concatenate the strings ``äter" and ``råttan".
-Note that if we wanted to add negation, ``katten äter inte råttan", or question
-"äter katten råttan", we cannot glue the object directly to the verb as done in
-\verb-UseV2-. Instead we would have to use a table, with one field for the verb
-and one for the object.
-
-For verbs, but also others, there is an important distinction for valency.
-Encoded in the lexicon. Categories as V,V2,VS...
-Either directly to VP if V, or needs to be given arguments. Can then add adverbs.
+The simplest idea would let the VP consist of one field where ``se" to ``katten" is
+glued together. But we also want to be able to express sentencens like
+``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
+The \verb-VP- category is hence implemented to ressembel the satsschema (Diedriches,xx), 
+it has multiple fields that may be put in the correct order later. There are fields
+for negation, different types of adverbs, objects, the finite verb and an infinite verb.
+The fields may in turn be tables showing the differnt forms they may have.
+
+\begin{verbatim}
+han (har   inte alltid tänkt på henne så)
+     finit  neg  adV    inf  comp obj adv
+\end{verbatim}
+
+The funciton \verb-SlashV2a- lets you create a \textit{VPSlash}, a verb phrase missing a
+noun phrase \cite{gazdar}. We combine the SlashVP with the object and get a complete
+verb phrase.
+\begin{verbatim}
+ presens negative  (du ser mig)
+  np.s ++ vp.fin ++ vp.neg ++ vp.obj ; 
+\end{verbatim}
+
+The slash-categories cat be combined with a subject to form a \verb-ClSlash-,
+a clause missing a noun phrase. This is a convenient way to form questions ,
+relative clauses
+and topicalized clauses.
+\begin{verbatim}
+wh-questions: (interrogativ pronoun + ClSlash)
+               vad                 (subject: Johan, vpslash: tittar på)
+relative clauses: object + relative pronoun + ClSlash
+                  katten     som             (subject: Johan, vpslash: tittar på)
+topicalized clauses: object + ClSlash
+                     katten  (subject: Johan, vpslash: tittar på)
+\end{verbatim}
+
+%Categories as V,V2,VS...
 %Maybe talk about ställa.
-Particles and prepositions, can't be seen on the trees. Can now see why the parse
-trees are not complete. difference shows when fronting.
-tycka om /tycka om -> om vad tycker du inte. fast bra
-Swedish verb may have up to five arguments according to Stymne
-'jag tar med mig den till Bo' : ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;
-but is really 'till' part of verb?
-VP is divided corresponding to satsschema (diderichsen) (renewed by SAG).
-If the object is not given immediatly, the grammar uses slash.
-"Speed up", instead of movement used slashes \cite{gazdar}.
-Slashes good for question (vilken katt ser du idag)
-relatives (katten som du tittar på) and topicalisation
-The grammar has VP,Cl, SSlash, all missing NP.
-
-Example, use variants {}.
+Swedish verbs may take up to five arguments \cite[p. xx]{stymne}.
+\enumsentence{Jag tar med mig den till honom}\label{sent:tamed}
+These may be particles, reflexive object, indirect objects and direct objects.
+In the grammmar these will be encoded in different ways. The particles are
+given in the lexicon, as well as the prepositions that are choosen by the verbs.
+The entry for sentence \ref{sent:tamed} looks like this, assuming that we already
+have an constant \verb-take_V- which produces the correct conjugation for ``ta":
+\verb|ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;|
+The category \verb-V3- tells us that the verb takes two arguments (``den" and ``honom").
+The preposition is set to ``till", reflV shows that the verb always is used with a reflexive
+pronoun and partV gives the particle.
+The fact that the choosen prepositions is given to the verb in the lexicon makes the 
+parse tree visualisation algorithm put them together. See parse tree \ref{fig:tittapa}.
+This is also the case for particles, see parse tree \ref{tittapart}, but even if
+the visualizations look the same the two cases are treated differently inside the grammar
+to keep the actual difference.
+\enumsentence{På pojken tittar du \\ *På när du springer tittar jag}
+As already stated, the visualized parse trees is not a complete represtation, and the reader
+should not be confused or angry at this.\\
+
+%
+%Verbs with particles or verbs that choose a special 
+%Particles and prepositions, can't be seen on the trees. Can now see why the parse
+%trees are not complete. difference shows when fronting.
+%tycka om /tycka om -> om vad tycker du inte. fast bra
+%Swedish verb may have up to five arguments according to Stymne
+%'jag tar med mig den till Bo' : 
+%but is really 'till' part of verb?
+%VP is divided corresponding to satsschema (diderichsen) (renewed by SAG).
+%If the object is not given immediatly, the grammar uses slash.
+%"Speed up", instead of movement used slashes \cite{gazdar}.
+%Slashes good for question (vilken katt ser du idag)
+%relatives (katten som du tittar på) and topicalisation
+%The grammar has VP,Cl, SSlash, all missing NP.
+%
+
+In some special cases, phrase cannot be put in all forms. 
+Give a good example!!
+It is not quite clear how to handle this at the moment. When
+the grammar is meant for parsing only, it is feasible to put
+a \textit{nonexist} pattern.
+example with this.\\
 
@@
 \label{gfSwe:parseable}
-The grammar can construct and parse the sentence in example \ref{gfSwe:parseable}
+The resource grammar can construct and parse the sentence in example
+\ref{gfSwe:parseable}
 and the parse tree is represented as in figure \ref{gfSwe:parsetree}.
@@
 \label{sec:Added}
-Summary of what have been added and motivation.
-
-\subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian have two ways of forming future tense. The standard
-form 'ska' is included in the resources as the translation to
-'will'. The second tense 'kommer att' is also very commonly used, and
-desereved to be covered by our grammar.
-Komma is used when we believe that we are sure that it will happen:
-'det kommer bli mörkt snart'
-and ska more of planning
-'jag ska äta mat nu'.
-since we believe it will happend, it can not be as in :
-'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
-
-Since two out of the three Scandinavian languages share this tense, it was 
-added to the Scandiniavin extra module.
-"jeg kommer til å savne deg"
-
-A table of result can be overview in figure \ref{fig:kommeratt}
-\begin{figure}[h]
-\begin{verbatim}
-s SFutKommer Simul Pos Main : jag kommer att se henne
-s SFutKommer Simul Pos Inv  : kommer jag att se henne
-s SFutKommer Simul Pos Sub  : jag kommer att se henne
-s SFutKommer Simul Neg Main : jag kommer inte att se henne
-s SFutKommer Simul Neg Inv  : kommer jag inte att se henne
-s SFutKommer Simul Neg Sub  : jag inte kommer att se henne
-s SFutKommer Anter Pos Main : jag kommer att ha sett henne
-s SFutKommer Anter Pos Inv  : kommer jag att ha sett henne
-s SFutKommer Anter Pos Sub  : jag kommer att ha sett henne
-s SFutKommer Anter Neg Main : jag kommer inte att ha sett henne
-s SFutKommer Anter Neg Inv  : kommer jag inte att ha sett henne
-s SFutKommer Anter Neg Sub  : jag inte kommer att ha sett henne
-\end{verbatim}
-\caption{The covered and accepted usage of future modal verb 'komma'}
-\label{gfSwe:parsetree}
-\end{figure}
+The resource grammar is meant to cover the most important grammatical structures
+present in all? languages. We have increased this by adding language specific
+constructions. Talbanken has been our inspiration.
+From the things listed in section \ref{sec:swedish}, the two first are basic
+enough to be covered by the resources. The others have been added.
 
 %%PASSIV
-\subsubsection{Passive}
->> why not VPSlash here??
+\subsection{The s-passive}
 Passive voice is often used in Swedish, especially the
@@
 The resource grammar for Scandinavian therefore ~~expressed~~ the passive using auxiliary verb. \\
->> change to:
 \verb|PassV2 : V2  -> VP | \\
 \verb|         ta -> blev tagen| \\
->> frånta, erbjuda works. add. check vp in urdu
->> how are they used with reflexives??
+%>> how are they used with reflexives??
 This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
@@
 than the default.
->> error!!! 
 \verb|PassVPSlash : VPSlash  -> VP | \\
@@
 form passives:
-\enumsentence{First place in two-place verbs\shortex{3}
+\enumsentence{First place in two-place verbs\\\shortex{3}
 {Jobbet & erbjöds & henne }
@@
 \label{ex:passV32}
-\enumsentence{Second place in two-place verbs\shortex{3}
+\enumsentence{Second place in two-place verbs\\\shortex{3}
 {Hon & erbjöds & jobbet}
@@
 
-%r
-%\enumsentence{\shortex{6}
-%{Betyget & 3 &ges &till &många &elever}
-%{Grade & 3 & give+\textbf{s}& to & many & students}
-%{``Grade 3 is given to many students"}}
-%\label{ex:passV3}
-%by letting the first object be `removed' and given by the subject. 
-%
-%A remark can be made about sentences where the second object is put first in the sentence, like 
-%\enumsentence{Till många elever ges betyget 3}
-%This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
-%parse this by applying \verb|FocObj|.
-%%
-
-%% Clauses
-\subsubsection{Impersonal constructions}%Presentering, , Topicalisering
+\subsection{Impersonal constructions}%Presentering, , Topicalisering
 \label{sec:Formal}
@@
 % 
-
-\subsubsection{Quantifiers, determiners and predeterminers}
+\subsection{Formalizing the rules for reflexive pronouns by using dependent types}
+%varandra and sin/sitt/sig.
+An important piece in a Swedish grammar is the reflexive pronouns and
+%both \emph{sin} and , eg. \emph{sin}.
+the possessive pronouns as described on p. \pageref{swe:refl}.
+The reflexive pronouns require an antecedent with which they agree in
+number and person.
+Our grammar should accept the following sentences:
+
+\enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
+He saw {\sc self's} children's shoes.}
+\enumsentence{Sina vantar hade han glömt på tåget.\\
+{\sc self's} gloves, he had forgotten on the train.}
+\enumsentence{Han ger sina pengar till sina syskon.\\
+He gives {\sc self's} money to {\sc self's} children.}
+\enumsentence{Hon ber alla sina kompisar att gå\\
+She asks all {\sc self's} friends to leave}\label{ex:ReflPredet}
+\enumsentence{Jag vill att han rakar sig.\\
+I want him to shave {\sc self}.}
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
+a. & Han är längre än sin kompis. & & b. &Han är här oftare än sin kompis.\\
+   & He is taller than {\sc self's} friend.
+   & & & He is here more often than {\sc self's} friend.
+\end{tabular}}\label{ex:ReflAP} \label{ex:ReflAdv}
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. & Hon tyckte om skolan och alla sina elever. & & b. & Han såg sina få böcker och sin penna.\\
+   & She liked the school and all {\sc self's} students. &&&
+ He saw {\sc self's} few books and {\sc self's} pencil.\\ 
+\end{tabular}}\label{ex:ReflConj} \label{ex:ReflConj2}
+When a reflexive pronoun is used in a noun phrase, it is hence not
+allowed as subject. The antecedent must not be outside of the smallest verb phrame??
+The grammar should not accept any of these sentences:
+\enumsentence{
+*Sina vantar var kvar på tåget.\\
+{\sc self's} gloves were left on the train.}\label{ex:ReflSubj}
+\enumsentence{*Hon och sin kompis läser en bok.\\
+She and {\sc self's} friend are reading a book. }
+\enumsentence{*Jag ger sina pengar till sina syskon.\\
+I give {\sc self's} money to {\sc self's}his siblings. }
+\enumsentence{*Han vill att jag ser på sig.\\
+He wants me to look at {\sc self}.}
+
+Apart from these restrictions, noun phrases containing reflexive pronouns
+may be used as any other \verb-NP-. They may beconjuncted (sentence
+\ref{ex:ReflConj} a and b) and determined (sentence \ref{ex:ReflPredet}).
+%and used as genitives (sentence \ref{ex:ReflGen}) as such.
+%Could be combined with 'sig', a ReflNP without another noun.
+In short, what is wanted can be summarized as follows:\\
+\begin{figure}[h]
+\begin{enumerate}
+\item{
+Construct a noun phrase from a common noun phrase \\
+katt $\rightarrow$ sin katt} % \\
+\item{Modify it like a \verb|NP|\\
+sina katter $\rightarrow$ alla sina katter}
+\item{Use it only as object\\
+sin katt $\rightarrow$ han såg sin katt }
+\end{enumerate}
+\caption{Rules for noun phrases with reflexive possessive pronouns}
+\label{fig:reflrestrict}
+\end{figure}
+If we ignore the second requirement, we might come up with a solution where
+special functions for using common nouns
+together with reflexive possessive pronouns are introduced:
+\begin{verbatim}
+ReflVP: CN -> VPSlash -> VP ;
+        mat + äta     -> äta sin mat 
+\end{verbatim}
+This way, using the phrases as subject is excluded, but neither of sentence
+\ref{ex:ReflConj}a, b or \ref{ex:ReflPredet} are allowed.\\
+The simplest way to fix this is to treat the phrases as normal \verb_NP_s.
+We can add a rule for using the possessives with common noun phrases:
+\begin{verbatim}
+ReflNP: CN -> NP ;
+        mat -> sin mat 
+PredetNP : Predet -> NP       -> NP ;
+           alla   + sina barn -> alla sina barn
+\end{verbatim}
+But by keeping them as \verb_NP_s we lose information, and there is no way of
+keeping them from being used as subjects. We have ignored the third restriction
+and are allowing sentences like \ref{ex:ReflSubj}.
+
+We get closer to a satisfying solution by introducing new object type,
+\verb_Obj_, which is
+identical to normal \verb_NP_ except for that it depends on 
+the antecedent. All functions where objects may be used in the same way
+as other \verb-NP-s are duplicated, one version is dealing with
+with objects and one with \verb-NP-.
+\begin{verbatim}
+ReflVP: Obj     -> VPSlash -> VP ;
+     -- sin mat +  äta     -> äta sin mat 
+NPtoObj : NP  -> Obj ;
+PredVP  : NP -> VP -> Cl ;  -- NP used as subject
+\end{verbatim}
+Any normal noun phrase may also be used as an object, whereas subjects only
+can be made up by \verb-NP-s.
+The drawback of this solution is the duplication of functions.
+\begin{verbatim}
+PredetNP  : Det -> NP  -> NP ;
+PredetObj : Det -> Obj -> Obj ;
+\end{verbatim}
+%We will need functions for conjugating objects with objects, 
+% say smt about overgeneration?
+Further, the dependence on the antecedent needs to be applied to adverbial phrases and
+adjective phrases as well. The adjective phrase in sentence \ref{ex:ReflAP}a and
+the adverbial phrase in sentence \ref{ex:ReflAdv}b are examples of this; if the
+subject was changed to 2nd person, they  would need to change too.
+\enumsentence{\textbf{Du} är längre än \textbf{din} kompis.\\
+You are taller than your friend.} \label{ex:ReflAP2}
+\vspace{-2mm}
+\enumsentence{*Du är här oftare än sin kompis.\\
+You are here more often than his friend.} \label{ex:ReflAdv2}
+The dependency is spreading through the grammar and in order to avoid all code
+duplication we should come up with a better idea.
+
+When looking at the type of
+the functions, we notice that they can be generalized:
+\begin{verbatim}
+PredetNP  : Det -> NP x  -> NP x;
+\end{verbatim}
+The solution given in this project is making use of this generalization, it
+introduces the use of dependent types in a resource grammar.
+Following the idea given above, we make a difference between subjects and objects, but not
+by giving them different types, but by treating them as instances of the \verb-NP- category.
+The type \verb-NP- \textit{dependes} on its argument, which may either be \verb-Subject- or
+\verb-Object-.
+\begin{verbatim}
+cat NP (Obj | Subj) ;
+PredVP : NP Obj -> VP -> Cl ;
+ComplSlash : VPSlash -> NP Obj -> VP ;
+PredetNP : (a : Boolean) -> Det -> NP a -> NP a ;
+\end{verbatim}
+
+The solution allows adverbial phrases may also be parts of new noun phrases.
+\enumsentence{\shortex{3}
+{Taket & på sitt hus &}
+{\scriptsize{\textbf{NP +}} & \scriptsize{\textbf{(Adv with Obj)}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{The roof of his house}}
+\enumsentence{\shortex{3}
+{Sitt hus & på berget &}
+{\scriptsize{\textbf{Obj +}} & \scriptsize{\textbf{Adv}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{His house on the mountain}}
+
+The use of dependent types separates our grammar from the other resource grammars. Such separation
+is not desirable in it self, but there are at least three reasons to why we believe it is feasible in this case:
+\begin{itemize}
+\item The goal of the project is to make a language specific grammar, and it would be surprising if this could be
+      done without leaving the pattern laid in by the multilingual abstract.
+\item The new grammar can be made compatible with the old implementation.
+\item The separation between object and subject could be useful in other languages. For example, the slavic languages
+      have similar use of reflexive pronouns.
+\end{itemize}
+
+The use of dependent types is interesting for other problems.
+The rules for reciprocal pronouns could be formalized using the same idea. As reflexive pronouns, reciprocals
+must not be used as subject and they may only be used when the number of the antecendent is plural.
+Sentences like
+'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
+We have a clause in which the subject must agree with the added pronoun in natural gender and
+in 
+%
+%Since the resource grammar did not make any distinction between
+%
+%subjects and objects, the new Swedish grammar needed to be separated from the old
+%structure. 
+%The first restriction in figure \ref{fig:reflrestrict} should be slightly
+%modified, to allow certain kind of \verb|NP|s to be used.
+%\begin{figure}[h]
+%\begin{enumerate}
+%\item{
+%Construct a noun phrase from a common noun phrase or certain noun phrases \\
+%få katter $\rightarrow$ sina få katter} 
+%\end{enumerate}
+%\caption{Modification of \ref{fig:reflrestrict}}
+%\end{figure}
+%
+%The kind of noun phrases that can be used seems to coincide with the ones
+%that can be used with formal subjects (see section `Impersonal constructions').
+%>> subject controlling verbs
+%>> why comlicated with doubled types, show simplicity (elegant) with dep.types, and
+%>> some problems (many fields).
+%>> samesame as Obj, but conjunction works too, sin is Quant but what i sig?
+%>> (show circle (sig : NP -> GenNP sig -> sin??)
+%>> how we can add dependent type for Num to use 'varandra'
+%>> CN should have dettyp too, 'avståndet till sin bil'
+%\verb- PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
+%  ajaj unknown here!!
+
+%>>Mention varandra. Same idea.
+%>>agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
+% F1,2,3 : NP -> VP -> Cl ; NP needs gender (Masc,Fem,Utr,Neutr)
+
+%Make a distinction between objects and subjects?
+%do not want to copy all rules for another category. Would like depent types!
+%work pretty well by letting all NP depend on the subject in the phrase.
+%The current version is nice and works for recursive genitivs!
+%Extra info, but maybe this is just the case about NPs.
+%Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
+%*Slash* : VPSlash -> NPR -> VP ;
+%Keep number of fields down, but introduces another category, and one more
+%label to the parse tree, is it interesting? How to combine it with genitive 
+%without making it ambiguous (Han såg barnens hund).
+%What if we add alot of other rules that uses NP, will they all need to be
+%duplicated then?
+
+%'han såg sin mammas bil', they same could be applied to
+%APs for example \ref{ex:ReflAP} or Adv for
+%\enumsentence{Han leker oftare än sin kompis}
+% (ComparAdvAdj)
+%Atm you can't focus this kind of object, then a ClSlash need info about the
+%subject. FunRP also tricky (and what does it mean, we can say
+%Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
+%Kort sagt, the information about the subject must be spread
+%in different parts of the sentence. 
+%Could also decide to always produc 'sin' -> jag såg sin bil
+%to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
+%in third person, and add nonexist for others. But this way we can also get rid
+%of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
+%'jag såg din bil'.
+%Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
+%not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
+%etc.
+%varandra, works the same but has no case for Sg.
+%%
+
+\subsection{A second future tense: ``Kommer att"}
+Swedish and Norwegian have two ways of forming future tense. The standard
+form 'ska' is included in the resources as the translation to
+'will'. The second tense 'kommer att' is also very commonly used, and
+desereved to be covered by our grammar.
+Komma is used when we believe that we are sure that it will happen:
+'det kommer bli mörkt snart'
+and ska more of planning
+'jag ska äta mat nu'.
+since we believe it will happend, it can not be as in :
+'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
+
+Since two out of the three Scandinavian languages share this tense, it was 
+added to the Scandiniavin extra module.
+"jeg kommer til å savne deg"
+
+A table of result can be overview in figure \ref{fig:kommeratt}
+\begin{figure}[h]
+\begin{verbatim}
+s SFutKommer Simul Pos Main : jag kommer att se henne
+s SFutKommer Simul Pos Inv  : kommer jag att se henne
+s SFutKommer Simul Pos Sub  : jag kommer att se henne
+s SFutKommer Simul Neg Main : jag kommer inte att se henne
+s SFutKommer Simul Neg Inv  : kommer jag inte att se henne
+s SFutKommer Simul Neg Sub  : jag inte kommer att se henne
+s SFutKommer Anter Pos Main : jag kommer att ha sett henne
+s SFutKommer Anter Pos Inv  : kommer jag att ha sett henne
+s SFutKommer Anter Pos Sub  : jag kommer att ha sett henne
+s SFutKommer Anter Neg Main : jag kommer inte att ha sett henne
+s SFutKommer Anter Neg Inv  : kommer jag inte att ha sett henne
+s SFutKommer Anter Neg Sub  : jag inte kommer att ha sett henne
+\end{verbatim}
+\caption{The covered and accepted usage of future modal verb 'komma'}
+\label{gfSwe:parsetree}
+\end{figure}
+
+%r
+%\enumsentence{\shortex{6}
+%{Betyget & 3 &ges &till &många &elever}
+%{Grade & 3 & give+\textbf{s}& to & many & students}
+%{``Grade 3 is given to many students"}}
+%\label{ex:passV3}
+%by letting the first object be `removed' and given by the subject. 
+%
+%A remark can be made about sentences where the second object is put first in the sentence, like 
+%\enumsentence{Till många elever ges betyget 3}
+%This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
+%parse this by applying \verb|FocObj|.
+%%
+\subsection{Relative clauses}
+The resource grammar already provided a number of ways to express relative and embedded clauses, such as
+\enumsentence{Frågan är vilken färg den har\\
+                   Pojken, som är blyg, tystnar\\
+                   Han såg kunden som tyckte om sallad\\
+                   Jag tänkte på huset i vilket hon bodde} % better example!!
+
+A few problematic things can be illuminated:
+Since the constructions are inspired by other languages, some of them may sound quite
+non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
+and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
+\emph{`She sleeps, which is good'} as
+\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
+embedded in the relative clause. This is easily fixed, by adding an extra
+parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
+Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
+The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
+\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
+This relative clause in the sentence is, in GF, put together by two constituents: \emph{`` katt"}
+of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause). They are
+glued together by \emph{``sådan att"} and \emph{``katt"} should agree with \emph{``den"}. 
+\enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
+However, this is not the case and sentence \ref{sent:detfaller} is currently
+accepted by the grammar.
+
+The grammar has been extended to accept
+\enumsentence{\shortex{7}
+{\textbf{De uppfattningar}& som& förs& fram ...} 
+{The opinions& that& put & forward ...}
+{``The opinions presented ..."}}
+where the definite artcile \emph{``de"} is used together with a indefinite
+noun. This is only correct for the definite articles and not determiners
+and requires that the noun phrase is followud by a restrictive
+relative clauses or together with the verb \emph{``finnas"}.
+In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
+\enumsentence{\textbf{De uppfattningarna} förs fram.\\
+``The opinions are presented."}\label{sent:debocker}
+% also ok in finnas!!
+
+%An more natural sounding alternative would be
+%\enumsentence{Jag vill ha en sådan katt som inte fäller.}
+%and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
+%'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
+%needed.
+%du är här, vilket jag vet.
+%i de äpplen/det äpple du äter bor en mask/ som du ser.\\
+
+\subsection{Modifying verb phrases}
+\subsubsection{Focusing adverbs}
+GF separates between two categories of adverbs, the normal ones \verb-Adv-, 
+eg. `nu' (\emph{`now'}) and
+special ones (name!) \verb-AdV- like `alltid' (\emph{`always'}).
+The \verb-AdV- attaches directly to the verb and have a special
+field in the \verb-VP--table.
+\enumsentence{\shortex{4}
+{Jag & äter & aldrig & fisk}
+{I & eat & never & fish}
+{`I never eat fish'}}
+\enumsentence{\shortex{4}
+{Jag & äter & fisk& nu }
+{I & eat & fish& now }
+{`I eat fish now'}}
+
+\begin{wrapfigure}{l}{0.3\textwidth}
+\begin{tabular}{|l|}
+\hline
+bara \\
+inte ens \\
+till och med \\
+\hline
+\end{tabular}
+\caption{Focusing adverbs}
+\label{fig:focadv}
+\end{wrapfigure}\\
+The focusing adverbs, like `bara' (\emph{`only' or `just'}) may put in both? of this
+positions and additionally, they may be used before the finite verb.
+
+\enumsentence{\shortex{3}
+{Han & bara & log }
+{he & only & smiles }
+{`He just smiles'}}
+
+In order to implement this, a new field was needed in the \verb-VP--table,
+which may be put before the finite verb.\\
+\verb|s SPres Simul Pos Main : han bara sover|\\
+compare to negation, \verb-AdV- and \verb-Adv-\\
+\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
+example in Sub or Inv, 
+When used together with modal verbs, the new field results in
+some not very natural sounding constructions:\\
+\verb|s SPres Anter Pos Main : han bara har sovit|\\
+but rather than changing to 'han har bara sovit', we keep it. Is a possible
+tolkning and the other case is covered by using it as an AdV, we don't want
+ambiguities.
+
+The foucusing adverbs may also be used as predeterminers:
+\enumsentence{\shortex{7}
+{Det & är & bara & barn & får & göra & så }
+{it & is & only & childern & may & do & so }
+{`Only children are allowed to do that'}}
+This is handled by the rule:
+\verb|PredetAdvF : AdvFoc -> Predet ;|
+
+\subsubsection{PPartAP}
+%The VP require many changes PPart, bara.
+PPartAP : V2 -> AP ; --VPSlash -> AP ;
+
+About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
+for the last one, we need VPSlash to allow adverb.
+
+\subsection{Miscellaneous}
+\subsubsection{Solving overgeneration}
+the extramodule contained GenNP for costructing genitvives. In Swedish
+done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
+severely overgenerating when started, and had to be restricted. Cycle:
+GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP
+      katt -> kattens      ->          kattens           det är kattens
+Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
+but still much better then the overgeneration.
+Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
+min with GenNP applied n number of times. solutions: bind s.
+or: A field should also been added to Quants, so that they can have a gen-form
+ : somligas.
+1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
+           fixing this include restructuring all???
+
+\subsubsection{Using Talbanken to assign correct categorise\\
+Quantifiers, determiners and predeterminers}
 \label{sec:gf.quant}
@@
 
-\subsubsection{Other small things}
-
-\textbf{Genitive}
-the extramodule contained GenNP for costructing genitvives. In Swedish
-done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
-severely overgenerating when started, and had to be restricted. Cycle:
-GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP
-      katt -> kattens      ->          kattens           det är kattens
-Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
-but still much better then the overgeneration.
-Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
-min with GenNP applied n number of times. solutions: bind s.
-or: A field should also been added to Quants, so that they can have a gen-form
- : somligas.
-1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
-           fixing this include restructuring all???
-
-\textbf{Det-utr}. no effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
- Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
- so added one more. Leads to ambiguities in some cases, but considering that we would want to use
- this for translation, neutr and utr may differ in the other language.
-
-\textbf{finiteness}
- 'de flesta de gula hästarna'
-
-\subsubsection{advs}
--- or is this really just AdVs? 
--- redan is Adv in abstract. Wrong!!
--- what about 'i regel'?
-to change position of adv:
-normal: after verb 'han äpplet redan'
-now want: 'äter redan äpplet'
-          'har redan ätit äpplet' 
-          'har ätit redan äpplet' :( which doesn't work either :)
-\subsubsection{Focusing adverbs}
-GF separates between two categories of adverbs, the normal ones \verb-Adv-, 
-eg. `nu' (\emph{`now'}) and
-special ones (name!) \verb-AdV- like `alltid' (\emph{`always'}).
-The \verb-AdV- attaches directly to the verb and have a special
-field in the \verb-VP--table.
-\enumsentence{\shortex{4}
-{Jag & äter & aldrig & fisk}
-{I & eat & never & fish}
-{`I never eat fish'}}
-\enumsentence{\shortex{4}
-{Jag & äter & fisk& nu }
-{I & eat & fish& now }
-{`I eat fish now'}}
-
-\begin{wrapfigure}{l}{0.3\textwidth}
-\begin{tabular}{|l|}
-\hline
-bara \\
-inte ens \\
-till och med \\
-\hline
-\end{tabular}
-\caption{Focusing adverbs}
-\label{fig:focadv}
-\end{wrapfigure}\\
-The focusing adverbs, like `bara' (\emph{`only' or `just'}) may put in both? of this
-positions and additionally, they may be used before the finite verb.
-
-\enumsentence{\shortex{3}
-{Han & bara & log }
-{he & only & smiles }
-{`He just smiles'}}
-
-In order to implement this, a new field was needed in the \verb-VP--table,
-which may be put before the finite verb.\\
-\verb|s SPres Simul Pos Main : han bara sover|\\
-compare to negation, \verb-AdV- and \verb-Adv-\\
-\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
-example in Sub or Inv, 
-When used together with modal verbs, the new field results in
-some not very natural sounding constructions:\\
-\verb|s SPres Anter Pos Main : han bara har sovit|\\
-but rather than changing to 'han har bara sovit', we keep it. Is a possible
-tolkning and the other case is covered by using it as an AdV, we don't want
-ambiguities.
-
-The foucusing adverbs may also be used as predeterminers:
-\enumsentence{\shortex{7}
-{Det & är & bara & barn & får & göra & så }
-{it & is & only & childern & may & do & so }
-{`Only children are allowed to do that'}}
-This is handled by the rule:
-\verb|PredetAdvF : AdvFoc -> Predet ;|
-
-\textbf{antalet}
-Also needed categories for 'pronouns' which can act in different ways.
-and 'antal','sorter' - need indetermined, antal plural, sorter either.
-Should be able to modify by a arbitrary antal adjective, and they must be determined.
-'(ett stort antal) människor' and not 'ett stort (antal människor)'
-Don't want them as a NPs or CNs since we don't want other Ns to be used this way
-'en flicka kaffe', this is then Appos? (Or do we?). Should also not be utsatta for what other
-NPs or CN can be. Dessutom, we want to make use of the comp-fiel. 'mamma till', but is this really
-the same??
-
-\textbf{VP}
-The VP require many changes PPart, bara.
-About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
-for the last one, we need VPSlash to allow adverb.
-
-\textbf{SuperlA}
-
-In general, many categories need more parameters and fields, tex Quant can be used as NP
-and therefor needs to depend on Case, de -> dem.
-
-here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
-\subsubsection{Reflexive pronouns}
-%varandra and sin/sitt/sig.
-An important piece in a Swedish grammar is the reflexive pronouns and
-%both \emph{sin} and , eg. \emph{sin}.
-the possessive pronouns as described in p. \pageref{swe:refl}.
-Those make up noun phrases which are not allowed in subjects, 
-and we want to be able to parse to following sentences:
-
-\enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
-He saw \sc{self's} children's shoes.}
-\enumsentence{Sina vantar hade han glömt på tåget.\\
-\sc{self's} gloves, he had forgotten on the train.}
-\enumsentence{Han ger sina pengar till sina syskon.\\
-He gives \sc{self's} money to \sc{self's} children.}
-\enumsentence{Han är längre än sin kompis.\\
-He is taller than \sc{self's} friend.} \label{ex:ReflAP}
-\vspace{-2mm}
-\enumsentence{Han är här oftare än sin kompis.\\
-He is here more often than \sc{self's} friend.} \label{ex:ReflAdv}
-\vspace{-2mm}
-\enumsentence{Hon tyckte om skolan och alla sina elever.\\
-She liked the school and all \sc{self's} students.}\label{ex:ReflConj}
-\enumsentence{Han såg sina få böcker och sin penna.\\
-He saw \sc{self's}his few books and \sc{self's} pencil.}\label{ex:ReflConj2}
-\vspace{-2mm}
-\enumsentence{Hon ber alla sina kompisar att gå\\
-She asks all \sc{self's} friends to leave}\label{ex:ReflPredet}
-\enumsentence{Jag vill att han rakar sig.\\
-I want him to shave \sc{self's}.}
-but not
-\enumsentence{*Sina vantar var kvar på tåget.\\
-\sc{self's} gloves were left on the train.}\label{ex:ReflSubj}
-\enumsentence{*Hon och sin kompis läser en bok.\\
-She and \sc{self's} friend are reading a book. }
-\enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give \sc{self's} money to \sc{self's}his siblings. }
-\enumsentence{*Han vill att jag rakar sig.\\
-He wants me to shave \sc{self's}.}
-
-The reflexive pronouns require an antecedent with which they agree in
-number and person. Apart from this restriction, they may be used as normal
-noun phrases and ~~conjuncted~~ (sentence \ref{ex:ReflConj} and \ref{ex:ReflConj2}),
-determined (sentence \ref{ex:ReflPredet}) and used as genitives (sentence
-\ref{ex:ReflGen}) as such.
-%Could be combined with 'sig', a ReflNP without another noun.
-In short, what is wanted can be summarized as follows:
-\begin{figure}
-\begin{enumerate}
-\item{
-Construct a noun phrase from a common noun phrase \\
-katt $\rightarrow$ sin katt} % \\
-\item{Modify it like a \verb|NP|\\
-sina katter $\rightarrow$ alla sina katter}
-\item{Use it only as object\\
-sin katt $\rightarrow$ han såg sin katt }
-\end{enumerate}
-\caption{Rules for noun phrases with reflexive possessive pronouns}
-\label{fig:reflrestrict}
-\end{figure}
-If we ignore the second requirement, special functions for using common nouns
-together with reflexive possessive pronouns could be introduced:
-\begin{verbatim}
-ReflVP: CN -> VPSlash -> VP ;
-        mat + äta     -> äta sin mat 
-\end{verbatim}
-This way, using the phrases as subject is excluded, but neither of sentence
-\ref{ex:ReflConj}, \ref{ex:ReflConj2} or \ref{ex:ReflPredet} is allowed.\\
-The simplest way of fixing this is to treat the phrases as normal \verb_NP_s.
-We can add a rule for using the possessives with common noun phrases:
-\begin{verbatim}
-ReflNP: CN -> NP ;
-        mat -> sin mat 
-PredetNP : Predet -> NP       -> NP ;
-           alla   + sina barn -> alla sina barn
-\end{verbatim}
-But by keeping them as \verb_NP_s we lose information, and there is no way of
-keeping them from being used as subjects. We have ignored the third restriction
-and are allowing sentences like \ref{ex:ReflSubj}.
-
-Another solution introduces a new object type, \verb_Obj_, which is
-identical to normal \verb_NP_ except for that it depends on 
-the antecedent. All functions using noun phrase as objects are changed to dealing
-with objects instead.
-\begin{verbatim}
-ReflVP: Obj     -> VPSlash -> VP ;
-     -- sin mat +  äta     -> äta sin mat 
-NPtoObj : NP  -> Obj ;
-PredVP  : NP -> VP -> Cl ;  -- NP used as subject
-\end{verbatim}
-Any normal noun phrase may also be used as an object, whereas subjects only
-can be made up by \verb-NP-s.
-Since the resource grammar did not make any distinction between
-subjects and objects, the new Swedish grammar needed to be separated from the old
-structure. 
-Further, the dependence on the subject needs to be applied to adverbial phrases and
-adjective phrases as well. The adjective phrase in sentence \ref{ex:ReflAP} and
-the adverbial phrase in sentence \ref{ex:ReflAdv} are examples of this; if the
-subject was changed to 2nd person, they  would need to change too.
-\enumsentence{\textbf{Du} är längre än \textbf{din} kompis.\\
-You are taller than your friend.} \label{ex:ReflAP2}
-\vspace{-2mm}
-\enumsentence{*Du är här oftare än sin kompis.\\
-You are here more often than his friend.} \label{ex:ReflAdv2}
-Adverbial phrases may also be parts of new noun phrases
-\enumsentence{\shortex{3}
-{Taket & på sitt hus &}
-{\scriptsize{\textbf{NP +}} & \scriptsize{\textbf{(Adv with Obj)}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
-{The roof of his house}}
-\enumsentence{\shortex{3}
-{Sitt hus & på berget &}
-{\scriptsize{\textbf{Obj +}} & \scriptsize{\textbf{Adv}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
-{His house on the mountain}}
-
-The first restriction in figure \ref{fig:reflrestrict} should be slightly
-modified, to allow certain kind of \verb|NP|s to be used.
-\begin{figure}[h]
-\begin{enumerate}
-\item{
-Construct a noun phrase from a common noun phrase or certain noun phrases \\
-få katter $\rightarrow$ sina få katter} 
-\end{enumerate}
-\caption{Modification of \ref{fig:reflrestrict}}
-\end{figure}
-
-The kind of noun phrases that can be used seems to coincide with the ones
-that can be used with formal subjects (see section `Impersonal constructions').
->> subject controlling verbs
->> why comlicated with doubled types, show simplicity (elegant) with dep.types, and
->> some problems (many fields).
->> samesame as Obj, but conjunction works too, sin is Quant but what i sig?
->> (show circle (sig : NP -> GenNP sig -> sin??)
->> how we can add dependent type for Num to use 'varandra'
->> CN should have dettyp too, 'avståndet till sin bil'
-\verb- PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
-  ajaj unknown here!!
-
->>Mention varandra. Same idea.
->>agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
-% F1,2,3 : NP -> VP -> Cl ; NP needs gender (Masc,Fem,Utr,Neutr)
-
-%Make a distinction between objects and subjects?
-%do not want to copy all rules for another category. Would like depent types!
-%work pretty well by letting all NP depend on the subject in the phrase.
-%The current version is nice and works for recursive genitivs!
-%Extra info, but maybe this is just the case about NPs.
-%Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
-%*Slash* : VPSlash -> NPR -> VP ;
-%Keep number of fields down, but introduces another category, and one more
-%label to the parse tree, is it interesting? How to combine it with genitive 
-%without making it ambiguous (Han såg barnens hund).
-%What if we add alot of other rules that uses NP, will they all need to be
-%duplicated then?
-
-%'han såg sin mammas bil', they same could be applied to
-%APs for example \ref{ex:ReflAP} or Adv for
-%\enumsentence{Han leker oftare än sin kompis}
-% (ComparAdvAdj)
-%Atm you can't focus this kind of object, then a ClSlash need info about the
-%subject. FunRP also tricky (and what does it mean, we can say
-%Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
-%Kort sagt, the information about the subject must be spread
-%in different parts of the sentence. 
-%Could also decide to always produc 'sin' -> jag såg sin bil
-%to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
-%in third person, and add nonexist for others. But this way we can also get rid
-%of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
-%'jag såg din bil'.
-%Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
-%not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
-%etc.
-%varandra, works the same but has no case for Sg.
-%%
-
-\subsubsection{Relative clauses}
-The resource grammar already provided a number of ways to express relative and embedded clauses, such as
-\enumsentence{Frågan är vilken färg den har\\
-                   Pojken, som är blyg, tystnar\\
-                   Han såg kunden som tyckte om sallad\\
-                   Jag tänkte på huset i vilket hon bodde} % better example!!
-
-A few problematic things can be illuminated:
-Since the constructions are inspired by other languages, some of them may sound quite
-non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
-and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
-\emph{`She sleeps, which is good'} as
-\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
-embedded in the relative clause. This is easily fixed, by adding an extra
-parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
-Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
-The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
-\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
-This relative clause in the sentence is, in GF, put together by two constituents: \emph{`` katt"}
-of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause). They are
-glued together by \emph{``sådan att"} and \emph{``katt"} should agree with \emph{``den"}. 
-\enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-However, this is not the case and sentence \ref{sent:detfaller} is currently
-accepted by the grammar.
-
-The grammar has been extended to accept
-\enumsentence{\shortex{7}
-{\textbf{De uppfattningar}& som& förs& fram ...} 
-{The opinions& that& put & forward ...}
-{``The opinions presented ..."}}
-where the definite artcile \emph{``de"} is used together with a indefinite
-noun. This is only correct for the definite articles and not determiners
-and requires that the noun phrase is followud by a restrictive
-relative clauses or together with the verb \emph{``finnas"}.
-In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
-\enumsentence{\textbf{De uppfattningarna} förs fram.\\
-``The opinions are presented."}\label{sent:debocker}
-% also ok in finnas!!
-
-%An more natural sounding alternative would be
-%\enumsentence{Jag vill ha en sådan katt som inte fäller.}
-%and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
-%'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
-%needed.
-%du är här, vilket jag vet.
-%i de äpplen/det äpple du äter bor en mask/ som du ser.\\
-
+%\subsubsection{Other small things}
+%
+%\textbf{Det-utr}. no effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
+% Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
+% so added one more. Leads to ambiguities in some cases, but considering that we would want to use
+% this for translation, neutr and utr may differ in the other language.
+%
+%\textbf{finiteness}
+% 'de flesta de gula hästarna'
+%
+%%\subsubsection{advs}
+%%-- or is this really just AdVs? 
+%%-- redan is Adv in abstract. Wrong!!
+%%-- what about 'i regel'?
+%%to change position of adv:
+%%normal: after verb 'han äpplet redan'
+%%now want: 'äter redan äpplet'
+%%          'har redan ätit äpplet' 
+%%          'har ätit redan äpplet' :( which doesn't work either :)
+%\textbf{antalet}
+%Also needed categories for 'pronouns' which can act in different ways.
+%and 'antal','sorter' - need indetermined, antal plural, sorter either.
+%Should be able to modify by a arbitrary antal adjective, and they must be determined.
+%'(ett stort antal) människor' and not 'ett stort (antal människor)'
+%Don't want them as a NPs or CNs since we don't want other Ns to be used this way
+%'en flicka kaffe', this is then Appos? (Or do we?). Should also not be utsatta for what other
+%NPs or CN can be. Dessutom, we want to make use of the comp-fiel. 'mamma till', but is this really
+%the same??
+%
+%\textbf{SuperlA}
+%
+%In general, many categories need more parameters and fields, tex Quant can be used as NP
+%and therefor needs to depend on Case, de -> dem.
+
+%here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
 \section{Testing}

### 2011-12-16 doc/report/report.tex
report on grammar and ramonas comments

@@
 %ska innehålla vettiga sökord
-\title{Scaling up a GF grammar for free parsing} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Towards a wide-coverage grammar for Swedish} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 \chapter{Introduction}
-% more intro to gf? Good starting sentence.
-The grammar formalism GF has so far been successful used for 
-controlled languages, but recent experiments have showed
-that it is also possible to use the framework for parsing free language.
+% more intro to gf. dependently typed...high-level formalism
+The grammar formalism GF \cite{gfbok} has so far been successfully used for 
+controlled languages \cite{cnl}, but recent experiments have showed
+that it is also possible to use the framework for parsing free language \cite{patent}.
+ >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
+ --> sentence about connection GF - grammar based parser.
 Parsing, that is the task of automatically identifying morphological and
@@
 To build a parser or a grammar from scratch would be not only time consuming but
-also a great theoretical and technical challenge. In order to not reinvent the wheel
+also a ~~great theoretical and technical challenge~~.
+>> not duplicate, already existing resources.
+In order to not reinvent the wheel
 we develop our parser by combining well-tested resources.
@@
 random generation, word alignment, morphological quizzes etc.
-%The following example shows how to parse a sentence and visualizing the parse tree.
-%, by using the \textit{pipe} \verb-|-, sending the result to 
+>> shell reference, servers, cloud. \footnote{http://www.grammaticalframework.org/demos/gfse/}
 \begin{figure}[h]
@@
 %>> use talk here and move key to other place??
+>> talk!
 The framework is also used in the
-European project MOLTO \cite{Molto} for online translation between up to 15 language.\\
+European project MOLTO \footnote{http://www.molto-project.eu/} for online
+translation between up to 15 language.\\
 
@@
 representing one language. As a comparison, the abstract and concrete syntax
-may be thought of as fstructures and cstructures in Lexical Functional Grammar (Bresnan and Kaplan, 1982).
+may be thought of as f-structures and c-structures in Lexical Functional Grammar (Bresnan and Kaplan, 1982).
 
@@
 The GF package provides a advantageous resource library \cite{gf-resource}, covering the
-fundamental morphology and syntactics of more than 20 languages.
+fundamental morphology and syntax of more than 20 languages.
 %is a very useful part of the framework. It is provided in the 
@@
 increased. There are research on how to make more use of the dependent
-types, for reasoning by using ontologies \cite{ramona} or generating natural
+types, for reasoning by using ontologies \cite{ramona}and krasse?? or generating natural
 language via Montague semantics \cite{montague}.
@@
 for ie. English and Korean, but, to our knowledge, not for Swedish.
+
+ParGram! More parallel development than Lingo.
+
 %However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
@@
 \section{Results}
+\label{sec:saldoRes}
 All words from Saldo were not be imported. There are a number of reasons
@@
 explains the development during this project.
+>> the grammar will not be totally covering, but interesting and useful.
 
@@
 \label{sec:swegf}
+>> diagram of modules, etc.
+>> gf has pos-driven, leads to part-of-sentence analyse. first have v2,a2,vs etc.
+>> they may be combined in traditional way, and so the types depend and relate to each other
+>> see slides from summer school.
+>> functions for predication, complementation.
+>> rather different from others syntactical theories. does not remember syntactical
+>> parts of sentences, like subj, obj, attributes.
+>> worked well until now, but particular langs. needs particular solutions. --> reflexives!
+>> no syntactic theory can cover a whole language, and so gf can't either.
+>> shares 80\% with scandinavian, mostly words that differ.
 The GF resource grammars contains a fundamental description of Swedish,
@@
 \end{verbatim}        
+>> compare NP to HPSG.
+>> CN may be different in other theories.
+>> how we have basic Pred, compl
+>> clase is a pre-sentence needs tense, polarity etc.
 Which combination that shoud be used is decided by the determiner. Hence all determiners
@@
 %Maybe talk about ställa.
-Swedish verbs may take up to five arguments \cite[p. xx]{stymne}.
+Swedish verbs may take up to four arguments \cite[p. 53]{stymne}.
 \enumsentence{Jag tar med mig den till honom}\label{sent:tamed}
@@
 to keep the actual difference.
-\enumsentence{På pojken tittar du \\ *På när du springer tittar jag}
+\enumsentence{På pojken tittar du \\ **På när du springer tittar jag}
 As already stated, the visualized parse trees is not a complete represtation, and the reader
@@
 In addition to the basic resource grammar, additional constructions had been added
+>> core is shared but particular lang. has particular features.
 to the module \verb|Extra|. The functions given here do not have to be translatable
 to all other language, but are meant to cover language specific constructions.
-%For example, the \verb|Extra| module included
 Among those were functions for topicalisation, although not all
@@
 enough to be covered by the resources. The others have been added.
+>> hard to estimate before, since no big use case/resource.
+>> after initial experiments, first evaluation. found much room for improvement. 
+>> complete morphology, limited syntax
+>> discovered constructions missing.
 
@@
 %The periphrastic passive is preferred, especially in spoken langauge.
-\enumsentence{\begin{tabular}{ll}
-Oppgaven ble skrevet av en student & [NO]\\
-Uppsatsen blev skriven av en student& [SE]\\
-\end{tabular}}
+\enumsentence{\shortexnt{7} %\begin{tabular}{ll}
+{Oppgaven &ble& skrevet& av& en& student & [NO]}
+{uppsatsen& blev& skriven& av& en & student& [SE]}}
+%\end{tabular}}
 The corresponding Swedish sentence is acceptable, but not as natural sounding as sentence
@@
 \label{sec:Formal}
-Formal subjects \cite[19]{SAG} is often used in Swedish.
+Formal subjects \cite[\textsection 19]{SAG} is often used in Swedish.
 \enumsentence{\shortex{6}
@@
 \subsection{Formalizing the rules for reflexive pronouns by using dependent types}
-%varandra and sin/sitt/sig.
+>> solution requires new strategies, combining 2 theories! (part of speech, part of sentence)
 An important piece in a Swedish grammar is the reflexive pronouns and
-%both \emph{sin} and , eg. \emph{sin}.
 the possessive pronouns as described on p. \pageref{swe:refl}.
@@
 {The roof of his house}}
+%{{\sc NP +} & {\sc (Adv with Obj)} & $\Rightarrow$ {\sc Obj}}
 \enumsentence{\shortex{3}
@@
 'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
-We have a clause in which the subject must agree with the added pronoun in natural gender and
-in 
+'min syster är inte dum, hon'
+We have a clause of which we need to know the natural gender and number of the subject, since it
+must agree with the added pronoun. sag s 439. fullständigt outforskade!
+
+%or just implement this and remove it here?
+The question is how many types we should give dependent types. CN could also need it,
+in order to say
+'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
+The analysis of this sentence is that
+avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
+and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
+Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
+
 %
@@
 \subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian have two ways of forming future tense. The standard
-form 'ska' is included in the resources as the translation to
-'will'. The second tense 'kommer att' is also very commonly used, and
-desereved to be covered by our grammar.
-Komma is used when we believe that we are sure that it will happen:
-'det kommer bli mörkt snart'
-and ska more of planning
-'jag ska äta mat nu'.
-since we believe it will happend, it can not be as in :
-'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
-
-Since two out of the three Scandinavian languages share this tense, it was 
-added to the Scandiniavin extra module.
-"jeg kommer til å savne deg"
-
-A table of result can be overview in figure \ref{fig:kommeratt}
+Swedish and Norwegian have two modal verbs for future tense
+\cite[p. 246]{H&H}.
+%\cite[Tempus \textsection 28]{SAG}.
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. & Det \textbf{kommer att} bli mörkt snart & & b. &Jag \textbf{ska} gå och lägga mig nu\\
+  & ``It will get dark soon" & & b. &``I'm going to bed now"\\
+\end{tabular}}
+\enumsentence{
+Jeg kommer til å savne deg ~~$[$NO$]$\\
+``I will miss you"}
+\emph{``ska"} signals an intention of committing the action, either from the subject or from
+the speaker. Cf.
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+a. & Du ska tycka om den & & b. &Du kommer tycka om den\\
+  & ``You \emph{shall} like it" & & b. &``You will like it."\\
+\end{tabular}}
+\emph{``kommer"} (normally used with the infinite marker \emph{att}) 
+does not signal any intention, but that the speaker has belief that 
+it will actually come true.\\
+The resource grammar included 
+\emph{``ska"} which was implemented as the standard way of forming future
+tense, and hence represents the translation of
+\emph{``will"}. 
+The new grammar also supports \emph{``kommer att"}, expressed as an alternative
+future tense.
+\begin{figure}[h]
+\begin{verbatim}
+(UseCl (TTAnt TFutKommer ASimul) PPos 
+       (ImpersCl (AdvVP (ComplVA become_VA (PositA dark_A)) soon_Adv))
+\end{verbatim}
+\caption{Result of parsing ``Det kommer att bli mörkt snart". The future tense
+         is marked by the constant TFutKommer}
+  \label{fig:kommeratt}
+\end{figure}
+
+%since we believe it will happend, it can not be as in :
+%'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
+Since two out of the three Scandinavian languages share this
+tense, it has been added to the Scandiniavin \verb-Extra- module.
+Table \ref{fig:kommeratt} shows how \emph{``kommer att"} may be used in different
+types of sentences.
 \begin{figure}[h]
@@
 \end{verbatim}
-\caption{The covered and accepted usage of future modal verb 'komma'}
-\label{gfSwe:parsetree}
+\caption{The covered and accepted usage of the future modal verb `komma'}
+  \label{fig:kommeratt}
 \end{figure}
@@
 %%
-\subsection{Relative clauses}
+\subsection{Modifying verb phrases}
+\subsubsection{Focusing adverbs}
+The GF analysis distinguishs between two categories of adverbs: \verb-Adv- 
+%eg. `nu' (\emph{`now'}) and
+and \verb-AdV-. %like `alltid' (\emph{`always'}).
+The \verb-AdV- (eg. `alltid' (\emph{`always'}) and `inte' (\emph{`not'}))
+attaches directly to the verb.
+%and have a special field in the \verb-VP--table.
+\enumsentence{Cf.\\\begin{tabular}[t]{@{}*{12}{l@{\ }}}
+a. & Jag & äter & \textbf{aldrig} & fisk & & b. &Jag & äter & fisk& \textbf{nu}\\ 
+&\emph{I} & \emph{eat} & \emph{never} & \emph{fish} &&&\emph{ I }&\emph{ eat} &
+\emph{fish}& \emph{now}
+\end{tabular}\\
+``\hspace{1mm} I never eat fish" \hspace{9mm} ``I eat fish now"}
+The difference is implemented by having separate fields in the \verb-VP- table
+for the two categories.
+\begin{verbatim}
+Main clause : subj ++ verb.fin ++ verb.adV ++ verb.inf ++ verb.adv
+              han     har         alltid      sovit       här 
+\end{verbatim}
+%\begin{wrapfigure}{l}{0.3\textwidth}
+%\begin{tabular}{|l|}
+%\hline
+%bara \\
+%inte ens \\
+%till och med \\
+%\hline
+%\end{tabular}
+%\caption{Focusing adverbs}
+%\label{fig:focadv}
+%\end{wrapfigure}\\
+The adverb `bara' may be put int both of these positions but also before the finite verb,
+when emphazing the verb itself.
+This is an example of a focusing adverb, another example is `inte ens' (\emph{`not even'}).
+\enumsentence{\shortex{9}
+{a. & Han & bara & log  && &b. & Hon till och med visslar}
+{&\emph{he} & \emph{only} &\emph{ smiled} &  & & &  }
+{``He just smiled"\hspace{8mm} ``She even whistles"}}
+Focusing adverbs are accepted by the new grammar implementation, where they have
+their own field in th \verb-VP-table.
+\begin{verbatim}
+Main clause : subj ++ verb.focAdv ++ verb.fin ++ verb.adV ++ verb.inf ++ verb.adv
+              han     bara           sover        -             -           -
+\end{verbatim}
+They may also be used as \verb-AdV-...
+\enumsentence{\shortexnt{4}
+{Han & har & bara  & sovit}
+{\emph{he} & \emph{has} & \emph{only}  &\emph{ slept}}}
+
+...or as predeterminers:
+\enumsentence{\shortex{8}
+{Det & är & bara & barn &som& får & göra & så }
+{it & is & only & childern &that& may & do & so }
+{`Only children are allowed to do that'}}
+Two more rules are consequentally needed
+\begin{verbatim}
+FocAdvAdV : FocAdv -> AdV ;
+PredetAdvF : AdvFoc -> Predet ;
+\end{verbatim}
+
+The focusing adverbs are usually not combined with modal verbs, since the modal
+verbs rarely are focused.
+\enumsentence{\shortexnt{5}
+{?&Han & bara&  har&  sovit}
+{&\emph{he} &\emph{ only} & \emph{ has}& \emph{ slept}}}
+We have however chosen to allow this, since it is grammatically correct although semantically
+dubious, and since \emph{``Han har bara sovit"} is covered by the rules using focusing
+adverbs as \verb-AdV-.
+Discussion: should han åt bara ett äpple be (bara ett äpple) focusing an NP?
+Then what about: han har bara sovit, han har bara ätit ett äpple.\\
+
+%Ask elisbet for other examples with inte ens och till och med, nätt och jämt.
+%In order to implement this, a new field was needed in the \verb-VP--table,
+%which may be put before the finite verb.\\
+%\verb|s SPres Simul Pos Main : han bara sover|\\
+%compare to negation, \verb-AdV- and \verb-Adv-\\
+%\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
+%example in Sub or Inv, 
+%When used together with modal verbs, the new field results in
+%some not very natural sounding constructions:\\
+%\verb|s SPres Anter Pos Main : han bara har sovit|\\
+%but rather than changing to 'han har bara sovit', we keep it. Is a possible
+%tolkning and the other case is covered by using it as an AdV, we don't want
+%ambiguities.
+\subsubsection{PPartAP}
+%The VP require many changes PPart, bara.
+
+PPartAP : V2 -> AP ; --VPSlash -> AP ;
+
+About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen
+funna'
+for the last one, we need VPSlash to allow adverb.
+
+\subsection{Miscellaneous}
+\subsubsection{Relative clauses}
 The resource grammar already provided a number of ways to express relative and embedded clauses, such as
-\enumsentence{Frågan är vilken färg den har\\
-                   Pojken, som är blyg, tystnar\\
-                   Han såg kunden som tyckte om sallad\\
-                   Jag tänkte på huset i vilket hon bodde} % better example!!
+\eenumsentence{
+\item[a.]Frågan är vilken färg den har
+\item[b.]Pojken, som är blyg, tystnar
+\item[c.]Han såg kunden som tyckte om sallad
+\item[d.]Jag tänkte på huset i vilket hon bodde} % better example!!
 
@@
 {\textbf{De uppfattningar}& som& förs& fram ...} 
-{The opinions& that& put & forward ...}
+{The opinions& that& put+{\sc s} & forward ...}
 {``The opinions presented ..."}}
@@
 
-\subsection{Modifying verb phrases}
-\subsubsection{Focusing adverbs}
-GF separates between two categories of adverbs, the normal ones \verb-Adv-, 
-eg. `nu' (\emph{`now'}) and
-special ones (name!) \verb-AdV- like `alltid' (\emph{`always'}).
-The \verb-AdV- attaches directly to the verb and have a special
-field in the \verb-VP--table.
-\enumsentence{\shortex{4}
-{Jag & äter & aldrig & fisk}
-{I & eat & never & fish}
-{`I never eat fish'}}
-\enumsentence{\shortex{4}
-{Jag & äter & fisk& nu }
-{I & eat & fish& now }
-{`I eat fish now'}}
-
-\begin{wrapfigure}{l}{0.3\textwidth}
-\begin{tabular}{|l|}
-\hline
-bara \\
-inte ens \\
-till och med \\
-\hline
-\end{tabular}
-\caption{Focusing adverbs}
-\label{fig:focadv}
-\end{wrapfigure}\\
-The focusing adverbs, like `bara' (\emph{`only' or `just'}) may put in both? of this
-positions and additionally, they may be used before the finite verb.
-
-\enumsentence{\shortex{3}
-{Han & bara & log }
-{he & only & smiles }
-{`He just smiles'}}
-
-In order to implement this, a new field was needed in the \verb-VP--table,
-which may be put before the finite verb.\\
-\verb|s SPres Simul Pos Main : han bara sover|\\
-compare to negation, \verb-AdV- and \verb-Adv-\\
-\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
-example in Sub or Inv, 
-When used together with modal verbs, the new field results in
-some not very natural sounding constructions:\\
-\verb|s SPres Anter Pos Main : han bara har sovit|\\
-but rather than changing to 'han har bara sovit', we keep it. Is a possible
-tolkning and the other case is covered by using it as an AdV, we don't want
-ambiguities.
-
-The foucusing adverbs may also be used as predeterminers:
-\enumsentence{\shortex{7}
-{Det & är & bara & barn & får & göra & så }
-{it & is & only & childern & may & do & so }
-{`Only children are allowed to do that'}}
-This is handled by the rule:
-\verb|PredetAdvF : AdvFoc -> Predet ;|
-
-\subsubsection{PPartAP}
-%The VP require many changes PPart, bara.
-PPartAP : V2 -> AP ; --VPSlash -> AP ;
-
-About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen funna'
-for the last one, we need VPSlash to allow adverb.
-
-\subsection{Miscellaneous}
-\subsubsection{Solving overgeneration}
-the extramodule contained GenNP for costructing genitvives. In Swedish
-done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
-severely overgenerating when started, and had to be restricted. Cycle:
+\subsubsection{An example of overgeneration and how it can be solved}
+The \verb-Extra- module contained a function for constructing genitives.
+\begin{verbatim}GenNP : NP -> Quant ;\end{verbatim}
+It selected the possessive form given in the \verb-NP- and changed its type
+to \verb-Quantifier-. 
+%\begin{verbatim}
+%(DetCN (DetQuant (GenNP (DetCN (DetQuant DefArt NumSg) (UseN boy_N))) NumSg) (UseN cat_N))
+%\end{verbatim}
+The \verb-GenNP- function was severly overgenerating when this project started, even
+though its implementation was not erroneous in it self. The overgeneration stemmed
+from two places.Quantifiers does not have any possessive forms, so when a quantifier is turned into
+a NP (via a Det), its only field was copied and used for both nominative and genitive.
+In combination, \verb-GenNP-
+introduced a cycle, a Quant may
+be turned into a determiner which can be turned into a NP, which can be turned
+into a Quant.
+\begin{verbatim}
 GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP
-      katt -> kattens      ->          kattens           det är kattens
-Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
-but still much better then the overgeneration.
-Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
-min with GenNP applied n number of times. solutions: bind s.
-or: A field should also been added to Quants, so that they can have a gen-form
- : somligas.
-1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
-           fixing this include restructuring all???
-
-\subsubsection{Using Talbanken to assign correct categorise\\
-Quantifiers, determiners and predeterminers}
+\end{verbatim}
+We know had that a noun phrase, say `katten' could be turned into a quantifier,
+by using its genitive form, `kattens'. This could be used as a determiner, and
+consequentelly as a noun phrase `Det är kattens'. But if we try to use `kattens'
+in genitive, we get `kattens' again. The function \verb-GenNP- could be applied
+any number of times without any change in the output.
+
+The problem originated in the function \verb-DetNP- that put the possessive form
+to be the same as the nominative. If we instead let it append an 
+\emph{s} to the word, \verb-GenNP- will stop overgenerating. Besides, the grammar 
+will allow genitive forms of quantifiers
+\enumsentence{Somligas skor står kvar i hallen\\
+``Some peoples shoes are still in the hall"??}
+
+This reduced the number of parse trees for ``kattens bils hus" from 100 to six.
+The six remaining trees describes different interpretations of the sentence,
+such as `hus' is in genitive or nominative and in either plural or singular.
+kattes bils hus är stort.
+kattens bils hus är stora.
+
+%the extramodule contained GenNP for costructing genitvives. In Swedish
+%done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
+%severely overgenerating when started, and had to be restricted. Cycle:
+%      katt -> kattens      ->          kattens           det är kattens
+%Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
+%but still much better then the overgeneration.
+%Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
+%min with GenNP applied n number of times. solutions: bind s.
+%or: A field should also been added to Quants, so that they can have a gen-form
+% : somligas.
+%1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
+%           fixing this include restructuring all???
+
+\subsubsection{Using Talbanken to assign correct categorisations}
 \label{sec:gf.quant}
-The GF analyse differs sometimes. Experiments of how to extract
-this information from Talbanken by using Saldo as a starting point.
 One of the biggest differences between the GF analysis and the one made in Talbanken
-is related to pronouns. While GF only recognizes the personal pronouns, 
-`sådan', `fler`, `somliga' etc. are all regarded as pronouns in Talbanken and Saldo.
-In the GF analyse they may be quantifiers, determiners or predeterminers. 
-All pronouns from Saldo was listed and by searching Talbanken for which tags
-they got we could sort out the ones acting either as
-determiners, as `sådana' in sentence \ref{ex:QuantDet}, or as adjectives as
-`sådan' in sentence \ref{ex:QuantAdj}. 
-\enumsentence{Sådana katter vill jag ha}\label{ex:QuantDet}
-\vspace{-3mm}
-\enumsentence{Han är sådan}\label{ex:QuantAdj}
-\vspace{-3mm}
-\enumsentence{Sådant är inte bra}\label{ex:QuantDet2}
-Since all determiners can be used alone as noun phrases, sentence
-\ref{ex:QuantDet2} is also accepted.
+is related to pronouns. 
+As noted in section \ref{sec:saldoRes} our GF grammar only classifies personal pronouns
+as pronouns. Many pronouns could not be imported from Saldo, since the correct category
+could not be identified automatically. Therefore, some experiments of how to extract this 
+information from Talbanken has been conducted. We were interested in words which can be used
+the same ways as `sådana' (\emph{`such'}), which acts either as
+determiners, as in sentence \ref{ex:QuantDet}, or as adjectives as
+ in sentence \ref{ex:QuantAdj}. 
+\eenumsentence{\item[a.]Sådana \label{ex:QuantDet}katter vill jag ha
+\item[b.]Han är \label{ex:QuantAdj}sådan\toplabel{sent:Quant}}
+
+By starting from a list containing of all pronouns  in Saldo and all forms they may occure in,
+Talbanken was harvested for the ones that both occured with the determiner tag \verb-DT-
+and with the present participle tag \verb-SP-. This way we are left with a limited number
+of words that could easily be analysed and added to the lexicon.
+
+Further, list of all pronous acting as determiners were extracted. This method is
+both faster and more reliable than just going through the list by hand and using
+intraspection or manually looking up each word.
+
+%Determiners can be used alone as noun phrases, as in sentence
+%\ref{ex:QuantDet2}.
+%\enumsentence{Sådant är inte bra}\label{ex:QuantDet2}
+%The GF analyse differs sometimes. Experiments of how to extract
+%this information from Talbanken by using Saldo as a starting point.
+%While GF only recognizes the personal pronouns, 
+%`sådan', `fler`, `somliga' etc. are all regarded as pronouns in Talbanken and Saldo.
+%In the GF analyse they may be quantifiers, determiners or predeterminers. 
+%All pronouns from Saldo was listed and by searching Talbanken for which tags
+%they got we could sort out the ones acting either as
+%determiners, as `sådana' in sentence \ref{ex:QuantDet}, or as adjectives as
+%`sådan' in sentence \ref{ex:QuantAdj}. 
+%Since all determiners can be used alone as noun phrases, sentence
+%\ref{ex:QuantDet2} is also accepted.
 %what is the difference really???
@@
 \section{Testing}
+>>evolution diagram? also removed unnecessary ambigutiies.
+>> how to handle ambigs: know them and controll them. using statistics or ontolgies.
+>> that is: disambiguitation by context or by meaning.
 For testing has been carried out by using a treebank, consisting of sentences used in 
@@
 >> reusablilty, freely available.
+>> have experiments of how to extract info from talbanken.
 %There is obviously a lot do still, will never get finish with NPL.

### 2011-12-16 doc/report/report.tex
all parts of report started

@@
 \subsubsection{Relative clauses}
-The resource grammar already provided a number of ways to express relative and embedded clauses, such as
+The resource grammar already gave a good coverage of relative clauses and embedded sentences.
+All sentences in example \ref{sent:relex} was accepted.
+%provided a number of ways to express relative and embedded clauses, such as
 \eenumsentence{
@@
 \item[c.]Han såg kunden som tyckte om sallad
-\item[d.]Jag tänkte på huset i vilket hon bodde} % better example!!
+\item[d.]Jag tänkte på huset i vilket hon bodde}\label{sent:relex} % better example!!
+
+The grammar has been extended to accept a few more constructions:
+\enumsentence{\shortex{7}
+{\textbf{De uppfattningar}& som& förs& fram ...} 
+{The opinions& that& put+{\sc s} & forward ...}
+{``The opinions presented ..."}}
+The definite artcile \emph{``de"} is used together with a indefinite
+noun. This is only correct for the definite articles and not determiners
+and requires that the noun phrase is followud by a restrictive
+relative clauses or together with the verb \emph{``finnas"}.
+In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
+\enumsentence{\textbf{De uppfattningarna} förs fram.\\
+``The opinions are presented."}\label{sent:debocker}
 
@@
 
-The grammar has been extended to accept
-\enumsentence{\shortex{7}
-{\textbf{De uppfattningar}& som& förs& fram ...} 
-{The opinions& that& put+{\sc s} & forward ...}
-{``The opinions presented ..."}}
-where the definite artcile \emph{``de"} is used together with a indefinite
-noun. This is only correct for the definite articles and not determiners
-and requires that the noun phrase is followud by a restrictive
-relative clauses or together with the verb \emph{``finnas"}.
-In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
-\enumsentence{\textbf{De uppfattningarna} förs fram.\\
-``The opinions are presented."}\label{sent:debocker}
 % also ok in finnas!!
@@
 \section{Testing}
+Througout the project regression testing has been used. Every time the grammar was updated
+it was tested against a treebank consisting of 155 sentences, and the number of parse trees
+compared to a standard.
+The purpose was first of all to
+make sure that the coverage was not decrease, but also to be sure we would notice
+if new ambiguites or overgenerating functions were introduced.
+When possible, the ambiguities could be removed and when not we could at least know about
+them. The grammar will always contain ambiguities, but by being aware helps to
+handle them in other ways, such by statistical means.
+%notice if the grammar starts to overgenerate. Avoiding ambiguities is 
+New functions were also tested by random generating and by creating tables showing
+all forms, such as table \ref{fig:kommeratt}.
+
 >>evolution diagram? also removed unnecessary ambigutiies.
@@
 >> that is: disambiguitation by context or by meaning.
-For testing has been carried out by using a treebank, consisting of sentences used in 
-sub-testsuites. This way we make sure that we doesn't decrease the coverage and will
-notice if the grammar starts to overgenerate. Also have some sentences that should not
-be parsed. 
-When adding new constructions, the table-function is very helpful, shows how a word will
-turn out in differnet cases. all tenses, for komma.
+%For testing has been carried out by using a treebank, consisting of sentences used in 
+%sub-testsuites. Also have some sentences that should not
+%be parsed. 
+%When adding new constructions, the table-function is very helpful, shows how a word will
+%turn out in differnet cases. all tenses, for komma.
 
@@
 
+These results are lower than the results of the mapping of Penn Treebank. The
+reason for this may lay in the fact that English is syntactically less complicated
+than Swedish. Also, the Penn Treebank contains a lower number of tags, 82 compared
+to more than 300 in Talbanken. Even if you exclude the tags describing the same word
+class as another tag, Talbanken still leaves us with more than 130 tags. The 
+number of ways to combine them are hence much increased, and so is the number of
+rules that would be needed to cover all constructions.
+
 We believe that the results could be enhanced by simply adding more rules 
@@
 \section{Evaluation}
+The project has resulted in
+\begin{itemize}
+\item one large-scale GF lexicon and a program to redo the importation when needed
+\item a comparison between GF and other annotations
+\item a deeper testing of the Swedish resource grammar and an estimation
+of how well GF can be used to describe larger parts of a language
+\item a study of how dependent types can be used in the resource grammars
+\end{itemize}
+
+The increase of the grammatical coverage is hard to compare, since
 The grammar in it self is not at all robust, and one unknown word will cause a whole sentence to 
-fail. When parsing text, we get low results, for this reason. Not yet interesting to talk about
+fail. 
+An estimation can be given by looking at the results from running the test suite.
+The first grammar could parse almost halv of the sentences, and the final
+grammar could parse 65\%. 
+When parsing text, we get low results, for this reason. Not yet interesting to talk about
 coverage, but focus on quality and ability to scale up. Our goal standarn is so far Talbanken,
@@
 Instead we have the luck to have an expert and her presence has increased the standard.
-
+No syntactic theory will ever cover all language, and this should just be seen as an important
+fragment.
+
+From the renewed import of Saldo, we got a much better coverage especially considering
+that the old version was missing many of the most common words. Bo, göra, se.
+However, the lexicon is now to big to use as a normal GF lexicon, since it will request
+too much memory to parse most sentences. We will need to handle this in a clever way,
+see future work.
 %Results from mapping, can probably be improved, bit by bit, depends much on words.
-%Valencys would therefor help.  
+%Valencys would therefore help.  
 %Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
@@
 \label{sec:future}
->> divide into consequences and future plans.
-What is still left to do? Description and direction of future work:
+A part of the project is finished, but the project is not complete, and will never be.
+Many interesting future directions. We will first look at the ones which are 
+direct consequences of what we did so far. 
+\subsection{Enhancments and completitions}
+\subsubsection{Grammar}
+The are still many constructions that are not covered by the grammar.
+As example we can mention
+\textbf{object shift}
+\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
+a. & jag & ser&  honom&  inte&~~ & b.&   * jag &ser &huset     &inte\\
+   & I   & see&  him  &  not & & b.&    I   &see &the house &not\\
+\end{tabular}}
+\textbf{bare indefinites}
+\enumsentence{\shortex{4}
+{Vi & ska & köpa & hund} 
+{we& will & buy & dog}
+{``We will buy a dog"}}
+\textbf{måttsattribut}
+\enumsentence{
+{Antalet bilar, ett antal bilar, ett stort antal}}
+Genitive; Not like 'sonen till mig' -> antalet bilars??\\
+
+Probably many more, interesting work, but also infinite.
+\subsubsection{Lexicon}
+The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
+Also to analyse what's wrong, or use the words with a word guesser. The code
+has been adapted for this.
+For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
+category. How this works, look at tags. Small amount, can be done manually.
+
+\subsection{Making the grammar robust}
+If we want to use the grammar for free parsing, this is now a very important part.
+Need to use stastics methods. The parser needs to be able to handle
+ellipses and long sentences, names and numbers.
 Handling unknown words and , handling unknown grammar, 
-how to make the parser robust, ellipses and long sentences, names and numbers.
-What is still needed in the grammar, how important that is. 
-\subsection{Grammar}
-köpa bil 
-object shift
-\subsection{Parallel parsing/Chunk parsing}
-\subsection{Lexicon}
+\subsubsection{Parallel parsing/Chunk parsing}
+\subsubsection{Lexicon}
 compounds
@@
 exchange words.
-%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
-%Also to analyse what's wrong, or use the words with a word guesser. The code
-%has been adapted for this.
-%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
-%category. How this works, look at tags. Small amount, can be done manually.
-
-\subsection{Valency}
+%process list from saldo.
+\subsubsection{Valency}
 \label{sec:futureValency}
@@
 
-\subsection{Probabilities}
+\subsubsection{Probabilities}
 \label{sec:futureProbabilities}
@@
 %Getting and using probabilities, to disambiguate.
-%process list from saldo.
+% ontologies?
+
 %Use multiligual treebank?
-%stop ambiguities by fixing so that Gen of a quant isn't a quant, or better, that all fields
-%have a genitive field.
-%More fields in the grammar? Or leave it all to the robust parser
 

### 2011-12-17 doc/report/report.tex
report: evaluation, future work

@@
 \begin{itemize}
-\item one large-scale GF lexicon and a program to redo the importation when needed
+\item a large-scale GF lexicon and a program to redo the importation when needed
 \item a comparison between GF and other annotations
@@
 
-The increase of the grammatical coverage is hard to compare, since
-The grammar in it self is not at all robust, and one unknown word will cause a whole sentence to 
-fail. 
-An estimation can be given by looking at the results from running the test suite.
-The first grammar could parse almost halv of the sentences, and the final
-grammar could parse 65\%. 
-When parsing text, we get low results, for this reason. Not yet interesting to talk about
-coverage, but focus on quality and ability to scale up. Our goal standarn is so far Talbanken,
-but considering that the low coverage of the grammar without statisticts, a comparison would not be
-describing anything, to little indata.
-Instead we have the luck to have an expert and her presence has increased the standard.
-No syntactic theory will ever cover all language, and this should just be seen as an important
+The grammar has been extended and enhanced, and its current status is
+a specialized extension of the resource grammar. When it comes to parsing,
+however, we do not get far without robustness.
+The grammar in itself is by no means robust, and just one unknown word,
+unexpected punctuation
+or missing phrase will cause the parsing of the whole sentence to fail. 
+Our goal is to parse Talbanken, but considering that no syntactic theory is capable
+of covering a whole natural language, our grammar should just be seen as an important
 fragment.
-
-From the renewed import of Saldo, we got a much better coverage especially considering
-that the old version was missing many of the most common words. Bo, göra, se.
-However, the lexicon is now to big to use as a normal GF lexicon, since it will request
-too much memory to parse most sentences. We will need to handle this in a clever way,
-see future work.
+A comparison would not be of much value of this state since the coverage of Talbanken
+is too small to do be able to get any interesting analysis.\\
+
+An estimation of the improvement can be given by looking at the results from
+running the testsuite used for the grammar development.
+The sentences given in the testsuite are short and the words are in most cases known
+to the lexicon, but there are also constructions that have not been implemented 
+during this project.
+The first grammar could parse almost half of the sentences, the result for the final
+grammar was 65\%. 
+Parsing bigger parts of Talbanken would give very low results, for the reasons stated.
+% try NPs?
+It is thus not yet interesting to talk about coverage, but about the quality and
+the ability to scale up, which has so far proved to be good. The constructions
+that we have focused on have all been possible to implement, although the amount
+of work to get it ~~right has varied.
+I further believe that the presence of a Swedish expert has
+increased the standard substantially.\\
+
+By the renewed import of Saldo, we doubled the size of the lexicon and thereby
+added many of the commonly used irregular words that were missing from the older
+version. This is of course a big improvement.
+However, the lexical part still requires much work before it can be made use of.
+We need valency information to make a good analysis. The lexicon is also too
+big to use in a trivial manner, its size makes the incremental parsing algorithm
+use too much memory. To solve this, we need to use the lexicon data more cleverly.
+
 %Results from mapping, can probably be improved, bit by bit, depends much on words.
@@
 \label{sec:future}
-A part of the project is finished, but the project is not complete, and will never be.
-Many interesting future directions. We will first look at the ones which are 
-direct consequences of what we did so far. 
+Although this one part of the project is finished, there are many interesting future
+directions. As a natural language processing-project ~ it will hardely veer be entirely
+complete.
+The future work described in this section is divided into two categories:
+the ones aiming at making the parser robust and the ones that can be
+seen  as extensions of the work done so far.
+
 \subsection{Enhancments and completitions}
@@
 >> why is this significant, other grammars can use this Swedish grammar.
->> well known (comp. ling) trade-off between depth and coverage. for now depth, robustness in in future
->> work leads to coverage.
+>> well known (comp. ling) trade-off between depth and coverage. for now depth,
+>> robustness in in future work leads to coverage.
 >> reusablilty, freely available.
 >> have experiments of how to extract info from talbanken.
+By combining GF with a large lexicon, we acquire a deep parser for 
+Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
+We have developed tools for extending the lexicon with words from Talbanken,
+and we show how to make use of the information in the manually tagged
+treebank.
+The usage of GF allows us to start from
+a well-defined system for describing grammar, as well as tools for
+parsing.
+
+We are building a parser which is based on grammatical rules rather than
+statistics. The parser will therefore validate that
+a given sentence is grammatically correct according to the rules defined.
+Chosen parts of the parser or grammar may be used in other applications
+which deal with controlled natural languages. Automatic language
+generation from the grammar is also provided by GF.
+
 %There is obviously a lot do still, will never get finish with NPL.

### 2011-12-18 doc/report/report.tex
report getting finished, figures and papers of pargram

@@
 \usepackage{tipa}
+\usepackage{textcomp}
 % exempel: 2) Inte vet jag
@@
 that it is also possible to use the framework for parsing free language \cite{patent}.
- >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
- --> sentence about connection GF - grammar based parser.
+% >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
+% --> sentence about connection GF - grammar based parser.
 Parsing, that is the task of automatically identifying morphological and
@@
 goal is
-to create a open source Swedish grammar and parser that can be continuously enhanced and
+to create a open source Swedish grammar, from which we derive a parser
+accepting all sentences described by the given rules.
+As a freely available resource, it can be continuously enhanced and
 made use of in other projects.\\
-To build a parser or a grammar from scratch would be not only time consuming but
-also a ~~great theoretical and technical challenge~~.
->> not duplicate, already existing resources.
+To build the grammar from scratch would be not only time consuming but
+would also mean that already existing resources were be reimplemented.
 In order to not reinvent the wheel
@@
 Chapters \ref{sec:prog.saldo} - \ref{sec:Mapping} present the methodology,
-implementation and some results. Finally, conclusions, an evaluation and a list
-of future work are presented in chapter \ref{sec:results}.
+implementation and some results. Finally, the conclusion and evaluation are
+presented in chapter \ref{sec:results} together with a list of future work .
 
@@
 random generation, word alignment, morphological quizzes etc.
->> shell reference, servers, cloud. \footnote{http://www.grammaticalframework.org/demos/gfse/}
+An intercative GF editor and the GF shell can be tried out online 
+\footnote{http://www.grammaticalframework.org/demos/gfse/}.
 \begin{figure}[h]
@@
 %>> use talk here and move key to other place??
->> talk!
-The framework is also used in the
+The project talk \cite{talk} used GF to describe dialogue grammars and
+the framework is further used in the
 European project MOLTO \footnote{http://www.molto-project.eu/} for online
@@
 \subsection{Writing a GF grammar}
+\label{sec:writegf}
 The key feature in GF is the distinction between
@@
 representing one language. As a comparison, the abstract and concrete syntax
-may be thought of as f-structures and c-structures in Lexical Functional Grammar (Bresnan and Kaplan, 1982).
+may be thought of as f-structures and c-structures in Lexical Functional
+Grammar (Bresnan and Kaplan, 1982).
 
@@
 language blabla (xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
-CLE is also reported to be very limited (ref).\\
+CLE is also reported to be very limited (Nivre,2002).\\
 
 In the TAG formalism (Joshi,1975), there are projects on getting open source, wide-coverage grammars
-for ie. English and Korean, but, to our knowledge, not for Swedish.
-
-ParGram! More parallel development than Lingo.
+for ie. English and Korean, but, to our knowledge, not for Swedish.  \\
+
+The ParGram project aims at making wide coverage grammars using the Lexicon Functional Grammar approach
+(Bresnan and Kaplan, 1982).
+The grammars are implemented in parallel in order to coordinate the analyses of
+different languages and there are now grammars for English, German, Japanese and Norweigan.
 
@@
 analysis of the most important Swedish constructions.
-The starting point has been the GF resource grammar and our new grammar is
+The starting point has been the GF resource grammar and the new implementation is
 still compatible with this.
@@
 explains the development during this project.
->> the grammar will not be totally covering, but interesting and useful.
 
@@
 \label{sec:swegf}
->> diagram of modules, etc.
->> gf has pos-driven, leads to part-of-sentence analyse. first have v2,a2,vs etc.
->> they may be combined in traditional way, and so the types depend and relate to each other
->> see slides from summer school.
+%>> gf has pos-driven, leads to part-of-sentence analyse. first have v2,a2,vs etc.
+%>> they may be combined in traditional way, and so the types depend and relate to each other
+%>> see slides from summer school.
 >> functions for predication, complementation.
@@
 >> parts of sentences, like subj, obj, attributes.
->> worked well until now, but particular langs. needs particular solutions. --> reflexives!
->> no syntactic theory can cover a whole language, and so gf can't either.
->> shares 80\% with scandinavian, mostly words that differ.
+%>> worked well until now, but particular langs. needs particular solutions. --> reflexives!
+%>> no syntactic theory can cover a whole language, and so gf can't either.
+>> shares 80\% with scandinavian, mostly words that differ.\\
+>> complete morphology, limited syntax
 The GF resource grammars contains a fundamental description of Swedish,
@@
 
-All categories need tables covering the forms the phrase can be used in. They also
+The analysis preformed by GF is driven by the parts of speech, which are combined into
+parts of sentences.
+Figure \ref{fig:gfpic} shows the different categories, or types, used in the
+resource grammars. 
+A lexicon normally consist of words from the open word-classes, shown in rectangular boxes in
+the picture. Each word has a type describing its word class, and the abstract grammar
+consists of rules for how the parts of speech can be used -- how the types relate
+to each other.
+\begin{figure}[h]
+\includegraphics[height=90mm]{categories.png}
+\caption{The types of the GF resource grammars.}
+\label{fig:gfpic}
+\end{figure}
+
+A category is represented by a table covering the forms the phrase can be used in.
+They may also
 need to carry information about inherent features that may matter to other phrases.
-Section \ref{gf:swe} contained an easy example with nouns. We used the categories \verb-N-
+Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
 and \verb-N'- to separated between nouns with repsectively without a set number.
@@
 \verb-CN-, common noun phrases which may be modified by adjectival phrases or conjucted.
+The \verb-N- category is implemented as an inflectiontable and field keeping information
+about the gender.
 A \verb-CN- may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
@@
 
+When a category has been formed, there is no longer any information available
+about how it was put together. This is a result of the functional approach
+of GF. In the case of the of the \verb-CN- above,
+we would like to know whether
+the it has been modified or not, that is, whether the rule
+\verb|AdjCN : AP -> CN -> CN ;| has been applied. Insted the information
+is passed on by an inherent feature, a flag telling set to tell
+if the \verb-AdjCN- was applied.
+
 In gf, only personal pronouns are considered to belong to PN. The rests, eg.
@@
 The resource grammar is meant to cover the most important grammatical structures
-present in all? languages. We have increased this by adding language specific
-constructions. Talbanken has been our inspiration.
-From the things listed in section \ref{sec:swedish}, the two first are basic
-enough to be covered by the resources. The others have been added.
->> hard to estimate before, since no big use case/resource.
->> after initial experiments, first evaluation. found much room for improvement. 
->> complete morphology, limited syntax
->> discovered constructions missing.
+present in many languages, but not the language specific constructions.
+It has earlier been hard to identify missing constructions of the Swedish
+implementation, since there was no big resource to evaluate it on.
+When first conducting tests in this project, we found much room for improvement.
+From the topcis listed in section \ref{sec:swedish}, the post-nominal artciles
+and the verb-second property
+were covered by the resources, as well as the periphrastic passive.
+A start on topicalisation was implemented, while
+the others are examples of what has been added during this project.
+%enough to be covered by the resources. The others have been added.
+%We have increased this by adding language specific
+%constructions. Talbanken has been our inspiration.
+Even though the grammar will not be totally covering, we belive that
+the work done is both interesting and useful.
+%>> the grammar will not be totally covering, but interesting and useful.
+%>> hard to estimate before, since no big use case/resource.
+%>> after initial experiments, first evaluation. found much room for improvement. 
+%>> discovered constructions missing.
 
@@
 than the default.
-\verb|PassVPSlash : VPSlash  -> VP | \\
-\verb|              ta -> togs| \\
-\verb|              erbjöd -> erbjöds| \\
 The grammar also allows not only V2 but all verb phrases missing an object to 
 form passives:
+\begin{verbatim}
+PassVPSlash : VPSlash -> VP ;
+              ta      -> togs
+              erbjöd  -> erbjöds
+\end{verbatim}
 \enumsentence{First place in two-place verbs\\\shortex{3}
 {Jobbet & erbjöds & henne }
-{the job & offered+\textbf{s}& henne}
+{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ henne}}
 {``The job was offered to her"}}
@@
 {Hon & erbjöds & jobbet}
-{she & offered+\textbf{s}& the job}
+{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
 {``She was offered the job"}}
@@
 {Huset & målades & rött}
-{the house & painted+\textbf{s}& red}
+{\emph{the house} & \emph{painted+\textbf{s}}& \emph{red}}
 {``The house was painted red"}}
@@
 {Det & sitter & en & fågel & på & taket}
-{It & sits & a & bird & on & the roof}
+{\emph{it }& \emph{sits }& \emph{a} & \emph{bird} &\emph{ on} & \emph{the roof}}
 {``There is a bird sitting on the roof''}}
@@
 {*Det & äter & en & fågel & frön & på & taket}
-{it & eats & a & bird & seeds & on & the roof}}
+{~\emph{it} & \emph{eats} & \emph{a} & \emph{bird} & \emph{seeds} & \emph{on }& \emph{the roof}}}
 
@@
 {Det&dricks&mycket & öl& nuförtiden}
-{It&drink+s&much & beer& nowadays}
+{\emph{It}&\emph{drink+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
 {A lot of beer is being drunk these days}}
@@
 {Det & finns & kaffe}
-{It & exist & coffee}
+{\emph{it }& \emph{exist }& \emph{coffee}}
 {''There is coffee"}}
@@
 {Det & saknas&  kaffe}
-{It & misses & coffee}
+{\emph{it }& \emph{misses} & \emph{coffee}}
 {''There is no coffee"}}
@@
 {Det & fattas & kaffe}
-{It & lacks & coffee}
+{\emph{it} & \emph{lacks} & \emph{coffee}}
 {''There is no coffee"}}
@@
 \subsection{Miscellaneous}
+This section covers some minor changes done in the grammar. They are described
+as documentation of the implementation, and to illustrate some problems
+and solution of more general nature.
 \subsubsection{Relative clauses}
 The resource grammar already gave a good coverage of relative clauses and embedded sentences.
-All sentences in example \ref{sent:relex} was accepted.
+All construction used in examples \ref{sent:relex}a-d was accepted.
 %provided a number of ways to express relative and embedded clauses, such as
@@
 
-The grammar has been extended to accept a few more constructions:
+The grammar has been extended to accept a definite article to be used with a
+noun in indefinite form whenever the noun phrase is followed by a restrictive
+relative clause.
 \enumsentence{\shortex{7}
-{\textbf{De uppfattningar}& som& förs& fram ...} 
-{The opinions& that& put+{\sc s} & forward ...}
+{De uppfattningar& som& förs& fram ...} 
+{\emph{the opinions}& \emph{that}&\emph{ put+{\textbf s}} & \emph{forward ...}}
 {``The opinions presented ..."}}
-The definite artcile \emph{``de"} is used together with a indefinite
-noun. This is only correct for the definite articles and not determiners
-and requires that the noun phrase is followud by a restrictive
-relative clauses or together with the verb \emph{``finnas"}.
-In other cases definitev form must be used, cf. sentence \ref{sent:debocker}.
-\enumsentence{\textbf{De uppfattningarna} förs fram.\\
+%or together with the verb \emph{``finnas"}.
+When no relative clause is present, the definitive form must be used, cf.
+sentence \ref{sent:debocker}.
+\enumsentence{De uppfattningar\textbf{na} förs fram.\\
 ``The opinions are presented."}\label{sent:debocker}
 
-A few problematic things can be illuminated:
-Since the constructions are inspired by other languages, some of them may sound quite
-non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
-and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
-\emph{`She sleeps, which is good'} as
-\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
-embedded in the relative clause. This is easily fixed, by adding an extra
-parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
-Another complication stems from \verb|RelCl| which express the english \emph{`such that'}.
-The Swedish version \emph{`sådan att'} sounds stelbent, when used outside of logic books.
+Apart for this, only corrections have been done, exemplified in
+this sentence.
+\enumsentence{
+Hon sover \textbf{som} är bra ~ \textrightarrow ~ Hon sover \textbf{vilket} är bra\\
+``She sleeps, which is good"}
+
+%{\emph{the opinions}& \emph{that}&\emph{ put+{\textbf s}} & \emph{forward ...}}
+%{``The opinions presented ..."}}
+%
+%A few problematic things can be illuminated:
+%Since the constructions are inspired by other languages, some of them may sound quite
+%non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
+%and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
+%\emph{`She sleeps, which is good'} as
+%\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
+%embedded in the relative clause. This is easily fixed, by adding an extra
+%parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
+As a paranthesis, some complications regarding the function \verb-RelCl- might
+be pointed out.
+%Another complication stems from \verb|RelCl| which express the 
+The english implementation of this construction is \emph{`such that'},
+and the Swedish version \emph{`sådan att'} sounds awkward, when used outside of
+logic books.
 \enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
-This relative clause in the sentence is, in GF, put together by two constituents: \emph{`` katt"}
-of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause). They are
-glued together by \emph{``sådan att"} and \emph{``katt"} should agree with \emph{``den"}. 
+This relative clause \emph{``en katt sådan att den inte fäller"} is, in GF, put
+together by two constituents: \emph{``katt"}
+of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause), glued
+together by \emph{``sådan att"}.
+As a subject \emph{`den'} is of type \verb-NP- and does not depend on any
+agreement, although it should agree with the antecedent, \emph{`katt'}. 
 \enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-However, this is not the case and sentence \ref{sent:detfaller} is currently
+Changing the type of \verb-NP- would amount in a bigger reconstruction which therefore
+would require a carfully prepared analysis and motivation. 
+Consequentelly, sentence \ref{sent:detfaller} is still
 accepted by the grammar.
@@
 The \verb-GenNP- function was severly overgenerating when this project started, even
-though its implementation was not erroneous in it self. The overgeneration stemmed
-from two places.Quantifiers does not have any possessive forms, so when a quantifier is turned into
-a NP (via a Det), its only field was copied and used for both nominative and genitive.
-In combination, \verb-GenNP-
-introduced a cycle, a Quant may
-be turned into a determiner which can be turned into a NP, which can be turned
-into a Quant.
+though its implementation was not erroneous in itself. The cause of the overgeneration 
+was a combination of two things:
+\begin{itemize}
+\item Quantifiers does not have any possessive
+forms, so when a quantifier is turned into
+a NP (by the functions \verb-DetQuant- and \verb-DetNP-), its genitive field gets
+the same form as its nomitative field.
+\item In combination, \verb-GenNP-
+introduced a cycle, a \verb-Quant- may be turned into a \verb-NP-, and now also
+a \verb-NP- can be turned into a Quant.
+\end{itemize}
 \begin{verbatim}
@@
 \end{verbatim}
-We know had that a noun phrase, say `katten' could be turned into a quantifier,
-by using its genitive form, `kattens'. This could be used as a determiner, and
-consequentelly as a noun phrase `Det är kattens'. But if we try to use `kattens'
-in genitive, we get `kattens' again. The function \verb-GenNP- could be applied
+We had that a noun phrase, eg \emph{`katten'} could be turned into a quantifier,
+by using its genitive form, \emph{``kattens'}. This could be used as a determiner,
+\enumsentence{Det är kattens mat}
+and consequentelly as a noun phrase:
+\enumsentence{Det är kattens.}
+The possessive form of this \verb-NP- is now the same as its nominative:
+\emph{`kattens'}. Hence, the function \verb-GenNP- could be applied
 any number of times without any change in the output.
 
-The problem originated in the function \verb-DetNP- that put the possessive form
-to be the same as the nominative. If we instead let it append an 
-\emph{s} to the word, \verb-GenNP- will stop overgenerating. Besides, the grammar 
-will allow genitive forms of quantifiers
-\enumsentence{Somligas skor står kvar i hallen\\
-``Some peoples shoes are still in the hall"??}
-
-This reduced the number of parse trees for ``kattens bils hus" from 100 to six.
-The six remaining trees describes different interpretations of the sentence,
-such as `hus' is in genitive or nominative and in either plural or singular.
-kattes bils hus är stort.
-kattens bils hus är stora.
+The problem can be solved by changing the implementation of the function
+\verb-DetNP-, that was responsible for creating the possessive form
+of quantifiers and determiners.
+Instead we let it append an extra
+\emph{s} to the word, so that applying \verb-GenNP- to \emph{`kattens'} results
+in \emph{`katten\textbf{ss}'}. Even though this generates form does not exist, it is perhaps
+the best approximation to what it would look like, given that it indeed was used in 
+Swedish. Most important is however that this changed stopped the \verb-GenNP-
+from introducing unnecessarily ambiguities.
+The number of parse trees for the phrase \emph{``Kattens bils hus"} was reducesed
+from over one hundered to six.
+Besides, the grammar will allow the actually existing genitive
+forms of quantifiers:
+\enumsentence{\shortex{6}
+{Somligas &skor& står& kvar& i& hallen}
+{Some's &shoes& stand& left& in& the hall}
+{``Some peoples shoes are still in the hall"}}
+
+%The six remaining trees describes different interpretations of the sentence,
+%such as `hus' is in genitive or nominative and in either plural or singular.
+%kattes bils hus är stort.
+%kattens bils hus är stora.
 
@@
 could not be identified automatically. Therefore, some experiments of how to extract this 
-information from Talbanken has been conducted. We were interested in words which can be used
-the same ways as `sådana' (\emph{`such'}), which acts either as
-determiners, as in sentence \ref{ex:QuantDet}, or as adjectives as
+information from Talbanken have conducted. We were interested in words which can be used
+the same ways as `sådana' (\emph{`such'}) that acts either as a
+determiner, as in sentence \ref{ex:QuantDet}, or as an adjective as
  in sentence \ref{ex:QuantAdj}. 
@@
 
-By starting from a list containing of all pronouns  in Saldo and all forms they may occure in,
+By starting from a list containing all pronouns in Saldo and all forms they may occure in,
 Talbanken was harvested for the ones that both occured with the determiner tag \verb-DT-
@@
 of words that could easily be analysed and added to the lexicon.
-
-Further, list of all pronous acting as determiners were extracted. This method is
-both faster and more reliable than just going through the list by hand and using
-intraspection or manually looking up each word.
+Further, list of all pronous acting as determiners were extracted.
+
+This method is
+both faster and more reliable than going through the list by hand and categories them
+by manually looking up each word or by using intraspection.
 
@@
 Througout the project regression testing has been used. Every time the grammar was updated
-it was tested against a treebank consisting of 155 sentences, and the number of parse trees
-compared to a standard.
+it was tested against a treebank consisting of 155 sentences, and the number of
+parse trees were compared to a standard.
 The purpose was first of all to
-make sure that the coverage was not decrease, but also to be sure we would notice
-if new ambiguites or overgenerating functions were introduced.
-When possible, the ambiguities could be removed and when not we could at least know about
-them. The grammar will always contain ambiguities, but by being aware helps to
-handle them in other ways, such by statistical means.
-%notice if the grammar starts to overgenerate. Avoiding ambiguities is 
+make sure that the coverage was not decreased, but also to make sure that we
+would notice if new ambiguites or overgenerating functions were introduced.
+When possible, the ambiguities could be removed and when not, we were at least made
+aware of them and could decide whether the increase of coverage were worth the
+increase in ambiguities.
+%The grammar will always contain ambiguities, but by being aware helps to
+%handle them in other ways, such by statistical means.
+
 New functions were also tested by random generating and by creating tables showing
@@
 
->>evolution diagram? also removed unnecessary ambigutiies.
->> how to handle ambigs: know them and controll them. using statistics or ontolgies.
->> that is: disambiguitation by context or by meaning.
 %For testing has been carried out by using a treebank, consisting of sentences used in 
@@
 than Swedish. Also, the Penn Treebank contains a lower number of tags, 82 compared
-to more than 300 in Talbanken. Even if you exclude the tags describing the same word
-class as another tag, Talbanken still leaves us with more than 130 tags. The 
+to more than 300 in Talbanken. Even if the tags describing the same word
+class as another tag are excluded, Talbanken still leaves us with more than 130 tags. The 
 number of ways to combine them are hence much increased, and so is the number of
@@
 
-\chapter{Results and discussion}
+\chapter{Discussion}
 \label{sec:results}
 \section{Evaluation}
+>>evolution diagram? also removed unnecessary ambigutiies.
+>> how to handle ambigs: know them and controll them. using statistics or ontolgies.
+>> that is: disambiguitation by context or by meaning.
 The project has resulted in
@@
 Our goal is to parse Talbanken, but considering that no syntactic theory is capable
-of covering a whole natural language, our grammar should just be seen as an important
-fragment.
-A comparison would not be of much value of this state since the coverage of Talbanken
-is too small to do be able to get any interesting analysis.\\
-
+of covering a whole natural language, our grammar when used without a
+robust layer should just be seen as an important
+fragment.\\
+
+A comparison with Talbanken would not be of much value of this state since the coverage 
+is yet too small to do be able to get any interesting analysis.
 An estimation of the improvement can be given by looking at the results from
@@
 Although this one part of the project is finished, there are many interesting future
-directions. As a natural language processing-project ~ it will hardely veer be entirely
+directions. As a natural language processing-project ~ it will hardely ever be entirely
 complete.
@@
 
-\subsection{Enhancments and completitions}
+\subsection{Enhancements and complements}
 \subsubsection{Grammar}
-The are still many constructions that are not covered by the grammar.
-As example we can mention
-\textbf{object shift}
+All imaginable grammatical constructs cannot be cover by our grammar.
+Some work must be left for the robustness, but some specific constructions
+are desirable and interesting to implement. A few examples are listed here:\\
+
+In many Germanic? languages \textit{object shifts} are common.
 \enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
-a. & jag & ser&  honom&  inte&~~ & b.&   * jag &ser &huset     &inte\\
-   & I   & see&  him  &  not & & b.&    I   &see &the house &not\\
+a. & Jag & ser&  honom& \textbf{inte}&~~ & b.&Jag &ser &\textbf{inte} &honom\\
+   & \emph{I} & \emph{see}&\emph{him}&\emph{not}&&& \emph{I} &\emph{see}&\emph{not}&\emph{him}\\
 \end{tabular}}
-\textbf{bare indefinites}
+Personal pronouns are allowed to shift position with the negation in main clauses.
+(read).
+\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
+a. & *Vi & ser&  varandra&  inte&~~ & b.& * Jag &ser &huset &inte\\
+   & \emph{~we} & \emph{see}&  \emph{each other}&  \emph{not} && b.& \emph{~I}&\emph{see} &\emph{the house} &\emph{not}\\
+\end{tabular}}
+
+This is common in relaxed? language, but since it is rarely found in
+Talbanken's part P, which we have been using, it has not been highly prioritised.\\
+
+Nouns used without any explicit definitness can at this stage be parsed as mass
+nouns.
+\enumsentence{\shortexnt{3}
+{Jag & dricker& vatten}
+{\emph{I}&\emph{drink}&\emph{water} }
+}
+This is not the correct analyse for `hund' in sentence \ref{sent:hund}.
 \enumsentence{\shortex{4}
-{Vi & ska & köpa & hund} 
-{we& will & buy & dog}
-{``We will buy a dog"}}
+{Vi & ska& köpa & hund}
+{\emph{we} &\emph{will}& \emph{buy} & \emph{dog} }
+{We are going to buy a dog}}\label{sent:hund}
+This sentence is instead an example of \textit{bare indefinites}, used for
+... what?\\
+
+\verb| X : NP x + CN + NP --better|\\
+\verb| X : CN x + CN + CN|
+check apposCN.
 \textbf{måttsattribut}
@@
 
-Probably many more, interesting work, but also infinite.
 \subsubsection{Lexicon}
-The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
-Also to analyse what's wrong, or use the words with a word guesser. The code
-has been adapted for this.
-For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
-category. How this works, look at tags. Small amount, can be done manually.
+Our lexicon still lacks some part that can be imported from Saldo.
+The multipleword verbs (\verb-vbm-) and multiple word nouns (\verb-nm-) should
+be imported to an idiom lexicon, which can be used as extensions to the main
+lexicon.
+For the words that we tried but failed to import, a tool for lexicon
+acquisition could be used. The tool developed in the previous part of
+this project\footnote{??} would be suitable.
+All in all, it should be ensured that we use as much information as
+possible from our resources.
+%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
+%Also to analyse what's wrong, or use the words with a word guesser. The code
+%has been adapted for this.
+%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
+%category. How this works, look at tags. Small amount, can be done manually.
 
 \subsection{Making the grammar robust}
-If we want to use the grammar for free parsing, this is now a very important part.
-Need to use stastics methods. The parser needs to be able to handle
-ellipses and long sentences, names and numbers.
-Handling unknown words and , handling unknown grammar, 
+To achieve our goal of free parsing, we need to use statistical methods.
+The parser needs to be able to handle
+ellipses and long sentences, names and numbers,
+compounds words and unknown constructions.
 \subsubsection{Parallel parsing/Chunk parsing}
+Instead of parsing a whole sentence in one go, we would like to
+parallely parse parts of it and then try to combine the pieces.
+This method should be faster then traditional parsing and would
+at the same time give robustness.
+Components that cannot be combined by any grammar rule are combined by a 
+meta function. We might end up with partial parse trees, but we would
+be able to parse at least parts of every sentence and give a more
+interesting result to the user.
+
 \subsubsection{Lexicon}
-compounds
-Speed up! : perhaps have one lexicon for morph. analyse and one for parsing.
-exchange words.
-%process list from saldo.
-\subsubsection{Valency}
 \label{sec:futureValency}
+No matter how much we increase the size of our lexicon, we will
+never cover all compounds. We therefore need to be able to do
+compounding analyses.
+
+We need to come up with better methods for using the lexicon when
+parsing, since its time and memory consumption is very high.
+One possibility is of course to do 
+deeper refinements of the parsing source code, which would be a substantial 
+project. Other solutions are to to either
+adapt the lexicon for its domain and hence making it smaller and faster
+or to adapt the input sentences to a smaller lexicon, by preprocessing .
+
+Further, the lexicon needs valencies since this information lays the
+foundation for the GF analysis. We have already analysed how Talbanken
+annotates valency information, and by using the rules made for the mapping
+described in section \ref{sec:Mapping}, we could extract data from Talbanken,
+showing how words are normally used. There should also be algorithms for
+automatically adding the information to the lexicons. 
+Many Swedish verbs can be used with different number of arguments. Adding
+all this information to the lexicon does not seem to be a good idea considering
+the ambiguities it will lead to and the already high time usage.
+The task should instead be left to the robust layer of the parser, possibly
+implemented by using external resources.
+
 %how to get valency information 'sitter och läser',
@@
 \label{sec:futureProbabilities}
+The grammar will always contain ambiguities but before returing a parse result
+to the user, some analysis should be done to find the most likely tree.
+When the size of the lexicon
+increases, so will the ambiguities as the number of word forms overlapping each other
+gets higher. Our new grammar also have many more rules which also will contribute
+to increase the number to interpret each sentence.
+
+GF already allows the lexicon to have probabilities, and by using our extracted
+GF treebank, we can get reliable statisics for our functions. 
+By implementing dependency probabilities, our results would be even better.
+
 % discuss ambiguites, will always have some and the big lexicon will increase them.
@@
 
-%Use multiligual treebank?
+%Use multiligual treebank
 
@@
 >> reusablilty, freely available.
->> have experiments of how to extract info from talbanken.
-By combining GF with a large lexicon, we acquire a deep parser for 
-Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
-We have developed tools for extending the lexicon with words from Talbanken,
-and we show how to make use of the information in the manually tagged
-treebank.
-The usage of GF allows us to start from
-a well-defined system for describing grammar, as well as tools for
-parsing.
-
-We are building a parser which is based on grammatical rules rather than
-statistics. The parser will therefore validate that
+>> have experiments of how to extract info from talbanken.\\
+We have developed the main components
+for a deep Swedish parser. %To abstract:  by combining resources.
+Our main resource is the grammar formalism GF,
+which allows us to start from a well-defined system for describing grammar.
+The parser is derived from the grammar and will therefore validate that
 a given sentence is grammatically correct according to the rules defined.
-Chosen parts of the parser or grammar may be used in other applications
-which deal with controlled natural languages. Automatic language
-generation from the grammar is also provided by GF.
+All parts, including the underlying grammar, are open-source and 
+may thus be used in applications dealing with controlled natural languages.
 

### 2011-12-19 doc/report/report.tex
report: grammar section revised

@@
 \label{sec:prog.grammar}
-An important part of this project has been to developed a grammar and
-adapting it to cover
-constructions used in Talbanken. A grammar can never be expected to give
-full coverage of a language, but we aim for a grammar fragment which gives a deep
+An important part of this project has been to develope the Swedish GF grammar and
+to adapt it to cover
+constructions used in Talbanken. As a grammar implementation never can be
+expected to give
+full coverage of a language, we aim for a grammar fragment which gives a deep
 analysis of the most important Swedish constructions.
@@
 >> shares 80\% with scandinavian, mostly words that differ.\\
->> complete morphology, limited syntax
+>> complete morphology, limited syntax\\
 The GF resource grammars contains a fundamental description of Swedish,
-covering the morphology and many phrase and clause constructions. Most of the implementation
-is shared with the other Scandinavian languages.
-Apart form describing morphology, the resource grammar deals with
+covering the morphology,
 word order, agreement, tense, basic conjunction etc.
-
+ %and many phrase and clause constructions.
+Due to the syntactic similarities of the Scandinavian languages, much of the implementation
+is shared with Norwegian and Danish. The modules that concern the lexical
+aspects are separate, while 85\% of the syntax description is shared.
+There are about 80 functions, which describe the rules for building phrases and clauses.
+\begin{verbatim}
+PredVP       : NP -> VP -> Cl ; -- Prediciation
+ComplVPSlash : VPSlash  -> VP ; -- Complementation
+\end{verbatim}
 The analysis preformed by GF is driven by the parts of speech, which are combined into
@@
 resource grammars. 
-A lexicon normally consist of words from the open word-classes, shown in rectangular boxes in
-the picture. Each word has a type describing its word class, and the abstract grammar
-consists of rules for how the parts of speech can be used -- how the types relate
-to each other.
+A lexicon normally consists of words from the open word-classes, shown in rectangular boxes in
+the picture. Each lexical entry is assigned with a type describing its word-class,
+and the abstract grammar can be viewed as a set rules showing
+%how the parts of speech can be used --
+how the types relate to each other.
 \begin{figure}[h]
@@
 
-A category is represented by a table covering the forms the phrase can be used in.
+The concrete grammar gives a \textit{linearization type} for each category,
+usually a record containing a table with the forms the phrase can be used in.
 They may also
-need to carry information about inherent features that may matter to other phrases.
+carry information about inherent features that matter to other phrases.
+
+\subsection{Noun phrases}
 Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
@@
 \verb-CN-, common noun phrases which may be modified by adjectival phrases or conjucted.
-The \verb-N- category is implemented as an inflectiontable and field keeping information
-about the gender.
-A \verb-CN- may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
-the number and definitness. Since
-all morphological information is generated by the smart paradigms and contained in the \verb-N-
-the construction of a \verb-NP- more or less consists of putting the pieces together.
-The two basic articles are \verb-IndefArt- and \verb-DefArt- which
-uses the post-modified form of the noun: katten.
-However, we need to consider the rules described in section \ref{sec:swedishnoun}. If we have
-a common noun phrase consisting of the parts \emph{`liten'} (\emph{`small'}) and \emph{`katt'}
-(\emph{`cat'}), there are three ways they can be combined:
+The \verb-N- category is implemented as an inflection table, generated by the smart paradigm,
+and field keeping information about the gender.
 \begin{verbatim}
-DetNP : Det -> CN -> NP ;
-        en    liten katt     -- DIndef
-        den   lilla katten   -- DDef Def
-        min   lilla katt     -- DDef Indef
+{s : Number => Species => Case => Str ; g : Gender}
+flicka_N = {s = {Sg Indef Nom =>  flicka
+                 Sg Indef Gen =>  flickas
+                 Sg Def Nom   => flickan
+                 Sg Def Gen   => flickans
+                 Pl Indef Nom => flickor
+                 Pl Indef Gen => flickors
+                 Pl Def Nom   => flickorna
+                 Pl Def Gen   => flickornas ;
+            g = utr }
 \end{verbatim}
->> compare NP to HPSG.
->> CN may be different in other theories.
->> how we have basic Pred, compl
->> clase is a pre-sentence needs tense, polarity etc.
-Which combination that shoud be used is decided by the determiner. Hence all determiners
-in our grammar must keep information of which definitness they allow. They record field
-has a field storing this and the parameters are called \verb-DIndef-, 
-\verb-DDef Indef- and \verb-DDef Def-. The determiners have arguments telling
-which number the noun phrase should be in.
-We also need to make sure that we do not allow constructions as given in
-\ref{sent:trottkatt}a.
-The definite article  \emph{``en"}is considered to be a quantifier, in plural it
-is the either \emph{``de"} or nothing:
+When a noun is turned into a \verb-CN-, this information is kept, and the noun may
+be modified.
 \begin{verbatim}
-DetNP : Det              -> CN        -> NP ;
-        definite, plural +  katt      =  katter
-        definite, plural +  stor katt =  de stora katterna
+AdjCN : AP -> CN -> CN ; 
 \end{verbatim}
-This difference depends on the common noun phrase, which must carry information
-about whether it has been modified or not.
-The resource grammar have differenciate between quantifiers -- \verb-Quant-,
+The function \verb-DetCN-, determination, creates a
+\verb-NP- by setting the number and definitness of a \verb-CN-.
+%may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
+%the number and definitness.
+%Since all morphological information is generated by the smart paradigms and contained in the \verb-N-
+%the construction of a \verb-NP- more or less consists of putting the pieces together.
+%The definitness is given by a determiner.
+
+\begin{figure}[h]
+\begin{verbatim}
+DetCN : Det                    -> CN     -> NP ;
+        indefinitive, singular -> flicka -> "en flicka"
+        definitive,singular    -> flicka -> "flickan"
+\end{verbatim}        
+\caption{}
+\label{gfcode:DetCN}
+\end{figure}
+
+%The two basic articles are \verb-IndefArt- and \verb-DefArt- which
+%uses the post-modified form of the noun: katten.
+Some rules for determination were described in section \ref{sec:swedishnoun};
+if we have a common noun phrase consisting of the parts \emph{`liten'}
+(\emph{`small'}) and \emph{`katt'} (\emph{`cat'}), there are three ways they
+can be combined:
+\begin{figure}
+\begin{tabular}{|lll|l|}
+\hline
+\textbf{Determiner} & \textbf{Adjective} & \textbf{Noun} & \textbf{DetSpecies}\\
+\hline
+en  &  liten $[${\sc -Def}$]$& katt   $[${\sc -Det}$]$& \textbf{ DIndef}\\
+min &  lilla $[${\sc +Def}$]$& katt   $[${\sc -Det}$]$  & \textbf{ DDef Indef} \\
+den &  lilla $[${\sc +Def}$]$& katten $[${\sc +Det}$]$& \textbf{ DDef Def} \\
+\hline
+\end{tabular}        
+\caption{}\label{tab:detdef}
+\end{figure}
+
+%Which combination that shoud be used is decided by the determiner. Hence all
+Hence, all determiners
+in our grammar must keep information of which definitness they allow, the
+\verb-DetSpecies- parameter. The number is also stored as an inherent 
+feature of the determiner.
+The resource grammar differenciates between quantifiers, 
 determiners \verb-Det- and predeterminers. Predeterminers modify \verb-NP-s
 while the other to modifies \verb-CN-s. The differences are shown in table
-\ref{tab:detquant}
+\ref{tab:detquant}. 
 \begin{figure}[h]
@@
 
+The definite article is considered to be a quantifier, which has the forms
+\emph{en} and \emph{ett} for singular in plural it
+is the either \emph{``de"} or nothing, depending on whether the noun phrase
+has been modified.
+This information must thus be stored in the \verb-NP-. 
+\begin{figure}[h]
+\begin{verbatim}
+DetNP : Det              -> CN        -> NP ;
+        definite, plural +  katt      =  katter
+        definite, plural +  stor katt =  de stora katterna
+\end{verbatim}        
+\caption{}\label{gfcode:defmod}
+\end{figure}
+
+%Their record
+%has a field storing the
+%which may  are called \verb-DIndef-, \verb-DDef Indef- and \verb-DDef Def-. 
+%Additionally, we need to make sure that sentences as 
+%\ref{sent:trottkatt}a are not allowed.
+
+Many words which are considered to be pronouns in other analyses are 
+classified as quantifiers and determiners in GF.
+Only personal pronouns are of the type \verb-Pron-.
+
+>> compare NP to HPSG.
+>> CN may be different in other theories.
+---------
 When a category has been formed, there is no longer any information available
 about how it was put together. This is a result of the functional approach
-of GF. In the case of the of the \verb-CN- above,
+of GF. In the case shown in figure \ref{gfcode:defmod},
 we would like to know whether
-the it has been modified or not, that is, whether the rule
-\verb|AdjCN : AP -> CN -> CN ;| has been applied. Insted the information
-is passed on by an inherent feature, a flag telling set to tell
+the it has been modified or not.
+%that is, whether the function \verb|AdjCN| has been applied.
+This information
+has to be passed on by an inherent feature, a flag set to tell
 if the \verb-AdjCN- was applied.
 
-In gf, only personal pronouns are considered to belong to PN. The rests, eg.
-sådan (\emph{``such"}) are either Quantifiers or Dets.
-
+\subsection{Verb phrases}
 In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
-ser katten". The verb ``såg" (see) takes one object: ``katten". Valency information
-is encoded into the GF categories. The lexicon is therefore a very important part of the
+ser katten". The verb ``såg" (see) takes one object: ``katten".
+In GF, this can be seen on the type of the verb,
+valency information
+is encoded into the lexicon. Transitive verbs have the type \verb-V2-, ditransitive
+\verb-V3- etc. There are types for verb which takes sentences (\verb-VS-) and
+adjectival (\verb-VA-) complements.
+The lexicon is therefore a very important part of the
 grammar. 
@@
 \end{verbatim}
-The simplest idea would let the VP consist of one field where ``se" to ``katten" is
-glued together. But we also want to be able to express sentencens like
-``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
-The \verb-VP- category is hence implemented to ressembel the satsschema (Diedriches,xx), 
-it has multiple fields that may be put in the correct order later. There are fields
+
+The funciton \verb-SlashV2a- lets you create a \verb-VPSlash-, a verb phrase missing a
+noun phrase \cite{gazdar}. We combine the \verb-SlashVP- with the object and
+get a complete verb phrase.
+Both the verb and its complements are stored in the verb phrase category, \verb-VP-.
+%A simple idea would let the \verb-VP- consist of one field where the verb and object are
+%glued together. But we also want to be able to express sentencens like
+%``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
+The \verb-VP- category resembles the Didrichsen's field model,
+it has multiple fields that may be put in the different orders. There are fields
 for negation, different types of adverbs, objects, the finite verb and an infinite verb.
-The fields may in turn be tables showing the differnt forms they may have.
-
+The fields may in turn be tables showing the differnt forms of the component.\\
+
+\begin{tabular}{|l|lllllll|}
+\hline
+\textbf{Subj} && &\textbf{VP}& & & & \\
+    & \sc finit& \sc neg& \sc adV & \sc inf& \sc comp& \sc obj& \sc adv \\
+\hline
+han & har &  inte&  alltid&  tänkt&  på&  henne&  så \\
+\hline
+\end{tabular}\\
+
+\vspace{5mm}
+The fields form the verb phrase are not put in the correct order until a sentence is
+created.\\
+
+Swedish verbs may take up to five arguments \cite[p. 53]{stymne}.
+\enumsentence{Jag tar med mig den till honom}\label{sent:tamed}
+These may be prepositions, particles, reflexive object, indirect objects and direct objects.
+In the grammmar these will be encoded in different ways. The particles are
+given in the lexicon, as well as the prepositions that are choosen by the verbs.
+The entry for `\emph{ta med}' as used in sentence \ref{sent:tamed}, is described as follows:
+%, assuming that we already have an constant \verb-take_V- which produces the correct conjugation for ``ta":
 \begin{verbatim}
-han (har   inte alltid tänkt på henne så)
-     finit  neg  adV    inf  comp obj adv
+ta_med_V3 = dirV3 (reflV (partV (take_V "med")) (mkPrep "till") ;
 \end{verbatim}
-
-The funciton \verb-SlashV2a- lets you create a \textit{VPSlash}, a verb phrase missing a
-noun phrase \cite{gazdar}. We combine the SlashVP with the object and get a complete
-verb phrase.
-\begin{verbatim}
- presens negative  (du ser mig)
-  np.s ++ vp.fin ++ vp.neg ++ vp.obj ; 
-\end{verbatim}
-
-The slash-categories cat be combined with a subject to form a \verb-ClSlash-,
-a clause missing a noun phrase. This is a convenient way to form questions ,
-relative clauses
-and topicalized clauses.
-\begin{verbatim}
-wh-questions: (interrogativ pronoun + ClSlash)
-               vad                 (subject: Johan, vpslash: tittar på)
-relative clauses: object + relative pronoun + ClSlash
-                  katten     som             (subject: Johan, vpslash: tittar på)
-topicalized clauses: object + ClSlash
-                     katten  (subject: Johan, vpslash: tittar på)
-\end{verbatim}
+The category \verb-V3- tells us that the verb takes two objects, % (``den" and ``honom").
+the second one is to be used with the preposition `\emph{till}'. \verb-reflV- shows
+that the verb always is used with a reflexive
+pronoun and \verb-partV- gives the particle.
+The fact that the choosen prepositions is attatched to the verb in the lexicon
+causes the parse tree visualisation algorithm group them together.
+This is also the case for particles, cf. parse tree \ref{pic:tittapart}
+and \ref{pic:tittapa}.
+As already stated, the visualized parse trees is not a complete represtation,
+even if the verb phrases in the the visualizations look the same, the two cases
+are treated and represented differently internally. The fronting of the
+preposition as in of sentence \ref{sent:tittapa}a. is accepted but
+not fronting of particles as in\ref{sent:tittapa}b.
+\eenumsentence{
+\item[a]På pojken tittar du 
+\item[b] *På när du springer tittar jag}\label{sent:tittapa}.
+\begin{figure}[h]
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=80mm]{tittapa.pdf}}
+\hspace{10mm}
+\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=90mm]{tittapart.pdf}}
+\caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
+\label{fig:translationtrees}
+\end{figure}
 
@@
 %Maybe talk about ställa.
-Swedish verbs may take up to four arguments \cite[p. 53]{stymne}.
-\enumsentence{Jag tar med mig den till honom}\label{sent:tamed}
-These may be particles, reflexive object, indirect objects and direct objects.
-In the grammmar these will be encoded in different ways. The particles are
-given in the lexicon, as well as the prepositions that are choosen by the verbs.
-The entry for sentence \ref{sent:tamed} looks like this, assuming that we already
-have an constant \verb-take_V- which produces the correct conjugation for ``ta":
-\verb|ta\_med\_V3 = dirV3 (reflV (partV (take\_V "med")) (mkPrep "till") ;|
-The category \verb-V3- tells us that the verb takes two arguments (``den" and ``honom").
-The preposition is set to ``till", reflV shows that the verb always is used with a reflexive
-pronoun and partV gives the particle.
-The fact that the choosen prepositions is given to the verb in the lexicon makes the 
-parse tree visualisation algorithm put them together. See parse tree \ref{fig:tittapa}.
-This is also the case for particles, see parse tree \ref{tittapart}, but even if
-the visualizations look the same the two cases are treated differently inside the grammar
-to keep the actual difference.
-\enumsentence{På pojken tittar du \\ **På när du springer tittar jag}
-As already stated, the visualized parse trees is not a complete represtation, and the reader
-should not be confused or angry at this.\\
+\subsection{Clauses}
+The category clause, \verb-Cl-, represents a pre-sentence, that does not yet have any tense, polarity
+or word order set.\\
+
+\begin{tabular}{|lll|ll|}
+\hline
+\textbf{Tense} &\textbf{Polarity} &\textbf{Word order} &\textbf{Cl} &\textbf{S} \\
+\hline
+Presens &Negative & Main        & du ser mig & 'Du ser inte mig'\\
+Perfect &Positive & Inverted    & du ser mig & 'Har du sett mig'\\
+Perfect &Negative & Subordinate & du ser mig & 'Du har sett mig'\\
+\hline
+\end{tabular}\\
+
+Like verb phrases, a clause may also be missing an object and then has the type
+\verb-ClSlash-. The \verb-ClSlash- is formed by a \verb-VPSlash- which is 
+given a subject.
+This is a convenient way to form questions , relative clauses
+and topicalized clauses.\\
+
+\begin{tabular}{|l|ll|}
+\hline
+\textbf{Cl}& Johan + tittar på \\
+\hline
+\textbf{Wh-Questions} & Interrogativ pronoun   & Vad tittar Johan på?\\
+\textbf{Relative clauses} & Object + Relative pronoun & Katten som Johan tittar på \\
+\textbf{Topicalized clauses} & Object  & Henne tittar Johan på \\
+\hline
+\end{tabular}\\
 
@@
 
-In some special cases, phrase cannot be put in all forms. 
-Give a good example!!
-It is not quite clear how to handle this at the moment. When
-the grammar is meant for parsing only, it is feasible to put
-a \textit{nonexist} pattern.
-example with this.\\
-
+%In some special cases, phrase cannot be put in all forms. 
+%Give a good example!!
+%It is not quite clear how to handle this at the moment. When
+%the grammar is meant for parsing only, it is feasible to put
+%a \textit{nonexist} pattern.
+%example with this.\\
+
+\subsection{Overview}
+\vspace{4mm}
+The resource grammar could express complex sentences
+as the one in figure \ref{gfSwe:parsetree}.
+%shows the parse tree for sentence 
+%\ref{gfSwe:parseable}.\\
 >> change dot file to have VP NP in correct order
-\enumsentence{\shortexnt{8}
-{Har& han& inte& ätit& de & gula& äpplena& idag?}
-{Has& he& not& eaten& the& yellow& apples& today?}} 
-\label{gfSwe:parseable}
-The resource grammar can construct and parse the sentence in example
-\ref{gfSwe:parseable}
-and the parse tree is represented as in figure \ref{gfSwe:parsetree}.
+%\enumsentence{\shortexnt{8}
+%{Har& han& inte& ätit& de & gula& äpplena& idag?}
+%{Has& he& not& eaten& the& yellow& apples& today?}} 
+%\label{gfSwe:parseable}
 %Covers sentences like
@@
 \label{gfSwe:parsetree}
-\end{figure}
+\end{figure}\\
 Even though the verb phrase \emph{``har inte ätit.."} is discontinuous, the whole
@@
 are connected in the tree, and the subject \emph{``han"} is put between the
-finite verb and the rest of the phrase. At the code level, this is implemented
-using records.
-The linearization rule for questions -- clauses with inverted order -- picks 
-the field for the finite verb and appends it to the subject followed by
-the negation, infinite part and the complement of the verb phrase: \\
-\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
-
-Relative clauses are also covered: 
-\enumsentence{\shortexnt{5}{Hon & ser & pojken & som  &sover}
-              {she & sees & the boy & who & sleeps}}
+finite verb and the rest of the phrase.
+%At the code level, this is implemented %using records.
+%The linearization rule for questions -- clauses with inverted order -- picks 
+%the field for the finite verb and appends it to the subject followed by
+%the negation, infinite part and the complement of the verb phrase: \\
+%\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
+
+The resource grammar covered relative clauses: 
+\eenumsentence{\item[a]\shortexnt{5}{Hon & ser & pojken & som  &sover}
+              {\emph{she} & \emph{sees} & \emph{the boy} & \emph{who} & \emph{sleeps}}
               %{`She sees the sleeping boy'}}
-
-\enumsentence{\shortexnt{6}{Han & ser &katten & han  &tycker & om}
-              {he & sees & the cat & he & likes &}}
+\item[b]\shortexnt{6}{Han & ser &katten & han  &tycker & om}
+              {\emph{he} & \emph{sees} & \emph{the cat} & \emph{he} & \emph{likes} &}}
               %{`She sees the sleeping boy'}}
-In addition to the basic resource grammar, additional constructions had been added
->> core is shared but particular lang. has particular features.
+
+In addition to the core of the resource grammars, which is shared with the other languages implemented in the
+library, 
+%In addition to the basic resource grammar,
+ language specific constructions may bee added
+%>> core is shared but particular lang. has particular features.
 to the module \verb|Extra|. The functions given here do not have to be translatable
 to all other language, but are meant to cover language specific constructions.
-Among those were functions for topicalisation, although not all
-of them were implemented for Swedish.
+Among those were functions for topicalisation: %$ although not all
+%of them were implemented for Swedish.
 %word order, agreement, tempus, adverbs, adjectivs...
@@
 %can be found.
-For example, \verb|FocObj| fronted the object as in sentence \ref{gfSwe:apple}.
+%For example, \verb|FocObj| fronted the object as in sentence \ref{gfSwe:apple}.
 \enumsentence{\shortex{6}
@@
 {`I don't want that apple'}}\label{gfSwe:apple}
-The module provided functions for preposition stranding:
+and for preposition stranding:
 \enumsentence{\begin{tabular}{ll}
@@
 \label{sec:Added}
-The resource grammar is meant to cover the most important grammatical structures
-present in many languages, but not the language specific constructions.
+%The resource grammar is meant to cover the most important grammatical structures
+%present in many languages, but not the language specific constructions.
 It has earlier been hard to identify missing constructions of the Swedish
-implementation, since there was no big resource to evaluate it on.
+implementation, since there was no large resource available to evaluate it on.
 When first conducting tests in this project, we found much room for improvement.
@@
 %constructions. Talbanken has been our inspiration.
-Even though the grammar will not be totally covering, we belive that
-the work done is both interesting and useful.
 %>> the grammar will not be totally covering, but interesting and useful.
@@
 %        {~~`A student wrote the essay.'}}
-Some studies suggest that the s-passive is used in more than 80\% of the times \cite{laanemets}.
+Some studies suggest that the s-passive is used in more than 80\% of the times
+\cite{laanemets}.
 It is not as common in the other Scandinavian languages however, 
-and it can not be used in all tenses for all words. The Norwegian 
+where not all words have passive forms for all tenses. The Norwegian 
 translation of sentence \ref{sent:skrevs} is:
@@
 \ref{sent:skrevs}.
-The resource grammar for Scandinavian therefore ~~expressed~~ the passive using auxiliary verb. \\
-\verb|PassV2 : V2  -> VP | \\
-\verb|         ta -> blev tagen| \\
+The resource grammar for Scandinavian therefore implemented the function for passive,
+\verb-PassV2-, by using auxiliary verb. \\
+\begin{verbatim}
+PassV2 : V2  -> VP ;
+         ta -> blev tagen
+\end{verbatim}
 %>> how are they used with reflexives??
-This rules allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
-turned into complete verb phrases; they no longer need an object.
-The s-passive has now been added as the standard case for the Swedish GF
+The function allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
+turned into complete verb phrases; they no longer need an object.\\
+
+During this project, the s-passive was added as the standard case for the Swedish GF
 grammar. Periphrastic passive is still allowed, but as an alternative rather
 than the default.
-The grammar also allows not only V2 but all verb phrases missing an object to 
+The grammar further allows not only V2 but all verb phrases that misses an object to 
 form passives:
@@
 \end{verbatim}
-\enumsentence{First place in two-place verbs\\\shortex{3}
+\enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
 {Jobbet & erbjöds & henne }
@@
 \label{ex:passV32}
-\enumsentence{Second place in two-place verbs\\\shortex{3}
+\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
 {Hon & erbjöds & jobbet}
@@
 \label{ex:passV33}
-\enumsentence{\shortex{3}
+\enumsentence{\textbf{V2A verb}\\\shortex{3}
 {Huset & målades & rött}
@@
 {~\emph{it} & \emph{eats} & \emph{a} & \emph{bird} & \emph{seeds} & \emph{on }& \emph{the roof}}}
+\label{sent:fagelfron}
 
@@
 {\emph{It}&\emph{drink+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
-{A lot of beer is being drunk these days}}
+{``A lot of beer is being drunk these days"}}
 A very common example of this is sentences with the verbs \emph{finnas},
@@
 \emph{saknas} (\emph{miss}) and \emph{fattas}(\emph{lack}).
-\enumsentence{\shortex{3}
-{Det & finns & kaffe}
-{\emph{it }& \emph{exist }& \emph{coffee}}
-{''There is coffee"}}
-\enumsentence{\shortex{3}
-{Det & saknas&  kaffe}
-{\emph{it }& \emph{misses} & \emph{coffee}}
-{''There is no coffee"}}
-\enumsentence{\shortex{3}
-{Det & fattas & kaffe}
-{\emph{it} & \emph{lacks} & \emph{coffee}}
-{''There is no coffee"}}
-
-A special GF category, \verb|SimpleVP|, was needed for verb phrases of the type
-described above.
-There are also restrictions on the real subject.
+\enumsentence{\begin{tabular}[t]{@{}*{14}{l@{\ }}}
+a. &Det & finns & kaffe &
+b. &Det & saknas&  kaffe &
+c. &Det & fattas & kaffe \\
+&\emph{it }& \emph{exist }& \emph{coffee} &
+ &\emph{it }& \emph{misses} & \emph{coffee} &
+ &\emph{it} & \emph{lacks} & \emph{coffee}\\
+\end{tabular}\\
+{~~''There is coffee"~~ ''There is no coffee"~~ ''There is no coffee"}}
+
+To implement this construction, we needed a
+special GF category, \verb|SimpleVP|, to exclude other verb phrases
+like the one in sentence \ref{sent:fagelfron}. Any intransitive verb can form
+a \verb-SimpleVP-, as can transitive verbs without their objects. The \verb-SimpleVP-
+may be modified by adverbs.
+
+There are also restrictions on the real subject, which is not allowed to be in
+definite form.
 \enumsentence{\begin{tabular}{ll}
@@
 \end{tabular}}
-Expressed in GF terms, the determiner must have definitness DIndef. That is, all determiners
-that requireds the noun and the adjective to be used in indefinite form may be used.
+Neither is sentence \ref{sent:dumfagel} correct.
+\enumsentence{
+*Det sitter min fågel på taket\\
+``There are my bird sitting on the roof''\\
+}\label{sent:dumfagel}
+The destinction lays in the determiner. The different types of determiners are
+showed in figure \ref{tab:detdef}. 
+For real subjects, the determiner must be of type \verb-DIndef-,
+that requires both the noun and the adjective are used in indefinite form.
 %The quantifier must have determiner must havenoun must be in indefinite form,
 %even though some determiners are allowed.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. & Det sitter några fåglar på taket    & & b. & *Det sitter fåglarna på taket\\
-   & ``There are some birds on the roof''& &    & ~``There are the birds on the roof''\\
-\end{tabular}}\label{sent:faglar}
-Sentence \ref{sent:gamlafagel}b exemplifies a determiner that uses the noun in indefinite form,
-but the adjective in definite form.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. & Det sitter en gammal fågel på taket.& & b. & * Det sitter min gamla fågel på taket.\\
-   & ``There are some birds on the roof''& &    &~ ``There is my old bird on the roof.''\\
-\end{tabular}}\label{sent:gamlafagel}
-
+%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+%a. & Det sitter några fåglar på taket    & & b. & Det sitter fåglarna på taket\\
+%   & ``There are some birds on the roof''& &    & ``There are the birds on the roof''\\
+%\end{tabular}}\label{sent:faglar}
+%Sentence \ref{sent:gamlafagel}b exemplifies a determiner that uses the noun in indefinite form,
+%but the adjective in definite form.
+%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+%a. & Det sitter en gammal fågel på taket.& & b. & * Det sitter min gamla fågel på taket.\\
+%   & ``There are some birds on the roof''& &    &~ ``There is my old bird on the roof.''\\
+%\end{tabular}}\label{sent:gamlafagel}
+%
 %%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
@@
 %%Semantic difference, atm we allow any NP. \\
+
 To form a clause with formal subject, we hence need to inspect the determiner.
-\verb|FormalSub : SimpleVP -> Det -> CN -> Cl ;|\\
-The function combines a verb phrase a determiner and a noun phrase in to a clause.
-If the determiner requires a noun in determined form, the clause is put to NONEXIST.
-However, this is not an ideal solution, but it cannot be seen on the type level
-which definiteness a noun phrase has and we therefore need to do it at the runtime? level.
+\begin{verbatim}
+FormalSub : SimpleVP -> Det -> CN -> Cl ;
+\end{verbatim}
+The function combines a verb phrase, a determiner and a noun phrase to a clause.
+If the determiner does not fulfill the requriments stated above, the clause is
+put to NONEXIST.
+This works well for parsing, but leads to problems if the grammar is used
+for random generation. The solution is thus not ideal,
+but since the definiteness of noun phrases or determiners cannot be seen on the
+type level, it cannot be known until runtime whether the determiner is accepted
+in the subject.
 
 % move to future?
-As a future direction it would be intresting to examine if the grammars noun phrases
-should have more information on the type level. In the reflexive objects are added in (see p.
-\pageref{swe:refl}) dependent types are used for showing if a phrase can be used as
-subject. We would also like to produce the sentences, where `av' should be used
+As a future direction it would be intresting to examine the consequences of
+letting the noun phrases have more information on the type level. In the
+implementation for reflexive objects  (see section \ref{sec:reflexives}),
+dependent types are used for showing if a noun phrase 
+needs an antecendant.
+We would also like to differenciate between the \verb-NP-s in sentence
+\ref{sent:mosthorses}a and b, where 
+`\emph{av}' should be used
 only when the noun has an explicit article or determiner.
@@
 \subsection{Formalizing the rules for reflexive pronouns by using dependent types}
->> solution requires new strategies, combining 2 theories! (part of speech, part of sentence)
+\label{sec:reflexives}
 An important piece in a Swedish grammar is the reflexive pronouns and
-the possessive pronouns as described on p. \pageref{swe:refl}.
-The reflexive pronouns require an antecedent with which they agree in
+the reflexive possessive pronouns as described on p. \pageref{swe:refl}.
+The reflexives require an antecedent with which they agree in
 number and person.
@@
 {\sc self's} gloves, he had forgotten on the train.}
-\enumsentence{Han ger sina pengar till sina syskon.\\
-He gives {\sc self's} money to {\sc self's} children.}
+%\enumsentence{Han ger sina pengar till sina syskon.\\
+%He gives {\sc self's} money to {\sc self's} children.}
 \enumsentence{Hon ber alla sina kompisar att gå\\
@@
 When a reflexive pronoun is used in a noun phrase, it is hence not
-allowed as subject. The antecedent must not be outside of the smallest verb phrame??
+allowed as subject. The antecedent must not be out of scope for the closest finite verb.
 The grammar should not accept any of these sentences:
@@
 \enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give {\sc self's} money to {\sc self's}his siblings. }
+I give {\sc self's} money to {\sc self's} siblings. }
 \enumsentence{*Han vill att jag ser på sig.\\
@@
 Apart from these restrictions, noun phrases containing reflexive pronouns
-may be used as any other \verb-NP-. They may beconjuncted (sentence
+may be used as any other \verb-NP-. They may be conjuncted (sentence
 \ref{ex:ReflConj} a and b) and determined (sentence \ref{ex:ReflPredet}).
@@
         mat -> sin mat 
+\end{verbatim}
+They may now be used with predeterminers:
+\begin{verbatim}
 PredetNP : Predet -> NP       -> NP ;
@@
 % say smt about overgeneration?
-Further, the dependence on the antecedent needs to be applied to adverbial phrases and
+The dependence on the antecedent needs to be applied to adverbial phrases and
 adjective phrases as well. The adjective phrase in sentence \ref{ex:ReflAP}a and
@@
 subject was changed to 2nd person, they  would need to change too.
-\enumsentence{\textbf{Du} är längre än \textbf{din} kompis.\\
-You are taller than your friend.} \label{ex:ReflAP2}
-\vspace{-2mm}
-\enumsentence{*Du är här oftare än sin kompis.\\
-You are here more often than his friend.} \label{ex:ReflAdv2}
+\enumsentence{\begin{tabular}{ll}
+\textbf{Du} är längre än \textbf{din} kompis. &*Du är här oftare än \textbf{sin} kompis.\\
+You are taller than your friend. &You are here more often than {\sc self's} friend.\\
+\end{tabular}} \label{ex:ReflAP2} \label{ex:ReflAdv2}
 The dependency is spreading through the grammar and in order to avoid all code
-duplication we should come up with a better idea.
-
+duplication we use another approach.
 When looking at the type of
@@
 \end{verbatim}
-The solution given in this project is making use of this generalization, it
+The solution given in this project is making use of this generalization and
 introduces the use of dependent types in a resource grammar.
 Following the idea given above, we make a difference between subjects and objects, but not
-by giving them different types, but by treating them as instances of the \verb-NP- category.
-The type \verb-NP- \textit{dependes} on its argument, which may either be \verb-Subject- or
+by giving them different types, but by % treating them as instances of the \verb-NP- category.
+letting the type \verb-NP- \textit{depend} on an argument, which may either be \verb-Subject- or
 \verb-Object-.
 \begin{verbatim}
-cat NP (Obj | Subj) ;
+cat NP Boolean ;
 PredVP     : NP Obj  -> VP     -> Cl ;
@@
 
-The solution allows adverbial phrases may also be parts of new noun phrases.
+The solution hence combines the-part-of-speech driven analyse normally preformed by
+GF with a part-of-sentence analyse.
+
+The types for adverbial and adjectival phrases are also turned into dependent
+types.
 \enumsentence{\shortex{3}
@@
 {\scriptsize{\textbf{NP +}} & \scriptsize{\textbf{(Adv with Obj)}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
-{The roof of his house}}
+{\emph{The roof of {\sc self's} house}}}
 %{{\sc NP +} & {\sc (Adv with Obj)} & $\Rightarrow$ {\sc Obj}}
@@
 {\scriptsize{\textbf{Obj +}} & \scriptsize{\textbf{Adv}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
-{His house on the mountain}}
+{\emph{{\sc self's} house on the mountain}}}
 
@@
 \item The goal of the project is to make a language specific grammar, and it would be surprising if this could be
-      done without leaving the pattern laid in by the multilingual abstract.
+      done while leaving the abstract grammar unchanged.
 \item The new grammar can be made compatible with the old implementation.
-\item The separation between object and subject could be useful in other languages. For example, the slavic languages
+\item The possiblity for noun phrases to agreemee with an antecedent 
+      could be useful in other languages. For example, the slavic languages
       have similar use of reflexive pronouns.
@@
 must not be used as subject and they may only be used when the number of the antecendent is plural.
-Sentences like
-'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
-'min syster är inte dum, hon'
-We have a clause of which we need to know the natural gender and number of the subject, since it
-must agree with the added pronoun. sag s 439. fullständigt outforskade!
-
-%or just implement this and remove it here?
-The question is how many types we should give dependent types. CN could also need it,
-in order to say
-'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
-The analysis of this sentence is that
-avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
-and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
-Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
-
 %
@@
 tense, it has been added to the Scandiniavin \verb-Extra- module.
-Table \ref{fig:kommeratt} shows how \emph{``kommer att"} may be used in different
-types of sentences.
+Table \ref{fig:kommeratt} shows how the grammar expresses new tense
+in different types of sentences.
 \begin{figure}[h]
@@
 adverbs as \verb-AdV-.
-Discussion: should han åt bara ett äpple be (bara ett äpple) focusing an NP?
-Then what about: han har bara sovit, han har bara ätit ett äpple.\\
+%The semantic aspect of focusing adverbs remains rather unclear in this
+%representation. 
+%Discussion: should han åt bara ett äpple be (bara ett äpple) focusing an NP?
+%Then what about: han har bara sovit, han har bara ätit ett äpple.\\
 
@@
 %ambiguities.
-\subsubsection{PPartAP}
+%\subsubsection{PPartAP}
 %The VP require many changes PPart, bara.
 
-PPartAP : V2 -> AP ; --VPSlash -> AP ;
-
-About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen
-funna'
-for the last one, we need VPSlash to allow adverb.
+%PPartAP : V2 -> AP ; --VPSlash -> AP ;
+%
+%About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen
+%funna'
+%for the last one, we need VPSlash to allow adverb.
 
 \subsection{Miscellaneous}
-This section covers some minor changes done in the grammar. They are described
+This section covers some minor changes of the grammar.
+They are described
 as documentation of the implementation, and to illustrate some problems
@@
 
-The grammar has been extended to accept a definite article to be used with a
-noun in indefinite form whenever the noun phrase is followed by a restrictive
+The grammar has been extended to accept a definite article for
+nouns in indefinite form, whenever the noun phrase is followed by a restrictive
 relative clause.
@@
 {De uppfattningar& som& förs& fram ...} 
-{\emph{the opinions}& \emph{that}&\emph{ put+{\textbf s}} & \emph{forward ...}}
+{\emph{the opinions}& \emph{that}&\emph{ put+{\sc passive}} & \emph{forward ...}}
 {``The opinions presented ..."}}
@@
 As a subject \emph{`den'} is of type \verb-NP- and does not depend on any
-agreement, although it should agree with the antecedent, \emph{`katt'}. 
+agreement. However, it shoud agree with its antecedent, \emph{`katt'}. 
 \enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-Changing the type of \verb-NP- would amount in a bigger reconstruction which therefore
-would require a carfully prepared analysis and motivation. 
+The dependent types introduced in section \ref{sec:reflexives} could probaly be
+useful.
+However a carfully prepared analysis and motiv are needed. %ationamount in a bigger reconstruction which therefore
+%would require a. 
 Consequentelly, sentence \ref{sent:detfaller} is still
@@
 %\end{verbatim}
-The \verb-GenNP- function was severly overgenerating when this project started, even
+The \verb-GenNP- function was severely overgenerating when this project started, even
 though its implementation was not erroneous in itself. The cause of the overgeneration 
@@
 the same form as its nomitative field.
-\item In combination, \verb-GenNP-
-introduced a cycle, a \verb-Quant- may be turned into a \verb-NP-, and now also
+\item\verb-GenNP-
+introduced a cycle: a \verb-Quant- may be turned into a \verb-NP-, and now also
 a \verb-NP- can be turned into a Quant.
@@
 The problem can be solved by changing the implementation of the function
-\verb-DetNP-, that was responsible for creating the possessive form
+\verb-DetNP-, that is responsible for creating the possessive form
 of quantifiers and determiners.
-Instead we let it append an extra
+We let it append an extra
 \emph{s} to the word, so that applying \verb-GenNP- to \emph{`kattens'} results
-in \emph{`katten\textbf{ss}'}. Even though this generates form does not exist, it is perhaps
+in \emph{`katten\textbf{ss}'}. Even though this generates a form that does not exist, it is perhaps
 the best approximation to what it would look like, given that it indeed was used in 
-Swedish. Most important is however that this changed stopped the \verb-GenNP-
+Swedish. Most important is however that this stopped the \verb-GenNP-
 from introducing unnecessarily ambiguities.
@@
 would notice if new ambiguites or overgenerating functions were introduced.
-When possible, the ambiguities could be removed and when not, we were at least made
+If so, the ambiguities could be removed when possible, and when not, we were at least made
 aware of them and could decide whether the increase of coverage were worth the
@@
 
+Sentences like
+'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
+'min syster är inte dum, hon'
+We have a clause of which we need to know the natural gender and number of the subject, since it
+must agree with the added pronoun. sag s 439. fullständigt outforskade!
+
+%or just implement this and remove it here?
+The question is how many types we should give dependent types. CN could also need it,
+in order to say
+'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
+The analysis of this sentence is that
+avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
+and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
+Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
+
 Nouns used without any explicit definitness can at this stage be parsed as mass
@@
 
+Even though the grammar will not be totally covering, we belive that
+the work done is both interesting and useful.
+
 %There is obviously a lot do still, will never get finish with NPL.

### 2011-12-19 doc/report/report.tex
try to parse nps

@@
 gives a full syntactic analysis as well as semantics represented in xx. A
-translation for conversations?? to English was implemented. The work was further developed in the spoken
-language blabla (xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
+translation for conversations to English was implemented. The work was further developed in the spoken
+language(xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
 CLE is also reported to be very limited (Nivre,2002).\\
@@
 however, we do not get far without robustness.
-The grammar in itself is by no means robust, and just one unknown word,
-unexpected punctuation
+The grammar in itself is by no means robust, and just one 
+unexpected punctuation mark, unknown word
 or missing phrase will cause the parsing of the whole sentence to fail. 
@@
 
-By the renewed import of Saldo, we doubled the size of the lexicon and thereby
-added many of the commonly used irregular words that were missing from the older
+By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
+added many of the commonly used words that were missing from the older
 version. This is of course a big improvement.
@@
 We need valency information to make a good analysis. The lexicon is also too
-big to use in a trivial manner, its size makes the incremental parsing algorithm
-use too much memory. To solve this, we need to use the lexicon data more cleverly.
+big to use trivially with the current techniques. Its size causes the
+incremental parsing algorithm use more heap memory than what normal computers
+allow. To solve this, we need to use the lexicon data more cleverly.
 
@@
 Although this one part of the project is finished, there are many interesting future
-directions. As a natural language processing-project ~ it will hardely ever be entirely
-complete.
+directions. As a natural language processing-project it will probably never be
+entirely complete.
 The future work described in this section is divided into two categories:
@@
 All imaginable grammatical constructs cannot be cover by our grammar.
-Some work must be left for the robustness, but some specific constructions
-are desirable and interesting to implement. A few examples are listed here:\\
-
-In many Germanic? languages \textit{object shifts} are common.
+Some work must be left for the robustness, but there are specific constructions
+that are desirable and interesting to implement. A few examples are listed here:\\
+
+\begin{itemize}
+\item
+\textit{Pronomial object shifts} are common in Swedish.
 \enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
@@
 \end{tabular}}
-Personal pronouns are allowed to shift position with the negation in main clauses.
-(read).
+Personal pronouns are allowed to shift position with the negation in main clauses, without
+the modal verbs.
+
 \enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
@@
 
-This is common in relaxed? language, but since it is rarely found in
-Talbanken's part P, which we have been using, it has not been highly prioritised.\\
-
-Sentences like
-'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
-'min syster är inte dum, hon'
-We have a clause of which we need to know the natural gender and number of the subject, since it
-must agree with the added pronoun. sag s 439. fullständigt outforskade!
+Although object shifts are frequently used, they are rarely found in
+Talbanken's part P, which have been the inspiration for this project.
+Therefore, it has so far not been prioritised.\\
+
+%Sentences like
+%'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
+%'min syster är inte dum, hon'
+%We have a clause of which we need to know the natural gender and number of the subject, since it
+%must agree with the added pronoun. sag s 439. fullständigt outforskade!
 
 %or just implement this and remove it here?
-The question is how many types we should give dependent types. CN could also need it,
-in order to say
-'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
-The analysis of this sentence is that
-avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
-and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
-Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
-
-Nouns used without any explicit definitness can at this stage be parsed as mass
-nouns.
+%The question is how many types we should give dependent types. CN could also need it,
+%in order to say
+%'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
+%The analysis of this sentence is that
+%avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
+%and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
+%Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
+
+\item
+\textit{Bare indefinites} are at this stage always parsed as mass nouns.
 \enumsentence{\shortexnt{3}
@@
 }
-This is not the correct analyse for `hund' in sentence \ref{sent:hund}.
+This however is not the correct analyse for `hund' in sentence \ref{sent:hund}.
 \enumsentence{\shortex{4}
@@
 {We are going to buy a dog}}\label{sent:hund}
-This sentence is instead an example of \textit{bare indefinites}, used for
-... what?\\
-
-\verb| X : NP x + CN + NP --better|\\
-\verb| X : CN x + CN + CN|
-check apposCN.
-\textbf{måttsattribut}
-\enumsentence{
-{Antalet bilar, ett antal bilar, ett stort antal}}
-Genitive; Not like 'sonen till mig' -> antalet bilars??\\
-
+
+\item
+The implementation of reflexive pronouns can be improved.  %both code-wise
+It should for example be possible to differenciate between 
+object and subject control verbs.
+
+\eenumsentence{
+\item[a]\shortex{7}
+{Han & ber & henne & att & städa & sitt & rum}
+{\emph{he} & \emph{begs} & \emph{her} & \emph{to} & \emph{clean} & {\sc self's} & \emph{room}}
+{He asks her clean her room}
+\item[b]\shortex{7}
+{Han& lovar& henne& att& städa &sitt & rum}
+{\emph{he}& \emph{promises}& \emph{her}& \emph{to}&  \emph{clean} &\sc self's&\emph{room}}
+  {He promises her to clean his room}}
+
+%
+%\verb| X : NP x + CN + NP --better|\\
+%\verb| X : CN x + CN + CN|
+%check apposCN.
+%\textbf{måttsattribut}
+%\enumsentence{
+%{Antalet bilar, ett antal bilar, ett stort antal}}
+%Genitive; Not like 'sonen till mig' -> antalet bilars??\\
+
+\end{itemize}
 \subsubsection{Lexicon}
@@
 ellipses and long sentences, names and numbers,
-compounds words and unknown constructions.
+compound words and unknown constructions.
 \subsubsection{Parallel parsing/Chunk parsing}
@@
 parallely parse parts of it and then try to combine the pieces.
-This method should be faster then traditional parsing and would
+This method should be faster then normal parsing and would
 at the same time give robustness.
@@
 
-We need to come up with better methods for using the lexicon when
+As noted, we additionally need to come up with better methods for using the lexicon when
 parsing, since its time and memory consumption is very high.
@@
 project. Other solutions are to to either
-adapt the lexicon for its domain and hence making it smaller and faster
-or to adapt the input sentences to a smaller lexicon, by preprocessing .
+adapt the automatically lexicon for its domain and hence making it smaller and
+faster, or to adapt the input sentences to a smaller lexicon, by preprocessing.
 
@@
 showing how words are normally used. There should also be algorithms for
-automatically adding the information to the lexicons. 
+automatically adding the information to the lexicon. \\
 Many Swedish verbs can be used with different number of arguments. Adding
@@
 \label{sec:futureProbabilities}
-The grammar will always contain ambiguities but before returing a parse result
-to the user, some analysis should be done to find the most likely tree.
+The grammar will always contain ambiguities but before returing the result
+to the user, some analysis should be done to find the most probable tree.
 When the size of the lexicon
@@
 GF already allows the lexicon to have probabilities, and by using our extracted
-GF treebank, we can get reliable statisics for our functions. 
+GF treebank, we can get more reliable statisics for our functions. 
 By implementing dependency probabilities, our results would be even better.
@@
 All parts, including the underlying grammar, are open-source and 
-may thus be used in applications dealing with controlled natural languages.
+may thus be used in other applications.
 

### 2011-12-19 testSuites/Parsing.hs

@@ -103,8 +103,8 @@ output s c str = do
 
 textType = fromJust $ readType "Text"
 uttType  = fromJust $ readType "Phr"
-lang     = fromJust $ readLanguage "BigTestSwe" --"BigSwe" --"BigNewSwe"
-pgfFile =  "../gf/BigTestOldest.pgf" -- "../gf/BigTest.pgf" -- "../gf/Big.pgf" -- "BigNew.pgf"
+lang     = fromJust $ readLanguage "BigSwe" --"BigSwe" --"BigNewSwe"
+pgfFile =  "../gf/Big.pgf" -- "../gf/BigTest.pgf" -- "../gf/Big.pgf" -- "BigNew.pgf"
 outFile = "testis.txt"
 
 

### 2011-12-19 mapping/Translate.hs
report files created

@@ -53,7 +53,7 @@ type PMonad = (RWS () [String] S)
 
 
 test = False
-usePGF = bigGr
+usePGF = testGr
 testGr = ("../gf/BigTest.pgf","BigTestSwe")
 bigGr  = ("../gf/Big.pgf","BigSwe")
 lang   = fromJust $ readLanguage "BigTestSwe"

### 2011-12-20 doc/report/report.tex
report: version to elisabet

@@
 \abstract{
-This is the report of the work of expanding and enhancing a GF grammar and preparing
-it for parsing of free Swedish. 
-Outside of this project, an equivalent project for English is being worked upon,
-as are techniques for making the parser robust.
-With this in mind, we have extended the current GF grammar to adapt it for
-Swedish specifically so that it can parse an extended set of basic
-constructions. We have completed a method for importing the electronic 
-lexicon Saldo to a GF format and established a connection to the treebank Talbanken.
-The imported lexicon consists of more than 100 000 words
-and the grammar now covers ??\% of the xx part of Talbanken. 
-66\% av easyeasy.
+%This is the report of the work of expanding and enhancing a GF grammar and preparing
+%it for parsing free Swedish. The grammar is a result of combining resources,
+%we have extended the current GF grammar to adapt it for Swedish specifically so
+%that it can parse an language specific constructions. 
+%We show how dependent types can be used to give a neater description of
+%syntax.
+%We have further completed a method for importing the electronic 
+%lexicon Saldo to a GF format and established a connection to the treebank Talbanken.
+%The imported lexicon consists of more than 100 000 words.
+
+%Outside of this project, an equivalent project for English is being worked upon,
+%as are techniques for making the parser robust.
+%With this in mind,
+
+%and the grammar now covers ??\% of the xx part of Talbanken. 
+%66 \% av easyeasy.
 % The mapping between the annotations of Talbanken and GF is automatic. 
@@
 % more intro to gf. dependently typed...high-level formalism
-The grammar formalism GF \cite{gfbok} has so far been successfully used for 
+Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism,
+developed for multilinguality. It is based on Martin-Löf type theory which allows
+reasoning within the programming language.
+GF has so far been successfully used for 
 controlled languages \cite{cnl}, but recent experiments have showed
@@
 % --> sentence about connection GF - grammar based parser.
+This project focuses on implementing a wide-coverage Swedish grammar, which also
+describes a parser accepting all sentences that can be expressed by the grammar.
+
 Parsing, that is the task of automatically identifying morphological and
-syntactical structures to text, is subject to increasing interest, especially
+syntactical structures, is subject to increasing interest, especially
 considering the steadily growing amount of data available online. 
-So far there is no freely available grammar-driven parser that gives a deep analyse for
-Swedish. The fact that %the existing rule-based Swedish
+So far there is no freely available grammar-driven parser that gives a deep
+analysis for Swedish. The fact that %the existing rule-based Swedish
 a parser or grammar is not freely available does not only restrict its
-usage but also its possibilities of being developed. Our
+usage but also its possibilities of being developed and our
 goal is
-to create a open source Swedish grammar, from which we derive a parser
-accepting all sentences described by the given rules.
+to create a open source Swedish grammar. % from which we derive a parser
+%accepting all sentences described by the given rules.
 As a freely available resource, it can be continuously enhanced and
@@
 To build the grammar from scratch would be not only time consuming but
-would also mean that already existing resources were be reimplemented.
-In order to not reinvent the wheel
-we develop our parser by combining well-tested resources.
+would also mean that already existing resources would reimplemented.
+In order to not reinvent the wheel we proceed from a combination of well-tested
+resources.
 We start from GF resource grammar and thereby get
@@
 Chapters \ref{sec:prog.saldo} - \ref{sec:Mapping} present the methodology,
-implementation and some results. Finally, the conclusion and evaluation are
+implementation and some results. First we describe and evaluate the implementation
+of Saldo in chapter \ref{sec:prog.saldo}. The work on the GF grammar is described in
+chapter \ref{sec:prog.grammar}.
+%, section \ref{sec:gfdevel} presents the construction that have been added.
+In chapter \ref{sec:Mapping} we account for an automatic mapping of Talbanken
+trees to GF. %which enables us to extract a large GF treebank.
+
+Finally, the conclusion and evaluation are
 presented in chapter \ref{sec:results} together with a list of future work .
@@
 using GF for parsing free Swedish. 
-In previous work\footnote{??}, the start on extending the Swedish GF grammar was
-made and a tool for lexical acquisition was developed.
+In previous work\footnote{??}, a start was made to extending the Swedish GF grammar
+and a tool for lexical acquisition was developed.
 We now construct a bigger
@@
 ML \cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
-and the built functionality for logic and reasoning 
+and the built-in functionality for logic and reasoning 
 is inspired by \textlambda Prolog \cite{prolog} and 
 by Agda \cite{agda}, with which GF also shares its type checking algorithm.
-The first version of the framework was implemented by the Xerox Research Center
-in Grenoble and it is now mainly developed in Gothenburg. One of the biggest
+The first version of the framework was implemented by Xerox Research Center
+in Grenoble and is now mainly developed in Gothenburg. One of the biggest
 achievements is a library covering the 
-basic morphological and syntactical structures of more than
+basic morphological and syntactic structures of more than
 20 languages (see section \ref{sec:resources}).
@@
 random generation, word alignment, morphological quizzes etc.
-An intercative GF editor and the GF shell can be tried out online 
+An interactive GF editor and the GF shell can be tried out online 
 \footnote{http://www.grammaticalframework.org/demos/gfse/}.
@@
 \caption{Example of how the GF shell can be used. The first command will parse the given
-sentence and visualize the resulting parse tree and the second will visualize the 
-abstract syntax tree.}
+sentence and visualize the resulting parse tree. The second will visualize the 
+abstract syntax tree. The trees will be showed in the pgf-format by the program evince}
 \label{fig:shellvp}
@@
 The reader should keep in mind that the visualized parse trees is not a
-complete representation. It does not reproduce all information from the abstract tree,
+complete representation. It does not reproduce all information from the abstract tree.
 %If the end node of a  branch has no corresponding word, it is excluded from
 %the visualized parse tree. This is the case for the 
-eg. the definiteness and number of the noun \emph{`katten'}, which is shown as
+For example, the definiteness and number of the noun \emph{`katten'}, which is shown as
 \verb-DetQuant DefArt NumSg- in the abstract tree.
@@
 In the corresponding English parse tree, figure \ref{pic:gfEtree},
-the article \emph{``the"} is explicit and therefore shows as a quantifier in the parse tree.\\
+the noun is explicitly quantified by the article \emph{``the"},
+and the determiner is therefore shows in the parse tree.\\
 
@@
 can be limited and we can ensure that the semantics is preserved during
-translation. Inherited ambiguities and the ones arising from multilinguality will
-of course remain, but can be controlled more easily. %Example of domainspecific word?
+translation. Inherited ambiguities as well as ambiguities arising from multilinguality will
+remain, but can be controlled more easily. %Example of domainspecific word?
 The use of controlled language thus gives a possibility of good and reliable translation
@@
 %>> use talk here and move key to other place??
-The project talk \cite{talk} used GF to describe dialogue grammars and
+The project talk \cite{talk} used GF to describe dialog grammars and
 the framework is further used in the
@@
 taking a noun and a verb we can form a sentence.  No information of how this is
-done is given, at this stage. The grammars also defines two words, the noun
+done is given at this stage. The grammar also defines two words, the noun
 \verb|cat_N| and the verb \verb|sleep_V|.
@@
 Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
-for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|,
-and \verb|Pred| simply glues the two strings ``katten" and ``sover" together:\\
+for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|.
+The function \verb|Pred| simply glues the two strings ``katten" and ``sover" together:\\
 \verb|Pred cat sleep = "katten sover"|.\\
@@
 %now we concentrate on the number only. Hence \verb|N| now means a noun in any of the number,
-and we introduce two function, \verb|NSg : N -> N'| and
-\verb|N'Pl : N -> N'| which sets the number.
+and we introduce two functions that set the number: \verb|NSg : N -> N'| and
+\verb|NPl : N -> N'|.
 \begin{figure}[h!]
@@
 When implementing a English version of the grammar, we encounter another
-problem: the verb form depends on the number of the noun. The \verb|N'| needs to carry
-information about which its number and \verb|Pred| must pass this on to the
-verb. Finally, the type of \verb|V| needs a table, depending on the number.
+problem: the verb form depends on the number of the noun. We solve this by letting
+\verb|N'| carry
+information about which its number and letting \verb|Pred| pass this on to the
+verb. Finally, the type of \verb|V| is put to a table, showing the verbs forms for each
+number.
 \begin{figure}[h!]
@@
 %>>Most general resource, first large grammar, covers fundamental syntactics.
-The GF package provides a advantageous resource library \cite{gf-resource}, covering the
+The GF package provides an advantageous resource library \cite{gf-resource}, covering the
 fundamental morphology and syntax of more than 20 languages.
+There is also a small test lexicon included, containing a few hundred common
+words.
 %is a very useful part of the framework. It is provided in the 
@@
 %they constitute  a very useful resource 
-when implementing application grammars. There is also a small test lexicon
-included, containing a few hundred common words.
+when implementing application grammars.
 
@@
 languages are added, the compiler is being improved, ways of using it in more 
-efficient and easygoing manners are added
+efficient and easy-going manners are added
 and the possibilities to use GF in different environments
 increased. There are research on how to make more use of the dependent
-types, for reasoning by using ontologies \cite{ramona}and krasse?? or generating natural
+types, for reasoning by using ontologies \cite{ramona} or generating natural
 language via Montague semantics \cite{montague}.
-Further, experiments on using GF for free parsing has been conducted
-\cite[\textsection 4.1]{gfMech}.
-This is also a future goal for our project and one of the subgoal has been to
-model an important part of Swedish in GF. The starting point was the resource
-grammar to which the result has a strong connection.
+%Further, experiments on using GF for free parsing has been conducted
+%\cite[\textsection 4.1]{gfMech}.
+%This is also a future goal for our project and one of the subgoal has been to
+%model an important part of Swedish in GF. The starting point was the resource
+%grammar to which the result has a strong connection.
 
@@
 In total it consists of four
-parts. Two of them are transcribitions of spoken language, one a collection of
+parts. Two of them are transcriptions of spoken language, one a collection of
 text written by high school students, and one, section P,
@@
 It is released in Malt\footnote{http://w3.msi.vxu.se/~nivre/research/MaltXML.html} 
-and Tiger\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html},
+and Tiger\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html}
 XML-formats
@@
 in section P has been used.
-The treebank has served as an inspiration and an evaluation source throughout this
+The treebank has served as an inspiration and an evaluation source throughout the
 project. An automatic mapping between its trees and the abstract trees from GF has been
@@
 ``växa" is however declined the same way for both interpretations of the word.
-Hence \textit{Saldo's morphological lexicon}, provided under LPGL in XML format,
-only keeps one entry of ``växa".
+Hence \textit{Saldo's morphological lexicon} only keeps one entry of ``växa".
 For the homonyms that do not share the whole inflection tables, different entries are 
@@
 
-For our purpose, only this morphological section was needed.
+For our purpose, only this morphological section, provided under LPGL in XML
+format, was needed.
 The data can be processed and translated to GF format as described in 
@@
 %which may be 
-\textit{neuter} or \textit{common}.\\
+\textit{neuter} or \textit{non-neuter}.\\
 %(or \textit{utrum}) 
@@
  \hline
-\textbf{Singular} & {en katt} & {katten} & \textit{common}\\
+\textbf{Singular} & {en katt} & {katten} & \textit{non-neuter}\\
 & \emph{a cat} & \emph{the cat} &\\
@@
 phrase; sentence \ref{sent:trottkatt}a is not grammatically correct.
-Since adjectives also marks definiteness, it is marked on three places
+As adjectives also marks definiteness, this is marked on three places
 in sentence \ref{sent:trottkatt}b. %, the definiteness is 
@@
 fronted \cite[\textsection 1027]{H&H}.
-Topicalisation is very common, especially for temporal and spatial adverbial phrases.
+This is called topicalisation and is very common, especially for temporal and spatial adverbial phrases.
 The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
@@
 \label{gfPass:s-pass}
-The s-passive is the most commonly used one for both written
+The s-passive is the more commonly used than periphrastic passive, for both written
 and Swedish \cite[Pass. \textsection 1]{SAG} and dominates especially when the subject is inanimate.
@@
 \subsection*{Impersonal constructions}
-Constructions with `det är/var' (\emph{`it is/was'}) is very common in Swedish \cite[\textsection 309d]{H&H}:
+Constructions with `det är/var' (\emph{`it is/was'}) \\is very common in Swedish
+\cite[\textsection 309d]{H&H}:
 \enumsentence{
@@
 {`I'm glad to hear that.'}}
-'Det' (\emph{it}) is also used as formal subject in presentational constructions where
+'Det' %(\emph{it})
+is also used as formal subject in presentational constructions where
 the real subject is put in the position of an object.
@@
 The Scandinavian language have special reflexive pronouns
-and possessive pronouns for the 3rd
-person \cite[\textsection 310 \& 319]{H&H}. They are distinct from the normal 3rd person forms.
+and reflexive possessive pronouns for the 3rd
+person \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd person forms.
 \enumsentence{
@@
 technology tools for Swedish.
-An example is the well-known parser is the data-driven Malt parser \cite{malt},
+An example is the well-known data-driven Malt parser \cite{malt},
 which has been trained on Talbanken. 
@@
 %for the  finding subject, objects, suubject complements (SC) and
-%to link them to their proper regents (main verbs) : over 90\% both precision and recall.
+%to link them to their proper regents (main verbs) : over 90 \% both precision and recall.
 
@@
 The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
-Structure Grammars \cite{hpsg} (HPSG) providing comparability with tools for
+Structure Grammars \cite{hpsg} (HPSG) providing compatibility with tools for
 parsing, evaluation, semantic representations etc.
@@
 gives a full syntactic analysis as well as semantics represented in xx. A
-translation for conversations to English was implemented. The work was further developed in the spoken
-language(xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
+translation to English  was implemented and the work was further developed in the spoken
+language translator (xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
 CLE is also reported to be very limited (Nivre,2002).\\
@@
 In the TAG formalism (Joshi,1975), there are projects on getting open source, wide-coverage grammars
-for ie. English and Korean, but, to our knowledge, not for Swedish.  \\
+for English and Korean, but, to our knowledge, not for Swedish.  \\
 
@@
 The grammars are implemented in parallel in order to coordinate the analyses of
-different languages and there are now grammars for English, German, Japanese and Norweigan.
+different languages and there are now grammars for English, German, Japanese and Norwegian.
 
@@
 %finding subject, objects, subject complements (SC) and
-%to link them to their proper regents (main verbs) : over 90\% both precision and recall.
+%to link them to their proper regents (main verbs) : over 90 \% both precision and recall.
 %uses a dependency grammar, but "The functional description of adverb phrases
@@
 As Saldo is continuously updated, the importing process has been designed to be fast
-and stable enough to be redone any time.
+and stable enough to be redone at any time.
 
@@
 %>> semantic. Not 1-1.
-The basic algorithm for importing Saldo was implemented by Angelov (2009?)
+%TODO when??
+The basic algorithm for importing Saldo was implemented by Angelov (2009)
 and  produces code for a GF lexicon.
@@
 
-When compiling the resulting lexicon, it will produce an inflection table
+All assumed paradigms are printed to a temporary lexicon, 
+which is when compiled produces an inflection table
 for every entry. The tables are compared to the information given
@@
 Each entry in Saldo has an identifier, eg. \verb-äta..vb-, which are used as 
-constant names in the GF lexicon. The identifiers given i Saldo may contain
-special characters that should be avoided in constants in GF. The importation therefore
+constant names in the GF lexicon. However, the identifier may need some renaming since
+the there are special characters  in the Saldo identifiers
+that should be avoided in GF function names. The importation therefore
 needed some renaming. The Swedish letters \emph{å,ä,ö} are translated into
@@
 \begin{verbatim}äta..vb  -> aeta_V \end{verbatim}
-The translation may cause two or more lemmas to share the same
-GF identifier, and to avoid name collition a number is added to the name
+This translation may cause two or more lemmas to share the same
+GF identifier, and to avoid name collision a number is added to the name
 whenever a translation have been done: 
@@
 \label{sec:saldoRes}
-All words from Saldo were not be imported. There are a number of reasons
-for this, the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions(\emph{pp}), numerals(\emph{nl}),
+The resulting dictionary contains more than 100 000 entries, approximately 80 \% 
+of the total size of Saldo.
+There are a number of reasons why some words were not imported,
+the most obvious one is that we do not want all categories from
+Saldo in the GF lexicon. Prepositions %(\emph{pp})
+, numerals %(\emph{nl}),
 personal pronouns etc.
 are assumed to be present in the resource grammars and should not be added again.
-Saldo contains may other pronouns,
-but they are not analysed the same way in GF (see section \ref{sec:swegf}) and
-can thus not simply be extracted -- we first need to find their correct GF category.
+Saldo contains many pronouns which
+are not analysed the same way in GF (see section \ref{sec:swegf}).
+They can thus not simply be extracted -- we first need to find their correct GF category.
 Some experiments on finding
@@
 %structures of the language.
-Categories involving several words (\emph{vbm, nnm})
-are usuallly handled as idioms and should be given in a separate lexicon. In
+Categories involving multiple words %(\emph{vbm, nnm})
+are usually handled as idioms and should be given in a separate lexicon. In
 total six categories were imported: \\
@@
 One reason why the importing phase would fail 
-is that Saldo, unlike GF, only contains the acctually used word forms.
+is that Saldo, unlike GF, only contains the actually used word forms.
 % does not contain information about ~~word forms that are not used~~.
@@
 noun \emph{``glasögon"} (\emph{``glasses"}).
-The smart paradigm requires a singular form, and since the program couldnot
+The smart paradigm requires a singular form, and since the program could not
 find this in Saldo, there was no way of adding the lemma to the lexicon. 
-When the program failed to import a noun, this was often the explanaiton.
+When the program failed to import a noun, this was often the explanation.
 %bbMany nouns fai, \emph{``kläder"} (\emph{``clothes"} etc.
-Words of this type could be added manually, for \emph{``glasögon"} by using the
-ostensibly correct singular form ``glasöga", although this
+Words of this type may be added manually, for \emph{``glasögon"} we could use
+the ostensibly correct singular form ``glasöga", although this
 has another meaning (``glass-eye").
@@
 (\emph{``synas"} \emph{(``show")} or \emph{umgås} \emph{(``socialize")})
-which made up 61.5\% of the failing verbs of type \verb_vb_.\\
+which made up 61.5 \% of the failing verbs of type \verb_vb_.\\
 %``frodas" (``flourish"), ``lyckas" (``succeeed").
@@
 
-The resulting dictionary contains more than 100 000 entries, approximately 80\% (retest!)
-of the total size of Saldo.
 %compared to the total size of Saldo this is .
 When testing the coverage of Talbanken,
-we found that there are around 2500 wordforms were still missing, excluding the one
+we found that there are around 2500 word forms still missing, excluding the one
 tagged as names and numbers. This number may seem very high, but %considering that
-4/5 of the wordforms are compounds. We should also take into consideration that 
+4/5 of the word forms are compounds. We should also take into consideration that 
 we cannot automatically find out how many actually stem from the same word, or
-how many abbreviation that are present (Talbanken also contains a small number of 
-spelling errors, probably are enumerated among our missing words). The majority
-of the words are only used once.\\
+how many abbreviations that are present (Talbanken also contains a small number of 
+spelling errors, which probably are enumerated among our missing words). The majority
+of the missing words are only used once.\\
 
@@
 \hline
-Missing words & $\sim$ 2500 wordforms\\
-Missing words, ignoring compounds & $\sim$ 500 wordforms\\
-Missing words used more than once & $\sim$ 500 wordforms\\
-Missing words used more than once, ignoring compounds & $\sim$ 150 wordforms\\
+Missing words & $\sim$ 2500 word-forms\\
+Missing words, ignoring compounds & $\sim$ 500 word-forms\\
+Missing words used more than once & $\sim$ 500 word forms\\
+Missing words used more than once, ignoring compounds & $\sim$ 150 word-forms\\
 \hline
@@
 composed, consisting of about 1600 entries. Many of those are
-acceptable, such as the ones in table \ref{tab:saldodiff} and reflects the difference 
-made in the analyses, but others are examples of words that are still missing from
+acceptable and reflects the difference 
+made in the analyses, such as the examples in table \ref{tab:saldodiff}.
+Others are examples of words that are still missing from
 the lexicon.
@@
 \label{sec:prog.grammar}
-An important part of this project has been to develope the Swedish GF grammar and
+An important part of this project has been to develop the Swedish GF grammar and
 to adapt it to cover
@@
 still compatible with this.
-Before describing the work progress we will give an introduction to the
+Before describing the work progress in section \ref{sec:Added}, we will give an introduction to the
 resource grammars in general and to the Swedish implementation in particular.
-This is found in section \ref{sec:swegf} while section \ref{sec:Added} shows and
-explains the development during this project.
+%This is found in section \ref{sec:swegf} while section \ref{sec:Added} shows and
+%explains the development during this project.
 
@@
 \section{The Swedish resource grammar}
+\label{sec:swegf}
 %>> start by easy examples, jag ser inte dig, du ser inte mig osv. 
@@
 %>> compare to others, ask elisabet.\\
-\label{sec:swegf}
 %>> gf has pos-driven, leads to part-of-sentence analyse. first have v2,a2,vs etc.
@@
 %>> see slides from summer school.
->> functions for predication, complementation.
->> rather different from others syntactical theories. does not remember syntactical
->> parts of sentences, like subj, obj, attributes.
 %>> worked well until now, but particular langs. needs particular solutions. --> reflexives!
 %>> no syntactic theory can cover a whole language, and so gf can't either.
->> shares 80\% with scandinavian, mostly words that differ.\\
->> complete morphology, limited syntax\\
-The GF resource grammars contains a fundamental description of Swedish,
+
+% TODO!
+%>> functions for predication, complementation.
+%>> rather different from others syntactical theories. does not remember syntactical
+%>> parts of sentences, like subj, obj, attributes.
+%>> shares 80 \% with scandinavian, mostly words that differ.\\
+%>> complete morphology, limited syntax\\
+
+The GF resource grammars gives a fundamental description of Swedish,
 covering the morphology,
@@
 is shared with Norwegian and Danish. The modules that concern the lexical
-aspects are separate, while 85\% of the syntax description is shared.
+aspects are separate, while 85 \% of the syntax description is shared.
 There are about 80 functions, which describe the rules for building phrases and clauses.
@@
 A lexicon normally consists of words from the open word-classes, shown in rectangular boxes in
-the picture. Each lexical entry is assigned with a type describing its word-class,
-and the abstract grammar can be viewed as a set rules showing
+the picture. Each lexical entry is assigned with a type describing its word-class.
+%and the abstract grammar can be viewed as a set rules showing
 %how the parts of speech can be used --
-how the types relate to each other.
+%how the types relate to each other.
 \begin{figure}[h]
@@
 The concrete grammar gives a \textit{linearization type} for each category,
-usually a record containing a table with the forms the phrase can be used in.
+usually a record containing a table with the forms the word or phrase can be used in.
 They may also
-carry information about inherent features that matter to other phrases.
+carry information about inherent features that matter to other parts of the sentence.
 
@@
 Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
-and \verb-N'- to separated between nouns with repsectively without a set number.
+and \verb-N'- to separated between nouns with respectively without a set number.
 The full analysis includes a separation between the categories \verb-N- -- simple nouns --
 and 
-\verb-CN-, common noun phrases which may be modified by adjectival phrases or conjucted.
+\verb-CN- -- common noun phrases -- and \verb-NP- -- noun phrases.
 The \verb-N- category is implemented as an inflection table, generated by the smart paradigm,
-and field keeping information about the gender.
+and a field keeping information about the gender.
+\begin{figure}[h]
 \begin{verbatim}
-{s : Number => Species => Case => Str ; g : Gender}
 flicka_N = {s = {Sg Indef Nom =>  flicka
@@
 \end{verbatim}
-When a noun is turned into a \verb-CN-, this information is kept, and the noun may
-be modified.
+\caption{Representation of the noun `flicka' (\emph{`girl'}) in GF}
+\label{fig:gfpic}
+\end{figure}
+When a noun is turned into a \verb-CN- %, this information is kept, and the noun may
+which may be modified by adjectival phrases or conjucted.
 \begin{verbatim}
@@
 The function \verb-DetCN-, determination, creates a
-\verb-NP- by setting the number and definitness of a \verb-CN-.
+\verb-NP- by setting the number and definiteness of a \verb-CN-.
 %may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
@@
 DetCN : Det                    -> CN     -> NP ;
-        indefinitive, singular -> flicka -> "en flicka"
+        indefinite, singular -> flicka -> "en flicka"
         definitive,singular    -> flicka -> "flickan"
@@
 (\emph{`small'}) and \emph{`katt'} (\emph{`cat'}), there are three ways they
-can be combined:
+can be combined, as shown in figure \ref{tab:detdef}
 \begin{figure}
@@
 Hence, all determiners
-in our grammar must keep information of which definitness they allow, the
-\verb-DetSpecies- parameter. The number is also stored as an inherent 
+in our grammar must keep information of which definiteness they allow; the
+\verb-DetSpecies- parameter stored as an inherent 
 feature of the determiner.
-The resource grammar differenciates between quantifiers, 
+The resource grammar differentiates between quantifiers, 
 determiners \verb-Det- and predeterminers. Predeterminers modify \verb-NP-s
@@
 \begin{tabular}{l|lll}
-                        &\textbf{Has number}& \textbf{Has definitness}& Example  \\ 
+                        &\textbf{Has number}& \textbf{Has definiteness}& Example  \\ 
 \hline
@@
 \textbf{Quantifiers}&-- &\checkmark  & min katt, mina katter, *min katten \\ 
-\textbf{Determiners}   &\checkmark &\checkmark & varje katt, *varje katten  \\ 
+\textbf{Determiners}   &\checkmark &\checkmark & varje katt, *varje katten, *varje katter  \\ 
 \end{tabular}
@@
 The definite article is considered to be a quantifier, which has the forms
-\emph{en} and \emph{ett} for singular in plural it
-is the either \emph{``de"} or nothing, depending on whether the noun phrase
-has been modified.
-This information must thus be stored in the \verb-NP-. 
+\emph{en} and \emph{ett} for singular. In plural it
+is the either \emph{``de"} or nothing, cf. sentence \ref{sent:trottkatt}a and \ref{sent:trottkatt}b.
+We need to know if the common noun phrase 
+has been modified, that is, whether the function \verb_AdjCN_ has been used.
+However, once a category has been formed, there is no longer any information available
+about how it was put together. This is a result of the functional approach
+of GF. Therefore, %that is, whether the function \verb|AdjCN| has been applied.
+this information
+has to be passed on by an inherent feature of the \verb-CN-, a flag set to tell
+if the \verb-AdjCN- was applied.
 \begin{figure}[h]
@@
 
-Many words which are considered to be pronouns in other analyses are 
-classified as quantifiers and determiners in GF.
-Only personal pronouns are of the type \verb-Pron-.
-
->> compare NP to HPSG.
->> CN may be different in other theories.
+The GF category for pronouns \verb-PN- only contains personal pronouns.
+Many words which are considered to be pronouns in other analyses are hence
+classified differently in GF.
+%Only personal pronouns are of the type \verb-Pron-.
+
+%TODO
+%>> compare NP to HPSG.
+%>> CN may be different in other theories.
 ---------
-When a category has been formed, there is no longer any information available
-about how it was put together. This is a result of the functional approach
-of GF. In the case shown in figure \ref{gfcode:defmod},
-we would like to know whether
-the it has been modified or not.
-%that is, whether the function \verb|AdjCN| has been applied.
-This information
-has to be passed on by an inherent feature, a flag set to tell
-if the \verb-AdjCN- was applied.
 
@@
 
-The funciton \verb-SlashV2a- lets you create a \verb-VPSlash-, a verb phrase missing a
-noun phrase \cite{gazdar}. We combine the \verb-SlashVP- with the object and
+The function \verb-SlashV2a- lets you create a \verb-VPSlash-, a verb phrase missing a
+noun phrase \cite{gazdar}. When combine the \verb-SlashVP- with the object we
 get a complete verb phrase.
@@
 %``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
-The \verb-VP- category resembles the Didrichsen's field model,
-it has multiple fields that may be put in the different orders. There are fields
+The \verb-VP- category resembles the Didrichsen's field model.
+%it has multiple fields that may be put in the different orders.
+There are fields
 for negation, different types of adverbs, objects, the finite verb and an infinite verb.
-The fields may in turn be tables showing the differnt forms of the component.\\
-
-\begin{tabular}{|l|lllllll|}
+The fields may in turn be tables showing the different forms of the component.\\
+
+\begin{tabular}{|l|llllllll|}
 \hline
-\textbf{Subj} && &\textbf{VP}& & & & \\
-    & \sc finit& \sc neg& \sc adV & \sc inf& \sc comp& \sc obj& \sc adv \\
+& & &&\textbf{VP}& & & & \\
+\textbf{VP field} && \sc finit& \sc neg& \sc adV & \sc inf& \sc comp& \sc obj& \sc adv \\
 \hline
-han & har &  inte&  alltid&  tänkt&  på&  henne&  så \\
+&\emph{(han)} & har &  inte&  alltid&  tänkt&  på&  henne&  så \\
 \hline
@@
 \vspace{5mm}
-The fields form the verb phrase are not put in the correct order until a sentence is
-created.\\
+The verb phrase fields are not put in their correct order until the tense, type of clause etc is
+known, ie. when a sentence is created.\\
 
@@
 These may be prepositions, particles, reflexive object, indirect objects and direct objects.
-In the grammmar these will be encoded in different ways. The particles are
-given in the lexicon, as well as the prepositions that are choosen by the verbs.
+In the grammar these will be encoded in different ways. The particles are
+given in the lexicon, as well as the prepositions that are chosen by the verbs.
 The entry for `\emph{ta med}' as used in sentence \ref{sent:tamed}, is described as follows:
@@
 that the verb always is used with a reflexive
-pronoun and \verb-partV- gives the particle.
-The fact that the choosen prepositions is attatched to the verb in the lexicon
-causes the parse tree visualisation algorithm group them together.
+pronoun and \verb-partV- gives the particle \emph{`med'}.
+The fact that the chosen prepositions is attached to the verb in the lexicon
+causes the parse tree visualization algorithm to group them together.
 This is also the case for particles, cf. parse tree \ref{pic:tittapart}
 and \ref{pic:tittapa}.
-As already stated, the visualized parse trees is not a complete represtation,
+As already stated, the visualized parse trees is not a complete representation,
 even if the verb phrases in the the visualizations look the same, the two cases
@@
 preposition as in of sentence \ref{sent:tittapa}a. is accepted but
-not fronting of particles as in\ref{sent:tittapa}b.
+fronting of particles as in\ref{sent:tittapa}b. is not.
 \eenumsentence{
@@
 \hline
-Presens &Negative & Main        & du ser mig & 'Du ser inte mig'\\
-Perfect &Positive & Inverted    & du ser mig & 'Har du sett mig'\\
-Perfect &Negative & Subordinate & du ser mig & 'Du har sett mig'\\
+present &negative & main        & du ser mig & 'Du ser inte mig'\\
+perfect &positive & inverted    & du ser mig & 'Har du sett mig'\\
+perfect &negative & subordinate & du ser mig & 'Du har sett mig'\\
 \hline
@@
 given a subject.
-This is a convenient way to form questions , relative clauses
+This is a convenient way to form questions, relative clauses
 and topicalized clauses.\\
@@
 \hline
-\textbf{Wh-Questions} & Interrogativ pronoun   & Vad tittar Johan på?\\
+\textbf{Wh-Questions} & Interrogative pronoun   & Vad tittar Johan på?\\
 \textbf{Relative clauses} & Object + Relative pronoun & Katten som Johan tittar på \\
@@
 \vspace{4mm}
-The resource grammar could express complex sentences
+The original resource grammar could express complex sentences
 as the one in figure \ref{gfSwe:parsetree}.
@@
 %\ref{gfSwe:parseable}.\\
->> change dot file to have VP NP in correct order
+
+%TODO
+%>> change dot file to have VP NP in correct order
+
 %\enumsentence{\shortexnt{8}
@@
 \label{gfSwe:parsetree}
-\end{figure}\\
-Even though the verb phrase \emph{``har inte ätit.."} is discontinuous, the whole
+\end{figure}
+Even though the verb phrase \emph{``har inte ätit de gula äpplena idag"} is discontinuous, the whole
 phrase is still treated as one constituent in GF. The parts
@@
 finite verb and the rest of the phrase.
+At the code level, This is done by putting the \verb-VP- fields in the correct order\\
+\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
 %At the code level, this is implemented %using records.
@@
 %the negation, infinite part and the complement of the verb phrase: \\
-%\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
-
-The resource grammar covered relative clauses: 
+
+The resource grammar also covered relative clauses: 
 \eenumsentence{\item[a]\shortexnt{5}{Hon & ser & pojken & som  &sover}
@@
 
-In addition to the core of the resource grammars, which is shared with the other languages implemented in the
+In addition to the core resource grammars, which is shared with the other languages implemented in the
 library, 
@@
 Stranded preposition & cf.\\
-Vem måste jag akta mig för? & För vem måste jag akta mig?\\
-Who do I need to watch out for? & For whom do I need to watch out?\\
+Vem måste jag akta mig \textbf{för}? & \textbf{För} vem måste jag akta mig?\\
+Who do I need to watch out \textbf{for}? & \textbf{For} whom do I need to watch out?\\
 \end{tabular}}
@@
 implementation, since there was no large resource available to evaluate it on.
-When first conducting tests in this project, we found much room for improvement.
-From the topcis listed in section \ref{sec:swedish}, the post-nominal artciles
+
+Our evaluations are based on Talbanken, and when first conducting tests,
+ we found much room for improvement.
+From the topics listed in section \ref{sec:swedish}, the post-nominal articles
 and the verb-second property
@@
 %        {~~`A student wrote the essay.'}}
-Some studies suggest that the s-passive is used in more than 80\% of the times
+Some studies suggest that the s-passive is used in more than 80 \% of the times
 \cite{laanemets}.
-It is not as common in the other Scandinavian languages however, 
+It is however not as common in the other Scandinavian languages, 
 where not all words have passive forms for all tenses. The Norwegian 
@@
 
-During this project, the s-passive was added as the standard case for the Swedish GF
-grammar. Periphrastic passive is still allowed, but as an alternative rather
+During this project, the s-passive was added as the standard case.
+Periphrastic passive is still allowed, but as an alternative rather
 than the default.
-The grammar further allows not only V2 but all verb phrases that misses an object to 
+The grammar further allows not only V2, but all verb phrases that misses an object, to 
 form passives:
@@
 \enumsentence{\begin{tabular}{ll}
-*Det sitter den fågeln på taket. & *Det sitter fågeln på taket.\\
+*Det sitter \textbf{den fågeln} på taket. & *Det sitter \textbf{fågeln} på taket.\\
 ``There are that bird sitting on the roof'' & ``There are the bird sitting on the roof'' \\
@@
 \enumsentence{
-*Det sitter min fågel på taket\\
+*Det sitter \textbf{min fågel} på taket\\
 ``There are my bird sitting on the roof''\\
 }\label{sent:dumfagel}
-The destinction lays in the determiner. The different types of determiners are
+The distinction lays in the determiner. The different types of determiners are
 showed in figure \ref{tab:detdef}. 
 For real subjects, the determiner must be of type \verb-DIndef-,
-that requires both the noun and the adjective are used in indefinite form.
+that requires both the noun and the adjective to be used in indefinite form.
 %The quantifier must have determiner must havenoun must be in indefinite form,
@@
 %%Semantic difference, atm we allow any NP. \\
-
 To form a clause with formal subject, we hence need to inspect the determiner.
@@
 The function combines a verb phrase, a determiner and a noun phrase to a clause.
-If the determiner does not fulfill the requriments stated above, the clause is
+If the determiner does not fulfill the requirements stated above, the clause is
 put to NONEXIST.
@@
 but since the definiteness of noun phrases or determiners cannot be seen on the
-type level, it cannot be known until runtime whether the determiner is accepted
+type level, it is not known until runtime whether the determiner is accepted
 in the subject.
@@
 % move to future?
-As a future direction it would be intresting to examine the consequences of
+As a future direction it would be interesting to examine the consequences of
 letting the noun phrases have more information on the type level. In the
@@
 dependent types are used for showing if a noun phrase 
-needs an antecendant.
-We would also like to differenciate between the \verb-NP-s in sentence
+needs an antecedent.
+We would also like to differentiate between the \verb-NP-s in sentence
 \ref{sent:mosthorses}a and b, where 
@@
 When a reflexive pronoun is used in a noun phrase, it is hence not
-allowed as subject. The antecedent must not be out of scope for the closest finite verb.
+allowed as subject. The antecedent must not be out of scope for the closest finite verb,
+as in sentence \ref{sent:sesig}.
 The grammar should not accept any of these sentences:
@@
 \enumsentence{*Han vill att jag ser på sig.\\
-He wants me to look at {\sc self}.}
+He wants me to look at {\sc self}.}\label{sent:sesig}
 
@@
 In short, what is wanted can be summarized as follows:\\
-\begin{figure}[h]
+%\begin{figure}[h]
 \begin{enumerate}
@@
 \end{enumerate}
-\caption{Rules for noun phrases with reflexive possessive pronouns}
-\label{fig:reflrestrict}
-\end{figure}
+%\caption{Rules for noun phrases with reflexive possessive pronouns}
+%\label{fig:reflrestrict}
+%\end{figure}
 If we ignore the second requirement, we might come up with a solution where
@@
 This way, using the phrases as subject is excluded, but neither of sentence
-\ref{ex:ReflConj}a, b or \ref{ex:ReflPredet} are allowed.\\
+\ref{ex:ReflConj}a, \ref{ex:ReflConj}b or \ref{ex:ReflPredet} are allowed.\\
 The simplest way to fix this is to treat the phrases as normal \verb_NP_s.
-We can add a rule for using the possessives with common noun phrases:
+We can add a rule that allows using the possessives with common noun phrases:
 \begin{verbatim}
@@
 identical to normal \verb_NP_ except for that it depends on 
-the antecedent. All functions where objects may be used in the same way
-as other \verb-NP-s are duplicated, one version is dealing with
+the antecedent. All functions where objects are modified in the same way
+as other \verb-NP-s are duplicated, one version is dealing 
 with objects and one with \verb-NP-.
@@
 {Taket & på sitt hus &}
-{\scriptsize{\textbf{NP +}} & \scriptsize{\textbf{(Adv with Obj)}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{\scriptsize{\textbf{NP a +}} & \scriptsize{\textbf{(Adv Obj)}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
 {\emph{The roof of {\sc self's} house}}}
@@
 {Sitt hus & på berget &}
-{\scriptsize{\textbf{Obj +}} & \scriptsize{\textbf{Adv}} & $\Rightarrow$ \scriptsize{\textbf{Obj}} }
+{\scriptsize{\textbf{NP Obj + }} & \scriptsize{\textbf{Adv a}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
 {\emph{{\sc self's} house on the mountain}}}
@@
 The use of dependent types separates our grammar from the other resource grammars. Such separation
-is not desirable in it self, but there are at least three reasons to why we believe it is feasible in this case:
+is not desirable in it self, but there are at least three reasons to why we
+believe it is appropriate in this case:
 \begin{itemize}
+\item The new grammar can be made compatible with the old implementation.
+\item The possibility for noun phrases to agree with an antecedent 
+      could be useful in other languages. For example, the Slavic languages
+      have similar reflexive pronouns.
 \item The goal of the project is to make a language specific grammar, and it would be surprising if this could be
       done while leaving the abstract grammar unchanged.
-\item The new grammar can be made compatible with the old implementation.
-\item The possiblity for noun phrases to agreemee with an antecedent 
-      could be useful in other languages. For example, the slavic languages
-      have similar use of reflexive pronouns.
 \end{itemize}
@@
 The rules for reciprocal pronouns could be formalized using the same idea. As reflexive pronouns, reciprocals
-must not be used as subject and they may only be used when the number of the antecendent is plural.
+must not be used as subject and further they may only be used when the number of the antecedent is plural.
 %
@@
 Swedish and Norwegian have two modal verbs for future tense
-\cite[p. 246]{H&H}.
+\cite[p. 246]{H&H}:
 %\cite[Tempus \textsection 28]{SAG}.
@@
 a. & Det \textbf{kommer att} bli mörkt snart & & b. &Jag \textbf{ska} gå och lägga mig nu\\
-  & ``It will get dark soon" & & b. &``I'm going to bed now"\\
+  & ``It will get dark soon" & &  &``I'm going to bed now"\\
 \end{tabular}}
@@
 ``I will miss you"}
-\emph{``ska"} signals an intention of committing the action, either from the subject or from
+The modal verb \emph{``ska"} signals an intention of committing the action, either from the subject or from
 the speaker. Cf.
@@
 a. & Du ska tycka om den & & b. &Du kommer tycka om den\\
-  & ``You \emph{shall} like it" & & b. &``You will like it."\\
+  & ``You \emph{shall} like it" & &  &``You will like it."\\
 \end{tabular}}
-\emph{``kommer"} (normally used with the infinite marker \emph{att}) 
+The verb \emph{``kommer"}, normally used with the infinite marker \emph{att},
 does not signal any intention, but that the speaker has belief that 
@@
 The resource grammar included 
-\emph{``ska"} which was implemented as the standard way of forming future
+\emph{``ska"}, which was implemented as the standard way of forming future
 tense, and hence represents the translation of
@@
 Since two out of the three Scandinavian languages share this
-tense, it has been added to the Scandiniavin \verb-Extra- module.
+tense, it has been added to the Scandinavian \verb-Extra- module.
 Table \ref{fig:kommeratt} shows how the grammar expresses new tense
@@
 \subsubsection{Focusing adverbs}
-The GF analysis distinguishs between two categories of adverbs: \verb-Adv- 
+The GF analysis distinguishes between two categories of adverbs: \verb-Adv- 
 %eg. `nu' (\emph{`now'}) and
 and \verb-AdV-. %like `alltid' (\emph{`always'}).
-The \verb-AdV- (eg. `alltid' (\emph{`always'}) and `inte' (\emph{`not'}))
+The \verb-AdV- (eg. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}))
 attaches directly to the verb.
@@
 Main clause : subj ++ verb.fin ++ verb.adV ++ verb.inf ++ verb.adv
-              han     har         alltid      sovit       här 
+              han     har         aldrig      varit       här 
 \end{verbatim}
@@
 %\end{wrapfigure}\\
-The adverb `bara' may be put int both of these positions but also before the finite verb,
-when emphazing the verb itself.
-This is an example of a focusing adverb, another example is `inte ens' (\emph{`not even'}).
+The adverb `bara' may be used as an \verb-AdV- but also before the finite verb,
+when emphasizing the verb itself.
+This is an example of a \textit{focusing adverb}, another example is `inte ens' (\emph{`not even'}).
 \enumsentence{\shortex{9}
@@
 Focusing adverbs are accepted by the new grammar implementation, where they have
-their own field in th \verb-VP-table.
+their own field in the \verb-VP-table.
 \begin{verbatim}
@@
 {Det & är & bara & barn &som& får & göra & så }
-{it & is & only & childern &that& may & do & so }
+{\emph{it} & \emph{is} & \emph{only} & \emph{children} &\emph{that}& \emph{may} & \emph{do} & \emph{so} }
 {`Only children are allowed to do that'}}
-Two more rules are consequentally needed
+Two more rules are consequently needed
 \begin{verbatim}
@@
 The resource grammar already gave a good coverage of relative clauses and embedded sentences.
-All construction used in examples \ref{sent:relex}a-d was accepted.
+All constructions used in examples \ref{sent:relex}a-d were accepted.
 %provided a number of ways to express relative and embedded clauses, such as
@@
 %parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
-As a paranthesis, some complications regarding the function \verb-RelCl- might
+As a side note, some complications regarding the function \verb-RelCl- can
 be pointed out.
 %Another complication stems from \verb|RelCl| which express the 
-The english implementation of this construction is \emph{`such that'},
-and the Swedish version \emph{`sådan att'} sounds awkward, when used outside of
-logic books.
-\enumsentence{Jag vill ha en katt sådan att den inte fäller.} 
-This relative clause \emph{``en katt sådan att den inte fäller"} is, in GF, put
-together by two constituents: \emph{``katt"}
-of type \verb-CN- and \emph{``den inte fäller} of type \verb-RCl- (relative clause), glued
+The English implementation of this construction is \emph{`such that'},
+and the Swedish version \emph{`sådan att'} sounds awkward, except when used in of
+logic and mathematics books.
+\enumsentence{\begin{tabular}{ll}
+\emph{From the resource grammar}& \emph{An alternative formulation} \\
+Jag vill ha en katt sådan att den inte fäller & Jag villa ha en sådan katt som inte fäller\\
+\end{tabular}} 
+In GF, this relative clause \emph{``en katt sådan att den inte fäller"} is constructed by
+two constituents: \emph{``katt"}
+of type \verb-CN- and \emph{``den inte fäller"} of type \verb-RCl- (relative clause), glued
 together by \emph{``sådan att"}.
 As a subject \emph{`den'} is of type \verb-NP- and does not depend on any
-agreement. However, it shoud agree with its antecedent, \emph{`katt'}. 
+agreement. However, it should agree with its antecedent, \emph{`katt'}. 
 \enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-The dependent types introduced in section \ref{sec:reflexives} could probaly be
+The dependent types introduced in section \ref{sec:reflexives} could probably be
 useful.
-However a carfully prepared analysis and motiv are needed. %ationamount in a bigger reconstruction which therefore
+However a carefully prepared analysis and motivation is needed. %ationamount in a bigger reconstruction which therefore
 %would require a. 
-Consequentelly, sentence \ref{sent:detfaller} is still
+Consequently, sentence \ref{sent:detfaller} is still
 accepted by the grammar.
@@
 \begin{itemize}
-\item Quantifiers does not have any possessive
+\item \verb-Quantifiers- does not have any possessive
 forms, so when a quantifier is turned into
-a NP (by the functions \verb-DetQuant- and \verb-DetNP-), its genitive field gets
-the same form as its nomitative field.
+a \verb-NP- (by the functions \verb-DetQuant- and \verb-DetNP-), its genitive field gets
+the same form as its nominative field.
 \item\verb-GenNP-
 introduced a cycle: a \verb-Quant- may be turned into a \verb-NP-, and now also
-a \verb-NP- can be turned into a Quant.
+a \verb-NP- can be turned into a \verb-Quant-.\\
+\verb|GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP|
 \end{itemize}
-\begin{verbatim}
-GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP
-\end{verbatim}
 We had that a noun phrase, eg \emph{`katten'} could be turned into a quantifier,
@@
 \enumsentence{Det är kattens mat}
-and consequentelly as a noun phrase:
+and thus as a noun phrase:
 \enumsentence{Det är kattens.}
@@
 The problem can be solved by changing the implementation of the function
-\verb-DetNP-, that is responsible for creating the possessive form
+\verb-DetNP-  which is responsible for creating the possessive form
 of quantifiers and determiners.
@@
 in \emph{`katten\textbf{ss}'}. Even though this generates a form that does not exist, it is perhaps
-the best approximation to what it would look like, given that it indeed was used in 
+the best approximation to what it would look like, given that it was indeed used in 
 Swedish. Most important is however that this stopped the \verb-GenNP-
-from introducing unnecessarily ambiguities.
-The number of parse trees for the phrase \emph{``Kattens bils hus"} was reducesed
-from over one hundered to six.
+from introducing unnecessary ambiguities.
+The number of parse trees for the phrase \emph{``Kattens bils hus"} was reduced
+from over one hundred to six.
 Besides, the grammar will allow the actually existing genitive
@@
 as pronouns. Many pronouns could not be imported from Saldo, since the correct category
-could not be identified automatically. Therefore, some experiments of how to extract this 
-information from Talbanken have conducted. We were interested in words which can be used
+could not be identified automatically. Therefore, some experiments were
+conducted on how to extract this 
+information from Talbanken. We were interested in words which can be used
 the same ways as `sådana' (\emph{`such'}) that acts either as a
@@
 
-By starting from a list containing all pronouns in Saldo and all forms they may occure in,
-Talbanken was harvested for the ones that both occured with the determiner tag \verb-DT-
+By starting from a list containing all pronouns in Saldo and all forms they may occur in,
+Talbanken was harvested for the ones that both occurred with the determiner tag \verb-DT-
 and with the present participle tag \verb-SP-. This way we are left with a limited number
 of words that could easily be analysed and added to the lexicon.
-Further, list of all pronous acting as determiners were extracted.
+%Further, list of all pronous acting as determiners were extracted.
 
 This method is
-both faster and more reliable than going through the list by hand and categories them
-by manually looking up each word or by using intraspection.
+both faster and more reliable than going through the list by hand and categorise them
+by manually looking up each word or by using introspection.
 
@@
 \section{Testing}
-Througout the project regression testing has been used. Every time the grammar was updated
+Throughout the project regression testing has been used. Every time the grammar was updated
 it was tested against a treebank consisting of 155 sentences, and the number of
@@
 make sure that the coverage was not decreased, but also to make sure that we
-would notice if new ambiguites or overgenerating functions were introduced.
+would notice if new ambiguities or overgenerating functions were introduced.
 If so, the ambiguities could be removed when possible, and when not, we were at least made
@@
 Talbanken contains much valuable information about phrase structures and word 
-usage. It is analysed with the MAMBA annotation scheme (Telemann, 1974) which is 
+usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974) which is 
 %>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
@@
 automatically transforms trees in Talbanken format to GF abstract syntax trees.
-We hence get a comparision between the two annotaitons and at the same time
+We hence get a comparison between the two annotations and at the same time
 we get methods to extract and translate information to GF notation. 
@@
 is preserved.
-\emph{``På bilen"} should still form a subtree attatched to \emph{``katten"}.
+\emph{``På bilen"} should still form a subtree attached to \emph{``katten"}.
 %produces a \verb-NP- from ``katten" this means that
@@
 Talbanken. Adaption was required for the differences in annotation as well as 
-syntactic differences between Swedish and English.
+for the syntactic differences between Swedish and English.
 
@@
 Talbanken05 uses three kinds of tags: categories, edge labels and pos-tags. 
-While the pos-tags are reserved for words, the categories give syntactic information
-for phrases, such as \verb|S|, \verb|NP| and \verb|VP|.
+%TODO part of speech??
+While the pos-tags are reserved for words, the categories give part-of-speech information,
+such as \verb|S|, \verb|NP| and \verb|VP|.
 The edge labels give the grammatical functions: \verb|SS| for subject, 
 \verb|OO| for object etc. Nearly 400 different tags are used, of which more than
-300 different are pos-tags. The high number is due to the detailedness, for
-example there are almost 50 tags for nouns, excluding proper names. The tags
+300 different are pos-tags. The high number is due to the detailedness,  there are for
+example almost 50 tags for nouns, excluding proper names. The tags
 show definiteness, case and whether the word is a compound. Some words also have
@@
 \begin{verbatim}
-map S  = np <- map NP
-         vp <- map VP
+translate S  = do
+             np <- find NP
+             vp <- find VP
              return (PredVP np vp)
@@
 
-At every leaf a word is given, %When a leaf is reached,and ato see if
-and by using the knowledge from the pos-tag, the word is looked up in the GF
-lexicon. If the word is not found
-a \textit{meta} variable -- \textit{?} -- will be put in its place.
+At every leaf a word is given. %When a leaf is reached,and ato see if
+It is looked-up in the GF lexicon and the result compared to the information of
+the pos-tag.
+If the word is not found in the lexicon
+a \textit{meta} variable will be put in its place, denoted `\verb-?-'.
 The meta variable is also used when we have found two translatable trees
@@
 
-The translation covers complex noun phrases which may consists of pronouns,
-reflexive objects or common nouns and thereby contain predeterminers, 
+The translation covers complex noun phrases which may consist of pronouns,
+reflexive objects or common nouns and contain predeterminers, 
 determiners, quantifiers and adjectival modifiers. Special rules were needed
-for each category of verb in order to find the right number of arguments.
+for each verb category in order to find the right number of arguments.
 The mapping also covers changed word order, such as questions and topicalisation.
@@
 In Talbanken the valency information is given implicitly by the complements
-of a word. If a verb is followed by an object, \verb-OO-, which contains a
-\verb-S-, we can conclude that this is a sentence-complemnt verb.
+of a word. If a verb is followed by an object, \verb-OO-, which contains an
+\verb-S-, we can conclude that this is a sentence-complement verb.
 In GF, the valency information is given for each entry in the lexicon.
@@
 sentence.
-A similar issue arises form the use of verb particles. A verb with a
+A similar issue arises from the use of verb particles. A verb with a
 particle is treated as two independent units in Talbanken but in GF they are
@@
 \end{figure}
-When translating a senctence like this, it is thus not possible to simply glue the
-subject to the verb with \verb|PredVP|. For each similair GF construction, an extra ruled
+When translating a sentence like this, it is thus not possible to simply glue the
+subject to the verb with \verb|PredVP|. For each similar GF construction, an extra rule
 would be needed to get full coverage. \\ %and the normal top-down? approach fails.\\
@@
 Regression testing was used to ensure that a new rule wouldn't violate the old ones
-or overgenerate to allow trees to be translated in an erronous way.
+or allow trees to be translated in erroneously.
 
@@
 Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
-which can group an infinite verb and an object adverbial together into a \verb|VP|.
+which groups an infinite verb and its object together into a \verb|VP|.
 This information can however be extracted from the flat version, and the results
@@
 The \textbf{PU} tag is used for graphic listings, and not for fluent text.
-In our grammar there is natuarally no corresponding function, 
+In our grammar there is naturally no corresponding function; 
 %We are not interested in sentences made up by list, since 
@@
 folders etc and out of scope for the grammar itself. The tags \textbf{XX} and \textbf{NP} are often used since
-Talbanken (unlike GF) makes a difference between subjects
+Talbanken -- unlike GF -- makes a difference between subjects
 and objects. %(which is not needed in GF (see section \ref{sec:swegf}))
-The analyse of senctences like
+The analyse of sentence \ref{sent:krav}
 \enumsentence{För stora krav.\\
-``Too high demands."}
+``Too high demands."}\label{sent:krav}
 contains the tags \textbf{XX} and \textbf{NP}, since it is not obvious
@@
 involved, we
-can restore more than 85\% of the nodes in the original tree.
+can restore more than 85 \% of the nodes in the original tree.
 If we lift all the restrictions excluding the \verb|PU|, we get
 %When tested the 100 first sentences from Talbanken, we get
-65\% coverage. 
+65 \% coverage. 
 If we test randomly collected sentences that does not contain any of the tags listed
-in figure \ref{fig:mapBadtag} 72\% can be restored. \\
-
+in figure \ref{fig:mapBadtag}, 72 \% can be restored. \\
+
+%TODO
+%All NPs: 72%
 \begin{tabular}{|ll|}
 \hline
-No list items & 65\%\\
-No special punctuation or bad tags& 72\%\\
-Short sentences with known words & 85\%\\
+No list items & 65 \%\\
+No special punctuation or bad tags& 72 \%\\
+Short sentences with known words & 85 \%\\
 \hline
@@
 These results are lower than the results of the mapping of Penn Treebank. The
-reason for this may lay in the fact that English is syntactically less complicated
+reason for this may be the fact that English is syntactically less complicated
 than Swedish. Also, the Penn Treebank contains a lower number of tags, 82 compared
 to more than 300 in Talbanken. Even if the tags describing the same word
-class as another tag are excluded, Talbanken still leaves us with more than 130 tags. The 
-number of ways to combine them are hence much increased, and so is the number of
+class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
+Thus the number combinations are extensively increased, and so is the number of
 rules that would be needed to cover all constructions.
@@
 %we get a slightly better result,
-%from 62.5\% to 65.4\%.  (see EvalText). The complemnt of the verbs should tell which category
+%from 62.5 \% to 65.4 \%.  (see EvalText). The complemnt of the verbs should tell which category
 %they are anyway.
@@
 \section{Evaluation}
->>evolution diagram? also removed unnecessary ambigutiies.
->> how to handle ambigs: know them and controll them. using statistics or ontolgies.
->> that is: disambiguitation by context or by meaning.
+%TODO
+%>>evolution diagram?
+%also removed unnecessary ambigutiies.
+%>> how to handle ambigs: know them and controll them. using statistics or ontolgies.
+%>> that is: disambiguitation by context or by meaning.
 The project has resulted in
@@
 \item a large-scale GF lexicon and a program to redo the importation when needed
-\item a comparison between GF and other annotations
+\item a comparison between GF and another another annotation
 \item a deeper testing of the Swedish resource grammar and an estimation
@@
 or missing phrase will cause the parsing of the whole sentence to fail.
-Our goal is to parse Talbanken, but considering that no syntactic theory is capable
-of covering a whole natural language, our grammar when used without a
-robust layer should just be seen as an important
-fragment.\\
-
-A comparison with Talbanken would not be of much value of this state since the coverage 
-is yet too small to do be able to get any interesting analysis.
+
+Parsing bigger parts of Talbanken would hence give very low results at this state, 
+%, for the reasons stated.
+and a comparison of the results would not be of much value,
+there would not be enough material %underlag is yet too small 
+to do be able to do any interesting analysis.
 An estimation of the improvement can be given by looking at the results from
-running the testsuite used for the grammar development.
-The sentences given in the testsuite are short and the words are in most cases known
+running the test suite used for the grammar development.
+The sentences given in the test suite are short and the words are in most cases known
 to the lexicon, but there are also constructions that have not been implemented 
@@
 The first grammar could parse almost half of the sentences, the result for the final
-grammar was 65\%. 
-Parsing bigger parts of Talbanken would give very low results, for the reasons stated.
+grammar was 66 \%. 
 % try NPs?
 It is thus not yet interesting to talk about coverage, but about the quality and
-the ability to scale up, which has so far proved to be good. The constructions
-that we have focused on have all been possible to implement, although the amount
-of work to get it ~~right has varied.
+the ability to scale up, which has so far proved to be good.
 I further believe that the presence of a Swedish expert has
@@
 big to use trivially with the current techniques. Its size causes the
-incremental parsing algorithm use more heap memory than what normal computers
+incremental parsing algorithm use more heap memory than normal computers
 allow. To solve this, we need to use the lexicon data more cleverly.
@@
 \label{sec:future}
-Although this one part of the project is finished, there are many interesting future
-directions. As a natural language processing-project it will probably never be
-entirely complete.
+Having finished this part of the project, we are left with 
+%Although this one part of the project is finished, there are 
+many interesting future directions.
 The future work described in this section is divided into two categories:
@@
 \subsubsection{Grammar}
-All imaginable grammatical constructs cannot be cover by our grammar.
-Some work must be left for the robustness, but there are specific constructions
+The grammar should cover the most prominent Swedish features.
+%All imaginable grammatical constructs cannot be cover by our grammar.
+Some work must be left for the robustness, but there are specific constructs
 that are desirable and interesting to implement. A few examples are listed here:\\
@@
 \item
-\textit{Pronomial object shifts} are common in Swedish.
+\textit{Pronominal object shifts} are common in Swedish.
 \enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
@@
 \end{tabular}}
-Personal pronouns are allowed to shift position with the negation in main clauses, without
+Personal pronouns are allowed to shift position with the negation in main clauses without
 the modal verbs.
@@
 Talbanken's part P, which have been the inspiration for this project.
-Therefore, it has so far not been prioritised.\\
+Therefore, its implementation has so far not been prioritized.\\
 
@@
 \item
-\textit{Bare indefinites} are at this stage always parsed as mass nouns.
+\textit{Bare indefinites} are at this point always parsed as mass nouns.
 \enumsentence{\shortexnt{3}
 {Jag & dricker& vatten}
-{\emph{I}&\emph{drink}&\emph{water} }
-}
-This however is not the correct analyse for `hund' in sentence \ref{sent:hund}.
+{\emph{I}&\emph{drink}&\emph{water}}
+}\label{sent:water}
+The parse tree of sentence \ref{sent:water} looks as follows:
+\begin{verbatim}
+PredVP (UsePron i_Pron) (ComplSlash (SlashV2a drink_V2) 
+                                      (MassNP (UseN water_N))) 
+\end{verbatim}
+
+This however is not the correct analysis for `hund' in sentence \ref{sent:hund}.
 \enumsentence{\shortex{4}
@@
 The implementation of reflexive pronouns can be improved.  %both code-wise
-It should for example be possible to differenciate between 
+It should for example be possible to differentiate between 
 object and subject control verbs.
@@
 Our lexicon still lacks some part that can be imported from Saldo.
-The multipleword verbs (\verb-vbm-) and multiple word nouns (\verb-nm-) should
-be imported to an idiom lexicon, which can be used as extensions to the main
+The multiple-word verbs (\verb-vbm-) and multiple-word nouns (\verb-nm-) should
+be imported to an idiom lexicon, which can be used as an extension to the main
 lexicon.
-For the words that we tried but failed to import, a tool for lexicon
+For the words that we tried but failed to import, another tool for lexicon
 acquisition could be used. The tool developed in the previous part of
@@
 
-\subsection{Making the grammar robust}
+\subsection{Making the parser robust}
 To achieve our goal of free parsing, we need to use statistical methods.
@@
 compound words and unknown constructions.
+
 \subsubsection{Parallel parsing/Chunk parsing}
 Instead of parsing a whole sentence in one go, we would like to
-parallely parse parts of it and then try to combine the pieces.
+parse parts of it in parallel and then try to combine the pieces.
 This method should be faster then normal parsing and would
-at the same time give robustness.
-Components that cannot be combined by any grammar rule are combined by a 
-meta function. We might end up with partial parse trees, but we would
-be able to parse at least parts of every sentence and give a more
-interesting result to the user.
+at the same time give robustness;
+components that cannot be combined by any grammar rule are combined by a 
+meta function. When an unknown construction is encountered, the parser
+will not fail but give a partial parse tree with the analysis for the rest of
+the sentence.
+The output given to the user will be more valuable, since partial parse
+trees are more interesting the no parse trees.
+%We might end up with partial parse trees, but we would
+%be able to parse at least parts of every sentence and give a more
+%interesting result to the user.
 
@@
 never cover all compounds. We therefore need to be able to do
-compounding analyses.
+compounding analyses. Saldo has tools for this, which might be usable
+in this project as well \cite{}.
 
@@
 One possibility is of course to do 
-deeper refinements of the parsing source code, which would be a substantial 
-project. Other solutions are to to either
+deeper refinements of the parsing source code, an extensive work which is far out of scope
+for this project. Other solutions are to to either
 adapt the automatically lexicon for its domain and hence making it smaller and
-faster, or to adapt the input sentences to a smaller lexicon, by preprocessing.
+faster, or to adapt the input sentences to a smaller lexicon by preprocessing.
 
 Further, the lexicon needs valencies since this information lays the
-foundation for the GF analysis. We have already analysed how Talbanken
-annotates valency information, and by using the rules made for the mapping
-described in section \ref{sec:Mapping}, we could extract data from Talbanken,
+foundation for the GF analysis. 
+During the mapping, described in section \ref{sec:Mapping}, we analysed how Talbanken
+annotates valency information. It should be possible to %extract data from %and by using the rules made for the mapping
+extract data from Talbanken,
 showing how words are normally used. There should also be algorithms for
 automatically adding the information to the lexicon. \\
-Many Swedish verbs can be used with different number of arguments. Adding
-all this information to the lexicon does not seem to be a good idea considering
+Many Swedish verbs can be used with different number of arguments.
+Having one lexical entry for every possible usage of a verb
+%all this information to the lexicon 
+does not seem to be a good idea considering
 the ambiguities it will lead to and the already high time usage.
@@
 implemented by using external resources.
+% TODO like what?
 
@@
 \label{sec:futureProbabilities}
-The grammar will always contain ambiguities but before returing the result
+The grammar will always contain ambiguities and before returning the result
 to the user, some analysis should be done to find the most probable tree.
@@
 increases, so will the ambiguities as the number of word forms overlapping each other
-gets higher. Our new grammar also have many more rules which also will contribute
-to increase the number to interpret each sentence.
-
-GF already allows the lexicon to have probabilities, and by using our extracted
-GF treebank, we can get more reliable statisics for our functions. 
-By implementing dependency probabilities, our results would be even better.
+gets higher. Our new grammar also have many more rules which contribute
+to increase the number of to interpretations of each sentence.
+
+GF already allows probabilities, and by using an large
+GF treebank, we can get more reliable statistics for our functions. 
+By implementing dependency probabilities \cite{}, our results would be even better.
+
+\subsubsection{Named entity recognition}
+The parser should be able to identify names of persons, companies, countries etc.
+This can be done by the Haskell program executing the parsing.
 
@@
 \section{Conclusion}
->> why is this significant, other grammars can use this Swedish grammar.
->> well known (comp. ling) trade-off between depth and coverage. for now depth,
->> robustness in in future work leads to coverage.
->> reusablilty, freely available.
->> have experiments of how to extract info from talbanken.\\
+%>> why is this significant, other grammars can use this Swedish grammar.
+%>> robustness in in future work leads to coverage.
+%>> reusablilty, freely available.
+%TODO
+%>> well known (comp. ling) trade-off between depth and coverage. for now depth,
+%>> have experiments of how to extract info from talbanken.\\
+
 We have developed the main components
-for a deep Swedish parser. %To abstract:  by combining resources.
-Our main resource is the grammar formalism GF,
-which allows us to start from a well-defined system for describing grammar.
-The parser is derived from the grammar and will therefore validate that
-a given sentence is grammatically correct according to the rules defined.
-All parts, including the underlying grammar, are open-source and 
-may thus be used in other applications.
-
-Even though the grammar will not be totally covering, we belive that
-the work done is both interesting and useful.
+for a deep Swedish parser; an extended grammar and lexicon and material for
+evaluation and disambiguation. %To abstract:  by combining resources.
+We started from a well-defined system for describing grammar as we built
+the grammar from the GF resource grammar.
+
+%When parser is derived from the grammar and will therefore validate that
+%a given sentence is grammatically correct according to the rules defined.
+
+Our final goal is to parse Talbanken, but considering that no syntactic theory is yet capable
+of covering a whole natural language, we are satisfied with %theconsider our grammar 
+%when used without a robust layer 
+the grammar implementation of an important fragment of Swedish.
+Following methods for making the parser robust as described in section \ref{sec:future}
+we still believe that it is possible to achieve the goal. 
+\\
+
+Being a natural language processing-project, it will probably never be
+entirely complete. Even so, we believe  
+%Even though the grammar will not be totally covering, we belive
+that the work done is both interesting and useful.
+All parts of the project are open-source and may thus be used in other applications.
+The grammar may be beneficial also when working with controlled languages,
+as it increases the coverage of the Swedish resource grammar.
+
+The constructions
+that we have focused on have all been possible to implement, with varying  amounts
+of work. Many of them could be done by utilizing and extending the resource library
+but in some cases we needed to part from the multilingual abstract and use other
+grammatical theories in order to be able to perform a good analysis.
 

### 2011-12-23 doc/report/report.tex
start on lexicon generation

@@
 The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
-gives a full syntactic analysis as well as semantics represented in xx. A
+gives a full syntactic analysis as well as semantics represented in `quasi logical form'. A
 translation to English  was implemented and the work was further developed in the spoken

### 2011-12-23 testSuites/MkLex.hs

@@ -1,0 +1,62 @@
+module MkLex where
+import PGF
+import Control.Monad
+import System.Process
+
+
+mkDir :: [String] -> Language -> FilePath -> IO (FilePath,Language)
+mkDir ws lang bigPGF = do
+   pgf <- readPGF bigPGF  
+   let morpho = buildMorpho pgf lang
+       allLemmas = map fst $ concatMap (lookupMorpho morpho) ws
+   (cnc,abs) <- extractLemmas allLemmas
+   rawSystem "gf --make" [cnc] -- must wait for process!
+   return (newPGF,newLang)
+
+extractLemmas :: [Lemma] -> IO (FilePath,FilePath)
+extractLemmas ws = do
+  writeHeaders
+  ex bigLexCnc newLexCnc
+  ex bigLexAbs newLexAbs
+  writeEnd
+  return (newLexCnc,newLexAbs)
+
+ where ex file newFile = do
+             lex <- liftM ((map words) . lines) $ readFile file
+             mapM_ (extract newFile) lex
+       extract newFile l@(x:xs)
+               | x `specelem` ws = appendFile newFile (unwords l++"\n")
+       extract _ _ = return ()
+
+       writeHeaders  = writeGF writeFile [headerCnc,headerAbs]
+       writeEnd      = writeGF appendFile (repeat "}\n")
+
+       writeGF f = zipWithM_ f [newLexCnc,newLexAbs]
+
+       headerCnc = "--# -path=.:abstract:alltenses:swedish:common:scandinavian\n"
+                   ++"concrete TestLex of TestLexAbs = CatSwe **\n"
+                   ++"open  Prelude, CommonScand, ParadigmsSwe, IrregSwe in {\n"
+                   ++"lin\n"
+       headerAbs = "--# -path=.:abstract:prelude:alltenses\n"
+                   ++"abstract TestLexAbs = Cat ** {\n"
+                   --smarter algorithm! have lexicon is lexical order
+
+bigLexCnc = "DictSwe.gf"
+bigLexAbs = "DictSweAbs.gf"
+newLexCnc = "TestLex.gf"
+newLexAbs = "TestLexAbs.gf"
+newPGF    = "TestLex.pgf"
+newLang   = read "TestLex"
+       
+specelem :: String -> [Lemma] -> Bool
+specelem w = elem w . map showCId
+
+-- use this, sort the lexicon (once), the input (should be fast)
+keep :: [[String]] -> [String] -> FilePath -> IO ()
[…]

### 2011-12-23 testSuites/Parsing.hs

@@ -12,6 +12,7 @@ import Control.Applicative
 import Control.Monad.State
 import PGF
 import XMLHelp hiding (words,parse)
+import MkLex
 
 -- use with tee -a outfile
 data Parsed = Pars {t :: Tree, s :: String, n :: Int}
@@ -26,36 +27,39 @@ timeLimit =  30*10^6
 
 run :: String -> Bool -> FilePath -> IO Count
 run file strict out = do  
-  putStr "Parsing input file..."
-  sents <- mainF file (return . map getSentence)
-  putStr " Ok\nReading PGF..."
-  pgf <- readPGF pgfFile
+  putStrLn "Parsing input file..."
+  sents <- liftM head $ mainF file (return . map getSentence)
+  putStrLn " Creating new dictionary..."
+  (pgf',lang') <- mkDir (words $ concatMap (map toLower . snd) sents) lang pgfFile 
+  putStrLn " Ok\nReading PGF..."
+  pgf <- readPGF pgf' --File
   putStrLn " Ok\n"
-  --absMap <- readAbbs 
-  let morpho = buildMorpho pgf lang
-  c <- execStateT (mapM_ (tryparse pgf morpho) (concat sents)) (newCount strict out)
+
+-- must compile a new grammar!! BigSwe with tempDict
+  let morpho = buildMorpho pgf lang'
+  c <- execStateT (mapM_ (tryparse pgf morpho lang) sents) (newCount strict out)
   print c
   return c
 
-tryparse :: PGF -> Morpho -> Sent -> CState ()
-tryparse pgf morpho sents = do 
+tryparse :: PGF -> Morpho -> Language -> Sent -> CState ()
+tryparse pgf morpho lang sents = do 
   count 
-  parse' pgf morpho sents
+  parse' pgf morpho lang sents
 
-parse' :: PGF -> Morpho -> Sent -> CState ()
-parse' pgf morpho (s,"") = addEmpty >> putEmptyMsg (s,"")
-parse' pgf morpho s = do
+parse' :: PGF -> Morpho -> Language -> Sent -> CState ()
+parse' pgf morpho lang (s,"") = addEmpty >> putEmptyMsg (s,"")
+parse' pgf morpho lang s = do
   let fix     = fixPunctuation $ replaceNumbers s 
       unknown = badWords morpho (snd fix)
   strict <- gets strict
   out    <- gets outputFile
-  if null unknown then parseOk pgf fix
[…]

### 2011-12-28 doc/report/report.tex
report, modification to passive
report: abstract, acknowledgments

@@
 %ska innehålla vettiga sökord
-\title{Towards a wide-coverage grammar for Swedish} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Towards a wide-coverage grammar for Swedish using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 \abstract{
+This thesis describes the work to achieve a covering Swedish grammar implemented
+in GF -- to model natural language by functional programming. 
+The idea has been to combine existing language technology resources with
+new techniques, and the aim to obtain a system for analysing written text.
+To reach this goal, problems of computational as well as linguistic nature had to be solved.
+\\The thesis explains the methods and technologies used and further inspects the possibilities
+of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
+combining our rule-based grammar with statistical methods.
+\\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
+to GF notation and we also give an implementation Swedish grammar. The grammar is based
+on the multilingual abstract syntax given in GF resource library, and now
+extends it to cover constructions specific to Swedish.
+We further give an example of the advantage of using dependent types when
+describing grammar and syntax, in this case reflexive pronouns.
+
 %This is the report of the work of expanding and enhancing a GF grammar and preparing
@@
 \newpage
-
+\pagestyle{empty}
 \section*{Acknowledgments}
+\vspace{5mm}
+Many people have helped me during this work and made this thesis possible.
+I would first of all like to thank
+Center of Language Technology, that has funded the project.\\
+
+Further, thanks to my excellent 
+%(~~devoted~~) 
+supervisor Ramona Enache for all her help and guidence 
+%(encouragement)
+in every phase and all aspects of the work.
+Thanks to Elisabet Engdahl for sharing her seemingly unlimited knowledge
+of Swedish grammar. She has also has acted as a second supervisor, and given me
+very helpful comments and suggestions. Thanks to Aarne Ranta for all his great
+ideas and for letting me do this project.
+
+I am also grateful to Krasimir Angelov, Markus Forsberg, Peter Ljunglöf, Lars Borin
+and many others who have contributed with ideas and inspiration, and for their
+interest in this work.\\
+
+Finally, I would also like to thank my friends and family. Special thanks to Dan Rosén
+for all his support, advice and patience and, most importantly, for being such a good friend. %for having listened to 
+%my monologues about the daily progress. 
 % Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc
@@
 \newpage
-
+%\pagestyle{empty}
+%\cleardoublepage
 \tableofcontents
+%
+\addtocontents{toc}{\protect\thispagestyle{empty}}
+%\pagestyle{empty}
+%\cleardoublepage
 \newpage
+\setcounter{page}{1} %reset the page counter
 \chapter{Introduction}
@@
 \begin{verbatim}
-PassVPSlash : VPSlash -> VP ;
+PassVP : VPSlash -> VP ;
          ta      -> togs

### 2011-12-28 doc/report/report.tex
report: elisabets corrections

@@
 \\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
-to GF notation and we also give an implementation Swedish grammar. The grammar is based
+to GF notation and a Swedish grammar implementation. The grammar is based
 on the multilingual abstract syntax given in GF resource library, and now
@@
 I would first of all like to thank
-Center of Language Technology, that has funded the project.\\
+Center of Language Technology, that has funded the project.
 
@@
 and many others who have contributed with ideas and inspiration, and for their
-interest in this work.\\
-
-Finally, I would also like to thank my friends and family. Special thanks to Dan Rosén
+interest in this work.
+
+Finally, I would also like to thank my friends and family. Special thanks to Dan % Rosén
 for all his support, advice and patience and, most importantly, for being such a good friend. %for having listened to 
@@
 % --> sentence about connection GF - grammar based parser.
-This project focuses on implementing a wide-coverage Swedish grammar, which also
+This project focuses on implementing a wide-coverage Swedish grammar, which
 describes a parser accepting all sentences that can be expressed by the grammar.
@@
 goal is
-to create a open source Swedish grammar. % from which we derive a parser
+to create an open source Swedish grammar. % from which we derive a parser
 %accepting all sentences described by the given rules.
@@
 To build the grammar from scratch would be not only time consuming but
-would also mean that already existing resources would reimplemented.
+would also mean that already existing resources would have to be reimplemented.
 In order to not reinvent the wheel we proceed from a combination of well-tested
 resources.
-We start from GF resource grammar and thereby get
+We start from a GF resource grammar and thereby get
 %The usage of GF allows us to start from
-a well-defined system for describing grammar, as well as a strong connection to
+a well-defined system for describing the grammar, as well as a strong connection to
 more than 20 other languages implemented in the framework. Further, we use
@@
 %>> enhance gf...
-The purpose of this project has been to prepare the GF grammar for free parsing.
+The purpose of this project has been to prepare the GF grammar for parsing of 
+unrestricted Swedish.
 This has meant to develop earlier techniques to fit for Swedish, create methods
 for achieving and keeping a large-scale lexicon and to adapt the existing
-resource grammar to fit for more language specific constrictions.
+resource grammar to fit more language specific constrictions.
 Three sub-goals has been identified:
 \begin{itemize}
-\item Enhancing the Swedish GF grammar
+\item Extending the Swedish GF grammar
 \item Importing the lexicon Saldo
-\item Create a translation between Talbanken and GF
+\item Creating translation between Talbanken and GF
 \end{itemize}
@@
 description of the Swedish resource grammar is given in section \ref{sec:swegf}.
-Section \ref{sec:swedish} gives an introduction Swedish, and
+Section \ref{sec:swedish} gives an introduction to the Swedish language, and
 brief presentations of Saldo and Talbanken are found in section \ref{sec:saldo} and
@@
 Finally, the conclusion and evaluation are
-presented in chapter \ref{sec:results} together with a list of future work .
+presented in chapter \ref{sec:results} together with some areas of future work.
 
@@
 \label{sec:background}
-The work described in this is thesis is part of a bigger project which aims at
-using GF for parsing free Swedish. 
-In previous work\footnote{??}, a start was made to extending the Swedish GF grammar
+The work described in this thesis is part of a bigger project which aims at
+using GF for parsing unrestricted Swedish. 
+In previous work\footnote{??}, a start was made to extend the Swedish GF grammar
 and a tool for lexical acquisition was developed.
@@
 by Agda \cite{agda}, with which GF also shares its type checking algorithm.
-The first version of the framework was implemented by Xerox Research Center
+The first version of the framework was implemented at Xerox Research Center
 in Grenoble and is now mainly developed in Gothenburg. One of the biggest
@@
 An interactive GF editor and the GF shell can be tried out online 
-\footnote{http://www.grammaticalframework.org/demos/gfse/}.
+\footnote{http://www.grammaticalframework.org/demos/gfse/}. See figure \ref{fig:shell},
+ The first command will parse the given
+sentence and visualize the resulting parse tree. The second will visualize the 
+abstract syntax tree. The trees will be showed in the pgf-format by the program evince.
 \begin{figure}[h]
@@
 \end{verbatim}
-\caption{Example of how the GF shell can be used. The first command will parse the given
-sentence and visualize the resulting parse tree. The second will visualize the 
-abstract syntax tree. The trees will be showed in the pgf-format by the program evince}
-\label{fig:shellvp}
+\label{fig:shell}
+\caption{Example of how the GF shell can be used.}\label{fig:shellvp}
 \end{figure}
@@
 \subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
-\caption{Parse tree and abstract tree for the sentence ``Jag ser katten".}
+\caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
 \label{fig:gftree1}
@@
 %the \textit{abstract tree}.
+The \textit{abstract tree} shows which functions GF uses
+to parse a sentence.
 The \textit{parse tree} (\ref{pic:gfStree}) shows which types
 that are assigned to the words and phrases. 
-The \textit{abstract tree} shows which functions GF uses
-to parse a sentence.
 % do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
@@
 the noun is explicitly quantified by the article \emph{``the"},
-and the determiner is therefore shows in the parse tree.\\
+and the determiner is therefore shown in the parse tree.\\
 
@@
 \newpage %as long as english picture bråks.
-The scope of GF grammars is usually controlled language. By
+The scope of GF grammars is usually controlled language, a limited language used in
+a restricted domain. By
 restricting the coverage, the number of ambiguities
@@
 WebAlt \cite{webalt}, a project which aims to develop language independent
-material for mathematics problems, and the formal software verification tool
+material for mathematics problems. The formal software verification tool
 KeY \cite{key}, in which GF is used for translating specification to natural language.
@@
 %>> use talk here and move key to other place??
-The project talk \cite{talk} used GF to describe dialog grammars and
+GF can be used for describe dialog grammars see \cite{talk}, and
 the framework is further used in the
@@
 
-When implementing a English version of the grammar, we encounter another
+When implementing an English version of the grammar, we encounter another
 problem: the verb form depends on the number of the noun. We solve this by letting
 \verb|N'| carry
-information about which its number and letting \verb|Pred| pass this on to the
-verb. Finally, the type of \verb|V| is put to a table, showing the verbs forms for each
+information about its number and letting \verb|Pred| pass this on to the
+verb. Finally, the type of \verb|V| is put into a table, showing the verbs forms for each
 number.
@@
 
-We now have two implementations of the abstract. The resulting GF grammar is able both to
-parse a string to a abstract tree and to go in the other direction; to produce
+We now have two implementations of the abstract, one for Swedish and one for English.
+The resulting GF grammar is able both to
+parse a string to an abstract tree and to go in the other direction; to produce
 a string of natural language given an abstract tree. This step is called linearization.
@@
 %>>Most general resource, first large grammar, covers fundamental syntactics.
-The GF package provides an advantageous resource library \cite{gf-resource}, covering the
+The GF package provides an useful resource library \cite{gf-resource}, covering the
 fundamental morphology and syntax of more than 20 languages.
@@
 %this can be found as libraries. So far there are 20 languages in the library, and
-The languages all share the same abstract and are valuable
+The languages all share the same abstract syntax which is  valuable
 %they constitute  a very useful resource 
-when implementing application grammars.
+when implementing multilingual application grammars.
 
@@
 \end{wrapfigure}
-This is the case for ``flicka", the correct declination
-can be guessed by using the first declination. % and noting that the word ends with an ``a". 
+This is the case for ``flicka", a noun belonging to the first declination. %the correct declination
+%can be guessed by using the first declination. % and noting that the word ends with an ``a". 
 %Like most Swedish nouns ending with an ``a", it follows the first declination.
@@
 and the possibilities to use GF in different environments
-increased. There are research on how to make more use of the dependent
+increased. There is research on how to make more use of the dependent
 types, for reasoning by using ontologies \cite{ramona} or generating natural
@@
 compare them against a reliable source.
-Being a freely available, manually annotated, large-scale treebank,
-Talbanken \cite{talbanken} was perfect for our purpose.
+Talbanken \cite{talbanken} was perfect for our purpose,
+being a freely available, manually annotated, large-scale treebank.
 %It was created in the 1970s at Lund University.
-In total it consists of four
+It is analyzed with the MAMBA annotation scheme (Teleman, 1974) and 
+consists of four
 parts. Two of them are transcriptions of spoken language, one a collection of
 text written by high school students, and one, section P,
-consists of professionally written Swedish gathered from newspapers, brochures and textbooks.\\
-Talbanken was also used to train the Swedish version of the Maltparser \cite{malt}
-and was then redistributed in a more modern version, Talbanken05 \cite{talbanken05}.
+consists of professionally written Swedish gathered from newspapers, brochures and textbooks.
+\\
+Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
+and was then redistributed in an updated version,
+Talbanken05 \cite{talbanken05}.
 %This work was done by Nivre, Hall and Nilsson in 2005.
@@
 the lexical MAMBA layer. \\
-In this project, as well as for the Malt parser, the more than 6000 sentences
-in section P has been used.
+The Malt parser was trained on section P of Talbanken, and these
+more than 6000 sentences have been used our project. 
 The treebank has served as an inspiration and an evaluation source throughout the
@@
 large electronic lexicon developed and maintained at Gothenburg University. It is
-built on Svenskt Associations Lexicon and contains information about more than 
+built on Svenskt Associationslexikon and contains information about more than 
 120 000 modern Swedish words.
 For each word there is semantic, syntactical and a morphological
-information. The user can find compounding analyses, examples of usage in corpora,
-and graphs of semantically connected words.\\
+information. The user can find examples of usage in corpora,
+graphs of semantically connected words and some suggestions for how to analyse compounds. \\
 \begin{figure}[h!]
@@
 
-The semantic aspect of Saldo require that words with multiple meanings are
+The semantic aspect of Saldo requires that words with multiple meanings are
 separated into different entries.
@@
 
-For our purpose, only this morphological section, provided under LPGL in XML
+For our purpose, only the morphological section, provided under LPGL in XML
 format, was needed.
@@
 \begin{tabular}{l|lll}
-&\textbf{Definite}& \textbf{Indefinite}& \textbf{Gender} \\
+                  &\textbf{Indefinite} & \textbf{Definite}& \textbf{Gender} \\
  \hline
@@
 
-The definiteness can be accentuated by using the definite articles.\\
-\begin{tabular}{llll}
-den katten & det huset & de katterna & de husen \\
-\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
-\end{tabular}\\
+%The definiteness can be accentuated by using the definite articles.\\
+%\begin{tabular}{llll}
+%den katten & det huset & de katterna & de husen \\
+%\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
+%\end{tabular}\\
 
 A definite article or a determiner is necessary when the noun is modified by adjectival
-phrase; sentence \ref{sent:trottkatt}a is not grammatically correct.
+phrase; sentence (\ref{sent:trottkatt}a) is not grammatically correct.
 As adjectives also marks definiteness, this is marked on three places
-in sentence \ref{sent:trottkatt}b. %, the definiteness is 
+in sentence (\ref{sent:trottkatt}b). %, the definiteness is 
 %The definiteness in then marked by the article, on the noun and on the adjective.
@@
 \enumsentence{\longexnt{3}{3}
-{Min &trötta& katt.} %&& Den där & trötta &katten}
-{`My & tired & cat'}
+{min &trötta& katt} %&& Den där & trötta &katten}
+{`my & tired & cat'}
 { $[${\sc+def}$]$ &$[${\sc+def}$]$& $[${\sc-def}$]$} % . && $[${\sc+def}$]$  & $[${\sc+def}$]$ &$[${\sc+def}$]$}
@@
 fronted \cite[\textsection 1027]{H&H}.
-This is called topicalisation and is very common, especially for temporal and spatial adverbial phrases.
+This is called topicalisation and is very common, especially for temporal and
+locative adverbial phrases.
 The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
 meaning, but vary in how the content is presented.
-%~~ although the fronted part of sentence \ref{ex:swedish-ovs} and \ref{ex:swedish-avso} are being
+%~~ although the fronted part of sentence (\ref{ex:swedish-ovs}) and (\ref{ex:swedish-avso}) are being
 %emphasized.
@@
 {Du & ser &  inte &mig.}
-%empha glossning!! ha samma överallt, ' eller ''
 {\emph{you} & \emph{see} & \emph{not} &\emph{ me}}
@@
 There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using a modal auxiliary verb \\
+\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli' (\emph{`become'}).\\
 %>> better with animate subjects
@@
 \label{gfPass:s-pass}
-The s-passive is the more commonly used than periphrastic passive, for both written
+The s-passive is more commonly used than periphrastic passive, for both written
 and Swedish \cite[Pass. \textsection 1]{SAG} and dominates especially when the subject is inanimate.
@@
 \subsection*{Impersonal constructions}
-Constructions with `det är/var' (\emph{`it is/was'}) \\is very common in Swedish
+Constructions with `det är/var' (\emph{`it is/was'}) \\are very common in Swedish
 \cite[\textsection 309d]{H&H}:
@@
 %There has been much research about parsing Swedish. INTRO!!
-Many years of interesting research has lead to many interesting language
+Many years of research has lead to many interesting language
 technology tools for Swedish.
@@
 The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
-gives a full syntactic analysis as well as semantics represented in `quasi logical form'. A
+gives a full syntactic analysis as well as semantics represented in `Quasi logical form'. A
 translation to English  was implemented and the work was further developed in the spoken
-language translator (xx,xx). Unfortunately, it is no longer available. The coverage of the Swedish
-CLE is also reported to be very limited (Nivre,2002).\\
+language translator \cite{spoken}. Unfortunately, it is no longer available. The coverage of the Swedish
+CLE is also reported to be very limited \cite[p. 134]{nivretrees}.\\
 
@@
 %TODO when??
-The basic algorithm for importing Saldo was implemented by Angelov (2009)
+The basic algorithm for importing Saldo was implemented by Angelov (2008)
 and  produces code for a GF lexicon.
@@
 to the the GF smart paradigms. For a verb, this will in most cases mean giving
-the present form. \\
+the present tense form, see figure \ref{fig:saldoknyt}. \\
 
@@
 \verb-mkV "knyter" ;-
-\caption{First code produced for the verb `knyta'}
+\caption{First code produced for the verb `knyta' (\emph{`tie'})}
 \label{fig:saldoknyt}
@@
 All assumed paradigms are printed to a temporary lexicon, 
-which is when compiled produces an inflection table
-for every entry. The tables are compared to the information given
+which will produce an inflection table for every entry when compiled.
+The tables are compared to the information given
 in Saldo, if the tables are equal the code for the word is saved. If the table
@@
 correct inflection table. In the next try both the present and the past tense
-is given:\\
+are given:\\
 
@@
 
-Figure \ref{pic:TabVax} shows a the information given by Saldo and by GF respectively.\\
+Figure \ref{pic:TabVax} shows the information given by Saldo and by GF respectively.\\
 \begin{figure}[h]
@@
 \end{figure}
-The tables does not overlap completely, there are some more forms in Saldo's table (eg. the 
+The tables does not overlap completely, there are some more forms in Saldo's table (e.g. the 
 compounding form ``väx-")
-while the one generated in GF contains some that are not in Saldo (eg. ``vuxits").
+while the one generated in GF contains some that are not in Saldo (e.g. ``vuxits").
+As GF concerns about grammar only, and not semantics, and  as the GF tables are automatically
+generated, they always contain all word forms, although some forms may never be
+used in the natural language.
 Saldo may also contain variants: ``växt" and ``vuxit" are both supine form.
@@
 
-The program has been made more robust than the previous version. It also prints
-log files providing
+During this project, the program has been made more robust than the previous
+version. It also prints log files providing
 information about the process of importing each word: which paradigms that have been
@@
 
-Each entry in Saldo has an identifier, eg. \verb-äta..vb-, which are used as 
+Each entry in Saldo has an identifier, e.g. \verb-äta..vb-, which is used as 
 constant names in the GF lexicon. However, the identifier may need some renaming since
@@
 the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions %(\emph{pp})
-, numerals %(\emph{nl}),
+Saldo in the GF lexicon. Prepositions, %(\emph{pp})
+numerals %(\emph{nl}),
 personal pronouns etc.
@@
 
+\begin{figure}[h]
 \begin{tabular}{|l|lll|}
@@
  Reflexive verbs  &\textbf{vbm}& \textbf{V} & raka sig (\emph{shave})\\
- Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till (\emph{perk})\\
+ Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till (\emph{perk up})\\
 \hline
-\end{tabular}\\
+\end{tabular}
+\caption{}
+\end{figure}
 
@@
 %different grammatical analyse, a lot of pronouns. Example. 
-Most, but not all words of the categories listed above have been imported.
+Most but not all words of the categories listed above have been imported.
 One reason why the importing phase would fail 
@@
 When testing the coverage of Talbanken,
-we found that there are around 2500 word forms still missing, excluding the one
+we found that there are around 2500 word forms still missing, excluding the ones
 tagged as names and numbers. This number may seem very high, but %considering that
-4/5 of the word forms are compounds. We should also take into consideration that 
+4/5 of the word forms are compounds and when performing the inteded parsing,
+these an additional analysis before being looked-up in the lexicon. We
+should also take into consideration that 
 we cannot automatically find out how many actually stem from the same word, or
@@
 
+\begin{figure}[h]
 \begin{tabular}{|l|l|}
@@
 \hline
-\end{tabular}\\
+\end{tabular}
+\caption{}
+\end{figure}
 
@@
 still compatible with this.
-Before describing the work progress in section \ref{sec:Added}, we will give an introduction to the
+Before describing the actual implementation in section \ref{sec:Added}, we will give an introduction to the
 resource grammars in general and to the Swedish implementation in particular.
@@
 \begin{verbatim}
-PredVP       : NP -> VP -> Cl ; -- Prediciation
+PredVP       : NP -> VP -> Cl ; -- Predication
 ComplVPSlash : VPSlash  -> VP ; -- Complementation
@@
 resource grammars. 
-A lexicon normally consists of words from the open word-classes, shown in rectangular boxes in
-the picture. Each lexical entry is assigned with a type describing its word-class.
+%A lexicon normally consists of 
+Words from the open word-classes are shown in rectangular boxes in
+the picture. Each lexical entry is assigned a type describing its word-class.
 %and the abstract grammar can be viewed as a set rules showing
@@
 \begin{figure}[h]
-\includegraphics[height=90mm]{categories.png}
+\hspace{-18mm}
+\includegraphics[height=130mm]{categories.png}
 \caption{The types of the GF resource grammars.}
@@
 
+\newpage
 \subsection{Noun phrases}
 Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
-and \verb-N'- to separated between nouns with respectively without a set number.
-The full analysis includes a separation between the categories \verb-N- -- simple nouns --
+and \verb-N'- to distinguish nouns with respectively without number information.
+The full analysis includes a distinction between the categories \verb-N- -- simple nouns --
 and 
@@
 The \verb-N- category is implemented as an inflection table, generated by the smart paradigm,
-and a field keeping information about the gender.
+and a field keeping information about the gender, see figure \ref{fig:gfflicka}.
 \begin{figure}[h]
@@
 \caption{Representation of the noun `flicka' (\emph{`girl'}) in GF}
-\label{fig:gfpic}
-\end{figure}
+\label{fig:gfflicka}
+\end{figure}\\
 When a noun is turned into a \verb-CN- %, this information is kept, and the noun may
-which may be modified by adjectival phrases or conjucted.
+which may be modified by adjectival phrases or conjoined.
 \begin{verbatim}
@@
 The function \verb-DetCN-, determination, creates a
-\verb-NP- by setting the number and definiteness of a \verb-CN-.
+\verb-NP- by setting the number and definiteness of a \verb-CN-, shown in figure
+\ref{gfcode:DetCN}.
 %may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
@@
 Hence, all determiners
-in our grammar must keep information of which definiteness they allow; the
-\verb-DetSpecies- parameter stored as an inherent 
+in our grammar must keep information about which definiteness they require; the
+\verb-DetSpecies- parameter is stored as an inherent 
 feature of the determiner.
-The resource grammar differentiates between quantifiers, 
+The resource grammar distinguishes between quantifiers, 
 determiners \verb-Det- and predeterminers. Predeterminers modify \verb-NP-s
@@
 \emph{en} and \emph{ett} for singular. In plural it
-is the either \emph{``de"} or nothing, cf. sentence \ref{sent:trottkatt}a and \ref{sent:trottkatt}b.
+is the either \emph{``de"} or nothing, cf. sentence (\ref{sent:trottkatt2}a) and
+\ref{sent:trottkatt2}b.
+\enumsentence{\begin{tabular}[t]{@{}*{9}{l@{\ }}}   %\begin{tabular}[t]{@{}*{lllllll}}
+a. & Katten        &sov. &&b. &\textbf{Den}  &gamla & katten  &sov.\\
+   &\emph{cat}\small\sc+def &\emph{slept}&& &\emph{the} & \emph{old} & \emph{cat}\small\sc+def &\emph{slept}\\
+\end{tabular}}\label{sent:trottkatt2}
+
 We need to know if the common noun phrase 
@@
 %>> CN may be different in other theories.
----------
 
@@
 In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
-ser katten". The verb ``såg" (see) takes one object: ``katten".
+ser katten". The verb `ser' (\emph{`see'}) takes one object: `katten'.
 In GF, this can be seen on the type of the verb,
@@
 The function \verb-SlashV2a- lets you create a \verb-VPSlash-, a verb phrase missing a
-noun phrase \cite{gazdar}. When combine the \verb-SlashVP- with the object we
+noun phrase \cite{gazdar}. When we combine the \verb-SlashVP- with the object we
 get a complete verb phrase.
@@
 \vspace{5mm}
-The verb phrase fields are not put in their correct order until the tense, type of clause etc is
-known, ie. when a sentence is created.\\
+The verb phrase fields are not put in their correct order until the tense and type
+of clause is determined, ie. when a sentence is created.\\
 
 Swedish verbs may take up to five arguments \cite[p. 53]{stymne}.
-\enumsentence{Jag tar med mig den till honom}\label{sent:tamed}
-These may be prepositions, particles, reflexive object, indirect objects and direct objects.
-In the grammar these will be encoded in different ways. The particles are
-given in the lexicon, as well as the prepositions that are chosen by the verbs.
-The entry for `\emph{ta med}' as used in sentence \ref{sent:tamed}, is described as follows:
+These may be prepositions, particles, reflexive object, indirect objects
+and direct objects. 
+\enumsentence{\shortex{7}
+{Jag &tar& med &mig& den& till& honom}
+{\emph{I} &\emph{take}&\emph{with} &\emph{me}& \emph{it}& \emph{to}& \emph{him}}
+{`I bring it to him'}
+\label{sent:tamed}}
+The verb `ta' in sentence (\ref{sent:tamed}) takes one particle (\emph{med}),
+one preposition (\emph{till}) and two objects: \emph{den} and \emph{honom}.
+
+In a GF lexicon this verb is given the
+category \verb-V3-, %classifing it as 
+a \textit{three-place verb}, taking two objects.
+The notion \verb-V3- is motivated by the formal translation: \verb-bring(I,it,him)-.\\
+%these will be encoded in different ways.
+Particles are given in the lexicon, as well as the prepositions that are
+chosen by the verb.
+The entry for `\emph{ta med}', as used in sentence (\ref{sent:tamed}), is
+described as follows:
 %, assuming that we already have an constant \verb-take_V- which produces the correct conjugation for ``ta":
@@
 \end{verbatim}
-The category \verb-V3- tells us that the verb takes two objects, % (``den" and ``honom").
-the second one is to be used with the preposition `\emph{till}'. \verb-reflV- shows
-that the verb always is used with a reflexive
+%The category \verb-V3- tells us that the verb takes two objects, % (``den" and ``honom").
+The function \verb-dirV3- creates a three-place verb, where the first object is direct and
+the second is to be used with the preposition given as the last argument: `\emph{till}'.
+\verb-reflV- shows that the verb always is used with a reflexive
 pronoun and \verb-partV- gives the particle \emph{`med'}.
@@
 are treated and represented differently internally. The fronting of the
-preposition as in of sentence \ref{sent:tittapa}a. is accepted but
-fronting of particles as in\ref{sent:tittapa}b. is not.
+preposition, as in of sentence (\ref{sent:tittapa}a)., is accepted but
+fronting of particles, as in \ref{sent:tittapa}b., is not.
+%TODO use png to get correct size
 \eenumsentence{
-\item[a]På pojken tittar du 
-\item[b] *På när du springer tittar jag}\label{sent:tittapa}.
+\item[a]\shortex{4}
+{På &pojken& tittar& du.}
+{\emph{on} &\emph{the boy}& \emph{look}& \emph{you}}
+{`You look at the boy.'}
+\item[b]\shortexnt{6}
+{*På& när& du& springer& tittar& jag}
+{\emph{~on}& \emph{when}& \emph{you}& \emph{run}& \emph{watch}& \emph{I}}
+\label{sent:tittapa}}
 \begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=80mm]{tittapa.pdf}}
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=90mm]{tittapa.pdf}}
 \hspace{10mm}
@@
 
+\newpage
 %Categories as V,V2,VS...
@@
 The category clause, \verb-Cl-, represents a pre-sentence, that does not yet have any tense, polarity
-or word order set.\\
-
+or word order set, see figure \ref{fig:clause}.\\
+
+\begin{figure}[h]
 \begin{tabular}{|lll|ll|}
@@
 \hline
-\end{tabular}\\
+\end{tabular}
+\caption{}
+\label{fig:clause}
+\end{figure}
 
@@
 This is a convenient way to form questions, relative clauses
-and topicalized clauses.\\
-
+and topicalized clauses, as in figure \ref{fig:clause2}.\\
+
+\begin{figure}[h]
 \begin{tabular}{|l|ll|}
@@
 \hline
-\end{tabular}\\
+\end{tabular}
+\caption{}
+\label{fig:clause2}
+\end{figure}
 
@@
 \includegraphics[width=70mm,height=70mm]{apples.pdf}
-\caption{Parse tree for ``Har han inte ätit de gula äpplena idag".}
+\caption{Parse tree for ``Har han inte ätit de gula äpplena idag?"}
 \label{gfSwe:parsetree}
@@
 \eenumsentence{\item[a]\shortexnt{5}{Hon & ser & pojken & som  &sover}
-              {\emph{she} & \emph{sees} & \emph{the boy} & \emph{who} & \emph{sleeps}}
+              {\emph{she} & \emph{sees} & \emph{the boy} & \emph{that} & \emph{sleeps}}
               %{`She sees the sleeping boy'}}
@@
 %can be found.
-%For example, \verb|FocObj| fronted the object as in sentence \ref{gfSwe:apple}.
+%For example, \verb|FocObj| fronted the object as in sentence (\ref{gfSwe:apple}).
 \enumsentence{\shortex{6}
 {Det & äpplet & vill & jag & inte & ha}
-{that & apple & want & I & not &}
+{that & apple & want & I & not &have}
 {`I don't want that apple'}}\label{gfSwe:apple}
 and for preposition stranding:
-\enumsentence{\begin{tabular}{ll}
-Stranded preposition & cf.\\
-Vem måste jag akta mig \textbf{för}? & \textbf{För} vem måste jag akta mig?\\
-Who do I need to watch out \textbf{for}? & \textbf{For} whom do I need to watch out?\\
+\enumsentence{
+\begin{tabular}{ll}
+~~Stranded preposition &~ cf.\\
+{\begin{tabular}{llllll}
+Vem& måste& jag& akta& mig &\textbf{för}? \\
+\emph{who}& \emph{must}& \emph{I}& \emph{watch out}& \emph{me} &\textbf{\emph{for}}? \\
+\end{tabular}}&
+{\begin{tabular}{llllll}
+\textbf{För}& vem& måste& jag& akta& mig?\\
+\emph{\textbf}{for}& \emph{who}& \emph{must}& \emph{I}& \emph{watch out}& \emph{me}?\\
+\end{tabular}}\\
+~~`Who do I need to watch out \textbf{for}?' &~ `\textbf{For} whom do I need to watch out?'\\
 \end{tabular}}
@@
 and the verb-second property
-were covered by the resources, as well as the periphrastic passive.
-A start on topicalisation was implemented, while
-the others are examples of what has been added during this project.
+were covered by the resource grammar, as well as the periphrastic passive and
+a limited form of topicalisation. 
+The other constructions have been added during this project.
 %enough to be covered by the resources. The others have been added.
@@
 \shortex{5}{Uppsatsen & skrev\textbf{s} & av & en & student.}
-        {the essay & wrote+\textbf{s} & by & a & student}
+        {\emph{the essay} & \emph{wrote+\textbf{s}} & \emph{by} & \emph{a} & \emph{student}}
         {`The essay was written by a student.'}}\label{sent:skrevs}
@@
 where not all words have passive forms for all tenses. The Norwegian 
-translation of sentence \ref{sent:skrevs} is:
+translation of sentence (\ref{sent:skrevs}) is:
 %The periphrastic passive is preferred, especially in spoken langauge.
@@
 The corresponding Swedish sentence is acceptable, but not as natural sounding as sentence
-\ref{sent:skrevs}.
+(\ref{sent:skrevs}).
 The resource grammar for Scandinavian therefore implemented the function for passive,
@@
 \end{verbatim}
+A \verb-V3- like `give' in sentence (\ref{sent:give2pass}) gives rise to two
+passives, (\ref{ex:passV32}) and (\ref{ex:passV33}).
+\enumsentence{\textbf{Active use of two-place verb}\\\shortex{4}
+{Vi & erbjöd& henne& jobbet }
+{\emph{we} & \emph{offered}& her&\emph{the job}}
+{`We offered her the job'}}
+\label{sent:give2pass}
 \enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
 {Jobbet & erbjöds & henne }
-{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ henne}}
-{``The job was offered to her"}}
+{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
+{`The job was offered to her'}}
 \label{ex:passV32}
@@
 {\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
-{``She was offered the job"}}
+{`She was offered the job'}}
 \label{ex:passV33}
@@
 \label{sec:Formal}
-Formal subjects \cite[\textsection 19]{SAG} is often used in Swedish.
+Formal subjects \cite[\textsection 19]{SAG} are often used in Swedish.
 \enumsentence{\shortex{6}
@@
 special GF category, \verb|SimpleVP|, to exclude other verb phrases
-like the one in sentence \ref{sent:fagelfron}. Any intransitive verb can form
+like the one in sentence (\ref{sent:fagelfron}). Any intransitive verb can form
 a \verb-SimpleVP-, as can transitive verbs without their objects. The \verb-SimpleVP-
@@
 definite form.
-\enumsentence{\begin{tabular}{ll}
-*Det sitter \textbf{den fågeln} på taket. & *Det sitter \textbf{fågeln} på taket.\\
-``There are that bird sitting on the roof'' & ``There are the bird sitting on the roof'' \\
+\enumsentence{\begin{tabular}{llllll}
+*Det     & sitter     & \textbf{den}& \textbf{fågeln}& på       & taket. \\
+%& *Det sitter \textbf{fågeln} på taket.\\
+\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}\\
+%& ``There are the bird sitting on the roof'' \\
 \end{tabular}}
-Neither is sentence \ref{sent:dumfagel} correct.
-\enumsentence{
-*Det sitter \textbf{min fågel} på taket\\
-``There are my bird sitting on the roof''\\
+Neither is sentence (\ref{sent:dumfagel}) correct.
+\enumsentence{\begin{tabular}{llllll}
+*Det& sitter& \textbf{min}& \textbf{fågel}& på& taket\\
+\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}\\
+%``There are my bird sitting %on the roof''\\
+\end{tabular}
 }\label{sent:dumfagel}
-The distinction lays in the determiner. The different types of determiners are
-showed in figure \ref{tab:detdef}. 
-For real subjects, the determiner must be of type \verb-DIndef-,
-that requires both the noun and the adjective to be used in indefinite form.
+In order to implement this, we utilized the 
+%The distinction lays in the determiner. The
+different types of determiners shown in figure \ref{tab:detdef}. 
+For postverbal subjects, the determiner must be of type \verb-DIndef-,
+that requires both the noun and the adjective to be indefinite.
 %The quantifier must have determiner must havenoun must be in indefinite form,
@@
 We would also like to differentiate between the \verb-NP-s in sentence
-\ref{sent:mosthorses}a and b, where 
+(\ref{sent:mosthorses}a,b), where 
 `\emph{av}' should be used
@@
 \label{sec:reflexives}
-An important piece in a Swedish grammar is the reflexive pronouns and
+An important area in a Swedish grammar is the treatment of the reflexive pronouns and
 the reflexive possessive pronouns as described on p. \pageref{swe:refl}.
@@
 \end{tabular}}\label{ex:ReflConj} \label{ex:ReflConj2}
-When a reflexive pronoun is used in a noun phrase, it is hence not
-allowed as subject. The antecedent must not be out of scope for the closest finite verb,
-as in sentence \ref{sent:sesig}.
+Reflexive pronouns can not be used in subject noun phrases of finite sentences,
+as shown by the ungrammatical examples in %, it is hence not
+%allowed as subject. The antecedent must not be out of scope for the closest finite verb,
+sentence (\ref{sent:sesig}) and (\ref{ex:ReflSubj}).
+The third person reflexives (`sig',`sin') requires a third person antecedent (see \ref{ex:refljag}).
+Furthermore, the antencedent must be within the same finite sentence as the reflexive
+pronoun, see (\ref{sent:sesig}).
 The grammar should not accept any of these sentences:
@@
 \enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give {\sc self's} money to {\sc self's} siblings. }
+I give {\sc self's} money to {\sc self's} siblings. }\label{ex:refljag}
 \enumsentence{*Han vill att jag ser på sig.\\
@@
 Apart from these restrictions, noun phrases containing reflexive pronouns
-may be used as any other \verb-NP-. They may be conjuncted (sentence
-\ref{ex:ReflConj} a and b) and determined (sentence \ref{ex:ReflPredet}).
+may be used as any other \verb-NP-. They may be conjoined 
+(\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
 %and used as genitives (sentence \ref{ex:ReflGen}) as such.
@@
 \end{verbatim}
-This way, using the phrases as subject is excluded, but neither of sentence
-\ref{ex:ReflConj}a, \ref{ex:ReflConj}b or \ref{ex:ReflPredet} are allowed.\\
+This way, using the phrases as subject is excluded, but none of sentence
+(\ref{ex:ReflConj} a,b) or (\ref{ex:ReflPredet}) are allowed.\\
 The simplest way to fix this is to treat the phrases as normal \verb_NP_s.
@@
 keeping them from being used as subjects. We have ignored the third restriction
-and are allowing sentences like \ref{ex:ReflSubj}.
+and the grammar would generate and accept sentence (\ref{ex:ReflSubj}).
 
@@
 \verb_Obj_, which is
-identical to normal \verb_NP_ except for that it depends on 
+identical to normal \verb_NP_ except that it depends on 
 the antecedent. All functions where objects are modified in the same way
@@
 The dependence on the antecedent needs to be applied to adverbial phrases and
-adjective phrases as well. The adjective phrase in sentence \ref{ex:ReflAP}a and
-the adverbial phrase in sentence \ref{ex:ReflAdv}b are examples of this; if the
+adjective phrases as well. The adjective phrase in (\ref{ex:ReflAP}a) and
+the adverbial phrase in (\ref{ex:ReflAdv}b) are examples of this; if the
 subject was changed to 2nd person, they  would need to change too.
-\enumsentence{\begin{tabular}{ll}
-\textbf{Du} är längre än \textbf{din} kompis. &*Du är här oftare än \textbf{sin} kompis.\\
-You are taller than your friend. &You are here more often than {\sc self's} friend.\\
+\enumsentence{\begin{tabular}{llll}
+a. &\textbf{Du} är längre än \textbf{din} kompis.& b. &*Du är här oftare än \textbf{sin} kompis.\\
+ & You are taller than your friend. &&You are here more often than {\sc self's} friend.\\
 \end{tabular}} \label{ex:ReflAP2} \label{ex:ReflAdv2}
-The dependency is spreading through the grammar and in order to avoid all code
+The dependency spreads through the grammar and in order to avoid all code
 duplication we use another approach.
@@
 \end{verbatim}
-The solution given in this project is making use of this generalization and
-introduces the use of dependent types in a resource grammar.
+The solution chosen in this project is to make use of this generalization and
+introduce the use of dependent types in a resource grammar.
 Following the idea given above, we make a difference between subjects and objects, but not
@@
 The solution hence combines the-part-of-speech driven analyse normally preformed by
-GF with a part-of-sentence analyse.
+GF with a part-of-sentence analysis.
 
@@
 \enumsentence{\shortex{3}
-{Taket & på sitt hus &}
+{taket & på sitt hus &}
 {\scriptsize{\textbf{NP a +}} & \scriptsize{\textbf{(Adv Obj)}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
-{\emph{The roof of {\sc self's} house}}}
+{\emph{the roof of {\sc self's} house}}}
 %{{\sc NP +} & {\sc (Adv with Obj)} & $\Rightarrow$ {\sc Obj}}
 \enumsentence{\shortex{3}
-{Sitt hus & på berget &}
+{sitt hus & på berget &}
 {\scriptsize{\textbf{NP Obj + }} & \scriptsize{\textbf{Adv a}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
@@
 The use of dependent types separates our grammar from the other resource grammars. Such separation
-is not desirable in it self, but there are at least three reasons to why we
+is not desirable in itself, but there are at least two reasons why we
 believe it is appropriate in this case:
@@
 \item The possibility for noun phrases to agree with an antecedent 
-      could be useful in other languages. For example, the Slavic languages
+      will be useful in other languages. For example, the Slavic languages
       have similar reflexive pronouns.
-\item The goal of the project is to make a language specific grammar, and it would be surprising if this could be
-      done while leaving the abstract grammar unchanged.
 \end{itemize}
+Moreover, the goal of the project is to make a language specific grammar, and it
+would be surprising if this could be done while leaving the abstract grammar unchanged.
 
@@
 The rules for reciprocal pronouns could be formalized using the same idea. As reflexive pronouns, reciprocals
-must not be used as subject and further they may only be used when the number of the antecedent is plural.
+must not be used as subjects and furthermore they may only be used when the
+number of the antecedent is plural.
 %
@@
 \subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian have two modal verbs for future tense
+Swedish and Norwegian use two auxilirary verbs for future tense
 \cite[p. 246]{H&H}:
@@
 \end{verbatim}
-\caption{The covered and accepted usage of the future modal verb `komma'}
+\caption{The covered and accepted usage of the future tense with `komma'}
   \label{fig:kommeratt}
@@
 The GF analysis distinguishes between two categories of adverbs: \verb-Adv- 
-%eg. `nu' (\emph{`now'}) and
+%e.g. `nu' (\emph{`now'}) and
 and \verb-AdV-. %like `alltid' (\emph{`always'}).
-The \verb-AdV- (eg. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}))
+The \verb-AdV- (e.g. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}))
 attaches directly to the verb.
@@
 when emphasizing the verb itself.
-This is an example of a \textit{focusing adverb}, another example is `inte ens' (\emph{`not even'}).
+This is an example of a \textit{focusing adverb}, others examples are `inte ens' (\emph{`not even'})
+and `till och med' (\emph{`even'}).
 \enumsentence{\shortex{9}
@@
 
-The focusing adverbs are usually not combined with modal verbs, since the modal
-verbs rarely are focused.
-\enumsentence{\shortexnt{5}
+The focusing adverbs are usually not combined with modal verbs, the copula or temporal auxiliries. 
+% since the modal verbs rarely are focused.
+\eenumsentence{\item[a]
+\shortexnt{5}
+{?&Han & bara&  är&  dum}
+{&\emph{he} &\emph{ only} & \emph{ is}& \emph{ stupid}}
+\item[b]
+\shortexnt{5}
 {?&Han & bara&  har&  sovit}
@@
 When no relative clause is present, the definitive form must be used, cf.
-sentence \ref{sent:debocker}.
+sentence (\ref{sent:debocker}).
 \enumsentence{De uppfattningar\textbf{na} förs fram.\\
@@
 %would require a. 
-Consequently, sentence \ref{sent:detfaller} is still
+Consequently, sentence (\ref{sent:detfaller}) is still
 accepted by the grammar.
@@
 the same ways as `sådana' (\emph{`such'}) that acts either as a
-determiner, as in sentence \ref{ex:QuantDet}, or as an adjective as
- in sentence \ref{ex:QuantAdj}. 
+determiner, as in sentence (\ref{ex:QuantDet}), or as an adjective as
+ in sentence (\ref{ex:QuantAdj}). 
 \eenumsentence{\item[a.]Sådana \label{ex:QuantDet}katter vill jag ha
@@
 %they got we could sort out the ones acting either as
-%determiners, as `sådana' in sentence \ref{ex:QuantDet}, or as adjectives as
-%`sådan' in sentence \ref{ex:QuantAdj}. 
+%determiners, as `sådana' in sentence (\ref{ex:QuantDet}), or as adjectives as
+%`sådan' in sentence (\ref{ex:QuantAdj}). 
 %Since all determiners can be used alone as noun phrases, sentence
@@
 \caption{Meta tree}
-\label{pic:gfEtree}
+\label{pic:gMetatree}
 \end{wrapfigure}
@@
 Another annotational difference that turned out to be problematic for the
-translation can be illustrated by sentence \ref{ex:mappMan}, containing
+translation can be illustrated by sentence (\ref{ex:mappMan}), containing
 the generic pronoun \emph{``man"}.
@@
 and objects. %(which is not needed in GF (see section \ref{sec:swegf}))
-The analyse of sentence \ref{sent:krav}
+The analyse of sentence (\ref{sent:krav})
 \enumsentence{För stora krav.\\
@@
 }\label{sent:water}
-The parse tree of sentence \ref{sent:water} looks as follows:
+The parse tree of sentence (\ref{sent:water}) looks as follows:
 \begin{verbatim}
@@
 
-This however is not the correct analysis for `hund' in sentence \ref{sent:hund}.
+This however is not the correct analysis for `hund' in sentence (\ref{sent:hund}).
 \enumsentence{\shortex{4}
@@
 in this project as well \cite{}.
-
+ >> The many word forms in the gF tables, as seen in section \ref{saldo}.
+ >>This is a possible source for overgeneration.
 As noted, we additionally need to come up with better methods for using the lexicon when

### 2011-12-29 doc/report/report.tex
report getting better, pics and format

@@
 \usepackage[margin=1.5in]{geometry}
+\usepackage{parsetree}
 \usepackage{lingmacros}
@@
 Finally, I would also like to thank my friends and family. Special thanks to Dan % Rosén
-for all his support, advice and patience and, most importantly, for being such a good friend. %for having listened to 
+for all his support, advice and patience and  -- most importantly -- for being such a good friend. %for having listened to 
 %my monologues about the daily progress. 
@@
 \end{figure}
-
+\newpage
 %The two basic articles are \verb-IndefArt- and \verb-DefArt- which
@@
 can be combined, as shown in figure \ref{tab:detdef}
-\begin{figure}
+\begin{figure}[h]
 \begin{tabular}{|lll|l|}
@@
 \begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=90mm]{tittapa.pdf}}
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=90mm]{tittapa.jpg}}
 \hspace{10mm}
@@
 \begin{figure}[h]
-\includegraphics[width=70mm,height=70mm]{apples.pdf}
+\includegraphics[width=70mm,height=70mm]{apples1.pdf}
 \caption{Parse tree for ``Har han inte ätit de gula äpplena idag?"}
@@
 \enumsentence{
+~~Stranded preposition \hspace{45mm} cf.\\
 \begin{tabular}{ll}
-~~Stranded preposition &~ cf.\\
 {\begin{tabular}{llllll}
@@
 \eenumsentence{
-\item[a.]Frågan är vilken färg den har
-\item[b.]Pojken, som är blyg, tystnar
-\item[c.]Han såg kunden som tyckte om sallad
-\item[d.]Jag tänkte på huset i vilket hon bodde}\label{sent:relex} % better example!!
+%\item[a.]Frågan är vilken färg den har
+\item[a.]Pojken, som är blyg, tystnar\\
+`The boy, who is shy, falls silent'
+\item[b.]Han såg kunden som tyckte om sallad\\
+`He saw the customer who liked salad'
+\item[c.]Jag tänkte på huset i vilket hon bodde\\
+`I thought about the house in which she lived'}\label{sent:relex} % better example!!
 
@@
 relative clause.
+Talbanken contained several examples where the modified noun is in the 
+indefinite form as in (\ref{sent:uppfatt}).
+%TODO ref to H&H
 \enumsentence{\shortex{7}
-{De uppfattningar& som& förs& fram ...} 
+{de uppfattningar& som& förs& fram ...} 
 {\emph{the opinions}& \emph{that}&\emph{ put+{\sc passive}} & \emph{forward ...}}
-{``The opinions presented ..."}}
+{``the opinions, that are presented ..."}}\label{sent:uppfatt}
 %or together with the verb \emph{``finnas"}.
-When no relative clause is present, the definitive form must be used, cf.
+When no relative clause is present, the definite form with the postnominal
+definite article must be used, cf.
 sentence (\ref{sent:debocker}).
-\enumsentence{De uppfattningar\textbf{na} förs fram.\\
-``The opinions are presented."}\label{sent:debocker}
+\eenumsentence{
+\item de uppfattningar\textbf{na} förs fram.\\
+``the opinions are presented."
+\item *de uppfattningar förs fram}\label{sent:debocker}
 
@@
 logic and mathematics books.
-\enumsentence{\begin{tabular}{ll}
-\emph{From the resource grammar}& \emph{An alternative formulation} \\
-Jag vill ha en katt sådan att den inte fäller & Jag villa ha en sådan katt som inte fäller\\
-\end{tabular}} 
+\eenumsentence{%\begin{tabular}{ll}
+\item\emph{From the resource grammar}\\
+Jag vill ha en katt sådan att den inte fäller\\
+`I want a cat such that it does not shed'
+\item
+\emph{An alternative formulation} \\ 
+Jag vill ha en sådan katt som inte fäller\\
+`I want such a cat that it does not shed'}
+%\end{tabular}} 
 In GF, this relative clause \emph{``en katt sådan att den inte fäller"} is constructed by
@@
 useful.
-However a carefully prepared analysis and motivation is needed. %ationamount in a bigger reconstruction which therefore
+However a carefully prepared analysis and motivation is needed, %ationamount in a bigger reconstruction which therefore
 %would require a. 
-Consequently, sentence (\ref{sent:detfaller}) is still
-accepted by the grammar.
+and sentence (\ref{sent:detfaller}) is still accepted by the grammar.
 
@@
 \subsubsection{An example of overgeneration and how it can be solved}
-The \verb-Extra- module contained a function for constructing genitives.
+The \verb-Extra- module contains a function for constructing genitives.
 \begin{verbatim}GenNP : NP -> Quant ;\end{verbatim}
-It selected the possessive form given in the \verb-NP- and changed its type
+It selects the possessive form given in the \verb-NP- and changed its type
 to \verb-Quantifier-. 
@@
 %\end{verbatim}
-The \verb-GenNP- function was severely overgenerating when this project started, even
+The \verb-GenNP- function %was %severely
+overgenerated when this project started, even
 though its implementation was not erroneous in itself. The cause of the overgeneration 
@@
 {Somligas &skor& står& kvar& i& hallen}
-{Some's &shoes& stand& left& in& the hall}
+{\emph{some's }&\emph{shoes}& \emph{stand}& \emph{left}& \emph{in}& \emph{the hall}}
 {``Some peoples shoes are still in the hall"}}
@@
 This method is
-both faster and more reliable than going through the list by hand and categorise them
-by manually looking up each word or by using introspection.
+both faster and more reliable than going through the list by hand and categorising them
+manually by looking up each word or by using introspection.
 
@@
 Talbanken contains much valuable information about phrase structures and word 
-usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974) which is 
+usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974).% which is 
 %>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
@@
 %annotation scheme
-commonly used, and hence very well-tested, for describing Swedish.
-One part of this project has focused translating trees from Talbanken to GF
+%commonly used, and hence very well-tested, for describing Swedish.
+One part of this project has focused on translating trees from Talbanken to GF
 by constructing a mapping, which 
-automatically transforms trees in Talbanken format to GF abstract syntax trees.
+automatically transforms trees in the Talbanken format to GF abstract syntax trees.
 We hence get a comparison between the two annotations and at the same time
-we get methods to extract and translate information to GF notation. 
+we get methods for extracting and translating to GF notation. 
 Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
-of the sentence \emph{``The cat on the car gets bigger"} and the result of translating it to GF.
+of the sentence ``Katten på bilen blir större" (\emph{``The cat on the car gets bigger"})
+and the result of translating it to GF.
 Both the POS tags as well as the syntactic information are needed
 during the translation, so that all the information shown in the abstract tree
-(\ref{pic:gftree}) can be extracted. We need to know that \emph{``katten"} is a noun
+(\ref{pic:gftree}) can be extracted. We need to know that \emph{katten} is a noun
 used in definite form (\verb-DetArt-), singular (\verb-NumSg-). Even though the
@@
 is preserved.
-\emph{``På bilen"} should still form a subtree attached to \emph{``katten"}.
+\emph{på bilen} should still form a subtree attached to \emph{katten}.
 %produces a \verb-NP- from ``katten" this means that
@@
 \verb|OO| for object etc. Nearly 400 different tags are used, of which more than
-300 different are pos-tags. The high number is due to the detailedness,  there are for
+300 different are pos-tags. The high number is due to the level of detail,  there are for
 example almost 50 tags for nouns, excluding proper names. The tags
@@
 \textbf{WVIV} & vilja & The verb \emph{want} in infinitive\\
-\textbf{KVSP} & kommit att & The verb \emph{will} in supine\\
+\textbf{KVSP} & kommit att & The future auxiliary verb \emph{kommer} in supine\\
 \textbf{GVPT  PA} & gjordes & The verb \emph{do, make}, present, passive \\
@@
 The implementation consists of a set of rules that translates
-one node of a Talbanken tree to a GF function, possibly by describing how the
-nodes subtree should look. 
+one node of a Talbanken tree to a GF function, possibly by describing what the
+nodes subtree should look like. 
 \begin{verbatim}
@@
 \ref{pic:tbtree}. The verb \emph{``bli"} is
-followed by an
-adjective as a present participle, \verb|SP|. GF admits \emph{``bli"} to have
+followed by a predicative 
+adjective modifying the subject, \verb|SP|. GF admits \emph{``bli"} to have
 type \verb|VA|, a verb with an adjective as complement, and the translator can
@@
 for each verb category in order to find the right number of arguments.
-The mapping also covers changed word order, such as questions and topicalisation.
+The mapping also covers different word order, such as questions and topicalisation.
 
@@
 \subsection{Differences in the notation}
-Some structural differences between GF and Talbanken were made obvious during the
+Some structural differences between GF and Talbanken became obvious during the
 translation.
@@
 most transitive verbs can also be used as intransitive. Both sentences in
-example \ref{ex:mappTransitive} are grammatically correct, and
+example \ref{ex:mappTransitive} are grammatically correct but only (\ref{ex:mappTransitive}b)
+is accepted by the GF grammar, and
 the grammar rules are simply too strict for accepting many Talbanken sentences.
@@
 %>> motivation, why so bad??\\
-The development was mostly example-driven, and at least one rule for each
-Talbanken-tag has been implemented.
+The development was mostly example-driven, and at least one rule for the translation of
+every Talbanken-tag has been implemented.
 Shorter sentences have been prioritized in order to get a coverage of the most
 fundamental constructions. 
-Regression testing was used to ensure that a new rule wouldn't violate the old ones
-or allow trees to be translated in erroneously.
+Regression testing was used to ensure that a new rule would not violate the old ones
+or allow trees to be translated erroneously.
 
@@
 \begin{figure}[h]
-\centering
-\subfloat[Deep]{\label{pic:mappDeep}\includegraphics[width=20mm]{nofile.png}}
-\subfloat[Flat]{\label{pic:mappFlat}\includegraphics[width=20mm]{nofile.png}}
-\caption{...}
+\begin{parsetree}
+    ( .CNP. 
+        (.CJ.  (.NP. ~ `oskifta dödsbon'))
+        ( .++. (.++OC. `och'))
+        (.CJ.  (.VN. `familjestiftelser'))
+    )
+\end{parsetree}
+\begin{parsetree}
+    ( .CNP. 
+        (.CJ.  (.NP. ~ `oskifta dödsbon'))
+        ( .CC.
+            (.CNP. 
+                (.++.  (.++OC. `och' ))
+                (.C+.  (.VN. `familjestiftelser'))
+         ))
+    )
+\end{parsetree}
+\caption{An example of the difference in the two analyses}
 \label{fig:mappDeepFlat}
@@
 Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
-which groups an infinite verb and its object together into a \verb|VP|.
+which groups an inifinitival verb and its object together into a \verb|VP|.
 This information can however be extracted from the flat version, and the results
@@
 One of the reasons why a node cannot be translated, is the 
-use of the tags show in table \ref{fig:mapBadtag}.
+use of the tags show in figure \ref{fig:mapBadtag}.
 The \textbf{PU} tag is used for graphic listings, and not for fluent text.
@@
 the listings are meant for making the text look nice in
-folders etc and out of scope for the grammar itself. The tags \textbf{XX} and \textbf{NP} are often used since
+folders etc and outside the scope for the grammar itself. The tags \textbf{XX} and \textbf{NP} are often used since
 Talbanken -- unlike GF -- makes a difference between subjects
 and objects. %(which is not needed in GF (see section \ref{sec:swegf}))
-The analyse of sentence (\ref{sent:krav})
+The analysis of elliptical expression in (\ref{sent:krav})
 \enumsentence{För stora krav.\\
@@
 whether the noun phrase is used as subject or an object.
-The tags shown in table \ref{fig:mapBadtag} occur quite frequently in the treebank and are always translated
+The tags shown in figure \ref{fig:mapBadtag} occur quite frequently in the treebank and are always translated
 to metas, which lowers our result. \\
@@
 Our focus has been on shorter sentences, with no idioms or conjunction.
-If we assure that the lexicon know the correct word class for all lemmas
+If we assure that the lexicon contains the correct word class for all lemmas
 involved, we
@@
 65 \% coverage. 
-If we test randomly collected sentences that does not contain any of the tags listed
-in figure \ref{fig:mapBadtag}, 72 \% can be restored. \\
+If we test randomly collected sentences that do not contain any of the tags listed
+in figure \ref{fig:mapBadtag}, 72 \% can be restored (see figure \ref{tab:mappres})
 
@@
 %All NPs: 72%
+\begin{figure}[h]
 \begin{tabular}{|ll|}
@@
 \hline
-\end{tabular}\\
-
-These results are lower than the results of the mapping of Penn Treebank. The
-reason for this may be the fact that English is syntactically less complicated
+\end{tabular}
+\caption{}
+\label{tab:mappres}
+\end{figure}
+
+A mapping between GF and Penn Treebank has earlier been conducted \cite{gfMech}.
+The percentage of restored nodes from Peen Treebank is higher than our results.
+%Our results are lower than the results of the mapping of Penn Treebank.
+The reason for this may be the fact that English is syntactically less complicated
 than Swedish. Also, the Penn Treebank contains a lower number of tags, 82 compared
@@
 class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
-Thus the number combinations are extensively increased, and so is the number of
+Thus the number of combinations are increased, and so is the number of
 rules that would be needed to cover all constructions.
@@
 
-Parsing bigger parts of Talbanken would hence give very low results at this state, 
+Parsing bigger parts of Talbanken would hence give very low results at this stage, 
 %, for the reasons stated.
-and a comparison of the results would not be of much value,
+and a comparison of the results would not be of much value as
 there would not be enough material %underlag is yet too small 
@@
 running the test suite used for the grammar development.
-The sentences given in the test suite are short and the words are in most cases known
-to the lexicon, but there are also constructions that have not been implemented 
+The sentences given in the test suite are short, four to 10 words, and the
+words are in most cases included in
+the lexicon, but there are also constructions that have not been implemented 
 during this project.
@@
 We need valency information to make a good analysis. The lexicon is also too
-big to use trivially with the current techniques. Its size causes the
-incremental parsing algorithm use more heap memory than normal computers
+big to use with the current techniques. Its size causes the
+incremental parsing algorithm to use more heap memory than normal computers
 allow. To solve this, we need to use the lexicon data more cleverly.
@@
 \label{sec:future}
-Having finished this part of the project, we are left with 
+At the end of the current project,
+%Having finished this part of the project,
+we are left with 
 %Although this one part of the project is finished, there are 
@@
 \item
-\textit{Pronominal object shifts} are common in Swedish.
+\textit{Pronominal object shift} is common in Swedish and obligatory in Danish and Norwegian.
 \enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
@@
 \end{tabular}}
-Personal pronouns are allowed to shift position with the negation in main clauses without
-the modal verbs.
+%TODO Sats adverbials?
+Personal pronouns are allowed to precede negation and some other adverbials --\textit{satsadverb} --
+in main clauses without auxiliary verbs.
 
@@
 
-Although object shifts are frequently used, they are rarely found in
-Talbanken's part P, which have been the inspiration for this project.
-Therefore, its implementation has so far not been prioritized.\\
+Although object shifts are frequently used, they are hardly found in
+Talbanken's part P, which has been the inspiration for this project.
+Therefore, this implementation has so far not been prioritized.\\
 
@@
 
-This however is not the correct analysis for `hund' in sentence (\ref{sent:hund}).
+This however is not the correct analysis for `hund' (\emph{`dog'}) in sentence (\ref{sent:hund}).
 \enumsentence{\shortex{4}
@@
 {\emph{he} & \emph{begs} & \emph{her} & \emph{to} & \emph{clean} & {\sc self's} & \emph{room}}
-{He asks her clean her room}
+{`He asks her to clean her room'}
 \item[b]\shortex{7}
@@
 {\emph{he}& \emph{promises}& \emph{her}& \emph{to}&  \emph{clean} &\sc self's&\emph{room}}
-  {He promises her to clean his room}}
+{`He promises her to clean his room'}}
 
@@
 \subsubsection{Lexicon}
-Our lexicon still lacks some part that can be imported from Saldo.
+Our lexicon still lacks some parts that can be imported from Saldo.
 The multiple-word verbs (\verb-vbm-) and multiple-word nouns (\verb-nm-) should
@@
 \subsection{Making the parser robust}
-To achieve our goal of free parsing, we need to use statistical methods.
+To achieve our goal of parsing unrestricted text, we need to use statistical methods.
 The parser needs to be able to handle
@@
 never cover all compounds. We therefore need to be able to do
-compounding analyses. Saldo has tools for this, which might be usable
-in this project as well \cite{}.
- >> The many word forms in the gF tables, as seen in section \ref{saldo}.
- >>This is a possible source for overgeneration.
+compounding analysis. Saldo has tools for this, which might be usable
+in this project as well \cite{fm}.
 As noted, we additionally need to come up with better methods for using the lexicon when
@@
 One possibility is of course to do 
-deeper refinements of the parsing source code, an extensive work which is far out of scope
-for this project. Other solutions are to to either
-adapt the automatically lexicon for its domain and hence making it smaller and
+deeper refinements of the parsing source code, an extensive work which is far
+outside the scope for this project. Other solutions are to either
+adapt the lexicon automatically for its domain and hence making it smaller and
 faster, or to adapt the input sentences to a smaller lexicon by preprocessing.
@@
 gets higher. Our new grammar also have many more rules which contribute
-to increase the number of to interpretations of each sentence.
-
-GF already allows probabilities, and by using an large
+to increasing the number of interpretations of each sentence.
+
+GF already allows probabilities, and by using a large
 GF treebank, we can get more reliable statistics for our functions. 
-By implementing dependency probabilities \cite{}, our results would be even better.
-
+By implementing dependency probabilities, our results would be even better.
+
+%TODO
+%The many word forms in the gF tables, as seen in section \ref{saldo}.
+%This is a possible source for overgeneration.
 \subsubsection{Named entity recognition}
@@
 but in some cases we needed to part from the multilingual abstract and use other
-grammatical theories in order to be able to perform a good analysis.
+grammatical theories in order to arrive at a good analysis.
 

### 2011-12-29 testSuites/MkLex.hs
better lexicon extraction

@@ -2,51 +2,69 @@ module MkLex where
 import PGF
 import Control.Monad
 import System.Process
+import qualified Data.Map as M
 
+type LexMap = (Lex,Lex)
+type Lex    = M.Map String String
 
-mkDir :: [String] -> Language -> FilePath -> IO (FilePath,Language)
-mkDir ws lang bigPGF = do
-   pgf <- readPGF bigPGF  
+mkDir :: LexMap -> [String] -> Language -> PGF -> IO (FilePath,Language)
+mkDir lexs ws lang pgf = do
+--   putStrLn "Reading pgf..."
+--   pgf <- readPGF bigPGF  
    let morpho = buildMorpho pgf lang
        allLemmas = map fst $ concatMap (lookupMorpho morpho) ws
-   (cnc,abs) <- extractLemmas allLemmas
-   rawSystem "gf --make" [cnc] -- must wait for process!
+   putStrLn "Extracting new lexicon..."
+   (cnc,abs) <- extractLemmas lexs allLemmas
+   putStrLn "Compiling new grammar..."
+   readProcess "gf" ["--make",bigGF] []
+   putStrLn "Done"
    return (newPGF,newLang)
 
-extractLemmas :: [Lemma] -> IO (FilePath,FilePath)
-extractLemmas ws = do
+extractLemmas :: LexMap -> [Lemma] -> IO (FilePath,FilePath)
+extractLemmas (cnc,abs) ws = do
   writeHeaders
-  ex bigLexCnc newLexCnc
-  ex bigLexAbs newLexAbs
+  mapM_ (extract newLexCnc cnc) ws
+  mapM_ (extract newLexAbs abs) ws
   writeEnd
   return (newLexCnc,newLexAbs)
-
- where ex file newFile = do
-             lex <- liftM ((map words) . lines) $ readFile file
-             mapM_ (extract newFile) lex
-       extract newFile l@(x:xs)
-               | x `specelem` ws = appendFile newFile (unwords l++"\n")
-       extract _ _ = return ()
-
+  where
+ --             mapM_ (extract newFile) lex
+       extract newFile lex w
+               | Just a <- M.lookup (showCId w) lex = appendFile newFile (a++"\n")
+--       extract newFile l@(x:xs)
+--               | x `specelem` ws = appendFile newFile (unwords l++"\n")
+       extract _ _ _ = return ()
        writeHeaders  = writeGF writeFile [headerCnc,headerAbs]
[…]

### 2011-12-29 testSuites/Parsing.hs

@@ -1,5 +1,6 @@
 module Parsing where
 import System.Timeout
+import System.Process
 import Data.Maybe
 import Data.Map hiding (map,null,filter)
 import Data.Ord
@@ -25,29 +26,51 @@ type Sent = (Id,String)
 -- 40 s, maybe too much
 timeLimit =  30*10^6
 
-run :: String -> Bool -> FilePath -> IO Count
-run file strict out = do  
-  putStrLn "Parsing input file..."
-  sents <- liftM head $ mainF file (return . map getSentence)
-  putStrLn " Creating new dictionary..."
-  (pgf',lang') <- mkDir (words $ concatMap (map toLower . snd) sents) lang pgfFile 
-  putStrLn " Ok\nReading PGF..."
-  pgf <- readPGF pgf' --File
-  putStrLn " Ok\n"
+play :: IO ()
+play = do
+  maps <- mkLexMap
+  pgf <- readPGF pgfFile
+  loop maps pgf
+ where loop maps pgf = do
+            putStrLn "Give a sentence"
+            str <- getLine 
+            (pgf',l') <- mkDir maps (prepare str) langOld pgf
+            pgfNew <- readPGF pgf'
+            let morpho = buildMorpho pgfNew l'
+            tree <- evalStateT (tryparse pgfNew morpho l' ("1",str)) (newCount False "pipfil")
+            case tree of
+                 Nothing -> putStrLn "no parse"
+                 Just t  -> do writeFile "tmptree" $ graphvizParseTree pgfNew l' t
+                               rawSystem "showdot" ["tmptree"]  
+                               putStrLn "hurray"
+            loop maps pgf
+       prepare = words . map toLower
 
--- must compile a new grammar!! BigSwe with tempDict
-  let morpho = buildMorpho pgf lang'
-  c <- execStateT (mapM_ (tryparse pgf morpho lang) sents) (newCount strict out)
-  print c
-  return c
 
-tryparse :: PGF -> Morpho -> Language -> Sent -> CState ()
+--run :: String -> Bool -> FilePath -> IO Count
+--run file strict out = do  
+--  putStrLn "Parsing input file..."
+--  sents <- liftM head $ mainF file (return . map getSentence)
+--  putStrLn " Creating new dictionary..."
[…]

### 2011-12-29 doc/report/report.tex
report: refs, only pics missing

@@
 \usepackage{textcomp}
+%\usepackage{apacite}
 % exempel: 2) Inte vet jag
@@
 %(~~devoted~~) 
-supervisor Ramona Enache for all her help and guidence 
+supervisor Ramona Enache for all her help and guidance 
 %(encouragement)
@@
 Finally, I would also like to thank my friends and family. Special thanks to Dan % Rosén
-for all his support, advice and patience and  -- most importantly -- for being such a good friend. %for having listened to 
+for all his support, advice and patience and  -- most importantly -- for being
+such a good friend. %for having listened to 
 %my monologues about the daily progress. 
@@
 stronger than mildly context-free grammars. The framework has  hence the same expressiveness
-as Tree Adjoining Grammars (Joshi,1975) and Head Grammars (Pollard,1984).
+as Tree Adjoining Grammars \cite{tag} and Head Grammars \cite{hg}.
 
@@
 \end{verbatim}
+\caption{Example of how the GF shell can be used.}\label{fig:shellvp}
 \label{fig:shell}
-\caption{Example of how the GF shell can be used.}\label{fig:shellvp}
 \end{figure}
-\begin{figure}[h]
-\centering
-\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
-\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
-\caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
-\label{fig:gftree1}
-\end{figure}
 
 % nice: test vt -format=pdf -view=evince -nofun
-\begin{wrapfigure}{l}{0.22\textwidth}
-\includegraphics[width=20mm]{gfETree.png}
-\caption{English parse tree}
-\label{pic:gfEtree}
-\end{wrapfigure}
 The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
@@
 that are assigned to the words and phrases. 
+\begin{figure}[h]
+\centering
+\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
+\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
+\caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
+\label{fig:gftree1}
+\end{figure}
+\vspace{10mm}
 % do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
@@
 and the determiner is therefore shown in the parse tree.\\
-
 %>> how parse trees work, not correct complete representation.
@@
 %>> have ambiguity section. 
-\newpage %as long as english picture bråks.
+\newpage
+\begin{wrapfigure}{l}{0.35\textwidth}
+\includegraphics[width=30mm]{gfETree.png}
+\caption{English parse tree}
+\label{pic:gfEtree}
+\end{wrapfigure}
+
 The scope of GF grammars is usually controlled language, a limited language used in
@@
 remain, but can be controlled more easily. %Example of domainspecific word?
+
 The use of controlled language thus gives a possibility of good and reliable translation
@@
 
+\vspace{10mm}
 \subsection{Writing a GF grammar}
@@
 may be thought of as f-structures and c-structures in Lexical Functional
-Grammar (Bresnan and Kaplan, 1982).
+Grammar \cite{lfg}.
 
@@
 
-In the TAG formalism (Joshi,1975), there are projects on getting open source, wide-coverage grammars
+In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
 for English and Korean, but, to our knowledge, not for Swedish.  \\
 
-The ParGram project aims at making wide coverage grammars using the Lexicon Functional Grammar approach
-(Bresnan and Kaplan, 1982).
+The ParGram \cite{pargram} project aims at making wide coverage grammars using
+the Lexical Functional Grammar approach \cite{lfg}.
 The grammars are implemented in parallel in order to coordinate the analyses of
@@
 %>> semantic. Not 1-1.
-%TODO when??
 The basic algorithm for importing Saldo was implemented by Angelov (2008)
@@
 
+\begin{figure}[h]
 \verb-mkV "knyter" "knöt" "knutit"-\\
+\caption{Final output for the verb `knyta'}
+\label{fig:saldoknyt3}
+\end{figure}
 
@@
 tagged as names and numbers. This number may seem very high, but %considering that
-4/5 of the word forms are compounds and when performing the inteded parsing,
+4/5 of the word forms are compounds and when performing the intended parsing,
 these an additional analysis before being looked-up in the lexicon. We
@@
 
-The function \verb-SlashV2a- lets you create a \verb-VPSlash-, a verb phrase missing a
-noun phrase \cite{gazdar}. When we combine the \verb-SlashVP- with the object we
+The function \verb-SlashV2a- lets you create a \verb-VPSlash-. The name
+\verb-SlashV2a- is inspired by categorical grammar and is a short
+hand for \verb-VP \ NP- -- a verb phrase missing a
+noun phrase..  When we combine the \verb-SlashVP- with the object we
 get a complete verb phrase.
@@
 %``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
+%TODO explain, diff from Gazdar
 The \verb-VP- category resembles the Didrichsen's field model.
@@
 \begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=90mm]{tittapa.jpg}}
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=80mm]{tittapa.jpg}}
 \hspace{10mm}
-\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=90mm]{tittapart.pdf}}
+\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=80mm]{tittapart.pdf}}
 \caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
@@
 Like verb phrases, a clause may also be missing an object and then has the type
-\verb-ClSlash-. The \verb-ClSlash- is formed by a \verb-VPSlash- which is 
-given a subject.
+\verb-ClSlash-.
+The \verb-ClSlash- is formed by a \verb-VPSlash- which is given a subject.
 This is a convenient way to form questions, relative clauses
-and topicalized clauses, as in figure \ref{fig:clause2}.\\
+and topicalized clauses, as introduced by \cite{gazdar} (see figure \ref{fig:clause2}).\\
 
@@
 %\ref{gfSwe:parseable}.\\
-
-%TODO
-%>> change dot file to have VP NP in correct order
 
@@
 definite form.
-\enumsentence{\begin{tabular}{llllll}
-*Det     & sitter     & \textbf{den}& \textbf{fågeln}& på       & taket. \\
+\enumsentence{\shortexnt{6}
+{*Det     & sitter     & \textbf{den}& \textbf{fågeln}& på       & taket.}
 %& *Det sitter \textbf{fågeln} på taket.\\
-\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}\\
+{\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}}
 %& ``There are the bird sitting on the roof'' \\
-\end{tabular}}
+}
 Neither is sentence (\ref{sent:dumfagel}) correct.
-\enumsentence{\begin{tabular}{llllll}
-*Det& sitter& \textbf{min}& \textbf{fågel}& på& taket\\
-\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}\\
+\enumsentence{\shortexnt{6}
+{*Det& sitter& \textbf{min}& \textbf{fågel}& på& taket}
+{\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}}
 %``There are my bird sitting %on the roof''\\
-\end{tabular}
 }\label{sent:dumfagel}
@@
 \end{tabular}}\label{ex:ReflAP} \label{ex:ReflAdv}
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. & Hon tyckte om skolan och alla sina elever. & & b. & Han såg sina få böcker och sin penna.\\
-   & She liked the school and all {\sc self's} students. &&&
- He saw {\sc self's} few books and {\sc self's} pencil.\\ 
-\end{tabular}}\label{ex:ReflConj} \label{ex:ReflConj2}
+\eenumsentence{%\begin{tabular}[t]{@{}*{5}{l@{\ }}}
+\item Hon tyckte om skolan och alla sina elever.\\
+     She liked the school and all {\sc self's} students.
+\item Han såg sina få böcker och sin penna.\\
+ He saw {\sc self's} few books and {\sc self's} pencil.
+}\label{ex:ReflConj} \label{ex:ReflConj2}
 Reflexive pronouns can not be used in subject noun phrases of finite sentences,
@@
 The third person reflexives (`sig',`sin') requires a third person antecedent (see \ref{ex:refljag}).
-Furthermore, the antencedent must be within the same finite sentence as the reflexive
+Furthermore, the antecedent must be within the same finite sentence as the reflexive
 pronoun, see (\ref{sent:sesig}).
@@
 subject was changed to 2nd person, they  would need to change too.
-\enumsentence{\begin{tabular}{llll}
+\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
 a. &\textbf{Du} är längre än \textbf{din} kompis.& b. &*Du är här oftare än \textbf{sin} kompis.\\
@@
 \subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian use two auxilirary verbs for future tense
+Swedish and Norwegian use two auxiliary verbs for future tense
 \cite[p. 246]{H&H}:
@@
 and `till och med' (\emph{`even'}).
-\enumsentence{\shortex{9}
-{a. & Han & bara & log  && &b. & Hon till och med visslar}
-{&\emph{he} & \emph{only} &\emph{ smiled} &  & & &  }
-{``He just smiled"\hspace{8mm} ``She even whistles"}}
+\enumsentence{\shortex{11}
+{a. & Han & bara & log  && &b. & Hon& till och med& visslar}
+{&\emph{he} & \emph{only} &\emph{ smiled} & & & &\emph{she} &\emph{even} & \emph{whistles}  }
+{~~``He just smiled"\hspace{8mm} ``She even whistles"}}
 Focusing adverbs are accepted by the new grammar implementation, where they have
@@
 
-The focusing adverbs are usually not combined with modal verbs, the copula or temporal auxiliries. 
+The focusing adverbs are usually not combined with modal verbs, the copula or temporal auxiliaries. 
 % since the modal verbs rarely are focused.
@@
  in sentence (\ref{ex:QuantAdj}). 
-\eenumsentence{\item[a.]Sådana \label{ex:QuantDet}katter vill jag ha
-\item[b.]Han är \label{ex:QuantAdj}sådan\toplabel{sent:Quant}}
+\eenumsentence{
+\item\shortex{5}
+{Sådana& \label{ex:QuantDet}katter& vill& jag& ha}
+{\emph{such}& \label{ex:QuantDet}\emph{cats}& \emph{want}& \emph{I}& \emph{have}}
+{`I want such cats'}
+\item\shortex{3}
+{Han& är& \label{ex:QuantAdj}sådan\toplabel{sent:Quant}}
+{\emph{he}& \emph{is}& \label{ex:QuantAdj}\emph{such}\toplabel{sent:Quant}}
+{`He's like that'}
+}
 
@@
 (\ref{pic:gftree}) can be extracted. We need to know that \emph{katten} is a noun
-used in definite form (\verb-DetArt-), singular (\verb-NumSg-). Even though the
+used in definite form (\verb-DefArt-), singular (\verb-NumSg-). Even though the
 notation is translated, the original analyse from Talbanken
@@
 \begin{figure}[h]
-\hspace{-10mm}
+\hspace{-30mm}
 \subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
-\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=60mm]{gfTree.pdf}}
+\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=90mm]{gfTree.pdf}}
 \subfloat[The corresponding GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
@@
 sentence.
+\begin{wrapfigure}{l}{0.52\textwidth}
+\includegraphics[width=80mm]{metatree.pdf}
+\caption{Abstract tree with meta}
+\label{pic:gMetatree}
+\end{wrapfigure}
+
 A similar issue arises from the use of verb particles. A verb with a
@@
 
-\begin{wrapfigure}{l}{0.22\textwidth}
-\includegraphics[width=20mm]{metatree.pdf}
-\caption{Meta tree}
-\label{pic:gMetatree}
-\end{wrapfigure}
 %In Talbanken, verbs and particles are treated as separate units,
@@
 the mapping cannot depend on the valencies given in GF only, 
-since the distinction between different types of verbs is not always clear in Swedish;
+\newpage
+
+\noindent since the distinction between different types of verbs is not always clear in Swedish;
 most transitive verbs can also be used as intransitive. Both sentences in
@@
 Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
-which groups an inifinitival verb and its object together into a \verb|VP|.
+which groups an infinitival verb and its object together into a \verb|VP|.
 This information can however be extracted from the flat version, and the results
@@
 
-Our final goal is to parse Talbanken, but considering that no syntactic theory is yet capable
+Our final goal is to be able to parse unrestricted text, but considering that
+no syntactic theory is yet capable
 of covering a whole natural language, we are satisfied with %theconsider our grammar 
@@
 the grammar implementation of an important fragment of Swedish.
-Following methods for making the parser robust as described in section \ref{sec:future}
+We have therefor focused on constructions that are frequent in Talbanken, 
+and following methods for making the parser robust as described in section \ref{sec:future}
 we still believe that it is possible to achieve the goal. 
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{apalikeurl}
 \bibliography{masterbib}

### 2011-12-31 doc/report/report.tex
last update to report before aarne

@@
 increased. There is research on how to make more use of the dependent
-types, for reasoning by using ontologies \cite{ramona} or generating natural
+types, for reasoning by using ontologies \cite{ontologies2} or generating natural
 language via Montague semantics \cite{montague}.
@@
 %TODO Sats adverbials?
-Personal pronouns are allowed to precede negation and some other adverbials --\textit{satsadverb} --
+Personal pronouns are allowed to precede negation and some other adverbials of place --\textit{satsadverbial} --
 in main clauses without auxiliary verbs.
 
-\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
-a. & *Vi & ser&  varandra&  inte&~~ & b.& * Jag &ser &huset &inte\\
-   & \emph{~we} & \emph{see}&  \emph{each other}&  \emph{not} && b.& \emph{~I}&\emph{see} &\emph{the house} &\emph{not}\\
+\enumsentence{\begin{tabular}[t]{@{}*{12}{l@{\ }}}
+a. & *Vi & har&  honom&  inte&sett&~~ & b.& * Jag &ser &huset &inte\\
+   & \emph{~we} & \emph{have}&  \emph{him}&  \emph{not} &  \emph{seen} && b.& \emph{~I}&\emph{see} &\emph{the house} &\emph{not}\\
 \end{tabular}}

### 2012-01-02 doc/report/report.tex
report: fix comments

@@
 %ska innehålla vettiga sökord
-\title{Towards a wide-coverage grammar for Swedish using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Towards a Wide-Coverage Grammar for Swedish Using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 %generalized context free grammar (GCFG) it is hence
-stronger than mildly context-free grammars. The framework has  hence the same expressiveness
-as Tree Adjoining Grammars \cite{tag} and Head Grammars \cite{hg}.
+stronger than mildly context-free grammars. 
+The framework's %has hence the same
+expressiveness is hence stronger than 
+Tree Adjoining Grammars \cite{tag} and Head Grammars \cite{hg}, and shown
+equivalent to Parallel Multiple Context-Free Grammar \cite{pmcfg} by \cite{peter}.
 

### 2012-01-02 doc/report/report.tex
report: write about generation

@@
 or missing phrase will cause the parsing of the whole sentence to fail. 
+Besides parsing, the grammar may well be used for language generation,
+which is fast even with an extensive lexicon.
+%TODO make nice!
+Although it cannot not been verified, we believe that the majority of the
+sentences generated are grammatically correct in a syntactical point of view.
+Without any semantical knowledge, nonsense phrases cannot be avoided in
+random generation. However, an abstract tree has a meaningful interpretation,
+the linearization should be correct. There are some cases when the correctness
+of the output has been ignored in order to increase the expressivity, such
+as ComplBareVV, SupCl. These functions should be excluded or enhanced when doing
+careful generation...
+``övriga datorer idag som ska vara inte skulle ha målats rödare än johan"
+PredVP (DetCN Subject (DetSub oovriga_Det) (RelCN (AdvCN (UseN computer_N) (AdvSub today_Adv)) (UseRCl (TTAnt TFut ASimul) PPos (RelVP IdRP UseCopula)))) (PassVP (SlashV2A paint_V2A (ComparA Object red_A (UsePN Object johan_PN))))
+
+%The use of the NONEXIST pattern, as described in \ref{sec:Formal}. When using
+%the grammar for generation, an extra
 

### 2012-01-03 doc/report/report.tex
new final report

--- 2012-01-02 doc/report/report.tex
@@
 nouns in indefinite form, whenever the noun phrase is followed by a restrictive
-relative clause.
+relative clause \cite[\textsection 329]{H&H}
 Talbanken contained several examples where the modified noun is in the 
 indefinite form as in (\ref{sent:uppfatt}).
-%TODO ref to H&H
 \enumsentence{\shortex{7}
@@
 the grammar rules are simply too strict for accepting many Talbanken sentences.
-\enumsentence{\shortexm{13}
-{a. &Han& sitter& och &läser.}
-{&\emph{he}& \emph{sits}& \emph{and} &\emph{reads}}{}
-{b. &Han& sitter& och& läser& en& bok.}
-{&\emph{he}& \emph{sits}& \emph{and}& \emph{reads}& \emph{a}& \emph{book}}{}}
+\enumsentence{\begin{tabular}[t]{@{}*{14}{l@{\ }}}
+a. &Han& sitter& och &läser.&& & b. &Han& sitter& och& läser& en& bok.\\
+&\emph{he}& \emph{sits}& \emph{and} &\emph{reads} &&&
+&\emph{he}& \emph{sits}& \emph{and}& \emph{reads}& \emph{a}& \emph{book}\\
+\end{tabular}}
 \label{ex:mappTransitive}
@@
 The grammar has been extended and enhanced, and its current status is
-a specialized extension of the resource grammar. When it comes to parsing,
-however, we do not get far without robustness.
+a specialized extension of the resource grammar.
+Besides parsing, the grammar may well be used for language generation,
+which is fast even with an extensive lexicon.
+%TODO make nice!
+Although it cannot not been verified, we believe that the majority of the
+sentences generated are grammatically correct in a syntactical point of view.
+Without any semantical knowledge, nonsense phrases cannot be avoided in
+random generation. However, given that the abstract tree has a meaningful
+interpretation,
+the linearization should be correct. There are some cases when the correctness
+of the output has been put aside in order to increase the expressivity, such
+as \verb-ComplBareVV-, which use a verb-complement verb without the inifite mark, 
+as in sentence (\ref{sent:complVV1}b).
+\eenumsentence{\item\shortex{5}
+{Jag& börjar&\textbf{att}& bli& hungrig}
+{I& begin& \textbf{to}&become& hungry}
+{I'm getting hungry}
+\item
+\shortex{4}
+{Jag& börjar& bli& hungrig}
+{I& begin& become& hungry}
+{I'm getting hungry}}\label{sent:complVV1}
+Since there is no information about which verbs that allows the inifinite mark
+to be left out, it will also allow the more questionable sentence (\ref{sent:complVV2}b).
+\eenumsentence{
+\item \shortex{5}
+{Jag & vill & \textbf{att}& du & går} 
+{I & want & \textbf{to} &you & leave}
+{I want you to leave}
+\item \shortexnt{4}
+{Jag & vill & du & går} 
+{I & want &you & leave}}\label{sent:complVV2}
+%ComplBareVV, SupCl
+For random generation, the sentences generated may be next to impossible
+to interpret without looking at the abstract tree, since they may contain
+too many attributes, relative clauses etc. 
+%\enumsentence{Kommer övriga datorer idag som ska vara inte att ha målats?}
+%\label{sent:what}
+%Sentence (\ref{sent:what})
+%PredVP (DetCN Subject (DetSub oovriga_Det) (RelCN (AdvCN (UseN computer_N) (AdvSub today_Adv)) (UseRCl (TTAnt TFut ASimul) PPos (RelVP IdRP UseCopula)))) (PassVP (SlashV2A paint_V2A (ComparA Object red_A (UsePN Object johan_PN))))
+%The use of the NONEXIST pattern, as described in \ref{sec:Formal}. When using
+%the grammar for generation, an extra
+
+When it comes to parsing, we do not get far without robustness.
 The grammar in itself is by no means robust, and just one 

### 2012-01-03 doc/report/report.tex
slides

@@
 %TODO Sats adverbials?
-Personal pronouns are allowed to precede negation and some other adverbials of place --\textit{satsadverbial} --
+Personal pronouns are allowed to precede negation and some other sentence adverbials --\textit{satsadverbial} --
 in main clauses without auxiliary verbs.

### 2012-01-04 testSuites/MkLex.hs
files from master

@@ -1,0 +1,87 @@
+module MkLex where
+import PGF
+import Control.Monad
+import System.Process
+import System.Exit
+import qualified Data.HashMap as M
+
+type LexMap = (Lex,Lex)
+type Lex    = M.Map String String
+
+mkDir :: LexMap -> [String] -> Language -> PGF -> IO (Maybe (FilePath,Language))
+mkDir lexs ws lang pgf = do
+--   putStrLn "Reading pgf..."
+--   pgf <- readPGF bigPGF  
+   let morpho = buildMorpho pgf lang
+       allLemmas = map fst $ concatMap (lookupMorpho morpho) ws
+   putStrLn "Extracting new lexicon..."
+   (cnc,abs) <- extractLemmas lexs allLemmas
+   putStrLn "Compiling new grammar..."
+   (ex,res,err) <- readProcessWithExitCode "gf" ["--make",bigGF] []
+   case ex of
+        ExitSuccess   -> do putStrLn "Done"
+                            return $ Just (newPGF,newLang)
+        ExitFailure _ -> do putStrLn "PGF error"
+                            putStrLn err
+                            return Nothing
+
+-- do not add fun or lin if no words extracted!!
+extractLemmas :: LexMap -> [Lemma] -> IO (FilePath,FilePath)
+extractLemmas (cnc,abs) ws = do
+  writeHeaders
+  mapM_ (extract newLexCnc cnc) ws
+  mapM_ (extract newLexAbs abs) ws
+  writeEnd
+  return (newLexCnc,newLexAbs)
+  where
+       extract newFile lex w
+               | Just a <- M.lookup (showCId w) lex =
+                         liftM Just $ appendFile newFile (a++"\n")
+       extract _ _ _ = return Nothing
+       writeHeaders  = writeGF writeFile [headerCnc,headerAbs]
+       writeEnd      = writeGF appendFile (repeat "}\n")
+       writeGF f = zipWithM_ f [newLexCnc,newLexAbs]
+
+       headerCnc = "--# -path=.:abstract:alltenses:swedish:common:scandinavian\n"
+                   ++"concrete TestLex of TestLexAbs = CatSwe **\n"
+                   ++"open  Prelude, CommonScand, ParadigmsSwe, IrregSwe in {\n"
+                   ++" flags optimize=values ; coding=utf8 ;"
+                   ++"lin\n"
+       headerAbs = "--# -path=.:abstract:prelude:alltenses\n"
[…]

### 2012-01-04 doc/report/report.tex
report: ramones comments

@@
 %ska innehålla vettiga sökord
+%TODO page numbers!!
 \title{Towards a Wide-Coverage Grammar for Swedish Using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
@@
 \abstract{
-This thesis describes the work to achieve a covering Swedish grammar implemented
+The thesis describes the work to ~~achieve a covering Swedish~~ grammar implemented
+%TODO first sentence should be more, no understatements!
 in GF -- to model natural language by functional programming. 
@@
 To reach this goal, problems of computational as well as linguistic nature had to be solved.
-\\The thesis explains the methods and technologies used and further inspects the possibilities
+\\We present the methods and technologies used and further inspects the possibilities
 of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
@@
 We further give an example of the advantage of using dependent types when
-describing grammar and syntax, in this case reflexive pronouns.
+describing grammar and syntax, in this case for dealing with reflexive pronouns.
+%the linguistic phenomena
 
@@
 \newpage
+%\pagenumberingstyle{plain}
 \setcounter{page}{1} %reset the page counter
@@
 Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism,
-developed for multilinguality. It is based on Martin-Löf type theory which allows
+with facilities for multilinguality. It is based on Martin-Löf type theory which allows
 reasoning within the programming language.
@@
 a parser or grammar is not freely available does not only restrict its
-usage but also its possibilities of being developed and our
+usage but also its possibilities of being further developed and our
 goal is
@@
 resources.
-We start from a GF resource grammar and thereby get
+%TODO verify this
+We start from a GF resource grammar, consisting of an abstract and a concrete syntax
+file defining a set of rules describing morpology and syntax.
+From now on, this is what we mean by grammar, as
+opposed to grammar in the traditional linguistic sense. 
+From GF, we get %and thereby get
 %The usage of GF allows us to start from
-a well-defined system for describing the grammar, as well as a strong connection to
+a well-defined system for describing the language, as well as a strong connection to
 more than 20 other languages implemented in the framework. Further, we use
@@
 using GF for parsing unrestricted Swedish. 
+%TODO LREC ref?
 In previous work\footnote{??}, a start was made to extend the Swedish GF grammar
@@
 Tree Adjoining Grammars \cite{tag} and Head Grammars \cite{hg}, and shown
-equivalent to Parallel Multiple Context-Free Grammar \cite{pmcfg} by \cite{peter}.
+equivalent to Parallel Multiple Context-Free Grammar \cite{pmcfg} in \cite{peter}.
 
@@
 
-The scope of GF grammars is usually controlled language, a limited language used in
-a restricted domain. By
+%TODO read again
+The scope of GF grammars has so far been controlled language, a limited
+language used in a restricted domain. By
 restricting the coverage, the number of ambiguities
@@
 \label{sec:saldo}
-A good parser needs a good lexicon. We have used of Saldo \cite{saldo}, a
+A good parser needs a good lexicon. We have used Saldo \cite{saldo}, a
 large electronic lexicon developed and maintained at Gothenburg University. It is
@@
 separated into different entries.
-%If one word has different meanings, there will be separate entries for each of them. 
+%TODO change word? uppskatta, känna, få
 The word ``växa" (``grow") for example, has two meanings: \emph{getting
@@
 %There has been much research about parsing Swedish. INTRO!!
-Many years of research has lead to many interesting language
+Many years of research have lead to many interesting language
 technology tools for Swedish.
@@
 while the one generated in GF contains some that are not in Saldo (e.g. ``vuxits").
-As GF concerns about grammar only, and not semantics, and  as the GF tables are automatically
+As GF concerns about syntax only, and not semantics, and  as the GF tables are automatically
 generated, they always contain all word forms, although some forms may never be
@@
 are not analysed the same way in GF (see section \ref{sec:swegf}).
-They can thus not simply be extracted -- we first need to find their correct GF category.
-Some experiments on finding
-the right GF-category for those have been done using Talbanken, see p. \pageref{sec:gf.quant}.
+Befor adding them to our lexicon, we need to do more analysing to find their correct GF-category.
+%TODO read again
+%They can thus not simply be extracted -- we first need to find their correct GF category.
+Some experiments on finding the category
+%the right GF-category for those
+have been done using Talbanken, see p. \pageref{sec:gf.quant}.
 
@@
 to adapt it to cover
-constructions used in Talbanken. As a grammar implementation never can be
+constructions used in Talbanken. As a grammar implementation can never be
 expected to give
@@
 \includegraphics[height=130mm]{categories.png}
-\caption{The types of the GF resource grammars.}
+\caption{The types of the GF resource grammars. }
+%TODO have footnote? where?
+%\footnote{Available at http://www.grammaticalframework.org/lib/doc/synopsis.html}
 \label{fig:gfpic}
@@
 and \verb-N'- to distinguish nouns with respectively without number information.
+%TODO read again
+The category \verb-N'- was a simplification and is not used in the resource grammar. and is not used in the resource grammar.
 The full analysis includes a distinction between the categories \verb-N- -- simple nouns --
@@
 
-The GF category for pronouns \verb-PN- only contains personal pronouns.
-Many words which are considered to be pronouns in other analyses are hence
-classified differently in GF.
+%TODO review, verify
+Due syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
+contains personal pronouns.
+Many words which are considered to be pronouns in other analyses, such as The Swedish
+Academy Grammar \cite{SAG}, only exist in 3rd person and are 
+classified differently in GF, usually as determiners or quantifiers.
 %Only personal pronouns are of the type \verb-Pron-.
+ %TODO
+ % Adjectives? About weak - strong - indefinite..
 
@@
 An important area in a Swedish grammar is the treatment of the reflexive pronouns and
+%TODO fix page ref
 the reflexive possessive pronouns as described on p. \pageref{swe:refl}.
@@
 (\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
+In the standard GF analysis, which is preformed bottom-up starting from the pos-tags, 
+all information about semantical roles are lost. There is no way
+of knowing whether a \verb-NP- acts as subject or object. For this reason,
+the formalization of reflexive pronouns required the use of a different analysis.
 %and used as genitives (sentence \ref{ex:ReflGen}) as such.
@@
 %\end{figure}
+
 If we ignore the second requirement, we might come up with a solution where
@@
 \end{verbatim}
-
-The solution hence combines the-part-of-speech driven analyse normally preformed by
-GF with a part-of-sentence analysis.
-
 The types for adverbial and adjectival phrases are also turned into dependent
@@
 
+We hence combine the-part-of-speech driven analyse normally preformed by
+GF with a part-of-sentence analysis, where the dependent types gives the semantical
+information we were missing.
+
 The use of dependent types separates our grammar from the other resource grammars. Such separation
@@
 \item The new grammar can be made compatible with the old implementation.
+%TODO verify
 \item The possibility for noun phrases to agree with an antecedent 
-      will be useful in other languages. For example, the Slavic languages
+      will be useful in other languages. For example, the Slavic and Romance languages
       have similar reflexive pronouns.
@@
 
-The use of dependent types is interesting for other problems.
-The rules for reciprocal pronouns could be formalized using the same idea. As reflexive pronouns, reciprocals
+Dependent types are also interesting for other problems.
+The rules for reciprocal pronouns could be formalized using the same idea. As
+reflexive pronouns, reciprocals (sentence \ref{sent:varandra})
 must not be used as subjects and furthermore they may only be used when the
 number of the antecedent is plural.
+\enumsentence{
+\shortex{3}
+{De &ser& varandra}
+{\emph{they} &\emph{see}& \emph{each other}}
+{`They see each other'}}\label{sent:varandra}
+%\item \shortexnt{3}
+%{*Han &ser& varandra}
+%{~he &sees& each other}}
 %
@@
 
+\newpage
 \subsection{Miscellaneous}
@@
 \begin{itemize}
-\item \verb-Quantifiers- does not have any possessive
+\item \verb-Quantifiers- do not have any possessive
 forms, so when a quantifier is turned into
@@
 \subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
-\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=90mm]{gfTree.pdf}}
+\hspace{-10mm}
+\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=100mm]{gfTree.pdf}}
 \subfloat[The corresponding GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
@@
 
-A mapping between GF and Penn Treebank has earlier been conducted \cite{gfMech}.
+A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
 The percentage of restored nodes from Peen Treebank is higher than our results.
-%Our results are lower than the results of the mapping of Penn Treebank.
+%TODO review
 The reason for this may be the fact that English is syntactically less complicated
-than Swedish. Also, the Penn Treebank contains a lower number of tags, 82 compared
+than Swedish.
+Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
+where idiomatic expressions are more likely and the language presumably less strict than
+in Wall Street Journal.
+Also, the Penn Treebank contains a lower number of tags, 82 compared
 to more than 300 in Talbanken. Even if the tags describing the same word
+%TODO review
 class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
-Thus the number of combinations are increased, and so is the number of
-rules that would be needed to cover all constructions.
+With more tags, we get more information, but as the number 
+%Thus the number of
+increase, so does the amount of work of finding the correct translation 
+for each combination of tags and writing rules that cover all constructions. 
+%iand so is the number of
+%rules that would be needed to cover all constructions.
 
@@
 interpretation,
+%TODO move to Misc?
 the linearization should be correct. There are some cases when the correctness
@@
 {Jag& börjar&\textbf{att}& bli& hungrig}
-{I& begin& \textbf{to}&become& hungry}
+{\emph{I}& \emph{begin}& \textbf{\emph{to}}&\emph{become}& \emph{hungry}}
 {I'm getting hungry}
@@
 {Jag& börjar& bli& hungrig}
-{I& begin& become& hungry}
+{\emph{I}& \emph{begin}& \emph{become}& \emph{hungry}}
 {I'm getting hungry}}\label{sent:complVV1}
@@
 unexpected punctuation mark, unknown word
-or missing phrase will cause the parsing of the whole sentence to fail.
+or ellipsis will cause the parsing of the whole sentence to fail.
 
@@
 the ability to scale up, which has so far proved to be good.
-I further believe that the presence of a Swedish expert has
+I further believe that the presence of a expert in Swedish, professor Elisabet
+Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl},
+has
 increased the standard substantially.\\
@@
 \end{tabular}}
-%TODO Sats adverbials?
 Personal pronouns are allowed to precede negation and some other sentence adverbials --\textit{satsadverbial} --
@@
 will not fail but give a partial parse tree with the analysis for the rest of
-the sentence.
+the sentence \cite{gfMech}.
 The output given to the user will be more valuable, since partial parse
@@
 The parser should be able to identify names of persons, companies, countries etc.
-This can be done by the Haskell program executing the parsing.
+This can be done by the Haskell program executing the parsing \cite{patent}.
 
@@
 evaluation and disambiguation. %To abstract:  by combining resources.
-We started from a well-defined system for describing grammar as we built
+We started from a well-defined system for describing grammars as we built
 the grammar from the GF resource grammar.
@@
 no syntactic theory is yet capable
-of covering a whole natural language, we are satisfied with %theconsider our grammar 
+of wholy covering a natural language, we are satisfied with %theconsider our grammar 
 %when used without a robust layer 
 the grammar implementation of an important fragment of Swedish.
-We have therefor focused on constructions that are frequent in Talbanken, 
+We have therefore focused on constructions that are frequent in Talbanken, 
 and following methods for making the parser robust as described in section \ref{sec:future}
@@
 
+%TODO more here
 Being a natural language processing-project, it will probably never be
@@
 
+%TODO Espana -> ~
 \bibliographystyle{apalikeurl}

### 2012-01-06 doc/report/report.tex
report

@@
 We start from a GF resource grammar, consisting of an abstract and a concrete syntax
-file defining a set of rules describing morpology and syntax.
+file defining a set of rules describing morphology and syntax.
 From now on, this is what we mean by grammar, as
@@
 sentence and visualize the resulting parse tree. The second will visualize the 
-abstract syntax tree. The trees will be showed in the pgf-format by the program evince.
+abstract syntax tree. The trees will be showed in the pdf-format by the program evince.
 \begin{figure}[h]
@@
 are not analysed the same way in GF (see section \ref{sec:swegf}).
-Befor adding them to our lexicon, we need to do more analysing to find their correct GF-category.
+Before adding them to our lexicon, we need to do more analysing to find their correct GF-category.
 %TODO read again
@@
 In the standard GF analysis, which is preformed bottom-up starting from the pos-tags, 
-all information about semantical roles are lost. There is no way
+all information about semantic roles are lost. There is no way
 of knowing whether a \verb-NP- acts as subject or object. For this reason,
@@
 We hence combine the-part-of-speech driven analyse normally preformed by
-GF with a part-of-sentence analysis, where the dependent types gives the semantical
+GF with a part-of-sentence analysis, where the dependent types gives the semantic
 information we were missing.
@@
 sentences generated are grammatically correct in a syntactical point of view.
-Without any semantical knowledge, nonsense phrases cannot be avoided in
+Without any semantic knowledge, nonsense phrases cannot be avoided in
 random generation. However, given that the abstract tree has a meaningful
@@
 of the output has been put aside in order to increase the expressivity, such
-as \verb-ComplBareVV-, which use a verb-complement verb without the inifite mark, 
+as \verb-ComplBareVV-, which use a verb-complement verb without the infinitive mark, 
 as in sentence (\ref{sent:complVV1}b).
@@
 {I'm getting hungry}}\label{sent:complVV1}
-Since there is no information about which verbs that allows the inifinite mark
+Since there is no information about which verbs that allows the infinitive mark
 to be left out, it will also allow the more questionable sentence (\ref{sent:complVV2}b).
@@
 \end{tabular}}
-Personal pronouns are allowed to precede negation and some other sentence adverbials --\textit{satsadverbial} --
+Personal pronouns are allowed to precede negation and some other sentence adverbials % --\textit{satsadverbial} --
 in main clauses without auxiliary verbs.
@@
 no syntactic theory is yet capable
-of wholy covering a natural language, we are satisfied with %theconsider our grammar 
+of wholly covering a natural language, we are satisfied with %theconsider our grammar 
 %when used without a robust layer 

### 2012-01-06 testSuites/Parsing.hs
presentation

@@ -1,5 +1,10 @@
 module Parsing where
 import System.Timeout
+import System.Process
+import System.Console.ANSI hiding (Color)
+import System.IO
+import System.Exit
+import qualified Data.ByteString.Lazy as BS
 import Data.Maybe
 import Data.Map hiding (map,null,filter)
 import Data.Ord
@@ -10,8 +15,10 @@ import Control.Concurrent
 import Control.Arrow
 import Control.Applicative
 import Control.Monad.State
+import Control.Exception
 import PGF
 import XMLHelp hiding (words,parse)
+import MkLex
 
 -- use with tee -a outfile
 data Parsed = Pars {t :: Tree, s :: String, n :: Int}
@@ -24,43 +31,108 @@ type Sent = (Id,String)
 -- 40 s, maybe too much
 timeLimit =  30*10^6
 
-run :: String -> Bool -> FilePath -> IO Count
-run file strict out = do  
-  putStr "Parsing input file..."
-  sents <- mainF file (return . map getSentence)
-  putStr " Ok\nReading PGF..."
+play :: IO ()
+play = do
+  (maps,pgf) <- play'
+  loop maps pgf
+ where loop maps pgf = do
+            putStrLn "Give a sentence"
+            str <- input 
+            tree <- processparse str pgf langOld maps
+            case tree of
+                 Nothing -> putStrLn "no parse"
+                 Just t  -> putStrLn "hurray"
+            loop maps pgf
+
+play' = do 
+  maps <- mkLexMap
   pgf <- readPGF pgfFile
-  putStrLn " Ok\n"
-  --absMap <- readAbbs 
-  let morpho = buildMorpho pgf lang
-  c <- execStateT (mapM_ (tryparse pgf morpho) (concat sents)) (newCount strict out)
-  print c
-  return c
+  return (maps,pgf)
 
-tryparse :: PGF -> Morpho -> Sent -> CState ()
-tryparse pgf morpho sents = do 
+processparse :: String -> PGF -> Language -> LexMap -> IO (Maybe FilePath)
+processparse str pgf langOld maps = do
+  (pgf',l') <- mkDir maps (prepare str) langOld pgf
+  pgfNew <- readPGF pgf'
+  let morpho = buildMorpho pgfNew l'
+  tree <- evalStateT (tryparse pgfNew morpho l' ("1",str)) (newCount False "pipfil")
+  maybe (return Nothing) (createFile pgf) tree
[…]

### 2012-01-07 doc/report/report.tex
small bugs fixed all over

@@
 \abstract{
-The thesis describes the work to ~~achieve a covering Swedish~~ grammar implemented
+The thesis describes the work formalising an important fragment of Swedish
+in the GF formalism and to compare and evaluate the possibilities of using GF
+for modelling large-scale grammars. Grammatical constructions specific to and
+frequent in Swedish have been investigated and formalised in the functional
+programming language.
+%~~achieve a covering Swedish~~ grammar implemented
 %TODO first sentence should be more, no understatements!
-in GF -- to model natural language by functional programming. 
+%in GF -- to model natural language by functional programming. 
 The idea has been to combine existing language technology resources with
@@
 resources.
-%TODO verify this
 We start from a GF resource grammar, consisting of an abstract and a concrete syntax
 file defining a set of rules describing morphology and syntax.
+%TODO verify this, check all occurences of grammar
 From now on, this is what we mean by grammar, as
@@
 %The usage of GF allows us to start from
-a well-defined system for describing the language, as well as a strong connection to
+a well-defined system for describing language, as well as a strong connection to
 more than 20 other languages implemented in the framework. Further, we use
@@
 %>>The grammar defines a parser, which accepts all constructions defined by the grammar.
+
+\newpage % until it looks nicer
 
@@
 
-\newpage % until it looks nicer
 \section{Outline}
@@
 \begin{verbatim}
-ta_med_V3 = dirV3 (reflV (partV (take_V "med")) (mkPrep "till") ;
+ta_med_V3 = dirV3 (reflV (partV (take_V "med"))) (mkPrep "till") ;
 \end{verbatim}
@@
 \enumsentence{\shortex{6}
-{Det & äpplet & vill & jag & inte & ha}
+{Det där & äpplet & vill & jag & inte & ha}
 {that & apple & want & I & not &have}
@@
 {\sc self's} gloves were left on the train.}\label{ex:ReflSubj}
-\enumsentence{*Hon och sin kompis läser en bok.\\
-She and {\sc self's} friend are reading a book. }
+\enumsentence{*Han och sin kompis läser en bok.\\
+He and {\sc self's} friend are reading a book. }
 \enumsentence{*Jag ger sina pengar till sina syskon.\\
@@
 \enumsentence{
-Hon sover \textbf{som} är bra ~ \textrightarrow ~ Hon sover \textbf{vilket} är bra\\
+Hon sover \textbf{som} är bra ~ \textrightarrow ~ Hon sover, \textbf{vilket} är bra\\
 ``She sleeps, which is good"}
@@
 \item a large-scale GF lexicon and a program to redo the importation when needed
-\item a comparison between GF and another another annotation
+\item a comparison between GF and another annotation
 \item a deeper testing of the Swedish resource grammar and an estimation
@@
 \item \shortex{5}
-{Jag & vill & \textbf{att}& du & går} 
-{I & want & \textbf{to} &you & leave}
-{I want you to leave}
+{Jag & låter bli & \textbf{att}& titta} 
+{I & let & be &\textbf{to} &watch}
+{I don't watch}
 \item \shortexnt{4}
-{Jag & vill & du & går} 
-{I & want &you & leave}}\label{sent:complVV2}
+{*Jag & låter & bli & titta} 
+{I & leave &be & watch}}\label{sent:complVV2}
 %ComplBareVV, SupCl
@@
 evaluation and disambiguation. %To abstract:  by combining resources.
-We started from a well-defined system for describing grammars as we built
-the grammar from the GF resource grammar.
+By starting from the GF resource grammar, we got a well-defined system for
+describing grammars. 
+%as we built the grammar from .
 
@@
 %when used without a robust layer 
-the grammar implementation of an important fragment of Swedish.
-We have therefore focused on constructions that are frequent in Talbanken, 
-and following methods for making the parser robust as described in section \ref{sec:future}
+the grammar implementation of an important fragment of Swedish and have
+focused on constructions that are frequent in Talbanken. 
+Being a natural language processing-project, it will probably never be
+entirely complete, but by combining the rule-based parser with statistical
+techniques, 
+%By following the methods for making the parser robust 
+as described in section \ref{sec:future},
 we still believe that it is possible to achieve the goal. 
@@
 %TODO more here
-Being a natural language processing-project, it will probably never be
-entirely complete. Even so, we believe  
+%. Even so, we believe  
 %Even though the grammar will not be totally covering, we belive
-that the work done is both interesting and useful.
+%that the work done is both interesting and useful.
 All parts of the project are open-source and may thus be used in other applications.

### 2012-01-07 testSuites/ParseLex.hs

@@ -121,9 +121,9 @@ langOld  = fromJust $ readLanguage "BigSwe" --"BigSwe" --"BigNewSwe"
 pgfFile =  "../gf/Big.pgf" -- "../gf/BigTest.pgf" -- "../gf/Big.pgf" -- "BigNew.pgf"
 outFile = "testisNP.txt"
 
-pgfBackUp = "../gf/Big.pgf"
+pgfBackUp = "../gf/BigTest.pgf"
 langBackUp :: Language
-langBackUp = read "BigSwe"
+langBackUp = read "BigTestSwe"
 
 pgfDict = "DictSweAbs.pgf"
 langDict :: Language

### 2012-01-07 doc/report/report.tex
report, pres update

@@
 \chapter{Introduction}
+\thispagestyle{plain}
 % more intro to gf. dependently typed...high-level formalism
@@
 %>> enhance gf...
+\protect\thispagestyle{plain}
 The purpose of this project has been to prepare the GF grammar for parsing of 
@@
 abstract syntax tree. The trees will be showed in the pdf-format by the program evince.
+\thispagestyle{plain}
 \begin{figure}[h]
@@
 \label{sec:swedish}
-Swedish \cite[Inl. \textsection 3]{SAG}is a North-Germanic language,
+Swedish \cite[Inl. \textsection 3]{SAG} is a North-Germanic language,
 closely related to Norwegian and Danish. The languages share most of their
@@
 
-\subsection*{Post-nominal articles} 
+\subsection{Post-nominal articles} 
 \label{sec:swedishnoun}
@@
 
-\subsection*{Verb second}
+\subsection{Verb second}
 Swedish is a verb-second language \cite[p.116]{gunlog}: the
@@
 
-\subsection*{Passive voice}
+\subsection{Passive voice}
 There are two ways of forming passive verb phrases in Swedish: the 
@@
 
-\subsection*{Impersonal constructions}
+\subsection{Impersonal constructions}
 Constructions with `det är/var' (\emph{`it is/was'}) \\are very common in Swedish
@@
 
-\subsection*{Reflexive pronouns}
+\subsection{Reflexive pronouns}
 \label{swe:refl}
@@
 \begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=80mm]{tittapa.jpg}}
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=72mm]{tittapa.jpg}}
 \hspace{10mm}
-\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=80mm]{tittapart.pdf}}
+\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=72mm]{tittapart.pdf}}
 \caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
@@
 An important area in a Swedish grammar is the treatment of the reflexive pronouns and
-%TODO fix page ref
-the reflexive possessive pronouns as described on p. \pageref{swe:refl}.
+the reflexive possessive pronouns as described in section \ref{swe:refl} 
 The reflexives require an antecedent with which they agree in
@@
 \subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian use two auxiliary verbs for future tense
+Swedish and Norwegian use two auxiliary verbs for future tense \\
 \cite[p. 246]{H&H}:
@@
 
-\newpage
 \subsection{Miscellaneous}
@@
 sentence (\ref{sent:debocker}).
+\newpage %until behave
 \eenumsentence{
@@
 by using its genitive form, \emph{``kattens'}. This could be used as a determiner,
+\newpage % until behave
 \enumsentence{Det är kattens mat}
@@
 %TODO part of speech??
-While the pos-tags are reserved for words, the categories give part-of-speech information,
+While the pos-tags are reserved for words, the categories give part-of-sentence information,
 such as \verb|S|, \verb|NP| and \verb|VP|.
@@
 By starting from the GF resource grammar, we got a well-defined system for
-describing grammars. 
+describing language. 
 %as we built the grammar from .

### 2012-01-07 testSuites/ParseLex.hs
parsing improvements

@@ -82,7 +82,6 @@ input = input' ""
        initSafe [] = []
        initSafe xs = init xs
 
---parse without state
 parseNormal pgf morpho lang s = do
   let fix     = fixPunctuation $ replaceNumbers s 
       unknown = badWords morpho (snd fix)
@@ -92,15 +91,24 @@ parseNormal pgf morpho lang s = do
          putStrLn $ "Parsing "++s
          tree <- timeout timeLimit $ return $ parse pgf lang textType (map toLower s)
          case tree of
-            Just xs@(x:_) -> return $ Just (x,length xs)
+            Just xs@(x:_) -> return $ Just (getBest xs,length xs)
             _             -> return Nothing
-       
+       getBest :: [Tree] -> Tree
+       getBest ts = fromMaybe (head ts) (rank ts) 
+       rank :: [Tree] -> Maybe Tree
+       rank = listToMaybe . filter (not .(\x -> any (`isInfixOf` x) bads) . showExpr [])
+       bads = ["NumCard","man_nn","faar_nn","en_nn","foer_nn"
+              ,"mycken_nn","nyy_nn","vill_ab","titt_nn","leva_nn","haer_nn"
+              ,"PrepCN"]
+ 
 
 testa :: String -> IO ()
 testa s = do
   pgf <- readPGF "BigParse.pgf"
   let morpho = buildMorpho pgf (read "BigParseSwe")
   print $ lookupMorpho morpho $ map toLower s
+  tree <- parseNormal pgf morpho (read "BigParseSwe") ("",s)
+  print tree 
   --print [w| w <- words s, null $ lookupMorpho morpho $ map toLower w]
 
 badWords :: Morpho -> String -> [String]

### 2012-01-07 doc/report/report.tex
pres, report

--- 2012-01-07 doc/report/report.tex
@@
 \abstract{
-The thesis describes the work formalising an important fragment of Swedish
+The thesis describes the work of achieving a covering Swedish grammar
 in the GF formalism and to compare and evaluate the possibilities of using GF
@@
 %TODO LREC ref?
-In previous work\footnote{??}, a start was made to extend the Swedish GF grammar
+In previous work\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf}, a start was made to extend the Swedish GF grammar
 and a tool for lexical acquisition was developed.
@@
 acquisition could be used. The tool developed in the previous part of
-this project\footnote{??} would be suitable.
+this project\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf} would be suitable.
 All in all, it should be ensured that we use as much information as

### 2012-01-08 testSuites/Gui.hs
new fancy gui

@@ -33,6 +33,7 @@ main = do
   (maps,pgf) <- play'
   initGUI
   window <- windowNew
+  widgetModifyBg window StateNormal (Color 25000 25000 45555)
   bigbox  <- vBoxNew True 0
 
   set window [windowDefaultWidth := 200, windowDefaultHeight := 200,

### 2012-01-08 testSuites/ParseLex.hs

@@ -46,15 +46,15 @@ play = do
 
 play' = do
   putStrLn "before pgf"
---  maps <- mkLexMap
-  pgf <- readPGF pgfBackUp --pgfDict
+  maps <- mkLexMap
+  pgf <- readPGF pgfDict
   putStrLn "after pgf"
-  return (error "undefined from play'",pgf)
+  return (maps,pgf)
 
 processparse :: String -> PGF -> Language -> LexMap -> IO (Maybe ((FilePath,FilePath),Int))
 processparse str pgfDict langDict maps = do
---  res <- mkDir maps (prepare str) langDict pgfDict
-  let (pgf',l') = maybe (pgfBackUp,langBackUp) id Nothing
+  res <- mkDir maps (prepare str) langDict pgfDict
+  let (pgf',l') = maybe (pgfBackUp,langBackUp) id res
   putStrLn "reading new PGF"
   pgfNew <- readPGF pgf'
   let morpho = buildMorpho pgfNew l'

### 2012-01-08 doc/report/report.tex
correction in report

@@
 For each word in Saldo, it decides which forms should be used as input
-to the the GF smart paradigms. For a verb, this will in most cases mean giving
+to the GF smart paradigms. For a verb, this will in most cases mean giving
 the present tense form, see figure \ref{fig:saldoknyt}. \\
@@
 constant names in the GF lexicon. However, the identifier may need some renaming since
-the there are special characters  in the Saldo identifiers
+there are special characters  in the Saldo identifiers
 that should be avoided in GF function names. The importation therefore
@@
 As already stated, the visualized parse trees is not a complete representation,
-even if the verb phrases in the the visualizations look the same, the two cases
+even if the verb phrases in the visualizations look the same, the two cases
 are treated and represented differently internally. The fronting of the

### 2012-01-12 doc/report/report.tex
report corrections

--- 2012-01-07 doc/report/report.tex
@@
 %ska innehålla vettiga sökord
-%TODO page numbers!!
-\title{Towards a Wide-Coverage Grammar for Swedish Using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
+\title{Towards a Wide-Coverage Grammar for Swedish\\ Using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
 \author{Malin Ahlberg}
@@
 \abstract{
-The thesis describes the work formalising an important fragment of Swedish
-in the GF formalism and to compare and evaluate the possibilities of using GF
-for modelling large-scale grammars. Grammatical constructions specific to and
-frequent in Swedish have been investigated and formalised in the functional
-programming language.
+The goal of this thesis is to enable open-domain parsing for Swedish
+by using the GF \\grammar formalism,
+%by %The idea has been to% 
+combining existing language technology resources with
+new techniques.
+The work includes the development of a covering grammar,
+%in the GF formalism
+to identify and formalize \\
+grammatical constructions %specific to and
+frequent in Swedish. % have been investigated and formalised in the functional
+%programming language.
+We compare and evaluate the possibilities of using GF
+for modelling large-scale grammars and for 
+%and the aim to reaching our goal;
+obtaining a system for analysing written text.
+To reach this goal, problems of computational as well as linguistic nature had to be solved.
 %~~achieve a covering Swedish~~ grammar implemented
@@
 %in GF -- to model natural language by functional programming. 
-The idea has been to combine existing language technology resources with
-new techniques, and the aim to obtain a system for analysing written text.
-To reach this goal, problems of computational as well as linguistic nature had to be solved.
-\\We present the methods and technologies used and further inspects the possibilities
+\\We present the methods and technologies used and further inspect the possibilities
 of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
@@
 \\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
-to GF notation and a Swedish grammar implementation. The grammar is based
+to GF notation and an extended Swedish grammar implementation. The grammar is based
 on the multilingual abstract syntax given in GF resource library, and now
-extends it to cover constructions specific to Swedish.
+also covers constructions specific to Swedish.
 We further give an example of the advantage of using dependent types when
@@
 \vspace{5mm}
-Many people have helped me during this work and made this thesis possible.
+Many people have helped me during this project and made work thesis possible.
 I would first of all like to thank
@@
 I am also grateful to Krasimir Angelov, Markus Forsberg, Peter Ljunglöf, Lars Borin
-and many others who have contributed with ideas and inspiration, and for their
+and many others who have contributed with ideas and inspiration and shown %for their
 interest in this work.
 
-Finally, I would also like to thank my friends and family. Special thanks to Dan % Rosén
+Finally, I would like to thank my friends and family. Special thanks to Dan % Rosén
 for all his support, advice and patience and  -- most importantly -- for being
@@
 % more intro to gf. dependently typed...high-level formalism
-Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism,
-with facilities for multilinguality. It is based on Martin-Löf type theory which allows
+Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
+%with facilities for multilinguality. 
+It is based on Martin-Löf type theory which allows
 reasoning within the programming language.
-GF has so far been successfully used for 
-controlled languages \cite{cnl}, but recent experiments have showed
+GF has strong support for multilinguality and has so far been successfully used for 
+controlled languages \cite{cnl}. Recent experiments have showed
 that it is also possible to use the framework for parsing free language \cite{patent}.
@@
 % --> sentence about connection GF - grammar based parser.
-This project focuses on implementing a wide-coverage Swedish grammar, which
-describes a parser accepting all sentences that can be expressed by the grammar.
-
+%This project focuses on implementing a wide-coverage Swedish grammar, which
+%describes a parser accepting all sentences that can be expressed by the grammar.
 Parsing, that is the task of automatically identifying morphological and
@@
 a parser or grammar is not freely available does not only restrict its
-usage but also its possibilities of being further developed and our
+usage but also its possibilities of being further developed. Our
 goal is
-to create an open source Swedish grammar. % from which we derive a parser
-%accepting all sentences described by the given rules.
+to create an open-source Swedish grammar from which we derive a parser
+accepting all sentences described by the given rules.
 As a freely available resource, it can be continuously enhanced and
 made use of in other projects.\\
+
 To build the grammar from scratch would be not only time consuming but
-would also mean that already existing resources would have to be reimplemented.
+would also mean that already existing systems would have to be reimplemented.
 In order to not reinvent the wheel we proceed from a combination of well-tested
-resources.
+sources.
 We start from a GF resource grammar, consisting of an abstract and a concrete syntax
-file defining a set of rules describing morphology and syntax.
+file defining a set of rules for morphology and syntax.
 %TODO verify this, check all occurences of grammar
-From now on, this is what we mean by grammar, as
+This is what is meant by grammar in this thesis, as
+%From now on, this is what we mean by grammar, as
 opposed to grammar in the traditional linguistic sense. 
@@
 %The usage of GF allows us to start from
-a well-defined system for describing language, as well as a strong connection to
+a well-defined system for describing language, as well as a strong connection
+to and possibility of translation between the 
 more than 20 other languages implemented in the framework. Further, we use
@@
 for achieving and keeping a large-scale lexicon and to adapt the existing
-resource grammar to fit more language specific constrictions.
-Three sub-goals has been identified:
+resource grammar to model more language specific constrictions.
+The project was divided into three subsections, aiming at
+%Three sub-goals has been identified:
 \begin{itemize}
@@
 \section{Outline}
-The thesis is divided into 6 chapters. We start by giving some background
-information to areas relevant to the project. %Chapter \ref{sec:background} 
+The thesis is divided into 6 chapters. We start by giving background
+information of areas relevant to the project. %Chapter \ref{sec:background} 
 Grammatical Framework is presented in section \ref{sec:gf} while a more profound
@@
 %programming languages like 
-ML \cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
+ML \\\cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
 and the built-in functionality for logic and reasoning 
@@
 random generation, word alignment, morphological quizzes etc.
-An interactive GF editor and the GF shell can be tried out online 
-\footnote{http://www.grammaticalframework.org/demos/gfse/}. See figure \ref{fig:shell},
- The first command will parse the given
-sentence and visualize the resulting parse tree. The second will visualize the 
-abstract syntax tree. The trees will be showed in the pdf-format by the program evince.
+The shell can be tried out online 
+together with an interactive GF editor %and the GF shell 
+\footnote{http://www.grammaticalframework.org/demos/gfse/}.
 \thispagestyle{plain}
@@
 \begin{verbatim} 
+> parse "jag ser katten" | visualize_tree  -format=pdf -view=evince
 > parse "jag ser katten" | visualize_parse -format=pdf -view=evince
-> parse "jag ser katten" | visualize_tree  -format=pdf -view=evince
 \end{verbatim}
@@
 \end{figure}
-
-% nice: test vt -format=pdf -view=evince -nofun
-The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
-Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
-%A GF parse tree has two visualized representations, the \textit{parse tree} and
-%the \textit{abstract tree}.
-The \textit{abstract tree} shows which functions GF uses
-to parse a sentence.
-The \textit{parse tree} (\ref{pic:gfStree}) shows which types
-that are assigned to the words and phrases. 
+Figure \ref{fig:shell} shows commands for parsing, where the results are shown as
+in figure \ref{fig:gftree1}. 
+GF uses Graphviz\footnote{http://www.graphviz.org/}
+to visualize the trees and the given commands (fig. \ref{fig:shell}) specify that the
+output format should be PDF-files and that these should be opened by the program
+Evince. The user can choose to see the \textit{parse tree} (fig. \ref{pic:gfStree})
+or the \textit{abstract tree} (fig. \ref{pic:gfAtree}). Abstract trees are more verbose and show all
+functions and types used for parsing the sentence. 
 \begin{figure}[h]
 \centering
+\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
 \subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
-\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
 \caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
@@
 \vspace{10mm}
-% do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
-The reader should keep in mind that the visualized parse trees is not a
-complete representation. It does not reproduce all information from the abstract tree.
-%If the end node of a  branch has no corresponding word, it is excluded from
-%the visualized parse tree. This is the case for the 
-For example, the definiteness and number of the noun \emph{`katten'}, which is shown as
-\verb-DetQuant DefArt NumSg- in the abstract tree.
-%and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
+
+Parse trees on the other hand show only the types assigned to the words and phrases.
+Information about tense, polarity etc. which are explicitly given in the abstract tree are
+not reproduced in the parse tree. Hence, parse trees do not give complete representations
+but model the parse resluts in a transparent manner. For our example, the definiteness and
+number of the noun `\emph{katten}' is shown as \verb-DetQuant DefArt NumSg- in
+the abstract tree while the parse tree only show that a the noun phrase consists of one noun. 
+% is used but not in the parse tree.
 In the corresponding English parse tree, figure \ref{pic:gfEtree},
-the noun is explicitly quantified by the article \emph{``the"},
-and the determiner is therefore shown in the parse tree.\\
+the noun is explicitly quantified by the article \emph{`the'},
+and the determiner, the first argument to the function \verb-DetCN-, is
+therefore shown in the parse tree.\\
+
+% The first command will parse the given
+%sentence and visualize the resulting parse tree. The second will visualize the 
+%abstract syntax tree. The trees will be showed in the pdf-format by the program evince.
+%% nice: test vt -format=pdf -view=evince -nofun
+%The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
+%Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
+%%A GF parse tree has two visualized representations, the \textit{parse tree} and
+%%the \textit{abstract tree}.
+%The \textit{abstract tree} shows which functions GF uses
+%to parse a sentence.
+%The \textit{parse tree} (\ref{pic:gfStree}) shows which types
+%that are assigned to the words and phrases. 
+%% do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
+%The reader should keep in mind that the visualized parse trees is not a
+%complete representation. It does not reproduce all information from the abstract tree.
+%%If the end node of a  branch has no corresponding word, it is excluded from
+%%the visualized parse tree. This is the case for the 
+%For example, the definiteness and number of the noun \emph{`katten'}, which is shown as
+%\verb-DetQuant DefArt NumSg- in the abstract tree.
+%%and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
+%In the corresponding English parse tree, figure \ref{pic:gfEtree},
+%the noun is explicitly quantified by the article \emph{`the'},
+%and the determiner is therefore shown in the parse tree.\\
 %>> how parse trees work, not correct complete representation.
@@
 %TODO read again
-The scope of GF grammars has so far been controlled language, a limited
-language used in a restricted domain. By
+The scope of GF grammars has so far been controlled language, a subset of a
+natural language used for a restricted domain. By
 restricting the coverage, the number of ambiguities
-can be limited and we can ensure that the semantics is preserved during
+can be limited and it can be ensured that the semantics is preserved during
 translation. Inherited ambiguities as well as ambiguities arising from multilinguality will
@@
 
-The use of controlled language thus gives a possibility of good and reliable translation
-and GF has successfully been used in project like %Some examples of projects using GF are 
-WebAlt \cite{webalt}, a project which aims to develop language independent
-material for mathematics problems. The formal software verification tool
-KeY \cite{key}, in which GF is used for translating specification to natural language.
+The use of controlled language thus gives the possibility of good and reliable
+translation and GF has successfully been used for this in a number of projects.
+%like %Some examples of projects using GF are 
+WebAlt \cite{webalt} aims to develop language independent
+material for mathematics problems by using GF grammars.
+The formal software verification tool
+KeY \cite{key} used GF for translating formal specification to
+natural language.
 %>> key, talk, hats, bom (see references in Ramonas thesis).
 %>> use talk here and move key to other place??
-GF can be used for describe dialog grammars see \cite{talk}, and
-the framework is further used in the
-European project MOLTO \footnote{http://www.molto-project.eu/} for online
-translation between up to 15 language.\\
+GF has further been used for describing dialogue grammars \cite{talk}, and
+the framework is one of the main component in the
+European project MOLTO \footnote{http://www.molto-project.eu/}, which develop
+online translation between to 15 languages.\\
 
@@
 for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|.
-The function \verb|Pred| simply glues the two strings ``katten" and ``sover" together:\\
+The function \verb|Pred| simply glues the two strings `katten' and `sover' together:\\
 \verb|Pred cat sleep = "katten sover"|.\\
@@
 %now we concentrate on the number only. Hence \verb|N| now means a noun in any of the number,
-and we introduce two functions that set the number: \verb|NSg : N -> N'| and
+and we introduce two functions for setting the number: \verb|NSg : N -> N'| and
 \verb|NPl : N -> N'|.
@@
 %this can be found as libraries. So far there are 20 languages in the library, and
-The languages all share the same abstract syntax which is  valuable
+Since the languages all share the same abstract syntax translation between any given
+pair is possible, which is valuable
 %they constitute  a very useful resource 
-when implementing multilingual application grammars.
+when implementing multilingual applications.
 
@@
 \end{wrapfigure}
-This is the case for ``flicka", a noun belonging to the first declination. %the correct declination
-%can be guessed by using the first declination. % and noting that the word ends with an ``a". 
-%Like most Swedish nouns ending with an ``a", it follows the first declination.
+This is the case for `flicka', a noun belonging to the first declination. %the correct declination
+%can be guessed by using the first declination. % and noting that the word ends with an `a'. 
+%Like most Swedish nouns ending with an `a', it follows the first declination.
 For others, like "hjärta", also the plural form "hjärtan" is needed.
-%, even though it ends with an ``a". 
-The worst case is nouns that need four forms, both singular and
+%, even though it ends with an `a'. 
+The worst case for nouns is four needed forms, both singular and
 plural in definite and indefinite form.
@@
 XML-formats
-where the trees had been made deeper and more detailed while still containing
+where the trees have been made deeper and more detailed while still containing
 the lexical MAMBA layer. \\
@@
 \centering
-\subfloat[Morphological info for ``katt"]{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
+\subfloat[Morphological info for `katt']{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
 \hspace{20mm}
-\subfloat[Graph showing the hyponyms of ``katt"]{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
+\subfloat[Graph showing the hyponyms of `katt']{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
 \caption{Information in Saldo}
@@
 %TODO change word? uppskatta, känna, få
-The word ``växa" (``grow") for example, has two meanings: \emph{getting
-bigger}
-\enumsentence{Gräset växte varje dag\\
-`The grass grew each day'}
-or \emph{being a plant}. 
-\enumsentence{Det växte en gammal ek där.\\
-`There stood an old oak.'}
-``växa" is however declined the same way for both interpretations of the word.
-Hence \textit{Saldo's morphological lexicon} only keeps one entry of ``växa".
+The word `uppskatta' for example, has two entries, with sligthly
+different semantics: \emph{enjoy}
+\enumsentence{Jag uppskattar teater. \\
+``I enjoy theater."}
+or \emph{estimate}. 
+\enumsentence{Värdet uppskattades till 5 miljoner.\\
+``The value was estimated to 5 million."}
+%\enumsentence{Barnen växte varje dag\\
+%``The children grew for each day"}
+%or \emph{vegetate}. 
+%\enumsentence{Det växte en gammal ek där.\\
+%``There stood an old oak."}
+`uppskatta' is however declined the same way for both interpretations of the word.
+Hence \textit{Saldo's morphological lexicon} only keeps one entry of `uppskatta'.
 For the homonyms that do not share the whole inflection tables, different entries are 
-kept also in the morphological lexicon.
-
+kept also in the morphological lexicon.\\
 For our purpose, only the morphological section, provided under LPGL in XML
@@
    & $[${\sc+def}$]$ & $[${\sc+def}$]$&    &&  & $[${\sc+def}$]$ & $[${\sc+def}$]$ & $[${\sc+def}$]$&  \\
-   &~~`The old       & cat           &slept.'&& &`The            & old             & cat            &slept.'\\
+   &~~``The old       & cat           &slept."&& &``The            & old             & cat            &slept."\\
 \end{tabular}}\label{sent:trottkatt}
@@
 {min &trötta& katt} %&& Den där & trötta &katten}
-{`my & tired & cat'}
+{``my & tired & cat"}
 { $[${\sc+def}$]$ &$[${\sc+def}$]$& $[${\sc-def}$]$} % . && $[${\sc+def}$]$  & $[${\sc+def}$]$ &$[${\sc+def}$]$}
@@
 {\emph{you} & \emph{see} & \emph{not} &\emph{ me}}
-{`You  don't see me'.}} \label{ex:swedish-svo}
+{``You  don't see me."}} \label{ex:swedish-svo}
 \vspace{-3mm}
@@
 {\emph{saw}  & \emph{you}&  \emph{me}& \emph{not}}
-{`Didn't you see me?'}}
+{``Didn't you see me?"}}
 %\enumsentence{Gick Johan på gatan?\\
@@
 { \emph{I} & \emph{saw} & \emph{Anna} & \emph{but} & \emph{she} & \emph{saw} & {\bf \emph{not}} &\emph{ me}}
-{`I saw Anna but she didn't see me'}}}
+{``I saw Anna but she didn't see me"}}}
 \enumsentence{Subordinate sentence:\\
@@
             {\emph{I} &  \emph{understood} & \emph{that}& \emph{Anna}& \textbf{\emph{not}}& \emph{saw}& \emph{me}}
-            {`I understood that Anna didn't see me'}}
+            {``I understood that Anna didn't see me"}}
 
@@
 {\emph{the} \emph{thief} & \emph{was} & \emph{taken} & \emph{by} & \emph{the police} }
-{`The thief was arrested by the police'}}
+{``The thief was arrested by the police"}}
 \label{gfPass:peri-pass}
@@
 {\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
-{`The thief was arrested by the police'}
+{``The thief was arrested by the police"}
 {&&Polisen & tog & tjuven}
 {&&\emph{the police} & \emph{took} & \emph{the thief}}
-{\hspace{-9mm}`The police arrested the thief'}}
+{\hspace{-9mm}``The police arrested the thief"}}
 \label{gfPass:s-pass}
@@
 \subsection{Impersonal constructions}
-Constructions with `det är/var' (\emph{`it is/was'}) \\are very common in Swedish
+Constructions with `det är/var' (\emph{`it is/was'}) are very common in Swedish\\
 \cite[\textsection 309d]{H&H}:
@@
 {\emph{it} &\emph{was} & \emph{nice} &\emph{to} &\emph{hear}.}
-{`I'm glad to hear that.'}}
+{``I'm glad to hear that."}}
 'Det' %(\emph{it})
@@
 { \emph{it} & \emph{stands} &\emph{a} & \emph{moose} &\emph{on} &\emph{the field}}
-{`There is a moose in the field.'}}
+{``There is a moose in the field."}}
 
@@
 The cascaded finite state parser CassSwe \cite{casswe} and
-The Swedish Constraint Grammar (Birn,1998) 
+The Swedish Constraint Grammar \cite{birn}
 give syntactic analyses. %birn: shallow
 Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
-(Tapanainen and Järvinen,1997), an extension of the Constraint Grammar
+\cite{fdg},
+%(Tapanainen and Järvinen,1997), 
+an extension of the Constraint Grammar
 formalism, and produces a dependency structure focusing on finding the nominal
@@
 The tables are compared to the information given
-in Saldo, if the tables are equal the code for the word is saved. If the table
+in Saldo and if the tables are equal the code for the word is saved. If the table
 is erroneous, another try is made
@@
 \end{figure}
+\newpage
 The program is run iteratively until the GF table matches the one given in Saldo,
 or until there are no more ways of using the smart paradigm. The verb 'knyta'
-will need tree forms:\\
+will need three forms:\\
 
@@
 \subfloat[GF]{\label{pic:gfVax}\includegraphics[height=60mm]{gfVax.png}}
-\caption{Inflection table for the verb ``växa" (grow)}
+\caption{Inflection table for the verb `växa' (grow)}
 \label{pic:TabVax}
@@
 The tables does not overlap completely, there are some more forms in Saldo's table (e.g. the 
-compounding form ``väx-")
-while the one generated in GF contains some that are not in Saldo (e.g. ``vuxits").
+compounding form `väx-')
+while the one generated in GF contains some that are not in Saldo (e.g. `vuxits').
 As GF concerns about syntax only, and not semantics, and  as the GF tables are automatically
@@
 used in the natural language.
-Saldo may also contain variants: ``växt" and ``vuxit" are both supine form.
+Saldo may also contain variants: `växt' and `vuxit' are both supine form.
 As far as possible, the program makes up for this by only comparing the overlapping forms
@@
 Saldo in the GF lexicon. Prepositions, %(\emph{pp})
-numerals %(\emph{nl}),
+numerals, %(\emph{nl}),
 personal pronouns etc.
@@
 Saldo contains many pronouns which
-are not analysed the same way in GF (see section \ref{sec:swegf}).
+are not analysed the same way in GF (see section \ref{sec:gfnouns}).
 Before adding them to our lexicon, we need to do more analysing to find their correct GF-category.
@@
 %the right GF-category for those
-have been done using Talbanken, see p. \pageref{sec:gf.quant}.
+have been done using Talbanken, see section \ref{sec:gf.quant}.
 
@@
 are usually handled as idioms and should be given in a separate lexicon. In
-total six categories were imported: \\
+total six types of words were considered for the extraction: \\
 
@@
 %different grammatical analyse, a lot of pronouns. Example. 
-Most but not all words of the categories listed above have been imported.
+Most but not all words of these categories have been imported.
 One reason why the importing phase would fail 
@@
 the plural tantum % \cite[44]{SAG} 
-noun \emph{``glasögon"} (\emph{``glasses"}).
+noun \emph{`glasögon'} (\emph{`glasses'}).
 The smart paradigm requires a singular form, and since the program could not
@@
 When the program failed to import a noun, this was often the explanation.
-%bbMany nouns fai, \emph{``kläder"} (\emph{``clothes"} etc.
-Words of this type may be added manually, for \emph{``glasögon"} we could use
-the ostensibly correct singular form ``glasöga", although this
-has another meaning (``glass-eye").
+%bbMany nouns fai, \emph{`kläder'} (\emph{`clothes'} etc.
+Words of this type may be added manually, for \emph{`glasögon'} we could use
+the ostensibly correct singular form `glasöga', although this
+has another meaning (`glass-eye').
 The same problem occurred for the irregular s-verbs,
-(\emph{``synas"} \emph{(``show")} or \emph{umgås} \emph{(``socialize")})
+(\emph{`synas'} \emph{(`show')} or \emph{umgås} \emph{(`socialize')})
 which made up 61.5 \% of the failing verbs of type \verb_vb_.\\
-%``frodas" (``flourish"), ``lyckas" (``succeeed").
+%`frodas' (`flourish'), `lyckas' (`succeeed').
 In a few cases the smart paradigms could not generate the correct declination.\\
@@
 4/5 of the word forms are compounds and when performing the intended parsing,
-these an additional analysis before being looked-up in the lexicon. We
+an additional analysis identifying compounds should be preformed before
+looking-up the words in the lexicon. We
 should also take into consideration that 
 we cannot automatically find out how many actually stem from the same word, or
-how many abbreviations that are present (Talbanken also contains a small number of 
-spelling errors, which probably are enumerated among our missing words). The majority
+how many abbreviations that are present. Talbanken also contains a small number of 
+spelling errors, which probably are enumerated among our missing words. The majority
 of the missing words are only used once.\\
@@
 \subsection{Noun phrases}
+\label{sec:gfnouns}
 Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
@@
 %TODO read again
-The category \verb-N'- was a simplification and is not used in the resource grammar. and is not used in the resource grammar.
+The category \verb-N'- was a simplification and is not used in the resource
+grammar. 
 The full analysis includes a distinction between the categories \verb-N- -- simple nouns --
@@
 \verb-CN- -- common noun phrases -- and \verb-NP- -- noun phrases.
-The \verb-N- category is implemented as an inflection table, generated by the smart paradigm,
+The category \verb-N- is implemented as an inflection table, generated by the smart paradigm,
 and a field keeping information about the gender, see figure \ref{fig:gfflicka}.
@@
 \end{figure}\\
-When a noun is turned into a \verb-CN- %, this information is kept, and the noun may
+Nouns may turned into common noun phrases \verb-CN-, %this information is kept, and the noun may
 which may be modified by adjectival phrases or conjoined.
@@
         indefinite, singular -> flicka -> "en flicka"
-        definitive,singular  -> flicka -> "flickan"
+        definitive, singular -> flicka -> "flickan"
 \end{verbatim}        
@@
 (\emph{`small'}) and \emph{`katt'} (\emph{`cat'}), there are three ways they
-can be combined, as shown in figure \ref{tab:detdef}
+can be combined, as shown in figure \ref{tab:detdef}. The noun may be used in
+definite or indefinite form and the adjective in its weak or strong form \cite[p. 31]{gunlog}.
 \begin{figure}[h]
@@
 feature of the determiner.
-The resource grammar distinguishes between quantifiers, 
-determiners \verb-Det- and predeterminers. Predeterminers modify \verb-NP-s
-while the other to modifies \verb-CN-s. The differences are shown in table
+The resource grammar distinguishes between quantifiers (\verb-Quant-), 
+determiners (\verb-Det-) and predeterminers (\verb-Predet-). Predeterminers modify \verb-NP-s
+while the other to modifies \verb-CN-s. The differences are further shown in table
 \ref{tab:detquant}. 
@@
 \emph{en} and \emph{ett} for singular. In plural it
-is the either \emph{``de"} or nothing, cf. sentence (\ref{sent:trottkatt2}a) and
+is the either \emph{`de'} or nothing, cf. sentence (\ref{sent:trottkatt2}a) and
 \ref{sent:trottkatt2}b.
@@
 
-We need to know if the common noun phrase 
+In order to get the correct plural form
+we need to know if the common noun phrase 
 has been modified, that is, whether the function \verb_AdjCN_ has been used.
@@
 %TODO review, verify
-Due syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
+Due to the syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
 contains personal pronouns.
-Many words which are considered to be pronouns in other analyses, such as The Swedish
-Academy Grammar \cite{SAG}, only exist in 3rd person and are 
-classified differently in GF, usually as determiners or quantifiers.
+Many words, like `somliga' (\emph{`some'}), are considered to be pronouns
+in other analyses, such as The Swedish
+Academy Grammar \cite{SAG},  but are classified differently in GF, usually
+as determiners or quantifiers as they may determine noun phrases
+(\ref{sent:somestud}).
+\enumsentence{Somliga studenter jobbar bara på nätterna\\
+``Some student only work at night time"}\label{sent:somestud}
 %Only personal pronouns are of the type \verb-Pron-.
@@
 valency information
-is encoded into the lexicon. Transitive verbs have the type \verb-V2-, ditransitive
-\verb-V3- etc. There are types for verb which takes sentences (\verb-VS-) and
+is encoded into the lexicon. Transitive verbs have the type \verb-V2- and ditransitive
+\verb-V3-. There are types for verb which takes sentences (\verb-VS-) and
 adjectival (\verb-VA-) complements.
@@
 {\emph{I} &\emph{take}&\emph{with} &\emph{me}& \emph{it}& \emph{to}& \emph{him}}
-{`I bring it to him'}
+{``I bring it to him"}
 \label{sent:tamed}}
-The verb `ta' in sentence (\ref{sent:tamed}) takes one particle (\emph{med}),
-one preposition (\emph{till}) and two objects: \emph{den} and \emph{honom}.
-
+The verb `ta' in sentence (\ref{sent:tamed}) takes one particle (\emph{`med'}),
+one preposition (\emph{`till'}) and two objects: \emph{`den'} and \emph{`honom'}.
 In a GF lexicon this verb is given the
@@
 Particles are given in the lexicon, as well as the prepositions that are
-chosen by the verb.
+chosen by the verb, since these may not vary.
 The entry for `\emph{ta med}', as used in sentence (\ref{sent:tamed}), is
 described as follows:
-%, assuming that we already have an constant \verb-take_V- which produces the correct conjugation for ``ta":
+%, assuming that we already have an constant \verb-take_V- which produces the correct conjugation for `ta':
 \begin{verbatim}
@@
 \end{verbatim}
-%The category \verb-V3- tells us that the verb takes two objects, % (``den" and ``honom").
+%The category \verb-V3- tells us that the verb takes two objects, % (`den' and `honom').
 The function \verb-dirV3- creates a three-place verb, where the first object is direct and
@@
 and \ref{pic:tittapa}.
+\begin{figure}[h]
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=70mm]{tittapa.jpg}}
+\hspace{10mm}
+\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=76mm]{tittapart.pdf}}
+\caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
+\label{fig:translationtrees}
+\end{figure}
+\newpage
 As already stated, the visualized parse trees is not a complete representation,
-even if the verb phrases in the the visualizations look the same, the two cases
+even if the verb phrases in the visualizations look the same, the two cases
 are treated and represented differently internally. The fronting of the
@@
 {\emph{on} &\emph{the boy}& \emph{look}& \emph{you}}
-{`You look at the boy.'}
+{``You look at the boy."}
 \item[b]\shortexnt{6}
@@
 \label{sent:tittapa}}
-\begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=72mm]{tittapa.jpg}}
-\hspace{10mm}
-\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=72mm]{tittapart.pdf}}
-\caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
-\label{fig:translationtrees}
-\end{figure}
-
-\newpage
+
 %Categories as V,V2,VS...
@@
 This is a convenient way to form questions, relative clauses
-and topicalized clauses, as introduced by \cite{gazdar} (see figure \ref{fig:clause2}).\\
+and topicalized clauses (see figure \ref{fig:clause2}), as introduced in \cite{gazdar}.\\
 
@@
 \hline
-\textbf{Wh-Questions} & Interrogative pronoun   & Vad tittar Johan på?\\
-\textbf{Relative clauses} & Object + Relative pronoun & Katten som Johan tittar på \\
-\textbf{Topicalized clauses} & Object  & Henne tittar Johan på \\
+\textbf{Wh-Questions} & Interrogative pronoun   & `Vad tittar Johan på?'\\
+\textbf{Relative clauses} & Object + Relative pronoun & `Katten som Johan tittar på' \\
+\textbf{Topicalized clauses} & Object  & `Henne tittar Johan på' \\
 \hline
@@
 phrase is still treated as one constituent in GF. The parts
-are connected in the tree, and the subject \emph{``han"} is put between the
+are connected in the tree, and the subject \emph{`han'} is put between the
 finite verb and the rest of the phrase.
-At the code level, This is done by putting the \verb-VP- fields in the correct order\\
+At the code level, this is done using the record type for \verb-VP-, wich
+consists of fields that can be put %the \verb-VP- fields 
+in different order.\\
 \verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
@@
 In addition to the core resource grammars, which is shared with the other languages implemented in the
-library, 
+library, there is also an extra module, simple called the \verb|Extra| module, where
 %In addition to the basic resource grammar,
@@
 %>> core is shared but particular lang. has particular features.
-to the module \verb|Extra|. The functions given here do not have to be translatable
+%to the module \verb|Extra|.
+The functions given here do not have to be translatable
 to all other language, but are meant to cover language specific constructions.
-Among those were functions for topicalisation: %$ although not all
+Among those were functions for topicalisation (\ref{gfSwe:apple}), %$ although not all
 %of them were implemented for Swedish.
@@
 {that & apple & want & I & not &have}
-{`I don't want that apple'}}\label{gfSwe:apple}
+{``I don't want that apple"}}\label{gfSwe:apple}
 and for preposition stranding:
@@
         {\emph{the essay} & \emph{wrote+\textbf{s}} & \emph{by} & \emph{a} & \emph{student}}
-        {`The essay was written by a student.'}}\label{sent:skrevs}
+        {``The essay was written by a student."}}\label{sent:skrevs}
 %{b. &En & student & skrev & uppsatsen.}
@@
 
-During this project, the s-passive was added as the standard case.
-Periphrastic passive is still allowed, but as an alternative rather
-than the default.
+During this project, the s-passive was added although the
+periphrastic passive is still allowed.
 The grammar further allows not only V2, but all verb phrases that misses an object, to 
@@
 \end{verbatim}
-A \verb-V3- like `give' in sentence (\ref{sent:give2pass}) gives rise to two
+A \verb-V3- like `give' in sentence (\ref{sent:give2pass}) hence gives rise to two
 passives, (\ref{ex:passV32}) and (\ref{ex:passV33}).
@@
 {Vi & erbjöd& henne& jobbet }
-{\emph{we} & \emph{offered}& her&\emph{the job}}
-{`We offered her the job'}}
+{\emph{we} & \emph{offered}& \emph{her}&\emph{the job}}
+{``We offered her the job"}}
 \label{sent:give2pass}
@@
 {\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
-{`The job was offered to her'}}
+{``The job was offered to her"}}
 \label{ex:passV32}
@@
 {\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
-{`She was offered the job'}}
+{``She was offered the job"}}
 \label{ex:passV33}
@@
 {Det&dricks&mycket & öl& nuförtiden}
-{\emph{It}&\emph{drink+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
+{\emph{It}&\emph{drinks+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
 {``A lot of beer is being drunk these days"}}
@@
 \emph{saknas} (\emph{miss}) and \emph{fattas}(\emph{lack}).
-\enumsentence{\begin{tabular}[t]{@{}*{14}{l@{\ }}}
-a. &Det & finns & kaffe &
-b. &Det & saknas&  kaffe &
+\enumsentence{\begin{tabular}[t]{@{}*{18}{l@{\ }}}
+a. &Det & finns & kaffe &&&
+b. &Det & saknas&  kaffe &&&
 c. &Det & fattas & kaffe \\
-&\emph{it }& \emph{exist }& \emph{coffee} &
- &\emph{it }& \emph{misses} & \emph{coffee} &
+&\emph{it }& \emph{exist }& \emph{coffee} && &
+ &\emph{it }& \emph{misses} & \emph{coffee} && &
  &\emph{it} & \emph{lacks} & \emph{coffee}\\
 \end{tabular}\\
-{~~''There is coffee"~~ ''There is no coffee"~~ ''There is no coffee"}}
+{~~''There is coffee"~~~~ ''There is no coffee"~~~~ ''There is no coffee"}}
 
@@
 a \verb-SimpleVP-, as can transitive verbs without their objects. The \verb-SimpleVP-
-may be modified by adverbs.
+may further be modified by adverbs.
 
@@
 (\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
-In the standard GF analysis, which is preformed bottom-up starting from the pos-tags, 
-all information about semantic roles are lost. There is no way
-of knowing whether a \verb-NP- acts as subject or object. For this reason,
+In the standard GF analysis, which is preformed bottom-up starting from the
+%TODO verify
+POS-tags, information about semantic roles are given by the functions, not by the
+categories. That is,
+we know that the first argument of the function \verb-PredVP- acts as the subject.
+The noun phrase itself does not carry information about its semantic role.
+Until it is given as an argument to a clause level function, no difference
+is made between subject and object noun phrases.
+For this reason,
 the formalization of reflexive pronouns required the use of a different analysis.
@@
 \end{verbatim}
-Any normal noun phrase may also be used as an object, whereas subjects only
+Any noun phrase may also be used as an object, whereas subjects only
 can be made up by \verb-NP-s.
@@
 subject was changed to 2nd person, they  would need to change too.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
-a. &\textbf{Du} är längre än \textbf{din} kompis.& b. &*Du är här oftare än \textbf{sin} kompis.\\
- & You are taller than your friend. &&You are here more often than {\sc self's} friend.\\
+\eenumsentence{%\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
+\item\textbf{Du} är längre än \textbf{din} kompis.\\
+``You are taller than your friend." 
+\item
+*\textbf{Du} är här oftare än \textbf{sin} kompis.\\
+``You are here more often than {\sc self's} friend."}
 \end{tabular}} \label{ex:ReflAP2} \label{ex:ReflAdv2}
+Adverbs containing reflexive possessive pronouns may further not be used
+to construct subject noun phrases.
+%TODO understandable?
+\enumsentence{\shortexnt{5}
+{*Taket & på & sitt & hus & läcker}
+{~~The roof & on & {\sc self's} & house & leaks}}
+\vspace{-2mm}
+\verb-      (NP    +(Adv (NP Obj)))-\\
+
 The dependency spreads through the grammar and in order to avoid all code
@@
 Following the idea given above, we make a difference between subjects and objects, but not
-by giving them different types, but by % treating them as instances of the \verb-NP- category.
+by giving them entirely different types, but by % treating them as instances of the \verb-NP- category.
 letting the type \verb-NP- \textit{depend} on an argument, which may either be \verb-Subject- or
@@
 \begin{verbatim}
-cat NP Boolean ;
+cat NP NPType ;
 PredVP     : NP Obj  -> VP     -> Cl ;
 ComplSlash : VPSlash -> NP Obj -> VP ;
-PredetNP   : (a : Boolean) -> Det -> NP a -> NP a ;
+PredetNP   : (a : NPType) -> Det -> NP a -> NP a ;
 \end{verbatim}
@@
 We hence combine the-part-of-speech driven analyse normally preformed by
-GF with a part-of-sentence analysis, where the dependent types gives the semantic
+GF with a part-of-sentence analysis, where the dependent types gives the 
 information we were missing.
@@
       have similar reflexive pronouns.
+      %and English reflexive pronouns (eg. `himself') cannot
+      %be used without antecedent.
 \end{itemize}
-Moreover, the goal of the project is to make a language specific grammar, and it
-would be surprising if this could be done while leaving the abstract grammar unchanged.
+Moreover, the goal of the project is to make a language specific grammar,
+whereas the common abstract is developed to be generalizable enough to 
+describe any language. It thus would be surprising if we could cover
+all syntactical aspects of one language while leaving the abstract grammar
+unchanged.
 
@@
 {\emph{they} &\emph{see}& \emph{each other}}
-{`They see each other'}}\label{sent:varandra}
+{``They see each other"}}\label{sent:varandra}
 %\item \shortexnt{3}
@@
 ``I will miss you"}
-The modal verb \emph{``ska"} signals an intention of committing the action, either from the subject or from
+The modal verb \emph{`ska'} signals an intention of committing the action, either from the subject or from
 the speaker. Cf.
@@
 \end{tabular}}
-The verb \emph{``kommer"}, normally used with the infinite marker \emph{att},
+The verb \emph{`kommer'} (\emph{`come'}), normally used with the infinite marker \emph{att},
 does not signal any intention, but that the speaker has belief that 
@@
 The resource grammar included 
-\emph{``ska"}, which was implemented as the standard way of forming future
+\emph{`ska'}, which was implemented as the standard way of forming future
 tense, and hence represents the translation of
-\emph{``will"}. 
+\emph{`will'}. 
 The new grammar also supports \emph{``kommer att"}, expressed as an alternative
@@
 and \verb-AdV-. %like `alltid' (\emph{`always'}).
-The \verb-AdV- (e.g. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}))
+The \verb-AdV-  -- e.g. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}) --
 attaches directly to the verb.
@@
 {&\emph{he} & \emph{only} &\emph{ smiled} & & & &\emph{she} &\emph{even} & \emph{whistles}  }
-{~~``He just smiled"\hspace{8mm} ``She even whistles"}}
+{``\hspace{2mm}He just smiled"\hspace{8mm} ``She even whistles"}}
 Focusing adverbs are accepted by the new grammar implementation, where they have
@@
 \end{verbatim}
-They may also be used as \verb-AdV-...
+They are also allowed as \verb-AdV-...
 \enumsentence{\shortexnt{4}
@@
 {\emph{it} & \emph{is} & \emph{only} & \emph{children} &\emph{that}& \emph{may} & \emph{do} & \emph{so} }
-{`Only children are allowed to do that'}}
+{``Only children are allowed to do that"}}
 Two more rules are consequently needed
@@
 We have however chosen to allow this, since it is grammatically correct although semantically
-dubious, and since \emph{``Han har bara sovit"} is covered by the rules using focusing
-adverbs as \verb-AdV-.
+dubious, and since the alternative linearization \emph{``Han har bara sovit"}
+is covered by the rules using focusing adverbs as \verb-AdV-.
 %The semantic aspect of focusing adverbs remains rather unclear in this
@@
 They are described
-as documentation of the implementation, and to illustrate some problems
+as documentation of the implementation and to illustrate some problems
 and solution of more general nature.
@@
 The resource grammar already gave a good coverage of relative clauses and embedded sentences.
-All constructions used in examples \ref{sent:relex}a-d were accepted.
+All constructions used in examples \ref{sent:relex}a-c were accepted.
 %provided a number of ways to express relative and embedded clauses, such as
@@
 \item[a.]Pojken, som är blyg, tystnar\\
-`The boy, who is shy, falls silent'
+``The boy, who is shy, falls silent"
 \item[b.]Han såg kunden som tyckte om sallad\\
-`He saw the customer who liked salad'
+``He saw the customer who liked salad"
 \item[c.]Jag tänkte på huset i vilket hon bodde\\
-`I thought about the house in which she lived'}\label{sent:relex} % better example!!
+``I thought about the house in which she lived"}\label{sent:relex} % better example!!
 
@@
 nouns in indefinite form, whenever the noun phrase is followed by a restrictive
-relative clause \cite[\textsection 329]{H&H}
+relative clause \cite[\textsection 329]{H&H}.
 Talbanken contained several examples where the modified noun is in the 
@@
 {``the opinions, that are presented ..."}}\label{sent:uppfatt}
-%or together with the verb \emph{``finnas"}.
+%or together with the verb \emph{`finnas'}.
 When no relative clause is present, the definite form with the postnominal
@@
 \eenumsentence{
-\item de uppfattningar\textbf{na} förs fram.\\
-``the opinions are presented."
+\item de uppfattningar\textbf{na} förs fram\\
+``the opinions are presented"
 \item *de uppfattningar förs fram}\label{sent:debocker}
@@
 Jag vill ha en katt sådan att den inte fäller\\
-`I want a cat such that it does not shed'
+``I want a cat such that it does not shed"
 \item
@@
 Jag vill ha en sådan katt som inte fäller\\
-`I want such a cat that it does not shed'}
+``I want such a cat that it does not shed"}
 %\end{tabular}} 
 In GF, this relative clause \emph{``en katt sådan att den inte fäller"} is constructed by
-two constituents: \emph{``katt"}
+two constituents: \emph{`katt'}
 of type \verb-CN- and \emph{``den inte fäller"} of type \verb-RCl- (relative clause), glued
@@
 \newpage % until behave
-\enumsentence{Det är kattens mat}
+\enumsentence{Det är kattens mat.\\``It is the cat's food."}
 and thus as a noun phrase:
-\enumsentence{Det är kattens.}
+\enumsentence{Det är kattens.\\``It is the cat's."}
 The possessive form of this \verb-NP- is now the same as its nominative:
@@
 {\emph{such}& \label{ex:QuantDet}\emph{cats}& \emph{want}& \emph{I}& \emph{have}}
-{`I want such cats'}
+{``I want such cats'}
 \item\shortex{3}
@@
 {\emph{he}& \emph{is}& \label{ex:QuantAdj}\emph{such}\toplabel{sent:Quant}}
-{`He's like that'}
+{``He's like that"}
 }
@@
 \emph{på bilen} should still form a subtree attached to \emph{katten}.
-%produces a \verb-NP- from ``katten" this means that
+%produces a \verb-NP- from `katten' this means that
 %DetCN : Quant                  -> CN    -> NP
@@
 \section{The Talbanken annotation}
-Talbanken05 uses three kinds of tags: categories, edge labels and pos-tags. 
-%TODO part of speech??
-While the pos-tags are reserved for words, the categories give part-of-sentence information,
+Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
+While the POS-tags are reserved for words, the categories give part of sentence information,
 such as \verb|S|, \verb|NP| and \verb|VP|.
@@
 \verb|OO| for object etc. Nearly 400 different tags are used, of which more than
-300 different are pos-tags. The high number is due to the level of detail,  there are for
+300 different are POS-tags. The high number is due to the level of detail,  there are for
 example almost 50 tags for nouns, excluding proper names. The tags
@@
 It is looked-up in the GF lexicon and the result compared to the information of
-the pos-tag.
+the POS-tag.
 If the word is not found in the lexicon
@@
 Lets look at the example in figure
-\ref{pic:tbtree}. The verb \emph{``bli"} is
+\ref{pic:tbtree}. The verb \emph{`bli'} is
 followed by a predicative 
-adjective modifying the subject, \verb|SP|. GF admits \emph{``bli"} to have
+adjective modifying the subject, \verb|SP|. GF admits \emph{`bli'} to have
 type \verb|VA|, a verb with an adjective as complement, and the translator can
@@
 %The grammar requires a adjective-complement verb,
-%\verb|VA|, to be used with one adjectival phrase. For \emph{``bli"} in example
+%\verb|VA|, to be used with one adjectival phrase. For \emph{`bli'} in example
 %\ref{pic:tbtree}, we know that we need to find an adjective phrase.
@@
 translation can be illustrated by sentence (\ref{ex:mappMan}), containing
-the generic pronoun \emph{``man"}.
+the generic pronoun \emph{`man'}.
 \enumsentence{
@@
 
-In Talbanken, \emph{``man"} is simply a personal pronoun, 
+In Talbanken, \emph{`man'} is simply a personal pronoun, 
 in GF it is represented by using the function \verb|GenericCl|, see figure
@@
 %The lexicon may contain a verb both with and without a particle,
-%for example \emph{tänka} (``think") and \emph{tänka till} (``consider").
+%for example \emph{tänka} (`think') and \emph{tänka till} (`consider').
 %\begin{verbatim}
@@
 \item a large-scale GF lexicon and a program to redo the importation when needed
+\item an extended grammar covering an importand part of Swedish
 \item a comparison between GF and another annotation
@@
 
+Besides being capable of reimporting Saldo, the lexicon extraction program could also
+be modified for importing other lexical resources. The only requirement is that the resource
+provides inflection tables.
+
 The grammar has been extended and enhanced, and its current status is
@@
 %TODO make nice!
-Although it cannot not been verified, we believe that the majority of the
+Although it is not been verified, we believe that the majority of the
 sentences generated are grammatically correct in a syntactical point of view.
@@
 {\emph{I}& \emph{begin}& \textbf{\emph{to}}&\emph{become}& \emph{hungry}}
-{I'm getting hungry}
+{``I'm getting hungry"}
 \item
@@
 {\emph{I}& \emph{begin}& \emph{become}& \emph{hungry}}
-{I'm getting hungry}}\label{sent:complVV1}
+{``I'm getting hungry"}}\label{sent:complVV1}
 Since there is no information about which verbs that allows the infinitive mark
@@
 \item \shortex{5}
-{Jag & låter bli & \textbf{att}& titta} 
-{I & let & be &\textbf{to} &watch}
-{I don't watch}
+{Jag & låter &bli & \textbf{att}& titta} 
+{\emph{I} & \emph{let} & \emph{be} &\textbf{\emph{to}} &\emph{watch}}
+{``I refrain from watching"}
 \item \shortexnt{4}
 {*Jag & låter & bli & titta} 
-{I & leave &be & watch}}\label{sent:complVV2}
+{~~\emph{I} & \emph{leave} &\emph{be} & \emph{watch}}}\label{sent:complVV2}
+As these functions provide stylistic flexiblity when parsing, they can be left out
+when generating language, and the grammar then provides text of good
+grammatical quality. \\
 %ComplBareVV, SupCl
-For random generation, the sentences generated may be next to impossible
-to interpret without looking at the abstract tree, since they may contain
-too many attributes, relative clauses etc. 
+%For random generation, the sentences generated may be next to impossible
+%to interpret without looking at the abstract tree, since they may contain
+%too many attributes, relative clauses etc.
+
 %\enumsentence{Kommer övriga datorer idag som ska vara inte att ha målats?}
@@
 %The analysis of this sentence is that
-%avståndet is a noun of type \verb-N3-, which chooses the prepositions ``från" and ``till"
+%avståndet is a noun of type \verb-N3-, which chooses the prepositions `från' and `till'
 %and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
@@
 {\emph{we} &\emph{will}& \emph{buy} & \emph{dog} }
-{We are going to buy a dog}}\label{sent:hund}
+{``We are going to buy a dog"}}\label{sent:hund}
 
@@
 {\emph{he} & \emph{begs} & \emph{her} & \emph{to} & \emph{clean} & {\sc self's} & \emph{room}}
-{`He asks her to clean her room'}
+{``He asks her to clean her room"}
 \item[b]\shortex{7}
@@
 {\emph{he}& \emph{promises}& \emph{her}& \emph{to}&  \emph{clean} &\sc self's&\emph{room}}
-{`He promises her to clean his room'}}
+{``He promises her to clean his room"}}
 
@@
 acquisition could be used. The tool developed in the previous part of
-this project\footnote{??} would be suitable.
+this project\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf} would be suitable.
 All in all, it should be ensured that we use as much information as
 possible from our resources.
-%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
+%The list of not-imported can be used for adding other kinds of lexicons - idioms, have POS-tags.
 %Also to analyse what's wrong, or use the words with a word guesser. The code
@@
 extract data from Talbanken,
-showing how words are normally used. There should also be algorithms for
+showing how words are normally used.
+Lexin\footnote{http://spraakbanken.gu.se/lexin/} provides list of valencies information,
+from which lexical information could be extracted, given that the data is freely available.
+There should also be algorithms for
 automatically adding the information to the lexicon. \\
@@
 
-%TODO Espana -> ~
 \bibliographystyle{apalikeurl}

### 2012-01-13 doc/report/report.tex
report: last comments fixed

@@
 ``You are here more often than {\sc self's} friend."}
-\end{tabular}} \label{ex:ReflAP2} \label{ex:ReflAdv2}
+ \label{ex:ReflAP2} \label{ex:ReflAdv2}
 Adverbs containing reflexive possessive pronouns may further not be used
@@
 describe any language. It thus would be surprising if we could cover
-all syntactical aspects of one language while leaving the abstract grammar
+all syntactical aspects of Swedish while leaving the abstract grammar
 unchanged.
@@
 The dependent types introduced in section \ref{sec:reflexives} could probably be
-useful.
+useful to solve this problem.
 However a carefully prepared analysis and motivation is needed, %ationamount in a bigger reconstruction which therefore
@@
 \begin{verbatim}GenNP : NP -> Quant ;\end{verbatim}
-It selects the possessive form given in the \verb-NP- and changed its type
+It selects the possessive form given in the \verb-NP- and changes its type
 to \verb-Quantifier-. 
@@
 \item\verb-GenNP-
-introduced a cycle: a \verb-Quant- may be turned into a \verb-NP-, and now also
+introduced a cycle: a \verb-Quant- may be turned into a \verb-NP-, and now
 a \verb-NP- can be turned into a \verb-Quant-.\\
@@
 is related to pronouns. 
-As noted in section \ref{sec:saldoRes} our GF grammar only classifies personal pronouns
+As noted in section \ref{sec:gfnouns} our GF grammar only classifies personal pronouns
 as pronouns. Many pronouns could not be imported from Saldo, since the correct category
@@
 the same ways as `sådana' (\emph{`such'}) that acts either as a
-determiner, as in sentence (\ref{ex:QuantDet}), or as an adjective as
- in sentence (\ref{ex:QuantAdj}). 
+determiner (\ref{ex:QuantDet}), or as an adjective agreeing with the noun
+(\ref{ex:QuantAdj}-c). 
 \eenumsentence{
@@
 {``He's like that"}
+\item\shortex{3}
+{Huset& är& \label{ex:QuantAdj2}sådan\textbf{t}}
+{\emph{the house}& \emph{is}& \label{ex:QuantAdj2}\emph{such}}
+{``The house is like that"}
 }
@@
 Talbanken contains much valuable information about phrase structures and word 
-usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974).% which is 
+usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974).
+% which is 
 %>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
@@
 Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
-While the POS-tags are reserved for words, the categories give part of sentence information,
+While the POS-tags are reserved for words, the categories give grammatictal information,
 such as \verb|S|, \verb|NP| and \verb|VP|.
-The edge labels give the grammatical functions: \verb|SS| for subject, 
+The edge labels give the
+part of sentence: \verb|SS| for subject, 
 \verb|OO| for object etc. Nearly 400 different tags are used, of which more than
@@
 \section{The translation process}
-The implementation consists of a set of rules that translates
-one node of a Talbanken tree to a GF function, possibly by describing what the
+The implementation consists of a set of rules that translate
+nodes of a Talbanken tree to GF functions, possibly by describing what the
 nodes subtree should look like. 
@@
 
-Lets look at the example in figure
-\ref{pic:tbtree}. The verb \emph{`bli'} is
+%Lets look at the example 
+In figure \ref{pic:tbtree} the verb \emph{`bli'} is
 followed by a predicative 
-adjective modifying the subject, \verb|SP|. GF admits \emph{`bli'} to have
-type \verb|VA|, a verb with an adjective as complement, and the translator can
-apply the GF function \verb|ComplVA| to combine the verb with its object. \\
+adjective modifying the subject tagged with \verb|SP|. Our GF grammar admits
+\emph{`bli'} to have type \verb|VA|, a verb with an adjective as complement,
+and the translator can hence apply the GF function \verb|ComplVA| to combine the verb
+with its object. \\
 
@@
 most transitive verbs can also be used as intransitive. Both sentences in
-example \ref{ex:mappTransitive} are grammatically correct but only (\ref{ex:mappTransitive}b)
-is accepted by the GF grammar, and
+example (\ref{ex:mappTransitive}) are grammatically correct but only (\ref{ex:mappTransitive}b)
+is accepted by the GF grammar: 
 the grammar rules are simply too strict for accepting many Talbanken sentences.
@@
 \label{ex:mappTransitive}
-See section \ref{sec:futureValency} for a longer discussion about this.\\
+%See section \ref{sec:futureValency} for a longer discussion about this.\\
 %the Talbankenwill allow words to be used in other ways than their
@@
 which groups an infinitival verb and its object together into a \verb|VP|.
-This information can however be extracted from the flat version, and the results
-get slightly better when using the flat version. \\
+This information can however be extracted from the flat implementation, and the results
+get slightly better when using this version. \\
 
@@
 the listings are meant for making the text look nice in
-folders etc and outside the scope for the grammar itself. The tags \textbf{XX} and \textbf{NP} are often used since
-Talbanken -- unlike GF -- makes a difference between subjects
-and objects. %(which is not needed in GF (see section \ref{sec:swegf}))
+folders etc and outside the scope for the grammar itself. The tags \textbf{XX}
+and \textbf{NAC} are often used since Talbanken makes a
+difference between subject
+and object noun phrases. %(which is not needed in GF (see section \ref{sec:swegf}))
 The analysis of elliptical expression in (\ref{sent:krav})
@@
 ``Too high demands."}\label{sent:krav}
-contains the tags \textbf{XX} and \textbf{NP}, since it is not obvious
+contains the tags \textbf{XX} and \textbf{NAC}, since it is not obvious
 whether the noun phrase is used as subject or an object.
@@
 Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
-where idiomatic expressions are more likely and the language presumably less strict than
-in Wall Street Journal.
+where idiomatic expressions are more likely the be appear and the language
+presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
 Also, the Penn Treebank contains a lower number of tags, 82 compared
@@
 Besides parsing, the grammar may well be used for language generation,
-which is fast even with an extensive lexicon.
+which works fast even when using an extensive lexicon.
 %TODO make nice!
-Although it is not been verified, we believe that the majority of the
+Although it is not been formally verified, we believe that the majority of the
 sentences generated are grammatically correct in a syntactical point of view.
@@
 {``I'm getting hungry"}}\label{sent:complVV1}
-Since there is no information about which verbs that allows the infinitive mark
-to be left out, it will also allow the more questionable sentence (\ref{sent:complVV2}b).
+Since there is no information in the grammar or lexicon about which verbs that
+allows the infinitive mark to be left out, it will also allow the more
+questionable sentence (\ref{sent:complVV2}b).
 \eenumsentence{
@@
 {~~\emph{I} & \emph{leave} &\emph{be} & \emph{watch}}}\label{sent:complVV2}
-As these functions provide stylistic flexiblity when parsing, they can be left out
-when generating language, and the grammar then provides text of good
+As these functions serve to provide stylistic flexiblity when parsing, they can be left out
+when generating language, and the grammar then generates text of good
 grammatical quality. \\
@@
 during this project.
-The first grammar could parse almost half of the sentences, the result for the final
+The first grammar could parse less than half of the sentences, the result for the final
 grammar was 66 \%. 
@@
 %All imaginable grammatical constructs cannot be cover by our grammar.
-Some work must be left for the robustness, but there are specific constructs
+Even as some work must be left for the robustness, there are specific constructs
 that are desirable and interesting to implement. A few examples are listed here:\\
@@
 The parser needs to be able to handle
-ellipses and long sentences, names and numbers,
-compound words and unknown constructions.
+ellipses and long and complex sentences, names and numbers,
+compound words and other unknown constructions.
 
 \subsubsection{Parallel parsing/Chunk parsing}
-Instead of parsing a whole sentence in one go, we would like to
-parse parts of it in parallel and then try to combine the pieces.
+Instead of parsing a whole sentence in one go, we intend
+parse parts of it in parallel and then try to combine the pieces, building
+the sentence bottom up.
 This method should be faster then normal parsing and would
@@
 from which lexical information could be extracted, given that the data is freely available.
-There should also be algorithms for
+We also need be algorithms for
 automatically adding the information to the lexicon. \\
@@
 as described in section \ref{sec:future},
-we still believe that it is possible to achieve the goal. 
+we still believe that it is possible to achieve our goal. 
 \\
@@
 All parts of the project are open-source and may thus be used in other applications.
-The grammar may be beneficial also when working with controlled languages,
+The grammar and the lexicon may be beneficial also when working with controlled languages,
 as it increases the coverage of the Swedish resource grammar.

### 2012-01-13 doc/report/report.tex
planning

@@
 \abstract{
-The goal of this thesis is to enable open-domain parsing for Swedish
-by using the GF \\grammar formalism,
+This thesis describes the work to develop a wide-coverage grammar for
+parsing and generating Swedish text.
+We do this by using the dependently typed grammar formalism GF,
+a functional programming language specialized at describing grammars.
+The idea is to 
+%The goal of this thesis is to enable open-domain parsing for Swedish
+%by using the GF \\grammar formalism,
 %by %The idea has been to% 
-combining existing language technology resources with
-new techniques.
-The work includes the development of a covering grammar,
+combine existing language technology resources with
+new techniques and our motivation is the goal of achieving a parser
+for unrestricted Swedish.
+To reach this goal, problems of computational as well as linguistic nature had
+to be solved.  The work includes the development of the grammar,
 %in the GF formalism
-to identify and formalize \\
+to identify and formalize 
 grammatical constructions %specific to and
-frequent in Swedish. % have been investigated and formalised in the functional
+frequent in Swedish, as well as methods for importing a large-scale lexicon and to evaluate
+the parser. % have been investigated and formalised in the functional
 %programming language.
-We compare and evaluate the possibilities of using GF
-for modelling large-scale grammars and for 
+%~~achieve a covering Swedish~~ grammar implemented
 %and the aim to reaching our goal;
-obtaining a system for analysing written text.
-To reach this goal, problems of computational as well as linguistic nature had to be solved.
-%~~achieve a covering Swedish~~ grammar implemented
 %TODO first sentence should be more, no understatements!
@@
 of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
-combining our rule-based grammar with statistical methods.
+combining our rule-based grammar with statistical means.
+We compare and discuss the advantages and problems of using GF
+for modeling large-scale grammars.
+%and for obtaining a system for analysing written text.
 \\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
@@
 not reproduced in the parse tree. Hence, parse trees do not give complete representations
-but model the parse resluts in a transparent manner. For our example, the definiteness and
+but model the parse results in a transparent manner. For our example, the definiteness and
 number of the noun `\emph{katten}' is shown as \verb-DetQuant DefArt NumSg- in
@@
 %TODO change word? uppskatta, känna, få
-The word `uppskatta' for example, has two entries, with sligthly
+The word `uppskatta' for example, has two entries, with slightly
 different semantics: \emph{enjoy}
@@
 finite verb and the rest of the phrase.
-At the code level, this is done using the record type for \verb-VP-, wich
+At the code level, this is done using the record type for \verb-VP-, which
 consists of fields that can be put %the \verb-VP- fields 
@@
 Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
-While the POS-tags are reserved for words, the categories give grammatictal information,
+While the POS-tags are reserved for words, the categories give grammatical information,
 such as \verb|S|, \verb|NP| and \verb|VP|.
@@
 \item a large-scale GF lexicon and a program to redo the importation when needed
-\item an extended grammar covering an importand part of Swedish
+\item an extended grammar covering an important part of Swedish
 \item a comparison between GF and another annotation
@@
 {~~\emph{I} & \emph{leave} &\emph{be} & \emph{watch}}}\label{sent:complVV2}
-As these functions serve to provide stylistic flexiblity when parsing, they can be left out
+As these functions serve to provide stylistic flexibility when parsing, they can be left out
 when generating language, and the grammar then generates text of good

### 2012-01-13 webpage/Main.hs

@@ -2,92 +2,114 @@
  
 import Network.Wai
 import Network.Wai.Handler.Warp
-import Network.HTTP.Types (status200)
-import Network.Wai.Application.Static
 import Network.HTTP.Types
 import Blaze.ByteString.Builder (copyByteString)
 import qualified Data.ByteString.UTF8 as BU
+import qualified Data.ByteString.Char8 as BC
 import qualified Data.ByteString as B
-import System.Process
 import System.IO
-import Text.XHtml.Transitional
 import Text.XHtml
 import Data.Monoid
 import Data.Enumerator (run_, enumList, ($$))
-import Control.Monad
 import Data.Maybe
 import Control.Monad.IO.Class
+
+import ParseLex
  
 main = do
     let port = 3000
     putStrLn $ "Listening on port " ++ show port
-    --run port app
-    run port parseF
+    gr <- play
+    run port (parseF gr True)
 
---parseF :: Request -> IO Response
-parseF req = do
+mainSmall = do
+    let port = 3000
+    putStrLn $ "Listening on port " ++ show port
+    gr <- return (undefined,undefined)
+    run port (parseF gr False)
+
+
+parseF :: MonadIO m => ParseData -> Bool -> Request -> m Response
+parseF gr b req = do
   liftIO $ putStrLn "start"
   let mn = queryString req
-  x <- liftIO $ findText req mn 
+  x <- liftIO $ findText gr req mn b 
   liftIO $ putStrLn "have returned"
   return x
 
-findText :: Request -> Query -> IO Response
-findText req mn = 
-         do let im = lookup "img" mn
-            case im of
-                Just (Just x) -> do
---                                 putStrLn "read file"
---                                 fil <- B.readFile  "images/quote_1.gif" --Process "cat" ["images/bg.jpg"] []
---                                 putStrLn $ show fil
---                                 putStrLn "read file -done"
-                                 hSetEncoding stdin latin1
-                                 return $
-                                   ResponseFile status200 [("Content-Type", "image/gif")] 
-                                         "/home/malin/Swedishproject/webpage/images/quote_1.gif" Nothing
-                                   --ResponseBuilder status200 [("Content-Type", "image/gif")] 
[…]

### 2012-01-13 testSuites/Gui.hs
small changes everywhere

@@ -84,8 +84,8 @@ main = do
                 putStrLn $ "Appended to notebook: " ++ show rP ++ " and " ++ show rA
                 notebookSetTabReorderable notebook scrollerP True
                 notebookSetTabReorderable notebook scrollerA True
-                notebookSetCurrentPage notebook rP
                 widgetShowAll notebook
+                notebookSetCurrentPage notebook rP
                 return [scrollerP,scrollerA]
            Nothing -> return []
        mapM_ maxScrollerV new

### 2012-01-13 testSuites/ParseLex.hs

@@ -99,7 +99,7 @@ parseNormal pgf morpho lang s = do
        getBest ts = fromMaybe (head ts) (rank ts)
        rank :: [Tree] -> Maybe Tree
        rank = listToMaybe . filter (not .(\x -> any (`isInfixOf` x) bads) . showExpr [])
-       bads = ["NumCard","man_nn","faar_nn","en_nn","foer_nn"
+       bads = ["NumCard","MassNP","Top","man_nn","faar_nn","en_nn","foer_nn"
               ,"mycken_nn","nyy_nn","vill_ab","titt_nn","leva_nn","haer_nn"
               ,"PrepCN"]
 

### 2012-01-15 doc/report/report.tex
report corrections

@@
 
+\newenvironment{changemargin}[2]{%
+\begin{list}{}{%
+\setlength{\topsep}{0pt}%
+\setlength{\leftmargin}{#1}%
+\setlength{\rightmargin}{#2}%
+\setlength{\listparindent}{\parindent}%
+\setlength{\itemindent}{\parindent}%
+\setlength{\parsep}{\parskip}%
+}%
+\item[]}{\end{list}}
+
 \begin{document}
@@
 \abstract{
-This thesis describes the work to develop a wide-coverage grammar for
+This thesis describes work towards a wide-coverage grammar for
 parsing and generating Swedish text.
@@
 %by %The idea has been to% 
-combine existing language technology resources with
-new techniques and our motivation is the goal of achieving a parser
+combine existing language resources with
+new techniques with an aim to achieve a parser
 for unrestricted Swedish.
 To reach this goal, problems of computational as well as linguistic nature had
-to be solved.  The work includes the development of the grammar,
+to be solved.  
+\\The work includes the development of the grammar --
 %in the GF formalism
@@
 grammatical constructions %specific to and
-frequent in Swedish, as well as methods for importing a large-scale lexicon and to evaluate
+frequent in Swedish -- as well as methods for importing a large-scale lexicon and for evaluating
 the parser. % have been investigated and formalised in the functional
@@
 %in GF -- to model natural language by functional programming. 
-\\We present the methods and technologies used and further inspect the possibilities
-of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
-combining our rule-based grammar with statistical means.
-We compare and discuss the advantages and problems of using GF
-for modeling large-scale grammars.
+We present the methods and technologies used 
+and discuss the advantages and problems of using GF
+for modeling large-scale grammars and 
+how our long-term goal can be reached 
+%the possibilities
+%of reaching our long-term goal % -- to accomplish a parser for unrestricted Swedish -- 
+by combining our rule-based grammar with statistical means.
+%We compare and discuss .
 %and for obtaining a system for analysing written text.
 \\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
-to GF notation and an extended Swedish grammar implementation. The grammar is based
-on the multilingual abstract syntax given in GF resource library, and now
+into the GF notation and an extended Swedish grammar implementation. The grammar is based
+on the multilingual abstract syntax given in the GF resource library, and now
 also covers constructions specific to Swedish.
@@
 reasoning within the programming language.
-GF has strong support for multilinguality and has so far been successfully used for 
+GF has strong support for multilinguality and has so far been used successfully for 
 controlled languages \cite{cnl}. Recent experiments have showed
@@
 Parsing, that is the task of automatically identifying morphological and
-syntactical structures, is subject to increasing interest, especially
+syntactical structures, is receiving increasing interest, especially
 considering the steadily growing amount of data available online. 
@@
 
-To build the grammar from scratch would be not only time consuming but
+\noindent To build the grammar from scratch would be not only time consuming but
 would also mean that already existing systems would have to be reimplemented.
@@
 
-The example in figure \ref{fig:gfAbstract1} shows an abstract grammar defining three categories, % \verb|Categories|,
+\noindent The example in figure \ref{fig:gfAbstract1} shows an abstract grammar defining three categories, % \verb|Categories|,
 one for nouns, one for verbs and one for sentences. The abstract grammar also gives
@@
 %>> move much of this to Swedish GF part?
-Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
+\noindent Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
 for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|.
@@
 \end{wrapfigure}
-This is the case for `flicka', a noun belonging to the first declination. %the correct declination
+
+\noindent This is the case for `flicka', a noun belonging to the first declination. %the correct declination
 %can be guessed by using the first declination. % and noting that the word ends with an `a'. 
@@
 
+\newpage
 \subsection{Frontiers of Grammatical Framework}
@@
 \\
-Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
+
+\noindent Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
 and was then redistributed in an updated version,
@@
 where the trees have been made deeper and more detailed while still containing
-the lexical MAMBA layer. \\
+the lexical MAMBA layer. 
 The Malt parser was trained on section P of Talbanken, and these
@@
 
-The semantic aspect of Saldo requires that words with multiple meanings are
+\noindent The semantic aspect of Saldo requires that words with multiple meanings are
 separated into different entries.
@@
 
+\begin{changemargin}{0cm}{-2cm}
 \section{Related work}
@@
 
-The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
+\noindent The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
 Structure Grammars \cite{hpsg} (HPSG) providing compatibility with tools for
@@
 grammar \cite{stymne}, also implemented using the Lingo Matrix,
-which focuses on describing and translating verb frames.\\ % and translating them to English. \\
+which focuses on describing and translating verb frames.\\ % and translating them to English.
 
@@
 
-The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
+\noindent The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
 gives a full syntactic analysis as well as semantics represented in `Quasi logical form'. A
@@
 
-In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
+\noindent In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
 for English and Korean, but, to our knowledge, not for Swedish.  \\
 
-The ParGram \cite{pargram} project aims at making wide coverage grammars using
+\noindent The ParGram \cite{pargram} project aims at making wide coverage grammars using
 the Lexical Functional Grammar approach \cite{lfg}.
@@
 
+\end{changemargin}
 %However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
@@
 
-During this project, the program has been made more robust than the previous
+\noindent During this project, the program has been made more robust than the previous
 version. It also prints log files providing
@@
 %compared to the total size of Saldo this is .
-When testing the coverage of Talbanken,
+\noindent When testing the coverage of Talbanken,
 we found that there are around 2500 word forms still missing, excluding the ones
@@
 has to be passed on by an inherent feature of the \verb-CN-, a flag set to tell
-if the \verb-AdjCN- was applied.
+if the \verb-AdjCN- was applied.\\
 \begin{figure}[h]
@@
 %TODO review, verify
-Due to the syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
+\noindent Due to the syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
 contains personal pronouns.
@@
 
-Swedish verbs may take up to five arguments \cite[p. 53]{stymne}.
+\noindent Swedish verbs may take up to five arguments \cite[p. 53]{stymne}.
 These may be prepositions, particles, reflexive object, indirect objects
@@
 \end{figure}
-Even though the verb phrase \emph{``har inte ätit de gula äpplena idag"} is discontinuous, the whole
+\noindent Even though the verb phrase \emph{``har inte ätit de gula äpplena idag"} is discontinuous, the whole
 phrase is still treated as one constituent in GF. The parts
@@
 
-During this project, the s-passive was added although the
+\noindent During this project, the s-passive was added although the
 periphrastic passive is still allowed.
@@
 \enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
+{Hon & erbjöds & jobbet}
+{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
+{``She was offered the job"}}
+\label{ex:passV33}
+\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
 {Jobbet & erbjöds & henne }
@@
 \label{ex:passV32}
-\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
-{Hon & erbjöds & jobbet}
-{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
-{``She was offered the job"}}
-\label{ex:passV33}
 \enumsentence{\textbf{V2A verb}\\\shortex{3}
@@
 sentence (\ref{sent:debocker}).
-\newpage %until behave
 \eenumsentence{
@@
 
-Apart for this, only corrections have been done, exemplified in
+Apart from this, only corrections have been done, exemplified in
 this sentence.
@@
 by using its genitive form, \emph{``kattens'}. This could be used as a determiner,
-\newpage % until behave
 \enumsentence{Det är kattens mat.\\``It is the cat's food."}
@@
 
+The mapping is based on 
+a translation of the English Penn Treebank \cite{gfpenn}.
+By modifying this program, we now have a translation that works for
+the Tiger XML-format of
+Talbanken. Adaption was required for the differences in annotation as well as 
+for the syntactic differences between Swedish and English.
+
 The mapping %connects GF with another annotations and 
@@
 
-The mapping is based on 
-a translation of the English Penn Treebank \cite{gfpenn}.
-By modifying this program, we now have a translation that works for
-the Tiger XML-format of
-Talbanken. Adaption was required for the differences in annotation as well as 
-for the syntactic differences between Swedish and English.
-
 %The mapping works for Talbanken05 in Tiger XML-format.
@@
 If the word is not found in the lexicon
-a \textit{meta} variable will be put in its place, denoted `\verb-?-'.
-The meta variable is also used when we have found two translatable trees
-but no rule that allows them to be combined.
+a \textit{meta variable} will be put in its place, denoted `\verb-?-'.
+The meta variable is also used as a function, needed when we have found two
+translatable trees but no rule that allows them to be combined.
 
@@
 
-The translation covers complex noun phrases which may consist of pronouns,
+\noindent The translation covers complex noun phrases which may consist of pronouns,
 reflexive objects or common nouns and contain predeterminers, 
@@
 \includegraphics[width=80mm]{metatree.pdf}
-\caption{Abstract tree with meta}
+\caption{Abstract tree with meta variable}
 \label{pic:gMetatree}
@@
 
-A similar issue arises from the use of verb particles. A verb with a
+\noindent A similar issue arises from the use of verb particles. A verb with a
 particle is treated as two independent units in Talbanken but in GF they are
@@
 
-When evaluating the mapping, the results strongly depend on which restrictions we
+\noindent When evaluating the mapping, the results strongly depend on which restrictions we
 put on the input. 
@@
 
-When it comes to parsing, we do not get far without robustness.
+\noindent When it comes to parsing, we do not get far without robustness.
 The grammar in itself is by no means robust, and just one 
@@
 
-By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
+\noindent By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
 added many of the commonly used words that were missing from the older
@@
 
+\newpage
 \section{Future Work}
@@
 
-\subsection{Enhancements and complements}
+\subsection{Enhancements and improvements}
 \subsubsection{Grammar}
@@
 from which lexical information could be extracted, given that the data is freely available.
-We also need be algorithms for
-automatically adding the information to the lexicon. \\
+This is also a source of information helping us separate between verbs that can
+be used without an infinitive marker from the ones that cannot, cf. (\ref{sent:complVV2}a-b).
+However, Lexin is relatively small and does not contain more than 10 000 verbs.
+%Furthermore, we also need be algorithms for
+%automatically adding the information to the lexicon. \\
+
 Many Swedish verbs can be used with different number of arguments.
@@
 %that the work done is both interesting and useful.
-All parts of the project are open-source and may thus be used in other applications.
+\noindent All parts of the project are open-source and may thus be used in other applications.
 The grammar and the lexicon may be beneficial also when working with controlled languages,
@@
 %Hard, but not finished! 
-
+\begin{changemargin}{-0.1cm}{0cm}
 \bibliographystyle{apalikeurl}
 \bibliography{masterbib}
+\end{changemargin}
 \end{document}

### 2012-01-17 doc/report/report.tex
last report?

@@
 combine existing language resources with
-new techniques with an aim to achieve a parser
+new techniques, with an aim to achieve a parser
 for unrestricted Swedish.
@@
 to be solved.  
-\\The work includes the development of the grammar --
+The work includes the development of the grammar -- \;
 %in the GF formalism
@@
 and discuss the advantages and problems of using GF
-for modeling large-scale grammars and 
-how our long-term goal can be reached 
+for modeling large-scale grammars.
+We further discuss how our long-term goal can be reached
 %the possibilities
@@
 \vspace{5mm}
-Many people have helped me during this project and made work thesis possible.
+Many people have helped me during this project and made this work possible.
 I would first of all like to thank
@@
 the framework is one of the main component in the
-European project MOLTO \footnote{http://www.molto-project.eu/}, which develop
+European project MOLTO\footnote{http://www.molto-project.eu/}, which develop
 online translation between to 15 languages.\\
@@
 \end{figure}
+\newpage % until behave
 
@@
 
-We now have two implementations of the abstract, one for Swedish and one for English.
+\noindent We now have two implementations of the abstract, one for Swedish and one for English.
 The resulting GF grammar is able both to
@@
 ``I enjoy theater."}
-or \emph{estimate}. 
+\newpage % until behave
+or \emph{estimate}: 
 \enumsentence{Värdet uppskattades till 5 miljoner.\\
@@
 %different grammatical analyse, a lot of pronouns. Example. 
-Most but not all words of these categories have been imported.
+\noindent Most but not all words of these categories have been imported.
 One reason why the importing phase would fail 
@@
 %       >> give reasons! should look good, not bad!
-A list of words that were given different labels in GF than in Talbanken has been
+\noindent A list of words that were given different labels in GF than in Talbanken has been
 composed, consisting of about 1600 entries. Many of those are
@@
 %The lexicon does not contain any information.
-Valency information, which is crucial for GF, is not given in Saldo and
+\noindent Valency information, which is crucial for GF, is not given in Saldo and
 hence not in the imported lexicon. It remains as future work to find methods 
@@
 %uses the post-modified form of the noun: katten.
-Some rules for determination were described in section \ref{sec:swedishnoun};
+\noindent Some rules for determination were described in section \ref{sec:swedishnoun};
 if we have a common noun phrase consisting of the parts \emph{`liten'}
@@
 \newpage
-As already stated, the visualized parse trees is not a complete representation,
+\noindent As already stated, the visualized parse trees is not a complete representation,
 even if the verb phrases in the visualizations look the same, the two cases
@@
 
-Like verb phrases, a clause may also be missing an object and then has the type
+\noindent Like verb phrases, a clause may also be missing an object and then has the type
 \verb-ClSlash-.
@@
 %coverage of standard Swedish.
-
+\newpage % until behave
 \section{Development of the grammar}
@@
 %%
-
+\newpage %until behave
 \subsection{A second future tense: ``Kommer att"}
@@
 
-The mapping %connects GF with another annotations and 
+The translation %connects GF with another annotations and 
 gives us means to evaluate our parser. By both parsing a Talbanken sentence and
@@
 The particle has therefore no effect on the abstract tree, its presence is
-announced by the verb itself.
+announced by the verb \\itself.
 
@@
 \end{figure}
-
-Some more information about VP are given by the tag \verb|VG| (\emph{verb group}),
+The focus on the mapping has been simple sentences, in order to cover the most
+fundamental structures. If the program is extended and adapted to cover complex
+conjunction, a deeper evaluation should be preformed to see which version that
+suits our needs the best.
+The deeper implementation also gives 
+some more information about verb phrases, by the tag \verb|VG| (\emph{verb group}),
 which groups an infinitival verb and its object together into a \verb|VP|.
@@
 
-Our focus has been on shorter sentences, with no idioms or conjunction.
+The main goal has been to be able to translate shorter sentences, with no
+idioms or conjunction.
 If we assure that the lexicon contains the correct word class for all lemmas
@@
 be used without an infinitive marker from the ones that cannot, cf. (\ref{sent:complVV2}a-b).
-However, Lexin is relatively small and does not contain more than 10 000 verbs.
+However, Lexin is relatively small and does not contain more than 10 000 entries.
 %Furthermore, we also need be algorithms for

### 2012-01-17 talbanktools/ExtractXML.hs

@@ -1,0 +1,117 @@
+import XMLHelp hiding (id,words)
+import qualified XMLHelp as X
+--import SaldoXML
+import Text.XML.HXT.Core hiding (Tree,trace)
+import Debug.Trace
+import Data.List
+import Data.Tree
+import Data.Char
+import Data.Function
+import Data.ByteString.Char8 (pack)
+import Control.Monad.State
+import System.Environment
+ 
+-- For extrating certian combination of tags from Talbanken
+
+mainet f src 
+    = do
+      runX ( xunpickleDocument xpSentences
+                                [withInputEncoding utf8
+                                , withRemoveWS yes] src
+	        >>>
+	        arrIO (return . f))
+
+input = "../Talbanken05_20060604/FPS/P.tiger.xml"
+
+mainEt 	:: String -> String -> IO ()
+mainEt src out
+    = do
+      runX ( xunpickleDocument xpSentences
+                                [withInputEncoding utf8
+                                , withRemoveWS yes] src
+	        >>>
+	        arrIO (return  . concatMap getTags)
+	        >>>
+            xpickleDocument xpSentences
+                            [withIndent yes,
+                             withInputEncoding utf8] (out++".xml"))
+
+      return ()
+main = do
+  (src:out:_) <- getArgs
+  putStrLn "Starting ..."
+  putStrLn "Reading xml ..."
+  xml <- mainEt src out
+  putStrLn "Done!"
+  --let avps = nub $ concatMap (extractTag "AVP") $ map snd $ concat xml
+
+containTag :: String -> Tree String -> Bool
+containTag t (Node cat [tr]) = posIs t tr 
+containTag t _               = False
+  
+-- returns the trees that returns words of type tag.
+-- the result contains the label, the cat, and then a tree containing the tag
+-- we prefer as long branches as possible 
+extractTagDeep :: String -> Tree String -> [Tree String]
+extractTagDeep tag tr@(Node cat [Node lab [Node cat2 ts]]) 
+               | any (containTag tag) ts = [tr]
+extractTagDeep tag tr@(Node cat [Node lab ts]) 
+               | any (containTag tag) ts = [tr]
+extractTagDeep tag (Node t ts) = concatMap (extractTagDeep tag) ts
+
+-- for extracting all trees of type tag.  (all subtrees starting with tag)
+extractTag :: String -> Tree String -> [Tree String]
[…]

### 2012-01-18 doc/report/report.tex
varios small fixes

@@
 %of reaching our long-term goal % -- to accomplish a parser for unrestricted Swedish -- 
-by combining our rule-based grammar with statistical means.
+by combining our rule-based grammar with statistical methods.
 %We compare and discuss .
@@
 GF has strong support for multilinguality and has so far been used successfully for 
-controlled languages \cite{cnl}. Recent experiments have showed
+controlled languages \cite{cnl}. Recent experiments have  however showed
 that it is also possible to use the framework for parsing free language \cite{patent}.
@@
 to visualize the trees and the given commands (fig. \ref{fig:shell}) specify that the
-output format should be PDF-files and that these should be opened by the program
+output format should be PDF and that these files should be opened by the program
 Evince. The user can choose to see the \textit{parse tree} (fig. \ref{pic:gfStree})
@@
 Parse trees on the other hand show only the types assigned to the words and phrases.
-Information about tense, polarity etc. which are explicitly given in the abstract tree are
+Information about tense, polarity etc., which are explicitly given in the abstract tree, are
 not reproduced in the parse tree. Hence, parse trees do not give complete representations
@@
 number of the noun `\emph{katten}' is shown as \verb-DetQuant DefArt NumSg- in
-the abstract tree while the parse tree only show that a the noun phrase consists of one noun. 
+the abstract tree while the parse tree only shows that the noun phrase consists of one noun. 
 % is used but not in the parse tree.

### 2012-01-18 robust/ne/ParseLexin.hs

@@ -1,75 +1,0 @@
-
-import Data.Attoparsec.Char8 as A
-import Control.Monad
-import Control.Applicative
-import Data.ByteString.Char8 as B 
-import Data.ByteString.UTF8 as BU
-
-parseStuff = parseOnly verbValency . fromString
-
-verbValency = do
-  xorA
-  theWord
-  v2
-  
-char' c = do
-  skipSpace
-  char c
-
-string' = string . fromString 
-
-subject = xorA
-
-xorA = do
-    char' 'A' <|> char' 'x' <|> char' 'y' <|> char' 'B'
-    mone (char' '/' >> (char' 'A' <|> char' 'x'))
-
-theWord = do
-   char' '&' 
-
---          reflexive,preposition,particle
-v2 :: Parser (Maybe ByteString,Maybe ByteString,Maybe ByteString)
-v2 = do
-  pa <- maybeP particle 
-  pr <- maybeP (do mone (char' '(') 
-                   skipSpace
-                   A.takeWhile $ notInClass "yzxAB/)( ") -- no prepositions with yzx..
-  rf <- maybeP sig 
-  xorA
-  mone (char' ')')
-  return (rf,trimJust pr,pa)
-
-  
-sig = do
-  char' '('
-  s <- string' "sig"
-  char ')'
-  return s 
-
-particle = do
-  char' '('
-  p <- A.takeWhile $ notInClass "xyzABC/)( "
-  char ')'
-  return p
-
-mone p = (p >> return ()) <|> return ()
-
-maybeP :: Parser a -> Parser (Maybe a)
-maybeP p = do
-   (p >>= return . Just)
-   <|>
-   return Nothing
-
-trimJust :: Maybe ByteString -> Maybe ByteString
-trimJust (Just x) | B.null x = Nothing
-trimJust x         = x
-
-{-
-    word <- str
-    char' ','
-    res <- liftM ((,) word) (list lemma)
-    char' ')'
-    return res
-    -}
-
-

### 2012-01-18 webpage/Main.hs

@@ -38,7 +38,7 @@ parseF gr b req = do
   let mn = queryString req
   id <- return "theMap" --get userid somehow! mkDir theMap, images/theMap
   liftIO $ print ("pathinfo: "++show (pathInfo req))
-  liftIO $ print ("query: "++show mn)
+  liftIO $ print ("all requst: "++show req)
   x <- liftIO $ findText id gr req mn b 
   liftIO $ putStrLn "have returned"
   return x

### 2012-01-19 doc/report/report.tex
last report update befor submit

@@
 GF has strong support for multilinguality and has so far been used successfully for 
-controlled languages \cite{cnl}. Recent experiments have  however showed
+controlled languages \cite{cnl}, while recent experiments have showed
 that it is also possible to use the framework for parsing free language \cite{patent}.
@@
 categories. That is,
-we know that the first argument of the function \verb-PredVP- acts as the subject.
-The noun phrase itself does not carry information about its semantic role.
+we know that the first argument of the function \verb-PredVP- acts as the
+subject, but
+the noun phrase itself does not carry information about its semantic role.
 Until it is given as an argument to a clause level function, no difference

### 2012-01-19 webpage/Main.hs
autocompletion to server

@@ -61,6 +61,10 @@ findText id gr req mn b
   | (Just _) <- lookup "firstform" mn = do
       putStrLn "have nothing, will return textfield"
       return inputForm
+  | (Just (Just txt)) <- lookup "input" mn = do
+       putStrLn $ "will complete "++ show txt
+       autoComplete txt
+
 findText id gr req mn b = do
   let path = map (filter (/='\"')) $ map show $ pathInfo req
   liftIO $ print (pathInfo req)
@@ -128,6 +132,14 @@ parseMore id txt = do
   return $
     ResponseBuilder status200 [("Content-Type", "text/html")] $ mconcat 
          $ map copyByteString html
+
+
+autoComplete :: B.ByteString -> IO Response
+autoComplete txt = do
+  w <- getNextWord (BU.toString txt) 10
+  putStrLn $ "autoComplet returns" ++ unwords w
+  return $ ResponseBuilder status200 [("Content-Type", "text/html")] $ mconcat $ map copyByteString $
+              ["<p>",BU.fromString "Alternatives:",BU.fromString $ unwords w,"</p>"]
 
 
    

### 2012-01-19 webpage/ParseLex.hs

@@ -4,6 +4,7 @@ module ParseLex
   ,processparse
   ,testa
   ,smallparse
+  ,getNextWord
   ,ParseData
   ,UserId) where
 import System.Timeout
@@ -17,6 +18,8 @@ import Control.Arrow
 import PGF
 import XMLHelp hiding (words,parse,id)
 import MkLex
+import qualified Data.Map as Map     
+import Data.Function
 
 -- use with tee -a outfile
 data Parsed = Pars {t :: Tree, s :: String, n :: Int}
@@ -154,4 +157,44 @@ pipeIt2graphviz id pgf lang t i = do
     writeFile dotFileA $ graphvizAbstractTree pgf (True,True) t
     readProcess "dot" ["-Tpng",dotFileA,"-o","images/"++pngFileA] []
     return (pngFileP,pngFileA)
+ 
+----------------------
 
+complete' :: PGF -> Language -> Type -> Maybe Int -> String
+         -> (BracketedString, String, [String])
+complete' pgf from typ mlimit input =
+  let (ws,prefix) = tokensAndPrefix input
+      ps0 = PGF.initState pgf from typ
+      (ps,ws') = loop ps0 ws
+      bs       = snd (PGF.getParseOutput ps typ Nothing)
+  in if not (null ws')
+       then (bs, unwords (if null prefix then ws' else ws'++[prefix]), [])
+       else (bs, prefix, maybe id takeBest mlimit $ order $ Map.keys (PGF.getCompletions ps prefix))
+  where
+    order = delete nonExist . sortBy (compare `on` nicety)
+    takeBest i xs | last input == ' ' = take i $ map head $ groupBy ((==) `on` (take 1)) xs
+                  | otherwise         = take i xs
+
+    tokensAndPrefix :: String -> ([String],String)
+    tokensAndPrefix s | not (null s) && isSpace (last s) = (ws, "")
+                      | null ws = ([],"")
+                      | otherwise = (init ws, last ws)
+        where ws = words s
+
+    loop ps []     = (ps,[])
+    loop ps (w:ws) = case PGF.nextState ps (PGF.simpleParseInput w) of
+                       Left  es -> (ps,w:ws)
+                       Right ps -> loop ps ws
+
+    nonExist = "#\191@\167X?X&%/"
+    nicety (i:xs) | not (isAlpha i)  = ('\247':i:xs)
+                  | otherwise        = map toLower (i:xs)
+
[…]

### 2012-01-30 doc/report/report.tex
cover and pics to report

@@
 \author{Malin Ahlberg}
-\maketitle
+%\maketitle
 \newpage
@@
 \centering
-\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
+\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTreec.pdf}}
 \subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
@@
 \end{figure}
-\vspace{10mm}
+%\vspace{10mm}
 
@@
 \begin{wrapfigure}{l}{0.35\textwidth}
-\includegraphics[width=30mm]{gfETree.png}
+\includegraphics[width=30mm]{gfETreec.pdf}
 \caption{English parse tree}
@@
 \centering
-\subfloat[Morphological info for `katt']{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
+\subfloat[Morphological info for `katt']{\label{pic:saldoTab}\includegraphics[width=40mm]{saldoTab.png}}
 \hspace{20mm}
-\subfloat[Graph showing the hyponyms of `katt']{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
+\subfloat[Graph showing the hyponyms of `katt']{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldoGraph.png}}
 \caption{Information in Saldo}
@@
   \begin{center}
-\subfloat[Saldo]{\label{pic:saldoVax}\includegraphics[height=90mm]{saldoVax.png}}
+\subfloat[Saldo]{\label{pic:saldoVax}\includegraphics[height=80mm]{saldovax.png}}
 \hspace{5mm}
-\subfloat[GF]{\label{pic:gfVax}\includegraphics[height=60mm]{gfVax.png}}
+\subfloat[GF]{\label{pic:gfVax}\includegraphics[height=60mm]{gfvax.png}}
 \caption{Inflection table for the verb `växa' (grow)}
@@
 \hspace{-18mm}
-\includegraphics[height=130mm]{categories.png}
+\includegraphics[height=130mm]{Categories.png}
 \caption{The types of the GF resource grammars. }
@@
 \begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=70mm]{tittapa.jpg}}
+\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=70mm]{tittapac.pdf}}
 \hspace{10mm}
-\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=76mm]{tittapart.pdf}}
+\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=70mm]{tittapartc.pdf}}
 \caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
@@
 and for preposition stranding:
-\enumsentence{
-~~Stranded preposition \hspace{45mm} cf.\\
-\begin{tabular}{ll}
+\eenumsentence{
+\item
+Stranded preposition \\ %\hspace{45mm} cf.\\
+%\begin{tabular}{ll}
 {\begin{tabular}{llllll}
@@
 \emph{who}& \emph{must}& \emph{I}& \emph{watch out}& \emph{me} &\textbf{\emph{for}}? \\
-\end{tabular}}&
-{\begin{tabular}{llllll}
+\end{tabular}}
+\item
+{cf.\\
+\begin{tabular}{llllll}
 \textbf{För}& vem& måste& jag& akta& mig?\\
@@
 ~~`Who do I need to watch out \textbf{for}?' &~ `\textbf{For} whom do I need to watch out?'\\
-\end{tabular}}
+%\end{tabular}}
+}
 
@@
 Instead of parsing a whole sentence in one go, we intend
-parse parts of it in parallel and then try to combine the pieces, building
-the sentence bottom up.
+parse parts of it in parallel and then try to combine the pieces into
+a parse tree for the whole sentence.
 This method should be faster then normal parsing and would

### 2012-01-31 robust/valens/ParseLexin.hs
last final version? of report

@@ -41,11 +41,11 @@ data VerbType = VT {vtype :: V , argument :: [Argument], preps :: [Preposition]}
   deriving (Show)
    -- X is unparseble type
 data V = V | V2 | V3 | X | VV Bool | VS | VQ | VA | V2S | V2Q | V2A -- V2V, very uncommon 
-  deriving (Show)
+  deriving (Show,Eq)
 data Argument = Part Text  --particles
               | Refl             --is reflexive
               -- | Inf Bool         --infinitival verb, True if inifinitive marker is used
-  deriving (Show)
+  deriving (Show,Eq)
 type Preposition = Maybe Text
 
 v :: Parser VerbType

### 2012-01-31 robust/valens/RunProg.hs

@@ -1,16 +1,19 @@
-import Control.Monad
 import Control.Applicative
 import Control.Arrow
 import Control.DeepSeq
+import Control.Monad
+import Data.Function
 import Data.List
 import Data.List.Utils
-import Data.Maybe
 import Data.Map as M hiding (map, delete, filter)
+import Data.Maybe
 import Data.Ord
 import qualified Data.Text as T 
 import Debug.Trace
-import System.IO
 import PGF
+import System.IO
+import System.IO.Unsafe
+import System.Process
 
 import ParseLexin
 
@@ -50,37 +53,41 @@ testa = do
 
 main = do
    putStrLn "Parsing lexin..."
+   runCommand "rm TestAbs.gf TestCnc.gf log"
    vs <- getLexin
-   trace (show vs) $ putStrLn "Reading pgf..."
+   putStrLn "Reading pgf..."
    pgf <- readPGF pgfFile
    putStrLn "Building morpho..."
    let morpho = buildMorpho pgf lang
    putStrLn "Creating map of lemmas..."
    lex <-  mkLexMap
    putStrLn "Extracting new dictionary..."
-   mapM (\v -> mkLexicon morpho pgf lex v >>= writeCode) vs
+   mapM_ (\v -> mkLexicon morpho pgf lex v >>= writeCode) vs
    cleanDicts
    return ()
 
 -- to avoid name clashes
 cleanDicts :: IO ()
 cleanDicts = do
-  abs <- getAndOrder newabs
-  cnc <- getAndOrder newcnc
-  writeFile "NewAbs.gf" abs
+  abs  <- getAndOrder newabs
+  cnc  <- getAndOrder newcnc
   writeFile "NewCnc.gf" cnc
+  writeFile "NewAbs.gf" abs
   writeGF $!! (abs,cnc)
 
  where getAndOrder :: FilePath -> IO String
        getAndOrder file = do
-            lst <- extractLex tail file   -- should remove replicate code
-            let sortIt = concatMap addIndicies . group . sort
+            lst <- extractLex tail file   
+            let sortIt = concatMap addIndicies . groupBy ((==) `on` fst) . f . sort
             return $ unlines $ map format $ sortIt lst
+         where f | file =="TestCnc.gf" = trace "nubs" nub   --remove duplicates (in cnc only..)
+                 | otherwise          = id 
        addIndicies [x] = [x]
        addIndicies xs = addI 1 xs
        -- TODO do not add index at the very end (V21 ?)
        --      empty files before rewriting
[…]

### 2012-02-06 doc/paper/Paper.tex
paper without old comments

@@
 
-\title{%First results of developing a Swedish parser based on a wide-coverage grammar\\
-        Combining language resources into a grammar-driven Swedish parser} 
-        %or Bootstrapping a Swedish parser \\
-        %or A grammar-based Swedish parser from already existing resources}
+\title{ Combining language resources into a grammar-driven Swedish parser} 
 \author{Malin Ahlberg \qquad\qquad Ramona Enache
-%\and
-%Co Author \qquad\qquad 
+
 \institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
@@
 \begin{abstract}
-This paper describes on-going work on a rule-based, open-source
-parser for Swedish. The central component is a wide-coverage grammar
-implemented in the GF formalism (Grammatical Framework), from which
-we obtain a %can be compiled to a % from which we derive?? 
-parser.
-In addition to GF, we use two other main resources: the Swedish
-treebank Talbanken and the electronic SALDO lexicon.
-The resulting parser 
-gives a full syntactic analysis of the input sentences.
-It will be highly reusable, freely available,
-and as GF provides libraries for compiling
-grammars to a number of programming languages,
-chosen parts of the the 
-grammar may be used in various NLP applications. % dealing with
-%controlled or free
-%natural language. 
-% allow easy interaction with other programs and grammars
-%>> something very commercial, interesting and outstanding here please!
+This paper describes on-going work on a rule-based, open-source parser for
+Swedish. The central component is a wide-coverage grammar implemented in the GF
+formalism (Grammatical Framework), from which we obtain a parser. In addition
+to GF, we use two other main resources: the Swedish treebank Talbanken and the
+electronic SALDO lexicon.
+The resulting parser gives a full syntactic analysis of the input sentences.
+It will be highly reusable, freely available, and as GF provides libraries for
+compiling grammars to a number of programming languages, chosen parts of the the 
+grammar may be used in various NLP applications.
 
@@
 thereby investigate how GF can be used for open-domain parsing.
-% better 
 We compile a large-scale grammar to a parser
@@
 be evaluated by experts. \\
-%s and by comparing the results of parsing sentences in a commonly used Swedish treebank, Talbanken.
 We chose our resources so that both our grammar and parser can 
@@
 and describe our results in Section \ref{sec:progress}.
-%We aim to implement % The goal is to be able to 
-%a parser for open domain language and 
-%This paper describes our work and results 
-%so far as well as the directions of future work.
 
@@
 
-%städa upp detta stycke!!
+Grammatical Framework\cite{ranta-2011} (GF) is a grammar formalism based on
+functional programming.
 
-Grammatical Framework\cite{ranta-2011} (GF) is a grammar formalism based on functional
-programming. %, designed for multilingual applications.
-
-The key idea is to divide a grammar into abstract and concrete parts. %% parts?
+The key idea is to divide a grammar into abstract and concrete parts. 
 The abstract grammar gives a logical representation of the semantics,
-% shorten this part
 modeled as abstract trees.
-%Complications occurring in natural languages, such as agreement, case and word
-%order, are abstracted away. 
-%That is, 
-%It contains no actual linguistic information, but
-%information about what categories are used and how to combine them into trees.
 The concrete grammars tell how to translate the abstract trees to a given language,
 and deal with issues such as word order, case and agreement. 
-%implements the abstract 
-%To add a new language it amounts to give a concrete grammar, that implements
-%the abstract grammar and describes how to translate abstract
-%trees into strings of the language.
-The framework enables us to parse strings into 
-abstract trees as well as linearize trees into strings.
-%Since all languages supported in an application share a common abstract syntax,
-%we get multilinguality for free.
+The framework enables us to parse strings into abstract trees as well as
+linearize trees into strings.
 
-The grammar acts as an independent module and
-%and may be used in different projects.
-reusability is further supported by the separation between resource grammars
-and application grammars. The resource library provided with GF 
-implements morphological and
-syntactical rules for more than 20 different languages.
-Hence the writer of an application grammar can start her work at a higher
-level and does not need to describe how to 
-form standard sentences, phrases or inflect words.
+The grammar acts as an independent module and reusability is further supported
+by the separation between resource grammars and application grammars. The
+resource library provided with GF implements morphological and syntactical
+rules for more than 20 different languages.  Hence the writer of an application
+grammar can start her work at a higher level and does not need to describe how
+to form standard sentences, phrases or inflect words.
 
@@
 they can share much of their implementations. This is usually done by using a
-\verb|Functor|, % , which lets a number of languages share parts of their implementation. 
-avoids code duplication and aids the code maintenance.
+\verb|Functor|, avoids code duplication and aids the code maintenance.
 
@@
 This project takes a different approach by using GF for open domain language.
-%While the resources have to be general and language independent,
-%language specific constructs such as stylistic changes, idioms, informal
-%expressions are given in a special module.
-%This module, together with 
 Using the Swedish resource grammar as our starting
-point, % of this project.
-%By starting from an already existing Swedish grammar written in GF,
-we get a basic description of the language. The framework provides
+point, we get a basic description of the language. The framework provides
 tools such as parsing, generation and
@@
 This part was also used when training the data-driven parser Maltparser \cite{malt}. \\
-%and the parsetree generated by our grammar will eventually be 
-%compared to those from the treebank.
 
@@
 
-%compare!!
 \section{Related work}
@@
 unification-based chart parser\cite{wiren},
-%correct reference?
 though none of them are freely available.
@@
 extraction compatible with GF.
-%For lexical extractions, there are tools like
-%Extract for supervised lexicon extraction and
-%FM for programming lexical resources.
 
@@
 guesses for 25\%.
-%is % estimated
-% whereas the user has to reject the others to force a new guess. %rest are found after an incorrect guess rejected by the user.\\
-%Extract uses a similar 
-%method but defines the paradigms itself.
-%rules are given in another format.
 
-%% Write more here!!
 \subsection{Mapping of trees}
-%R: mapping - very nice - more about it
 The information from the tags in Talbanken can be used for many purposes.
@@
 currently used in and this may be used by the lexical extraction
-tools. % Those can be enhanced if they are given more data. 
+tools.
 The translated trees enable us to extract probabilities for how often
@@
 can restore 85\% of the nodes. 
-%, given that all words are
-%known to the grammar.
 If we lift the restrictions, we can restore 63\% of the tags.
-%mapping, a number that would probably increase substantially when working with a bigger lexicon.
-%is an interesting way of
-%% an straightforward
-%accomplishing an evaluation.
-
-% add evaluation
 
@@
 For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
-languages. % >> When trying to achieve a lightweight, usable code? Not
-%due to the resemblances between the languages, they share
-%>> using a functor for sharing code, for when having the same structure but
-%>> but only different words. 
+languages. 
 However, if we aim for a deeper and more comprehensive analysis of Swedish,
 the implementation of the languages needs to be more independent.
-%<<in order to avoid ad-hoc solutions and 
-%Some constructions that are standard in
-%Norwegian and Danish may be very rare in Swedish and while extending the
-%grammar there may occur situations where a deeper restructuring of the
-%implementation is needed. \\
 The resource grammar gives a good start and our present grammar 
@@
 \emph{Han såg \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
-%>> not critical, does not have to be in resource grammar
-%>> also not expressible in all languages
 Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
-%the possibility for this has been added to the grammar:\\
 \emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
@@
 
-%In order to keep the grammar clean and to keep it syntactically correct,
-%much care is taken to develop it as neatly as possible.
-
 \section{Evaluation and future work}
-%Compared to a statistical parser which operates like a black box,
-%a rule-based one is not only theoretically interesting,
-%but could also give a more explanatory output since the rules
-%are given their names by a human and hence act as informative labels.
 We are building a parser which is based on grammatical rules rather than
@@
 
-%% more explanation of probabilities
 We aim to make the parser robust by equipping it with techniques such as chunk
 parsing, named entity recognition and methods for handling unknown 
-grammatical constructions such as idioms and ellipses. %etc. 
+grammatical constructions such as idioms and ellipses. 
 It is already possible to add probabilities for GF functions,
@@
 the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
-%an expert in Swedish grammar.
 She also evaluates the intermediate results.
@@
 \section{Conclusion}
-%We intend to implement a robust deep parser for Swedish, by first developing
-%a large scale-grammar. 
 By combining GF with a large lexicon, we acquire a deep parser for 
@@
 
-%We have shown how to combine existing resources, we build a large-scale grammar for Swedish.
-%The grammar can be compiled into a
-%parser by GF, and we equip this with a extensive lexicon. 
-%and we will equip this with named entity recognition, a
-%voluminous lexicon, statistical information for disambiguation,
-%a method of parallel chunk parsing etc.
-%So far we have been working on the grammar, the lexical resources and
-%a translation of the treebank Talbanken to GF.
-% adventage to what??
-
 \section{Acknowledgments}
 The work has been funded by Center of Language Technology.
-We would also like to give special thanks to Aarne Ranta, % Ramona Enache,
+We would also like to give special thanks to Aarne Ranta,
 Elisabet Engdahl, Krasimir Angelov, Olga Caprotti, Lars Borin and John Camilleri for their help

### 2012-03-05 doc/paper/PaperFin.tex
work on final version of lrec paper

\documentclass[10pt, a4paper]{article}
\usepackage{lrec2006}
\usepackage{lingmacros}
\usepackage{graphicx}
%\usepackage[utf8]{inputenc}

\title{Combining Language Resources Into A Grammar-Driven Swedish parser}

\name{Malin Ahlberg, Ramona Enache}

\address{ Department of Computer Science \& Engineering, Gothenburg University \\
               Address1 \\
               author1@xxx.yy, author2@zzz.edu, author3@hhh.com\\}

% 150-200 words here
\abstract{This paper describes on-going work on a rule-based, open-source parser for
Swedish. The central component is a wide-coverage grammar implemented in the GF
formalism (Grammatical Framework), a dependently typed grammar formalism
based on Martin-L{\"o}f type theory. GF has strong support for multilinguality and
has so far been used successfully for
controlled languages \cite{cnl} and recent experiments have showed
that it is also possible to use the framework for parsing free language \cite{patent}.
In addition
to GF, we use two other main resources: the Swedish treebank Talbanken and the
electronic SALDO lexicon. By combining the grammar together with
a lexicon extracted from SALDO we obtain a parser accepting all sentences
described by the given rules. We develop and test this
on examples from Talbanken.
The resulting parser gives a full syntactic analysis of the input sentences.
It will be highly reusable, freely available, and as GF provides libraries for
compiling grammars to a number of programming languages, chosen parts of the the 
grammar may be used in various NLP applications. 
\\ \newline \Keywords{GF, Swedish, Parsing}}

\begin{document}

\maketitleabstract

\section{Introduction}
% Copied from Master's Thesis. ok to do that?
Swedish is a North-Germanic language,
closely related to Norwegian and Danish. The languages share most of their
grammatical structures and are mutually intelligible. Swedish is also 
one of the official languages in Finland and altogether spoken by approximately 9
million people.
Swedish syntax is often similar to English, but the  morphology is richer and the
word order slightly more intricate.
It is a verb-second language: the second constituent of a declarative main
clause must consist of a verb.
The first constituent of the clause is usually made up of the subject,
although it likewise could consist of adverbial phrases or objects.
Fronting the finite verb marks questions.

Our goal is to implement a wide-coverage grammar and parser for Swedish
using the GF formalism, and 
thereby investigate how GF can be used for open-domain parsing.
We compile a large-scale grammar to a parser
and combine it with the extensive lexicon SALDO. The Swedish treebank 
Talbanken provides manually tagged trees which we use for improving and evaluating
the grammar. The parser will additionally
be evaluated by experts. \\
We chose our resources so that both our grammar and parser can 
be freely available and open-source.

This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
and describe our results in Section \ref{sec:progress}.

\section{Background}
\label{sec:background}
\subsection{Grammatical Framework}
\label{sec:gf}

Grammatical Framework \cite{ranta-2011} (GF) is a grammar formalism based on
functional programming.

The key idea is to divide a grammar into abstract and concrete parts. 
The abstract grammar gives a logical representation of the semantics,
modeled as abstract trees.
The concrete grammars tell how to translate the abstract trees to a given language,
and deal with issues such as word order, case and agreement. 
The framework enables us to parse strings into abstract trees as well as
linearize trees into strings.

The grammar acts as an independent module and reusability is further supported
by the separation between resource grammars and application grammars. The
resource library provided with GF implements morphological and syntactical
rules for more than 20 different languages.  Hence the writer of an application
grammar can start her work at a higher level and does not need to describe how
to form standard sentences, phrases or inflect words.

Since many of the languages in the GF library resemble each other grammatically,
they can share much of their implementations. This is usually done by using a
\verb|Functor|, avoids code duplication and aids the code maintenance.

GF has so far been used in a number of projects,
MOLTO\footnote{http://www.molto-project.eu/}, TALK \cite{talk}
and WebAlt \cite{webalt} to mention a few. 
All those are special domain applications, dealing with controlled natural
language.
This project takes a different approach by using GF for open domain language.
Using the Swedish resource grammar as our starting
point, we get a basic description of the language. The framework provides
tools such as parsing, generation and
a well-tested interpretation of the parse trees. Furthermore, there are tools
for using GF grammars in a number of programming languages like Haskell
and Java.

\subsection{Talbanken}
For development and evaluation, we use the Swedish treebank
Talbanken \cite{talbanken}.
It was assembled in the 1970s at Lund University and modernized
in 2005 \cite{talbanken05} and
enriched with annotation for a full phrase structure analysis.  \\
Although Talbanken contains both written and spoken Swedish,
only the prose material, consisting of 6316 sentences, will be used in this project.
This part was also used when training the data-driven parser Maltparser \cite{malt}. \\

\subsection{Saldo}
SALDO \cite{saldo} is an open source lexicon resource
based on Svenskt Associationslexikon. It is
developed at Spr\{aa}kbanken at Gothenburg University
and intended for usage in language technology
research. 
We are developing tools for extracting GF lexicons
from SALDO, which have resulted in a dictionary containing
over 98000 entries including information about verbs that
are reflexive or needs a particle.
The importing method should be fast and reliable
enough to allow us to always have a fresh version of the dictionary
in GF.

\section{Related work}
%Copied form thesis!!!! 
% Add: how are these related to this work?
% Add :detaild comparison of outputs of various parsers of Sw.
Many years of research have lead to many interesting language
technology tools for Swedish.
An example is the well-known data-driven Malt parser \cite{malt},
which has been trained on Talbanken. 
There are also a number of grammar-based parsers, although none is freely available.
The cascaded finite state parser CassSwe \cite{casswe} and
The Swedish Constraint Grammar \cite{birn}
give syntactic analyses. 
Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
\cite{fdg},
an extension of the Constraint Grammar
formalism, and produces a dependency structure focusing on finding the nominal
arguments. \\
The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building
Head-Driven Phrase Structure Grammars \cite{hpsg} (HPSG) providing
compatibility with tools for parsing, evaluation, semantic representations etc.
Translation is supported by using Minimal Recursion
Semantics \cite{mrs} as an interlingua. \\
There is a collection of grammars implemented in this framework, giving broad-coverage
descriptions of 
English, Japanese and German. 
The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
Scandinavian, while Norsource (Hellan, 2003) describes Norwegian. A Swedish version
was based upon this (SweCore, Ahrenberg) covering the morphology and some
differences between Swedish and Norwegian. Further, there is the BiTSE 
grammar \cite{stymne}, also implemented using the Lingo Matrix,
which focuses on describing and translating verb frames.\\

The Swedish version of the Core Language Engine (CLE) \cite{gamback}
gives a full syntactic analysis as well as semantics represented in `Quasi logical form'. A
translation to English  was implemented and the work was further developed in the spoken
language translator \cite{spoken}. Unfortunately, it is no longer available. The coverage of the Swedish
CLE is also reported to be very limited \cite[p. 134]{nivreTrees}.\\

In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
for English and Korean, but, to our knowledge, not for Swedish.  \\

The ParGram \cite{pargram} project aims at making wide coverage grammars using
The grammars are implemented in parallel in order to coordinate the analyses of
different languages and there are now grammars for English, German, Japanese and Norwegian.

%old related work:
%Language technology for Swedish is an area of much interesting research.
%There are two other deep grammar-based parser for Swedish:
% put in references!!
%Swedish Core Language Engine (Raynen and Gambäck, 1992) and a
% does this really exist??
%unification-based chart parser \cite{wiren},
%though none of them are freely available.
%Among other parsers for Swedish, the statistical MaltParser \cite{malt},
%trained on Talbanken, is worth mentioning. 
% put in references!!
%CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
% put in references!!
%whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
%a context free grammar. Both GTA and CassSwe operate on POS tagged text.\\
%Extract and FM \cite{MarkusForsberg2007} are tools for supervised lexicon
%extraction compatible with GF.

%\section{Results and progress}
%\label{sec:progress}

\subsection{Extracting a large lexicon}
The lexicon provided with the GF resources is far too small for open-domain
parsing.
This section describes the process of importing Saldo, which is 
compatible with GF, and easily translated to GF format.
As Saldo is continuously updated, the importing process has been designed to be fast
and stable enough to be redone at any time.

\subsubsection{Implementation}
The basic algorithm for importing Saldo was implemented by Angelov (2008)
and  produces code for a GF lexicon.
For each word in Saldo, it decides which forms should be used as input
to the GF smart paradigms. For a verb, this will in most cases mean giving
the present tense form, see figure \ref{fig:saldoknyt}. \\

\begin{figure}[h]
\begin{center}
\verb-mkV "knyter" ;-
\includegraphics[scale=0.5]{image1.eps} 
\caption{First code produced for the verb `knyta' (\emph{`tie'})}
\label{fig:saldoknyt}
\end{center}
\end{figure}

All assumed paradigms are printed to a temporary lexicon, 
which will produce an inflection table for every entry when compiled.
The tables are compared to the information given
in Saldo and if the tables are equal the code for the word is saved. If the table
is erroneous, another try is made
by giving more forms to the smart paradigm.
For example \ref{fig:saldoknyt}, the smart paradigm will fail to calculate the
correct inflection table. In the next try both the present and the past tense
are given:\\

\begin{figure}[h]
\begin{center}
\verb-mkV "knyter" "kn{\"o}t" ;-
\caption{Second output for the verb `knyta'}
\label{fig:saldoknyt2}
\end{center}
\end{figure}

The program is run iteratively until the GF table matches the one given in Saldo,
or until there are no more ways of using the smart paradigm. The verb 'knyta'
will need three forms:\\

\begin{figure}[h]
\begin{center}
\verb-mkV "knyter" "kn{\"o}t" "knutit"-\\
\caption{Final output for the verb `knyta'}
\label{fig:saldoknyt3}
\end{center}
\end{figure}

\subsubsection{Results}
\label{sec:saldoRes}
The resulting dictionary contains more than 100 000 entries, approximately 80 \% 
of the total size of Saldo.
There are a number of reasons why some words were not imported,
the most obvious one is that we do not want all categories from
Saldo in the GF lexicon. Prepositions, 
numerals, 
personal pronouns etc.
are assumed to be present in the resource grammars and should not be added again.
Saldo contains many pronouns which
are not analysed the same way in GF %TODO see where??
(see section \ref{sec:gfnouns}).
Before adding them to our lexicon, we need to do more analysing to find their
correct GF-category. 
%Some experiments on finding the category
%have been done using Talbanken, see section \ref{sec:gf.quant}.

Categories involving multiple words 
are usually handled as idioms and should be given in a separate lexicon. In
total six types of words were considered for the extraction: \\

\begin{figure}[h]
\begin{tabular}{|l|ll|}
\hline
& Saldo$/$ GF tag & Example \\
\hline
 Adverb & \textbf{ab} $/$\textbf{Adv} & ofta (\emph{often})\\
 Adjective&\textbf{av} $/$    \textbf{A} & gul (\emph{yellow})\\
 Noun & \textbf{nn} $/$\textbf{N} & hus (\emph{house})\\
 Verb & \textbf{vb} $/$\textbf{V} & springa (\emph{run})\\
 Reflexive verbs  & \textbf{vbm}$/$\textbf{V} & raka sig (\emph{shave})\\
 Verbs with particles &\textbf{vbm}$/$\textbf{V}  &  piggna till (\emph{perk up})\\
\hline
\end{tabular}
\caption{}
\end{figure}

Most but not all words of these categories have been imported.
One reason why the importing phase would fail 
is that Saldo, unlike GF, only contains the actually used word forms.
For technical reasons, the smart paradigm might need forms never used.
Consider for example the plural tantum 
noun \emph{`glas{\"o}gon'} (\emph{`glasses'}).
The smart paradigm requires a singular form, and since the program could not
find this in Saldo, there was no way of adding the lemma to the lexicon. 
When the program failed to import a noun, this was often the explanation.
Words of this type may be added manually, for \emph{`glas{\"o}gon'} we could use
the ostensibly correct singular form `glas{\"o}ga', although this
has another meaning (`glass-eye').
The same problem occurred for the irregular s-verbs,
(\emph{`synas'} \emph{(`show')} or \emph{umg{\aa}s} \emph{(`socialize')})
which made up 61.5 \% of the failing verbs of type \verb_vb_.\\
In a few cases the smart paradigms could not generate the correct declination.\\

When testing the coverage of Talbanken,
we found that there are around 2500 word forms still missing, excluding the ones
tagged as names and numbers. This number may seem very high, but 
4/5 of the word forms are compounds and when performing the intended parsing,
an additional analysis identifying compounds should be preformed before
looking-up the words in the lexicon. We
should also take into consideration that 
we cannot automatically find out how many actually stem from the same word, or
how many abbreviations that are present. Talbanken also contains a small number of 
spelling errors, which probably are enumerated among our missing words. The majority
of the missing words are only used once.\\

\begin{figure}[h]
\begin{tabular}{|l|l|}
\hline
Missing words &$\sim$ 2500 word-forms\\
\hline
\hline
Ignoring compounds & $\sim$ 500 word-forms\\
Used more than once & $\sim$ 500 word forms\\
Used more than once,& \\
\hspace{2mm} ignoring compounds & $\sim$ 150 word-forms\\
\hline
\end{tabular}
\caption{}
\end{figure}

A list of words that were given different labels in GF than in Talbanken has been
composed, consisting of about 1600 entries. Many of those are
acceptable and reflects the difference 
made in the analyses, such as the examples in figure \ref{tab:saldodiff}.
Others are examples of words that are still missing from
the lexicon.

%Since Saldo does not give any valency information, neither does the imported lexicon.
%The lexicon does not contain any information.
\noindent Valency information, which is crucial for GF, is not given in Saldo and
hence not in the imported lexicon. Instead we extract this from
Lexin\footnote{http://spraakbanken.gu.se/lexin/}.

\subsection{A tool for lexical acquisition}
As we extract our main lexicon from Saldo, we have also created a tool for
semi-automatic acquisition to complement the lexicon. We have
tested it on verbs, with good results. It makes use of
the \emph{smart paradigm} given in the Swedish resource grammar.
The smart paradigm is a function which given one form of a word, can
infer which paradigm it most likely belongs to.
For verbs, the paradigm accepts words in present tense indicative form.
If needed, it also accepts more verb forms showing the correct inflection.

By combining this method with the information from the tags in Talbanken,
the tool interactively generates GF lexical entries. 
Given a list of words, it iteratively
tries to figure out how to conjugate each of them. If several forms of a word are 
given, the program will try to identify the one that carries the most linguistic
information, put this in a form recognized by the smart paradigm and ask GF to output
a table with the resulting inflection. 
If the table contains all other conjugations from the input list,
the program will ask the  user to
validate the claimed paradigm. The user may now either
allow the word to be added to the lexicon, remove it or request another guess.
The user hence only needs to decide if each paradigm is correct or not, and
the program will do the rest.

The tool has been tested on verbs. Although using simple techniques it 
manages to assign the correct paradigm to 70-75 \% of the given lemmas.
A smaller test has shown that out of the accepted lemmas, the correct guess is
made directly in 75\% of the cases, whereas the user has to reject one or more
guesses for 25\%. Hence the program gets a correctness rate of 56 \% if it is
run without interference from the user. This kind of use is however not intended.

%add: ?
%From this a lexicon with about 1000 verbs have been extracted, manually verified.

\section{Translating The Treebank Talbanken}
The information from the tags in Talbanken can be used for many purposes.
We are currently working on an automatic transformation of Talbanken trees 
to trees in GF format. The translation makes use of the POS tags as well as
the syntactic information, and the mapping has so far turned out to be unambiguous.

The mapping gives information about which form a word is
currently used in and this may be used by the lexical extraction
tools.
The translated trees enable us to extract probabilities for how often
different functions are used, a feature that will enable disambiguation.
Furthermore, the translation makes it easy to identify grammatical constructions
missing from the GF grammar and shows how the GF analysis differs from the one made
in Talbanken.
Another important use of the mapping is evaluation of the parser, which can be
accomplished by comparing the parse trees and the trees from the transformer.

We have so far concentrated on shorter sentences, without idioms and
conjunction. 
Since GF demands information 
that is not given by the Talbanken tags -- such as the valency of verbs -- a
full translation requires that all words used are already known to GF. 
The evaluation of trees of this sort shows that the mapping
can restore 85\% of the nodes. 
If we lift the restrictions, we can restore 63\% of the tags.

\section{Development of the grammar}
An important part of this project has been to develop the Swedish GF grammar and
to adapt it to cover constructions used in Talbanken. As a grammar
implementation can never be expected to give
full coverage of a language, we aim for a grammar fragment which gives a deep
analysis of the most important Swedish constructions.
The starting point has been the GF resource grammar and the new implementation is
still compatible with this.
It has earlier been hard to identify missing constructions of the Swedish
implementation, since there was no large resource available to evaluate it on.

Our evaluations are based on Talbanken, and when first conducting tests,
we found much room for improvement.

For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
languages. 
However, if we aim for a deeper and more comprehensive analysis of Swedish,
the implementation of the languages needs to be more independent.
The resource grammar gives a good start and our present grammar 
% don't say inverted word order
covers constructions such as declarative sentences with normal or inverted
word order, questions, passives, imperatives, relative clauses, cleft
constructions etc.
A number of constructs that are generally not present in other languages and
therefore not given by the resources, have also been added. 
%These include the use of
%the reflexive pronoun \emph{sitt}: \\
%\emph{Han s{\aa}g \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw his (own) house}) $\;$
%as opposed to \\
%\emph{Han s{\aa}g \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
%Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
%\emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
%This sort of rephrasing is not given by the resource grammar, since it
%has little effect on the logical representation.
\subsection{The s-passive}
Passive voice is often used in Swedish, especially the
\textit{s-passive}.
\enumsentence{
\shortex{5}{Uppsatsen & skrev\textbf{s} & av & en & student.}
        {\emph{the essay} & \emph{wrote+\textbf{s}} & \emph{by} & \emph{a} & \emph{student}}
        {``The essay was written by a student."}}\label{sent:skrevs}
%{b. &En & student & skrev & uppsatsen.}
%        {&a &student &wrote &the essay}
%        {~~`A student wrote the essay.'}}
Some studies suggest that the s-passive is used in more than 80 \% of the times
\cite{laanemets}.
It is however not as common in the other Scandinavian languages, 
The resource grammar for Scandinavian therefore implemented a function for passive,
\verb-PassV2-, by using auxiliary verb. \\

During this project, the s-passive was added although the
periphrastic passive is still allowed.
The grammar further allows not only two-place verbs, but all verb phrases missing an object, to 
form passive:
\enumsentence{\textbf{Active use of two-place verb}\\\shortex{4}
{Vi & erbj{\"o}d& henne& jobbet }
{\emph{we} & \emph{offered}& \emph{her}&\emph{the job}}
{``We offered her the job"}}
\label{sent:give2pass}
\enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
{Hon & erbj{\"o}ds & jobbet}
{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
{``She was offered the job"}}
\label{ex:passV33}
\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
{Jobbet & erbj{\"o}ds & henne }
{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
{``The job was offered to her"}}
\label{ex:passV32}
\enumsentence{\textbf{V2A verb}\\\shortex{3}
{Huset & m{\aa}lades & r{\"o}tt}
{\emph{the house} & \emph{painted+\textbf{s}}& \emph{red}}
{``The house was painted red"}}
\label{ex:passV2A}

\subsection{Impersonal constructions}
\label{sec:Formal}
Formal subjects \cite[\textsection 19]{SAG} are often used in
Swedish.
\enumsentence{\shortex{6}
{Det & sitter & en & f{\aa}gel & p{\aa} & taket}
{\emph{it }& \emph{sits }& \emph{a} & \emph{bird} &\emph{ on} & \emph{the roof}}
{``There is a bird sitting on the roof''}}
\emph{`Det'} has the position of the subject, and the real subject, 
\emph{`en f{\aa}gel'} the one of an object.

Our grammar implementation allows this for intransitive verbs
and transitive verbs in passive form
\enumsentence{\shortexnt{7}
{*Det & äter & en & f{\aa}gel & fr{\"o}n & p{\aa} & taket}
{~\emph{it} & \emph{eats} & \emph{a} & \emph{bird} & \emph{seeds} & \emph{on }& \emph{the roof}}}
\label{sent:fagelfron}

\enumsentence{\shortex{5}
{Det&dricks&mycket & {\"o}l& nuf{\"o}rtiden}
{\emph{It}&\emph{drinks+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
{``A lot of beer is being drunk these days"}}

The grammar also respects the restrictions on the real subject, which is not
allowed to be in definite form or with an adjective in definite form \cite{Cooper}
\enumsentence{\shortexnt{6}
{*Det     & sitter     & \textbf{den}& \textbf{f{\aa}geln}& p{\aa}       & taket.}
{\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}}
}
\enumsentence{\shortexnt{6}
{*Det& sitter& \textbf{min}& \textbf{f{\aa}gel}& p{\aa}& taket}
{\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}}
}\label{sent:dumfagel}

\subsection{Formalizing the rules for reflexive pronouns by using dependent types}
\label{sec:reflexives}
An important area in a Swedish grammar is the treatment of the reflexive pronouns and
the reflexive possessive pronouns as described in section \ref{swe:refl} 
The reflexives require an antecedent with which they agree in
number and person.
Our grammar accept the following sentences:

\enumsentence{Han s{\aa}g sina barns skor.\\ \label{ex:ReflGen}
He saw {\sc self's} children's shoes.}
\enumsentence{Sina vantar hade han gl{\"o}mt p{\aa} t{\aa}get.\\
{\sc self's} gloves, he had forgotten on the train.}
\enumsentence{Hon ber alla sina kompisar att g{\aa}\\
She asks all {\sc self's} friends to leave}\label{ex:ReflPredet}
\enumsentence{Jag vill att han rakar sig.\\
I want him to shave {\sc self}.}
%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
%a. & Han är längre än sin kompis. & & b. &Han är här oftare än sin kompis.\\
%   & He is taller than {\sc self's} friend.
%   & & & He is here more often than {\sc self's} friend.
%\end{tabular}}\label{ex:ReflAP} \label{ex:ReflAdv}
\eenumsentence{
\item Hon tyckte om skolan och alla sina elever.\\
     She liked the school and all {\sc self's} students.
\item Han s{\aa}g sina f{\aa} b{\"o}cker och sin penna.\\
 He saw {\sc self's} few books and {\sc self's} pencil.
}\label{ex:ReflConj} \label{ex:ReflConj2}
Reflexive pronouns can not be used in subject noun phrases of finite sentences,
as shown by the ungrammatical examples in 
sentence (\ref{sent:sesig}) and (\ref{ex:ReflSubj}).
The third person reflexives (`sig',`sin') requires a third person antecedent (see \ref{ex:refljag}).
Furthermore, the antecedent must be within the same finite sentence as the reflexive
pronoun, see (\ref{sent:sesig}).
The grammar does not accept any of these sentences:
\enumsentence{
*Sina vantar var kvar p{\aa} t{\aa}get.\\
{\sc self's} gloves were left on the train.}\label{ex:ReflSubj}
\enumsentence{*Han och sin kompis läser en bok.\\
He and {\sc self's} friend are reading a book. }
\enumsentence{*Jag ger sina pengar till sina syskon.\\
I give {\sc self's} money to {\sc self's} siblings. }\label{ex:refljag}
\enumsentence{*Han vill att jag ser p{\aa} sig.\\
He wants me to look at {\sc self}.}\label{sent:sesig}

In order to implement a covering use of reflexive pronouns, we needed to diverge
from the old abstract implementation. To capture the agreement between the
antecedent and the reflexive pronoun and while allowing the grammar to accept reflexive
constructions embedded in abverbial and adjectival phrases, in fronted positions etc,
we make use of the dependent types of GF. This is the first use of the dependent
types in the resource grammars and a could be extended to other languages having
similar dependecies.

\subsection{A second future tense: ``Kommer att"}
Swedish and Norwegian use two auxiliary verbs for future tense \\
\cite[p. 246]{H&H}:
%\cite[Tempus \textsection 28]{SAG}.
\eenumsentence{
\item {
Det \textbf{kommer att} bli m{\"o}rkt snart\\
``It will get dark soon" }
\item {
Jag \textbf{ska} g{\aa} och lägga mig nu\\
``I'm going to bed now"\\}
}
The resource grammar included 
\emph{`ska'}, which was implemented as the standard way of forming future
tense, and hence represents the translation of
\emph{`will'}. 
The new grammar also supports \emph{``kommer att"}, expressed as an alternative
future tense.
\begin{figure}[h]
\begin{verbatim}
(UseCl (TTAnt TFutKommer ASimul) PPos 
       (ImpersCl (AdvVP (ComplVA become_VA 
                (PositA dark_A)) soon_Adv))
\end{verbatim}
\caption{Result of parsing ``Det kommer att bli m{\"o}rkt snart". The future tense
         is marked by the constant TFutKommer}
  \label{fig:kommeratt}
\end{figure}

Other goals: keep down overgeneration.
\subsection{Testing}
Throughout the project regression testing has been used. Every time the grammar was updated
it was tested against a treebank consisting of 155 sentences, and the number of
parse trees were compared to a standard.
The purpose was first of all to
make sure that the coverage was not decreased, but also to make sure that we
would notice if new ambiguities or overgenerating functions were introduced.
If so, the ambiguities could be removed when possible, and when not, we were at least made
aware of them and could decide whether the increase of coverage were worth the
increase in ambiguities.

New functions were also tested by random generating and by creating tables showing
all forms of the added construction.

\section{Evaluation and Future Work}
The project has so far resulted in
\begin{itemize}
\item a large-scale GF lexicon and a program to redo the importation when needed
\item an extended grammar covering an important part of Swedish
\item a comparison between GF and another annotation
\item a deeper testing of the Swedish resource grammar and an estimation
%TODO lexin, chunking?
of how well GF can be used to describe larger parts of a language
\item a study of how dependent types can be used in the resource grammars
\end{itemize}

Besides being capable of reimporting Saldo, the lexicon extraction program could also
be modified for importing other lexical resources. The only requirement is that
the resource provides inflection tables.

The grammar has been extended and enhanced, and its current status is
a specialized extension of the resource grammar.
Besides parsing, the grammar may well be used for language generation,
which works fast even when using an extensive lexicon.
Although it is not been formally verified, we believe that the majority of the
sentences generated are grammatically correct in a syntactical point of view.
Without any semantic knowledge, nonsense phrases cannot be avoided in
random generation. However, given that the abstract tree has a meaningful
interpretation,
the linearization should be correct.

Without methods for robust parsing, %TODO evaluate??
it is not interesting to talk about coverage, but about the quality and
the ability to scale up, which has so far proved to be good.
We further believe that the presence of a expert in Swedish, professor Elisabet
Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl},
has increased the standard substantially.\\

By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
added many of the commonly used words that were missing from the older
version. This is of course a big improvement.
However, the lexical part still requires some work before it can be made use of.
The lexicon is too big to use with the current techniques. Its size causes the
incremental parsing algorithm to use more heap memory than normal computers
allow. To solve this, we need to use the lexicon data more cleverly.

%\section{Evaluation and future work}
%We are building a parser which is based on grammatical rules rather than
%statistics. The parser will therefore validate that
%a given sentence is grammatically correct according to the rules defined.
%Chosen parts of the parser or grammar may be used in other applications
%which deal with controlled natural languages. Automatic language
%generation from the grammar is also provided by GF.
%
%%TODO already done?
%We aim to make the parser robust by equipping it with techniques such as chunk
%parsing, named entity recognition and methods for handling unknown 
%grammatical constructions such as idioms and ellipses. 
%It is already possible to add probabilities for GF functions,
%which rank the parse trees.
%By adding dependency probabilities, we aim to improve the
%potential for disambiguation.

Use 10 \% of Talbanken as evaluation material, not seen by grammar developer.
We intend to evaluate the parser both automatically -- by comparing the output of
the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
She also evaluates the intermediate results.

% Add: valencies extracted, compounds from språkbanken
\section{Conclusion}
We have developed the main components
for a deep Swedish parser; an extended grammar and lexicon and material for
evaluation and disambiguation.
By starting from the GF resource grammar, we got a well-defined system for
describing language.

Our final goal is to be able to parse unrestricted text, but considering that
no syntactic theory is yet capable
of wholly covering a natural language, we are satisfied with 
the grammar implementation of an important fragment of Swedish and have
focused on constructions that are frequent in Talbanken. 
Being a natural language processing-project, it will probably never be
entirely complete, but by combining the rule-based parser with statistical
techniques, 
we still believe that it is possible to achieve our goal. 
\\

All parts of the project are open-source and may thus be used in other applications.
The grammar and the lexicon may be beneficial also when working with controlled languages,
as it increases the coverage of the Swedish resource grammar.

The constructions
that we have focused on have all been possible to implement, with varying  amounts
of work. Many of them could be done by utilizing and extending the resource library
but in some cases we needed to part from the multilingual abstract and use other
grammatical theories in order to arrive at a good analysis.

By combining GF with a large lexicon, we acquire a deep parser for 
Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
We have developed tools for extending the lexicon with words from Talbanken,
and we show how to make use of the information in the manually tagged
treebank.
The usage of GF allows us to start from
a well-defined system for describing grammar, as well as tools for
parsing.

\section{Acknowledgments}
We would like to thank Center of Language Technology for funding this project.
Further, special thanks to Aarne Ranta,
Elisabet Engdahl, Krasimir Angelov, Olga Caprotti, Lars Borin and John Camilleri for their help
and support.
%\nocite{*}

\bibliographystyle{lrec2006}
\bibliography{FinBib}

\end{document}

### 2012-03-05 doc/paper/PaperFin.tex
second draft of final lrec paper

@@
 
-\address{ Department of Computer Science \& Engineering, Gothenburg University \\
-               Address1 \\
-               author1@xxx.yy, author2@zzz.edu, author3@hhh.com\\}
+\address{ Department of Computer Science \& Engineering\\
+          Gothenburg University \\
+          Box 8718, SE-402 75 Gothenburg, Sweden\\
+          ahlberg.malin@gmail.com, enache@chalmers.se \\}
 
 % 150-200 words here
-\abstract{This paper describes on-going work on a rule-based, open-source parser for
+\abstract{This paper describes work on a rule-based, open-source parser for
 Swedish. The central component is a wide-coverage grammar implemented in the GF
@@
 controlled languages \cite{cnl} and recent experiments have showed
-that it is also possible to use the framework for parsing free language \cite{patent}.
+that it is also possible to use the framework for parsing free language.
 In addition
 to GF, we use two other main resources: the Swedish treebank Talbanken and the
-electronic SALDO lexicon. By combining the grammar together with
+electronic SALDO lexicon. By combining the grammar with
 a lexicon extracted from SALDO we obtain a parser accepting all sentences
@@
 % Copied from Master's Thesis. ok to do that?
-Swedish is a North-Germanic language,
+Our goal is to implement a wide-coverage grammar and parser for Swedish
+using the GF formalism, and 
+thereby investigate how GF can be used for open-domain parsing.
+We compile a large-scale grammar to a parser
+and combine it with the extensive lexicon SALDO. The Swedish treebank 
+Talbanken provides manually tagged trees which we use for improving and evaluating
+the grammar. The parser will additionally
+be evaluated by experts. \\
+We chose our resources so that both our grammar and parser can 
+be freely available and open-source. 
+The target language is Swedish, a North-Germanic language
 closely related to Norwegian and Danish. The languages share most of their
@@
 
-Our goal is to implement a wide-coverage grammar and parser for Swedish
-using the GF formalism, and 
-thereby investigate how GF can be used for open-domain parsing.
-We compile a large-scale grammar to a parser
-and combine it with the extensive lexicon SALDO. The Swedish treebank 
-Talbanken provides manually tagged trees which we use for improving and evaluating
-the grammar. The parser will additionally
-be evaluated by experts. \\
-We chose our resources so that both our grammar and parser can 
-be freely available and open-source.
-
 This paper will briefly introduce GF, Talbanken and SALDO in Section \ref{sec:background}
-and describe our results in Section \ref{sec:progress}.
+Section \ref{sec:extractsaldo} explains how we extract a lexicon, section \ref{sec:mapping}
+how we translate Talbanken to GF annotation and section \ref{sec:grammar}
+presents the work on the grammar implementation.
 
@@
 
-Grammatical Framework \cite{ranta-2011} (GF) is a grammar formalism based on
+Grammatical Framework \cite{ranta-2011} is a grammar formalism based on
 functional programming.
@@
 they can share much of their implementations. This is usually done by using a
-\verb|Functor|, avoids code duplication and aids the code maintenance.
+\verb|Functor| by which we avoid code duplication and which aids the code maintenance.
 
@@
 language.
-This project takes a different approach by using GF for open domain language.
+This project takes a different approach by using GF for open domain language,
+as done in recently conducted work on translating patents \cite{patent}.
+
 Using the Swedish resource grammar as our starting
@@
 Although Talbanken contains both written and spoken Swedish,
-only the prose material, consisting of 6316 sentences, will be used in this project.
+only the prose material, consisting of 6316 sentences, is used in this
+project.
 This part was also used when training the data-driven parser Maltparser \cite{malt}. \\
 
-\subsection{Saldo}
+\subsection{SALDO}
 SALDO \cite{saldo} is an open source lexicon resource
 based on Svenskt Associationslexikon. It is
-developed at Spr\{aa}kbanken at Gothenburg University
+developed at Spr{\aa}kbanken at Gothenburg University
 and intended for usage in language technology
 research. 
-We are developing tools for extracting GF lexicons
-from SALDO, which have resulted in a dictionary containing
-over 98000 entries including information about verbs that
-are reflexive or needs a particle.
-The importing method should be fast and reliable
-enough to allow us to always have a fresh version of the dictionary
-in GF.
+We have developde tools for extracting GF lexicons
+from SALDO, described in \ref{sec:extractsaldo}
+%which have resulted in a dictionary containing
+%over 98000 entries including information about verbs that
+%are reflexive or needs a particle.
+%The importing method should be fast and reliable
+%enough to allow us to always have a fresh version of the dictionary
+%in GF.
 
@@
 An example is the well-known data-driven Malt parser \cite{malt},
-which has been trained on Talbanken. 
+trained on Talbanken. 
 There are also a number of grammar-based parsers, although none is freely available.
@@
 arguments. \\
+
 The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building
@@
 Translation is supported by using Minimal Recursion
-Semantics \cite{mrs} as an interlingua. \\
+Semantics \cite{mrs} as an interlingua. 
 There is a collection of grammars implemented in this framework, giving broad-coverage
@@
 Scandinavian, while Norsource (Hellan, 2003) describes Norwegian. A Swedish version
-was based upon this (SweCore, Ahrenberg) covering the morphology and some
+was based upon this by Ahrenberg, covering the morphology and some
 differences between Swedish and Norwegian. Further, there is the BiTSE 
@@
 
-\subsection{Extracting a large lexicon}
+\section{Extracting a large lexicon}
+\label{sec:extractsaldo}
 The lexicon provided with the GF resources is far too small for open-domain
 parsing.
-This section describes the process of importing Saldo, which is 
+This section describes the process of importing SALDO, which is 
 compatible with GF, and easily translated to GF format.
-As Saldo is continuously updated, the importing process has been designed to be fast
+As SALDO is continuously updated, the importing process has been designed to be fast
 and stable enough to be redone at any time.
@@
 \subsubsection{Implementation}
-The basic algorithm for importing Saldo was implemented by Angelov (2008)
+The basic algorithm for importing SALDO was implemented by Angelov (2008)
 and  produces code for a GF lexicon.
-For each word in Saldo, it decides which forms should be used as input
-to the GF smart paradigms. For a verb, this will in most cases mean giving
+For each word in SALDO, it decides which forms should be used as input
+to the GF smart paradigms. The smart paradigm is a function which given one
+form of a word, can infer which paradigm it most likely belongs to.
+For verbs, this will in most cases mean giving
 the present tense form, see figure \ref{fig:saldoknyt}. \\
@@
 The tables are compared to the information given
-in Saldo and if the tables are equal the code for the word is saved. If the table
+in SALDO and if the tables are equal the code for the word is saved. If the table
 is erroneous, another try is made
@@
 \begin{center}
-\verb-mkV "knyter" "kn{\"o}t" ;-
+\verb-mkV "knyter" "knöt" ;-
 \caption{Second output for the verb `knyta'}
@@
 
-The program is run iteratively until the GF table matches the one given in Saldo,
+The program is run iteratively until the GF table matches the one given in SALDO,
 or until there are no more ways of using the smart paradigm. The verb 'knyta'
@@
 \begin{center}
-\verb-mkV "knyter" "kn{\"o}t" "knutit"-\\
+\verb-mkV "knyter" "knöt" "knutit"-\\
 \caption{Final output for the verb `knyta'}
@@
 The resulting dictionary contains more than 100 000 entries, approximately 80 \% 
-of the total size of Saldo.
+of the total size of SALDO.
 There are a number of reasons why some words were not imported,
 the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions, 
+SALDO in the GF lexicon. Prepositions, 
 numerals, 
@@
 are assumed to be present in the resource grammars and should not be added again.
-Saldo contains many pronouns which
-are not analysed the same way in GF %TODO see where??
-(see section \ref{sec:gfnouns}).
+SALDO contains many pronouns which
+are not analysed the same way in GF. %TODO see where??
 Before adding them to our lexicon, we need to do more analysing to find their
@@
 %have been done using Talbanken, see section \ref{sec:gf.quant}.
-
 Categories involving multiple words 
@@
 \hline
-& Saldo$/$ GF tag & Example \\
+& SALDO$/$GF & Example \\
 \hline
- Adverb & \textbf{ab} $/$\textbf{Adv} & ofta (\emph{often})\\
- Adjective&\textbf{av} $/$    \textbf{A} & gul (\emph{yellow})\\
- Noun & \textbf{nn} $/$\textbf{N} & hus (\emph{house})\\
- Verb & \textbf{vb} $/$\textbf{V} & springa (\emph{run})\\
- Reflexive verbs  & \textbf{vbm}$/$\textbf{V} & raka sig (\emph{shave})\\
- Verbs with particles &\textbf{vbm}$/$\textbf{V}  &  piggna till (\emph{perk up})\\
+ Adverb & \textbf{ab}$/$\textbf{Adv} & ofta (\emph{often})\\
+ Adjective&\textbf{av}$/$\textbf{A} & gul (\emph{yellow})\\
+ Noun & \textbf{nn}$/$\textbf{N} & hus (\emph{house})\\
+ Verb & \textbf{vb}$/$\textbf{V} & springa (\emph{run})\\
+ Reflexive verbs &\textbf{vbm}$/$\textbf{V} & raka sig (\emph{shave})\\
+ Particles verbs &\textbf{vbm}$/$\textbf{V}  &  piggna till (\emph{perk up})\\
 \hline
@@
 One reason why the importing phase would fail 
-is that Saldo, unlike GF, only contains the actually used word forms.
+is that SALDO, unlike GF, only contains the actually used word forms.
 For technical reasons, the smart paradigm might need forms never used.
@@
 The smart paradigm requires a singular form, and since the program could not
-find this in Saldo, there was no way of adding the lemma to the lexicon. 
+find this in SALDO, there was no way of adding the lemma to the lexicon. 
 When the program failed to import a noun, this was often the explanation.
@@
 4/5 of the word forms are compounds and when performing the intended parsing,
-an additional analysis identifying compounds should be preformed before
+an additional analysis identifying compounds is preformed before
 looking-up the words in the lexicon. We
@@
 acceptable and reflects the difference 
-made in the analyses, such as the examples in figure \ref{tab:saldodiff}.
-Others are examples of words that are still missing from
+made in the analyses, while
+others are examples of words that are still missing from
 the lexicon.
 
-%Since Saldo does not give any valency information, neither does the imported lexicon.
+%Since SALDO does not give any valency information, neither does the imported lexicon.
 %The lexicon does not contain any information.
-\noindent Valency information, which is crucial for GF, is not given in Saldo and
+\noindent Valency information, which is crucial for GF, is not given in SALDO and
 hence not in the imported lexicon. Instead we extract this from
@@
 \subsection{A tool for lexical acquisition}
-As we extract our main lexicon from Saldo, we have also created a tool for
-semi-automatic acquisition to complement the lexicon. We have
-tested it on verbs, with good results. It makes use of
-the \emph{smart paradigm} given in the Swedish resource grammar.
-The smart paradigm is a function which given one form of a word, can
-infer which paradigm it most likely belongs to.
-For verbs, the paradigm accepts words in present tense indicative form.
-If needed, it also accepts more verb forms showing the correct inflection.
-
-By combining this method with the information from the tags in Talbanken,
-the tool interactively generates GF lexical entries. 
-Given a list of words, it iteratively
-tries to figure out how to conjugate each of them. If several forms of a word are 
+As we extract our main lexicon from SALDO, we have also created a tool for
+semi-automatic acquisition to complement the lexicon.
+%We have tested it on verbs, with good results.
+Like the SALDO importer it makes use of
+the smart paradigm given in the Swedish resource grammar.
+Unlike the SALDO importer, this tool does not require any particular
+word form, but operates on any verb form and 
+%By combining this method with the information from the tags in Talbanken,
+%the tool interactively generates GF lexical entries. 
+%Given a list of words, it 
+iteratively tries to figure out how to conjugate each of them. If several forms
+of a word are 
 given, the program will try to identify the one that carries the most linguistic
@@
 the program will ask the  user to
-validate the claimed paradigm. The user may now either
+validate the claimed paradigm. 
+This step is needed since the input may not provide information enough
+to automatically do the validation. The user may now either
 allow the word to be added to the lexicon, remove it or request another guess.
@@
 made directly in 75\% of the cases, whereas the user has to reject one or more
-guesses for 25\%. Hence the program gets a correctness rate of 56 \% if it is
-run without interference from the user. This kind of use is however not intended.
+guesses for 25\%. 
+%Hence the program gets a correctness rate of 56 \% if it is
+%run without help from the user. This kind of use is however not intended,
 
@@
 \section{Translating The Treebank Talbanken}
+\label{sec:mapping}
 The information from the tags in Talbanken can be used for many purposes.
-We are currently working on an automatic transformation of Talbanken trees 
+We have developed an automatic transformation of Talbanken trees 
 to trees in GF format. The translation makes use of the POS tags as well as
-the syntactic information, and the mapping has so far turned out to be unambiguous.
+the syntactic information. 
+\begin{figure}[h!]
+\begin{center}
+\hspace{-30mm}
+\subfloat{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
+\hspace{-10mm}
+\subfloat{\label{pic:gftree}\includegraphics[width=90mm]{gfTree.pdf}}
+%\subfloat{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
+\caption{The Talbanken tree and the GF tree for the sentence ``Katten p{\aa} bilen blir st{\"o}rre".}
+\label{fig:translationtrees}
+\end{center}
+\end{figure}
 
-The mapping gives information about which form a word is
-currently used in and this may be used by the lexical extraction
-tools.
-The translated trees enable us to extract probabilities for how often
-different functions are used, a feature that will enable disambiguation.
-Furthermore, the translation makes it easy to identify grammatical constructions
-missing from the GF grammar and shows how the GF analysis differs from the one made
+Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
+of the sentence ``Katten p{\aa} bilen blir st{\"o}rre" 
+(\emph{``The cat on the car gets bigger"}).
+The mapping is based on 
+a translation of the English Penn Treebank \cite{gfpenn}.
+By modifying this program, we now have a translation that works for
+the Tiger XML-format of
+Talbanken. Adaption was required for the differences in annotation as well as 
+for the syntactic differences between Swedish and English.
+
+The translation %connects GF with another annotations and 
+gives us means to evaluate our parser. By both parsing a Talbanken sentence and
+transforming its annotated tree, we can easily inspect if the results are
+equal.
+Additionally, the mapping shows which grammatical constructions that are still missing
+from the GF grammar and shows how the GF analysis differs from the one made
 in Talbanken.
-Another important use of the mapping is evaluation of the parser, which can be
-accomplished by comparing the parse trees and the trees from the transformer.
+If there are words missing from our dictionary, the rich
+POS-tags may help us to automatically find the correct declination and add it to the
+lexicon. Further, our parser will need probabilities of how often a function is
+used. The GF treebank we achieve from the translation is a good source for this
+information.\\
 
-We have so far concentrated on shorter sentences, without idioms and
-conjunction. 
-Since GF demands information 
-that is not given by the Talbanken tags -- such as the valency of verbs -- a
-full translation requires that all words used are already known to GF. 
-The evaluation of trees of this sort shows that the mapping
-can restore 85\% of the nodes. 
-If we lift the restrictions, we can restore 63\% of the tags.
+%The mapping gives information about which form a word is
+%currently used in and this may be used by the lexical extraction
+%tools.
+%The translated trees enable us to extract probabilities for how often
+%different functions are used, a feature that will enable disambiguation.
+%Furthermore, the translation makes it easy to identify grammatical constructions
+%missing from the GF grammar and shows how the GF analysis differs from the one made
+%in Talbanken.
+%Another important use of the mapping is evaluation of the parser, which can be
+%accomplished by comparing the parse trees and the trees from the transformer.
+%\subsection{Differences in the notation}
+%Some structural differences between GF and Talbanken became obvious during the
+%translation.
+%In Talbanken the valency information is given implicitly by the complements
+%of a word. If a verb is followed by an object, \verb-OO-, which contains an
+%\verb-S-, we can conclude that this is a sentence-complement verb.
+%In GF, the valency information is given for each entry in the lexicon.
+%A sentence-complement verb has the type \verb-VS- and from this we know 
+%that the function \verb-ComplVS- should be used to combine the verb and with a 
+%sentence.
+%\begin{wrapfigure}{l}{0.52\textwidth}
+%\includegraphics[width=80mm]{metatree.pdf}
+%\caption{Abstract tree with meta variable}
+%\label{pic:gMetatree}
+%\end{wrapfigure}
+%
+%\noindent A similar issue arises from the use of verb particles. A verb with a
+%particle is treated as two independent units in Talbanken but in GF they are
+%considered to be one constituent. % which means that 
+%The particle has therefore no effect on the abstract tree, its presence is
+%announced by the verb \\itself.
+%
+%
+%We have so far concentrated on shorter sentences, without idioms and
+%conjunction. 
+%Since GF demands information 
+%that is not given by the Talbanken tags -- such as the valency of verbs -- a
+%full translation requires that all words used are already known to GF. 
+%The evaluation of trees of this sort shows that the mapping
+%can restore 85\% of the nodes. 
+%If we lift the restrictions, we can restore 63\% of the tags.
+\subsection{Results and evaluation}
+%>> more! replace words, how much better results (both grammar and mapping) 
+%>> more results
+%>> motivation, why so bad??\\
+The development was mostly example-driven, and at least one rule for the translation of
+every Talbanken-tag has been implemented.
+Shorter sentences have been prioritized in order to get a coverage of the most
+fundamental constructions. 
+Regression testing was used to ensure that a new rule would not violate the old ones
+or allow trees to be translated erroneously.
+
+The flat version of Talbanken05 was used for developing the mapping, but
+the program works for both the flat and the deep Tiger-XML versions.
+%The deeper contains six more tags, although they are not always useful for our       
+%purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
+%% show piece of s5 here!
+%\begin{figure}[h]
+%\begin{parsetree}
+%    ( .CNP. 
+%        (.CJ.  (.NP. ~ `oskifta dödsbon'))
+%        ( .++. (.++OC. `och'))
+%        (.CJ.  (.VN. `familjestiftelser'))
+%    )
+%\end{parsetree}
+%\begin{parsetree}
+%    ( .CNP. 
+%        (.CJ.  (.NP. ~ `oskifta dödsbon'))
+%        ( .CC.
+%            (.CNP. 
+%                (.++.  (.++OC. `och' ))
+%                (.C+.  (.VN. `familjestiftelser'))
+%         ))
+%    )
+%\end{parsetree}
+%\caption{An example of the difference in the two analyses}
+%\label{fig:mappDeepFlat}
+%\end{figure}
+The focus on the mapping has been simple sentences, in order to cover the most
+fundamental structures. 
+%If the program is extended and adapted to cover complex
+%conjunction, a deeper evaluation should be preformed to see which version that
+%suits our needs the best.
+%The deeper implementation also gives 
+%some more information about verb phrases, by the tag \verb|VG| (\emph{verb group}),
+%which groups an infinitival verb and its object together into a \verb|VP|.
+%This information can however be extracted from the flat implementation, and the results
+%get slightly better when using this version. \\
+
+When evaluating the mapping, the results strongly depend on which restrictions we
+put on the input. 
+One of the reasons why a node cannot be translated, is the 
+use of the tags show in figure \ref{fig:mapBadtag}.
+The \textbf{PU} tag is used for graphic listings, and not for fluent text.
+In our grammar there is naturally no corresponding function; 
+%We are not interested in sentences made up by list, since 
+the listings are meant for making the text look nice in
+folders etc and outside the scope for the grammar itself. The tags \textbf{XX}
+and \textbf{NAC} are often used since Talbanken makes a
+difference between subject
+and object noun phrases. %(which is not needed in GF (see section \ref{sec:swegf}))
+The analysis of elliptical expression in (\ref{sent:krav})
+\enumsentence{F\"or stora krav.\\
+``Too high demands."}\label{sent:krav}
+contains the tags \textbf{XX} and \textbf{NAC}, since it is not obvious
+whether the noun phrase is used as subject or an object.
+The tags shown in figure \ref{fig:mapBadtag} occur quite frequently in the treebank and are always translated
+to metas, which lowers our result. \\
+%How we cannot cover all cases. 
+\begin{figure}[h]
+\begin{tabular}{ll}
+\textbf{NAC} & Not a constituent\\
+\textbf{XP} & Other (non-coordinated) phrase\\
+\textbf{DB} & Doubled function\\
+\textbf{PU} & List item\\
+\textbf{XX} & Unclassifiable part-of-speech\\
+\end{tabular}
+\caption{}\label{fig:mapBadtag}
+\end{figure}
+
+The main goal has been to be able to translate shorter sentences, with no
+idioms or conjunction.
+If we assure that the lexicon contains the correct word class for all lemmas
+involved, we
+can restore more than 85 \% of the nodes in the original tree.
+If we lift all the restrictions excluding the \verb|PU|, we get
+%When tested the 100 first sentences from Talbanken, we get
+65 \% coverage. 
+If we test randomly collected sentences that do not contain any of the tags listed
+in figure \ref{fig:mapBadtag}, 72 \% can be restored (see figure \ref{tab:mappres})
+
+%TODO
+%All NPs: 72%
+\begin{figure}[h]
+\begin{tabular}{|ll|}
+\hline
+No list items & 65 \%\\
+No special punctuation or bad tags& 72 \%\\
+Short sentences with known words & 85 \%\\
+\hline
+\end{tabular}
+\caption{}
+\label{tab:mappres}
+\end{figure}
+
+A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
+The percentage of restored nodes from Peen Treebank is higher than our results.
+%TODO review
+The reason for this may be the fact that English is syntactically less complicated
+than Swedish.
+Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
+where idiomatic expressions are more likely the be appear and the language
+presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
+Also, the Penn Treebank contains a lower number of tags, 82 compared
+to more than 300 in Talbanken. Even if the tags describing the same word
+%TODO review
+class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
+With more tags, we get more information, but as the number 
+%Thus the number of
+increase, so does the amount of work of finding the correct translation 
+for each combination of tags and writing rules that cover all constructions. 
+%iand so is the number of
+%rules that would be needed to cover all constructions.
+
+We believe that the results could be enhanced by simply adding more rules 
+and in this way get a wider coverage. There are many special cases that require
+special rules. Since we are not aware of any formal specification of how the tags may be
+combined in Talbanken, the only way of finding all possibilities are to manually look
+for different patterns. Another option would be to make the mapping more
+robust, but the robustness must not interfere with the correctness.
 
 \section{Development of the grammar}
+\label{sec:grammar}
 An important part of this project has been to develop the Swedish GF grammar and
@@
 still compatible with this.
+
 It has earlier been hard to identify missing constructions of the Swedish
 implementation, since there was no large resource available to evaluate it on.
-
 Our evaluations are based on Talbanken, and when first conducting tests,
 we found much room for improvement.
-
-For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
-languages. 
-However, if we aim for a deeper and more comprehensive analysis of Swedish,
+%For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
+%languages. 
+%However, if we aim for a deeper and more comprehensive analysis of Swedish,
 the implementation of the languages needs to be more independent.
@@
 % don't say inverted word order
-covers constructions such as declarative sentences with normal or inverted
-word order, questions, passives, imperatives, relative clauses, cleft
-constructions etc.
-A number of constructs that are generally not present in other languages and
-therefore not given by the resources, have also been added. 
-%These include the use of
-%the reflexive pronoun \emph{sitt}: \\
-%\emph{Han s{\aa}g \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw his (own) house}) $\;$
-%as opposed to \\
-%\emph{Han s{\aa}g \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
-%Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
-%\emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
-%This sort of rephrasing is not given by the resource grammar, since it
-%has little effect on the logical representation.
+covers constructions such as declarative sentences, questions, passives,
+imperatives, relative clauses, cleft constructions etc.
+During our work the grammar has been extended to cover language specific
+constructions such as topicalisation and the features listed below. We further
+put effort on avoiding and removing overgeneration, which might pass unnoticed
+when parsing controlled text but which cause an explosion of ambiguous parse trees
+when used for free text and with a large lexicon.
 \subsection{The s-passive}
-Passive voice is often used in Swedish, especially the
-\textit{s-passive}.
+Passive voice is commonly used in Swedish.
+There are two ways of forming passive verb phrases in Swedish: the 
+\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli'
+(\emph{`become'}) and the \textbf{s-passive} which is formed by adding an
+\emph{s} to the verb: \\
 \enumsentence{
@@
 It is however not as common in the other Scandinavian languages,
-The resource grammar for Scandinavian therefore implemented a function for passive,
-\verb-PassV2-, by using auxiliary verb. \\
+and the resource grammar for Scandinavian therefore only implemented a function for
+passive by using auxiliary verb.
 
-During this project, the s-passive was added although the
+During this project, the s-passive was added and the
 periphrastic passive is still allowed.
@@
 
-Our grammar implementation allows this for intransitive verbs
-and transitive verbs in passive form
-\enumsentence{\shortexnt{7}
-{*Det & äter & en & f{\aa}gel & fr{\"o}n & p{\aa} & taket}
-{~\emph{it} & \emph{eats} & \emph{a} & \emph{bird} & \emph{seeds} & \emph{on }& \emph{the roof}}}
-\label{sent:fagelfron}
+Our grammar implementation allows impersonal constructions 
+and does not allow ungrammatical variants such as sentence
+(\ref{sent:fagelfron}) where a transitive verb is used.
 
@@
 
-The grammar also respects the restrictions on the real subject, which is not
-allowed to be in definite form or with an adjective in definite form \cite{Cooper}
+The grammar also respects the restrictions on the real subject, which cannot
+contain a determiner requiring the noun or modifying adjective to be definite form \cite{Cooper}
 \enumsentence{\shortexnt{6}
@@
 An important area in a Swedish grammar is the treatment of the reflexive pronouns and
-the reflexive possessive pronouns as described in section \ref{swe:refl} 
+the reflexive possessive pronouns.
 The reflexives require an antecedent with which they agree in
@@
 
-In order to implement a covering use of reflexive pronouns, we needed to diverge
-from the old abstract implementation. To capture the agreement between the
-antecedent and the reflexive pronoun and while allowing the grammar to accept reflexive
-constructions embedded in abverbial and adjectival phrases, in fronted positions etc,
-we make use of the dependent types of GF. This is the first use of the dependent
-types in the resource grammars and a could be extended to other languages having
-similar dependecies.
+In order to achieve a scalable implementation of reflexive pronouns, we needed
+to diverge from the old abstract implementation. To capture the agreement
+between the antecedent and the reflexive pronoun, while still allowing the grammar
+to accept reflexive constructions embedded in abverbial and adjectival phrases,
+in fronted positions etc. we make use of the dependent types of GF. This is the
+first use of the dependent types in the resource grammars and a could be
+extended to other languages having similar dependecies.
 
 \subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian use two auxiliary verbs for future tense \\
+Swedish use two auxiliary verbs for future tense \\
 \cite[p. 246]{H&H}:
@@
 \item {
-Jag \textbf{ska} g{\aa} och lägga mig nu\\
+Jag \textbf{ska} g{\aa} och l{\"a}gga mig nu\\
 ``I'm going to bed now"\\}
@@
 
-Other goals: keep down overgeneration.
 \subsection{Testing}
@@
 would notice if new ambiguities or overgenerating functions were introduced.
-If so, the ambiguities could be removed when possible, and when not, we were at least made
+If so, the ambiguities could be removed when possible, and when not, we were made
 aware of them and could decide whether the increase of coverage were worth the
@@
 
-Besides being capable of reimporting Saldo, the lexicon extraction program could also
+Besides being capable of reimporting SALDO, the lexicon extraction program could also
 be modified for importing other lexical resources. The only requirement is that
@@
 
+When it comes to parsing, we do not get far without robustness.
+The grammar in itself is by no means robust, and just one 
+unexpected punctuation mark, unknown word
+or ellipsis will cause the parsing of the whole sentence to fail.
+
+Parsing bigger parts of Talbanken would hence give very low results at this stage, 
+%, for the reasons stated.
+and a comparison of the results would not be of much value as
+there would not be enough material %underlag is yet too small 
+to do be able to do any interesting analysis.
+
 Without methods for robust parsing, %TODO evaluate??
@@
 
-By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
+By the renewed import of SALDO, we have doubled the size of the lexicon and thereby
 added many of the commonly used words that were missing from the older
@@
 allow. To solve this, we need to use the lexicon data more cleverly.
+
+We are currently working on chunk parsing and disambiguation using probabilities
+extracted from our translation of Talbanken. We combine this with simple named
+entity recognition and compounding analysis. For evaluation, we intend to use
+10 \% of Talbanken, chosen so that is does not infer with our test data.
 
@@
 
-Use 10 \% of Talbanken as evaluation material, not seen by grammar developer.
-We intend to evaluate the parser both automatically -- by comparing the output of
+%Use 10 \% of Talbanken as evaluation material, not seen by grammar developer.
+The result will be evaluated both automatically -- by comparing the output of
 the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
@@
 
-Our final goal is to be able to parse unrestricted text, but considering that
-no syntactic theory is yet capable
-of wholly covering a natural language, we are satisfied with 
-the grammar implementation of an important fragment of Swedish and have
-focused on constructions that are frequent in Talbanken. 
-Being a natural language processing-project, it will probably never be
-entirely complete, but by combining the rule-based parser with statistical
-techniques, 
-we still believe that it is possible to achieve our goal. 
+We have implemented a grammar covering an important fragment of Swedish by
+focusing on constructions that are frequent in Talbanken.
 \\
@@
 
-By combining GF with a large lexicon, we acquire a deep parser for 
-Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
-We have developed tools for extending the lexicon with words from Talbanken,
-and we show how to make use of the information in the manually tagged
-treebank.
+%By combining GF with a large lexicon, we acquire a deep parser for 
+%Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
+%We have developed tools for extending the lexicon with words from Talbanken,
+%and we show how to make use of the information in the manually tagged
+%treebank.
 The usage of GF allows us to start from

### 2012-03-05 doc/paper/final/PaperFin.tex
third final lrec

--- 2012-03-05 doc/paper/PaperFin.tex
@@
 \usepackage{graphicx}
-%\usepackage[utf8]{inputenc}
 
-\title{Combining Language Resources Into A Grammar-Driven Swedish parser}
+\title{Combining Language Resources Into A Grammar-Driven Swedish Parser}
 
@@
 
+% TODO add links to the website?
 \section{Introduction}
@@
 research. 
-We have developde tools for extracting GF lexicons
+We have developed tools for extracting GF lexicons
 from SALDO, described in \ref{sec:extractsaldo}
-%which have resulted in a dictionary containing
-%over 98000 entries including information about verbs that
-%are reflexive or needs a particle.
-%The importing method should be fast and reliable
-%enough to allow us to always have a fresh version of the dictionary
-%in GF.
 
@@
 % Add: how are these related to this work?
-% Add :detaild comparison of outputs of various parsers of Sw.
+% Add: detaild comparison of outputs of various parsers of Sw.
 Many years of research have lead to many interesting language
@@
 different languages and there are now grammars for English, German, Japanese and Norwegian.
-
-%old related work:
-%Language technology for Swedish is an area of much interesting research.
-%There are two other deep grammar-based parser for Swedish:
-% put in references!!
-%Swedish Core Language Engine (Raynen and Gambäck, 1992) and a
-% does this really exist??
-%unification-based chart parser \cite{wiren},
-%though none of them are freely available.
-%Among other parsers for Swedish, the statistical MaltParser \cite{malt},
-%trained on Talbanken, is worth mentioning. 
-% put in references!!
-%CassSwe (Kokkinakis and Johansson-Kokkinakis, 1999) is based on finite state cascades,
-% put in references!!
-%whereas the shallow parser GTA (Knutsson, Bigert and Kann, 2003) relies on rules of 
-%a context free grammar. Both GTA and CassSwe operate on POS tagged text.\\
-%Extract and FM \cite{MarkusForsberg2007} are tools for supervised lexicon
-%extraction compatible with GF.
-
-%\section{Results and progress}
-%\label{sec:progress}
 
@@
 SALDO contains many pronouns which
-are not analysed the same way in GF. %TODO see where??
-Before adding them to our lexicon, we need to do more analysing to find their
+are not analyzed the same way in GF. 
+Before adding them to our lexicon, we need to do more analyzing to find their
 correct GF-category. 
-%Some experiments on finding the category
-%have been done using Talbanken, see section \ref{sec:gf.quant}.
 Categories involving multiple words 
@@
 
-%Since SALDO does not give any valency information, neither does the imported lexicon.
-%The lexicon does not contain any information.
-\noindent Valency information, which is crucial for GF, is not given in SALDO and
+Valency information, which is crucial for GF, is not given in SALDO and
 hence not in the imported lexicon. Instead we extract this from
@@
 semi-automatic acquisition to complement the lexicon.
-%We have tested it on verbs, with good results.
 Like the SALDO importer it makes use of
@@
 word form, but operates on any verb form and 
-%By combining this method with the information from the tags in Talbanken,
-%the tool interactively generates GF lexical entries. 
-%Given a list of words, it 
 iteratively tries to figure out how to conjugate each of them. If several forms
@@
 \subfloat{\label{pic:gftree}\includegraphics[width=90mm]{gfTree.pdf}}
-%\subfloat{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
 \caption{The Talbanken tree and the GF tree for the sentence ``Katten p{\aa} bilen blir st{\"o}rre".}
@@
 
-The translation %connects GF with another annotations and 
+The translation 
 gives us means to evaluate our parser. By both parsing a Talbanken sentence and
@@
 
-%The mapping gives information about which form a word is
-%currently used in and this may be used by the lexical extraction
-%tools.
-%The translated trees enable us to extract probabilities for how often
-%different functions are used, a feature that will enable disambiguation.
-%Furthermore, the translation makes it easy to identify grammatical constructions
-%missing from the GF grammar and shows how the GF analysis differs from the one made
-%in Talbanken.
-%Another important use of the mapping is evaluation of the parser, which can be
-%accomplished by comparing the parse trees and the trees from the transformer.
-%\subsection{Differences in the notation}
-%Some structural differences between GF and Talbanken became obvious during the
-%translation.
-%In Talbanken the valency information is given implicitly by the complements
-%of a word. If a verb is followed by an object, \verb-OO-, which contains an
-%\verb-S-, we can conclude that this is a sentence-complement verb.
-%In GF, the valency information is given for each entry in the lexicon.
-%A sentence-complement verb has the type \verb-VS- and from this we know 
-%that the function \verb-ComplVS- should be used to combine the verb and with a 
-%sentence.
-%\begin{wrapfigure}{l}{0.52\textwidth}
-%\includegraphics[width=80mm]{metatree.pdf}
-%\caption{Abstract tree with meta variable}
-%\label{pic:gMetatree}
-%\end{wrapfigure}
-%
-%\noindent A similar issue arises from the use of verb particles. A verb with a
-%particle is treated as two independent units in Talbanken but in GF they are
-%considered to be one constituent. % which means that 
-%The particle has therefore no effect on the abstract tree, its presence is
-%announced by the verb \\itself.
-%
-%
-%We have so far concentrated on shorter sentences, without idioms and
-%conjunction. 
-%Since GF demands information 
-%that is not given by the Talbanken tags -- such as the valency of verbs -- a
-%full translation requires that all words used are already known to GF. 
-%The evaluation of trees of this sort shows that the mapping
-%can restore 85\% of the nodes. 
-%If we lift the restrictions, we can restore 63\% of the tags.
 \subsection{Results and evaluation}
-%>> more! replace words, how much better results (both grammar and mapping) 
-%>> more results
-%>> motivation, why so bad??\\
 The development was mostly example-driven, and at least one rule for the translation of
@@
 the program works for both the flat and the deep Tiger-XML versions.
-%The deeper contains six more tags, although they are not always useful for our       
-%purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
-%% show piece of s5 here!
-%\begin{figure}[h]
-%\begin{parsetree}
-%    ( .CNP. 
-%        (.CJ.  (.NP. ~ `oskifta dödsbon'))
-%        ( .++. (.++OC. `och'))
-%        (.CJ.  (.VN. `familjestiftelser'))
-%    )
-%\end{parsetree}
-%\begin{parsetree}
-%    ( .CNP. 
-%        (.CJ.  (.NP. ~ `oskifta dödsbon'))
-%        ( .CC.
-%            (.CNP. 
-%                (.++.  (.++OC. `och' ))
-%                (.C+.  (.VN. `familjestiftelser'))
-%         ))
-%    )
-%\end{parsetree}
-%\caption{An example of the difference in the two analyses}
-%\label{fig:mappDeepFlat}
-%\end{figure}
 The focus on the mapping has been simple sentences, in order to cover the most
 fundamental structures.
-%If the program is extended and adapted to cover complex
-%conjunction, a deeper evaluation should be preformed to see which version that
-%suits our needs the best.
-%The deeper implementation also gives 
-%some more information about verb phrases, by the tag \verb|VG| (\emph{verb group}),
-%which groups an infinitival verb and its object together into a \verb|VP|.
-%This information can however be extracted from the flat implementation, and the results
-%get slightly better when using this version. \\
 
@@
 In our grammar there is naturally no corresponding function; 
-%We are not interested in sentences made up by list, since 
 the listings are meant for making the text look nice in
@@
 difference between subject
-and object noun phrases. %(which is not needed in GF (see section \ref{sec:swegf}))
+and object noun phrases.
 The analysis of elliptical expression in (\ref{sent:krav})
@@
 to metas, which lowers our result. \\
-%How we cannot cover all cases. 
 \begin{figure}[h]
@@
 \end{tabular}
-\caption{}\label{fig:mapBadtag}
+\caption{Untranslatable tags}\label{fig:mapBadtag}
 \end{figure}
@@
 If we lift all the restrictions excluding the \verb|PU|, we get
-%When tested the 100 first sentences from Talbanken, we get
 65 \% coverage. 
@@
 
-%TODO
-%All NPs: 72%
 \begin{figure}[h]
@@
 \end{tabular}
-\caption{}
+%TODO is this correct???
+\caption{Number of nodes in each translated tree not put to meta}
 \label{tab:mappres}
@@
 
-A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
+A mapping between GF and the Wall Street Journal of Penn Treebank has earlier
+been conducted \cite{gfMech}.
 The percentage of restored nodes from Peen Treebank is higher than our results.
-%TODO review
 The reason for this may be the fact that English is syntactically less complicated
@@
 to more than 300 in Talbanken. Even if the tags describing the same word
-%TODO review
 class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
 With more tags, we get more information, but as the number 
-%Thus the number of
 increase, so does the amount of work of finding the correct translation 
 for each combination of tags and writing rules that cover all constructions.
-%iand so is the number of
-%rules that would be needed to cover all constructions.
 
@@
 we found much room for improvement.
-%For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
-%languages. 
-%However, if we aim for a deeper and more comprehensive analysis of Swedish,
 the implementation of the languages needs to be more independent.
 The resource grammar gives a good start and our present grammar 
-% don't say inverted word order
 covers constructions such as declarative sentences, questions, passives,
@@
         {``The essay was written by a student."}}\label{sent:skrevs}
-%{b. &En & student & skrev & uppsatsen.}
-%        {&a &student &wrote &the essay}
-%        {~~`A student wrote the essay.'}}
 Some studies suggest that the s-passive is used in more than 80 \% of the times
@@
 I want him to shave {\sc self}.}
-%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
-%a. & Han är längre än sin kompis. & & b. &Han är här oftare än sin kompis.\\
-%   & He is taller than {\sc self's} friend.
-%   & & & He is here more often than {\sc self's} friend.
-%\end{tabular}}\label{ex:ReflAP} \label{ex:ReflAdv}
 \eenumsentence{
@@
 between the antecedent and the reflexive pronoun, while still allowing the grammar
-to accept reflexive constructions embedded in abverbial and adjectival phrases,
+to accept reflexive constructions embedded in adverbial and adjectival phrases,
 in fronted positions etc. we make use of the dependent types of GF. This is the
 first use of the dependent types in the resource grammars and a could be
-extended to other languages having similar dependecies.
+extended to other languages having similar dependencies.
 
@@
 \cite[p. 246]{H&H}:
-%\cite[Tempus \textsection 28]{SAG}.
 \eenumsentence{
@@
 Parsing bigger parts of Talbanken would hence give very low results at this stage, 
-%, for the reasons stated.
 and a comparison of the results would not be of much value as
-there would not be enough material %underlag is yet too small 
+there would not be enough material
 to do be able to do any interesting analysis.
@@
 
-%\section{Evaluation and future work}
-%We are building a parser which is based on grammatical rules rather than
-%statistics. The parser will therefore validate that
-%a given sentence is grammatically correct according to the rules defined.
-%Chosen parts of the parser or grammar may be used in other applications
-%which deal with controlled natural languages. Automatic language
-%generation from the grammar is also provided by GF.
-%
-%%TODO already done?
-%We aim to make the parser robust by equipping it with techniques such as chunk
-%parsing, named entity recognition and methods for handling unknown 
-%grammatical constructions such as idioms and ellipses. 
-%It is already possible to add probabilities for GF functions,
-%which rank the parse trees.
-%By adding dependency probabilities, we aim to improve the
-%potential for disambiguation.
-
-%Use 10 \% of Talbanken as evaluation material, not seen by grammar developer.
 The result will be evaluated both automatically -- by comparing the output of
@@
 
-%By combining GF with a large lexicon, we acquire a deep parser for 
-%Swedish. All parts, including the underlying grammar, are open-source and can be re-used.
-%We have developed tools for extending the lexicon with words from Talbanken,
-%and we show how to make use of the information in the manually tagged
-%treebank.
 The usage of GF allows us to start from

### 2012-03-08 doc/paper/final/PaperFin.tex
smaller verison of paper

@@
 \usepackage{graphicx}
+\usepackage[utf8]{inputenc}
 
@@
 has so far been used successfully for
-controlled languages \cite{cnl} and recent experiments have showed
-that it is also possible to use the framework for parsing free language.
+controlled languages \cite{cnl}. Recent experiments have showed
+that it is also possible to use the framework for parsing unrestricted language.
 In addition
@@
 grammar may be used in various NLP applications. 
-\\ \newline \Keywords{GF, Swedish, Parsing}}
+\\ \newline \Keywords{GF, Swedish, Parsing, Grammar Formalisms, Lexical Acquisition}}
 
@@
 \section{Introduction}
-% Copied from Master's Thesis. ok to do that?
 Our goal is to implement a wide-coverage grammar and parser for Swedish
@@
 the grammar. The parser will additionally
-be evaluated by experts. \\
+be evaluated by an expert. \\
 We chose our resources so that both our grammar and parser can 
@@
 This project takes a different approach by using GF for open domain language,
-as done in recently conducted work on translating patents \cite{patent}.
+similar to the recently conducted work on translation of patents \cite{patent}.
 
@@
 Talbanken \cite{talbanken}.
-It was assembled in the 1970s at Lund University and modernized
-in 2005 \cite{talbanken05} and
-enriched with annotation for a full phrase structure analysis.  \\
+It was assembled in the 1970s at Lund University and later
+%in 2005 \cite{talbanken05} and
+enriched with annotation for a full phrase structure analysis \cite{talbanken05}.  \\
 Although Talbanken contains both written and spoken Swedish,
@@
 We have developed tools for extracting GF lexicons
-from SALDO, described in \ref{sec:extractsaldo}
+from SALDO, described in section \ref{sec:extractsaldo}
 
 \section{Related work}
-%Copied form thesis!!!! 
-% Add: how are these related to this work?
 % Add: detaild comparison of outputs of various parsers of Sw.
@@
 arguments. \\
-
 The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building
@@
 which focuses on describing and translating verb frames.\\ 
-
 The Swedish version of the Core Language Engine (CLE) \cite{gamback}
@@
 translation to English  was implemented and the work was further developed in the spoken
-language translator \cite{spoken}. Unfortunately, it is no longer available. The coverage of the Swedish
-CLE is also reported to be very limited \cite[p. 134]{nivreTrees}.\\
-
-In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
+language translator \cite{spoken}. Unfortunately, it is no longer available.\\
+%The coverage of the Swedish
+%CLE is also reported to be very limited \cite[p. 134]{nivreTrees}.\\
+In the TAG formalism \cite{tag}, there are projects on getting open-source, wide-coverage grammars
 for English and Korean, but, to our knowledge, not for Swedish.  \\
-
 The ParGram \cite{pargram} project aims at making wide coverage grammars using
+the Lexical Functional Grammar approach \cite{lfg}.
 The grammars are implemented in parallel in order to coordinate the analyses of
 different languages and there are now grammars for English, German, Japanese and Norwegian.
+
+Extract \cite{MarkusForsberg2007} is a tool for lexicon
+%TODO
+extraction compatible with GF, sharing its basic ideas with our lexical
+acquisition tool. Extract does however not consider parts-of-speech and our
+tool is developed closer to GF, while still
+having additional mechanisms for being robust and human support to be accurate.
 
@@
 In a few cases the smart paradigms could not generate the correct declination.\\
-
 When testing the coverage of Talbanken,
@@
 an additional analysis identifying compounds is preformed before
-looking-up the words in the lexicon. We
-should also take into consideration that 
-we cannot automatically find out how many actually stem from the same word, or
-how many abbreviations that are present. Talbanken also contains a small number of 
+looking-up the words in the lexicon. 
+Talbanken also contains a small number of 
 spelling errors, which probably are enumerated among our missing words. The majority
 of the missing words are only used once.\\
-
 \begin{figure}[h]
@@
 \end{figure}
-
 A list of words that were given different labels in GF than in Talbanken has been
@@
 Valency information, which is crucial for GF, is not given in SALDO and
-hence not in the imported lexicon. Instead we extract this from
+hence not in the imported lexicon. Instead we are working on methods
+of extracting this from
 Lexin\footnote{http://spraakbanken.gu.se/lexin/}.
@@
 
+%TODO 
+%our largest set of testdata came from talbanken, about 2000? words. Not to large,
+%could be better if more forms appeared more often.
 The tool has been tested on verbs. Although using simple techniques it 
@@
 
-\section{Translating The Treebank Talbanken}
+\section{Mapping of Talbanken trees}
 \label{sec:mapping}
@@
 of the sentence ``Katten p{\aa} bilen blir st{\"o}rre" 
-(\emph{``The cat on the car gets bigger"}).
-The mapping is based on 
-a translation of the English Penn Treebank \cite{gfpenn}.
-By modifying this program, we now have a translation that works for
-the Tiger XML-format of
-Talbanken. Adaption was required for the differences in annotation as well as 
-for the syntactic differences between Swedish and English.
+(\emph{``The cat on the car gets bigger"}) and its translation to GF.
 
@@
 every Talbanken-tag has been implemented.
-Shorter sentences have been prioritized in order to get a coverage of the most
-fundamental constructions. 
-Regression testing was used to ensure that a new rule would not violate the old ones
-or allow trees to be translated erroneously.
-
-The flat version of Talbanken05 was used for developing the mapping, but
-the program works for both the flat and the deep Tiger-XML versions.
-The focus on the mapping has been simple sentences, in order to cover the most
-fundamental structures.
+Shorter sentences, with less than 10 words, have been prioritized in order to
+get a coverage of the most fundamental constructions.
 
@@
 
-A mapping between GF and the Wall Street Journal of Penn Treebank has earlier
-been conducted \cite{gfMech}.
-The percentage of restored nodes from Peen Treebank is higher than our results.
-The reason for this may be the fact that English is syntactically less complicated
-than Swedish.
-Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
-where idiomatic expressions are more likely the be appear and the language
-presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
-Also, the Penn Treebank contains a lower number of tags, 82 compared
-to more than 300 in Talbanken. Even if the tags describing the same word
-class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
-With more tags, we get more information, but as the number 
-increase, so does the amount of work of finding the correct translation 
-for each combination of tags and writing rules that cover all constructions.
-
-We believe that the results could be enhanced by simply adding more rules 
-and in this way get a wider coverage. There are many special cases that require
-special rules. Since we are not aware of any formal specification of how the tags may be
-combined in Talbanken, the only way of finding all possibilities are to manually look
-for different patterns. Another option would be to make the mapping more
-robust, but the robustness must not interfere with the correctness.
-
 \section{Development of the grammar}
@@
 
-It has earlier been hard to identify missing constructions of the Swedish
-implementation, since there was no large resource available to evaluate it on.
-Our evaluations are based on Talbanken, and when first conducting tests,
-we found much room for improvement.
+For Swedish, about 85\% of the GF resource code is shared with the other Scandinavian
+languages. 
+However, if we aim for a deeper and more comprehensive analysis of Swedish,
 the implementation of the languages needs to be more independent.
@@
 imperatives, relative clauses, cleft constructions etc.
-During our work the grammar has been extended to cover language specific
-constructions such as topicalisation and the features listed below. We further
-put effort on avoiding and removing overgeneration, which might pass unnoticed
-when parsing controlled text but which cause an explosion of ambiguous parse trees
-when used for free text and with a large lexicon.
-\subsection{The s-passive}
-Passive voice is commonly used in Swedish.
-There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli'
-(\emph{`become'}) and the \textbf{s-passive} which is formed by adding an
-\emph{s} to the verb: \\
-\enumsentence{
-\shortex{5}{Uppsatsen & skrev\textbf{s} & av & en & student.}
-        {\emph{the essay} & \emph{wrote+\textbf{s}} & \emph{by} & \emph{a} & \emph{student}}
-        {``The essay was written by a student."}}\label{sent:skrevs}
-Some studies suggest that the s-passive is used in more than 80 \% of the times
-\cite{laanemets}.
-It is however not as common in the other Scandinavian languages,
-and the resource grammar for Scandinavian therefore only implemented a function for
-passive by using auxiliary verb.
-
-During this project, the s-passive was added and the
-periphrastic passive is still allowed.
-The grammar further allows not only two-place verbs, but all verb phrases missing an object, to 
-form passive:
-\enumsentence{\textbf{Active use of two-place verb}\\\shortex{4}
-{Vi & erbj{\"o}d& henne& jobbet }
-{\emph{we} & \emph{offered}& \emph{her}&\emph{the job}}
-{``We offered her the job"}}
-\label{sent:give2pass}
-\enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
-{Hon & erbj{\"o}ds & jobbet}
-{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
-{``She was offered the job"}}
-\label{ex:passV33}
-\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
-{Jobbet & erbj{\"o}ds & henne }
-{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
-{``The job was offered to her"}}
-\label{ex:passV32}
-\enumsentence{\textbf{V2A verb}\\\shortex{3}
-{Huset & m{\aa}lades & r{\"o}tt}
-{\emph{the house} & \emph{painted+\textbf{s}}& \emph{red}}
-{``The house was painted red"}}
-\label{ex:passV2A}
-
-\subsection{Impersonal constructions}
-\label{sec:Formal}
-Formal subjects \cite[\textsection 19]{SAG} are often used in
-Swedish.
-\enumsentence{\shortex{6}
-{Det & sitter & en & f{\aa}gel & p{\aa} & taket}
-{\emph{it }& \emph{sits }& \emph{a} & \emph{bird} &\emph{ on} & \emph{the roof}}
-{``There is a bird sitting on the roof''}}
-\emph{`Det'} has the position of the subject, and the real subject, 
-\emph{`en f{\aa}gel'} the one of an object.
-
-Our grammar implementation allows impersonal constructions 
-and does not allow ungrammatical variants such as sentence
-(\ref{sent:fagelfron}) where a transitive verb is used.
-
-\enumsentence{\shortex{5}
-{Det&dricks&mycket & {\"o}l& nuf{\"o}rtiden}
-{\emph{It}&\emph{drinks+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
-{``A lot of beer is being drunk these days"}}
-
-The grammar also respects the restrictions on the real subject, which cannot
-contain a determiner requiring the noun or modifying adjective to be definite form \cite{Cooper}
-\enumsentence{\shortexnt{6}
-{*Det     & sitter     & \textbf{den}& \textbf{f{\aa}geln}& p{\aa}       & taket.}
-{\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}}
-}
-\enumsentence{\shortexnt{6}
-{*Det& sitter& \textbf{min}& \textbf{f{\aa}gel}& p{\aa}& taket}
-{\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}}
-}\label{sent:dumfagel}
-
-\subsection{Formalizing the rules for reflexive pronouns by using dependent types}
-\label{sec:reflexives}
-An important area in a Swedish grammar is the treatment of the reflexive pronouns and
-the reflexive possessive pronouns.
-The reflexives require an antecedent with which they agree in
-number and person.
-Our grammar accept the following sentences:
-
-\enumsentence{Han s{\aa}g sina barns skor.\\ \label{ex:ReflGen}
-He saw {\sc self's} children's shoes.}
-\enumsentence{Sina vantar hade han gl{\"o}mt p{\aa} t{\aa}get.\\
-{\sc self's} gloves, he had forgotten on the train.}
-\enumsentence{Hon ber alla sina kompisar att g{\aa}\\
-She asks all {\sc self's} friends to leave}\label{ex:ReflPredet}
-\enumsentence{Jag vill att han rakar sig.\\
-I want him to shave {\sc self}.}
-\eenumsentence{
-\item Hon tyckte om skolan och alla sina elever.\\
-     She liked the school and all {\sc self's} students.
-\item Han s{\aa}g sina f{\aa} b{\"o}cker och sin penna.\\
- He saw {\sc self's} few books and {\sc self's} pencil.
-}\label{ex:ReflConj} \label{ex:ReflConj2}
-Reflexive pronouns can not be used in subject noun phrases of finite sentences,
-as shown by the ungrammatical examples in 
-sentence (\ref{sent:sesig}) and (\ref{ex:ReflSubj}).
-The third person reflexives (`sig',`sin') requires a third person antecedent (see \ref{ex:refljag}).
-Furthermore, the antecedent must be within the same finite sentence as the reflexive
-pronoun, see (\ref{sent:sesig}).
-The grammar does not accept any of these sentences:
-\enumsentence{
-*Sina vantar var kvar p{\aa} t{\aa}get.\\
-{\sc self's} gloves were left on the train.}\label{ex:ReflSubj}
-\enumsentence{*Han och sin kompis läser en bok.\\
-He and {\sc self's} friend are reading a book. }
-\enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give {\sc self's} money to {\sc self's} siblings. }\label{ex:refljag}
-\enumsentence{*Han vill att jag ser p{\aa} sig.\\
-He wants me to look at {\sc self}.}\label{sent:sesig}
-
-In order to achieve a scalable implementation of reflexive pronouns, we needed
-to diverge from the old abstract implementation. To capture the agreement
-between the antecedent and the reflexive pronoun, while still allowing the grammar
-to accept reflexive constructions embedded in adverbial and adjectival phrases,
-in fronted positions etc. we make use of the dependent types of GF. This is the
-first use of the dependent types in the resource grammars and a could be
-extended to other languages having similar dependencies.
-
-\subsection{A second future tense: ``Kommer att"}
-Swedish use two auxiliary verbs for future tense \\
-\cite[p. 246]{H&H}:
-\eenumsentence{
-\item {
-Det \textbf{kommer att} bli m{\"o}rkt snart\\
-``It will get dark soon" }
-\item {
-Jag \textbf{ska} g{\aa} och l{\"a}gga mig nu\\
-``I'm going to bed now"\\}
-}
-The resource grammar included 
-\emph{`ska'}, which was implemented as the standard way of forming future
-tense, and hence represents the translation of
-\emph{`will'}. 
-The new grammar also supports \emph{``kommer att"}, expressed as an alternative
-future tense.
-\begin{figure}[h]
-\begin{verbatim}
-(UseCl (TTAnt TFutKommer ASimul) PPos 
-       (ImpersCl (AdvVP (ComplVA become_VA 
-                (PositA dark_A)) soon_Adv))
-\end{verbatim}
-\caption{Result of parsing ``Det kommer att bli m{\"o}rkt snart". The future tense
-         is marked by the constant TFutKommer}
-  \label{fig:kommeratt}
-\end{figure}
-
-\subsection{Testing}
-Throughout the project regression testing has been used. Every time the grammar was updated
-it was tested against a treebank consisting of 155 sentences, and the number of
-parse trees were compared to a standard.
-The purpose was first of all to
-make sure that the coverage was not decreased, but also to make sure that we
-would notice if new ambiguities or overgenerating functions were introduced.
-If so, the ambiguities could be removed when possible, and when not, we were made
-aware of them and could decide whether the increase of coverage were worth the
-increase in ambiguities.
-
-New functions were also tested by random generating and by creating tables showing
-all forms of the added construction.
+A number of constructs that are generally not present in other languages and
+therefore not given by the resources, have also been added. These include the use of
+the reflexive pronoun \emph{sitt}: \\
+\emph{Han s{\aa}g \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw SELF's house}) $\;$ \\
+as opposed to \\
+\emph{Han s{\aa}g \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
+Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
+\emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
+This sort of rephrasing is not given by the resource grammar, since it
+has little effect on the logical representation.
 
@@
 \item a comparison between GF and another annotation
-\item a deeper testing of the Swedish resource grammar and an estimation
-%TODO lexin, chunking?
-of how well GF can be used to describe larger parts of a language
-\item a study of how dependent types can be used in the resource grammars
 \end{itemize}
@@
 which works fast even when using an extensive lexicon.
-Although it is not been formally verified, we believe that the majority of the
-sentences generated are grammatically correct in a syntactical point of view.
-Without any semantic knowledge, nonsense phrases cannot be avoided in
-random generation. However, given that the abstract tree has a meaningful
-interpretation,
-the linearization should be correct.
+By the renewed import of SALDO, we have doubled the size of the lexicon and thereby
+added many of the commonly used words that were missing from the older
+version. This is of course a big improvement.
+However, the lexical part still requires some work before it can be made use of.
+The lexicon is too big to use with the current techniques as its size exhausts the
+the current incremental parsing algorithm. However, new research is being conducted
+to improve the GF run-time system.
+%TODO can I write that?
 
@@
 has increased the standard substantially.\\
-
-By the renewed import of SALDO, we have doubled the size of the lexicon and thereby
-added many of the commonly used words that were missing from the older
-version. This is of course a big improvement.
-However, the lexical part still requires some work before it can be made use of.
-The lexicon is too big to use with the current techniques. Its size causes the
-incremental parsing algorithm to use more heap memory than normal computers
-allow. To solve this, we need to use the lexicon data more cleverly.
-
 We are currently working on chunk parsing and disambiguation using probabilities
@@
 
-% Add: valencies extracted, compounds from språkbanken
 \section{Conclusion}
@@
 describing language. 
-
-We have implemented a grammar covering an important fragment of Swedish by
-focusing on constructions that are frequent in Talbanken.
+We have developed tools for extending the lexicon with words from Talbanken,
+and we show how to make use of the information in the manually tagged
+treebank.
 \\
-
 All parts of the project are open-source and may thus be used in other applications.
@@
 as it increases the coverage of the Swedish resource grammar.
-
-The constructions
-that we have focused on have all been possible to implement, with varying  amounts
-of work. Many of them could be done by utilizing and extending the resource library
-but in some cases we needed to part from the multilingual abstract and use other
-grammatical theories in order to arrive at a good analysis.
 
@@
 \section{Acknowledgments}
-We would like to thank Center of Language Technology for funding this project.
-Further, special thanks to Aarne Ranta,
+The work has been funded by Center of Language Technology.
+We would also like to give special thanks to Aarne Ranta,
 Elisabet Engdahl, Krasimir Angelov, Olga Caprotti, Lars Borin and John Camilleri for their help
 and support.
-%\nocite{*}
 

### 2012-03-12 robust/chunk/MkLex.hs
chunking better but has weird bug

@@ -13,19 +13,21 @@ import ReadCommand
 type LexMap = (Lex,Lex)
 type Lex    = M.TMap String String
 
-mkDir ::LexMap -> [Lemma] -> IO (Maybe (FilePath,Language))
-mkDir lexs lem = do
+mkDir ::LexMap -> [Lemma] -> Int -> IO (Maybe (FilePath,Language))
+mkDir lexs lem i = do
    let lems = map head $ group $ sort lem
    putStrLn $ "Lemmas wanted "++show lems
    putStrLn "Extracting new lexicon..."
-   (t,(cnc,abs)) <- timeItT $ extractLemmas lexs lems
+   (t,(cnc,abs)) <- timeItT $ extractLemmas lexs lems 
    putStrLn $ "Compiling new grammar, extraction: "++ show t
    (t',(ex,res,err)) <- timeItT $ readProcessWithExitCode' "." "gf"
                               ["--make",parseGF] []
    putStrLn $ "compilation: "++ show t'
    case ex of
         ExitSuccess   -> do putStrLn "Done"
-                            return $ Just (newPGF,newLang)
+                            let thisPGF = "BigParse"++show i++".pgf"
+                            runCommand $ "mv \"BigParse.pgf\" "++ thisPGF
+                            return $ Just (thisPGF,newLang)
         ExitFailure _ -> do putStrLn "PGF error"
                             putStrLn err
                             return Nothing

### 2012-03-12 robust/chunk/NewerChunk.hs

@@ -8,6 +8,7 @@ import Data.Ord
 import qualified Data.Map as M
 import Data.Either
 import Data.Tree
+import System.IO
 import Types
 
 data PState = PS {recTypes      :: [[Type]]       --stack of types which unknown words should be parsed as
@@ -19,6 +20,8 @@ data PState = PS {recTypes      :: [[Type]]       --stack of types which unknown
                  ,replaced      :: [(Int,[Expr])]
                  ,counter       :: Int
                  ,pgf           :: PGF
+                 ,isInnerS      :: [()]
+                 ,gettingPieces :: Bool
                  ,trace         :: [String]
                  }
 type Parser = State PState
@@ -36,19 +39,20 @@ limit = 500
 chunklimit = 200
 disambiglimit = 700
 
--- TODO chunkade bitar ska disambigueras på en gång
 parseText :: Tree String -> PGF -> Language -> Type -> IO [Expr]
 parseText tree pgf lang startType = do
+  hSetBuffering stdout LineBuffering
   let (success,st) = runState (parseX tree) parser 
-  appendFile "outputdis" (unlines $ trace st++["\n"])
-  limitOutput success st 
+  res <- limitOutput success st 
+  appendFile "outputdis2" (unlines $ trace st++["\n"])
+  return res
   --appendFile "outputdis" (unlines $ trace st'++["\n"])
   --getOutput success st'
 
  where parser     = PS {recTypes = [[sent]], recState =  [[Ok startState]], skip = [False]
                        ,currentStates = [Ok startState], pieces =  []
                        ,emptyPState = initState pgf lang
-                       ,replaced = [], counter = 0
+                       ,replaced = [], counter = 0, isInnerS = [], gettingPieces = False
                        ,pgf = pgf, trace = []}
        startState = initState pgf lang startType
       
@@ -73,8 +77,8 @@ parseText tree pgf lang startType = do
                isOkParse _           = False
         
                rankThem trees = map fst $ rankTreesByProbs pgf trees
-               getOutput success st | success && canProduce = return $ take limit trees
-                                    | otherwise             = take limit <$> putPiecesTogether st
+               getOutput success st {-| success && canProduce = return $ take limit trees
+                                    | otherwise             -}= take limit <$> putPiecesTogether st
                                                               --TODO limit shouldn't be needed here
 
[…]

### 2012-03-12 robust/chunk/NewerTest.hs

@@ -16,17 +16,23 @@ import System.IO
 import System.TimeIt
 
 main = do hSetBuffering stdout LineBuffering
-          timeIt $ tryAll "hardtests.xml"
+          timeIt $ tryAll "stest.xml"
 
 tryAll fil = do
   putStrLn $ "Start, reading saldo ..."
-  lex <- getSaldo 
+  saldo <- getSaldo 
   putStrLn $ "Reading pgf ... "
   pgf <- readPGF pgfBigFile
   putStrLn $ "Building morpho ... "
   let morpho = buildMorpho pgf langBig
   putStrLn $ "Parsing xml ... "
-  input <- fmap concat $ Form.parse fil
+  input <- splitFile $ fmap concat $ Form.parse fil
+  --input <- fmap concat $ Form.parse fil
+
+  pgfs <- mapM (extractDicts saldo morpho) [x | x <- zip input [0..]]  --write to file, input included
+  mapM processPartition pgfs
+
+  {-
   putStrLn $ "Processing the tree "
   let ziptrees        = map (getFirstWord . fromTree . snd) input
       (newsTrees,res) = runWriter (mapM (processTree True lex morpho 0) ziptrees)
@@ -34,15 +40,30 @@ tryAll fil = do
   writeFile "namesEx" $ formatNames names
   lex <- mkLexMap
   Just (pgf',newLang) <- mkDir lex lms 
-  newPgf <- readPGF pgf' 
-  probs <- readProbabilitiesFromFile "ProbsNewSpec" newPgf
-  let goodPgf = setProbabilities probs newPgf
-  putStrLn $ "Parsing the tree "
-  mapM (processAndWrite goodPgf newLang) $ zip (map fst input) (map toTree newsTrees)
+  -}
+
  where processAndWrite pgf lang (i,tree) = do
-         res <- parseText tree pgf lang startType
-         appendFile "disambigtest3.txt" $ showRes (i,res) ++"\n"
+          res <- parseText tree pgf lang startType
+          appendFile "disambigtest4.txt" $ showRes (i,res) ++"\n"
         where ziptree = getFirstWord $ fromTree tree
+
+       extractDicts saldo morpho (input,i) = do
+          putStrLn $ "Processing the tree "
+          let ziptrees        = map (getFirstWord . fromTree . snd) input
+              (newsTrees,res) = runWriter (mapM (processTree True saldo morpho 0) ziptrees)
+              (lms,names)     = res
[…]

### 2012-03-12 robust/chunk/ProcessTree.hs

@@ -13,6 +13,7 @@ import Compound
 
 -- compound: kattpälsen  -> [(katt &+ pälsen)]
 -- ne      : Johan Ros   -> ["X1",""](
+--TODO write names to state!!
 
 type ProcessState = Writer ([Lemma], [(Int,String)])
 

### 2012-03-12 robust/chunk/Types.hs

@@ -63,7 +63,8 @@ advs   = [adv,adV,iadv]
 --pol    = fromJust $ readType "Pol"  --no, this must be some special, just string "inte"
 --advs = [advsub,advobj]
 nps  = [npsub,npobj,np]
-verbs   = [v,v2] -- should contain all, but never used...
+verbs   = [v,v2] --,v3,vs,vq,va,v2a,v2q,vv,v2v,v2s] 
+phrText = fromJust $ readType "PhrText"
 --aps  = [apsub,apobj]
 toGFStr :: [Type] -> Maybe String
 toGFStr xs  | xs == advs  = return "?advs"

### 2012-03-12 robust/chunk/imp.sh

@@ -1,0 +1,3 @@
+export GHC_INCLUDE_DIRS='../../mapping:../../webpage/:../../'
+echo "it does smth"
+echo $GHC_INCLUDE_DIRS

### 2012-03-14 mapping/Translate.hs
paper to tsd on the way

@@ -3,31 +3,53 @@ module Translate where
 import MonadSP
 import Idents
 import Test  
-import Structure
-import qualified State as S
 import qualified Format as Form
-import GraphTree
 
-import Control.Arrow hiding ((<+>))
+import PGF hiding (Tree,parse)
+import Control.Arrow
 import Control.Monad
-import Control.Monad.RWS hiding (gets,modify,local)
-import Data.Maybe
-import Data.Char
-import Data.IORef
-import qualified Data.Label as P 
-import Data.Label.PureM hiding (local)
-import Data.List
-import Data.Tree
-import PGF hiding (parse)
+import Control.Monad.RWS hiding (put,gets,get)
 import System.IO
 import System.Process
+import System.FilePath
+import Data.Maybe
+import Data.List
+import Data.IORef
+import Data.Char
+import Data.Tree
+import Data.Label 
+import Data.Label.PureM 
+import GraphTree
 
 -- Test by runnig mainTest. Use testGr, otherwise very slow
 
 -- rember to fix new NP2
 
-type PMonad = (RWS () [String] S.State)
-type Translator a = P String Expr PMonad a
+data SentenceType = Q | D | F
+  deriving (Show,Eq)
+data VPForm  = Cop | Sup | VV | VA 
+             | V | V2 | V2A | V2Pass 
+             | Fut | FutKommer
+             | VS  
+             | V3  | VQ | V2V | V2S  --new
+
+
+
+  deriving (Eq,Show)
+vForms    = [Cop,Sup,Fut,FutKommer,VV,VA,V2A,V2,V2Pass,VS,V,  V3,VQ,V2V,V2S]
+gfvForms  = ["VV","VA","V2A","V2","VS","V","V3","VQ","V2V","V2S"] -- need more?
+
+
+data S = S { _isReflGenVP  :: Bool
+           , _isExist      :: Bool
+           , _iquant       :: Bool
+           , _passive      :: Bool
+           , _sentenceType :: SentenceType
+           , _complement   :: (VPForm,[Maybe Expr])
+           , _object       :: Maybe Expr
+           }
+
+type PMonad = (RWS () [String] S)
 
 
 test = False
@@ -37,7 +59,20 @@ bigGr  = ("../gf/Big.pgf","BigSwe")
 lang   = fromJust $ readLanguage "BigTestSwe"
 paint  = False
 
-tb = "../Talbanken05_20060604/FPS/P.tiger.xml"
[…]

### 2012-03-14 robust/chunk/NewerChunk.hs

@@ -1,4 +1,3 @@
-{-# LANGUAGE ViewPatterns #-}
 module NewerChunk where 
 import PGF hiding (Tree)
 import Control.Monad.State hiding (ap)
@@ -9,7 +8,6 @@ import Data.Ord
 import qualified Data.Map as M
 import Data.Either
 import Data.Tree
-import System.IO
 import Types
 
 data PState = PS {recTypes      :: [[Type]]       --stack of types which unknown words should be parsed as
@@ -18,80 +16,49 @@ data PState = PS {recTypes      :: [[Type]]       --stack of types which unknown
                  ,currentStates :: [Result]       --the current parse state
                  ,pieces        :: [[Expr]]       --relevant pieces parsed so far
                  ,emptyPState   :: Type -> ParseState --an empty parse state
-                 ,replaced      :: [(Int,[Expr])]  --list of expressions to replace the ints with later
-                 ,counter       :: Int             --counter for replacing
-                 ,pgf           :: PGF             --the pgf
-                 ,isInnerS      :: [()]            --stack keeping track of if a sentence is embedded or not
-                 ,gettingPieces :: Bool            --are we parsing chunks or the whole sentence?
-                 ,trace         :: [String]        --logging trace
+                 ,trace         :: [String]
                  }
 type Parser = State PState
 data Result = Ok ParseState | Recover ParseState | Failed 
 instance Eq Result where
     (Ok _)      == (Ok _)      = True
     (Recover _) == (Recover _) = True
-    (Failed)    == (Failed)  = True
+    (Failed)  == (Failed)  = True
     _           == _           = False
 
 type Success = Bool
 
-limit, chunklimit, disambiglimit :: Int
-limit = 500
-chunklimit = 200
-disambiglimit = 700
+limit = take 500
 
 parseText :: Tree String -> PGF -> Language -> Type -> IO [Expr]
 parseText tree pgf lang startType = do
-  hSetBuffering stdout LineBuffering
-  let (success,st) = runState (parseX tree) parser 
-  res <- limitOutput success st 
-  appendFile "outputdis2" (unlines $ trace st++["\n"])
-  return res
-  --appendFile "outputdis" (unlines $ trace st'++["\n"])
-  --getOutput success st'
+  let startState = initState pgf lang startType
+      parser     = PS {recTypes = [[sent]], recState =  [[Ok startState]], skip = [False]
+                      ,currentStates = [Ok startState], pieces =  []
[…]

### 2012-03-14 doc/paper/final/PaperFin.tex
renewed planning
planning
final lrec paper!!

@@
 
+%TODO check ref-list
+
 % 150-200 words here
@@
 has so far been used successfully for
-controlled languages \cite{cnl}. Recent experiments have showed
+controlled languages \cite{cnl} and recent experiments have showed
 that it is also possible to use the framework for parsing unrestricted language.
@@
 to GF, we use two other main resources: the Swedish treebank Talbanken and the
-electronic SALDO lexicon. By combining the grammar with
+electronic lexicon SALDO. By combining the grammar with
 a lexicon extracted from SALDO we obtain a parser accepting all sentences
@@
 grammar may be used in various NLP applications. 
-\\ \newline \Keywords{GF, Swedish, Parsing, Grammar Formalisms, Lexical Acquisition}}
+\\ \newline \Keywords{GF, Swedish, Parsing}}
 
@@
 %in 2005 \cite{talbanken05} and
-enriched with annotation for a full phrase structure analysis \cite{talbanken05}.  \\
+enriched with annotation for a full phrase structure analysis \cite{talbanken05}.  
 Although Talbanken contains both written and spoken Swedish,
@@
 technology tools for Swedish.
-An example is the well-known data-driven Malt parser \cite{malt},
+An example is the well-known data-driven Maltparser \cite{malt},
 trained on Talbanken. 
@@
 The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
-Scandinavian, while Norsource (Hellan, 2003) describes Norwegian. A Swedish version
+Scandinavian, while Norsource \cite{norsource} describes Norwegian. A Swedish version
 was based upon this by Ahrenberg, covering the morphology and some
@@
 %CLE is also reported to be very limited \cite[p. 134]{nivreTrees}.\\
-In the TAG formalism \cite{tag}, there are projects on getting open-source, wide-coverage grammars
+In the TAG formalism \cite{tag}, there are projects in getting open-source, wide-coverage grammars
 for English and Korean, but, to our knowledge, not for Swedish.  \\
@@
 tool is developed closer to GF, while still
-having additional mechanisms for being robust and human support to be accurate.
+having additional mechanisms for robustness and human support for accuracy.
 
@@
 
-\subsubsection{Implementation}
+\subsection{Implementation}
 The basic algorithm for importing SALDO was implemented by Angelov (2008)
@@
 \verb-mkV "knyter" ;-
-\includegraphics[scale=0.5]{image1.eps} 
-\caption{First code produced for the verb `knyta' (\emph{`tie'})}
+\caption{First code produced for the verb \emph{knyta} (`tie')}
 \label{fig:saldoknyt}
@@
 \verb-mkV "knyter" "knöt" ;-
-\caption{Second output for the verb `knyta'}
+\caption{Second output for the verb \emph{knyta}}
 \label{fig:saldoknyt2}
@@
 The program is run iteratively until the GF table matches the one given in SALDO,
-or until there are no more ways of using the smart paradigm. The verb 'knyta'
+or until there are no more ways of using the smart paradigm. The verb \emph{knyta}
 will need three forms:\\
@@
 \verb-mkV "knyter" "knöt" "knutit"-\\
-\caption{Final output for the verb `knyta'}
+\caption{Final output for the verb \emph{knyta}}
 \label{fig:saldoknyt3}
@@
 
-\subsubsection{Results}
+\subsection{Results}
 \label{sec:saldoRes}
@@
 \hline
- Adverb & \textbf{ab}$/$\textbf{Adv} & ofta (\emph{often})\\
- Adjective&\textbf{av}$/$\textbf{A} & gul (\emph{yellow})\\
- Noun & \textbf{nn}$/$\textbf{N} & hus (\emph{house})\\
- Verb & \textbf{vb}$/$\textbf{V} & springa (\emph{run})\\
- Reflexive verbs &\textbf{vbm}$/$\textbf{V} & raka sig (\emph{shave})\\
- Particles verbs &\textbf{vbm}$/$\textbf{V}  &  piggna till (\emph{perk up})\\
+ Adverb & \textbf{ab}$/$\textbf{Adv} & \emph{ofta} (`often')\\
+ Adjective&\textbf{av}$/$\textbf{A} & \emph{gul} (`yellow')\\
+ Noun & \textbf{nn}$/$\textbf{N} & \emph{hus} (`house')\\
+ Verb & \textbf{vb}$/$\textbf{V} & \emph{springa} (`run')\\
+ Reflexive verbs &\textbf{vbm}$/$\textbf{V} & \emph{raka sig} (`shave')\\
+ Particle verbs &\textbf{vbm}$/$\textbf{V}  &  \emph{piggna till} (`perk up')\\
 \hline
 \end{tabular}
-\caption{}
+\caption{Word classes imported from SALDO}
 \end{figure}
@@
 Consider for example the plural tantum 
-noun \emph{`glas{\"o}gon'} (\emph{`glasses'}).
+noun \emph{glas{\"o}gon} (`glasses').
 The smart paradigm requires a singular form, and since the program could not
@@
 When the program failed to import a noun, this was often the explanation.
-Words of this type may be added manually, for \emph{`glas{\"o}gon'} we could use
-the ostensibly correct singular form `glas{\"o}ga', although this
+Words of this type may be added manually, for \emph{glas{\"o}gon} we could use
+the ostensibly correct singular form \emph{glas{\"o}ga}, although this
 has another meaning (`glass-eye').
-The same problem occurred for the irregular s-verbs,
-(\emph{`synas'} \emph{(`show')} or \emph{umg{\aa}s} \emph{(`socialize')})
+The same problem occurred for the irregular s-verbs
+(\emph{synas} (`show') or \emph{umg{\aa}s} (`socialize'))
 which made up 61.5 \% of the failing verbs of type \verb_vb_.\\
@@
 \end{tabular}
-\caption{}
+\caption{Number of Talbanken words still missing}
 \end{figure}
@@
 allow the word to be added to the lexicon, remove it or request another guess.
-The user hence only needs to decide if each paradigm is correct or not, and
-the program will do the rest.
+The user hence only needs to decide if each paradigm is correct or not.
 
@@
 the syntactic information. 
-\begin{figure}[h!]
+\begin{figure}[!h]
 \begin{center}
@@
 \subfloat{\label{pic:gftree}\includegraphics[width=90mm]{gfTree.pdf}}
-\caption{The Talbanken tree and the GF tree for the sentence ``Katten p{\aa} bilen blir st{\"o}rre".}
+\caption{The Talbanken tree and the GF tree for the sentence \emph{``Katten
+p{\aa} bilen blir st{\"o}rre".}}
 \label{fig:translationtrees}
@@
 Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
-of the sentence ``Katten p{\aa} bilen blir st{\"o}rre" 
-(\emph{``The cat on the car gets bigger"}) and its translation to GF.
+of the sentence \emph{``Katten p{\aa} bilen blir st{\"o}rre"}
+(``The cat on the car gets bigger") and its translation to GF.
 
@@
 the listings are meant for making the text look nice in
-folders etc and outside the scope for the grammar itself. The tags \textbf{XX}
+folders etc and are outside the scope for the grammar itself. The tags \textbf{XX}
 and \textbf{NAC} are often used since Talbanken makes a
@@
 \enumsentence{F\"or stora krav.\\
-``Too high demands."}\label{sent:krav}
+\emph{``Too high demands."}}\label{sent:krav}
 contains the tags \textbf{XX} and \textbf{NAC}, since it is not obvious
@@
 the reflexive pronoun \emph{sitt}: \\
-\emph{Han s{\aa}g \textbf{sitt} hus} $\; \; \; \;$ (\emph{He saw SELF's house}) $\;$ \\
+\emph{Han s{\aa}g \textbf{sitt} hus} $\; \; \; \;$ (``He saw SELF's house") $\;$ \\
 as opposed to \\
-\emph{Han s{\aa}g \textbf{hans} hus} $\;$ (\emph{He saw his (another person's) house}) \\
+\emph{Han s{\aa}g \textbf{hans} hus} $\;$ (``He saw his (another person's) house") \\
 Fronting words or phrases is very common in Swedish and are now allowed by the grammar:\\
-\emph{\textbf{Glad} var han inte.} $\;$ (\emph{\textbf{Happy} was he not}). \\
+\emph{\textbf{Glad} var han inte.} $\;$ (``\textbf{Happy} was he not"). \\
 This sort of rephrasing is not given by the resource grammar, since it
@@
 \item an extended grammar covering an important part of Swedish
-\item a comparison between GF and another annotation
+\item a comparison and translation between GF and another annotation
 \end{itemize}
@@
 a specialized extension of the resource grammar.
-Besides parsing, the grammar may well be used for language generation,
-which works fast even when using an extensive lexicon.
+Besides parsing, the grammar may well be used for language generation.
+%which works fast even when using an extensive lexicon.
 By the renewed import of SALDO, we have doubled the size of the lexicon and thereby
 added many of the commonly used words that were missing from the older
-version. This is of course a big improvement.
-However, the lexical part still requires some work before it can be made use of.
+version. This is of course a big improvement
+but the lexical part still requires some work before it can be made good use of.
 The lexicon is too big to use with the current techniques as its size exhausts the
-the current incremental parsing algorithm. However, new research is being conducted
+current incremental parsing algorithm. However, new research is being conducted
 to improve the GF run-time system.
@@
 
-Without methods for robust parsing, %TODO evaluate??
-it is not interesting to talk about coverage, but about the quality and
-the ability to scale up, which has so far proved to be good.
-We further believe that the presence of a expert in Swedish, professor Elisabet
-Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl},
-has increased the standard substantially.\\
-We are currently working on chunk parsing and disambiguation using probabilities
+We are currently working on chunk parsing by which we get robustness and a possibility
+to limit the memory usage. We perform disambiguation by using probabilities
 extracted from our translation of Talbanken. We combine this with simple named
@@
 
+%At the current stage, without methods for robust parsing, %TODO evaluate??
+%it is not interesting to talk about coverage, but about the quality and
+%the ability to scale up, which has so far proved to be good.
+%We further believe that the presence of a expert in Swedish, professor Elisabet
+%Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl},
+%has increased the standard substantially.\\
+
 The result will be evaluated both automatically -- by comparing the output of
-the translated trees from Talbanken -- and manually by professor Elisabet Engdahl.
+the translated trees from Talbanken -- and manually by professor Elisabet Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl}.
 She also evaluates the intermediate results.
@@
 treebank.
-\\
+The usage of GF allows us to start from
+a well-defined system for describing grammar, as well as tools for
+parsing.
 All parts of the project are open-source and may thus be used in other applications.
@@
 as it increases the coverage of the Swedish resource grammar.
-
-The usage of GF allows us to start from
-a well-defined system for describing grammar, as well as tools for
-parsing.
 

### 2012-03-14 mapping/Extract.hs
mapping corrected (65%)

@@ -2,5 +2,5 @@ module Main where
 --import LastTranslate
 import Translate hiding (main)
 
-main = do res <- main' "../Talbanken05_20060604/FPS/P.tiger.xml"
-          writeFile "AllTBNew" $ unlines $ getRes res 
+main = do res <- main' tb --"TBSimplified.xml"
+          writeFile "AllTBNewBig" $ unlines $ getRes res 

### 2012-03-14 mapping/TestMapping.hs

@@ -1,7 +1,8 @@
 import Translate hiding (main)
 import System.Environment
 
-main = do
+main1 = do
   fls <- getArgs
   sequence [evaluations file ("EvalResult"++show n) | (file,n) <- zip fls [0..]]
 
+main = main' tb 

### 2012-03-14 mapping/Translate.hs

@@ -3,76 +3,42 @@ module Translate where
 import MonadSP
 import Idents
 import Test  
+import Structure
+import qualified State as S
 import qualified Format as Form
+import GraphTree
 
-import PGF hiding (Tree,parse)
-import Control.Arrow
+import Control.Arrow hiding ((<+>))
 import Control.Monad
-import Control.Monad.RWS hiding (put,gets,get)
+import Control.Monad.RWS hiding (gets,modify,local)
+import Data.Maybe
+import Data.Char
+import Data.IORef
+import qualified Data.Label as P 
+import Data.Label.PureM hiding (local)
+import Data.List
+import Data.Tree
+import PGF hiding (parse)
 import System.IO
 import System.Process
-import System.FilePath
-import Data.Maybe
-import Data.List
-import Data.IORef
-import Data.Char
-import Data.Tree
-import Data.Label 
-import Data.Label.PureM 
-import GraphTree
 
 -- Test by runnig mainTest. Use testGr, otherwise very slow
 
 -- rember to fix new NP2
 
-data SentenceType = Q | D | F
-  deriving (Show,Eq)
-data VPForm  = Cop | Sup | VV | VA 
-             | V | V2 | V2A | V2Pass 
-             | Fut | FutKommer
-             | VS  
-             | V3  | VQ | V2V | V2S  --new
-
-
-
-  deriving (Eq,Show)
-vForms    = [Cop,Sup,Fut,FutKommer,VV,VA,V2A,V2,V2Pass,VS,V,  V3,VQ,V2V,V2S]
-gfvForms  = ["VV","VA","V2A","V2","VS","V","V3","VQ","V2V","V2S"] -- need more?
-
-
-data S = S { _isReflGenVP  :: Bool
-           , _isExist      :: Bool
-           , _iquant       :: Bool
-           , _passive      :: Bool
-           , _sentenceType :: SentenceType
-           , _complement   :: (VPForm,[Maybe Expr])
-           , _object       :: Maybe Expr
-           }
-
-type PMonad = (RWS () [String] S)
+type PMonad = (RWS () [String] S.State)
+type Translator a = P String Expr PMonad a
 
 
 test = False
-usePGF = testGr
+usePGF = newBigGr
 testGr = ("../gf/BigTest.pgf","BigTestSwe")
 bigGr  = ("../gf/Big.pgf","BigSwe")
[…]

### 2012-03-19 robust/chunk/Eval.hs
ner ok, np ration counting and adverbial counting

@@ -1,0 +1,157 @@
+import Control.Applicative
+import Control.Arrow hiding (first)
+import Control.Monad
+import Data.Tree
+import Data.Char
+import Data.List
+import Data.List.Split
+import Data.Maybe
+import qualified Format as Form
+import System.IO
+
+type Id = String
+data Result = Result {idN :: Id, isText :: Bool, total :: Bool, mets :: [String], first :: String
+                     ,tooBig :: Bool}
+ deriving (Show)
+
+tb = "../../Talbanken05_20060604/FPS/P.tiger.xml" 
+
+-- counts the number of sentences for which we could parse the structure, how many that were
+-- totally parsed and how many metas that were in them
+main = do
+   inp <- readFile "helaTBAdv"
+   let ss   = splitWhen null $ lines inp
+       oks  = map mkLine $ filter (not . null) ss
+       res  = (length oks,length (filter isText oks),length (filter (\x -> isText x && total x) oks)
+              ,map (head &&& length) $ group $ sort $ concatMap mets oks)
+   appendFile "Results2" $ show res ++"\n"++ show oks
+
+isOk l = okStart l -- || hasChunk l
+whole x = not $ "Meta" `isInfixOf` x
+       
+-- for a sentence, returns the relevant data (considering only the first parse tree) 
+mkLine :: [String] -> Result  
+mkLine (i:xs) = let nice     = any isOk xs
+                    semiNice = any hasChunk xs
+                    which    = if not nice && semiNice then last else head
+                in  Result i nice (whole $ which xs) (if nice then metas $ which xs else [])
+                           (which xs) semiNice
+ where metas         = filter ("Meta" `isInfixOf`)  . words
+
+
+okStart l = "isText" `isPrefixOf` l || "isUtt" `isPrefixOf` l
+
+
+-- hasChunk also counts the sentences for which there were more parse trees than could be
+-- rated, but have instead been just limited and ranked by the first (20-1000) sentences
+hasChunk l = let (s,rest ) = span (=='s') l
+                 (ns,rest') = span isDigit rest
+                 ok         = "_501" `isPrefixOf` rest'
+                 end        = drop 6 rest'
+             in length s ==1 && not (null ns) && ok && okStart end
+
+-- counts the number of words in each skipped chunk
[…]

### 2012-03-19 robust/chunk/NewerTest.hs

@@ -19,7 +19,7 @@ import System.IO
 import System.TimeIt
 
 main = do hSetBuffering stdout LineBuffering
-          timeIt $ tryAll tb --"hardtests.xml" {- try "stest.xml"-}
+          timeIt $ tryAll  "hardtests.xml" {- tb --try "stest.xml"-}
 
 tb = "../../Talbanken05_20060604/FPS/P.tiger.xml" 
 
@@ -33,14 +33,16 @@ tryAll fil = do
   putStrLn $ "Building morpho ... "
   let morpho = buildMorpho pgf langBig
   putStrLn $ "Parsing xml ... "
-  input <- take 190 . drop 190 . splitFile . concat <$> Form.parseIdTree fil
-  pgfs <- mapM (extractDicts saldo morpho) [x | x <- zip input [191..]]   --OBS
+  input <- take 43 . drop 190 .{- take 22 .-} splitFile . concat <$> Form.parseIdTree fil
+  pgfs <- mapM (extractDicts saldo morpho) [x | x <- zip input [0..]] 
   mapM processPartition pgfs
 
  where processAndWrite :: PGF -> Language -> (String,ChunkTree) -> IO ()
        processAndWrite pgf lang (i,tree) = do
           res <- parseText tree pgf lang startType
-          appendFile "helaTBpart2" $ showRes (i,res) ++"\n"
+          let fil = "night/advTest2"
+          putStrLn $ "write to "++fil
+          appendFile fil $ showRes (i,res) ++"\n"
         where ziptree = getFirstWord $ fromTree tree
 
        extractDicts :: Saldo -> Morpho -> ([(String,ChunkTree)],Int) -> IO ([String],[WorkingTree],Maybe (FilePath,Language))

### 2012-03-19 robust/chunk/NewestChunk.hs

@@ -33,11 +33,13 @@ data PState = PS {currentStates :: [String]       --the current parse state
                  ,saveChunks    :: [(Id,[Expr])]  
                  ,counter       :: Int             --counter for replacing
                  ,lastExpr      :: Maybe (Id,[Expr])
+                 ,npRate        :: (Int,Int) --NPInfo
 --                 ,splitChunks   :: [[String]]
 --                 ,backups       :: 
                  }
 type Parser = WriterT [String] (State PState)
 type Id     = String
+--data NPInfo = NP {tot :: Int, ok :: Int, good :: Int, sents :: [String]}
 
 
 toplimit = 500
@@ -52,7 +54,8 @@ parseText :: Tree (Id,String) -> PGF -> Language -> Type -> IO [Expr]
 parseText tree pgf lang startType = do
   hSetBuffering stdout LineBuffering
   let (trejs,st) = runState (execWriterT (parseX tree)) parser 
-  appendFile "TBUt" (unlines $ trejs++["\n"])
+  appendFile "night/advTest2" (unlines $ trejs++["\n"])
+  appendFile "night/advTest2"   (show $ npRate st)
   let res = lastExpr st
   -- testa att byta alla adv mot adV
 --  ok <- limitAndRank <$> findParse {-(sortBy (comparing noMetas)-} strs
@@ -99,6 +102,7 @@ initPState pgf lang startType =
          ,rankThem   = map fst . rankTreesByProbs pgf
          ,lastExpr   = Nothing 
          ,morpho     = buildMorpho pgf lang
+         ,npRate     = (0,0) --NP 0 0 0 []
          }
 
 startState pgf lang startType = initState pgf lang startType
@@ -147,6 +151,7 @@ parseX' (Node (i,"S") ts)
 parseX' (Node (i,x) ts) 
      | "NN" `isPrefixOf` x && length x>2 = parseX (Node (i,"NN") ts) -- put all variants of NNxxx to NN
 
+     | isAdverbNode x      = putTrace "fixed an adverb" >> return "?advs"
      | isSaveNode x        =  {-# SCC "parseXSave" #-}
                               putTrace ("saves node" ++show x)
                            >> saveState (getCat x) i ts
@@ -165,7 +170,7 @@ parseX' (Node (i,x) (t:ts)) = do
 
 isSaveNode = (`elem` map fst saveNodes)
 
-
+isNP = not . null . (`intersect` nps) 
[…]

### 2012-03-19 robust/chunk/ProcessTree.hs

@@ -68,7 +68,7 @@ exchangeNames i tree | lookNamish word  = moveOnTree (setLabel (id,name)) dropNa
                      | otherwise        = tree
    where word    = snd $ label tree
          id      = fst $ label tree
-         name    = if List.last word == 's' then "XPN"++show i else "YPN"++show i
+         name    = if List.last word == 's' then "XPN "++show i else "YPN "++show i
 
 dropName  :: WorkingTree -> WorkingTree
 dropName tree | lookNamish word || isNameSpec word = moveOnTree (setLabel (id,"")) dropName tree

### 2012-03-19 tsd/TSD2012/paper.tex
tsd paper, future work

--- 2012-01-18 doc/report/report.tex
@@
-\documentclass{report} 
+\documentclass[runningheads,a4paper]{llncs}
+
+\usepackage{amssymb}
 \usepackage[utf8]{inputenc}
-\usepackage{ amssymb }
+\usepackage{subfig}
+\setcounter{tocdepth}{3}
 \usepackage{graphicx}
-\usepackage[margin=1.5in]{geometry}
-\usepackage{parsetree}
+
+\usepackage{url}
 \usepackage{lingmacros}
-\usepackage{verbatim}
-\usepackage{subfig}
-\usepackage{wrapfig}
-\usepackage{tipa}
-\usepackage{textcomp}
-%\usepackage{apacite}
-% exempel: 2) Inte vet jag
-%              not know I
-%             I don't know
-
-\newenvironment{changemargin}[2]{%
-\begin{list}{}{%
-\setlength{\topsep}{0pt}%
-\setlength{\leftmargin}{#1}%
-\setlength{\rightmargin}{#2}%
-\setlength{\listparindent}{\parindent}%
-\setlength{\itemindent}{\parindent}%
-\setlength{\parsep}{\parskip}%
-}%
-\item[]}{\end{list}}
+\newcommand{\keywords}[1]{\par\addvspace\baselineskip
+\noindent\keywordname\enspace\ignorespaces#1}
 
 \begin{document}
-% ska gå att förstå och locka till läsning
-%ska tydligt beskriva vad rapporten handlar om
-%ska innehålla vettiga sökord
-\title{Towards a Wide-Coverage Grammar for Swedish\\ Using GF} %A nice titleScaling up a GF grammar and preparing it for cool stuff for Swedish}
-\author{Malin Ahlberg}
+
+%passwords to the submission: ID 467 klasse
+\mainmatter  % start of an individual contribution
+
+% first the title is needed
+\title{A Type-Theoretical Wide-Coverage Computational Grammar for Swedish}
+
+\author{}
+\institute{}
+
 \maketitle
-\newpage
-
-\abstract{
-This thesis describes work towards a wide-coverage grammar for
-parsing and generating Swedish text.
-We do this by using the dependently typed grammar formalism GF,
-a functional programming language specialized at describing grammars.
-The idea is to 
-%The goal of this thesis is to enable open-domain parsing for Swedish
-%by using the GF \\grammar formalism,
-%by %The idea has been to% 
-combine existing language resources with
-new techniques, with an aim to achieve a parser
-for unrestricted Swedish.
-To reach this goal, problems of computational as well as linguistic nature had
-to be solved.  
-The work includes the development of the grammar -- \;
-%in the GF formalism
-to identify and formalize 
-grammatical constructions %specific to and
-frequent in Swedish -- as well as methods for importing a large-scale lexicon and for evaluating
-the parser. % have been investigated and formalised in the functional
-%programming language.
-%~~achieve a covering Swedish~~ grammar implemented
-%and the aim to reaching our goal;
-%TODO first sentence should be more, no understatements!
-%in GF -- to model natural language by functional programming. 
-We present the methods and technologies used 
-and discuss the advantages and problems of using GF
-for modeling large-scale grammars.
-We further discuss how our long-term goal can be reached
-%the possibilities
-%of reaching our long-term goal % -- to accomplish a parser for unrestricted Swedish -- 
-by combining our rule-based grammar with statistical methods.
-%We compare and discuss .
-%and for obtaining a system for analysing written text.
-\\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
-into the GF notation and an extended Swedish grammar implementation. The grammar is based
-on the multilingual abstract syntax given in the GF resource library, and now
-also covers constructions specific to Swedish.
-We further give an example of the advantage of using dependent types when
-describing grammar and syntax, in this case for dealing with reflexive pronouns.
-%the linguistic phenomena
-
-%This is the report of the work of expanding and enhancing a GF grammar and preparing
-%it for parsing free Swedish. The grammar is a result of combining resources,
-%we have extended the current GF grammar to adapt it for Swedish specifically so
-%that it can parse an language specific constructions. 
-%We show how dependent types can be used to give a neater description of
-%syntax.
-%We have further completed a method for importing the electronic 
-%lexicon Saldo to a GF format and established a connection to the treebank Talbanken.
-%The imported lexicon consists of more than 100 000 words.
-
-%Outside of this project, an equivalent project for English is being worked upon,
-%as are techniques for making the parser robust.
-%With this in mind,
-
-%and the grammar now covers ??\% of the xx part of Talbanken. 
-%66 \% av easyeasy.
-% The mapping between the annotations of Talbanken and GF is automatic. 
-% The goal is to build a robust parser dealing with uncontrolled natural
-%language, a parser which can parse all of Talbanken.
-
-% sf dream to be able to work with npl. Parser important, statistics working
-% on n-grams fail, closer to the human brain with rules? 
-% Describe the basic.. of Swedish, in a format compatible with 20 other languages.
-% example of sentence and how it can be parsed, stepping out of the controlled lang.
-% Combine other resources, create free software.
-% Equivalent for English and other languages coming.
-% We aim for free parsing, this is preparation and up-scaling, finding difficulties.
-}
-
-\newpage
-\pagestyle{empty}
-\section*{Acknowledgments}
-\vspace{5mm}
-Many people have helped me during this project and made this work possible.
-I would first of all like to thank
-Center of Language Technology, that has funded the project.
-
-Further, thanks to my excellent 
-%(~~devoted~~) 
-supervisor Ramona Enache for all her help and guidance 
-%(encouragement)
-in every phase and all aspects of the work.
-Thanks to Elisabet Engdahl for sharing her seemingly unlimited knowledge
-of Swedish grammar. She has also has acted as a second supervisor, and given me
-very helpful comments and suggestions. Thanks to Aarne Ranta for all his great
-ideas and for letting me do this project.
-
-I am also grateful to Krasimir Angelov, Markus Forsberg, Peter Ljunglöf, Lars Borin
-and many others who have contributed with ideas and inspiration and shown %for their
-interest in this work.
-
-Finally, I would like to thank my friends and family. Special thanks to Dan % Rosén
-for all his support, advice and patience and  -- most importantly -- for being
-such a good friend. %for having listened to 
-%my monologues about the daily progress. 
-% Ramona, Aarne, Elisabet, Markus, Peter?, Lars, Dan, Krasimir etc
-% Funding!
-
-% check references! lncs is good maybe. lnai
-\newpage
-%\pagestyle{empty}
-%\cleardoublepage
-\tableofcontents
-%
-\addtocontents{toc}{\protect\thispagestyle{empty}}
-%\pagestyle{empty}
-%\cleardoublepage
-\newpage
-%\pagenumberingstyle{plain}
-\setcounter{page}{1} %reset the page counter
-\chapter{Introduction}
-\thispagestyle{plain}
-% more intro to gf. dependently typed...high-level formalism
-Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
-%with facilities for multilinguality. 
-It is based on Martin-Löf type theory which allows
-reasoning within the programming language.
+
+\begin{abstract}
+The work describes a wide-coverage computational grammar for Swedish. 
+It is developed using GF(Grammatical Framework), a dependently-typed 
+functional language, specialized for grammar programming. We trained
+and evaluated the grammar by using Talbanken, the largest treebank for 
+Swedish. As a result 65 $\%$ of the Talbanken trees were translated
+into the GF format in the training stage. Moreover, we obtained a language
+model for Swedish which we use for disambiguation when parsing free text
+with the grammar. 
+\keywords{grammar-formalism, computational grammar, treebank}
+\end{abstract}
+
+\section{Introduction}
+
+%Motivation + brief information about Swedish+resources.
+
+%Introduction + Aim from Master thesis(Chapter 1 + 1.1).
+
+%Outline of the paper.
+
+\textbf{Ramona: Copied from the thesis, could be improved!}
 GF has strong support for multilinguality and has so far been used successfully for 
-controlled languages \cite{cnl}. Recent experiments have  however showed
+controlled languages \cite{cnl}, while recent experiments have showed
 that it is also possible to use the framework for parsing free language \cite{patent}.
-% >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
-% --> sentence about connection GF - grammar based parser.
-%This project focuses on implementing a wide-coverage Swedish grammar, which
-%describes a parser accepting all sentences that can be expressed by the grammar.
-Parsing, that is the task of automatically identifying morphological and
-syntactical structures, is receiving increasing interest, especially
-considering the steadily growing amount of data available online. 
 So far there is no freely available grammar-driven parser that gives a deep
-analysis for Swedish. The fact that %the existing rule-based Swedish
-a parser or grammar is not freely available does not only restrict its
-usage but also its possibilities of being further developed. Our
-goal is
+analysis for Swedish. The fact that a parser or grammar is not freely available
+does not only restrict its usage but also its possibilities of being further
+developed. Our goal is
 to create an open-source Swedish grammar from which we derive a parser
@@
 made use of in other projects.\\
-
 \noindent To build the grammar from scratch would be not only time consuming but
@@
 file defining a set of rules for morphology and syntax.
-%TODO verify this, check all occurences of grammar
 This is what is meant by grammar in this thesis, as
-%From now on, this is what we mean by grammar, as
 opposed to grammar in the traditional linguistic sense. 
-From GF, we get %and thereby get
-%The usage of GF allows us to start from
-a well-defined system for describing language, as well as a strong connection
-to and possibility of translation between the 
+From GF, we get a well-defined system for describing language, as well as a
+strong connection to and possibility of translation between the 
 more than 20 other languages implemented in the framework. Further, we use
 the extensive lexicon Saldo and the treebank Talbanken.
-%>>The grammar defines a parser, which accepts all constructions defined by the grammar.
-
-\newpage % until it looks nicer
-
-\section{Aims}
-%>> enhance gf...
-\protect\thispagestyle{plain}
-The purpose of this project has been to prepare the GF grammar for parsing of 
-unrestricted Swedish.
-This has meant to develop earlier techniques to fit for Swedish, create methods
-for achieving and keeping a large-scale lexicon and to adapt the existing
-resource grammar to model more language specific constrictions.
-The project was divided into three subsections, aiming at
-%Three sub-goals has been identified:
-\begin{itemize}
-\item Extending the Swedish GF grammar
-\item Importing the lexicon Saldo
-\item Creating translation between Talbanken and GF
-\end{itemize}
-%We hence needed to extend the Swedish grammar to cover
-%the language specific constructions. Further a large lexicon was
-%needed, as well as a method for evaluation.
-%During this project, we have worked on and explored those things and
-%while identifying problems and future work.
-
-\section{Outline}
-The thesis is divided into 6 chapters. We start by giving background
-information of areas relevant to the project. %Chapter \ref{sec:background} 
-Grammatical Framework is presented in section \ref{sec:gf} while a more profound
-description of the Swedish resource grammar is given in section \ref{sec:swegf}.
-Section \ref{sec:swedish} gives an introduction to the Swedish language, and
-brief presentations of Saldo and Talbanken are found in section \ref{sec:saldo} and
-\ref{sec:talbanken} respectively.
-A summary of related work can be found in section \ref{sec:related}.
-
-Chapters \ref{sec:prog.saldo} - \ref{sec:Mapping} present the methodology,
-implementation and some results. First we describe and evaluate the implementation
-of Saldo in chapter \ref{sec:prog.saldo}. The work on the GF grammar is described in
-chapter \ref{sec:prog.grammar}.
-%, section \ref{sec:gfdevel} presents the construction that have been added.
-In chapter \ref{sec:Mapping} we account for an automatic mapping of Talbanken
-trees to GF. %which enables us to extract a large GF treebank.
-
-Finally, the conclusion and evaluation are
-presented in chapter \ref{sec:results} together with some areas of future work.
-
-%We have extended the Swedish resource grammar in GF, to give a better coverage and imported
-%a large lexicon. Work to compare the trees from Talbanken to GF and to translate them.
-
-%Ultimate goal: want computers to 'understand' language. Not good with flat strings.
-%We need to translate strings of natural language to a richer structure,
-%a parse tree. 
-%There are parsers which put only put part-of-speach tags to
-%the text, and others that gives a deeper analyse. 
-%There are statistical parser, and parser which only are deal with controlled language.
-%The GF parser are one of those. 
-%What is parsing, flat strings to trees, example picture, what is deep parsing.
-
-%What is controlled language,
-%what is robust parsing and why we need it for free langugae, benefits of a grammar.
-%GF is grammar formalism, many languages, many projects for controlled lang.
-%GF grammar for Swedish, how big, not intended for parsing, but for building
-%application grammars. Experiments of using the grammar for free parsing, methods
-%being developped for this. Benefits of starting from GF. Free parsing is
-%interesting. To do this, we need a better (bigger grammar), a large scale
-%lexicon, and an evaluation method. This project has explored those things. The
-%grammar also have other usages, important in itself, so we wanted to make it as
-%covering as possible without being overallowing or too ambiguous.
-
-%What we could not cover by this project, what have been covered before (in
-%the spring project), and why we choose those parts.
-
-\chapter{Background}  
-\label{sec:background}
-The work described in this thesis is part of a bigger project which aims at
-using GF for parsing unrestricted Swedish. 
-%TODO LREC ref?
-In previous work\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf}, a start was made to extend the Swedish GF grammar
-and a tool for lexical acquisition was developed.
-We now construct a bigger
-and more expressive grammar as well as a large scale GF lexicon.
-As all GF grammars, this one defines a parser, and we develop it by getting
-examples, ideas and test material from the treebank Talbanken.
-The project is hence heavily depending on three resources, which will be described
-in this section.
-
-\section{Grammatical Framework}
+
+Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
+into the GF notation and an extended Swedish grammar implementation. The grammar is based
+on the multilingual abstract syntax given in the GF resource library, and now
+also covers constructions specific to Swedish.
+We further give an example of the advantage of using dependent types when
+describing grammar and syntax, in this case for dealing with reflexive pronouns.
+Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
+It is based on Martin-Löf type theory which allows
+reasoning within the programming language.
+
+Our goal is a wide-coverage grammar for Swedish which -- combined with statistical 
+resources -- can be used for parsing open-domain text. The grammar implementation gives
+a deep analysis, in the trade-off between coverage and depth we aim for depth
+while getting coverage from statistical methods.\\
+
+Outline ...
+
+\section{Background}
+
+\subsection{Swedish}
+
+% Chapter 2.4(with emphasis on the syntactic structures)
+\textbf{Too long?}
+Swedish is a North-Germanic language, sharing most of its grammatical structure 
+with Norwegian and Danish. It is also one of the official languages in Finland
+and is spoken by approximately 9 million people.
+Although Swedish syntax often is similar to English, the  morphology is richer and the
+word order more intricate.
+It is a verb-second language \cite[p.116]{gunlog}: the
+second constituent of a declarative main clause must consist of a verb.
+The normal word order is subject-verb-object, but fronting other constituents
+-- topicalisation -- is very common,  especially for temporal and
+locative adverbial phrases \cite[\textsection 1027]{H&H}.
+Fronting the finite verb marks questions.
+
+Swedish has two ways of forming passive verb phrases: the 
+\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli' (\emph{`become'})
+and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
+\enumsentence{Passive \hspace{45mm} Active\\
+\shortexm{12}
+{Tjuven & togs&  av & polisen}
+{\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
+{``The thief was arrested by the police"}
+{&&Polisen & tog & tjuven}
+{&&\emph{the police} & \emph{took} & \emph{the thief}}
+{\hspace{-9mm}``The police arrested the thief"}}
+\label{gfPass:s-pass}
+
+Special reflexive pronouns and reflexive possessive pronouns for the 3rd
+exist \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd person forms.
+\enumsentence{
+\begin{tabular}{llll}
+a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
+&\emph{He hit \sc self.} && \emph{He saw {\sc self's} child.}
+\end{tabular}}
+\enumsentence{
+\begin{tabular}{llll}
+a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
+&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
+\end{tabular}}
+
+\subsection{GF}
 \label{sec:gf}
-The main component of the project is the Grammatical Framework (GF) \cite{gfbok}, %It is
+\textbf{Ramona: this is copied too, probably much room for improvement,
+and it also has to be shortened}
+%Chapter 2.1 + reference to Krasimir's work on the Penn Treebank
+%(+ 2.1.3), nothing about resource library here
+The main component of the project is the Grammatical Framework (GF) \cite{gfbok},
 a grammar formalism based on Martin-Löf type theory \cite{martinlof}. GF is
 designed for multilingual applications and represents a formalism
-%equivalent to Parallel Multiple Context-Free Grammars (Seki \& al. 1991), an instance of
-%generalized context free grammar (GCFG) it is hence
 stronger than mildly context-free grammars.
-The framework's %has hence the same
-expressiveness is hence stronger than 
-Tree Adjoining Grammars \cite{tag} and Head Grammars \cite{hg}, and shown
-equivalent to Parallel Multiple Context-Free Grammar \cite{pmcfg} in \cite{peter}.
 
 GF is a strongly typed functional programming language, inspired by
-%programming languages like 
 ML \\\cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
 and the built-in functionality for logic and reasoning 
-is inspired by \textlambda Prolog \cite{prolog} and 
+is inspired by $\lambda$ Prolog \cite{prolog} and 
 by Agda \cite{agda}, with which GF also shares its type checking algorithm.
@@
 
-%> dependent and implicit types : agda. (read gf book and krasses book)
 A grammar written in GF can be used for both parsing and generation.
@@
 The grammars can be compiled into portable grammar format (PGF) \cite{pgf},
-supported by %There are libraries for using PGF files with
-Java, Java script, C and Haskell libraries. 
-%>> Haskell only complete, also or C\# and C.  %may also be used in Android applications.
+supported by Java, Java script, C and Haskell libraries. 
 The interpreter, the GF shell, allows
@@
 The shell can be tried out online 
-together with an interactive GF editor %and the GF shell 
+together with an interactive GF editor 
 \footnote{http://www.grammaticalframework.org/demos/gfse/}.
@@
 functions and types used for parsing the sentence. 
-\begin{figure}[h]
+\begin{figure}[!h]
 \centering
-\subfloat[Swedish parse tree]{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTree.png}}
-\subfloat[GF abstract tree]{\label{pic:gfAtree}\includegraphics[width=100mm]{gfTree.pdf}}
+\subfloat[Swedish parse tree]
+{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTreec.pdf}}
+\label{fig:gftree1}
+\subfloat[English parse tree]
+{\includegraphics[width=30mm]{gfETreec.pdf}
+\label{pic:gfEtree}}
+\subfloat[GF abstract tree]
+{\label{pic:gfAtree}\includegraphics[width=30mm]{gfTree.pdf}}
 \caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
@@
 \end{figure}
-\vspace{10mm}
-
 Parse trees on the other hand show only the types assigned to the words and phrases.
@@
 not reproduced in the parse tree. Hence, parse trees do not give complete representations
-but model the parse results in a transparent manner. For our example, the definiteness and
-number of the noun `\emph{katten}' is shown as \verb-DetQuant DefArt NumSg- in
-the abstract tree while the parse tree only shows that the noun phrase consists of one noun. 
-% is used but not in the parse tree.
-In the corresponding English parse tree, figure \ref{pic:gfEtree},
-the noun is explicitly quantified by the article \emph{`the'},
-and the determiner, the first argument to the function \verb-DetCN-, is
-therefore shown in the parse tree.\\
-
-% The first command will parse the given
-%sentence and visualize the resulting parse tree. The second will visualize the 
-%abstract syntax tree. The trees will be showed in the pdf-format by the program evince.
-%% nice: test vt -format=pdf -view=evince -nofun
-%The visualization of the trees is done with graphviz\footnote{http://www.graphviz.org/}.
-%Figure \ref{fig:gftree1} shows the results from running the commands in figure \ref{fig:shellvp}
-%%A GF parse tree has two visualized representations, the \textit{parse tree} and
-%%the \textit{abstract tree}.
-%The \textit{abstract tree} shows which functions GF uses
-%to parse a sentence.
-%The \textit{parse tree} (\ref{pic:gfStree}) shows which types
-%that are assigned to the words and phrases. 
-%% do not know about abstract/concrete yet: and is consequently the same for all concrete syntaxes.
-%The reader should keep in mind that the visualized parse trees is not a
-%complete representation. It does not reproduce all information from the abstract tree.
-%%If the end node of a  branch has no corresponding word, it is excluded from
-%%the visualized parse tree. This is the case for the 
-%For example, the definiteness and number of the noun \emph{`katten'}, which is shown as
-%\verb-DetQuant DefArt NumSg- in the abstract tree.
-%%and the information \verb-DetQuant DefArt NumSg- is not shown in figure \ref{pic:gfStree}.
-%In the corresponding English parse tree, figure \ref{pic:gfEtree},
-%the noun is explicitly quantified by the article \emph{`the'},
-%and the determiner is therefore shown in the parse tree.\\
-%>> how parse trees work, not correct complete representation.
-%>> not a constituent structure. compare to lfg, fstructure, cstructure.
-
-%>>the scope is controlled language, but possible ambiguities (discuss this in future
-%>>work, saldo, conclusion). Ambiguities: from multilinguality, or inherited:
-%>>syntactical (man with the telescope), lexical (youSg/Pl, växa).
-%>> have ambiguity section. 
-\newpage
-\begin{wrapfigure}{l}{0.35\textwidth}
-\includegraphics[width=30mm]{gfETree.png}
-\caption{English parse tree}
-\label{pic:gfEtree}
-\end{wrapfigure}
-
-%TODO read again
+but model the parse results in a transparent manner.
+
 The scope of GF grammars has so far been controlled language, a subset of a
@@
 translation. Inherited ambiguities as well as ambiguities arising from multilinguality will
-remain, but can be controlled more easily. %Example of domainspecific word?
+remain, but can be controlled more easily.
 
@@
 translation and GF has successfully been used for this in a number of projects.
-%like %Some examples of projects using GF are 
 WebAlt \cite{webalt} aims to develop language independent
@@
 natural language.
-%>> key, talk, hats, bom (see references in Ramonas thesis).
-%>> use talk here and move key to other place??
 GF has further been used for describing dialogue grammars \cite{talk}, and
@@
 
-\vspace{10mm}
-\subsection{Writing a GF grammar}
-\label{sec:writegf}
-The key feature in GF is the distinction between
-\textit{abstract} and \textit{concrete} syntax. The abstract syntax represents
-the internal structure and models the semantics without concern for language
-specific features such as agreement or word order.
-An abstract grammar can be implemented by a set of concrete grammars, each
-representing one language. As a comparison, the abstract and concrete syntax
-may be thought of as f-structures and c-structures in Lexical Functional
-Grammar \cite{lfg}.
-
-%Being developed with multilinguality in mind, GF makes an important 
-\begin{figure}[h]
-\begin{verbatim}
- abstract TestGrammar = {
-  cat N ; V ; S ;
-
-  fun 
-    Pred : N -> V -> S ;
-    cat_N : N ;
-    sleep_V : V ;
- }
-\end{verbatim}
-\caption{A small example of an abstract syntax}
-\label{fig:gfAbstract1}
-\end{figure}
-\newpage % until behave
-
-\noindent The example in figure \ref{fig:gfAbstract1} shows an abstract grammar defining three categories, % \verb|Categories|,
-one for nouns, one for verbs and one for sentences. The abstract grammar also gives
-the function types. In this case we have \verb|Pred|, which tells us that by
-taking a noun and a verb we can form a sentence.  No information of how this is
-done is given at this stage. The grammar also defines two words, the noun
-\verb|cat_N| and the verb \verb|sleep_V|.
-
-% or how any of the categories should look.
-\begin{figure}[h]
-\begin{verbatim}
-concrete TestGrammarSwe of TestGrammar = {
-  lincat N, V, S = Str ;
-
-  lin Pred n v   = n ++ v ;
-      cat_N   = "katten"  ;
-      sleep_V = "sover"  ;
-}
-\end{verbatim}
-\caption{A Swedish concrete grammar}
-\label{fig:gfSweCnc1}
-\end{figure}
-
-%>> move much of this to Swedish GF part?
-\noindent Figure \ref{fig:gfSweCnc1} shows how the abstract grammar can be implemented
-for Swedish. Nouns, verbs and sentences are all defined as strings, \verb|Str|.
-The function \verb|Pred| simply glues the two strings `katten' and `sover' together:\\
-\verb|Pred cat sleep = "katten sover"|.\\
-We get a more complicated example if we allow the nouns to be used in both
-plural and singular. We add a category \verb|N'| to the
-abstract, which represents a noun with a fixed number,
-%Note that \verb-N'- is normally used for nouns with a fixed definitness, but for 
-%now we concentrate on the number only. Hence \verb|N| now means a noun in any of the number,
-and we introduce two functions for setting the number: \verb|NSg : N -> N'| and
-\verb|NPl : N -> N'|.
-\begin{figure}[h!]
-\begin{verbatim}              
- abstract TestGrammar = {          concrete TestGrammarSwe of TestGrammar = {
-  cat N ; N' ; V ; S ;               lincat V, S, N' = Str ;
-                                            N = {s : Num => Str} ;
-  fun                                lin   
-    Pred : N' -> V -> S ;              Pred n v = n ++ v ;
-    NSg : N -> N' ;                    NPl n = n.s ! Pl ;
-    NPl : N -> N' ;                    NSg n = n.s ! Sg ;
-    cat_N : N ;                        cat_N = {s = table {Sg => "katten" ;
-    sleep_V : V ;                                          Pl => "katterna"}};
- }                                     sleep_V = "sover"  ;
-                                     param Num = Sg | Pl ;
-                                     }
-\end{verbatim}           
-\caption{A modified grammar}
-\label{fig:gfTest2}
-\end{figure}
-Figure \ref{fig:gfTest2} introduces some new concepts: records, tables and parameters.
-In the concrete syntax, \verb|N| is defined to be a record consisting of the field
-\verb|s|. The type of \verb|s|, \verb-Num => Str- shows that it is a table, which given a
-parameter of type \verb|Num| returns a string. \verb|Num| is defined 
-%as a parameter, which can
-to either have value \verb|Sg| or \verb|Pl|. 
-The dot (\verb-.-) is used for projection and the bang (\verb-!-) as a selection operator.
-%In \verb|NPl| and \verb|NSg|,
-\verb|n.s ! Sg| thus means that we use the branch for \verb-Sg- in
-field \verb|s| of \verb|n|.
-%, and then the selection operator \verb|!| is used to select a branch in the table.
-
-When implementing an English version of the grammar, we encounter another
-problem: the verb form depends on the number of the noun. We solve this by letting
-\verb|N'| carry
-information about its number and letting \verb|Pred| pass this on to the
-verb. Finally, the type of \verb|V| is put into a table, showing the verbs forms for each
-number.
-\begin{figure}[h!]
-\begin{verbatim}              
-concrete TestGrammarEng of TestGrammar = {
-  lincat S  = Str ;
-         V  = {s : Num => Str} ;
-         N  = {s : Num => Str} ;
-         N' = {s : Str ; num : Num} ;
-  lin   
-    Pred n v = n.s ++ v.s ! n.num ;
-    NPl n = {s = n.s ! Pl ; num = Pl} ;
-    NSg n = {s = n.s ! Sg ; num = Sg} ;
-    cat_N = {s = table {Sg => "the cat" ;
-                        Pl => "the cats"}};
-    sleep_V = {s = table {Sg => "sleeps" ; 
-                          Pl => "sleep"}};
-  param Num = Sg | Pl ;
-  }
-\end{verbatim}           
-\caption{English implementation}
-\label{fig:gfTestEng}
-\end{figure}
-
-%maybe continue example with swedish and english, have DetNP : N -> N wich
-%gives 'the cat', 'katten'. Thereby introduce tables, parameters, fields.
-%Show abstract tree.
-
-\noindent We now have two implementations of the abstract, one for Swedish and one for English.
-The resulting GF grammar is able both to
-parse a string to an abstract tree and to go in the other direction; to produce
-a string of natural language given an abstract tree. This step is called linearization.
-Translation is a consequence of this, we can parse a Swedish string and then 
-linearize the resulting abstract tree to English.
-
-\subsection{The resource library}
-\label{sec:resources}
-%>>Best part? Biggest library?
-%>>Most general resource, first large grammar, covers fundamental syntactics.
-The GF package provides an useful resource library \cite{gf-resource}, covering the
-fundamental morphology and syntax of more than 20 languages.
-There is also a small test lexicon included, containing a few hundred common
-words.
-%is a very useful part of the framework. It is provided in the 
-%, they give a good start for all who want to write an application grammar.
-%of a language for each application, all
-%this can be found as libraries. So far there are 20 languages in the library, and
-Since the languages all share the same abstract syntax translation between any given
-pair is possible, which is valuable
-%they constitute  a very useful resource 
-when implementing multilingual applications.
-
-The resource grammars describe how to construct phrases and sentences and how to
-decline words. The latter is done by smart paradigms: functions that analyse
-some given forms of a word to find out how the inflection table should look.
-For example, the declination of many Swedish nouns can be determined by looking
-at the singular form only. 
-\begin{wrapfigure}{l}{0.4\textwidth}
-\begin{tabular}{| l | l |}
-\hline
-1st declination & 5th declination \\
-\hline
-flick\textbf{a}    &     hjärt\textbf{a}   \\
-flick\textbf{an}    &    hjärt\textbf{at}  \\
-flick\textbf{or}    &    hjärt\textbf{an}  \\
-flick\textbf{orna}  &    hjärt\textbf{ana} \\
-\hline
-\end{tabular}
-\end{wrapfigure}
-
-\noindent This is the case for `flicka', a noun belonging to the first declination. %the correct declination
-%can be guessed by using the first declination. % and noting that the word ends with an `a'. 
-%Like most Swedish nouns ending with an `a', it follows the first declination.
-For others, like "hjärta", also the plural form "hjärtan" is needed.
-%, even though it ends with an `a'. 
-The worst case for nouns is four needed forms, both singular and
-plural in definite and indefinite form.
-Section \ref{sec:swegf} will give a more thorough description of the Swedish resource grammar.
-
-\newpage
-\subsection{Frontiers of Grammatical Framework}
 As an open source-project, GF is constantly being developed and improved. New
@@
 language via Montague semantics \cite{montague}.
-%Further, experiments on using GF for free parsing has been conducted
-%\cite[\textsection 4.1]{gfMech}.
-%This is also a future goal for our project and one of the subgoal has been to
-%model an important part of Swedish in GF. The starting point was the resource
-%grammar to which the result has a strong connection.
-
-%Functor
-%Main ideas, abstract and concrete, translation, parse- and abstract trees.
-%We say that this is a function that takes as argument and returns..
-%Categories, valency built in (problems with ställa - på, under, vid). Parameters, fields.
-%Example of a piece with Swedish, the reader gets familiar with how it looks.
-%Also show how information 'disappears', become strings.
-%What kind of linguistic analyse, what is the purpose to cover. Mention other uses,
-%ideas of anaphores, montague? dependent types, ontologies, cool pictures by Krasse.
-%
-%The resources, why, how many, how to use, projects. functors.
-%
-%Still under big development, evolving, growing.
-%For this project, only (mostly) important to model Swedish without abandoning the
-%resources too much. Use of the big grammar for translation, why it might be hard with
-%semantics and lexicon.
-%Research about how to parse with gf by Krasimir, experiments with English.
-
-%inte vet jag. foc + v + resten
-
-\section{Talbanken}
+
+\subsection{Talbanken}
+
+%Chapter 2.2
 \label{sec:talbanken}
@@
 compare them against a reliable source.
-Talbanken \cite{talbanken} was perfect for our purpose,
-being a freely available, manually annotated, large-scale treebank.
-%It was created in the 1970s at Lund University.
+We have used Talbanken \cite{talbanken},
+a freely available, manually annotated, large-scale treebank.
 It is analyzed with the MAMBA annotation scheme (Teleman, 1974) and 
@@
 \\
-
 \noindent Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
@@
 Talbanken05 \cite{talbanken05}.
-%This work was done by Nivre, Hall and Nilsson in 2005.
 It is released in Malt\footnote{http://w3.msi.vxu.se/~nivre/research/MaltXML.html} 
@@
 The treebank has served as an inspiration and an evaluation source throughout the
-project. An automatic mapping between its trees and the abstract trees from GF has been
-done, which will be explained in section \ref{sec:Mapping}.
-%Differences in analyse, do we want a similar? 
-%Uses of Talbanken in project, mapping to evaluate and test.  
-%Look at output from Maltparser, could also comparing our parse trees to this.
-
-\section{Saldo}
-\label{sec:saldo}
-A good parser needs a good lexicon. We have used Saldo \cite{saldo}, a
-large electronic lexicon developed and maintained at Gothenburg University. It is
-built on Svenskt Associationslexikon and contains information about more than 
-120 000 modern Swedish words.
-For each word there is semantic, syntactical and a morphological
-information. The user can find examples of usage in corpora,
-graphs of semantically connected words and some suggestions for how to analyse compounds. \\
-\begin{figure}[h!]
-\centering
-\subfloat[Morphological info for `katt']{\label{pic:saldoTab}\includegraphics[width=40mm]{saldotab.png}}
-\hspace{20mm}
-\subfloat[Graph showing the hyponyms of `katt']{\label{pic:saldoWheel}\includegraphics[width=70mm]{saldograph.png}}
-\caption{Information in Saldo}
-\label{fig:saldo}
-\end{figure}
-
-\noindent The semantic aspect of Saldo requires that words with multiple meanings are
-separated into different entries.
-%TODO change word? uppskatta, känna, få
-The word `uppskatta' for example, has two entries, with slightly
-different semantics: \emph{enjoy}
-\enumsentence{Jag uppskattar teater. \\
-``I enjoy theater."}
-\newpage % until behave
-or \emph{estimate}: 
-\enumsentence{Värdet uppskattades till 5 miljoner.\\
-``The value was estimated to 5 million."}
-%\enumsentence{Barnen växte varje dag\\
-%``The children grew for each day"}
-%or \emph{vegetate}. 
-%\enumsentence{Det växte en gammal ek där.\\
-%``There stood an old oak."}
-`uppskatta' is however declined the same way for both interpretations of the word.
-Hence \textit{Saldo's morphological lexicon} only keeps one entry of `uppskatta'.
-For the homonyms that do not share the whole inflection tables, different entries are 
-kept also in the morphological lexicon.\\
-For our purpose, only the morphological section, provided under LPGL in XML
-format, was needed.
-The data can be processed and translated to GF format as described in 
-section \ref{sec:prog.saldo}.
-
-\section{Swedish}
-\label{sec:swedish}
-Swedish \cite[Inl. \textsection 3]{SAG} is a North-Germanic language,
-closely related to Norwegian and Danish. The languages share most of their
-grammatical structures and are mutually intelligible. Swedish is also 
-one of the official languages in Finland and altogether spoken by approximately 9
-million people.
-Swedish syntax is often similar to English, but the  morphology is richer and the
-word order slightly more intricate.
-
-\subsection{Post-nominal articles} 
-\label{sec:swedishnoun}
-Unlike most of the world's languages Swedish express not only the number but also
-the definiteness of a noun by suffixes. The endings are decided by the noun's gender,
-%which may be 
-\textit{neuter} or \textit{non-neuter}.\\
-%(or \textit{utrum}) 
-\begin{tabular}{l|lll}
-                  &\textbf{Indefinite} & \textbf{Definite}& \textbf{Gender} \\
- \hline
-\textbf{Singular} & {en katt}        & {katten}           & \textit{non-neuter}\\
-                  & \emph{a cat}     & \emph{the cat}     &\\
-\textbf{Plural}   & katter           & katterna           &\\
-                  &\emph{cats}       & \emph{the cats}    &\\
- \hline
-\textbf{Singular} & {ett hus}        & {huset}            & \textit{neuter}\\
-                  & \emph{a house}   & \emph{the house}   & \\
-\textbf{Plural}   & hus              & husen              & \\
-                  & \emph{houses}    & \emph{the houses}  & \\
-\end{tabular}\\
-
-%The definiteness can be accentuated by using the definite articles.\\
-%\begin{tabular}{llll}
-%den katten & det huset & de katterna & de husen \\
-%\emph{the cat} & \emph{the house} & \emph{the cats} & \emph{the houses} \\
-%\end{tabular}\\
-
-A definite article or a determiner is necessary when the noun is modified by adjectival
-phrase; sentence (\ref{sent:trottkatt}a) is not grammatically correct.
-As adjectives also marks definiteness, this is marked on three places
-in sentence (\ref{sent:trottkatt}b). %, the definiteness is 
-%The definiteness in then marked by the article, on the noun and on the adjective.
-
-\enumsentence{\begin{tabular}[t]{@{}*{10}{l@{\ }}}   %\begin{tabular}[t]{@{}*{lllllll}}
-a. & *Gamla          & katten        &sov. &&b. &Den             &gamla            & katten         &sov.\\
-   & $[${\sc+def}$]$ & $[${\sc+def}$]$&    &&  & $[${\sc+def}$]$ & $[${\sc+def}$]$ & $[${\sc+def}$]$&  \\
-   &~~``The old       & cat           &slept."&& &``The            & old             & cat            &slept."\\
-\end{tabular}}\label{sent:trottkatt}
-However, for some determiners, ie. `min' (\emph{`my'}), the noun should be in indefinite form.
-%the definitness of the noun is set by the determiner. For some,
-%like `min' , the noun should be used in indefinite form.
-
-\enumsentence{\longexnt{3}{3}
-{min &trötta& katt} %&& Den där & trötta &katten}
-{``my & tired & cat"}
-{ $[${\sc+def}$]$ &$[${\sc+def}$]$& $[${\sc-def}$]$} % . && $[${\sc+def}$]$  & $[${\sc+def}$]$ &$[${\sc+def}$]$}
-{}}
-
-%a.& Katten& sov. & b
-% &$[${\sc+def}$]$ & %can be superdoubled
-%  &`The cat& slept.'%example: den där boken
-
-\subsection{Verb second}
-Swedish is a verb-second language \cite[p.116]{gunlog}: the
-second constituent of a declarative main clause must consist of a verb.
-The normal word order is subject-verb-object, but any syntactic category can be
-fronted \cite[\textsection 1027]{H&H}.
-This is called topicalisation and is very common, especially for temporal and
-locative adverbial phrases.
-The examples \ref{ex:swedish-svo} - \ref{ex:swedish-avso} all have the same propositional
-meaning, but vary in how the content is presented.
-%~~ although the fronted part of sentence (\ref{ex:swedish-ovs}) and (\ref{ex:swedish-avso}) are being
-%emphasized.
-\enumsentence{
-\shortex{4}
-{Du & ser &  inte &mig.}
-{\emph{you} & \emph{see} & \emph{not} &\emph{ me}}
-{``You  don't see me."}} \label{ex:swedish-svo}
-\vspace{-3mm}
-\enumsentence{
-\shortexnt{4}
-{Mig  & ser &  du & inte.}
-{\emph{me} & \emph{see} & \emph{you} &\emph{ not}}}\label{ex:swedish-ovs} 
-\vspace{-3mm}
-\enumsentence{
-\shortexnt{4}
-{Inte & ser& du &mig.}
-{\emph{not}&  \emph{see}& \emph{you}& \emph{me}}} \label{ex:swedish-avso}
-
-%Alla syntaktiska kategorier kan be fronted utom obetonade satsadverbial (ju,väl).
-%\enumsentence{
-%\begin{tabular}{lll}
-%  Johan & gick & sakta på gatan.\\
-%\emph{Johan} & \emph{walked} & \emph{slowely on the street.}\\
-%  Sakta & gick & Johan på gatan.\\
-%\emph{Slowly} & \emph{walked} & \emph{Johan on the street.}\\
-%  På gatan & gick & Johan sakta.\\
-%\emph{On the street} & \emph{walked}  & \emph{Johan slowly.}\\
-%\end{tabular}}
-
-Inverted word order marks questions
-\enumsentence{
-\shortex{4}
-{Såg & du & mig & inte?}
-{\emph{saw}  & \emph{you}&  \emph{me}& \emph{not}}
-{``Didn't you see me?"}}
-%\enumsentence{Gick Johan på gatan?\\
-%        \emph{Did Johan walk on the street?}}
-The word order in subordinate clauses in also slightly modified,
-%>> and stricter
-\enumsentence{Main sentences:\\
-{\shortex{8}{Jag & såg & Anna & men & hon & såg & {\bf inte} & mig }
-{ \emph{I} & \emph{saw} & \emph{Anna} & \emph{but} & \emph{she} & \emph{saw} & {\bf \emph{not}} &\emph{ me}}
-{``I saw Anna but she didn't see me"}}}
-\enumsentence{Subordinate sentence:\\
-\shortex{7}{Jag & förstod & att& Anna& \textbf{inte}& såg& mig.}
-            {\emph{I} &  \emph{understood} & \emph{that}& \emph{Anna}& \textbf{\emph{not}}& \emph{saw}& \emph{me}}
-            {``I understood that Anna didn't see me"}}
-
-\subsection{Passive voice}
-There are two ways of forming passive verb phrases in Swedish: the 
-\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli' (\emph{`become'}).\\
-%>> better with animate subjects
-\enumsentence{
-\shortex{5}
-{Tjuven &blev& tagen& av & polisen}
-{\emph{the} \emph{thief} & \emph{was} & \emph{taken} & \emph{by} & \emph{the police} }
-{``The thief was arrested by the police"}}
-\label{gfPass:peri-pass}
-and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
-\enumsentence{Passive \hspace{45mm} Active\\
-\shortexm{12}
-{Tjuven & togs&  av & polisen}
-{\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
-{``The thief was arrested by the police"}
-{&&Polisen & tog & tjuven}
-{&&\emph{the police} & \emph{took} & \emph{the thief}}
-{\hspace{-9mm}``The police arrested the thief"}}
-\label{gfPass:s-pass}
-The s-passive is more commonly used than periphrastic passive, for both written
-and Swedish \cite[Pass. \textsection 1]{SAG} and dominates especially when the subject is inanimate.
-%cf.
-%\enumsentence{
-%\shortex{5}{Ett & lejon & jagade & honom.}
-%        {A &lion &hunted &him.}
-%        {`A lion was hunting him.'}}
-
-\subsection{Impersonal constructions}
-Constructions with `det är/var' (\emph{`it is/was'}) are very common in Swedish\\
-\cite[\textsection 309d]{H&H}:
-\enumsentence{
-\shortex{6}{Det & var & roligt & att & höra.}
-{\emph{it} &\emph{was} & \emph{nice} &\emph{to} &\emph{hear}.}
-{``I'm glad to hear that."}}
-'Det' %(\emph{it})
-is also used as formal subject in presentational constructions where
-the real subject is put in the position of an object.
-\enumsentence{
-\shortex{6}{Det &står &en &älg &på &fältet.}
-{ \emph{it} & \emph{stands} &\emph{a} & \emph{moose} &\emph{on} &\emph{the field}}
-{``There is a moose in the field."}}
-
-\subsection{Reflexive pronouns}
-\label{swe:refl}
-The Scandinavian language have special reflexive pronouns
-and reflexive possessive pronouns for the 3rd
-person \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd person forms.
-\enumsentence{
-\begin{tabular}{llll}
-a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
-&\emph{He hit him self.} && \emph{He saw his (own) child.}
-\end{tabular}}
-\enumsentence{
-\begin{tabular}{llll}
-a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
-&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
-\end{tabular}}
-The 1st and 2nd persons use the normal personal pronoun in object form as reflexive
-pronouns.
-
-%Mention varandra. Same idea.
-%Basic info about Swedish. V2 lang, inverted word order and subordinate clauses.
-%Passive, reflexive 'sitt', 'det är', 'det sitter en katt där', prepare the reader
-%for what will be written in the grammar part.
-%Adjectives but not verbs congruate with nouns. De är stora, They are big, Jag
-%är stor, I am big
-%Since the s-passive is the most common one in written Swedish \cite{SAG-34-1}, it has
-
-\begin{changemargin}{0cm}{-2cm}
-\section{Related work}
-\label{sec:related}
-%There has been much research about parsing Swedish. INTRO!!
-Many years of research have lead to many interesting language
-technology tools for Swedish.
-An example is the well-known data-driven Malt parser \cite{malt},
+project.
+
+\section{Related Work}
+\textbf{Ramona: also just copied, if you want, improve it}
+%Chapter 2.5 - w/o the references about lexical resources
+%
+%Mention the LREC paper and what it does, without any direct reference
+%between the authors
+There are many other parsers for Swedish,
+like the well-known data-driven Malt parser \cite{malt},
 which has been trained on Talbanken. 
@@
 The Swedish Constraint Grammar \cite{birn}
-give syntactic analyses. %birn: shallow
+give syntactic analyses. 
 Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
 \cite{fdg},
-%(Tapanainen and Järvinen,1997), 
 an extension of the Constraint Grammar
 formalism, and produces a dependency structure focusing on finding the nominal
-arguments. \\
-%for the  finding subject, objects, suubject complements (SC) and
-%to link them to their proper regents (main verbs) : over 90 \% both precision and recall.
-
-%Tag, Tree adjoining grammar, XTAG: English and Korean, Chinese? English:
-%ongoing to get widecoverage grammar
-%under GPL. Translation to Korean.
-
-\noindent The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
+arguments.
+
+The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
 Structure Grammars \cite{hpsg} (HPSG) providing compatibility with tools for
@@
 There is a collection of grammars implemented in this framework, giving broad-coverage
-descriptions of %broad-coverage
-English, Japanese and German. %, Greece, Spanish etc. 
+descriptions of
+English, Japanese and German.
 The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
-Scandinavian, while Norsource (Hellan, 2003) describes Norwegian. A Swedish version
+Scandinavian, while Norsource \cite{norsource} describes Norwegian. A Swedish version
 was based upon this (SweCore, Ahrenberg) covering the morphology and some
@@
 grammar \cite{stymne}, also implemented using the Lingo Matrix,
-which focuses on describing and translating verb frames.\\ % and translating them to English.
-
-%It should be pointed out that grammars implemented using Lingo Matrix
-%do not share any common abstract, like the resources grammars of GF?\\
-
-\noindent The Swedish version of the Core Language Engine (CLE) \cite{gamback} % \cite{cle}
+which focuses on describing and translating verb frames.
+
+The Swedish version of the Core Language Engine (CLE) \cite{gamback}
 gives a full syntactic analysis as well as semantics represented in `Quasi logical form'. A
@@
 language translator \cite{spoken}. Unfortunately, it is no longer available. The coverage of the Swedish
-CLE is also reported to be very limited \cite[p. 134]{nivretrees}.\\
-
-\noindent In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
-for English and Korean, but, to our knowledge, not for Swedish.  \\
-
-\noindent The ParGram \cite{pargram} project aims at making wide coverage grammars using
+
+In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
+for English and Korean, but, to our knowledge, not for Swedish.
+
+The ParGram \cite{pargram} project aims at making wide coverage grammars using
 the Lexical Functional Grammar approach \cite{lfg}.
@@
 
-\end{changemargin}
-%However has been two, about CLE \cite{cle} and what happened to it. Logic and translation
-%to Swedish, S-CLE (Gambäck,1997) full analyse and semantics, translation,
-%but coverage is too limited (nivres papper), The
-%spoken lang. translator \cite{cle2}.
-
-%>> Much more here, important!! comparisons.
-%>> HPSG and LFG are context sensitive
-%Computational ling is active area for Swedish, there are many other Swedish parser,
-%but grammars?
-%Many of the parsers not ruled based, the most known one Malt \cite{malt} ... .
-%Wiren \cite{wiren}?? chartparser, but for
-%swedish. bit more information than just pos: cassSwe \cite{casswe}
-%Other Sw. parsers and grammars, evaluation of them? 
-%Other grammar frameworks.
-%SweCG (birn) annotation, parser Swedish FDG  (voultanien,2001)which is not
-%freely available.
-%finding subject, objects, subject complements (SC) and
-%to link them to their proper regents (main verbs) : over 90 \% both precision and recall.
-%uses a dependency grammar, but "The functional description of adverb phrases
-%and prepositional phrases (e.g. agent, source, goal, benefactive, time) remains
-%to be described in a future version." 
-%HPSG \cite{hpsg} and lingo matrix , English, Japanese, German,
-%Greece, Spanish, Norweigan (NorSource, Hellan 2003),( not famous: SweCore
-%(Ahrenberg) based on NorSource)
-%also Swedish: BiTSE \cite{stymne}, Stymne 2006, Swedish English translation,
-%and one for the common parts of scandinavian (Haugereid etc) in Nordisk sprog.
-%focusing on verb phrames, built on HPSG.
-%S\o rgaard and Haugereid, Scandinavian Grammar Matrix \cite{scandmatrix}
-% related and references work: CLE, Krasimirs book, CassSwe, GF book,SAG, gunlög,
-% the multilingual treebank, Homse&Hinchcliff, Aarne i LILT
-
-\chapter{Importing Saldo}
-\label{sec:prog.saldo}
-The lexicon provided with the GF resources is far too small for open-domain parsing.
-Experiments have been made to use an interactive tool for lexical acquisition,
-%but the dictionary was still much to small. The acquisition tool may still
-but this should be used for complementing rather than creating the lexicon.
-This section describes the process of importing Saldo, which is 
-compatible with GF, and easily translated to GF format.
-As Saldo is continuously updated, the importing process has been designed to be fast
-and stable enough to be redone at any time.
-
-\section{Implementation}
-%>> syntactic information too. Mention that GF does not have this, only lexical or
-%>> semantic. Not 1-1.
-The basic algorithm for importing Saldo was implemented by Angelov (2008)
-and  produces code for a GF lexicon.
-For each word in Saldo, it decides which forms should be used as input
-to the GF smart paradigms. For a verb, this will in most cases mean giving
-the present tense form, see figure \ref{fig:saldoknyt}. \\
-
-\begin{figure}[h]
-\verb-mkV "knyter" ;-
-\caption{First code produced for the verb `knyta' (\emph{`tie'})}
-\label{fig:saldoknyt}
-\end{figure}
-
-All assumed paradigms are printed to a temporary lexicon, 
-which will produce an inflection table for every entry when compiled.
-The tables are compared to the information given
-in Saldo and if the tables are equal the code for the word is saved. If the table
-is erroneous, another try is made
-by giving more forms to the smart paradigm.
-For example \ref{fig:saldoknyt}, the smart paradigm will fail to calculate the
-correct inflection table. In the next try both the present and the past tense
-are given:\\
-
-\begin{figure}[h]
-\verb-mkV "knyter" "knöt" ;-
-\caption{Second output for the verb `knyta'}
-\label{fig:saldoknyt2}
-\end{figure}
-\newpage
-The program is run iteratively until the GF table matches the one given in Saldo,
-or until there are no more ways of using the smart paradigm. The verb 'knyta'
-will need three forms:\\
-
-\begin{figure}[h]
-\verb-mkV "knyter" "knöt" "knutit"-\\
-\caption{Final output for the verb `knyta'}
-\label{fig:saldoknyt3}
-\end{figure}
-
-%Find the paper!
-%The program will try to create a lexicon by finding a number of forms for every
-%word and feeding them to the smart paradigms. As few forms as possible should be
-%used, and the resulting table generated by GF should be equivalent to the one found
-%in Saldo. This process is done iteratively,
-%starting from the easiest case. For each attempt the results are compared and when
-%the correct way of using a word has been found, the code is saved.
-
-Figure \ref{pic:TabVax} shows the information given by Saldo and by GF respectively.\\
-\begin{figure}[h]
-  \begin{center}
-\subfloat[Saldo]{\label{pic:saldoVax}\includegraphics[height=90mm]{saldoVax.png}}
-\hspace{5mm}
-\subfloat[GF]{\label{pic:gfVax}\includegraphics[height=60mm]{gfVax.png}}
-\caption{Inflection table for the verb `växa' (grow)}
-\label{pic:TabVax}
-  \end{center}
-\end{figure}
-The tables does not overlap completely, there are some more forms in Saldo's table (e.g. the 
-compounding form `väx-')
-while the one generated in GF contains some that are not in Saldo (e.g. `vuxits').
-As GF concerns about syntax only, and not semantics, and  as the GF tables are automatically
-generated, they always contain all word forms, although some forms may never be
-used in the natural language.
-Saldo may also contain variants: `växt' and `vuxit' are both supine form.
-As far as possible, the program makes up for this by only comparing the overlapping forms
-and only requiring that GF generates one variant whenever alternatives are given. \\
-%The program continues to generate new lexicons as long as there are more
-%paradigms to choose from and compare the results after each run.
-
-\noindent During this project, the program has been made more robust than the previous
-version. It also prints log files providing
-information about the process of importing each word: which paradigms that have been
-tried, the results from the comparisons of the inflection tables and finally listings
-of the words that could not be imported.
-
-Each entry in Saldo has an identifier, e.g. \verb-äta..vb-, which is used as 
-constant names in the GF lexicon. However, the identifier may need some renaming since
-there are special characters  in the Saldo identifiers
-that should be avoided in GF function names. The importation therefore
-needed some renaming. The Swedish letters \emph{å,ä,ö} are translated into
-\emph{aa,ae,oe} respectively. 
-\begin{verbatim}äta..vb  -> aeta_V \end{verbatim}
-This translation may cause two or more lemmas to share the same
-GF identifier, and to avoid name collision a number is added to the name
-whenever a translation have been done: 
-\begin{tabular}{lll}
-kältisk & $\rightarrow$ & kaeltisk\_1 \\
-kaeltisk & $\rightarrow$ & kaeltisk \\
-entrecôte &$\rightarrow$ & entrecoote\_1 \\ 
-\end{tabular}\\
-
-%Problems with naming convention, special
-%characters (å,ä,ö,apostrop,entrecôte.) needs to be translated , but avoid
-%classhes 'kaeltisk' 'kältisk', gf will crash by this, so we add a 1
-%in the end whenever we have changed a letter.
-%,é
-%Costs very much memory, the saldo file is big, the generated lexicon is big.
-%Saldo is divided to save memory. Errors should just be reported, but if something oförutsett
-%happens and one part fail, the separation ensures that the others wont, also
-%possible to restart.
-
-%Explanation of the algorithm: map categories, try each paradigm, compare the table of forms
-%created. If several forms in saldo, gf should pick on of them. Saldo has forms for
-%compounds, gf does not. A grammar is written, the correct ones saved, the others 
-%retried. Try all paradigms which can be formed from saldo. 
-%In Saldo there is also semantic info, so
-%We need the numbers from saldo if there are to lemmas with the same
-%identifier. växa (bli större), växa (vara en planta) vs. sluta\_V and sluta2\_V
-%(slöt). However gf does not want more than one table
-%if the forms are identical, saldo's morpho neither, but there are five in big saldo. 
-%We are just interested in morphology, so far. Dont want too big lexicon.
-
-%see notes to add more about this, pronouns and the importing itself
-%choices for adjectives vs. verbs for particips
-
-\section{Results}
-\label{sec:saldoRes}
-The resulting dictionary contains more than 100 000 entries, approximately 80 \% 
-of the total size of Saldo.
-There are a number of reasons why some words were not imported,
-the most obvious one is that we do not want all categories from
-Saldo in the GF lexicon. Prepositions, %(\emph{pp})
-numerals, %(\emph{nl}),
-personal pronouns etc.
-are assumed to be present in the resource grammars and should not be added again.
-Saldo contains many pronouns which
-are not analysed the same way in GF (see section \ref{sec:gfnouns}).
-Before adding them to our lexicon, we need to do more analysing to find their correct GF-category.
-%TODO read again
-%They can thus not simply be extracted -- we first need to find their correct GF category.
-Some experiments on finding the category
-%the right GF-category for those
-have been done using Talbanken, see section \ref{sec:gf.quant}.
-
-%and are usually not kept in the standard lexicon, but in modules for the basic
-%structures of the language.
-Categories involving multiple words %(\emph{vbm, nnm})
-are usually handled as idioms and should be given in a separate lexicon. In
-total six types of words were considered for the extraction: \\
-
-\begin{figure}[h]
-\begin{tabular}{|l|lll|}
-\hline
-& Saldo tag & GF category & Example \\
-\hline
- Adverb & \textbf{ab} &\textbf{Adv} & ofta (\emph{often})\\
- Adjective&\textbf{av} &    \textbf{A} & gul (\emph{yellow})\\
- Noun & \textbf{nn} &\textbf{N} & hus (\emph{house})\\
- Verb & \textbf{vb} &\textbf{V} & springa (\emph{run})\\
- Reflexive verbs  &\textbf{vbm}& \textbf{V} & raka sig (\emph{shave})\\
- Verbs with particles &\textbf{vbm}& \textbf{V}  &  piggna till (\emph{perk up})\\
-\hline
-\end{tabular}
-\caption{}
-\end{figure}
-
-%
-%There are a number of differences between and saldo Till skillnad från GF,
-%saldo does not contain forms that aren't used gf generates all forms. A form
-%needed to generate the gf table may not be given in saldo such as singular
-%forms for 'byxor'. Also contains idioms, 'hålla huvudet kallt' and has
-%different grammatical analyse, a lot of pronouns. Example. 
-\noindent Most but not all words of these categories have been imported.
-One reason why the importing phase would fail 
-is that Saldo, unlike GF, only contains the actually used word forms.
-% does not contain information about ~~word forms that are not used~~.
-For technical reasons, the smart paradigm might need forms never used.
-%for this is Deponent verbs
-%are for example given with their (often non-existing) active form, al
-%In some cases those forms are needed by the smart paradigm. 
-Consider for example
-the plural tantum % \cite[44]{SAG} 
-noun \emph{`glasögon'} (\emph{`glasses'}).
-The smart paradigm requires a singular form, and since the program could not
-find this in Saldo, there was no way of adding the lemma to the lexicon. 
-When the program failed to import a noun, this was often the explanation.
-%bbMany nouns fai, \emph{`kläder'} (\emph{`clothes'} etc.
-Words of this type may be added manually, for \emph{`glasögon'} we could use
-the ostensibly correct singular form `glasöga', although this
-has another meaning (`glass-eye').
-The same problem occurred for the irregular s-verbs,
-(\emph{`synas'} \emph{(`show')} or \emph{umgås} \emph{(`socialize')})
-which made up 61.5 \% of the failing verbs of type \verb_vb_.\\
-%`frodas' (`flourish'), `lyckas' (`succeeed').
-In a few cases the smart paradigms could not generate the correct declination.\\
-
-%although this happened seldom. The passive preteritum is an exapmle of a form 
-%oftenly because the one form could not be generated. anbd for passive preteritum could not be guessed.
-%that seems hard to guess for some verb.
-% >> should we add smth for this?
-%\begin{tabular}{3}
-%&\textbf{Saldo}& \textbf{GF}
-%\textbf{Passive preteritum} & fryses & frys \\
-%\end{tabular}
-%>> exapmles: % frysa,fisa passive preterium
-
-%compared to the total size of Saldo this is .
-\noindent When testing the coverage of Talbanken,
-we found that there are around 2500 word forms still missing, excluding the ones
-tagged as names and numbers. This number may seem very high, but %considering that
-4/5 of the word forms are compounds and when performing the intended parsing,
-an additional analysis identifying compounds should be preformed before
-looking-up the words in the lexicon. We
-should also take into consideration that 
-we cannot automatically find out how many actually stem from the same word, or
-how many abbreviations that are present. Talbanken also contains a small number of 
-spelling errors, which probably are enumerated among our missing words. The majority
-of the missing words are only used once.\\
-
-\begin{figure}[h]
-\begin{tabular}{|l|l|}
-\hline
-Missing words & $\sim$ 2500 word-forms\\
-Missing words, ignoring compounds & $\sim$ 500 word-forms\\
-Missing words used more than once & $\sim$ 500 word forms\\
-Missing words used more than once, ignoring compounds & $\sim$ 150 word-forms\\
-\hline
-\end{tabular}
-\caption{}
-\end{figure}
-
-%       The table suggests that many 
-%       >> 2537 without IDs and only the ones built of Alpha,only 552 without compounds, still contains abbrev.
-%       >> 494 without abbr. spelling errors
-%       >> 142 more then one time.
-%       numerals and proper names, but some names are still there). In addition, there were 1817 words
-%       >> 223 were IDS.++146 HS.
-%       >> 426 more than one time, with compounds.
-%       >> give reasons! should look good, not bad!
-\noindent A list of words that were given different labels in GF than in Talbanken has been
-composed, consisting of about 1600 entries. Many of those are
-acceptable and reflects the difference 
-made in the analyses, such as the examples in table \ref{tab:saldodiff}.
-Others are examples of words that are still missing from
-the lexicon.
-\begin{figure}[h]
-\begin{tabular}{|lll|}
-\hline
-Word    & Talbanken tag & GF category \\
-\hline
-måste   & MVPS          & VV \\
-allting & POTP          & N \\
-få      & POZP          & Det \\
-\hline
-\end{tabular}
-\caption{}
-\label{tab:saldodiff}
-\end{figure}
-
-%Since Saldo does not give any valency information, neither does the imported lexicon.
-%The lexicon does not contain any information.
-\noindent Valency information, which is crucial for GF, is not given in Saldo and
-hence not in the imported lexicon. It remains as future work to find methods 
-%using the results from the mapping (see section \ref{sec:Mapping})
-to extract this information from Talbanken and to automatically build
-it into the lexicon. 
-%The saddest thing: no valency info. Crucial for GF. Only reflexive and verbs with particles
-%have any information. The earlier explained techniques may be used for this in Talbanken.
-%To get even bigger material, use Korp, but then no guarantee.
-
-\chapter{The grammar}
-\label{sec:prog.grammar}
-An important part of this project has been to develop the Swedish GF grammar and
-to adapt it to cover
-constructions used in Talbanken. As a grammar implementation can never be
-expected to give
-full coverage of a language, we aim for a grammar fragment which gives a deep
-analysis of the most important Swedish constructions.
-The starting point has been the GF resource grammar and the new implementation is
-still compatible with this.
-Before describing the actual implementation in section \ref{sec:Added}, we will give an introduction to the
-resource grammars in general and to the Swedish implementation in particular.
-%This is found in section \ref{sec:swegf} while section \ref{sec:Added} shows and
-%explains the development during this project.
-
-%We can't expect to get a grammar giving full coverage, instead we aim for a
-%grammar that gives
-%a deep analysis while having a satisfactory coverage.
-%The starting point has been the Swedish
-%resource grammar and the new grammar should still be compatible with this.
-%%A description of the Swedish implementation of the resource grammar is given
-%%in the next section, while
-%Before describing what has been added, will give a short introductions to the
-%the cats of resource grammar and the sw implementation in particular.
-%This is done in section%Explain what's in the resources, constructions present in most languages. 
-%Been working on a grammar fragment based on Talbanken,
-%like cool people like Montague. May also be interesting for other purposes.
-%Swedish shares with Nor and Dan. Extra module, covering things like sw is a
-%v2-lang, focusing parts of sentences.
-
-\section{The Swedish resource grammar}
-\label{sec:swegf}
-%>> start by easy examples, jag ser inte dig, du ser inte mig osv. 
-%>> plocka från svenska delen 1 2 4 
-%>> show simple trees
-%>> explain all in Sw. section.
-%>> show abstract and concrete tree, compare. (or in Gf sect)
-%>> no subjects or object, analyse is more from the lexicon,
-%>> compare to others, ask elisabet.\\
-%>> gf has pos-driven, leads to part-of-sentence analyse. first have v2,a2,vs etc.
-%>> they may be combined in traditional way, and so the types depend and relate to each other
-%>> see slides from summer school.
-%>> worked well until now, but particular langs. needs particular solutions. --> reflexives!
-%>> no syntactic theory can cover a whole language, and so gf can't either.
-
-% TODO!
-%>> functions for predication, complementation.
-%>> rather different from others syntactical theories. does not remember syntactical
-%>> parts of sentences, like subj, obj, attributes.
-%>> shares 80 \% with scandinavian, mostly words that differ.\\
-%>> complete morphology, limited syntax\\
-
-The GF resource grammars gives a fundamental description of Swedish,
-covering the morphology,
+\section{Grammar}
+
+%Discussion about the two dimensions of a grammar - lexicon and syntax 
+There are to aspects of on a computational grammar clearly distinctive (?)  in
+our work- the work on the lexicon and the work regarding syntax.
+For this article, we will focus on the syntactical dimension, the description
+and formalization of grammar rules and the usage of Talbanken.
+
+\subsection{Resource grammar}
+
+%Chapter 2.1.2
+The GF package provides an useful resource library \cite{gf-resource}, covering the
+fundamental morphology and syntax of more than 20 languages.
+There is also a small test lexicon included, containing a few hundred common
+words.
+Since the languages all share the same abstract syntax translation between any given
+pair is possible, which is valuable
+when implementing multilingual applications.
+
+The resource grammars describe how to construct phrases and sentences and how to
+decline words. The grammars cover the morphology,
 word order, agreement, tense, basic conjunction etc.
- %and many phrase and clause constructions.
 Due to the syntactic similarities of the Scandinavian languages, much of the implementation
@@
 parts of sentences.
-Figure \ref{fig:gfpic} shows the different categories, or types, used in the
-resource grammars. 
-%A lexicon normally consists of 
-Words from the open word-classes are shown in rectangular boxes in
-the picture. Each lexical entry is assigned a type describing its word-class.
-%and the abstract grammar can be viewed as a set rules showing
-%how the parts of speech can be used --
-%how the types relate to each other.
-\begin{figure}[h]
-\hspace{-18mm}
-\includegraphics[height=130mm]{categories.png}
-\caption{The types of the GF resource grammars. }
-%TODO have footnote? where?
-%\footnote{Available at http://www.grammaticalframework.org/lib/doc/synopsis.html}
-\label{fig:gfpic}
-\end{figure}
-
-The concrete grammar gives a \textit{linearization type} for each category,
-usually a record containing a table with the forms the word or phrase can be used in.
-They may also
-carry information about inherent features that matter to other parts of the sentence.
-
-\newpage
-\subsection{Noun phrases}
-\label{sec:gfnouns}
-Section \ref{sec:writegf} contained an easy example with nouns. We used the categories \verb-N-
-and \verb-N'- to distinguish nouns with respectively without number information.
-%TODO read again
-The category \verb-N'- was a simplification and is not used in the resource
-grammar. 
-The full analysis includes a distinction between the categories \verb-N- -- simple nouns --
-and 
-\verb-CN- -- common noun phrases -- and \verb-NP- -- noun phrases.
-The category \verb-N- is implemented as an inflection table, generated by the smart paradigm,
-and a field keeping information about the gender, see figure \ref{fig:gfflicka}.
-\begin{figure}[h]
-\begin{verbatim}
-flicka_N = {s = {Sg Indef Nom =>  flicka
-                 Sg Indef Gen =>  flickas
-                 Sg Def Nom   => flickan
-                 Sg Def Gen   => flickans
-                 Pl Indef Nom => flickor
-                 Pl Indef Gen => flickors
-                 Pl Def Nom   => flickorna
-                 Pl Def Gen   => flickornas ;
-            g = utr }
-\end{verbatim}
-\caption{Representation of the noun `flicka' (\emph{`girl'}) in GF}
-\label{fig:gfflicka}
-\end{figure}\\
-Nouns may turned into common noun phrases \verb-CN-, %this information is kept, and the noun may
-which may be modified by adjectival phrases or conjoined.
-\begin{verbatim}
-AdjCN : AP -> CN -> CN ; 
-\end{verbatim}
-The function \verb-DetCN-, determination, creates a
-\verb-NP- by setting the number and definiteness of a \verb-CN-, shown in figure
-\ref{gfcode:DetCN}.
-%may be turned into a \verb-NP- by the rule \verb-DetNP- which sets
-%the number and definitness.
-%Since all morphological information is generated by the smart paradigms and contained in the \verb-N-
-%the construction of a \verb-NP- more or less consists of putting the pieces together.
-%The definitness is given by a determiner.
-
-\begin{figure}[h]
-\begin{verbatim}
-DetCN : Det                  -> CN     -> NP ;
-        indefinite, singular -> flicka -> "en flicka"
-        definitive, singular -> flicka -> "flickan"
-\end{verbatim}        
-\caption{}
-\label{gfcode:DetCN}
-\end{figure}
-\newpage
-%The two basic articles are \verb-IndefArt- and \verb-DefArt- which
-%uses the post-modified form of the noun: katten.
-\noindent Some rules for determination were described in section \ref{sec:swedishnoun};
-if we have a common noun phrase consisting of the parts \emph{`liten'}
-(\emph{`small'}) and \emph{`katt'} (\emph{`cat'}), there are three ways they
-can be combined, as shown in figure \ref{tab:detdef}. The noun may be used in
-definite or indefinite form and the adjective in its weak or strong form \cite[p. 31]{gunlog}.
-\begin{figure}[h]
-\begin{tabular}{|lll|l|}
-\hline
-\textbf{Determiner} & \textbf{Adjective} & \textbf{Noun} & \textbf{DetSpecies}\\
-\hline
-en  &  liten $[${\sc -Def}$]$& katt   $[${\sc -Det}$]$& \textbf{ DIndef}\\
-min &  lilla $[${\sc +Def}$]$& katt   $[${\sc -Det}$]$  & \textbf{ DDef Indef} \\
-den &  lilla $[${\sc +Def}$]$& katten $[${\sc +Det}$]$& \textbf{ DDef Def} \\
-\hline
-\end{tabular}        
-\caption{}\label{tab:detdef}
-\end{figure}
-
-%Which combination that shoud be used is decided by the determiner. Hence all
-Hence, all determiners
-in our grammar must keep information about which definiteness they require; the
-\verb-DetSpecies- parameter is stored as an inherent 
-feature of the determiner.
-The resource grammar distinguishes between quantifiers (\verb-Quant-), 
-determiners (\verb-Det-) and predeterminers (\verb-Predet-). Predeterminers modify \verb-NP-s
-while the other to modifies \verb-CN-s. The differences are further shown in table
-\ref{tab:detquant}. 
-\begin{figure}[h]
-\begin{tabular}{l|lll}
-                        &\textbf{Has number}& \textbf{Has definiteness}& Example  \\ 
-\hline
-\textbf{Predeterminers} & --  & --  & alla katter, all maten \\ 
-\textbf{Quantifiers}&-- &\checkmark  & min katt, mina katter, *min katten \\ 
-\textbf{Determiners}   &\checkmark &\checkmark & varje katt, *varje katten, *varje katter  \\ 
-\end{tabular}
-\caption{}\label{tab:detquant}
-\end{figure}
-
-The definite article is considered to be a quantifier, which has the forms
-\emph{en} and \emph{ett} for singular. In plural it
-is the either \emph{`de'} or nothing, cf. sentence (\ref{sent:trottkatt2}a) and
-\ref{sent:trottkatt2}b.
-\enumsentence{\begin{tabular}[t]{@{}*{9}{l@{\ }}}   %\begin{tabular}[t]{@{}*{lllllll}}
-a. & Katten        &sov. &&b. &\textbf{Den}  &gamla & katten  &sov.\\
-   &\emph{cat}\small\sc+def &\emph{slept}&& &\emph{the} & \emph{old} & \emph{cat}\small\sc+def &\emph{slept}\\
-\end{tabular}}\label{sent:trottkatt2}
-
-In order to get the correct plural form
-we need to know if the common noun phrase 
-has been modified, that is, whether the function \verb_AdjCN_ has been used.
-However, once a category has been formed, there is no longer any information available
-about how it was put together. This is a result of the functional approach
-of GF. Therefore, %that is, whether the function \verb|AdjCN| has been applied.
-this information
-has to be passed on by an inherent feature of the \verb-CN-, a flag set to tell
-if the \verb-AdjCN- was applied.\\
-\begin{figure}[h]
-\begin{verbatim}
-DetNP : Det              -> CN        -> NP ;
-        definite, plural +  katt      =  katter
-        definite, plural +  stor katt =  de stora katterna
-\end{verbatim}        
-\caption{}\label{gfcode:defmod}
-\end{figure}
-
-%Their record
-%has a field storing the
-%which may  are called \verb-DIndef-, \verb-DDef Indef- and \verb-DDef Def-. 
-%Additionally, we need to make sure that sentences as 
-%\ref{sent:trottkatt}a are not allowed.
-
-%TODO review, verify
-\noindent Due to the syntax oriented analysis in GF, the GF category for pronouns \verb-PN- only
-contains personal pronouns.
-Many words, like `somliga' (\emph{`some'}), are considered to be pronouns
-in other analyses, such as The Swedish
-Academy Grammar \cite{SAG},  but are classified differently in GF, usually
-as determiners or quantifiers as they may determine noun phrases
-(\ref{sent:somestud}).
-\enumsentence{Somliga studenter jobbar bara på nätterna\\
-``Some student only work at night time"}\label{sent:somestud}
-%Only personal pronouns are of the type \verb-Pron-.
- %TODO
- % Adjectives? About weak - strong - indefinite..
-
-%TODO
-%>> compare NP to HPSG.
-%>> CN may be different in other theories.
-
-\subsection{Verb phrases}
-In figure \ref{fig:gftree1} we saw the GF representation of the sentence ``Jag
-ser katten". The verb `ser' (\emph{`see'}) takes one object: `katten'.
-In GF, this can be seen on the type of the verb,
-valency information
-is encoded into the lexicon. Transitive verbs have the type \verb-V2- and ditransitive
-\verb-V3-. There are types for verb which takes sentences (\verb-VS-) and
-adjectival (\verb-VA-) complements.
-The lexicon is therefore a very important part of the
-grammar. 
-\begin{verbatim}
-see_V2 ; 
-SlashV2a   : V2 -> VPSlash ;
-ComplSlash : VPSlash -> NP -> VP ;
-\end{verbatim}
-
-The function \verb-SlashV2a- lets you create a \verb-VPSlash-. The name
-\verb-SlashV2a- is inspired by categorical grammar and is a short
-hand for \verb-VP \ NP- -- a verb phrase missing a
-noun phrase..  When we combine the \verb-SlashVP- with the object we
-get a complete verb phrase.
-Both the verb and its complements are stored in the verb phrase category, \verb-VP-.
-%A simple idea would let the \verb-VP- consist of one field where the verb and object are
-%glued together. But we also want to be able to express sentencens like
-%``Du ser inte mig" and ``Mig ser du inte", where the object is separated from the verb.
-%TODO explain, diff from Gazdar
-The \verb-VP- category resembles the Didrichsen's field model.
-%it has multiple fields that may be put in the different orders.
-There are fields
-for negation, different types of adverbs, objects, the finite verb and an infinite verb.
-The fields may in turn be tables showing the different forms of the component.\\
-
-\begin{tabular}{|l|llllllll|}
-\hline
-& & &&\textbf{VP}& & & & \\
-\textbf{VP field} && \sc finit& \sc neg& \sc adV & \sc inf& \sc comp& \sc obj& \sc adv \\
-\hline
-&\emph{(han)} & har &  inte&  alltid&  tänkt&  på&  henne&  så \\
-\hline
-\end{tabular}\\
-
-\vspace{5mm}
-The verb phrase fields are not put in their correct order until the tense and type
-of clause is determined, ie. when a sentence is created.\\
-
-\noindent Swedish verbs may take up to five arguments \cite[p. 53]{stymne}.
-These may be prepositions, particles, reflexive object, indirect objects
-and direct objects. 
-\enumsentence{\shortex{7}
-{Jag &tar& med &mig& den& till& honom}
-{\emph{I} &\emph{take}&\emph{with} &\emph{me}& \emph{it}& \emph{to}& \emph{him}}
-{``I bring it to him"}
-\label{sent:tamed}}
-The verb `ta' in sentence (\ref{sent:tamed}) takes one particle (\emph{`med'}),
-one preposition (\emph{`till'}) and two objects: \emph{`den'} and \emph{`honom'}.
-In a GF lexicon this verb is given the
-category \verb-V3-, %classifing it as 
-a \textit{three-place verb}, taking two objects.
-The notion \verb-V3- is motivated by the formal translation: \verb-bring(I,it,him)-.\\
-%these will be encoded in different ways.
-Particles are given in the lexicon, as well as the prepositions that are
-chosen by the verb, since these may not vary.
-The entry for `\emph{ta med}', as used in sentence (\ref{sent:tamed}), is
-described as follows:
-%, assuming that we already have an constant \verb-take_V- which produces the correct conjugation for `ta':
-\begin{verbatim}
-ta_med_V3 = dirV3 (reflV (partV (take_V "med"))) (mkPrep "till") ;
-\end{verbatim}
-%The category \verb-V3- tells us that the verb takes two objects, % (`den' and `honom').
-The function \verb-dirV3- creates a three-place verb, where the first object is direct and
-the second is to be used with the preposition given as the last argument: `\emph{till}'.
-\verb-reflV- shows that the verb always is used with a reflexive
-pronoun and \verb-partV- gives the particle \emph{`med'}.
-The fact that the chosen prepositions is attached to the verb in the lexicon
-causes the parse tree visualization algorithm to group them together.
-This is also the case for particles, cf. parse tree \ref{pic:tittapart}
-and \ref{pic:tittapa}.
-\begin{figure}[h]
-\subfloat[Verb with a chosen preposition]{\label{pic:tittapa}\includegraphics[height=70mm]{tittapa.jpg}}
-\hspace{10mm}
-\subfloat[Verb with particle]{\label{pic:tittapart}\includegraphics[height=76mm]{tittapart.pdf}}
-\caption{The visualized parse trees do not show the internal difference of chosen prepositions and particles}
-\label{fig:translationtrees}
-\end{figure}
-\newpage
-\noindent As already stated, the visualized parse trees is not a complete representation,
-even if the verb phrases in the visualizations look the same, the two cases
-are treated and represented differently internally. The fronting of the
-preposition, as in of sentence (\ref{sent:tittapa}a)., is accepted but
-fronting of particles, as in \ref{sent:tittapa}b., is not.
-%TODO use png to get correct size
-\eenumsentence{
-\item[a]\shortex{4}
-{På &pojken& tittar& du.}
-{\emph{on} &\emph{the boy}& \emph{look}& \emph{you}}
-{``You look at the boy."}
-\item[b]\shortexnt{6}
-{*På& när& du& springer& tittar& jag}
-{\emph{~on}& \emph{when}& \emph{you}& \emph{run}& \emph{watch}& \emph{I}}
-\label{sent:tittapa}}
-
-%Categories as V,V2,VS...
-%Maybe talk about ställa.
-\subsection{Clauses}
-The category clause, \verb-Cl-, represents a pre-sentence, that does not yet have any tense, polarity
-or word order set, see figure \ref{fig:clause}.\\
-
-\begin{figure}[h]
-\begin{tabular}{|lll|ll|}
-\hline
-\textbf{Tense} &\textbf{Polarity} &\textbf{Word order} &\textbf{Cl} &\textbf{S} \\
-\hline
-present &negative & main        & du ser mig & 'Du ser inte mig'\\
-perfect &positive & inverted    & du ser mig & 'Har du sett mig'\\
-perfect &negative & subordinate & du ser mig & 'Du har sett mig'\\
-\hline
-\end{tabular}
-\caption{}
-\label{fig:clause}
-\end{figure}
-
-\noindent Like verb phrases, a clause may also be missing an object and then has the type
-\verb-ClSlash-.
-The \verb-ClSlash- is formed by a \verb-VPSlash- which is given a subject.
-This is a convenient way to form questions, relative clauses
-and topicalized clauses (see figure \ref{fig:clause2}), as introduced in \cite{gazdar}.\\
-
-\begin{figure}[h]
-\begin{tabular}{|l|ll|}
-\hline
-\textbf{Cl}& Johan + tittar på \\
-\hline
-\textbf{Wh-Questions} & Interrogative pronoun   & `Vad tittar Johan på?'\\
-\textbf{Relative clauses} & Object + Relative pronoun & `Katten som Johan tittar på' \\
-\textbf{Topicalized clauses} & Object  & `Henne tittar Johan på' \\
-\hline
-\end{tabular}
-\caption{}
-\label{fig:clause2}
-\end{figure}
-
-%
-%Verbs with particles or verbs that choose a special  
-%Particles and prepositions, can't be seen on the trees. Can now see why the parse
-%trees are not complete. difference shows when fronting.
-%tycka om /tycka om -> om vad tycker du inte. fast bra
-%Swedish verb may have up to five arguments according to Stymne
-%'jag tar med mig den till Bo' : 
-%but is really 'till' part of verb?
-%VP is divided corresponding to satsschema (diderichsen) (renewed by SAG).
-%If the object is not given immediatly, the grammar uses slash.
-%"Speed up", instead of movement used slashes \cite{gazdar}.
-%Slashes good for question (vilken katt ser du idag)
-%relatives (katten som du tittar på) and topicalisation
-%The grammar has VP,Cl, SSlash, all missing NP.
-%
-
-%In some special cases, phrase cannot be put in all forms. 
-%Give a good example!!
-%It is not quite clear how to handle this at the moment. When
-%the grammar is meant for parsing only, it is feasible to put
-%a \textit{nonexist} pattern.
-%example with this.\\
-
-\subsection{Overview}
-\vspace{4mm}
-The original resource grammar could express complex sentences
-as the one in figure \ref{gfSwe:parsetree}.
-%shows the parse tree for sentence 
-%\ref{gfSwe:parseable}.\\
-
-%\enumsentence{\shortexnt{8}
-%{Har& han& inte& ätit& de & gula& äpplena& idag?}
-%{Has& he& not& eaten& the& yellow& apples& today?}} 
-%\label{gfSwe:parseable}
-%Covers sentences like
-% better sentence where the verb phrase is split up?
-\begin{figure}[h]
-\includegraphics[width=70mm,height=70mm]{apples1.pdf}
-\caption{Parse tree for ``Har han inte ätit de gula äpplena idag?"}
-\label{gfSwe:parsetree}
-\end{figure}
-\noindent Even though the verb phrase \emph{``har inte ätit de gula äpplena idag"} is discontinuous, the whole
-phrase is still treated as one constituent in GF. The parts
-are connected in the tree, and the subject \emph{`han'} is put between the
-finite verb and the rest of the phrase.
-At the code level, this is done using the record type for \verb-VP-, which
-consists of fields that can be put %the \verb-VP- fields 
-in different order.\\
-\verb|table { Inv => verb.fin ++ subj ++ verb.neg ++ verb.inf ++ verb.compl ; ... | \\
-%At the code level, this is implemented %using records.
-%The linearization rule for questions -- clauses with inverted order -- picks 
-%the field for the finite verb and appends it to the subject followed by
-%the negation, infinite part and the complement of the verb phrase: \\
-
-The resource grammar also covered relative clauses: 
-\eenumsentence{\item[a]\shortexnt{5}{Hon & ser & pojken & som  &sover}
-              {\emph{she} & \emph{sees} & \emph{the boy} & \emph{that} & \emph{sleeps}}
-              %{`She sees the sleeping boy'}}
-\item[b]\shortexnt{6}{Han & ser &katten & han  &tycker & om}
-              {\emph{he} & \emph{sees} & \emph{the cat} & \emph{he} & \emph{likes} &}}
-              %{`She sees the sleeping boy'}}
-
 In addition to the core resource grammars, which is shared with the other languages implemented in the
 library, there is also an extra module, simple called the \verb|Extra| module, where
-%In addition to the basic resource grammar,
  language specific constructions may bee added
-%>> core is shared but particular lang. has particular features.
-%to the module \verb|Extra|.
 The functions given here do not have to be translatable
 to all other language, but are meant to cover language specific constructions.
-Among those were functions for topicalisation (\ref{gfSwe:apple}), %$ although not all
-%of them were implemented for Swedish.
-%word order, agreement, tempus, adverbs, adjectivs...
-%and relatives
-%Also info in the Extra module, where more language specific constructions
-%can be found.
-%For example, \verb|FocObj| fronted the object as in sentence (\ref{gfSwe:apple}).
-\enumsentence{\shortex{6}
-{Det där & äpplet & vill & jag & inte & ha}
-{that & apple & want & I & not &have}
-{``I don't want that apple"}}\label{gfSwe:apple}
-and for preposition stranding:
-\enumsentence{
-~~Stranded preposition \hspace{45mm} cf.\\
-\begin{tabular}{ll}
-{\begin{tabular}{llllll}
-Vem& måste& jag& akta& mig &\textbf{för}? \\
-\emph{who}& \emph{must}& \emph{I}& \emph{watch out}& \emph{me} &\textbf{\emph{for}}? \\
-\end{tabular}}&
-{\begin{tabular}{llllll}
-\textbf{För}& vem& måste& jag& akta& mig?\\
-\emph{\textbf}{for}& \emph{who}& \emph{must}& \emph{I}& \emph{watch out}& \emph{me}?\\
-\end{tabular}}\\
-~~`Who do I need to watch out \textbf{for}?' &~ `\textbf{For} whom do I need to watch out?'\\
-\end{tabular}}
-
-%jag ser dig
-%dig ser jag
-%dig har jag inte sett
-%jag har inte sett dig.
-%how the parse tree is generated, the finite verb, the subject, the rest. usw.
-
-%honom ville hon inte tänka på.
-
-%What was in the resources
-%description of the existing grammar, the types, the possibilities
-%VP, NP.
-%valency built in (problems with ställa - på, under, vid).
-%Variants, nonexist.
-%coverage of standard Swedish.
-\newpage % until behave
-\section{Development of the grammar}
-\label{sec:Added}
-%The resource grammar is meant to cover the most important grammatical structures
-%present in many languages, but not the language specific constructions.
+
+\subsection{Lexicon}
+
+%Everything about Saldo, Lexin, etc, compressed.
+%
+%We use the resources from the other paper, so to say :)
+\textbf{Some more here}
+Our main lexical resource is SALDO \cite{}. We use the tools
+described in \cite{} to extract a large-scale lexicon and annotate
+this with valencies from Lexin. This gives us a dictionary with more than 
+100 000 entries, covering all but 500 words from Talbanken ignoring names,
+numbers and compounds.
+
+\subsection{New Features}
+
+%Section 4.2
+%
+%Selection of additions to the grammar 
+%
+%- reflexive possesive pronouns
+%
+%- s-passive
+%
+%- impersonal constructions
+%
+%- miscellaneous 
 It has earlier been hard to identify missing constructions of the Swedish
-implementation, since there was no large resource available to evaluate it on.
-
-Our evaluations are based on Talbanken, and when first conducting tests,
- we found much room for improvement.
-From the topics listed in section \ref{sec:swedish}, the post-nominal articles
-and the verb-second property
-were covered by the resource grammar, as well as the periphrastic passive and
-a limited form of topicalisation. 
-The other constructions have been added during this project.
-%enough to be covered by the resources. The others have been added.
-%We have increased this by adding language specific
-%constructions. Talbanken has been our inspiration.
-%>> the grammar will not be totally covering, but interesting and useful.
-%>> hard to estimate before, since no big use case/resource.
-%>> after initial experiments, first evaluation. found much room for improvement. 
-%>> discovered constructions missing.
-
-%%PASSIV
-\subsection{The s-passive}
+implementation since there was no large resource available to evaluate it on.
+
+Our evaluations are based on Talbanken and when first conducting tests
+ we found much room for improvement. The items listed below are examples
+ of constructions implemented during this work.
+
+\textbf{shorten these sections}
+\subsubsection{The s-passive}
 Passive voice is often used in Swedish, especially the
-\textit{s-passive}.
-\enumsentence{
-\shortex{5}{Uppsatsen & skrev\textbf{s} & av & en & student.}
-        {\emph{the essay} & \emph{wrote+\textbf{s}} & \emph{by} & \emph{a} & \emph{student}}
-        {``The essay was written by a student."}}\label{sent:skrevs}
-%{b. &En & student & skrev & uppsatsen.}
-%        {&a &student &wrote &the essay}
-%        {~~`A student wrote the essay.'}}
+\textit{s-passive} as described in ....
 Some studies suggest that the s-passive is used in more than 80 \% of the times
@@
 It is however not as common in the other Scandinavian languages, 
-where not all words have passive forms for all tenses. The Norwegian 
-translation of sentence (\ref{sent:skrevs}) is:
-%The periphrastic passive is preferred, especially in spoken langauge.
-\enumsentence{\shortexnt{7} %\begin{tabular}{ll}
-{Oppgaven &ble& skrevet& av& en& student & [NO]}
-{uppsatsen& blev& skriven& av& en & student& [SE]}}
-%\end{tabular}}
-The corresponding Swedish sentence is acceptable, but not as natural sounding as sentence
-(\ref{sent:skrevs}).
-The resource grammar for Scandinavian therefore implemented the function for passive,
-\verb-PassV2-, by using auxiliary verb. \\
-\begin{verbatim}
-PassV2 : V2  -> VP ;
-         ta -> blev tagen
-\end{verbatim}
-%>> how are they used with reflexives??
+where not all words have passive forms for all tenses. The resource grammar for
+Scandinavian therefore implemented the function for passive,
+\verb-PassV2-, by using auxiliary verb.
 The function allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
-turned into complete verb phrases; they no longer need an object.\\
-
-\noindent During this project, the s-passive was added although the
+turned into complete verb phrases; they no longer need an object.
+During this project, the s-passive was added although the
 periphrastic passive is still allowed.
-The grammar further allows not only V2, but all verb phrases that misses an object, to 
-form passives:
-\begin{verbatim}
-PassVP : VPSlash -> VP ;
-         ta      -> togs
-         erbjöd  -> erbjöds
-\end{verbatim}
-A \verb-V3- like `give' in sentence (\ref{sent:give2pass}) hence gives rise to two
+The grammar further allows not only transitive verbs, but all verb phrases that
+misses an object, to form passives.
+A ditransitive verb -- `give' (\ref{sent:give2pass}) -- hence gives rise to two
 passives, (\ref{ex:passV32}) and (\ref{ex:passV33}).
@@
 \label{ex:passV32}
-\enumsentence{\textbf{V2A verb}\\\shortex{3}
-{Huset & målades & rött}
-{\emph{the house} & \emph{painted+\textbf{s}}& \emph{red}}
-{``The house was painted red"}}
-\label{ex:passV2A}
-
-\subsection{Impersonal constructions}%Presentering, , Topicalisering
+
+\subsubsection{Impersonal constructions}
 \label{sec:Formal}
@@
 
-Transitive verbs may not be used like this
-\enumsentence{\shortexnt{7}
-{*Det & äter & en & fågel & frön & på & taket}
-{~\emph{it} & \emph{eats} & \emph{a} & \emph{bird} & \emph{seeds} & \emph{on }& \emph{the roof}}}
-\label{sent:fagelfron}
-
-unless their in passive form
-\enumsentence{\shortex{5}
-{Det&dricks&mycket & öl& nuförtiden}
-{\emph{It}&\emph{drinks+s}&\emph{much} &\emph{ beer}&\emph{ nowadays}}
-{``A lot of beer is being drunk these days"}}
-A very common example of this is sentences with the verbs \emph{finnas},
+Both the verb and the noun phrase have some restrictions: the determiner of the
+noun phrase must be of such type that it requires both the noun and its modifiers
+to appear in indefinite form. The verb phrase must consists of a intransitive verb
+or be in passive form.
+A very common example of impersonal constructions is sentences with the verbs \emph{finnas},
 (\emph{exist})
@@
 
-To implement this construction, we needed a
-special GF category, \verb|SimpleVP|, to exclude other verb phrases
-like the one in sentence (\ref{sent:fagelfron}). Any intransitive verb can form
-a \verb-SimpleVP-, as can transitive verbs without their objects. The \verb-SimpleVP-
-may further be modified by adverbs.
-
-There are also restrictions on the real subject, which is not allowed to be in
-definite form.
-\enumsentence{\shortexnt{6}
-{*Det     & sitter     & \textbf{den}& \textbf{fågeln}& på       & taket.}
-%& *Det sitter \textbf{fågeln} på taket.\\
-{\emph{it}& \emph{sits}& \emph{the}  & \emph{bird}    & \emph{on}& \emph{the roof}}
-%& ``There are the bird sitting on the roof'' \\
-}
-Neither is sentence (\ref{sent:dumfagel}) correct.
-\enumsentence{\shortexnt{6}
-{*Det& sitter& \textbf{min}& \textbf{fågel}& på& taket}
-{\emph{it}&\emph{sits}&\emph{my}&\emph{bird}&\emph{on}&\emph{the roof}}
-%``There are my bird sitting %on the roof''\\
-}\label{sent:dumfagel}
-In order to implement this, we utilized the 
-%The distinction lays in the determiner. The
-different types of determiners shown in figure \ref{tab:detdef}. 
-For postverbal subjects, the determiner must be of type \verb-DIndef-,
-that requires both the noun and the adjective to be indefinite.
-%The quantifier must have determiner must havenoun must be in indefinite form,
-%even though some determiners are allowed.
-%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-%a. & Det sitter några fåglar på taket    & & b. & Det sitter fåglarna på taket\\
-%   & ``There are some birds on the roof''& &    & ``There are the birds on the roof''\\
-%\end{tabular}}\label{sent:faglar}
-%Sentence \ref{sent:gamlafagel}b exemplifies a determiner that uses the noun in indefinite form,
-%but the adjective in definite form.
-%\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-%a. & Det sitter en gammal fågel på taket.& & b. & * Det sitter min gamla fågel på taket.\\
-%   & ``There are some birds on the roof''& &    &~ ``There is my old bird on the roof.''\\
-%\end{tabular}}\label{sent:gamlafagel}
-%
-%%Quantifiers like \emph{`denna'} or determiners like \emph{`samtliga'} may
-%%not be used, but the determiners \emph{`många'}, \emph{`några'} and the quantifier
-%%\emph{`ingen'} work well.
-%%Semantic difference, atm we allow any NP. \\
-To form a clause with formal subject, we hence need to inspect the determiner.
-\begin{verbatim}
-FormalSub : SimpleVP -> Det -> CN -> Cl ;
-\end{verbatim}
-The function combines a verb phrase, a determiner and a noun phrase to a clause.
-If the determiner does not fulfill the requirements stated above, the clause is
-put to NONEXIST.
-This works well for parsing, but leads to problems if the grammar is used
-for random generation. The solution is thus not ideal,
-but since the definiteness of noun phrases or determiners cannot be seen on the
-type level, it is not known until runtime whether the determiner is accepted
-in the subject.
-
-% move to future?
-As a future direction it would be interesting to examine the consequences of
-letting the noun phrases have more information on the type level. In the
-implementation for reflexive objects  (see section \ref{sec:reflexives}),
-dependent types are used for showing if a noun phrase 
-needs an antecedent.
-We would also like to differentiate between the \verb-NP-s in sentence
-(\ref{sent:mosthorses}a,b), where 
-`\emph{av}' should be used
-only when the noun has an explicit article or determiner.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. &De flesta hästarna  && b.& De flesta \textbf{av} de där hästarna.\\
- &``Most of the horses"  && b.&  ``Most of those horses."\\
-\end{tabular}}\label{sent:mosthorses}
-
-%There are other examples of when we need to know the definiteness of the noun phrase 
-%%, for example when using %the reflexive possessive pronoun `sin' (see section `Reflexive pronouns' below)
-%or when using some predeterminers (see section `Quantifiers'). Not true, not the same :(
-
-%finnas: här får man ha definit: det finns de som inte vill. Samma RelCl som
-%jag just la till, de äppplen. (de + relsats)
-
-%'det är kul att..' Näst vanligaste två-ordskombinationen i talad svenska.
-%%
-%embedded - relative, 'sådan att' not really nice for normal swedish, and no agreement
-%between subject 'en katt sådan att det regnar'. Kan säga 'en sån katt som sover, med 'sådan' + Rel
-%har lagt till Det äpple + Rel (RelCNNP). *Det äpple är rött. men Det äpple jag äter är rött.
-% 
-\subsection{Formalizing the rules for reflexive pronouns by using dependent types}
+\subsubsection{Formalizing the rules for reflexive pronouns by using dependent types}
 \label{sec:reflexives}
 An important area in a Swedish grammar is the treatment of the reflexive pronouns and
-the reflexive possessive pronouns as described in section \ref{swe:refl} 
+the reflexive possessive pronouns.
 The reflexives require an antecedent with which they agree in
 number and person.
-Our grammar should accept the following sentences:
-
-\enumsentence{Han såg sina barns skor.\\ \label{ex:ReflGen}
-He saw {\sc self's} children's shoes.}
-\enumsentence{Sina vantar hade han glömt på tåget.\\
-{\sc self's} gloves, he had forgotten on the train.}
-%\enumsentence{Han ger sina pengar till sina syskon.\\
-%He gives {\sc self's} money to {\sc self's} children.}
-\enumsentence{Hon ber alla sina kompisar att gå\\
-She asks all {\sc self's} friends to leave}\label{ex:ReflPredet}
-\enumsentence{Jag vill att han rakar sig.\\
-I want him to shave {\sc self}.}
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
-a. & Han är längre än sin kompis. & & b. &Han är här oftare än sin kompis.\\
-   & He is taller than {\sc self's} friend.
-   & & & He is here more often than {\sc self's} friend.
-\end{tabular}}\label{ex:ReflAP} \label{ex:ReflAdv}
-\eenumsentence{%\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-\item Hon tyckte om skolan och alla sina elever.\\
-     She liked the school and all {\sc self's} students.
-\item Han såg sina få böcker och sin penna.\\
- He saw {\sc self's} few books and {\sc self's} pencil.
-}\label{ex:ReflConj} \label{ex:ReflConj2}
-Reflexive pronouns can not be used in subject noun phrases of finite sentences,
-as shown by the ungrammatical examples in %, it is hence not
-%allowed as subject. The antecedent must not be out of scope for the closest finite verb,
-sentence (\ref{sent:sesig}) and (\ref{ex:ReflSubj}).
-The third person reflexives (`sig',`sin') requires a third person antecedent (see \ref{ex:refljag}).
+The third person reflexives requires a third person antecedent and
+reflexive pronouns can not be used in subject noun phrases of finite sentences,
+as shown by the ungrammatical examples in 
+sentence (\ref{ex:ReflGenB}) and (\ref{ex:ReflSubj}).
 Furthermore, the antecedent must be within the same finite sentence as the reflexive
 pronoun, see (\ref{sent:sesig}).
-The grammar should not accept any of these sentences:
-\enumsentence{
-*Sina vantar var kvar på tåget.\\
-{\sc self's} gloves were left on the train.}\label{ex:ReflSubj}
-\enumsentence{*Han och sin kompis läser en bok.\\
+
+Still our grammar should accept the sentences (\ref{ex:ReflGen}),
+(\ref{sent:rakasig}) and (\ref{ex:ReflAdv}) :
+
+\eenumsentence{
+\item
+{Sina vantar hade han glömt på tåget.\\
+{\sc self's} gloves, he had forgotten on the train.}
+\label{ex:ReflGen}
+\item
+{*Sina vantar var kvar på tåget.\\
+{~~ \sc self's} gloves was left on the train.}
+\label{ex:ReflGenB}}
+\eenumsentence{
+\item{
+Jag vill att han rakar sig.\\
+I want him to shave {\sc self}.}
+\label{sent:rakasig}
+\item
+{*Han vill att jag ser på sig.\\
+~~He wants me to look at {\sc self}.}\label{sent:sesig}
+}
+\eenumsentence{
+\item{
+Han är längre än sin kompis.\\
+He is taller than {\sc self's} friend.
+}\label{ex:ReflAP} \label{ex:ReflAdv}
+\item{*Han och sin kompis läser en bok.\\
 He and {\sc self's} friend are reading a book. }
-\enumsentence{*Jag ger sina pengar till sina syskon.\\
-I give {\sc self's} money to {\sc self's} siblings. }\label{ex:refljag}
-\enumsentence{*Han vill att jag ser på sig.\\
-He wants me to look at {\sc self}.}\label{sent:sesig}
-
-Apart from these restrictions, noun phrases containing reflexive pronouns
-may be used as any other \verb-NP-. They may be conjoined 
-(\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
+\label{ex:ReflSubj}
+}
+
 In the standard GF analysis, which is preformed bottom-up starting from the
-%TODO verify
 POS-tags, information about semantic roles are given by the functions, not by the
-categories. That is,
-we know that the first argument of the function \verb-PredVP- acts as the subject.
-The noun phrase itself does not carry information about its semantic role.
-Until it is given as an argument to a clause level function, no difference
-is made between subject and object noun phrases.
-For this reason,
-the formalization of reflexive pronouns required the use of a different analysis.
-%and used as genitives (sentence \ref{ex:ReflGen}) as such.
-%Could be combined with 'sig', a ReflNP without another noun.
-In short, what is wanted can be summarized as follows:\\
-%\begin{figure}[h]
-\begin{enumerate}
+categories. That is, we know that the first argument of the function
+\verb-PredVP- acts as the subject, but the noun phrase itself does not carry
+information about its semantic role.
+As noun phrases containing reflexive pronouns may be used as ordinary
+noun phrases, apart from the restrictions stated above,
+we do not want them to differentiate
+from other \verb-NP- on the type level since we would then need to duplicate
+our code, due to the rich type system in GF. Still, we need the type system to
+help preventing noun phrases containing reflexive pronouns to be used as subject.
+This does not only concern noun phrases, but also adverbial and adjectival phrases 
+(\ref{ex:adjTyp} a,b).
+\eenumsentence{
 \item{
-Construct a noun phrase from a common noun phrase \\
-katt $\rightarrow$ sin katt} % \\
-\item{Modify it like a \verb|NP|\\
-sina katter $\rightarrow$ alla sina katter}
-\item{Use it only as object\\
-sin katt $\rightarrow$ han såg sin katt }
-\end{enumerate}
-%\caption{Rules for noun phrases with reflexive possessive pronouns}
-%\label{fig:reflrestrict}
-%\end{figure}
-
-If we ignore the second requirement, we might come up with a solution where
-special functions for using common nouns
-together with reflexive possessive pronouns are introduced:
-\begin{verbatim}
-ReflVP: CN -> VPSlash -> VP ;
-        mat + äta     -> äta sin mat 
-\end{verbatim}
-This way, using the phrases as subject is excluded, but none of sentence
-(\ref{ex:ReflConj} a,b) or (\ref{ex:ReflPredet}) are allowed.\\
-The simplest way to fix this is to treat the phrases as normal \verb_NP_s.
-We can add a rule that allows using the possessives with common noun phrases:
-\begin{verbatim}
-ReflNP: CN -> NP ;
-        mat -> sin mat 
-\end{verbatim}
-They may now be used with predeterminers:
-\begin{verbatim}
-PredetNP : Predet -> NP       -> NP ;
-           alla   + sina barn -> alla sina barn
-\end{verbatim}
-But by keeping them as \verb_NP_s we lose information, and there is no way of
-keeping them from being used as subjects. We have ignored the third restriction
-and the grammar would generate and accept sentence (\ref{ex:ReflSubj}).
-
-We get closer to a satisfying solution by introducing new object type,
-\verb_Obj_, which is
-identical to normal \verb_NP_ except that it depends on 
-the antecedent. All functions where objects are modified in the same way
-as other \verb-NP-s are duplicated, one version is dealing 
-with objects and one with \verb-NP-.
-\begin{verbatim}
-ReflVP: Obj     -> VPSlash -> VP ;
-     -- sin mat +  äta     -> äta sin mat 
-NPtoObj : NP  -> Obj ;
-PredVP  : NP -> VP -> Cl ;  -- NP used as subject
-\end{verbatim}
-Any noun phrase may also be used as an object, whereas subjects only
-can be made up by \verb-NP-s.
-The drawback of this solution is the duplication of functions.
-\begin{verbatim}
-PredetNP  : Det -> NP  -> NP ;
-PredetObj : Det -> Obj -> Obj ;
-\end{verbatim}
-%We will need functions for conjugating objects with objects, 
-% say smt about overgeneration?
-The dependence on the antecedent needs to be applied to adverbial phrases and
-adjective phrases as well. The adjective phrase in (\ref{ex:ReflAP}a) and
-the adverbial phrase in (\ref{ex:ReflAdv}b) are examples of this; if the
-subject was changed to 2nd person, they  would need to change too.
-\eenumsentence{%\begin{tabular}[t]{@{}*{5}{l@{\ }}} 
-\item\textbf{Du} är längre än \textbf{din} kompis.\\
-``You are taller than your friend." 
-\item
-*\textbf{Du} är här oftare än \textbf{sin} kompis.\\
-``You are here more often than {\sc self's} friend."}
- \label{ex:ReflAP2} \label{ex:ReflAdv2}
-Adverbs containing reflexive possessive pronouns may further not be used
-to construct subject noun phrases.
-%TODO understandable?
-\enumsentence{\shortexnt{5}
-{*Taket & på & sitt & hus & läcker}
-{~~The roof & on & {\sc self's} & house & leaks}}
-\vspace{-2mm}
-\verb-      (NP    +(Adv (NP Obj)))-\\
-
-The dependency spreads through the grammar and in order to avoid all code
-duplication we use another approach.
-When looking at the type of
-the functions, we notice that they can be generalized:
-\begin{verbatim}
-PredetNP  : Det -> NP x  -> NP x;
-\end{verbatim}
-The solution chosen in this project is to make use of this generalization and
-introduce the use of dependent types in a resource grammar.
-Following the idea given above, we make a difference between subjects and objects, but not
-by giving them entirely different types, but by % treating them as instances of the \verb-NP- category.
+taket på sitt hus\\
+\emph{the roof of {\sc self's} house}}
+\item{
+längre än sin kompis\\
+{\emph{taller than {\sc self's} friend}}}}
+\label{ex:adjTyp}
+
+The solution chosen introduce the use of dependent types.
+We make a difference between subjects and objects, but not
+by giving them entirely different types, but by 
 letting the type \verb-NP- \textit{depend} on an argument, which may either be \verb-Subject- or
@@
 \begin{verbatim}
-cat NP NPType ;
-PredVP     : NP Obj  -> VP     -> Cl ;
-ComplSlash : VPSlash -> NP Obj -> VP ;
-PredetNP   : (a : NPType) -> Det -> NP a -> NP a ;
+PredVP       : NP Subject -> VP        -> Cl ; -- predication
+ComplVPSlash : VPSlash    -> NP Object -> VP ; -- complementation
+
+-- adverbial phrases 
+PrepNP      : (a : NPType) -> Prep -> NPTyped a -> AdvTyped a ; 
 \end{verbatim}
-The types for adverbial and adjectival phrases are also turned into dependent
-types.
-\enumsentence{\shortex{3}
-{taket & på sitt hus &}
-{\scriptsize{\textbf{NP a +}} & \scriptsize{\textbf{(Adv Obj)}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
-{\emph{the roof of {\sc self's} house}}}
-%{{\sc NP +} & {\sc (Adv with Obj)} & $\Rightarrow$ {\sc Obj}}
-\enumsentence{\shortex{3}
-{sitt hus & på berget &}
-{\scriptsize{\textbf{NP Obj + }} & \scriptsize{\textbf{Adv a}} & $\Rightarrow$ \scriptsize{\textbf{NP Obj}} }
-{\emph{{\sc self's} house on the mountain}}}
 
@@
 
-The use of dependent types separates our grammar from the other resource grammars. Such separation
-is not desirable in itself, but there are at least two reasons why we
-believe it is appropriate in this case:
-\begin{itemize}
-\item The new grammar can be made compatible with the old implementation.
-%TODO verify
-\item The possibility for noun phrases to agree with an antecedent 
-      will be useful in other languages. For example, the Slavic and Romance languages
-      have similar reflexive pronouns.
-      %and English reflexive pronouns (eg. `himself') cannot
-      %be used without antecedent.
-\end{itemize}
-Moreover, the goal of the project is to make a language specific grammar,
-whereas the common abstract is developed to be generalizable enough to 
-describe any language. It thus would be surprising if we could cover
-all syntactical aspects of Swedish while leaving the abstract grammar
-unchanged.
-
-Dependent types are also interesting for other problems.
-The rules for reciprocal pronouns could be formalized using the same idea. As
-reflexive pronouns, reciprocals (sentence \ref{sent:varandra})
-must not be used as subjects and furthermore they may only be used when the
-number of the antecedent is plural.
-\enumsentence{
-\shortex{3}
-{De &ser& varandra}
-{\emph{they} &\emph{see}& \emph{each other}}
-{``They see each other"}}\label{sent:varandra}
-%\item \shortexnt{3}
-%{*Han &ser& varandra}
-%{~he &sees& each other}}
-%
-%Since the resource grammar did not make any distinction between
-%
-%subjects and objects, the new Swedish grammar needed to be separated from the old
-%structure. 
-%The first restriction in figure \ref{fig:reflrestrict} should be slightly
-%modified, to allow certain kind of \verb|NP|s to be used.
-%\begin{figure}[h]
-%\begin{enumerate}
-%\item{
-%Construct a noun phrase from a common noun phrase or certain noun phrases \\
-%få katter $\rightarrow$ sina få katter} 
-%\end{enumerate}
-%\caption{Modification of \ref{fig:reflrestrict}}
-%\end{figure}
-%
-%The kind of noun phrases that can be used seems to coincide with the ones
-%that can be used with formal subjects (see section `Impersonal constructions').
-%>> subject controlling verbs
-%>> why comlicated with doubled types, show simplicity (elegant) with dep.types, and
-%>> some problems (many fields).
-%>> samesame as Obj, but conjunction works too, sin is Quant but what i sig?
-%>> (show circle (sig : NP -> GenNP sig -> sin??)
-%>> how we can add dependent type for Num to use 'varandra'
-%>> CN should have dettyp too, 'avståndet till sin bil'
-%\verb- PhrUtt NoPConj (UttQS (UseQCl (TTAnt TPres ASimul) PPos (QuestVP whatSg_IP (ComplSlash (NPObj ?10 (DetCN ?10 (DetQuant ?10 (IndefArt ?10) NumSg) (UseN woman_N))) (SlashV2a fight_V2))))) NoVoc-
-%  ajaj unknown here!!
-
-%>>Mention varandra. Same idea.
-%>>agreement to subj: Also good for double functions : Flickan, hon var inte dum, hon?
-% F1,2,3 : NP -> VP -> Cl ; NP needs gender (Masc,Fem,Utr,Neutr)
-
-%Make a distinction between objects and subjects?
-%do not want to copy all rules for another category. Would like depent types!
-%work pretty well by letting all NP depend on the subject in the phrase.
-%The current version is nice and works for recursive genitivs!
-%Extra info, but maybe this is just the case about NPs.
-%Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
-%*Slash* : VPSlash -> NPR -> VP ;
-%Keep number of fields down, but introduces another category, and one more
-%label to the parse tree, is it interesting? How to combine it with genitive 
-%without making it ambiguous (Han såg barnens hund).
-%What if we add alot of other rules that uses NP, will they all need to be
-%duplicated then?
-
-%'han såg sin mammas bil', they same could be applied to
-%APs for example \ref{ex:ReflAP} or Adv for
-%\enumsentence{Han leker oftare än sin kompis}
-% (ComparAdvAdj)
-%Atm you can't focus this kind of object, then a ClSlash need info about the
-%subject. FunRP also tricky (and what does it mean, we can say
-%Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
-%Kort sagt, the information about the subject must be spread
-%in different parts of the sentence. 
-%Could also decide to always produc 'sin' -> jag såg sin bil
-%to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
-%in third person, and add nonexist for others. But this way we can also get rid
-%of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
-%'jag såg din bil'.
-%Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
-%not 'alla sina', usw. Want it to handle conjuction as in example \ref{ex:refGen}
-%etc.
-%varandra, works the same but has no case for Sg.
-%%
-\newpage %until behave
-\subsection{A second future tense: ``Kommer att"}
-Swedish and Norwegian use two auxiliary verbs for future tense \\
-\cite[p. 246]{H&H}:
-%\cite[Tempus \textsection 28]{SAG}.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. & Det \textbf{kommer att} bli mörkt snart & & b. &Jag \textbf{ska} gå och lägga mig nu\\
-  & ``It will get dark soon" & &  &``I'm going to bed now"\\
-\end{tabular}}
-\enumsentence{
-Jeg kommer til å savne deg ~~$[$NO$]$\\
-``I will miss you"}
-The modal verb \emph{`ska'} signals an intention of committing the action, either from the subject or from
-the speaker. Cf.
-\enumsentence{\begin{tabular}[t]{@{}*{5}{l@{\ }}}
-a. & Du ska tycka om den & & b. &Du kommer tycka om den\\
-  & ``You \emph{shall} like it" & &  &``You will like it."\\
-\end{tabular}}
-The verb \emph{`kommer'} (\emph{`come'}), normally used with the infinite marker \emph{att},
-does not signal any intention, but that the speaker has belief that 
-it will actually come true.\\
-The resource grammar included 
-\emph{`ska'}, which was implemented as the standard way of forming future
-tense, and hence represents the translation of
-\emph{`will'}. 
-The new grammar also supports \emph{``kommer att"}, expressed as an alternative
-future tense.
-\begin{figure}[h]
-\begin{verbatim}
-(UseCl (TTAnt TFutKommer ASimul) PPos 
-       (ImpersCl (AdvVP (ComplVA become_VA (PositA dark_A)) soon_Adv))
-\end{verbatim}
-\caption{Result of parsing ``Det kommer att bli mörkt snart". The future tense
-         is marked by the constant TFutKommer}
-  \label{fig:kommeratt}
-\end{figure}
-
-%since we believe it will happend, it can not be as in :
-%'det ska (kommer*) bli fint väder imorn, men de kanske har fel'.
-Since two out of the three Scandinavian languages share this
-tense, it has been added to the Scandinavian \verb-Extra- module.
-Table \ref{fig:kommeratt} shows how the grammar expresses new tense
-in different types of sentences.
-\begin{figure}[h]
-\begin{verbatim}
-s SFutKommer Simul Pos Main : jag kommer att se henne
-s SFutKommer Simul Pos Inv  : kommer jag att se henne
-s SFutKommer Simul Pos Sub  : jag kommer att se henne
-s SFutKommer Simul Neg Main : jag kommer inte att se henne
-s SFutKommer Simul Neg Inv  : kommer jag inte att se henne
-s SFutKommer Simul Neg Sub  : jag inte kommer att se henne
-s SFutKommer Anter Pos Main : jag kommer att ha sett henne
-s SFutKommer Anter Pos Inv  : kommer jag att ha sett henne
-s SFutKommer Anter Pos Sub  : jag kommer att ha sett henne
-s SFutKommer Anter Neg Main : jag kommer inte att ha sett henne
-s SFutKommer Anter Neg Inv  : kommer jag inte att ha sett henne
-s SFutKommer Anter Neg Sub  : jag inte kommer att ha sett henne
-\end{verbatim}
-\caption{The covered and accepted usage of the future tense with `komma'}
-  \label{fig:kommeratt}
-\end{figure}
-
-%r
-%\enumsentence{\shortex{6}
-%{Betyget & 3 &ges &till &många &elever}
-%{Grade & 3 & give+\textbf{s}& to & many & students}
-%{``Grade 3 is given to many students"}}
-%\label{ex:passV3}
-%by letting the first object be `removed' and given by the subject. 
-%
-%A remark can be made about sentences where the second object is put first in the sentence, like 
-%\enumsentence{Till många elever ges betyget 3}
-%This is a variant of example \ref{ex:passV3}, \emph{`många elever'} is still the object but put in focus. The grammar
-%parse this by applying \verb|FocObj|.
-%%
-\subsection{Modifying verb phrases}
-\subsubsection{Focusing adverbs}
-The GF analysis distinguishes between two categories of adverbs: \verb-Adv- 
-%e.g. `nu' (\emph{`now'}) and
-and \verb-AdV-. %like `alltid' (\emph{`always'}).
-The \verb-AdV-  -- e.g. `aldrig' (\emph{`never'}) and `inte' (\emph{`not'}) --
-attaches directly to the verb.
-%and have a special field in the \verb-VP--table.
-\enumsentence{Cf.\\\begin{tabular}[t]{@{}*{12}{l@{\ }}}
-a. & Jag & äter & \textbf{aldrig} & fisk & & b. &Jag & äter & fisk& \textbf{nu}\\ 
-&\emph{I} & \emph{eat} & \emph{never} & \emph{fish} &&&\emph{ I }&\emph{ eat} &
-\emph{fish}& \emph{now}
-\end{tabular}\\
-``\hspace{1mm} I never eat fish" \hspace{9mm} ``I eat fish now"}
-The difference is implemented by having separate fields in the \verb-VP- table
-for the two categories.
-\begin{verbatim}
-Main clause : subj ++ verb.fin ++ verb.adV ++ verb.inf ++ verb.adv
-              han     har         aldrig      varit       här 
-\end{verbatim}
-%\begin{wrapfigure}{l}{0.3\textwidth}
-%\begin{tabular}{|l|}
-%\hline
-%bara \\
-%inte ens \\
-%till och med \\
-%\hline
-%\end{tabular}
-%\caption{Focusing adverbs}
-%\label{fig:focadv}
-%\end{wrapfigure}\\
-The adverb `bara' may be used as an \verb-AdV- but also before the finite verb,
-when emphasizing the verb itself.
-This is an example of a \textit{focusing adverb}, others examples are `inte ens' (\emph{`not even'})
-and `till och med' (\emph{`even'}).
-\enumsentence{\shortex{11}
-{a. & Han & bara & log  && &b. & Hon& till och med& visslar}
-{&\emph{he} & \emph{only} &\emph{ smiled} & & & &\emph{she} &\emph{even} & \emph{whistles}  }
-{``\hspace{2mm}He just smiled"\hspace{8mm} ``She even whistles"}}
-Focusing adverbs are accepted by the new grammar implementation, where they have
-their own field in the \verb-VP-table.
-\begin{verbatim}
-Main clause : subj ++ verb.focAdv ++ verb.fin ++ verb.adV ++ verb.inf ++ verb.adv
-              han     bara           sover        -             -           -
-\end{verbatim}
-They are also allowed as \verb-AdV-...
-\enumsentence{\shortexnt{4}
-{Han & har & bara  & sovit}
-{\emph{he} & \emph{has} & \emph{only}  &\emph{ slept}}}
-
-...or as predeterminers:
-\enumsentence{\shortex{8}
-{Det & är & bara & barn &som& får & göra & så }
-{\emph{it} & \emph{is} & \emph{only} & \emph{children} &\emph{that}& \emph{may} & \emph{do} & \emph{so} }
-{``Only children are allowed to do that"}}
-Two more rules are consequently needed
-\begin{verbatim}
-FocAdvAdV : FocAdv -> AdV ;
-PredetAdvF : AdvFoc -> Predet ;
-\end{verbatim}
-
-The focusing adverbs are usually not combined with modal verbs, the copula or temporal auxiliaries. 
-% since the modal verbs rarely are focused.
-\eenumsentence{\item[a]
-\shortexnt{5}
-{?&Han & bara&  är&  dum}
-{&\emph{he} &\emph{ only} & \emph{ is}& \emph{ stupid}}
-\item[b]
-\shortexnt{5}
-{?&Han & bara&  har&  sovit}
-{&\emph{he} &\emph{ only} & \emph{ has}& \emph{ slept}}}
-We have however chosen to allow this, since it is grammatically correct although semantically
-dubious, and since the alternative linearization \emph{``Han har bara sovit"}
-is covered by the rules using focusing adverbs as \verb-AdV-.
-%The semantic aspect of focusing adverbs remains rather unclear in this
-%representation. 
-%Discussion: should han åt bara ett äpple be (bara ett äpple) focusing an NP?
-%Then what about: han har bara sovit, han har bara ätit ett äpple.\\
-
-%Ask elisbet for other examples with inte ens och till och med, nätt och jämt.
-%In order to implement this, a new field was needed in the \verb-VP--table,
-%which may be put before the finite verb.\\
-%\verb|s SPres Simul Pos Main : han bara sover|\\
-%compare to negation, \verb-AdV- and \verb-Adv-\\
-%\verb|s SPres Simul Neg Main : han bara sover inte alltid här|\\
-%example in Sub or Inv, 
-%When used together with modal verbs, the new field results in
-%some not very natural sounding constructions:\\
-%\verb|s SPres Anter Pos Main : han bara har sovit|\\
-%but rather than changing to 'han har bara sovit', we keep it. Is a possible
-%tolkning and the other case is covered by using it as an AdV, we don't want
-%ambiguities.
-%\subsubsection{PPartAP}
-%The VP require many changes PPart, bara.
-
-%PPartAP : V2 -> AP ; --VPSlash -> AP ;
-%
-%About PPart, but do we need it with new Saldo? 'de är stuckna' , 'den nyligen
-%funna'
-%for the last one, we need VPSlash to allow adverb.
-
-\subsection{Miscellaneous}
-This section covers some minor changes of the grammar.
-They are described
-as documentation of the implementation and to illustrate some problems
-and solution of more general nature.
-\subsubsection{Relative clauses}
-The resource grammar already gave a good coverage of relative clauses and embedded sentences.
-All constructions used in examples \ref{sent:relex}a-c were accepted.
-%provided a number of ways to express relative and embedded clauses, such as
-\eenumsentence{
-%\item[a.]Frågan är vilken färg den har
-\item[a.]Pojken, som är blyg, tystnar\\
-``The boy, who is shy, falls silent"
-\item[b.]Han såg kunden som tyckte om sallad\\
-``He saw the customer who liked salad"
-\item[c.]Jag tänkte på huset i vilket hon bodde\\
-``I thought about the house in which she lived"}\label{sent:relex} % better example!!
-
-The grammar has been extended to accept a definite article for
-nouns in indefinite form, whenever the noun phrase is followed by a restrictive
-relative clause \cite[\textsection 329]{H&H}.
-Talbanken contained several examples where the modified noun is in the 
-indefinite form as in (\ref{sent:uppfatt}).
-\enumsentence{\shortex{7}
-{de uppfattningar& som& förs& fram ...} 
-{\emph{the opinions}& \emph{that}&\emph{ put+{\sc passive}} & \emph{forward ...}}
-{``the opinions, that are presented ..."}}\label{sent:uppfatt}
-%or together with the verb \emph{`finnas'}.
-When no relative clause is present, the definite form with the postnominal
-definite article must be used, cf.
-sentence (\ref{sent:debocker}).
-\eenumsentence{
-\item de uppfattningar\textbf{na} förs fram\\
-``the opinions are presented"
-\item *de uppfattningar förs fram}\label{sent:debocker}
-
-Apart from this, only corrections have been done, exemplified in
-this sentence.
-\enumsentence{
-Hon sover \textbf{som} är bra ~ \textrightarrow ~ Hon sover, \textbf{vilket} är bra\\
-``She sleeps, which is good"}
-
-%{\emph{the opinions}& \emph{that}&\emph{ put+{\textbf s}} & \emph{forward ...}}
-%{``The opinions presented ..."}}
-%
-%A few problematic things can be illuminated:
-%Since the constructions are inspired by other languages, some of them may sound quite
-%non-natural in Swedish, such as the \verb|RelS|. This combines a sentence \emph{`Hon sover'}
-%and a relative sentence \emph{`som/vilket är bra'} which wrongly expressed 
-%\emph{`She sleeps, which is good'} as
-%\emph{`Hon sover som är bra'}. The error comes from the fact that \emph{`som'} is
-%embedded in the relative clause. This is easily fixed, by adding an extra
-%parameter was needed to tell whether \emph{`som'} or \emph{`vilket'} should be used. \\
-As a side note, some complications regarding the function \verb-RelCl- can
-be pointed out.
-%Another complication stems from \verb|RelCl| which express the 
-The English implementation of this construction is \emph{`such that'},
-and the Swedish version \emph{`sådan att'} sounds awkward, except when used in of
-logic and mathematics books.
-\eenumsentence{%\begin{tabular}{ll}
-\item\emph{From the resource grammar}\\
-Jag vill ha en katt sådan att den inte fäller\\
-``I want a cat such that it does not shed"
-\item
-\emph{An alternative formulation} \\ 
-Jag vill ha en sådan katt som inte fäller\\
-``I want such a cat that it does not shed"}
-%\end{tabular}} 
-In GF, this relative clause \emph{``en katt sådan att den inte fäller"} is constructed by
-two constituents: \emph{`katt'}
-of type \verb-CN- and \emph{``den inte fäller"} of type \verb-RCl- (relative clause), glued
-together by \emph{``sådan att"}.
-As a subject \emph{`den'} is of type \verb-NP- and does not depend on any
-agreement. However, it should agree with its antecedent, \emph{`katt'}. 
-\enumsentence{*Jag vill ha en katt sådan att \textbf{det} inte fäller.} \label{sent:detfaller}
-The dependent types introduced in section \ref{sec:reflexives} could probably be
-useful to solve this problem.
-However a carefully prepared analysis and motivation is needed, %ationamount in a bigger reconstruction which therefore
-%would require a. 
-and sentence (\ref{sent:detfaller}) is still accepted by the grammar.
-
-% also ok in finnas!!
-
-%An more natural sounding alternative would be
-%\enumsentence{Jag vill ha en sådan katt som inte fäller.}
-%and hopefully this will be implemented soon, otherwise: why is it hard and how we can say
-%'en katt sådan att det regnar'. It is relative but has on empty field, still congruation is
-%needed.
-%du är här, vilket jag vet.
-%i de äpplen/det äpple du äter bor en mask/ som du ser.\\
-
-\subsubsection{An example of overgeneration and how it can be solved}
-The \verb-Extra- module contains a function for constructing genitives.
-\begin{verbatim}GenNP : NP -> Quant ;\end{verbatim}
-It selects the possessive form given in the \verb-NP- and changes its type
-to \verb-Quantifier-. 
-%\begin{verbatim}
-%(DetCN (DetQuant (GenNP (DetCN (DetQuant DefArt NumSg) (UseN boy_N))) NumSg) (UseN cat_N))
-%\end{verbatim}
-The \verb-GenNP- function %was %severely
-overgenerated when this project started, even
-though its implementation was not erroneous in itself. The cause of the overgeneration 
-was a combination of two things:
-\begin{itemize}
-\item \verb-Quantifiers- do not have any possessive
-forms, so when a quantifier is turned into
-a \verb-NP- (by the functions \verb-DetQuant- and \verb-DetNP-), its genitive field gets
-the same form as its nominative field.
-\item\verb-GenNP-
-introduced a cycle: a \verb-Quant- may be turned into a \verb-NP-, and now
-a \verb-NP- can be turned into a \verb-Quant-.\\
-\verb|GenNP : NP -> Quant, DetQuant : Quant -> Det; DetNP : Det -> NP|
-\end{itemize}
-We had that a noun phrase, eg \emph{`katten'} could be turned into a quantifier,
-by using its genitive form, \emph{``kattens'}. This could be used as a determiner,
-\enumsentence{Det är kattens mat.\\``It is the cat's food."}
-and thus as a noun phrase:
-\enumsentence{Det är kattens.\\``It is the cat's."}
-The possessive form of this \verb-NP- is now the same as its nominative:
-\emph{`kattens'}. Hence, the function \verb-GenNP- could be applied
-any number of times without any change in the output.
-
-The problem can be solved by changing the implementation of the function
-\verb-DetNP-  which is responsible for creating the possessive form
-of quantifiers and determiners.
-We let it append an extra
-\emph{s} to the word, so that applying \verb-GenNP- to \emph{`kattens'} results
-in \emph{`katten\textbf{ss}'}. Even though this generates a form that does not exist, it is perhaps
-the best approximation to what it would look like, given that it was indeed used in 
-Swedish. Most important is however that this stopped the \verb-GenNP-
-from introducing unnecessary ambiguities.
-The number of parse trees for the phrase \emph{``Kattens bils hus"} was reduced
-from over one hundred to six.
-Besides, the grammar will allow the actually existing genitive
-forms of quantifiers:
-\enumsentence{\shortex{6}
-{Somligas &skor& står& kvar& i& hallen}
-{\emph{some's }&\emph{shoes}& \emph{stand}& \emph{left}& \emph{in}& \emph{the hall}}
-{``Some peoples shoes are still in the hall"}}
-
-%The six remaining trees describes different interpretations of the sentence,
-%such as `hus' is in genitive or nominative and in either plural or singular.
-%kattes bils hus är stort.
-%kattens bils hus är stora.
-
-%the extramodule contained GenNP for costructing genitvives. In Swedish
-%done by adding s to the end of the word. In Gf, Uses the NPPoss form of the noun. 
-%severely overgenerating when started, and had to be restricted. Cycle:
-%      katt -> kattens      ->          kattens           det är kattens
-%Restricted by removing the sp-field which is used by DetNP. Disallows ex 'det or kattens',
-%but still much better then the overgeneration.
-%Quants have no gen-form. min -> mins?? but when NPs they must have! otherwise min could be
-%min with GenNP applied n number of times. solutions: bind s.
-%or: A field should also been added to Quants, so that they can have a gen-form
-% : somligas.
-%1 cannot be alone (fixed) 2 some words have id. (no) gen form (fixed)
-%           fixing this include restructuring all???
-
-\subsubsection{Using Talbanken to assign correct categorisations}
-\label{sec:gf.quant}
-One of the biggest differences between the GF analysis and the one made in Talbanken
-is related to pronouns. 
-As noted in section \ref{sec:gfnouns} our GF grammar only classifies personal pronouns
-as pronouns. Many pronouns could not be imported from Saldo, since the correct category
-could not be identified automatically. Therefore, some experiments were
-conducted on how to extract this 
-information from Talbanken. We were interested in words which can be used
-the same ways as `sådana' (\emph{`such'}) that acts either as a
-determiner (\ref{ex:QuantDet}), or as an adjective agreeing with the noun
-(\ref{ex:QuantAdj}-c). 
-\eenumsentence{
-\item\shortex{5}
-{Sådana& \label{ex:QuantDet}katter& vill& jag& ha}
-{\emph{such}& \label{ex:QuantDet}\emph{cats}& \emph{want}& \emph{I}& \emph{have}}
-{``I want such cats'}
-\item\shortex{3}
-{Han& är& \label{ex:QuantAdj}sådan\toplabel{sent:Quant}}
-{\emph{he}& \emph{is}& \label{ex:QuantAdj}\emph{such}\toplabel{sent:Quant}}
-{``He's like that"}
-\item\shortex{3}
-{Huset& är& \label{ex:QuantAdj2}sådan\textbf{t}}
-{\emph{the house}& \emph{is}& \label{ex:QuantAdj2}\emph{such}}
-{``The house is like that"}
-}
-
-By starting from a list containing all pronouns in Saldo and all forms they may occur in,
-Talbanken was harvested for the ones that both occurred with the determiner tag \verb-DT-
-and with the present participle tag \verb-SP-. This way we are left with a limited number
-of words that could easily be analysed and added to the lexicon.
-%Further, list of all pronous acting as determiners were extracted.
-
-This method is
-both faster and more reliable than going through the list by hand and categorising them
-manually by looking up each word or by using introspection.
-
-%Determiners can be used alone as noun phrases, as in sentence
-%\ref{ex:QuantDet2}.
-%\enumsentence{Sådant är inte bra}\label{ex:QuantDet2}
-%The GF analyse differs sometimes. Experiments of how to extract
-%this information from Talbanken by using Saldo as a starting point.
-%While GF only recognizes the personal pronouns, 
-%`sådan', `fler`, `somliga' etc. are all regarded as pronouns in Talbanken and Saldo.
-%In the GF analyse they may be quantifiers, determiners or predeterminers. 
-%All pronouns from Saldo was listed and by searching Talbanken for which tags
-%they got we could sort out the ones acting either as
-%determiners, as `sådana' in sentence (\ref{ex:QuantDet}), or as adjectives as
-%`sådan' in sentence (\ref{ex:QuantAdj}). 
-%Since all determiners can be used alone as noun phrases, sentence
-%\ref{ex:QuantDet2} is also accepted.
-%what is the difference really???
-%The techniques of searching Talbanken can be used
-%for this purpose as well. 
-%combining the information in Saldo and Talbanken
-%pronouns/Quant. Many, created categories for 'sådan', and 'fler'. Not many in each,
-%has to decide, one category for each?
-
-%\subsubsection{Other small things}
-%
-%\textbf{Det-utr}. no effect for the some Dets, like 'sådan', no difference between utrum and neutrum.
-% Can vary in number, since is a NP. Do not want to cut connection to abstract but changing them,
-% so added one more. Leads to ambiguities in some cases, but considering that we would want to use
-% this for translation, neutr and utr may differ in the other language.
-%
-%\textbf{finiteness}
-% 'de flesta de gula hästarna'
-%
-%%\subsubsection{advs}
-%%-- or is this really just AdVs? 
-%%-- redan is Adv in abstract. Wrong!!
-%%-- what about 'i regel'?
-%%to change position of adv:
-%%normal: after verb 'han äpplet redan'
-%%now want: 'äter redan äpplet'
-%%          'har redan ätit äpplet' 
-%%          'har ätit redan äpplet' :( which doesn't work either :)
-%\textbf{antalet}
-%Also needed categories for 'pronouns' which can act in different ways.
-%and 'antal','sorter' - need indetermined, antal plural, sorter either.
-%Should be able to modify by a arbitrary antal adjective, and they must be determined.
-%'(ett stort antal) människor' and not 'ett stort (antal människor)'
-%Don't want them as a NPs or CNs since we don't want other Ns to be used this way
-%'en flicka kaffe', this is then Appos? (Or do we?). Should also not be utsatta for what other
-%NPs or CN can be. Dessutom, we want to make use of the comp-fiel. 'mamma till', but is this really
-%the same??
-%
-%\textbf{SuperlA}
-%
-%In general, many categories need more parameters and fields, tex Quant can be used as NP
-%and therefor needs to depend on Case, de -> dem.
-
-%here or in robust: SupCl, det as VP or VV or S for S (remember Foc)
-\section{Testing}
-Throughout the project regression testing has been used. Every time the grammar was updated
-it was tested against a treebank consisting of 155 sentences, and the number of
-parse trees were compared to a standard.
-The purpose was first of all to
-make sure that the coverage was not decreased, but also to make sure that we
-would notice if new ambiguities or overgenerating functions were introduced.
-If so, the ambiguities could be removed when possible, and when not, we were at least made
-aware of them and could decide whether the increase of coverage were worth the
-increase in ambiguities.
-%The grammar will always contain ambiguities, but by being aware helps to
-%handle them in other ways, such by statistical means.
-
-New functions were also tested by random generating and by creating tables showing
-all forms, such as table \ref{fig:kommeratt}.
-
-%For testing has been carried out by using a treebank, consisting of sentences used in 
-%sub-testsuites. Also have some sentences that should not
-%be parsed. 
-%When adding new constructions, the table-function is very helpful, shows how a word will
-%turn out in differnet cases. all tenses, for komma.
-
-% One part about the grammar. The state of the original grammar, what was covered and not. 
-% What parts that I have been working on and why, 
-% what constructions I wanted to add, examples of sentences that should work
-% and of sentences that shouldn't be covered. How to separate Swedish from
-% Scandinavian, bigger reconstructions in the grammar. 
-% sammanhängande explanation of the swedish grammar, choices I have made, why
-% regression testing
-
-\chapter{Extracting a GF treebank} %bootstraping/extracting a GF treebank
-\label{sec:Mapping}
-%>> no subjects or object, analyse is more from the lexicon,
-%>> compare to others, ask elisabet.
-%>> diffs: valencies, pp (gf ingen skillnad direkt/indirekt i träd, pga multiling)
-%>> mamba very used and tested for swedish, standard.
-%>> focus on what we DO cover.\\
-Talbanken contains much valuable information about phrase structures and word 
-usage. It is analyzed with the MAMBA annotation scheme (Telemann, 1974).
-% which is 
-%>> cite mamba? annotation: well intergrated in sw. tradition, descriptive, used
-%>> for large scale projects.
-%annotation scheme
-%commonly used, and hence very well-tested, for describing Swedish.
-One part of this project has focused on translating trees from Talbanken to GF
-by constructing a mapping, which 
-automatically transforms trees in the Talbanken format to GF abstract syntax trees.
-We hence get a comparison between the two annotations and at the same time
-we get methods for extracting and translating to GF notation. 
-Figure \ref{fig:translationtrees} shows an example of a visualized Talbanken05 tree
-of the sentence ``Katten på bilen blir större" (\emph{``The cat on the car gets bigger"})
-and the result of translating it to GF.
-Both the POS tags as well as the syntactic information are needed
-during the translation, so that all the information shown in the abstract tree
-(\ref{pic:gftree}) can be extracted. We need to know that \emph{katten} is a noun
-used in definite form (\verb-DefArt-), singular (\verb-NumSg-). Even though the
-notation is translated, the original analyse from Talbanken
-is preserved.
-\emph{på bilen} should still form a subtree attached to \emph{katten}.
-%produces a \verb-NP- from `katten' this means that
-%DetCN : Quant                  -> CN    -> NP
-%DetCN  (DetQuant DefArt NumSg)    cat_N = katten
-
-\begin{figure}[h]
-\hspace{-30mm}
-\subfloat[Talbanken tree]{\label{pic:tbtree}\includegraphics[width=60mm]{Talbankentree.pdf}}
-\hspace{-10mm}
-\subfloat[GF abstract tree]{\label{pic:gftree}\includegraphics[width=100mm]{gfTree.pdf}}
-\subfloat[The corresponding GF parse tree]{\label{pic:gfptree}\includegraphics[width=50mm]{gfPTree.pdf}}
-\caption{Trees for the sentence ``Katten på bilen blir större".}
-\label{fig:translationtrees}
-\end{figure}
-
-The mapping is based on 
-a translation of the English Penn Treebank \cite{gfpenn}.
-By modifying this program, we now have a translation that works for
-the Tiger XML-format of
-Talbanken. Adaption was required for the differences in annotation as well as 
-for the syntactic differences between Swedish and English.
-
-The translation %connects GF with another annotations and 
-gives us means to evaluate our parser. By both parsing a Talbanken sentence and
-transforming its annotated tree, we can easily inspect if the results are
-equal.
-Additionally, the mapping shows which grammatical constructions that are still missing
-from the GF grammar and shows how the GF analysis differs from the one made
-in Talbanken.
-If there are words missing from our dictionary, the rich
-POS-tags may help us to automatically find the correct declination and add it to the
-lexicon. Further, our parser will need probabilities (see section
-\ref{sec:future}) of how often a function is used. The GF treebank we
-achieve from the translation is a good source for this information.\\
-%missing words. Lexical aquisation, probabilities, see section .
-%The mapping gives information about which form a word is
-%currently used in and this may be used by the lexical extraction
-%tools. % Those can be enhanced if they are given more data. 
-%The translated trees enable us to extract those probabilities. %The mapping gives us a source for this.
-%Furthermore, the translation makes it easy to identify grammatical constructions
-%missing from the GF grammar Another important use of the mapping is evaluation of the parser, which can be
-%accomplished by comparing the parse trees and the trees from the transformer.
-
-%The mapping works for Talbanken05 in Tiger XML-format.
-\section{The Talbanken annotation}
+The dependent types in GF have been used for logical reasoning before but this 
+is the first example of using it in a grammar describing pure syntax of
+a natural language. Other languages having similar relations between 
+noun phrases and antecedent could be extended to use the type system in similar
+ways.
+
+\subsubsection{Overgenerating}
+Another aspect of the grammar implementation was the avoidance of overgeneration. As
+the Swedish resource grammar had not been used for larger projects, many examples of
+overgeneration were present that did not cause problems when working with small
+lexicons and controlled languages, but that caused explosions of ambiguities when 
+tested on Talbanken. For the function \verb-GenNP-, used for genitive, we managed
+to decrease the number of parse trees for some sentences to 10 \%.
+% Could also be here:
+%Experiments were also conducted on how we can use Talbanken for assigning correct
+%types to words which we couldnot import directly from SALDO, due to annotational
+%differeneces.
+
+\section{A GF Treebank for Swedish}
+\label{sec:mapping}
+
+%Chapter 5 - short introduction about Talbanken tags + difference in
+%the tags + 5.2(the translation process - first part) + example of
+%2 trees  
+Why we do this...... when is meta used
 Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
@@
 \hline
-\textbf{SV} & ska & The verb \emph{shall}\\
-\textbf{WVIV} & vilja & The verb \emph{want} in infinitive\\
-\textbf{KVSP} & kommit att & The future auxiliary verb \emph{kommer} in supine\\
-\textbf{GVPT  PA} & gjordes & The verb \emph{do, make}, present, passive \\
-\textbf{NNDDHSGG} & familjemedlemmarnas & Noun, definite, compound (person), genitive \\
-\textbf{PU} & * \; 1) \; a) & List item\\
+\textbf{SV} & \emph{ska} (`shall') & The verb \emph{shall}\\
+\textbf{NNDDHSGG} & \emph{familjemedlemmarnas} & Noun, definite, compound (person),\\
+&(`the family members' ') &genitive \\
+\textbf{XX} & & Unclassifiable part-of-speech\\
 \end{tabular}\\
 
-%Pos continues
-%to either new tree or a word. Example S -> PP -> (Hd -> Id , Hd -> Id). Show
-%corresponding GF tree.
-
-\section{The translation process}
-The implementation consists of a set of rules that translate
-nodes of a Talbanken tree to GF functions, possibly by describing what the
-nodes subtree should look like. 
-\begin{verbatim}
-translate S  = do
-             np <- find NP
-             vp <- find VP
-             return (PredVP np vp)
-\end{verbatim}
-This rule tells that we may map the category \verb-S- to the function 
-\verb-PredVP-, given that we find a noun phrase and a verb phrase in
-the subtree of \verb-S-.
-
-At every leaf a word is given. %When a leaf is reached,and ato see if
-It is looked-up in the GF lexicon and the result compared to the information of
-the POS-tag.
-If the word is not found in the lexicon
-a \textit{meta variable} will be put in its place, denoted `\verb-?-'.
-The meta variable is also used as a function, needed when we have found two
-translatable trees but no rule that allows them to be combined.
-
-%Lets look at the example 
-In figure \ref{pic:tbtree} the verb \emph{`bli'} is
-followed by a predicative 
-adjective modifying the subject tagged with \verb|SP|. Our GF grammar admits
-\emph{`bli'} to have type \verb|VA|, a verb with an adjective as complement,
-and the translator can hence apply the GF function \verb|ComplVA| to combine the verb
-with its object. \\
-
-\noindent The translation covers complex noun phrases which may consist of pronouns,
-reflexive objects or common nouns and contain predeterminers, 
-determiners, quantifiers and adjectival modifiers. Special rules were needed
-for each verb category in order to find the right number of arguments.
-The mapping also covers different word order, such as questions and topicalisation.
-
-%The translation works section-wise, ie. if it finds a tree of category 
-%it will start to look for a subject \verb|SS| and a finite verb \verb|FV|. If it
-%can find those, they can be combine with the GF function \verb|PredVP|.
-%It is not always the case that we first find the subject, then the verb and 
-%finally the objects. Often the order is Verb-Subject-Object or Object-Subject-Verb. \\
-%During the mapping process, the translator also needs to find out whether the sentence is
-%negated and what tense is used. If it finds two trees of type, say, \verb|VP| but does not
-%have any rules for combining them, the \textit{meta function} \verb-?- will be used. 
-%Figure xx shows an example of this.\\
-
-%How we try to build up a 'abstract syntax tree', a S with a NP VP for example. 
-%Continue with dividing VP into Verb Objs. More possible word orders (or at
-%least more commonly used). 'katten åt sin mat fort' 'åt katten sin mat fort' 'sin mat åt katten fort'
-% If can't find, add a '?', a meta.
-%This is also what happens when the translator finds a combination it doesn't know about.
-%Example of tree with meta.
-
-%Description of most important tags from Talbanken, their translation? appendix?
-\subsection{Differences in the notation}
-Some structural differences between GF and Talbanken became obvious during the
-translation.
-%Since the translation is done word by word, ambiguities may confuse the translator.\\
-%shown during the translation.
-%cause some translation problems.
-In Talbanken the valency information is given implicitly by the complements
-of a word. If a verb is followed by an object, \verb-OO-, which contains an
+The mapping could in some cases be performed tag-by-tag, but annotational
+differences complicated the translation.
+One example is valency information, which is given implicitly by the complements
+of a word in Talbanken. If a verb is followed by an object, \verb-OO-, which contains an
 \verb-S-, we can conclude that this is a sentence-complement verb.
@@
 sentence.
-\begin{wrapfigure}{l}{0.52\textwidth}
-\includegraphics[width=80mm]{metatree.pdf}
-\caption{Abstract tree with meta variable}
-\label{pic:gMetatree}
-\end{wrapfigure}
-
-\noindent A similar issue arises from the use of verb particles. A verb with a
-particle is treated as two independent units in Talbanken but in GF they are
-considered to be one constituent. % which means that 
-The particle has therefore no effect on the abstract tree, its presence is
-announced by the verb \\itself.
-
-%In Talbanken, verbs and particles are treated as separate units,
-%and the particle is marked with the tag \verb-PL- and attached to. 
-%Talbanken has no valence information while this is crucial for GF.
-%The grammar requires a adjective-complement verb,
-%\verb|VA|, to be used with one adjectival phrase. For \emph{`bli'} in example
-%\ref{pic:tbtree}, we know that we need to find an adjective phrase.
-%%know what kind of verb we are looking for
-%%when consulting the GF lexicon.
-%If the GF lexicon assings different types to verb, the correct type 
-%should be showed by the complements.
-%This way %\enumsentence{Han gillar det\\He likes it} 
-When constructing the abstract trees, we must of course follow and use the
-GF analysis. However,
-the mapping cannot depend on the valencies given in GF only, 
-\newpage
-
-\noindent since the distinction between different types of verbs is not always clear in Swedish;
-most transitive verbs can also be used as intransitive. Both sentences in
-example (\ref{ex:mappTransitive}) are grammatically correct but only (\ref{ex:mappTransitive}b)
-is accepted by the GF grammar: 
-the grammar rules are simply too strict for accepting many Talbanken sentences.
-\enumsentence{\begin{tabular}[t]{@{}*{14}{l@{\ }}}
-a. &Han& sitter& och &läser.&& & b. &Han& sitter& och& läser& en& bok.\\
-&\emph{he}& \emph{sits}& \emph{and} &\emph{reads} &&&
-&\emph{he}& \emph{sits}& \emph{and}& \emph{reads}& \emph{a}& \emph{book}\\
-\end{tabular}}
-\label{ex:mappTransitive}
-%See section \ref{sec:futureValency} for a longer discussion about this.\\
-%the Talbankenwill allow words to be used in other ways than their
-%GF valencies prescribes.
-
 Another annotational difference that turned out to be problematic for the
-translation can be illustrated by sentence (\ref{ex:mappMan}), containing
-the generic pronoun \emph{`man'}.
-\enumsentence{
-\shortex{3}{Man & uppskattar & dem}
-{\emph{one} & \emph{appreciates} & \emph{them}}
-{``They are appreciated"}}\label{ex:mappMan} %  (s216).
-
+translation can be illustrated by analysis of the generic pronoun \emph{`man'}.
 In Talbanken, \emph{`man'} is simply a personal pronoun, 
-in GF it is represented by using the function \verb|GenericCl|, see figure
-\ref{fig:mappMan}. \verb|GenericCl| turns a \verb-VP- into a clause and has no
-subject visible in the abstract tree. 
-\begin{figure}[h]
-\centering
-\includegraphics[width=80mm]{man.pdf}
-\caption{Abstract tree for ``Man uppskattar dem".}
-\label{fig:mappMan}
-\end{figure}
+in GF it is represented by using the function \verb|GenericCl| which turns a
+\verb-VP- into a clause and has no subject visible in the abstract tree.
+
 When translating a sentence like this, it is thus not possible to simply glue the
 subject to the verb with \verb|PredVP|. For each similar GF construction, an extra rule
-would be needed to get full coverage. \\ %and the normal top-down? approach fails.\\
-%Therefore, the translator must change the type of the 
-%sentence when inside the SS.
-
-%The lexicon may contain a verb both with and without a particle,
-%for example \emph{tänka} (`think') and \emph{tänka till} (`consider').
-%\begin{verbatim}
-%taenka_V  : V ;
-%taenka_till_V : V ;
-%\end{verbatim}
-%When the translator looks up a the form \emph{tänka}, both of the two verbs
-%will be among the results. %It is not known until later which one 
-%%The first option will be selected, and if the
-%%particle isn't directly attached to the verb
-%%There is, at this point, no easy way of knowing which one of them wants a particle and the
-%the transformation might confuse the two words. This is also the case when translating
-%all sorts of idioms that GF considers to be just one bit.\\
-%
-%Difficulties in looking up 'idioms' like this
-%in the gf lexicon. Therefore the translator might get confused.
-%Sentences are also differntly deep. Example of a conjuncted sentence.
-
-\section{Results and evaluation}
-%>> more! replace words, how much better results (both grammar and mapping) 
-%>> more results
-%>> motivation, why so bad??\\
-The development was mostly example-driven, and at least one rule for the translation of
-every Talbanken-tag has been implemented.
-Shorter sentences have been prioritized in order to get a coverage of the most
-fundamental constructions. 
-Regression testing was used to ensure that a new rule would not violate the old ones
-or allow trees to be translated erroneously.
-
-The flat version of Talbanken05 was used for developing the mapping, but
-the program works for both the flat and the deep Tiger-XML versions.
-The deeper contains six more tags, although they are not always useful for our       
-purpose. Figure \ref{fig:mappDeepFlat} shows the difference for a conjunction.
-% show piece of s5 here!
-\begin{figure}[h]
-\begin{parsetree}
-    ( .CNP. 
-        (.CJ.  (.NP. ~ `oskifta dödsbon'))
-        ( .++. (.++OC. `och'))
-        (.CJ.  (.VN. `familjestiftelser'))
-    )
-\end{parsetree}
-\begin{parsetree}
-    ( .CNP. 
-        (.CJ.  (.NP. ~ `oskifta dödsbon'))
-        ( .CC.
-            (.CNP. 
-                (.++.  (.++OC. `och' ))
-                (.C+.  (.VN. `familjestiftelser'))
-         ))
-    )
-\end{parsetree}
-\caption{An example of the difference in the two analyses}
-\label{fig:mappDeepFlat}
-\end{figure}
-The focus on the mapping has been simple sentences, in order to cover the most
-fundamental structures. If the program is extended and adapted to cover complex
-conjunction, a deeper evaluation should be preformed to see which version that
-suits our needs the best.
-The deeper implementation also gives 
-some more information about verb phrases, by the tag \verb|VG| (\emph{verb group}),
-which groups an infinitival verb and its object together into a \verb|VP|.
-This information can however be extracted from the flat implementation, and the results
-get slightly better when using this version. \\
-
-\noindent When evaluating the mapping, the results strongly depend on which restrictions we
-put on the input. 
-One of the reasons why a node cannot be translated, is the 
-use of the tags show in figure \ref{fig:mapBadtag}.
-The \textbf{PU} tag is used for graphic listings, and not for fluent text.
-In our grammar there is naturally no corresponding function; 
-%We are not interested in sentences made up by list, since 
-the listings are meant for making the text look nice in
-folders etc and outside the scope for the grammar itself. The tags \textbf{XX}
-and \textbf{NAC} are often used since Talbanken makes a
+would be needed to get full coverage.
+
+The tags \textbf{XX} and \textbf{NAC} (\verb-Not a constituent-) are often used since Talbanken makes a
 difference between subject
-and object noun phrases. %(which is not needed in GF (see section \ref{sec:swegf}))
+and object noun phrases.
 The analysis of elliptical expression in (\ref{sent:krav})
@@
 ``Too high demands."}\label{sent:krav}
-contains the tags \textbf{XX} and \textbf{NAC}, since it is not obvious
+contains both these tags, since it is not obvious
 whether the noun phrase is used as subject or an object.
-The tags shown in figure \ref{fig:mapBadtag} occur quite frequently in the treebank and are always translated
+These tags occur quite frequently in the treebank and are always translated
 to metas, which lowers our result. \\
-%How we cannot cover all cases. 
-\begin{figure}[h]
-\begin{tabular}{ll}
-\textbf{NAC} & Not a constituent\\
-\textbf{XP} & Other (non-coordinated) phrase\\
-\textbf{DB} & Doubled function\\
-\textbf{PU} & List item\\
-\textbf{XX} & Unclassifiable part-of-speech\\
-\end{tabular}
-\caption{}\label{fig:mapBadtag}
-\end{figure}
-
-The main goal has been to be able to translate shorter sentences, with no
-idioms or conjunction.
-If we assure that the lexicon contains the correct word class for all lemmas
-involved, we
-can restore more than 85 \% of the nodes in the original tree.
-If we lift all the restrictions excluding the \verb|PU|, we get
-%When tested the 100 first sentences from Talbanken, we get
-65 \% coverage. 
-If we test randomly collected sentences that do not contain any of the tags listed
-in figure \ref{fig:mapBadtag}, 72 \% can be restored (see figure \ref{tab:mappres})
-
-%TODO
-%All NPs: 72%
-\begin{figure}[h]
-\begin{tabular}{|ll|}
-\hline
-No list items & 65 \%\\
-No special punctuation or bad tags& 72 \%\\
-Short sentences with known words & 85 \%\\
-\hline
-\end{tabular}
-\caption{}
-\label{tab:mappres}
-\end{figure}
-
-A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
-The percentage of restored nodes from Peen Treebank is higher than our results.
-%TODO review
-The reason for this may be the fact that English is syntactically less complicated
-than Swedish.
-Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
-where idiomatic expressions are more likely the be appear and the language
-presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
-Also, the Penn Treebank contains a lower number of tags, 82 compared
-to more than 300 in Talbanken. Even if the tags describing the same word
-%TODO review
-class as another tag are excluded, Talbanken still leaves us with more than 130 tags.
-With more tags, we get more information, but as the number 
-%Thus the number of
-increase, so does the amount of work of finding the correct translation 
-for each combination of tags and writing rules that cover all constructions. 
-%iand so is the number of
-%rules that would be needed to cover all constructions.
-
-We believe that the results could be enhanced by simply adding more rules 
-and in this way get a wider coverage. There are many special cases that require
-special rules. Since we are not aware of any formal specification of how the tags may be
-combined in Talbanken, the only way of finding all possibilities are to manually look
-for different patterns. Another option would be to make the mapping more
-robust, but the robustness must not interfere with the correctness.
-
-% compared to penn treebank: 82 tags. talbanken: at least 130
-% 99.1 % of Penn sentences have 6 or fewer clauses
-
-%If we don't take the verb valency into account, allow all verbs to have Gf category V,
-%we get a slightly better result,
-%from 62.5 \% to 65.4 \%.  (see EvalText). The complemnt of the verbs should tell which category
-%they are anyway.
-%Evaluation. -> Works for both flat and deep, but with sligthly better results for the flat one.
-
-% hard examples like
-% ger sig till henne -> ReflSlash ge henne in gf
-% questions with objects that should be moved
-% vilken katt såg du
-%This part explains
-%the tags in Talbanken and gives examples of how they are translated,
-%examples of what can be parsed and what cannot.
-%about how to make the code work for swedish, differences
-%regression testing
-
-\chapter{Discussion}
-\label{sec:results}
+
+The evaluation measures the numbers of \textit{metas} in each translated tree.
+If the finite verb is untranslatable, we get metas for the verb itself and for the
+tense.
+For all of Talbanken, we get a number of 65 \%, and if we limit our input to
+uncomplicated sentences with no more than 10 words where all words are known to
+our lexicon, we can reach 85 \%. When translating the whole treebank, 78 sentences
+could not be mapped at all and resulted in \textit{metas} only.
+
+\section{Chunk Parsing}
+
+%New section - insights about latest work
+The grammar, which is hand written, cannot be expected to cover all
+possible expressions of the language.
+Hence, we combine it with statistical means to get good coverage. Our approach
+makes use of the rich annotation in Talbanken -- we perform chunk parsing
+relying on some of the tags given in the treebank. 
+Unknown words and grammatical constructions, idioms or ellipses cannot be
+handled from within the grammar.
+Therefore, whenever a sentence is not recognized by the grammar, it is 
+parsed chunk by chunk.
+Noun phrases, adverbial phrases and ? are considered. At the first stage,
+we try to parse the basic sentence structure, and therefore allow the parser
+to parse complicated chunks as dummy words. This way, we don't loose the 
+high-level sentence analysis given by GF.
+When this is done, we try to separately parse the chunks,
+until as much of the sentence as possible is covered.
+
+This way we get good coverage -- from ca 10 \% of Talbanken -- to ??
+
+\section{Language Model-Based Disambiguation}
+
+ Maybe move this to future work, we have it but not really sure
+ if it is actually used atm.
+
+%New section - fill in with the basic principle + new results(when available)
+The translation of Talbanken to GF gives us a GF treebank consisting of more than
+6000 sentences. The translation process is explained in section
+\ref{sec:mapping} and even though most trees are incomplete, that is contain
+meta-variables, they still give valuable 
+information about which words from our lexicon that have been used
+and which functions that are needed to parse the sentence.
+GF has built-in support for probabilities, and when feeding the output
+from the mapping to our grammar, we get a model of the language. We believe
+this data can be trusted as it is extracted from manually annotated 
+real-world sentences.
+
+This model is used for disambiguation: the parse trees for each sentence
+are ranked by the sum of the probability of each function used.
+
 \section{Evaluation}
-%TODO
-%>>evolution diagram?
-%also removed unnecessary ambigutiies.
-%>> how to handle ambigs: know them and controll them. using statistics or ontolgies.
-%>> that is: disambiguitation by context or by meaning.
-The project has resulted in
-\begin{itemize}
-\item a large-scale GF lexicon and a program to redo the importation when needed
-\item an extended grammar covering an important part of Swedish
-\item a comparison between GF and another annotation
-\item a deeper testing of the Swedish resource grammar and an estimation
-of how well GF can be used to describe larger parts of a language
-\item a study of how dependent types can be used in the resource grammars
-\end{itemize}
-
-Besides being capable of reimporting Saldo, the lexicon extraction program could also
-be modified for importing other lexical resources. The only requirement is that the resource
-provides inflection tables.
-
-The grammar has been extended and enhanced, and its current status is
-a specialized extension of the resource grammar.
-Besides parsing, the grammar may well be used for language generation,
-which works fast even when using an extensive lexicon.
-%TODO make nice!
-Although it is not been formally verified, we believe that the majority of the
-sentences generated are grammatically correct in a syntactical point of view.
-Without any semantic knowledge, nonsense phrases cannot be avoided in
-random generation. However, given that the abstract tree has a meaningful
-interpretation,
-%TODO move to Misc?
-the linearization should be correct. There are some cases when the correctness
-of the output has been put aside in order to increase the expressivity, such
-as \verb-ComplBareVV-, which use a verb-complement verb without the infinitive mark, 
-as in sentence (\ref{sent:complVV1}b).
-\eenumsentence{\item\shortex{5}
-{Jag& börjar&\textbf{att}& bli& hungrig}
-{\emph{I}& \emph{begin}& \textbf{\emph{to}}&\emph{become}& \emph{hungry}}
-{``I'm getting hungry"}
-\item
-\shortex{4}
-{Jag& börjar& bli& hungrig}
-{\emph{I}& \emph{begin}& \emph{become}& \emph{hungry}}
-{``I'm getting hungry"}}\label{sent:complVV1}
-Since there is no information in the grammar or lexicon about which verbs that
-allows the infinitive mark to be left out, it will also allow the more
-questionable sentence (\ref{sent:complVV2}b).
-\eenumsentence{
-\item \shortex{5}
-{Jag & låter &bli & \textbf{att}& titta} 
-{\emph{I} & \emph{let} & \emph{be} &\textbf{\emph{to}} &\emph{watch}}
-{``I refrain from watching"}
-\item \shortexnt{4}
-{*Jag & låter & bli & titta} 
-{~~\emph{I} & \emph{leave} &\emph{be} & \emph{watch}}}\label{sent:complVV2}
-As these functions serve to provide stylistic flexibility when parsing, they can be left out
-when generating language, and the grammar then generates text of good
-grammatical quality. \\
-%ComplBareVV, SupCl
-%For random generation, the sentences generated may be next to impossible
-%to interpret without looking at the abstract tree, since they may contain
-%too many attributes, relative clauses etc.
-
-%\enumsentence{Kommer övriga datorer idag som ska vara inte att ha målats?}
-%\label{sent:what}
-%Sentence (\ref{sent:what})
-%PredVP (DetCN Subject (DetSub oovriga_Det) (RelCN (AdvCN (UseN computer_N) (AdvSub today_Adv)) (UseRCl (TTAnt TFut ASimul) PPos (RelVP IdRP UseCopula)))) (PassVP (SlashV2A paint_V2A (ComparA Object red_A (UsePN Object johan_PN))))
-%The use of the NONEXIST pattern, as described in \ref{sec:Formal}. When using
-%the grammar for generation, an extra
-
-\noindent When it comes to parsing, we do not get far without robustness.
-The grammar in itself is by no means robust, and just one 
-unexpected punctuation mark, unknown word
-or ellipsis will cause the parsing of the whole sentence to fail.
-
-Parsing bigger parts of Talbanken would hence give very low results at this stage, 
-%, for the reasons stated.
-and a comparison of the results would not be of much value as
-there would not be enough material %underlag is yet too small 
-to do be able to do any interesting analysis.
-An estimation of the improvement can be given by looking at the results from
-running the test suite used for the grammar development.
-The sentences given in the test suite are short, four to 10 words, and the
-words are in most cases included in
-the lexicon, but there are also constructions that have not been implemented 
-during this project.
-The first grammar could parse less than half of the sentences, the result for the final
-grammar was 66 \%. 
-% try NPs?
-It is thus not yet interesting to talk about coverage, but about the quality and
-the ability to scale up, which has so far proved to be good.
-I further believe that the presence of a expert in Swedish, professor Elisabet
-Engdahl\footnote{http://svenska.gu.se/om-oss/personal/elisabet-engdahl},
-has
-increased the standard substantially.\\
-
-\noindent By the renewed import of Saldo, we have doubled the size of the lexicon and thereby
-added many of the commonly used words that were missing from the older
-version. This is of course a big improvement.
-However, the lexical part still requires much work before it can be made use of.
-We need valency information to make a good analysis. The lexicon is also too
-big to use with the current techniques. Its size causes the
-incremental parsing algorithm to use more heap memory than normal computers
-allow. To solve this, we need to use the lexicon data more cleverly.
-
-%Results from mapping, can probably be improved, bit by bit, depends much on words.
-%Valencys would therefore help.  
-%Saldo, made a big difference to renew the lexicon. Tested what was missing now from talbanken,
-%results. Mostly compound nouns, which we can't expect in the dictionary.
-%Is too slow, can't be used for parsing right away
-%
-%Grammar - hard to evaluate automatically, but Elibet is an expert who has been involved
-%in the process to verify the solutions. Testing trees against talbanken.
-%The big lexicon makes it very slow, eats all memory
-%
-%Discussion of the results and methods, why are the results good, why are they not better.
-%What other methods could have been used? What did we/I expect and what happend?
-
-\newpage
+
+Section 5.3 + the evaluation part from 6.1
+
+Latest results from chunk parsing and disambiguation
+
 \section{Future Work}
-\label{sec:future}
-At the end of the current project,
-%Having finished this part of the project,
-we are left with 
-%Although this one part of the project is finished, there are 
+%From 6.2 subtract the things that are already done, maybe more future
+%work related to the results for chunk parsing and disambiguation.
+
+abbrivations, idioms, learning adVs.
+\textbf{too much about grammar details? there should be other things here as well}
+At the end of the current project, we are left with 
 many interesting future directions.
-The future work described in this section is divided into two categories:
-the ones aiming at making the parser robust and the ones that can be
-seen as extensions of the work done so far.
-
-\subsection{Enhancements and improvements}
+
 \subsubsection{Grammar}
 The grammar should cover the most prominent Swedish features.
-%All imaginable grammatical constructs cannot be cover by our grammar.
 Even as some work must be left for the robustness, there are specific constructs
@@
 
-%Sentences like
-%'hon är inte dum, hon.' 'hon är inte dum, flickan' 'min syster, hon är inte dum, hon'
-%'min syster är inte dum, hon'
-%We have a clause of which we need to know the natural gender and number of the subject, since it
-%must agree with the added pronoun. sag s 439. fullständigt outforskade!
-
-%or just implement this and remove it here?
-%The question is how many types we should give dependent types. CN could also need it,
-%in order to say
-%'han mäter avståndet från sitt hus till sitt jobb' but now `avståndet från sitt hus är stort'.
-%The analysis of this sentence is that
-%avståndet is a noun of type \verb-N3-, which chooses the prepositions `från' and `till'
-%and needs to noun phrases to form a \verb-CN- -- ``avståndet från sitt hus till".
-%Now also \verb-CN- needs to carry about whether it is possible to use it as a subject.
-
 \item
-\textit{Bare indefinites} are at this point always parsed as mass nouns.
+\textit{Bare indefinites} are at this point always parsed as mass nouns
     \enumsentence{\shortexnt{3}
@@
 }\label{sent:water}
-The parse tree of sentence (\ref{sent:water}) looks as follows:
-\begin{verbatim}
-PredVP (UsePron i_Pron) (ComplSlash (SlashV2a drink_V2) 
-                                      (MassNP (UseN water_N))) 
-\end{verbatim}
-
-This however is not the correct analysis for `hund' (\emph{`dog'}) in sentence (\ref{sent:hund}).
+although this clearly is a unsatisfying analysis for bare indefinites as in (\ref{sent:hund})
 \enumsentence{\shortex{4}
@@
 {``We are going to buy a dog"}}\label{sent:hund}
+\vspace{3mm}
 
 \item
-The implementation of reflexive pronouns can be improved.  %both code-wise
+\textit{Object and subject control verbs}
+The implementation of reflexive pronouns can be improved.  
 It should for example be possible to differentiate between 
@@
 {``He promises her to clean his room"}}
-
-%
-%\verb| X : NP x + CN + NP --better|\\
-%\verb| X : CN x + CN + CN|
-%check apposCN.
-%\textbf{måttsattribut}
-%\enumsentence{
-%{Antalet bilar, ett antal bilar, ett stort antal}}
-%Genitive; Not like 'sonen till mig' -> antalet bilars??\\
-
 \end{itemize}
-\subsubsection{Lexicon}
-Our lexicon still lacks some parts that can be imported from Saldo.
-The multiple-word verbs (\verb-vbm-) and multiple-word nouns (\verb-nm-) should
-be imported to an idiom lexicon, which can be used as an extension to the main
-lexicon.
-For the words that we tried but failed to import, another tool for lexicon
-acquisition could be used. The tool developed in the previous part of
-this project\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf} would be suitable.
-All in all, it should be ensured that we use as much information as
-possible from our resources.
-%The list of not-imported can be used for adding other kinds of lexicons - idioms, have POS-tags.
-%Also to analyse what's wrong, or use the words with a word guesser. The code
-%has been adapted for this.
-%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
-%category. How this works, look at tags. Small amount, can be done manually.
-
-\subsection{Making the parser robust}
-To achieve our goal of parsing unrestricted text, we need to use statistical methods.
-The parser needs to be able to handle
-ellipses and long and complex sentences, names and numbers,
-compound words and other unknown constructions.
-
-\subsubsection{Parallel parsing/Chunk parsing}
-Instead of parsing a whole sentence in one go, we intend
-parse parts of it in parallel and then try to combine the pieces, building
-the sentence bottom up.
-This method should be faster then normal parsing and would
-at the same time give robustness;
-components that cannot be combined by any grammar rule are combined by a 
-meta function. When an unknown construction is encountered, the parser
-will not fail but give a partial parse tree with the analysis for the rest of
-the sentence \cite{gfMech}.
-The output given to the user will be more valuable, since partial parse
-trees are more interesting the no parse trees.
-%We might end up with partial parse trees, but we would
-%be able to parse at least parts of every sentence and give a more
-%interesting result to the user.
 
@@
 \label{sec:futureValency}
-No matter how much we increase the size of our lexicon, we will
-never cover all compounds. We therefore need to be able to do
-compounding analysis. Saldo has tools for this, which might be usable
-in this project as well \cite{fm}.
-As noted, we additionally need to come up with better methods for using the lexicon when
-parsing, since its time and memory consumption is very high.
-One possibility is of course to do 
-deeper refinements of the parsing source code, an extensive work which is far
-outside the scope for this project. Other solutions are to either
-adapt the lexicon automatically for its domain and hence making it smaller and
-faster, or to adapt the input sentences to a smaller lexicon by preprocessing.
-
-Further, the lexicon needs valencies since this information lays the
-foundation for the GF analysis. 
-During the mapping, described in section \ref{sec:Mapping}, we analysed how Talbanken
-annotates valency information. It should be possible to %extract data from %and by using the rules made for the mapping
-extract data from Talbanken,
-showing how words are normally used.
-Lexin\footnote{http://spraakbanken.gu.se/lexin/} provides list of valencies information,
-from which lexical information could be extracted, given that the data is freely available.
-This is also a source of information helping us separate between verbs that can
-be used without an infinitive marker from the ones that cannot, cf. (\ref{sent:complVV2}a-b).
-However, Lexin is relatively small and does not contain more than 10 000 entries.
-%Furthermore, we also need be algorithms for
-%automatically adding the information to the lexicon. \\
-
+\begin{itemize}
+\item When using GF for parsing with lexicon of our size, we currently
+suffer from memory usage problems. The parsing algorithm is not fully
+capable of evaluating all possible parse trees, which is necessary in order
+to find the best one.
+
+\item
 Many Swedish verbs can be used with different number of arguments.
 Having one lexical entry for every possible usage of a verb
-%all this information to the lexicon 
 does not seem to be a good idea considering
@@
 implemented by using external resources.
-% TODO like what?
-
-%how to get valency information 'sitter och läser',
-%is this doable for swedish?
-%description of extracting this from talbanken/korp
-%or handling it by the robust parser or the grammar itself (leaveout obj?).
-%make the parser faster by using smaller lexicon.
-%Lexicon tool on the fly.
-
-\subsubsection{Probabilities}
-\label{sec:futureProbabilities}
-The grammar will always contain ambiguities and before returning the result
-to the user, some analysis should be done to find the most probable tree.
-When the size of the lexicon
-increases, so will the ambiguities as the number of word forms overlapping each other
-gets higher. Our new grammar also have many more rules which contribute
-to increasing the number of interpretations of each sentence.
-
-GF already allows probabilities, and by using a large
-GF treebank, we can get more reliable statistics for our functions. 
-By implementing dependency probabilities, our results would be even better.
-
-%TODO
-%The many word forms in the gF tables, as seen in section \ref{saldo}.
-%This is a possible source for overgeneration.
-\subsubsection{Named entity recognition}
-The parser should be able to identify names of persons, companies, countries etc.
-This can be done by the Haskell program executing the parsing \cite{patent}.
-
-% discuss ambiguites, will always have some and the big lexicon will increase them.
-% exapmle 'vad händer under äktenskapet?'
-%Getting and using probabilities, to disambiguate.
-% ontologies?
-
-%Use multiligual treebank
+
+\end{itemize}
 
 \section{Conclusion}
-%>> why is this significant, other grammars can use this Swedish grammar.
-%>> robustness in in future work leads to coverage.
-%>> reusablilty, freely available.
-%TODO
-%>> well known (comp. ling) trade-off between depth and coverage. for now depth,
-%>> have experiments of how to extract info from talbanken.\\
-
-We have developed the main components
-for a deep Swedish parser; an extended grammar and lexicon and material for
-evaluation and disambiguation. %To abstract:  by combining resources.
-By starting from the GF resource grammar, we got a well-defined system for
-describing language. 
-%as we built the grammar from .
-
-%When parser is derived from the grammar and will therefore validate that
-%a given sentence is grammatically correct according to the rules defined.
-
-Our final goal is to be able to parse unrestricted text, but considering that
-no syntactic theory is yet capable
-of wholly covering a natural language, we are satisfied with %theconsider our grammar 
-%when used without a robust layer 
-the grammar implementation of an important fragment of Swedish and have
-focused on constructions that are frequent in Talbanken. 
-Being a natural language processing-project, it will probably never be
-entirely complete, but by combining the rule-based parser with statistical
-techniques, 
-%By following the methods for making the parser robust 
-as described in section \ref{sec:future},
-we still believe that it is possible to achieve our goal. 
-\\
-
-%TODO more here
-%. Even so, we believe  
-%Even though the grammar will not be totally covering, we belive
-%that the work done is both interesting and useful.
-\noindent All parts of the project are open-source and may thus be used in other applications.
-The grammar and the lexicon may be beneficial also when working with controlled languages,
-as it increases the coverage of the Swedish resource grammar.
-
-The constructions
-that we have focused on have all been possible to implement, with varying  amounts
-of work. Many of them could be done by utilizing and extending the resource library
-but in some cases we needed to part from the multilingual abstract and use other
-grammatical theories in order to arrive at a good analysis.
-
-%There is obviously a lot do still, will never get finish with NPL.
-%Grammar may need even bigger changes, to allow enough but not too much.
-%Different from writing a grammar for generation.
-%Results from mapping shows that the translation is doable, but becomes harder
-%since talbanken don't have formal rules.
-%The saldo shows how easy to make use of utomstående resources, promising if
-%we want to extract other information. 
-%Hard, but not finished! 
-\begin{changemargin}{-0.1cm}{0cm}
-\bibliographystyle{apalikeurl}
-\bibliography{masterbib}
-\end{changemargin}
+%From 6.1 - the discussion part + something about the latest work
+
+The project has resulted in
+\begin{itemize}
+\item an extended grammar covering an important part of Swedish
+\item a comparison between GF and another annotation
+\item a deeper testing of the Swedish resource grammar and an estimation
+of how well GF can be used to describe larger parts of a language
+\item a study of how dependent types can be used in the resource grammars
+\end{itemize}
+
+The grammar has been extended and enhanced, and its current status is
+a specialized extension of the resource grammar.
+Besides parsing, the grammar may well be used for language generation,
+which works fast even when using an extensive lexicon.
+Although it is not been formally verified, we believe that the majority of the
+sentences generated are grammatically correct in a syntactical point of view.
+Without any semantic knowledge, nonsense phrases cannot be avoided in
+random generation. However, given that the abstract tree has a meaningful
+interpretation,
+the linearization should be correct. \\
+
+Here we have some nice new results... (todo..)
+
+\bibliographystyle{splncs}
+\bibliography{FinBib}
+
 \end{document}

### 2012-03-20 robust/chunk/Eval.hs
chunking: chunk info added

@@ -9,6 +9,8 @@ import Data.Maybe
 import qualified Format as Form
 import System.IO
 
+import qualified NpChunk as NP
+
 type Id = String
 data Result = Result {idN :: Id, isText :: Bool, total :: Bool, mets :: [String], first :: String
                      ,tooBig :: Bool}
@@ -19,12 +21,12 @@ tb = "../../Talbanken05_20060604/FPS/P.tiger.xml"
 -- counts the number of sentences for which we could parse the structure, how many that were
 -- totally parsed and how many metas that were in them
 main = do
-   inp <- readFile "helaTBAdv"
+   inp <- readFile "night/advTest1Res" --"helaTBAdv"
    let ss   = splitWhen null $ lines inp
        oks  = map mkLine $ filter (not . null) ss
        res  = (length oks,length (filter isText oks),length (filter (\x -> isText x && total x) oks)
-              ,map (head &&& length) $ group $ sort $ concatMap mets oks)
-   appendFile "Results2" $ show res ++"\n"++ show oks
+              ,length (filter tooBig oks),map (head &&& length) $ group $ sort $ concatMap mets oks)
+   appendFile "night/ResultsAdv1" $ show res ++"\n"++ show oks 
 
 isOk l = okStart l -- || hasChunk l
 whole x = not $ "Meta" `isInfixOf` x
@@ -33,10 +35,18 @@ whole x = not $ "Meta" `isInfixOf` x
 mkLine :: [String] -> Result  
 mkLine (i:xs) = let nice     = any isOk xs
                     semiNice = any hasChunk xs
-                    which    = if not nice && semiNice then last else head
+                    which    = if not nice && semiNice then (firstS . last) else head
                 in  Result i nice (whole $ which xs) (if nice then metas $ which xs else [])
-                           (which xs) semiNice
+                           (which xs) (not nice && semiNice)
  where metas         = filter ("Meta" `isInfixOf`)  . words
+   
+       firstS :: String -> String
+       firstS string = let (start,(p:rest)) = span (/='(') string
+                       in start++(p:takeParent "" rest 1)
+          where takeParent str rest 0 = str
+                takeParent str (x:xs) n | x=='('    = takeParent (str++[x]) xs (n+1)
+                                        | x==')'    = takeParent (str++[x]) xs (n-1)
[…]

### 2012-03-20 robust/chunk/MkLex.hs

@@ -26,7 +26,7 @@ mkDir lexs lem i = do
    putStrLn $ "compilation: "++ show t'
    case ex of
         ExitSuccess   -> do putStrLn "Done"
-                            let thisPGF = "pgfs/"++"BigParse"++show i++".pgf"
+                            let thisPGF = "pgfs2/"++"BigParse"++show i++".pgf"
                             x <- readProcess "mv" ["BigParse.pgf",thisPGF] []
                             putStrLn $ "Moved the pgf "++x
                             return   $ Just (thisPGF,newLang)

### 2012-03-20 robust/chunk/NewerTest.hs

@@ -19,7 +19,7 @@ import System.IO
 import System.TimeIt
 
 main = do hSetBuffering stdout LineBuffering
-          timeIt $ tryAll  "hardtests.xml" {- tb --try "stest.xml"-}
+          timeIt $ tryAll tb {-tb "hardtests.xml" --try "stest.xml"-}
 
 tb = "../../Talbanken05_20060604/FPS/P.tiger.xml" 
 
@@ -33,17 +33,23 @@ tryAll fil = do
   putStrLn $ "Building morpho ... "
   let morpho = buildMorpho pgf langBig
   putStrLn $ "Parsing xml ... "
-  input <- take 43 . drop 190 .{- take 22 .-} splitFile . concat <$> Form.parseIdTree fil
+  input <- {-take 43 . drop 190 . -}take 22 . splitFile . concat <$> Form.parseIdTree fil
   pgfs <- mapM (extractDicts saldo morpho) [x | x <- zip input [0..]] 
   mapM processPartition pgfs
 
  where processAndWrite :: PGF -> Language -> (String,ChunkTree) -> IO ()
        processAndWrite pgf lang (i,tree) = do
           res <- parseText tree pgf lang startType
-          let fil = "night/advTest2"
+          let fil = "night/advTest1Res"
           putStrLn $ "write to "++fil
           appendFile fil $ showRes (i,res) ++"\n"
         where ziptree = getFirstWord $ fromTree tree
+
+       chunkAndWrite :: PGF -> Language -> (String,ChunkTree) -> IO ChunkInfo
+       chunkAndWrite  pgf lang (i,tree) = do
+          res <- countChunks tree pgf lang startType
+          appendFile "night/chunkingAllPart1" $ show res
+          return res
 
        extractDicts :: Saldo -> Morpho -> ([(String,ChunkTree)],Int) -> IO ([String],[WorkingTree],Maybe (FilePath,Language))
        extractDicts saldo morpho (input,i) = do
@@ -63,7 +69,10 @@ tryAll fil = do
           probs  <- readProbabilitiesFromFile "ProbsNewSpec" newPgf
           let goodPgf = setProbabilities probs newPgf
           putStrLn $ "Parsing the tree "
-          mapM_ (processAndWrite goodPgf newLang) $ zip ids (map toTree newsTrees)
+--          mapM_ (processAndWrite goodPgf newLang) $ zip ids (map toTree newsTrees)
+          
[…]

### 2012-03-20 robust/chunk/NewestChunk.hs

@@ -31,15 +31,27 @@ data PState = PS {currentStates :: [String]       --the current parse state
                  ,chunks        :: [(Id,[Expr])]  --list of expressions to replace the ints with later
                  ,morpho        :: Morpho
                  ,saveChunks    :: [(Id,[Expr])]  
-                 ,counter       :: Int             --counter for replacing
                  ,lastExpr      :: Maybe (Id,[Expr])
-                 ,npRate        :: (Int,Int) --NPInfo
---                 ,splitChunks   :: [[String]]
---                 ,backups       :: 
+                 ,npRate        :: (Int,Int,[Int]) --NPInfo
+                 ,npCover       :: (Int,Int,[Int])
+                 ,apRate        :: (Int,Int,[Int]) --NPInfo
+                 ,apCover       :: (Int,Int,[Int])
+                 ,advRate       :: (Int,Int,[Int]) --NPInfo
+                 ,advCover      :: (Int,Int,[Int])
                  }
 type Parser = WriterT [String] (State PState)
 type Id     = String
---data NPInfo = NP {tot :: Int, ok :: Int, good :: Int, sents :: [String]}
+data Info   = I {ok, tot :: Int, okLength :: [Int]}
+    deriving Show
+data ChunkInfo =  CI {npR :: Info, apR :: Info, advR :: Info
+                     ,npC :: Info, apC :: Info, advC :: Info}
+    deriving Show
+
+chunkInfo :: PState -> ChunkInfo
+chunkInfo st = CI (info $ npRate st)  (info $ apRate st)  (info $ advRate st)
+                  (info $ npCover st) (info $ apCover st) (info $ advCover st)
+    where info (i,j,l) = I i j l
+                        
 
 
 toplimit = 500
@@ -50,25 +62,25 @@ disambiglimit = 700
 backuplimit = 20
 saveStatelimit = 8 
 
+countChunks :: Tree (Id,String) -> PGF -> Language -> Type -> IO ChunkInfo
+countChunks tree pgf lang startType = do
+  hSetBuffering stdout LineBuffering
+  let st = execState (execWriterT (parseX tree)) parser 
+  return $ chunkInfo st
+ where parser     = initPState pgf lang startType
+
 parseText :: Tree (Id,String) -> PGF -> Language -> Type -> IO [Expr]
 parseText tree pgf lang startType = do
   hSetBuffering stdout LineBuffering
[…]

### 2012-03-20 tsd/TSD2012/paper.tex
tsd paper, evaluation

@@
 word order more intricate.
-It is a verb-second language \cite[p.116]{gunlog}: the
+It is a verb-second language: the
 second constituent of a declarative main clause must consist of a verb.
@@
 \textbf{Some more here}
-Our main lexical resource is SALDO \cite{}. We use the tools
-described in \cite{} to extract a large-scale lexicon and annotate
+Our main lexical resource is SALDO \cite{saldo}. We use the tools
+described in \cite{ahlbergEnache} to extract a large-scale lexicon and annotate
 this with valencies from Lexin. This gives us a dictionary with more than 
@@
 %2 trees  
-Why we do this...... when is meta used
+Talbanken contains much valuable information about phrase structures and word 
+usage. One part of this project has focused on translating trees from Talbanken to GF
+by constructing a mapping, which automatically transforms trees in the
+Talbanken format to GF abstract syntax trees.
+We hence get a comparison between the two annotations and at the same time
+we get methods for extracting and translating to GF notation.
+
 Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
@@
 
-The evaluation measures the numbers of \textit{metas} in each translated tree.
-If the finite verb is untranslatable, we get metas for the verb itself and for the
-tense.
-For all of Talbanken, we get a number of 65 \%, and if we limit our input to
-uncomplicated sentences with no more than 10 words where all words are known to
-our lexicon, we can reach 85 \%. When translating the whole treebank, 78 sentences
-could not be mapped at all and resulted in \textit{metas} only.
+If a node fails to be translated, a \textit{meta variable} is in its place. This is also the
+case when two sister nodes can be translated, but no rule is applicable for joining them
+into one tree.
 
 \section{Chunk Parsing}
-
+\label{sec:chunk}
 %New section - insights about latest work
@@
 possible expressions of the language.
-Hence, we combine it with statistical means to get good coverage. Our approach
-makes use of the rich annotation in Talbanken -- we perform chunk parsing
+Hence, we combine it with statistical means to get good coverage. 
+The parser uses named entity recognition and a compounding analysis.
+Further we need to be able to give output
+even when encountering unknown words and grammatical constructions 
+as well as idioms or ellipses.
+Our approach makes use of the rich annotation in Talbanken -- we perform chunk parsing
 relying on some of the tags given in the treebank. 
-Unknown words and grammatical constructions, idioms or ellipses cannot be
-handled from within the grammar.
 Therefore, whenever a sentence is not recognized by the grammar, it is 
 parsed chunk by chunk.
-Noun phrases, adverbial phrases and ? are considered. At the first stage,
+Noun phrases, adverbial phrases and adjectival phrases are considered. Due to 
+differences in the annotation between Talbanken and GF, verbs are not considered
+as parsable chunks. The verbs need their particles and preposition to be parsed
+correctly, and these are usually not in the same Talbanken chunk. Therefore,
+we focus on whole sentences and noun phrases. At the first stage,
 we try to parse the basic sentence structure, and therefore allow the parser
@@
 
-This way we get good coverage -- from ca 10 \% of Talbanken -- to ??
+%This way we get good coverage -- from ca 10 \% of Talbanken -- to ??
 
 \section{Language Model-Based Disambiguation}
-
- Maybe move this to future work, we have it but not really sure
- if it is actually used atm.
 
@@
 
-This model is used for disambiguation: the parse trees for each sentence
+The model is used for disambiguation: the parse trees for each sentence
 are ranked by the sum of the probability of each function used.
+Both whole sentence and chunks -- phrases -- can be ranked using this
+principle.
 
@@
 
-Section 5.3 + the evaluation part from 6.1
+For the extraction of a GF Treebank from Talbanken, (section \ref{sec:mapping})
+the evaluation measures the numbers of \textit{metas} in each translated tree.
+If the finite verb is untranslatable, we get metas for the verb itself and for the
+tense.
+For all of Talbanken, we get a number of 65 \%, and if we limit our input to
+uncomplicated sentences with no more than 10 words where all words are known to
+our lexicon, we can reach 85 \%. When translating the whole treebank, 78 sentences
+could not be mapped at all and resulted in \textit{metas} only.
 
-Latest results from chunk parsing and disambiguation
+A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
+The percentage of restored nodes from Peen Treebank is higher than our results.
+The reason for this may be the fact that English is syntactically less complicated
+than Swedish.
+Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
+where idiomatic expressions are more likely the be appear and the language
+presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
+Also, the Penn Treebank contains a lower number of tags, 82 compared
+to more than 300 in Talbanken. With more tags, we get more information, but as the number 
+increase, so does the amount of work of finding the correct translation 
+for each combination of tags and writing rules that cover all constructions. \\
+
+The evaluation of the parser is performed by testing it on Talbanken sentences
+and is still being done. So far, our evaluation results cover 643 sentences -- about 10 \%
+of the treebank. Without the robustness given by the chunk parser,
+we can only parse 17.9 \% of the sentences. For noun phrases, we cover 76 \%.
+By using the chunk parsing described in section
+\ref{sec:chunk}, we identify the structure on 65 \% of the sentences. That
+is, the parser can identify the verbs and how they relate to the noun phrases,
+prepositions and particles. The noun phrases and adverbs, however, may not be
+completely parsed. If a verb is unknown to the grammar, or if it is used with
+another valency, other prepositions or particles than the lexicon has assigned
+it, the sentence structure can not be identified. In that case
+the parser returns chunks, showing the parse trees of identifiable sub-phrases.
+
+%Section 5.3 + the evaluation part from 6.1 
+%
+%Latest results from chunk parsing and disambiguation
 
@@
 
-abbrivations, idioms, learning adVs.
-\textbf{too much about grammar details? there should be other things here as well}
 At the end of the current project, we are left with 
 many interesting future directions. 
+We should make sure to make as much use of resources as possible -- Saldo
+and Lexin provides more information than we so far have been able to extract.
+For example, data about abbreviations and idioms should
+be added to the parser. This could be extracted from Saldo as well as Talbanken.
+As the valency information is so important to the GF grammar, the lexicon should
+be annotated with more of this in order to get a good parser.
 
@@
 \end{tabular}}
-Personal pronouns are allowed to precede negation and some other sentence adverbials % --\textit{satsadverbial} --
-in main clauses without auxiliary verbs.
-
-\enumsentence{\begin{tabular}[t]{@{}*{12}{l@{\ }}}
-a. & *Vi & har&  honom&  inte&sett&~~ & b.& * Jag &ser &huset &inte\\
-   & \emph{~we} & \emph{have}&  \emph{him}&  \emph{not} &  \emph{seen} && b.& \emph{~I}&\emph{see} &\emph{the house} &\emph{not}\\
-\end{tabular}}
-
-Although object shifts are frequently used, they are hardly found in
-Talbanken's part P, which has been the inspiration for this project.
-Therefore, this implementation has so far not been prioritized.\\
+%Personal pronouns are allowed to precede negation and some other sentence adverbials % --\textit{satsadverbial} --
+%in main clauses without auxiliary verbs.
+%
+%\enumsentence{\begin{tabular}[t]{@{}*{12}{l@{\ }}}
+%a. & *Vi & har&  honom&  inte&sett&~~ & b.& * Jag &ser &huset &inte\\
+%   & \emph{~we} & \emph{have}&  \emph{him}&  \emph{not} &  \emph{seen} && b.& \emph{~I}&\emph{see} &\emph{the house} &\emph{not}\\
+%\end{tabular}}
+%
+%Although object shifts are frequently used, they are hardly found in
+%Talbanken's part P, which has been the inspiration for this project.
+%Therefore, this implementation has so far not been prioritized.\\
 
 \item
-\textit{Bare indefinites} are at this point always parsed as mass nouns
-    \enumsentence{\shortexnt{3}
-{Jag & dricker& vatten}
-{\emph{I}&\emph{drink}&\emph{water}}
-}\label{sent:water}
-although this clearly is a unsatisfying analysis for bare indefinites as in (\ref{sent:hund})
+\textit{Bare indefinites} 
+%are at this point always parsed as mass nouns
+%    \enumsentence{\shortexnt{3}
+%{Jag & dricker& vatten}
+%{\emph{I}&\emph{drink}&\emph{water}}
+%}\label{sent:water}
+%although this clearly is a unsatisfying analysis for bare indefinites as in (\ref{sent:hund})
 \enumsentence{\shortex{4}
@@
 \textit{Object and subject control verbs}
-The implementation of reflexive pronouns can be improved.  
-It should for example be possible to differentiate between 
-object and subject control verbs.
-
+%It should for example be possible to differentiate between 
+%object and subject control verbs.
+%
 \eenumsentence{
@@
 
+A thorougly evaluation is still to be done, this should be carried out by
+measuring the agreement between all parsed chunks and their respective
+translation from Talbanken. The evaluation should also be done manually.
+For this, we rely on an expert in Swedish.
+
 \section{Conclusion}
@@
 which works fast even when using an extensive lexicon.
-Although it is not been formally verified, we believe that the majority of the
-sentences generated are grammatically correct in a syntactical point of view.
-Without any semantic knowledge, nonsense phrases cannot be avoided in
-random generation. However, given that the abstract tree has a meaningful
-interpretation,
-the linearization should be correct. \\
 
-Here we have some nice new results... (todo..)
-
+Open source, transparent, modular, explores the strengths and weakneses
+of GF.
 \bibliographystyle{splncs}

### 2012-03-20 tsd/TSD2012/paper.tex
tsd paper 9 pages

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usepackage{tikz-qtree}
 
@@
 functional language, specialized for grammar programming. We trained
-and evaluated the grammar by using Talbanken, the largest treebank for 
+and evaluated the grammar by using Talbanken, one of the largest treebanks for 
 Swedish. As a result 65 $\%$ of the Talbanken trees were translated
-into the GF format in the training stage. Moreover, we obtained a language
+into the GF format in the training stage and 75 \% of the noun phrases
+was parsed during the evaluation. Moreover, we obtained a language
 model for Swedish which we use for disambiguation when parsing free text
@@
 
-\textbf{Ramona: Copied from the thesis, could be improved!}
-GF has strong support for multilinguality and has so far been used successfully for 
-controlled languages \cite{cnl}, while recent experiments have showed
-that it is also possible to use the framework for parsing free language \cite{patent}.
-So far there is no freely available grammar-driven parser that gives a deep
-analysis for Swedish. The fact that a parser or grammar is not freely available
-does not only restrict its usage but also its possibilities of being further
-developed. Our goal is
-to create an open-source Swedish grammar from which we derive a parser
-accepting all sentences described by the given rules.
-As a freely available resource, it can be continuously enhanced and
-made use of in other projects.\\
-\noindent To build the grammar from scratch would be not only time consuming but
-would also mean that already existing systems would have to be reimplemented.
-In order to not reinvent the wheel we proceed from a combination of well-tested
-sources.
-We start from a GF resource grammar, consisting of an abstract and a concrete syntax
-file defining a set of rules for morphology and syntax.
-This is what is meant by grammar in this thesis, as
-opposed to grammar in the traditional linguistic sense. 
-From GF, we get a well-defined system for describing language, as well as a
-strong connection to and possibility of translation between the 
-more than 20 other languages implemented in the framework. Further, we use
-the extensive lexicon Saldo and the treebank Talbanken.
+Two main approaches divide the research in computational linguistics
+in the last years. The first one is the wide-coverage method, usually based on
+statistics methods, which compensates its usually shallow analysis of the
+input with its coverage. The other approach is the opposite - provides
+a deep analysis of the data, but it's coverage is limited to
+very specialized input. Our work tries to unite the two approaches by
+developing a wide-coverage computational grammar for Swedish. We start
+with a minimal model of the language and extend its coverage
+significantly, without affecting the quality. The grammar was
+developed and evaluated against Talbanken\cite{talbanken}, one of the
+largest Swedish treebanks.
 
-Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
-into the GF notation and an extended Swedish grammar implementation. The grammar is based
-on the multilingual abstract syntax given in the GF resource library, and now
-also covers constructions specific to Swedish.
-We further give an example of the advantage of using dependent types when
-describing grammar and syntax, in this case for dealing with reflexive pronouns.
-Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
-It is based on Martin-Löf type theory which allows
-reasoning within the programming language.
-
-Our goal is a wide-coverage grammar for Swedish which -- combined with statistical 
-resources -- can be used for parsing open-domain text. The grammar implementation gives
-a deep analysis, in the trade-off between coverage and depth we aim for depth
-while getting coverage from statistical methods.\\
-
-Outline ...
+%The paper is organized as follows: .... \textbf{maybe leave out if no space}
 
 \section{Background}
+
+\subsection{GF}
+\label{sec:gf}
+
+The computational grammar was developed within GF(Grammatical
+Framework) \cite{gfbok}. GF is a dependently-typed grammar formalism,
+based on Martin-Löf type theory, which is mainly used for
+multilingual natural language applications.
+
+GF grammars are composed of an abstract syntax, interlingua of the
+grammar, describing the semantics on a language-independent level and
+a number of concrete syntaxes, usually corresponding to natural
+languages. These describe how the concepts from the abstract syntax
+are described in the language. 
+By defining such a grammar, one also obtains a parser for the language
+fragment that is described by a concrete grammar and a natural
+language generation tool that is able to generate constructions from
+the same language fragment. 
+In this way, due to the division of the GF
+grammars into abstract and concrete, it is possible to achieve 
+semantics-preserving translation between any pair of languages from 
+the grammar.
+
+However, because the grammar is strict, we can only deal with
+constructions that the grammar defines and nothing more. One could
+classify the limitations of the natural language grammar in two
+categories --- missing lexical items and missing syntactic
+structures. For this reasons GF was mainly used so far for controlled
+languages\cite{cnl} and, dialogues systems \cite{talk} and interactive 
+systems where the user is guided to stay
+within the bounds of the grammar by a predictive
+parser\cite{parser}, such as a multilingual tourist
+phrasebook\cite{phrasebook}. 
+However, there is some recent work on parsing
+free text --- a grammar for patent claims from the biomedical
+domain \cite{patent}. This work highlighted a
+number of caveats of real-world text, but in the same time gives
+valuable insights that future work on wide-coverage GF grammars can
+benefit from.
+
+The current work aims at overcoming the second limitation of GF
+grammars --- adding support for the most common syntactic
+constructions that could appear in Swedish texts. For some of these we
+use dependent types to encode the syntactic phenomena in an elegant
+way. The lexicon limitation is not a problem, as we use an existing 
+large Swedish lexicon \cite{ahlbergEnache}, which is being continuously
+developed and contains more than 120 000 entries.
+%... \textbf{Malin, can you say a bit more ?But not more than a couple
+%  of lines :)}
 
@@
 % Chapter 2.4(with emphasis on the syntactic structures)
-\textbf{Too long?}
 Swedish is a North-Germanic language, sharing most of its grammatical structure 
@@
 word order more intricate.
-It is a verb-second language: the
-second constituent of a declarative main clause must consist of a verb.
+It is a verb-second language: %\cite[p.116]{gunlog}: 
+the second constituent of a declarative main clause must consist of a verb.
 The normal word order is subject-verb-object, but fronting other constituents
@@
 Fronting the finite verb marks questions.
-
-Swedish has two ways of forming passive verb phrases: the 
-\textbf{periphrastic passive}, formed by using the modal auxiliary verb `bli' (\emph{`become'})
-and the \textbf{s-passive} which is formed by adding an \emph{s} to the verb: \\
-\enumsentence{Passive \hspace{45mm} Active\\
-\shortexm{12}
-{Tjuven & togs&  av & polisen}
-{\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
-{``The thief was arrested by the police"}
-{&&Polisen & tog & tjuven}
-{&&\emph{the police} & \emph{took} & \emph{the thief}}
-{\hspace{-9mm}``The police arrested the thief"}}
-\label{gfPass:s-pass}
-
+%\enumsentence{Passive \hspace{45mm} Active\\
+%\shortexm{12}
+%{Tjuven & togs&  av & polisen}
+%{\emph{the}\emph{ thief}&  \emph{took}+\textbf{s}&  \emph{by}&\emph{the police}}
+%{``The thief was arrested by the police"}
+%{&&Polisen & tog & tjuven}
+%{&&\emph{the police} & \emph{took} & \emph{the thief}}
+%{\hspace{-9mm}``The police arrested the thief"}}
+%\label{gfPass:s-pass}
 Special reflexive pronouns and reflexive possessive pronouns for the 3rd
 exist \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd person forms.
-\enumsentence{
-\begin{tabular}{llll}
-a. & Han slog \textbf{sig}.&\hspace{20mm}b. & Han såg \textbf{sitt} barn.\\
-&\emph{He hit \sc self.} && \emph{He saw {\sc self's} child.}
-\end{tabular}}
-\enumsentence{
-\begin{tabular}{llll}
-a. &Han slog \textbf{honom}. &b.& Han såg \textbf{hans} barn.\\
-&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
-\end{tabular}}
-
-\subsection{GF}
-\label{sec:gf}
-\textbf{Ramona: this is copied too, probably much room for improvement,
-and it also has to be shortened}
-%Chapter 2.1 + reference to Krasimir's work on the Penn Treebank
-%(+ 2.1.3), nothing about resource library here
-The main component of the project is the Grammatical Framework (GF) \cite{gfbok},
-a grammar formalism based on Martin-Löf type theory \cite{martinlof}. GF is
-designed for multilingual applications and represents a formalism
-stronger than mildly context-free grammars.
-
-GF is a strongly typed functional programming language, inspired by
-ML \\\cite{ml} and Haskell \cite{haskell}. It is also a logical framework,
-and the built-in functionality for logic and reasoning 
-is inspired by $\lambda$ Prolog \cite{prolog} and 
-by Agda \cite{agda}, with which GF also shares its type checking algorithm.
-The first version of the framework was implemented at Xerox Research Center
-in Grenoble and is now mainly developed in Gothenburg. One of the biggest
-achievements is a library covering the 
-basic morphological and syntactic structures of more than
-20 languages (see section \ref{sec:resources}).
-
-A grammar written in GF can be used for both parsing and generation.
-The parsing algorithm is incremental and has polynomial time and space
-complexity \cite{gfMech}. 
-The GF package also provides various tools for working with and using the grammars:
-a compiler, an interpreter and a runtime system.
-The grammars can be compiled into portable grammar format (PGF) \cite{pgf},
-supported by Java, Java script, C and Haskell libraries. 
-The interpreter, the GF shell, allows
-the user to test grammars by commands for parsing, visualization of parse trees,
-random generation, word alignment, morphological quizzes etc.
-The shell can be tried out online 
-together with an interactive GF editor 
-\footnote{http://www.grammaticalframework.org/demos/gfse/}.
-\thispagestyle{plain}
-\begin{figure}[h]
-\begin{verbatim} 
-> parse "jag ser katten" | visualize_tree  -format=pdf -view=evince
-> parse "jag ser katten" | visualize_parse -format=pdf -view=evince
-\end{verbatim}
-\caption{Example of how the GF shell can be used.}\label{fig:shellvp}
-\label{fig:shell}
-\end{figure}
-Figure \ref{fig:shell} shows commands for parsing, where the results are shown as
-in figure \ref{fig:gftree1}. 
-GF uses Graphviz\footnote{http://www.graphviz.org/}
-to visualize the trees and the given commands (fig. \ref{fig:shell}) specify that the
-output format should be PDF and that these files should be opened by the program
-Evince. The user can choose to see the \textit{parse tree} (fig. \ref{pic:gfStree})
-or the \textit{abstract tree} (fig. \ref{pic:gfAtree}). Abstract trees are more verbose and show all
-functions and types used for parsing the sentence. 
-\begin{figure}[!h]
-\centering
-\subfloat[Swedish parse tree]
-{\label{pic:gfStree}\includegraphics[width=30mm]{gfSTreec.pdf}}
-\label{fig:gftree1}
-\subfloat[English parse tree]
-{\includegraphics[width=30mm]{gfETreec.pdf}
-\label{pic:gfEtree}}
-\subfloat[GF abstract tree]
-{\label{pic:gfAtree}\includegraphics[width=30mm]{gfTree.pdf}}
-\caption{Abstract tree and parse tree for the sentence ``Jag ser katten".}
-\label{fig:gftree1}
-\end{figure}
-Parse trees on the other hand show only the types assigned to the words and phrases.
-Information about tense, polarity etc., which are explicitly given in the abstract tree, are
-not reproduced in the parse tree. Hence, parse trees do not give complete representations
-but model the parse results in a transparent manner.
-
-The scope of GF grammars has so far been controlled language, a subset of a
-natural language used for a restricted domain. By
-restricting the coverage, the number of ambiguities
-can be limited and it can be ensured that the semantics is preserved during
-translation. Inherited ambiguities as well as ambiguities arising from multilinguality will
-remain, but can be controlled more easily.
-
-The use of controlled language thus gives the possibility of good and reliable
-translation and GF has successfully been used for this in a number of projects.
-WebAlt \cite{webalt} aims to develop language independent
-material for mathematics problems by using GF grammars.
-The formal software verification tool
-KeY \cite{key} used GF for translating formal specification to
-natural language.
-GF has further been used for describing dialogue grammars \cite{talk}, and
-the framework is one of the main component in the
-European project MOLTO\footnote{http://www.molto-project.eu/}, which develop
-online translation between to 15 languages.\\
-
-As an open source-project, GF is constantly being developed and improved. New
-languages are added, the compiler is being improved, ways of using it in more 
-efficient and easy-going manners are added
-and the possibilities to use GF in different environments
-increased. There is research on how to make more use of the dependent
-types, for reasoning by using ontologies \cite{ontologies2} or generating natural
-language via Montague semantics \cite{montague}.
+%\enumsentence{
+%\begin{tabular}{llll}
+%a. & Han slog \textbf{sig}.&\hspace{6mm}b. & Han såg \textbf{sitt} barn.\\
+%&\emph{He hit \sc self.} && \emph{He saw {\sc self's} child.}\\
+%c. &Han slog \textbf{honom}. &\hspace{6mm}d.& Han såg \textbf{hans} barn.\\
+%&\emph{He hit him (another person).}&&\emph{He saw his (another person's) child.}
+%\end{tabular}}
 
@@
 It is analyzed with the MAMBA annotation scheme (Teleman, 1974) and 
-consists of four
-parts. Two of them are transcriptions of spoken language, one a collection of
-text written by high school students, and one, section P,
+the more than 6000 sentences we have been using
+%consists of four
+%parts. Two of them are transcriptions of spoken language, one a collection of
+%text written by high school students, and one, section P,
 consists of professionally written Swedish gathered from newspapers, brochures and textbooks.
-\\
-\noindent Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
+
+Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
 and was then redistributed in an updated version,
 Talbanken05 \cite{talbanken05}.
-It is released in Malt\footnote{http://w3.msi.vxu.se/~nivre/research/MaltXML.html} 
-and Tiger\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html}
-XML-formats
-where the trees have been made deeper and more detailed while still containing
-the lexical MAMBA layer. 
-The Malt parser was trained on section P of Talbanken, and these
-more than 6000 sentences have been used our project. 
-The treebank has served as an inspiration and an evaluation source throughout the
-project.
+%It is released in Malt\footnote{http://w3.msi.vxu.se/~nivre/research/MaltXML.html} 
+%and Tiger\footnote{http://www.ims.uni-stuttgart.de/projekte/TIGER/TIGERSearch/doc/html/TigerXML.html}
+%XML-formats
+%where the trees have been made deeper and more detailed while still containing
+%the lexical MAMBA layer. 
+%The Malt parser was trained on section P of Talbanken, and these
+%more than 6000 sentences have been used our project. 
+%The treebank has served as an inspiration and an evaluation source throughout the
+%project.
 
 \section{Related Work}
-\textbf{Ramona: also just copied, if you want, improve it}
+%\textbf{Ramona: also just copied, if you want, improve it}
 %Chapter 2.5 - w/o the references about lexical resources
@@
 %between the authors
-There are many other parsers for Swedish,
-like the well-known data-driven Malt parser \cite{malt},
-which has been trained on Talbanken. 
-There are also a number of grammar-based parsers, although none is freely available.
-The cascaded finite state parser CassSwe \cite{casswe} and
-The Swedish Constraint Grammar \cite{birn}
-give syntactic analyses. 
-Swedish FDG (Voultanien,2001) uses the Functional Dependency Grammar
-\cite{fdg},
-an extension of the Constraint Grammar
-formalism, and produces a dependency structure focusing on finding the nominal
-arguments.
+There exist a number of parser for Swedish already, such as the
+data-driven Malt parser \cite{malt}, also trained on Talbanken, the 
+cascaded finite state parser CassSwe \cite{casswe}, % and
+The Swedish Constraint Grammar \cite{birn} and the Swedish FDG, which
+uses the Functional Dependency Grammar \cite{fdg}.
+%which gives syntactic analyses. \textbf{I don't really know if the
+  %last 2 do... could you rephrase please ?}
 
-The LinGO Grammar Matrix \cite{matrix}, is a starter-kit for building Head-Driven Phrase
-Structure Grammars \cite{hpsg} (HPSG) providing compatibility with tools for
-parsing, evaluation, semantic representations etc.
-Translation is supported by using Minimal Recursion
-Semantics \cite{mrs} as an interlingua. \\
-There is a collection of grammars implemented in this framework, giving broad-coverage
-descriptions of
-English, Japanese and German.
-The Scandinavian Grammar Matrix \cite{scandmatrix} covers common parts of
-Scandinavian, while Norsource \cite{norsource} describes Norwegian. A Swedish version
-was based upon this (SweCore, Ahrenberg) covering the morphology and some
-differences between Swedish and Norwegian. Further, there is the BiTSE 
-grammar \cite{stymne}, also implemented using the Lingo Matrix,
-which focuses on describing and translating verb frames.
+Among computational grammars for Swedish, we mention the Swedish
+version of the Core Language Engine \cite{gamback}, which
+provides a comprehensive description of syntax and semantics, as well
+as a translation to English. Unfortunately, the resource is not
+available for comparison. Other grammars are BiTSE \cite{stymne}, a
+Swedish grammar that uses the HPSG format \cite{hpsg} and developed
+within the LinGO Matrix library \cite{matrix}.
 
-The Swedish version of the Core Language Engine (CLE) \cite{gamback}
-gives a full syntactic analysis as well as semantics represented in `Quasi logical form'. A
-translation to English  was implemented and the work was further developed in the spoken
-language translator \cite{spoken}. Unfortunately, it is no longer available. The coverage of the Swedish
+The current work uses the extensive Swedish
+lexicon developed in \cite{ahlbergEnache} and the Swedish resource
+grammar as a starting point.
 
-In the TAG formalism \cite{tag}, there are projects on getting open source, wide-coverage grammars
-for English and Korean, but, to our knowledge, not for Swedish.
+Currently, GF is the leading technology in the European project
+MOLTO (Multilingual On-Line Translation)
+\footnote{http://www.molto-project.eu/}, which aims at developing
+ tools for translating between 15 languages in real-time and with high quality.
 
-The ParGram \cite{pargram} project aims at making wide coverage grammars using
-the Lexical Functional Grammar approach \cite{lfg}.
-The grammars are implemented in parallel in order to coordinate the analyses of
-different languages and there are now grammars for English, German, Japanese and Norwegian.
+Previous examples of larger type-theoretical GF grammars that use 
+dependent-types are SUMO-GF \cite{ontologies2}, a GF representation of
+SUMO \cite{sumo}, the largest open-source ontology and  a natural
+language generation grammar via Montague semantics \cite{montague}.
 
@@
 %Discussion about the two dimensions of a grammar - lexicon and syntax 
-There are to aspects of on a computational grammar clearly distinctive (?)  in
-our work- the work on the lexicon and the work regarding syntax.
-For this article, we will focus on the syntactical dimension, the description
-and formalization of grammar rules and the usage of Talbanken.
+There are two aspects of on a computational grammar -- the work on the lexicon
+and the work regarding syntax.  For this article, we will focus on the
+syntactical dimension of our work, the description and formalization of grammar
+rules and the usage of Talbanken.
 
@@
 words.
-Since the languages all share the same abstract syntax translation between any given
-pair is possible, which is valuable
-when implementing multilingual applications.
-
-The resource grammars describe how to construct phrases and sentences and how to
-decline words. The grammars cover the morphology,
+%Since the languages all share the same abstract syntax translation between any given
+%pair is possible, which is valuable
+%when implementing multilingual applications.
+The grammars describe how to construct phrases and sentences and how to
+decline words. They cover the morphology,
 word order, agreement, tense, basic conjunction etc.
-Due to the syntactic similarities of the Scandinavian languages, much of the implementation
-is shared with Norwegian and Danish. The modules that concern the lexical
-aspects are separate, while 85 \% of the syntax description is shared.
-There are about 80 functions, which describe the rules for building phrases and clauses.
+Due to the syntactic similarities between the Scandinavian languages, much of
+the implementation for Swedish is shared with Norwegian and Danish. The modules
+that concern the lexical aspects are separate, while 85 \% of the syntax
+description is shared.  There are about 80 functions, which describe the rules
+for building phrases and clauses.
 \begin{verbatim}
@@
 \end{verbatim}
-The analysis preformed by GF is driven by the parts of speech, which are combined into
-parts of sentences.
-In addition to the core resource grammars, which is shared with the other languages implemented in the
-library, there is also an extra module, simple called the \verb|Extra| module, where
- language specific constructions may bee added
-The functions given here do not have to be translatable
-to all other language, but are meant to cover language specific constructions.
+%The analysis preformed by GF is driven by the parts of speech, which are combined into
+%parts of sentences.
+In addition to the core resource grammars, which is shared with all other
+languages implemented in the library, there is also an extra module
+%, simple called the \verb|Extra| module,
+where language specific constructions may bee added
+%The functions given here do not have to be translatable
+%to all other language, but are meant to cover language specific constructions.
 
@@
 %We use the resources from the other paper, so to say :)
-\textbf{Some more here}
-Our main lexical resource is SALDO \cite{saldo}. We use the tools
-described in \cite{ahlbergEnache} to extract a large-scale lexicon and annotate
-this with valencies from Lexin. This gives us a dictionary with more than 
-100 000 entries, covering all but 500 words from Talbanken ignoring names,
-numbers and compounds.
+As mentioned before, our main lexical resource is SALDO \cite{saldo} from
+which a GF lexicon has been extracted \cite{ahlbergEnache}.
+Valency information, which is a key feature to good parsing using GF,
+are extracted from the lexicon
+Lexin\footnote{http://spraakbanken.gu.se/lexin/}.
+This gives us a dictionary with more than 100 000 entries, covering all but 500
+words from Talbanken (excluding names, compounds and numerical expressions).
 
@@
 implementation since there was no large resource available to evaluate it on.
-
 Our evaluations are based on Talbanken and when first conducting tests
  we found much room for improvement. The items listed below are examples
- of constructions implemented during this work.
+ of constructions implemented during this work.\\
 
-\textbf{shorten these sections}
-\subsubsection{The s-passive}
-Passive voice is often used in Swedish, especially the
-\textit{s-passive} as described in ....
-Some studies suggest that the s-passive is used in more than 80 \% of the times
+\noindent\textbf{The s-passive}
+Swedish has two ways of forming passive verb phrases: the 
+periphrastic passive, formed by using the modal auxiliary verb `bli'
+(\emph{`become'})
+and the s-passive which is formed by adding an \emph{s} to the verb. 
+Passive voice is often used, especially the
+s-passive, % (sentence \ref{gfPass:s-pass}).
+which is used in more than 80 \% of the times according to some studies 
+%Some studies suggest that the s-passive is 
 \cite{laanemets}.
@@
 where not all words have passive forms for all tenses. The resource grammar for
-Scandinavian therefore implemented the function for passive,
-\verb-PassV2-, by using auxiliary verb.
-The function allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
-turned into complete verb phrases; they no longer need an object.
+Scandinavian therefore only implemented the periphrastic passive.
+%\verb-PassV2-, by using auxiliary verb.
+%The function allows two-place verbs to be used in passive by using \emph{bli} (\emph{become}), and thereby
+%turned into complete verb phrases; they no longer need an object.
 During this project, the s-passive was added although the
 periphrastic passive is still allowed.
-The grammar further allows not only transitive verbs, but all verb phrases that
-misses an object, to form passives.
-A ditransitive verb -- `give' (\ref{sent:give2pass}) -- hence gives rise to two
-passives, (\ref{ex:passV32}) and (\ref{ex:passV33}).
-\enumsentence{\textbf{Active use of two-place verb}\\\shortex{4}
+%The grammar further allows not only transitive verbs, but 
+%All verb phrases that misses an object are allowed to form passives.
+A ditransitive verb -- \emph{ge} (`give') (\ref{sent:give2pass}) -- gives rise to two
+passives, (\ref{ex:passV33}) and (\ref{ex:passV32}), both covered by our new implementation.
+\eenumsentence{
+\item{\textbf{Active use of two-place verb}\\\shortex{4}
 {Vi & erbjöd& henne& jobbet }
@@
 \label{sent:give2pass}
-\enumsentence{\textbf{First place in two-place verbs}\\\shortex{3}
-{Hon & erbjöds & jobbet}
-{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}}
-{``She was offered the job"}}
+\item{\textbf{Passive use, first place \hspace{3mm}Passive use, second place}\\\shortex{7}
+{Hon & erbjöds & jobbet && \hspace{3mm}Jobbet & erbjöds & henne }
+{\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}
+&&  \hspace{3mm}\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
+{``She was offered the job"}
+~``The job was offered to her"}}
 \label{ex:passV33}
-\enumsentence{\textbf{Second place in two-place verbs}\\\shortex{3}
-{Jobbet & erbjöds & henne }
-{\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
-{``The job was offered to her"}}
-\label{ex:passV32}
+%\item{\textbf{Second place in two-place verbs}\\\shortex{3}
+%{Jobbet & erbjöds & henne }
+%{``The job was offered to her"}}
+\label{ex:passV32}}
 
@@
 {``There is a bird sitting on the roof''}}
-\emph{`Det'} has the position of the subject, and the real subject, 
-\emph{`en fågel'} the one of an object.
-
-Both the verb and the noun phrase have some restrictions: the determiner of the
+%\emph{`Det'} has the position of the subject, and the real subject, 
+%\emph{`en fågel'} the one of an object.
+Restrictions on both the verb and the noun phrase are covered by the grammar:
+the determiner of the
 noun phrase must be of such type that it requires both the noun and its modifiers
-to appear in indefinite form. The verb phrase must consists of a intransitive verb
+to appear in indefinite form. The verb phrase must consist of an intransitive verb
 or be in passive form.
-A very common example of impersonal constructions is sentences with the verbs \emph{finnas},
-(\emph{exist})
-\emph{saknas} (\emph{miss}) and \emph{fattas}(\emph{lack}).
-\enumsentence{\begin{tabular}[t]{@{}*{18}{l@{\ }}}
-a. &Det & finns & kaffe &&&
-b. &Det & saknas&  kaffe &&&
-c. &Det & fattas & kaffe \\
-&\emph{it }& \emph{exist }& \emph{coffee} && &
- &\emph{it }& \emph{misses} & \emph{coffee} && &
- &\emph{it} & \emph{lacks} & \emph{coffee}\\
-\end{tabular}\\
-{~~''There is coffee"~~~~ ''There is no coffee"~~~~ ''There is no coffee"}}
+%A very common example of impersonal constructions is sentences with the verbs \emph{finnas},
+%(\emph{exist})
+%\emph{saknas} (\emph{miss}) and \emph{fattas}(\emph{lack}).
+%\enumsentence{\begin{tabular}[t]{@{}*{18}{l@{\ }}}
+%a. &Det & finns & kaffe &&&
+%b. &Det & saknas&  kaffe &&&
+%c. &Det & fattas & kaffe \\
+%&\emph{it }& \emph{exist }& \emph{coffee} && &
+% &\emph{it }& \emph{misses} & \emph{coffee} && &
+% &\emph{it} & \emph{lacks} & \emph{coffee}\\
+%\end{tabular}\\
+%{\indent~~''There is coffee"~~~~ ''There is no coffee"~~~~ ''There is no coffee"}}
 
@@
 The reflexives require an antecedent with which they agree in
-number and person.
-The third person reflexives requires a third person antecedent and
-reflexive pronouns can not be used in subject noun phrases of finite sentences,
+number and person. They must 
+not be used in subject noun phrases of finite sentences,
 as shown by the ungrammatical examples in 
 sentence (\ref{ex:ReflGenB}) and (\ref{ex:ReflSubj}).
-Furthermore, the antecedent must be within the same finite sentence as the reflexive
-pronoun, see (\ref{sent:sesig}).
-
-Still our grammar should accept the sentences (\ref{ex:ReflGen}),
-(\ref{sent:rakasig}) and (\ref{ex:ReflAdv}) :
+%Furthermore, the antecedent must be within the same finite sentence as the reflexive
+%pronoun, see (\ref{sent:sesig}).
+Still our grammar should accept the sentences (\ref{ex:ReflGen})
+%(\ref{sent:rakasig}) 
+and (\ref{ex:ReflAdv}) :
 
@@
 \label{ex:ReflGenB}}
-\eenumsentence{
-\item{
-Jag vill att han rakar sig.\\
-I want him to shave {\sc self}.}
-\label{sent:rakasig}
-\item
-{*Han vill att jag ser på sig.\\
-~~He wants me to look at {\sc self}.}\label{sent:sesig}
-}
+%\eenumsentence{
+%\item{
+%Jag vill att han rakar sig.\\
+%I want him to shave {\sc self}.}
+%\label{sent:rakasig}
+%\item
+%{*Han vill att jag ser på sig.\\
+%~~He wants me to look at {\sc self}.}\label{sent:sesig}
+%}
 \eenumsentence{
@@
 POS-tags, information about semantic roles are given by the functions, not by the
-categories. That is, we know that the first argument of the function
-\verb-PredVP- acts as the subject, but the noun phrase itself does not carry
+categories. For example, the first argument of the function 
+\verb-PredVP- always acts as the subject, but the noun phrase itself does not carry
 information about its semantic role.
@@
 from other \verb-NP- on the type level since we would then need to duplicate
-our code, due to the rich type system in GF. Still, we need the type system to
-help preventing noun phrases containing reflexive pronouns to be used as subject.
+our code.
+%due to the rich type system in GF. 
+Still, the type system should prevent noun phrases containing reflexive
+pronouns from being used as subjects.
 This does not only concern noun phrases, but also adverbial and adjectival phrases 
-(\ref{ex:adjTyp} a,b).
-\eenumsentence{
-\item{
-taket på sitt hus\\
-\emph{the roof of {\sc self's} house}}
-\item{
-längre än sin kompis\\
-{\emph{taller than {\sc self's} friend}}}}
-\label{ex:adjTyp}
+(\ref{ex:ReflAP}).
+%\eenumsentence{
+%\item{
+%taket på sitt hus\\
+%\emph{the roof of {\sc self's} house}}
+%\item{
+%längre än sin kompis\\
+%{\emph{taller than {\sc self's} friend}}}}
+%\label{ex:adjTyp}
 
-The solution chosen introduce the use of dependent types.
-We make a difference between subjects and objects, but not
-by giving them entirely different types, but by 
-letting the type \verb-NP- \textit{depend} on an argument, which may either be \verb-Subject- or
+Our solution introduces the use of dependent types.
+We make a difference between subjects and objects
+%but not by giving them entirely different types, but
+by letting the type \verb-NP- depend on an argument, 
+which may either be \verb-Subject- or
 \verb-Object-.
@@
 ComplVPSlash : VPSlash    -> NP Object -> VP ; -- complementation
-
+PrepNP       : (a : NPType) -> Prep -> NP a -> Adv a ; 
                                            -- adverbial phrases 
-PrepNP      : (a : NPType) -> Prep -> NPTyped a -> AdvTyped a ; 
 \end{verbatim}
@@
 information we were missing.
-
-The dependent types in GF have been used for logical reasoning before but this 
-is the first example of using it in a grammar describing pure syntax of
-a natural language. Other languages having similar relations between 
-noun phrases and antecedent could be extended to use the type system in similar
-ways.
+%The dependent types in GF have been used for logical reasoning before, but
+This is the first example in GF making use of the dependent types for 
+describing pure syntax of a natural language. Our approach could be extended
+to describing similar relations between 
+noun phrases and antecedent in other languages.
+%having could be extended to use the type system in similar ways.
 
@@
 tested on Talbanken. For the function \verb-GenNP-, used for genitive, we managed
-to decrease the number of parse trees for some sentences to 10 \%.
+to decrease the number of parse trees for problematic, although short, sentences to 10 \%.
 % Could also be here:
@@
 %2 trees
+
 Talbanken contains much valuable information about phrase structures and word 
-usage. One part of this project has focused on translating trees from Talbanken to GF
+usage.
+One part of this project has focused on translating trees from Talbanken to GF
 by constructing a mapping, which automatically transforms trees in the
-Talbanken format to GF abstract syntax trees.
-We hence get a comparison between the two annotations and at the same time
-we get methods for extracting and translating to GF notation.
+Talbanken format to GF abstract syntax trees. We hence get a comparison 
+%\vspace{-9mm}
+\begin{figure}[!h]
+\centering
+\subfloat[Talbanken tree]
+{\begin{tikzpicture}
+\Tree [.S [.SS [.POOP Det ]]
+          [.FV [.AVPS \"ar ]]
+          [.NA [.ABNA inte ] ]
+          [.SP [.AJ l\"att ]]
+          [.IP [.IP . ] ]
+      ]
+\end{tikzpicture}
+}
+\subfloat[Translation to GF]
+{\includegraphics[width=60mm]{latt2.pdf}}
+\caption{The sentence \emph{Det är inte lätt} (``It is not easy"). If the word
+         \emph{lätt} (`easy') is unknown to the grammar, a meta variable is put
+         in its place.}
+\label{fig:gftree1}
+\vspace{-5mm}
+\end{figure}
+between the two annotations and at the same time we get methods for extracting
+and translating to GF notation. 
+If a node fails to be translated, a meta variable  --annotated \verb-?- -- is put in its place.
+This is also the case when two sister nodes can be translated, but no rule is
+applicable for joining them into one tree.
 
@@
 such as \verb|S|, \verb|NP| and \verb|VP|.
-The edge labels give the
-part of sentence: \verb|SS| for subject, 
-\verb|OO| for object etc. Nearly 400 different tags are used, of which more than
-300 different are POS-tags. The high number is due to the level of detail,  there are for
-example almost 50 tags for nouns, excluding proper names. The tags
-show definiteness, case and whether the word is a compound. Some words also have
-their own tags.\\
-
-\begin{tabular}{lll}
-\textbf{Tag} & \textbf{Example} & \textbf{Explanation} \\
-\hline
-\textbf{SV} & \emph{ska} (`shall') & The verb \emph{shall}\\
-\textbf{NNDDHSGG} & \emph{familjemedlemmarnas} & Noun, definite, compound (person),\\
-&(`the family members' ') &genitive \\
-\textbf{XX} & & Unclassifiable part-of-speech\\
-\end{tabular}\\
-
+The edge labels give the part of sentence: \verb|SS| for subject, 
+\verb|OO| for object etc. 
+%
+%\begin{tabular}{lll}
+%\textbf{Tag} & \textbf{Example} & \textbf{Explanation} \\
+%\hline
+%\textbf{SV} & \emph{ska} (`shall') & The verb \emph{shall}\\
+%\textbf{NNDDHSGG} & \emph{familjemedlemmarnas} & Noun, definite, compound (person),\\
+%&(`the family members' ') &genitive \\
+%\textbf{XX} & & Unclassifiable part-of-speech\\
+%\end{tabular}\\
 The mapping could in some cases be performed tag-by-tag, but annotational
@@
 In GF, the valency information is given for each entry in the lexicon.
-A sentence-complement verb has the type \verb-VS- and from this we know 
-that the function \verb-ComplVS- should be used to combine the verb and with a 
+A sentence-complement verb has the type \verb-VS- and must the 
+function \verb-ComplVS- must be used for combining the verb and with a 
 sentence.
-Another annotational difference that turned out to be problematic for the
-translation can be illustrated by analysis of the generic pronoun \emph{`man'}.
-In Talbanken, \emph{`man'} is simply a personal pronoun, 
-in GF it is represented by using the function \verb|GenericCl| which turns a
-\verb-VP- into a clause and has no subject visible in the abstract tree.
+%Another annotational difference that turned out to be problematic for the
+%translation can be illustrated by analysis of the generic pronoun \emph{`man'}.
+%In Talbanken, \emph{`man'} is simply a personal pronoun, 
+%in GF it is represented by using the function \verb|GenericCl| which turns a
+%\verb-VP- into a clause and has no subject visible in the abstract tree. 
+%When translating a sentence like this, it is thus not possible to simply glue the
+%subject to the verb with \verb|PredVP|. For each similar GF construction, an extra rule
+%would be needed to get full coverage.
 
-When translating a sentence like this, it is thus not possible to simply glue the
-subject to the verb with \verb|PredVP|. For each similar GF construction, an extra rule
-would be needed to get full coverage.
-
-The tags \textbf{XX} and \textbf{NAC} (\verb-Not a constituent-) are often used since Talbanken makes a
-difference between subject
-and object noun phrases.
-The analysis of elliptical expression in (\ref{sent:krav})
-\enumsentence{För stora krav.\\
-``Too high demands."}\label{sent:krav}
-contains both these tags, since it is not obvious
-whether the noun phrase is used as subject or an object.
-These tags occur quite frequently in the treebank and are always translated
-to metas, which lowers our result. \\
-
-If a node fails to be translated, a \textit{meta variable} is in its place. This is also the
-case when two sister nodes can be translated, but no rule is applicable for joining them
-into one tree.
+%The tags \textbf{XX} and \textbf{NAC} (\verb-Not a constituent-) are often used since Talbanken makes a
+%difference between subject
+%and object noun phrases.
+%The analysis of elliptical expression in (\ref{sent:krav})
+%\enumsentence{För stora krav.\\
+%``Too high demands."}\label{sent:krav}
+%contains both these tags, since it is not obvious
+%whether the noun phrase is used as subject or an object.
+%These tags occur quite frequently in the treebank and are always translated
+%to metas, which lowers our result. \\
 
@@
 Hence, we combine it with statistical means to get good coverage. 
-The parser uses named entity recognition and a compounding analysis.
-Further we need to be able to give output
+%The parser uses named entity recognition and a compounding analysis.
+The parser should give as much output as possible 
+%Further we need to be able to give output
 even when encountering unknown words and grammatical constructions 
-as well as idioms or ellipses.
+or idioms and ellipses.
 Our approach makes use of the rich annotation in Talbanken -- we perform chunk parsing
-relying on some of the tags given in the treebank. 
-Therefore, whenever a sentence is not recognized by the grammar, it is 
-parsed chunk by chunk.
-Noun phrases, adverbial phrases and adjectival phrases are considered. Due to 
-differences in the annotation between Talbanken and GF, verbs are not considered
-as parsable chunks. The verbs need their particles and preposition to be parsed
-correctly, and these are usually not in the same Talbanken chunk. Therefore,
-we focus on whole sentences and noun phrases. At the first stage,
+relying on the tags given in the treebank. 
+Consequentially, whenever a sentence is not recognized by the grammar, it is 
+parsed chunk by chunk and the parse trees for each chunk is returned.
+Noun phrases, adverbial phrases and adjectival phrases are considered.
+Due to the differences in the annotation between Talbanken and
+GF mentioned is section \ref{sec:mapping} verbs are not considered
+as parsable chunks. %The GF parser depend on requires verb particles and prepositions to give
+%a correct output, and which are usually not in the same Talbanken chunk. 
+Therefore, our focus is on whole sentences and noun phrases. At the first stage,
 we try to parse the basic sentence structure, and therefore allow the parser
@@
 6000 sentences. The translation process is explained in section
-\ref{sec:mapping} and even though most trees are incomplete, that is contain
-meta-variables, they still give valuable 
+\ref{sec:mapping} and the trees give valuable 
 information about which words from our lexicon that have been used
@@
 GF has built-in support for probabilities, and when feeding the output
-from the mapping to our grammar, we get a model of the language. We believe
-this data can be trusted as it is extracted from manually annotated 
-real-world sentences.
-
-The model is used for disambiguation: the parse trees for each sentence
-are ranked by the sum of the probability of each function used.
-Both whole sentence and chunks -- phrases -- can be ranked using this
-principle.
+from the mapping to our grammar, we get a model of the language. 
+After the parsing is completed, all trees are ranked measuring the
+probability of their constituents.
+As the data is extracted from manually annotated, real-world sentences,
+it constitutes a reliable source for disambiguation.
+%The model is used for disambiguation: the parse trees for each sentence
+%are ranked by the sum of the probability of each function used.
+%Both whole sentence and chunks -- phrases -- can be ranked using this
+%principle.
 
@@
 For the extraction of a GF Treebank from Talbanken, (section \ref{sec:mapping})
-the evaluation measures the numbers of \textit{metas} in each translated tree.
-If the finite verb is untranslatable, we get metas for the verb itself and for the
-tense.
+the evaluation measures the numbers of meta variables in each translated tree.
+%If the finite verb is untranslatable, we get meta variables for the verb itself and for the
+%tense.
 For all of Talbanken, we get a number of 65 \%, and if we limit our input to
 uncomplicated sentences with no more than 10 words where all words are known to
-our lexicon, we can reach 85 \%. When translating the whole treebank, 78 sentences
-could not be mapped at all and resulted in \textit{metas} only.
+our lexicon, we can reach 85 \%. 
+%When translating the whole treebank, only 78 sentences
+%could not be mapped at all and resulted in meta variables only.
 
-A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
-The percentage of restored nodes from Peen Treebank is higher than our results.
-The reason for this may be the fact that English is syntactically less complicated
-than Swedish.
-Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
-where idiomatic expressions are more likely the be appear and the language
-presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
-Also, the Penn Treebank contains a lower number of tags, 82 compared
-to more than 300 in Talbanken. With more tags, we get more information, but as the number 
-increase, so does the amount of work of finding the correct translation 
-for each combination of tags and writing rules that cover all constructions. \\
+%A mapping between GF and the Wall Street Journal of Penn Treebank has earlier been conducted \cite{gfMech}.
+%The percentage of restored nodes from Peen Treebank is higher than our results.
+%The reason for this may be the fact that English is syntactically less complicated
+%than Swedish.
+%Furthermore, the text in Talbanken are from various brochures, newspapers and text books,
+%where idiomatic expressions are more likely the be appear and the language
+%presumably less strict than in Wall Street Journal\footnote{www.wsj.com}.
+%Also, the Penn Treebank contains a lower number of tags, 82 compared
+%to more than 300 in Talbanken. With more tags, we get more information, but as the number 
+%increase, so does the amount of work of finding the correct translation 
+%for each combination of tags and writing rules that cover all constructions. \\
 
 The evaluation of the parser is performed by testing it on Talbanken sentences
-and is still being done. So far, our evaluation results cover 643 sentences -- about 10 \%
-of the treebank. Without the robustness given by the chunk parser,
-we can only parse 17.9 \% of the sentences. For noun phrases, we cover 76 \%.
-By using the chunk parsing described in section
+and is still being carried out. So far, our evaluation results cover
+600 sentences -- about 10 \% of the treebank. 
+%Without the robustness given by the chunk parser, we can only parse 17.9 \% of the sentences.
+For noun phrases, we cover 76 \%.  By using the chunk parsing described in section
 \ref{sec:chunk}, we identify the structure on 65 \% of the sentences. That
 is, the parser can identify the verbs and how they relate to the noun phrases,
-prepositions and particles. The noun phrases and adverbs, however, may not be
+prepositions and particles. Some subtrees may  however not be
 completely parsed. If a verb is unknown to the grammar, or if it is used with
@@
 
-At the end of the current project, we are left with 
-many interesting future directions. 
-We should make sure to make as much use of resources as possible -- Saldo
-and Lexin provides more information than we so far have been able to extract.
-For example, data about abbreviations and idioms should
-be added to the parser. This could be extracted from Saldo as well as Talbanken.
-As the valency information is so important to the GF grammar, the lexicon should
-be annotated with more of this in order to get a good parser.
+~~~~~~\textbf{Evaluation}
+A thoroughly evaluation is still to be done. This should be carried out by
+measuring the agreement between all parsed chunks and their respective
+translation from Talbanken. Evaluation should also be done manually.
+For this, we rely on an expert in Swedish.
 
-\subsubsection{Grammar}
+%\textbf{Using the resources}
+%We should make sure to make as much use of resources as possible -- Saldo,
+%Talbanken and Lexin provides more information than we so far have been able to
+%extract. For example, data about abbreviations and idioms should
+%be added to the parser. This could be extracted from Saldo as well as Talbanken.
+%As the valency information is so important to the GF grammar, the lexicon should
+%be annotated with more of this in order to get a good parser.
+
+\textbf{Grammar}
 The grammar should cover the most prominent Swedish features.
-Even as some work must be left for the robustness, there are specific constructs
-that are desirable and interesting to implement. A few examples are listed here:\\
-
-\begin{itemize}
-\item
-\textit{Pronominal object shift} is common in Swedish and obligatory in Danish and Norwegian.
-\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
-a. & Jag & ser&  honom& \textbf{inte}&~~ & b.&Jag &ser &\textbf{inte} &honom\\
-   & \emph{I} & \emph{see}&\emph{him}&\emph{not}&&& \emph{I} &\emph{see}&\emph{not}&\emph{him}\\
-\end{tabular}}
+While statistical methods compensates for constructions not covered by
+the grammar, we still aim for a even wider grammar coverage. 
+As examples of constructions to be added, we mention
+pronominal object shift, bare indefinites and a differentiation between
+object and subject control verbs.
+%\vspace{-2mm}
+%\begin{itemize}
+%\item
+%\textit{} % is common in Swedish and obligatory in Danish and Norwegian.
+%\enumsentence{\begin{tabular}[t]{@{}*{11}{l@{\ }}}
+%a. & Jag & ser&  honom& \textbf{inte}&~~ & b.&Jag &ser &\textbf{inte} &honom\\
+%   & \emph{I} & \emph{see}&\emph{him}&\emph{not}&&& \emph{I} &\emph{see}&\emph{not}&\emph{him}\\
+%\end{tabular}}
 %Personal pronouns are allowed to precede negation and some other sentence adverbials % --\textit{satsadverbial} --
@@
 
-\item
-\textit{Bare indefinites} 
+%\item
+%\textit{Bare indefinites} 
 %are at this point always parsed as mass nouns
@@
 %although this clearly is a unsatisfying analysis for bare indefinites as in (\ref{sent:hund})
-\enumsentence{\shortex{4}
-{Vi & ska& köpa & hund}
-{\emph{we} &\emph{will}& \emph{buy} & \emph{dog} }
-{``We are going to buy a dog"}}\label{sent:hund}
-\vspace{3mm}
+%\enumsentence{\shortex{4}
+%{Vi & ska& köpa & hund}
+%{\emph{we} &\emph{will}& \emph{buy} & \emph{dog} }
+%{``We are going to buy a dog"}}\label{sent:hund}
+%\vspace{3mm}
 
-\item
-\textit{Object and subject control verbs}
+%\item
+%\textit{Object and subject control verbs}
 %It should for example be possible to differentiate between 
@@
 %
-\eenumsentence{
-\item[a]\shortex{7}
-{Han & ber & henne & att & städa & sitt & rum}
-{\emph{he} & \emph{begs} & \emph{her} & \emph{to} & \emph{clean} & {\sc self's} & \emph{room}}
-{``He asks her to clean her room"}
-\item[b]\shortex{7}
-{Han& lovar& henne& att& städa &sitt & rum}
-{\emph{he}& \emph{promises}& \emph{her}& \emph{to}&  \emph{clean} &\sc self's&\emph{room}}
-{``He promises her to clean his room"}}
-\end{itemize}
+%\eenumsentence{
+%\item[a]\shortex{7}
+%{Han & ber & henne & att & städa & sitt & rum}
+%{\emph{he} & \emph{begs} & \emph{her} & \emph{to} & \emph{clean} & {\sc self's} & \emph{room}}
+%{``He asks her to clean her room"}
+%\item[b]\shortex{7}
+%{Han& lovar& henne& att& städa &sitt & rum}
+%{\emph{he}& \emph{promises}& \emph{her}& \emph{to}&  \emph{clean} &\sc self's&\emph{room}}
+%{``He promises her to clean his room"}}
+%\end{itemize}
 
-\subsubsection{Lexicon}
-\label{sec:futureValency}
-\begin{itemize}
-\item When using GF for parsing with lexicon of our size, we currently
+\textbf{Improving memory and time consumption}
+When using GF for parsing with lexicon of our size, we currently
 suffer from memory usage problems. The parsing algorithm is not fully
@@
 
-\item
-Many Swedish verbs can be used with different number of arguments.
-Having one lexical entry for every possible usage of a verb
-does not seem to be a good idea considering
-the ambiguities it will lead to and the already high time usage.
-The task should instead be left to the robust layer of the parser, possibly
-implemented by using external resources.
-
-\end{itemize}
-
-A thorougly evaluation is still to be done, this should be carried out by
-measuring the agreement between all parsed chunks and their respective
-translation from Talbanken. The evaluation should also be done manually.
-For this, we rely on an expert in Swedish.
+%\textbf{Dealing with valency information}
+%Many Swedish verbs can be used with different number of arguments.
+%Having one lexical entry for every possible usage of a verb
+%does not seem to be a good idea considering
+%the ambiguities it will lead to and the already high time usage.
+%Still valencies is crucial when parsing, and methods for handling 
+%this within the robust layer of the parser are to be implemented, possibly
+%implemented by using external resources.
 
@@
 %From 6.1 - the discussion part + something about the latest work
+Our work resulted in a wide-coverage grammar for Swedish which --
+combined with statistical resources -- can be used for parsing 
+open-domain text. The grammar covers a large number of syntactic 
+phenomena and uses an extensive morphological lexicon. To
+our knowledge, it is the most comprehensive computational grammar for
+Swedish and the most complex GF grammar to date.
+Besides parsing, the grammar may well be used for language generation.
 
-The project has resulted in
-\begin{itemize}
-\item an extended grammar covering an important part of Swedish
-\item a comparison between GF and another annotation
-\item a deeper testing of the Swedish resource grammar and an estimation
-of how well GF can be used to describe larger parts of a language
-\item a study of how dependent types can be used in the resource grammars
-\end{itemize}
+All parts of the project is open-source and the grammar and other parts
+are modular and may well be reused and developed.
 
-The grammar has been extended and enhanced, and its current status is
-a specialized extension of the resource grammar.
-Besides parsing, the grammar may well be used for language generation,
-which works fast even when using an extensive lexicon.
-
-Open source, transparent, modular, explores the strengths and weakneses
-of GF.
+%Open source, transparent, modular, explores the strengths and weakneses
+%of GF.
 \bibliographystyle{splncs}

### 2012-03-21 tsd/TSD2012/paper.tex
tsd updates

@@
 Swedish. As a result 65 $\%$ of the Talbanken trees were translated
-into the GF format in the training stage and 75 \% of the noun phrases
+into the GF format in the training stage and 76 \% of the noun phrases
 was parsed during the evaluation. Moreover, we obtained a language
@@
 way. The lexicon limitation is not a problem, as we use an existing 
-large Swedish lexicon \cite{ahlbergEnache}, which is being continuously
-developed and contains more than 120 000 entries.
+large Swedish lexicon \cite{ahlbergEnache}, which is imported to GF
+format from an electronic dictionary and which contains more than
+%being continuously
+%developed and contains more than 
+100 000 entries.
 %... \textbf{Malin, can you say a bit more ?But not more than a couple
@@
 Special reflexive pronouns and reflexive possessive pronouns for the 3rd
-exist \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd person forms.
+person exist \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd
+person forms.
 %\enumsentence{
@@
 a freely available, manually annotated, large-scale treebank.
-It is analyzed with the MAMBA annotation scheme (Teleman, 1974) and 
-the more than 6000 sentences we have been using
-%consists of four
+It is analyzed with the MAMBA annotation scheme and the section used for our
+work contains more than 6000 sentences
+%consists of four we have been using
 %parts. Two of them are transcriptions of spoken language, one a collection of
 %text written by high school students, and one, section P,
-consists of professionally written Swedish gathered from newspapers, brochures and textbooks.
+of professionally written Swedish gathered from newspapers, brochures and textbooks.
 
@@
 %Discussion about the two dimensions of a grammar - lexicon and syntax 
-There are two aspects of on a computational grammar -- the work on the lexicon
-and the work regarding syntax.  For this article, we will focus on the
-syntactical dimension of our work, the description and formalization of grammar
-rules and the usage of Talbanken.
+There are two aspects of a computational grammar -- the work on the lexicon
+and the work regarding syntax.  
+This article focuses on the syntactical dimension and the following section
+illustrates the work on describing grammatical constructions in the GF 
+formalism.
+%For this article, we will focus on the
+%of our work, the description and formalization of grammar
+%rules and the usage of Talbanken.
 
@@
 Swedish has two ways of forming passive verb phrases: the 
-periphrastic passive, formed by using the modal auxiliary verb `bli'
-(\emph{`become'})
+periphrastic passive, formed by using the modal auxiliary verb \emph{bli}
+(`become')
 and the s-passive which is formed by adding an \emph{s} to the verb. 
@@
 %turned into complete verb phrases; they no longer need an object.
-During this project, the s-passive was added although the
-periphrastic passive is still allowed.
+During this project, an implementation of   the s-passive was added
+%although the periphrastic passive is still allowed.
 %The grammar further allows not only transitive verbs, but 
@@
 A ditransitive verb -- \emph{ge} (`give') (\ref{sent:give2pass}) -- gives rise to two
-passives, (\ref{ex:passV33}) and (\ref{ex:passV32}), both covered by our new implementation.
+passives, (\ref{ex:passV32}a) and (\ref{ex:passV32}b), both covered by our grammar.
 \eenumsentence{
@@
 {\emph{we} & \emph{offered}& \emph{her}&\emph{the job}}
-{``We offered her the job"}}
-\label{sent:give2pass}
-\item{\textbf{Passive use, first place \hspace{3mm}Passive use, second place}\\\shortex{7}
-{Hon & erbjöds & jobbet && \hspace{3mm}Jobbet & erbjöds & henne }
+{``We offered her the job"}
+\label{sent:give2pass}}
+\item{\textbf{Passive use, first place} &c. \textbf{Passive use, second place}}
+\\\shortex{6}
+{Hon & erbjöds & jobbet &\hspace{18pt}Jobbet & erbjöds & henne }
 {\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}
-&&  \hspace{3mm}\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
+&  \hspace{18pt}\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
 {``She was offered the job"}
-~``The job was offered to her"}}
-\label{ex:passV33}
+~~``The job was offered to her"}
+\label{ex:passV33}}
 %\item{\textbf{Second place in two-place verbs}\\\shortex{3}
@@
 noun phrase must be of such type that it requires both the noun and its modifiers
-to appear in indefinite form. The verb phrase must consist of an intransitive verb
+to appear in indefinite form \cite{cooper}. The verb phrase must consist of an intransitive verb
 or be in passive form.
@@
 As noun phrases containing reflexive pronouns may be used as ordinary
-noun phrases, apart from the restrictions stated above,
+noun phrases -- apart from the restrictions stated above --
 we do not want them to differentiate
-from other \verb-NP- on the type level since we would then need to duplicate
-our code.
+from other \verb-NP- on the type level since this would require code duplication.
 %due to the rich type system in GF. 
@@
 
-\subsubsection{Overgenerating}
+\subsubsection{Overgeneration}
 Another aspect of the grammar implementation was the avoidance of overgeneration. As
@@
 overgeneration were present that did not cause problems when working with small
-lexicons and controlled languages, but that caused explosions of ambiguities when 
+lexicons and controlled language, but that caused explosions of ambiguities when 
 tested on Talbanken. For the function \verb-GenNP-, used for genitive, we managed
 to decrease the number of parse trees for problematic, although short, sentences to 10 \%.
+
 % Could also be here:
@@
 Talbanken format to GF abstract syntax trees. We hence get a comparison 
-%\vspace{-9mm}
 \begin{figure}[!h]
@@
 \subfloat[Talbanken tree]
-{\begin{tikzpicture}
-\Tree [.S [.SS [.POOP Det ]]
-          [.FV [.AVPS \"ar ]]
-          [.NA [.ABNA inte ] ]
-          [.SP [.AJ l\"att ]]
-          [.IP [.IP . ] ]
-      ]
-\end{tikzpicture}
-}
+{\includegraphics[width=60mm]{TBTree_cropped.pdf}}
 \subfloat[Translation to GF]
@@
 \label{fig:gftree1}
-\vspace{-5mm}
 \end{figure}
-between the two annotations and at the same time we get methods for extracting
-and translating to GF notation. 
+between the two annotations and at the same time we extract a GF treebank.
 If a node fails to be translated, a meta variable  --annotated \verb-?- -- is put in its place.
@@
 Talbanken05 uses three kinds of tags: categories, edge labels and POS-tags. 
-While the POS-tags are reserved for words, the categories give grammatical information,
-such as \verb|S|, \verb|NP| and \verb|VP|.
+While the POS-tags are reserved for words, the categories give grammatical information:
+ \verb|S|, \verb|NP| and \verb|VP|.
 The edge labels give the part of sentence: \verb|SS| for subject, 
@@
 \verb-S-, we can conclude that this is a sentence-complement verb.
-In GF, the valency information is given for each entry in the lexicon.
-A sentence-complement verb has the type \verb-VS- and must the 
-function \verb-ComplVS- must be used for combining the verb and with a 
-sentence.
+In GF, the valency information is given explicitly for each entry in the lexicon.
+A sentence-complement verb has the type \verb-VS- and the 
+function \verb-ComplVS-  must be used for combining the verb and  a 
+sentence into a complete verb phrase.
 %Another annotational difference that turned out to be problematic for the
@@
 %Further we need to be able to give output
-even when encountering unknown words and grammatical constructions 
-or idioms and ellipses.
+even when encountering unknown words and grammatical constructions,
+idioms and ellipses.
 Our approach makes use of the rich annotation in Talbanken -- we perform chunk parsing
@@
 Due to the differences in the annotation between Talbanken and
-GF mentioned is section \ref{sec:mapping} verbs are not considered
+GF mentioned is section \ref{sec:mapping}, verbs are not considered
 as parsable chunks. %The GF parser depend on requires verb particles and prepositions to give
@@
 
-For the extraction of a GF Treebank from Talbanken, (section \ref{sec:mapping})
+For the extraction of a GF Treebank from Talbanken (section \ref{sec:mapping}),
 the evaluation measures the numbers of meta variables in each translated tree.
@@
 While statistical methods compensates for constructions not covered by
-the grammar, we still aim for a even wider grammar coverage. 
+the grammar, we still aim for an even broader grammar coverage. 
 As examples of constructions to be added, we mention
@@
 All parts of the project is open-source and the grammar and other parts
-are modular and may well be reused and developed.
+are modular and may well be reused and further developed.
 

### 2012-03-22 tsd/TSD2012/paper.tex
final tsd

@@
 \usepackage{amssymb}
+\usepackage{wrapfig}
 \usepackage[utf8]{inputenc}
@@
 The work describes a wide-coverage computational grammar for Swedish. 
-It is developed using GF(Grammatical Framework), a dependently-typed 
-functional language, specialized for grammar programming. We trained
+It is developed using GF (Grammatical Framework), a %dependently-typed 
+functional language specialized for grammar programming. We trained
 and evaluated the grammar by using Talbanken, one of the largest treebanks for 
-Swedish. As a result 65 $\%$ of the Talbanken trees were translated
-into the GF format in the training stage and 76 \% of the noun phrases
+Swedish. As a result 65$\%$ of the Talbanken trees were translated
+into the GF format in the training stage and 76\% of the noun phrases
 was parsed during the evaluation. Moreover, we obtained a language
@@
 %Outline of the paper. 
+Two main approaches have divided the research in computational linguistics.
+The first and most prominent  one is the
+wide-coverage method, mostly based on
+statistical methods, which compensates its usually shallow analysis of the
+input with its coverage. The second one is the opposite - provides
+a deep analysis, but has a limited coverage. Our work on a
+computational grammar from Swedish falls into the 2nd category, but
+aims at getting the best of both worlds by enlarging the coverage in
+order to fit free text, without affecting the quality of the model.
 
-Two main approaches divide the research in computational linguistics
-in the last years. The first one is the wide-coverage method, usually based on
-statistics methods, which compensates its usually shallow analysis of the
-input with its coverage. The other approach is the opposite - provides
-a deep analysis of the data, but it's coverage is limited to
-very specialized input. Our work tries to unite the two approaches by
-developing a wide-coverage computational grammar for Swedish. We start
-with a minimal model of the language and extend its coverage
-significantly, without affecting the quality. The grammar was
-developed and evaluated against Talbanken\cite{talbanken}, one of the
-largest Swedish treebanks.
-
+%
+%Two main approaches divide the research in computational linguistics
+%in the last years. The first and most prominent  one is the
+%wide-coverage method, usually based on
+%statistical methods, which compensates its usually shallow analysis of the
+%input with its coverage. The second one is the opposite - provides
+%a deep analysis, but has a limited coverage. Our work on a
+%computational grammar from Swedish falls into the 2nd category, but
+%aims at getting the best of both worlds by enlarging the coverage in
+%order to fit "real-world'' text, without affecting the quality of the model.
+%
 %The paper is organized as follows: .... \textbf{maybe leave out if no space}
@@
 
-\subsection{GF}
-\label{sec:gf}
+%\subsubsection{GF}
+%\label{sec:gf}
 
-The computational grammar was developed within GF(Grammatical
+%\textbf{GF}
+The computational grammar was developed within \textbf{GF} (Grammatical
 Framework) \cite{gfbok}. GF is a dependently-typed grammar formalism,
@@
 fragment that is described by a concrete grammar and a natural
-language generation tool that is able to generate constructions from
+language generation tool for constructions from
 the same language fragment. 
@@
 grammars into abstract and concrete, it is possible to achieve 
-semantics-preserving translation between any pair of languages from 
-the grammar.
+semantics-preserving translation between any language pair.
+%of languages from the grammar.
 
@@
 structures. For this reasons GF was mainly used so far for controlled
-languages\cite{cnl} and, dialogues systems \cite{talk} and interactive 
+languages\cite{cnl}, dialogues systems\cite{talk} and interactive 
 systems where the user is guided to stay
 within the bounds of the grammar by a predictive
-parser\cite{parser}, such as a multilingual tourist
+parser, such as a multilingual tourist
 phrasebook\cite{phrasebook}. 
-However, there is some recent work on parsing
+However, there is recent work on parsing
 free text --- a grammar for patent claims from the biomedical
 domain \cite{patent}. This work highlighted a
-number of caveats of real-world text, but in the same time gives
-valuable insights that future work on wide-coverage GF grammars can
-benefit from.
+number of caveats of real-world text, but in the same time gave
+valuable insights for future work on wide-coverage GF grammars.
+%that future work on wide-coverage GF grammars can benefit from.
 
@@
 %developed and contains more than 
-100 000 entries.
+100 000 entries.\\
 %... \textbf{Malin, can you say a bit more ?But not more than a couple
@@
 
-\subsection{Swedish}
-
+\textbf{Swedish} 
 % Chapter 2.4(with emphasis on the syntactic structures)
-Swedish is a North-Germanic language, sharing most of its grammatical structure 
-with Norwegian and Danish. It is also one of the official languages in Finland
-and is spoken by approximately 9 million people.
-Although Swedish syntax often is similar to English, the  morphology is richer and the
+ is a North-Germanic language, sharing most of its grammatical structure 
+with Norwegian and Danish. It spoken by approximately nine million
+people, in Sweden and parts of Finland. 
+%is also one of the official languages in Finland, altogether spoken .
+Although Swedish syntax is often similar to English, the  morphology is richer and the
 word order more intricate.
@@
 The normal word order is subject-verb-object, but fronting other constituents
--- topicalisation -- is very common,  especially for temporal and
-locative adverbial phrases \cite[\textsection 1027]{H&H}.
+(topicalisation) is very common,  especially for temporal and
+locative adverbial phrases. %\cite[\textsection 1027]{H&H}.
 Fronting the finite verb marks questions.
@@
 %\label{gfPass:s-pass}
-Special reflexive pronouns and reflexive possessive pronouns for the 3rd
-person exist \cite[\textsection 310 \& 319]{H&H}, distinct from the normal 3rd
-person forms.
+Special reflexive pronouns and reflexive possessive pronouns for the third
+person exist, % \cite[\textsection 310 \& 319]{H&H}, 
+distinct from the normal third person forms.\\
 %\enumsentence{
@@
 
-\subsection{Talbanken}
-
+%\textbf{Talbanken}
 %Chapter 2.2
-\label{sec:talbanken}
-For testing and evaluation of the grammar and lexicon, we needed to be able to
-compare them against a reliable source.
-We have used Talbanken \cite{talbanken},
+%\label{sec:talbanken}
+For testing and evaluation of the grammar and lexicon, we needed 
+a reliable source for comparison.
+%to be able to
+%compare them against 
+We have used \textbf{Talbanken} \cite{talbanken},
 a freely available, manually annotated, large-scale treebank.
-It is analyzed with the MAMBA annotation scheme and the section used for our
+%It is analyzed with the MAMBA annotation scheme and 
+The section used for our
 work contains more than 6000 sentences
@@
 of professionally written Swedish gathered from newspapers, brochures and textbooks.
-
-Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
-and was then redistributed in an updated version,
+%Talbanken was also used to train the Swedish version of the Malt parser \cite{malt}
+It is also redistributed in an updated version,
 Talbanken05 \cite{talbanken05}.
@@
 %between the authors
-There exist a number of parser for Swedish already, such as the
+There exist a number of parsers for Swedish already, such as the
 data-driven Malt parser \cite{malt}, also trained on Talbanken, the 
@@
   %last 2 do... could you rephrase please ?} 
-
 Among computational grammars for Swedish, we mention the Swedish
@@
 
-The current work uses the extensive Swedish
-lexicon developed in \cite{ahlbergEnache} and the Swedish resource
-grammar as a starting point.
-
-Currently, GF is the leading technology in the European project
-MOLTO (Multilingual On-Line Translation)
-\footnote{http://www.molto-project.eu/}, which aims at developing
+%The current work uses the extensive Swedish
+%lexicon developed in \cite{ahlbergEnache} and the Swedish resource
+%grammar as a starting point. 
+Besides the usages mentioned before, GF is currently the leading technology in the European project
+MOLTO\footnote{http://www.molto-project.eu/},
+which aims at developing
 tools for translating between 15 languages in real-time and with high quality. 
-
 Previous examples of larger type-theoretical GF grammars that use 
-dependent-types are SUMO-GF \cite{ontologies2}, a GF representation of
-SUMO \cite{sumo}, the largest open-source ontology and  a natural
-language generation grammar via Montague semantics \cite{montague}.
+dependent-types are SUMO-GF \cite{ontologies2},
+a GF representation of SUMO\footnote{http://www.ontologyportal.org/}, the largest open-source ontology
+and  a natural language generation grammar via Montague semantics \cite{montague}.
 
@@
 %Discussion about the two dimensions of a grammar - lexicon and syntax 
-There are two aspects of a computational grammar -- the work on the lexicon
-and the work regarding syntax.  
-This article focuses on the syntactical dimension and the following section
+%There are two aspects of a computational grammar -- the work on the lexicon and the work regarding syntax.  
+The work focuses on the syntactical dimension and the following section
 illustrates the work on describing grammatical constructions in the GF 
-formalism.
+formalism.\\
 %For this article, we will focus on the
@@
 
-\subsection{Resource grammar}
-
+\textbf{Resource grammar}
 %Chapter 2.1.2
@@
 The grammars describe how to construct phrases and sentences and how to
-decline words. They cover the morphology,
+decline words. They cover the 
 word order, agreement, tense, basic conjunction etc.
@@
 the implementation for Swedish is shared with Norwegian and Danish. The modules
-that concern the lexical aspects are separate, while 85 \% of the syntax
+that concern the lexical aspects are separate, while 85\% of the syntax
 description is shared.  There are about 80 functions, which describe the rules
-for building phrases and clauses.
-\begin{verbatim}
-PredVP       : NP -> VP -> Cl ; -- Predication
-ComplVPSlash : VPSlash  -> VP ; -- Complementation
-\end{verbatim}
+for building phrases and clauses, such as functions for predication and
+complementation:\\
+\small \texttt{
+\indent PredVP ~~~~~ : NP -> VP -> Cl ; -- Predication\\
+\indent ComplVPSlash : VPSlash  -> VP ;~ -- Complementation}\\
+\normalsize
+%\small
+%\begin{verbatim}
+%PredVP       : NP -> VP -> Cl ; -- Predication
+%ComplVPSlash : VPSlash  -> VP ; -- Complementation
+%\end{verbatim}
+%\normalsize
 %The analysis preformed by GF is driven by the parts of speech, which are combined into
@@
 %, simple called the \verb|Extra| module,
-where language specific constructions may bee added
+where language specific constructions may be added.\\
 %The functions given here do not have to be translatable
@@
 
-\subsection{Lexicon}
-
+\textbf{Lexicon}
 %Everything about Saldo, Lexin, etc, compressed.
-%
 %We use the resources from the other paper, so to say :)
-As mentioned before, our main lexical resource is SALDO \cite{saldo} from
-which a GF lexicon has been extracted \cite{ahlbergEnache}.
+As mentioned before, our main lexical resource is SALDO\footnote{http://spraakbanken.gu.se/resurs/saldo}
+%\cite{saldo}
+from which a GF lexicon has been extracted \cite{ahlbergEnache}.
 Valency information, which is a key feature to good parsing using GF,
-are extracted from the lexicon
+is extracted from the lexicon
 Lexin\footnote{http://spraakbanken.gu.se/lexin/}.
@@
 %- miscellaneous 
-It has earlier been hard to identify missing constructions of the Swedish
+Earlier it has been hard to identify missing constructions of the Swedish
 implementation since there was no large resource available to evaluate it on.
@@
 periphrastic passive, formed by using the modal auxiliary verb \emph{bli}
-(`become')
+(``become")
 and the s-passive which is formed by adding an \emph{s} to the verb. 
 Passive voice is often used, especially the
-s-passive, % (sentence \ref{gfPass:s-pass}).
-which is used in more than 80 \% of the times according to some studies 
+s-passive. % (sentence \ref{gfPass:s-pass}).
+%which is used in more than 80\% of the times according to some studies 
 %Some studies suggest that the s-passive is 
-\cite{laanemets}.
-It is however not as common in the other Scandinavian languages, 
+%\cite{laanemets}.
+It is however not so common in the other Scandinavian languages, 
 where not all words have passive forms for all tenses. The resource grammar for
@@
 %turned into complete verb phrases; they no longer need an object.
-During this project, an implementation of   the s-passive was added
+During this project, an implementation of   the s-passive was added.
 %although the periphrastic passive is still allowed.
@@
 %All verb phrases that misses an object are allowed to form passives.
-A ditransitive verb -- \emph{ge} (`give') (\ref{sent:give2pass}) -- gives rise to two
-passives, (\ref{ex:passV32}a) and (\ref{ex:passV32}b), both covered by our grammar.
+A ditransitive verb -- \emph{ge} (``give") (\ref{sent:give2pass}) -- gives rise to two
+passives, (\ref{ex:passV33}b) and (\ref{ex:passV33}c), both covered by our grammar.
 \eenumsentence{
-\item{\textbf{Active use of two-place verb}\\\shortex{4}
+\item{\textbf{Active use of two-place verb}\\\shortexnt{4}
 {Vi & erbjöd& henne& jobbet }
 {\emph{we} & \emph{offered}& \emph{her}&\emph{the job}}
-{``We offered her the job"}
+{(We offered her the job)}
 \label{sent:give2pass}}
-\item{\textbf{Passive use, first place} &c. \textbf{Passive use, second place}}
+\item{\textbf{Passive use, first place} ~c. \textbf{Passive use, second place}}
 \\\shortex{6}
-{Hon & erbjöds & jobbet &\hspace{18pt}Jobbet & erbjöds & henne }
+{Hon & erbjöds & jobbet &\hspace{22pt}Jobbet & erbjöds & henne }
 {\emph{she} &\emph{ offered+\textbf{s}}&\emph{ the job}
-&  \hspace{18pt}\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
-{``She was offered the job"}
-~~``The job was offered to her"}
-\label{ex:passV33}}
+&  \hspace{22pt}\emph{the job} & \emph{offered+\textbf{s}}&\emph{ her}}
+{(She was offered the job) ~~~(The job was offered to her)}}
+\label{ex:passV33}
 %\item{\textbf{Second place in two-place verbs}\\\shortex{3}
@@
 %{``The job was offered to her"}}
-\label{ex:passV32}}
 
-\subsubsection{Impersonal constructions}
+\noindent \textbf{Impersonal constructions}
 \label{sec:Formal}
-Formal subjects \cite[\textsection 19]{SAG} are often used in Swedish.
-\enumsentence{\shortex{6}
+Formal subjects %\cite[\textsection 309d]{H&H}
+are often used in Swedish.
+\enumsentence{\shortexnt{6}
 {Det & sitter & en & fågel & på & taket}
 {\emph{it }& \emph{sits }& \emph{a} & \emph{bird} &\emph{ on} & \emph{the roof}}
-{``There is a bird sitting on the roof''}}
+{(There is a bird sitting on the roof)}}
 %\emph{`Det'} has the position of the subject, and the real subject, 
@@
 noun phrase must be of such type that it requires both the noun and its modifiers
-to appear in indefinite form \cite{cooper}. The verb phrase must consist of an intransitive verb
-or be in passive form.
+to appear in indefinite form. % \cite{cooper}.
+The verb phrase must consist of an intransitive verb
+or be in passive form.\\
 %A very common example of impersonal constructions is sentences with the verbs \emph{finnas},
@@
 
-\subsubsection{Formalizing the rules for reflexive pronouns by using dependent types}
+\noindent \textbf{Formalizing the rules for reflexive pronouns by using dependent types}
 \label{sec:reflexives}
@@
 not be used in subject noun phrases of finite sentences,
-as shown by the ungrammatical examples in 
-sentence (\ref{ex:ReflGenB}) and (\ref{ex:ReflSubj}).
+as shown by the ungrammatical examples  
+(\ref{ex:ReflGenB}) and (\ref{ex:ReflSubj}).
 %Furthermore, the antecedent must be within the same finite sentence as the reflexive
@@
 Still our grammar should accept the sentences (\ref{ex:ReflGen})
-%(\ref{sent:rakasig}) 
 and (\ref{ex:ReflAdv}) :
@@
 \item
-{Sina vantar hade han glömt på tåget.\\
-{\sc self's} gloves, he had forgotten on the train.}
+{Sina vantar hade han tappat. 
+~~({\sc self's} gloves, he had lost.)}
 \label{ex:ReflGen}
 \item
-{*Sina vantar var kvar på tåget.\\
-{~~ \sc self's} gloves was left on the train.}
+{*Sina vantar var kvar på tåget.
+({\sc self's} gloves was left on the train.)}
 \label{ex:ReflGenB}}
-%\eenumsentence{
-%\item{
-%Jag vill att han rakar sig.\\
-%I want him to shave {\sc self}.}
-%\label{sent:rakasig}
-%\item
-%{*Han vill att jag ser på sig.\\
-%~~He wants me to look at {\sc self}.}\label{sent:sesig}
-%}
 \eenumsentence{
 \item{
-Han är längre än sin kompis.\\
-He is taller than {\sc self's} friend.
+Han är längre än sin kompis.
+(He is taller than {\sc self's} friend.)
 }\label{ex:ReflAP} \label{ex:ReflAdv}
-\item{*Han och sin kompis läser en bok.\\
-He and {\sc self's} friend are reading a book. }
+\item{*Han och sin kompis leker.
+~~(He and {\sc self's} friend are playing.) }
 \label{ex:ReflSubj}
@@
 
-In the standard GF analysis, which is preformed bottom-up starting from the
-POS-tags, information about semantic roles are given by the functions, not by the
+In the standard GF analysis, which is performed bottom-up starting from the
+POS-tags, information about syntactic roles are given by the functions, not by the
 categories. For example, the first argument of the function 
 \verb-PredVP- always acts as the subject, but the noun phrase itself does not carry
-information about its semantic role.
+information about its syntactic role.
 As noun phrases containing reflexive pronouns may be used as ordinary
@@
 by letting the type \verb-NP- depend on an argument, 
-which may either be \verb-Subject- or
-\verb-Object-.
-\begin{verbatim}
-PredVP       : NP Subject -> VP        -> Cl ; -- predication
-ComplVPSlash : VPSlash    -> NP Object -> VP ; -- complementation
-PrepNP       : (a : NPType) -> Prep -> NP a -> Adv a ; 
-                                           -- adverbial phrases 
-\end{verbatim}
-
-We hence combine the-part-of-speech driven analyse normally preformed by
-GF with a part-of-sentence analysis, where the dependent types gives the 
+which may either be \verb-Subject- or \verb-Object-.\\
+\small \texttt{
+\indent PredVP  ~~~~~ : NP Subject -> VP        -> Cl ; ~~~ -- predication\\
+\indent ComplVPSlash : VPSlash    -> NP Object -> VP ; -- complementation\\
+\indent PrepNP       : (a: NPType) -> Prep -> NP a -> Adv a ; -- adverbial phrase \\
+}
+\normalsize
+We hence combine the-part-of-speech driven analysis normally performed by
+GF with a part-of-sentence analysis, where the dependent types give the 
 information we were missing.
@@
 This is the first example in GF making use of the dependent types for 
-describing pure syntax of a natural language. Our approach could be extended
-to describing similar relations between 
-noun phrases and antecedent in other languages.
+describing pure syntax of a natural language. The approach could be extended
+to describing similar relations
+%between noun phrases and antecedent 
+in other languages.\\
 %having could be extended to use the type system in similar ways.
 
-\subsubsection{Overgeneration}
+\noindent \textbf{Overgeneration}
 Another aspect of the grammar implementation was the avoidance of overgeneration. As
@@
 tested on Talbanken. For the function \verb-GenNP-, used for genitive, we managed
-to decrease the number of parse trees for problematic, although short, sentences to 10 \%.
+to decrease the number of parse trees for short, yet problematic sentences to 10\%.
 
@@
 
-\section{A GF Treebank for Swedish}
+\section{A GF Treebank and Model-Based Disambiguation}
 \label{sec:mapping}
@@
 
-Talbanken contains much valuable information about phrase structures and word 
+Talbanken contains very valuable information about phrase structures and word 
 usage.
@@
 {\includegraphics[width=60mm]{TBTree_cropped.pdf}}
-\subfloat[Translation to GF]
+\subfloat[Translation to GF abstract tree]
 {\includegraphics[width=60mm]{latt2.pdf}}
-\caption{The sentence \emph{Det är inte lätt} (``It is not easy"). If the word
-         \emph{lätt} (`easy') is unknown to the grammar, a meta variable is put
-         in its place.}
+\caption{The sentence \emph{Det är inte lätt} (``It is not easy").}
 \label{fig:gftree1}
@@
 between the two annotations and at the same time we extract a GF treebank.
-If a node fails to be translated, a meta variable  --annotated \verb-?- -- is put in its place.
-This is also the case when two sister nodes can be translated, but no rule is
-applicable for joining them into one tree.
+Nodes that fail to be translated are represented by a meta variable,
+annotated \verb-?-.  In figure \ref{fig:gftree1}, the word \emph{lätt} (``easy")
+was unknown and therefore represented by a meta variable in the 
+translated tree. Meta variables are also used for connecting 
+sister nodes that can be translated on their own, but not joined into one tree.
 
@@
 One example is valency information, which is given implicitly by the complements
-of a word in Talbanken. If a verb is followed by an object, \verb-OO-, which contains an
+of a word in Talbanken. If a verb is followed by an object, \verb-OO- containing an
 \verb-S-, we can conclude that this is a sentence-complement verb.
@@
 %to metas, which lowers our result. \\
+The translation of Talbanken to GF gives us a GF treebank consisting of more than
+6000 sentences.
+Moreover, the treebank could be mapped to another language from the 
+resource library provided that one defines transfer rules for lexical
+items and implements the extra syntactic constructions.
 
-\section{Chunk Parsing}
-\label{sec:chunk}
-%New section - insights about latest work
-The grammar, which is hand written, cannot be expected to cover all
-possible expressions of the language.
-Hence, we combine it with statistical means to get good coverage. 
-%The parser uses named entity recognition and a compounding analysis.
-The parser should give as much output as possible 
-%Further we need to be able to give output
-even when encountering unknown words and grammatical constructions,
-idioms and ellipses.
-Our approach makes use of the rich annotation in Talbanken -- we perform chunk parsing
-relying on the tags given in the treebank. 
-Consequentially, whenever a sentence is not recognized by the grammar, it is 
-parsed chunk by chunk and the parse trees for each chunk is returned.
-Noun phrases, adverbial phrases and adjectival phrases are considered.
-Due to the differences in the annotation between Talbanken and
-GF mentioned is section \ref{sec:mapping}, verbs are not considered
-as parsable chunks. %The GF parser depend on requires verb particles and prepositions to give
-%a correct output, and which are usually not in the same Talbanken chunk. 
-Therefore, our focus is on whole sentences and noun phrases. At the first stage,
-we try to parse the basic sentence structure, and therefore allow the parser
-to parse complicated chunks as dummy words. This way, we don't loose the 
-high-level sentence analysis given by GF.
-When this is done, we try to separately parse the chunks,
-until as much of the sentence as possible is covered.
-
-%This way we get good coverage -- from ca 10 \% of Talbanken -- to ??
-
-\section{Language Model-Based Disambiguation}
-
-%New section - fill in with the basic principle + new results(when available)
-The translation of Talbanken to GF gives us a GF treebank consisting of more than
-6000 sentences. The translation process is explained in section
-\ref{sec:mapping} and the trees give valuable 
-information about which words from our lexicon that have been used
-and which functions that are needed to parse the sentence.
+The trees give valuable 
+information about which lexical entries and functions
+are needed to parse the sentence.
 GF has built-in support for probabilities, and when feeding the output
@@
 
+\section{Chunk Parsing}
+\label{sec:chunk}
+%New section - insights about latest work
+The grammar, which is hand written, cannot be expected to cover all
+possible expressions of the language.
+Hence, we combine it with statistical means to get an even better coverage. 
+%The parser uses named entity recognition and a compounding analysis.
+The parser should give as much output as possible 
+%Further we need to be able to give output
+even when encountering unknown words and grammatical constructions,
+idioms and ellipses.
+Our approach uses of the rich annotation in Talbanken -- we perform chunk parsing
+relying on the tags given in the treebank. 
+Consequentially, whenever a sentence is not recognized by the grammar, it is 
+parsed chunk by chunk and the parse trees for each chunk are returned.
+Noun, adverbial and adjectival phrases are considered.
+Due to the differences in the annotation between Talbanken and
+GF mentioned is Section \ref{sec:mapping}, verbs are not considered
+as parsable chunks. %The GF parser depend on requires verb particles and prepositions to give
+%a correct output, and which are usually not in the same Talbanken chunk. 
+Therefore, our focus is on whole sentences and noun phrases. At the first stage,
+we try to parse the basic sentence structure, and therefore allow the parser
+to treat complicated chunks as dummy words. This way, we don not lose the 
+high-level sentence analysis given by GF.
+Subsequently, we try to separately parse the chunks,
+until as much of the sentence as possible is covered.
+
+%This way we get good coverage -- from ca 10\% of Talbanken -- to ??
+
+%\section{Language Model-Based Disambiguation}
+
+%New section - fill in with the basic principle + new results(when available)
+
 \section{Evaluation}
 
-For the extraction of a GF Treebank from Talbanken (section \ref{sec:mapping}),
+For the extraction of a GF treebank from Talbanken (Section \ref{sec:mapping}),
 the evaluation measures the numbers of meta variables in each translated tree.
@@
 %tense.
-For all of Talbanken, we get a number of 65 \%, and if we limit our input to
-uncomplicated sentences with no more than 10 words where all words are known to
-our lexicon, we can reach 85 \%. 
+Overall we get a number of 65\%, and if we limit our input to
+simple sentences with no more than 10 words, all known to
+our lexicon, we reach 85\%. 
 %When translating the whole treebank, only 78 sentences
@@
 and is still being carried out. So far, our evaluation results cover
-600 sentences -- about 10 \% of the treebank. 
-%Without the robustness given by the chunk parser, we can only parse 17.9 \% of the sentences.
-For noun phrases, we cover 76 \%.  By using the chunk parsing described in section
-\ref{sec:chunk}, we identify the structure on 65 \% of the sentences. That
+600 sentences -- about 10\% of the treebank. 
+%Without the robustness given by the chunk parser, we can only parse 17.9\% of the sentences.
+For noun phrases, we cover 76\%.  By using the chunk parsing described in Section
+\ref{sec:chunk}, we identify the structure on 65\% of the sentences. That
 is, the parser can identify the verbs and how they relate to the noun phrases,
@@
 another valency, other prepositions or particles than the lexicon has assigned
-it, the sentence structure can not be identified. In that case
-the parser returns chunks, showing the parse trees of identifiable sub-phrases.
+it, the sentence structure cannot be identified. In that case
+the parser returns chunks, showing the parse trees of identifiable subphrases.
 
@@
 ~~~~~~\textbf{Evaluation}
-A thoroughly evaluation is still to be done. This should be carried out by
+A more thorough evaluation is still to be done. This should be carried out by
 measuring the agreement between all parsed chunks and their respective
@@
 The grammar should cover the most prominent Swedish features.
-While statistical methods compensates for constructions not covered by
+While statistical methods compensate for constructions not covered by
 the grammar, we still aim for an even broader grammar coverage. 
@@
 
-\textbf{Improving memory and time consumption}
-When using GF for parsing with lexicon of our size, we currently
-suffer from memory usage problems. The parsing algorithm is not fully
-capable of evaluating all possible parse trees, which is necessary in order
-to find the best one.
+%\textbf{Improving memory and time consumption}
+%When using GF for parsing with lexicon of our size, we currently
+%suffer from memory usage problems. The parsing algorithm is not fully
+%capable of evaluating all possible parse trees, which is necessary in order
+%to find the best one.
 
@@
 Besides parsing, the grammar may well be used for language generation.
-
-All parts of the project is open-source and the grammar and other parts
-are modular and may well be reused and further developed.
+All parts of the project are open-source and the grammar and other parts
+are modular and could be reused and further developed.
 

### 2012-05-19 tsd/TSD2012/paper.tex
comments from jury

@@
 \begin{abstract}
+%\textbf{typos, omissions, grammar, proof read!}
 The work describes a wide-coverage computational grammar for Swedish. 
@@
 decline words. They cover the 
-word order, agreement, tense, basic conjunction etc.
+word order, agreement, tense, basic conjunction, etc.
 Due to the syntactic similarities between the Scandinavian languages, much of
@@
 complementation:\\
+%\textbf{explain notation, why ComplVPSlash?}\\
 \small \texttt{
-\indent PredVP ~~~~~ : NP -> VP -> Cl ; -- Predication\\
-\indent ComplVPSlash : VPSlash  -> VP ;~ -- Complementation}\\
+\indent PredVP ~~~ : NP -> VP -> Cl ;~~~~~~ -- Predication\\
+\indent ComplSlash : VPSlash  -> NP -> VP ;~ -- Complementation}\\
 \normalsize
+Here the types for the functions are given -- \texttt{PredVP} will return a clause when 
+given its input arguments: a noun phrase and a verb phrase.
 %\small
@@
 \small \texttt{
-\indent PredVP  ~~~~~ : NP Subject -> VP        -> Cl ; ~~~ -- predication\\
-\indent ComplVPSlash : VPSlash    -> NP Object -> VP ; -- complementation\\
+\indent PredVP  ~~~ : NP Subject -> VP        -> Cl ; ~~~ -- predication\\
+\indent ComplSlash : VPSlash    -> NP Object -> VP ; -- complementation\\
 \indent PrepNP       : (a: NPType) -> Prep -> NP a -> Adv a ; -- adverbial phrase \\
@@
 \noindent \textbf{Overgeneration}
+%\textbf{how many at first, how about other sentences?}
 Another aspect of the grammar implementation was the avoidance of overgeneration. As
@@
 
+%\textbf{About the underlying probalistic model of GF}
 Talbanken contains very valuable information about phrase structures and word 
@@
 The edge labels give the part of sentence: \verb|SS| for subject, 
-\verb|OO| for object etc. 
+\verb|OO| for object, etc. 
 %
@@
 \label{sec:chunk}
+%\textbf{what does eval in terms of metas mean? is gf not necessary when chunking}
 %New section - insights about latest work
@@
 Consequentially, whenever a sentence is not recognized by the grammar, it is 
-parsed chunk by chunk and the parse trees for each chunk are returned.
+parsed chunk by chunk. Each chunk is analysed by the grammar-driven parser, and
+resulting trees are returned.
 Noun, adverbial and adjectival phrases are considered.
@@
 \section{Evaluation}
+%\textbf{limited, how many rules? 45 functions added, plus
+%categories, updates, and tense}
 
@@
 \section{Conclusion}
+%\textbf{only minimal success? not ripe enough?}
 %From 6.1 - the discussion part + something about the latest work
@@
 Swedish and the most complex GF grammar to date.
+%\textbf{based on what?}
 Besides parsing, the grammar may well be used for language generation.
